### 2016-04-15 guidelines.tex

\documentclass[10pt]{report}

\usepackage[T1]{fontenc}
\usepackage{amsmath}
\begin{document}
	\textbf{Information about the model code in MATLAB}\\

	1. Model generation is carried out from a excel/text file using modelgen.m\\

	2. Using FBA to calculate the steady state fluxes is optional (what if it is a toy model?)\\

		2a. If using FBA or PFBA, fix uptake fluxes, reaction bounds and essential and nonessential exchange reactions in the model.\\

		2b. For toy model steady state fluxes can be alternatively specified in the excel/text file\\

	3. Specify metabolite concentrations that you may know\\

		3a. Sample the other metabolite concentrations using ACHR within a constrained space.\\

		3b. If not sampling, assign metabolites a concentration either using a file or in the script file.\\

\end{document}

### 2017-05-31 KineticModel/estimation/docs/formulation/noisy_estimate.tex
optim formulation for noisy data

\documentclass[10pt]{report}

\usepackage{amsmath}
\usepackage{array}
\usepackage{tabu}
\usepackage{lmodern}

\begin{document}
	Optimization problem for estimation of kinetic parameters from noisy experimental data. The noisy data is generated through the use of additive noise in flux equations that use a known kinetic model to predict the responses of a metabolic network to different perturbations.\\
	\begin{equation}
	v^* = g(x,p) + {\cal{N}}(0,1)
	\end{equation}
	where ${\cal{N}}(0,1)$ is Gaussian noise with zero mean and a standard deviation of 1.
	\begin{equation}\label{eq:ode1}
	\frac{d}{dt}pep=v_1-v_2-v_4
	\end{equation}
	\begin{equation}\label{eq:ode2}
	\frac{d}{dt}fdp=v_2-v_3
	\end{equation}
	\begin{equation}\label{eq:ode3}
	\frac{d}{dt}E=v_{e,max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right) - d E
	\end{equation}
	The kinetic expressions for fluxes $v_1$ through $v_4$ are given below. The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics. The acetate flux through $v_1$ is also governed by the quantity of available enzyme E. 
	\begin{equation}\label{eq:flux1}
	v_1 = k_{1}^{cat}E\frac{acetate}{acetate+K_{1}^{acetate}}
	\end{equation}	
	\begin{equation}\label{eq:flux2}
	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
	\end{equation}
	\begin{equation}\label{eq:flux3}
	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
	\end{equation}
	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$. The added flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
	\begin{equation}\label{eq:flux4}
	v_4 = k_{4}^{cat}.pep
	\end{equation}

	We denote the known noisy flux and concentration information with the superscript $^*$. Accordingly, fluxes $v^*$ and concentrations $x^*$ are the noisy information that will be used for estimation. Fluxes ($v$) and concentrations ($x$) that predicted by the model (without noise) are denoted without the superscript.

	The optimization formulation to estimate fluxes given below is based on the minimization of the tolerance for the L2-norm of the difference between the model predicted flux and the noisy flux. 
	\begin{center}
		\begin{subequations}\label{eq:noisyopt}
			\begin{align}
			\underset{x,p}{\mathrm{min}} & \text{      }\varepsilon\\
			\mathrm{st} & \text{      }|v-v^*| \le \varepsilon\\
			& \text{      }|x-x^*| \le \varepsilon\\
			& \text{      }Sv = 0\\
			& \text{      }v = f(x,p)\\
			& \text{      }x_{min}\le x \le x_{max}\\
			& \text{      }p_{min} \le p \le p_{max}
			\end{align}
		\end{subequations}		
	\end{center}

	In Equation (\ref{eq:noisyopt}) above, $\varepsilon$ is the absolute tolerance between the predicted and the measured quantities. The relationship between the predicted and the measured quantities are given by the nonlinear constraints in Equations (\ref{eq:noisyopt}b) and (\ref{eq:noisyopt}c). The stoichiometric constraints for steady state are given by Equation (\ref{eq:noisyopt}d), and Equation (\ref{eq:noisyopt}e) specifies the kinetic rate laws for the predicted fluxes as a function of the concentrations ($x$) and the parameters ($p$). Equations (\ref{eq:noisyopt}f) and (\ref{eq:noisyopt}g) are the bounds for the concentrations and parameters that are to be estimated by the optimization problem.

\end{document}

### 2017-05-31 KineticModel/estimation/docs/formulation/noisy_estimate.tex
include comments from Krishna on model formulation

@@
 			\begin{align}
-			\underset{x,p}{\mathrm{min}} & \text{      }\varepsilon\\
+			\underset{x,p}{\mathrm{min}} & \text{      }e\\
 			\mathrm{st} & \text{      }|v-v^*| \le \varepsilon\\
@@
 			& \text{      }Sv = 0\\
-			& \text{      }v = f(x,p)\\
+			& \text{      }v = f(x,p) + e\\
 			& \text{      }x_{min}\le x \le x_{max}\\
@@
 Equations (\ref{eq:noisyopt}f) and (\ref{eq:noisyopt}g) are the bounds for the concentrations and parameters that are to be estimated by the optimization problem.
+The above optimization problem is to be solved for fixed values of $\varepsilon \in \{\text{.01, .05, .1}\}$.
 

### 2017-06-13 KineticModel/estimation/est_noise.m
add code for other fluxes - incomplete

@@ -71,7 +71,19 @@ opts.opt_x0 = x0;
 % opts.opt_alg = 'GN_ISRES';
 opts.solver = 'scip';
 odep_opt = odep_bkp;
-opt_sol2 = runoptimp(opts,plist,odep_opt,optimopts,@optimize_p_noisy);    
+opt_sol_f1 = runoptimp(opts,plist,odep_opt,optimopts,@optimize_flux1);   
+
+%% flux 2 estimation
+p_id = [5,13];
+xi = sol(1).xss(:,5);
+pi = opts.odep; % pi(11) = 1;
+x0 = [xi;pi(p_id)';.1];
+opts.opt_x0 = x0;
+
+opts.solver = 'scip';
+odep_opt = odep_bkp;
+opt_sol_f2 = runoptimp(opts,plist,odep_opt,optimopts,@optimize_p_noisy);   
+
 
 
 

### 2017-06-13 KineticModel/estimation/optim_scripts/optimize_flux1.m
functional representation for flux 1 parameter opt

@@ -1,0 +1,18 @@
+function optsol = optimize_flux1(opts,xss,fss,plist,odep_bkp)
+
+opts.odep = odep_bkp;
+optsol = struct([]); 
+
+ts1 = tic;
+
+% 'K1ac','k1cat'
+fprintf('Optimizing with a single set of perturbations\n');
+fprintf('\nOptimizing parameters for flux 1.....\n');
+[x_opt_4_1,opt_id_4_1,~,fval_4_1] = flux1_k_noisy(opts,xss,fss,plist);
+optsol(1).x_opt = x_opt_4_1;
+optsol(1).opt_id = opt_id_4_1;
+optsol(1).fval = fval_4_1;
+optsol(1).time = toc(ts1);
+fprintf('Time to estimate flux 1 parameters %4.3g\n',toc(ts1));
+
+fprintf('\nTotal time for parameter estimation %4.3g\n',toc(ts1));

### 2017-06-13 KineticModel/estimation/docs/formulation/noisy_estimate.tex
changes made to forumaltion based on scip results

@@
 			\begin{align}
-			\underset{x,p}{\mathrm{min}} & \text{      }e\\
+			\underset{x,p,e}{\mathrm{min}} & \text{      }e\\
 			\mathrm{st} & \text{      }|v-v^*| \le \varepsilon\\
@@
 			& \text{      }x_{min}\le x \le x_{max}\\
-			& \text{      }p_{min} \le p \le p_{max}			
+			& \text{      }p_{min} \le p \le p_{max}\\	
+			& \text{      }e_{min} \le e \le e_{max}						
 			\end{align}
@@
 
+	\paragraph{Changes in problem formulation due to feasibility issues:}
+	Since the above problem (Equation \ref{eq:noisyopt}) does not result in feasible global optimal solutions (using SCIP), the constraints are relaxed and the feasible problem is presented below.
+
+	\begin{center}
+		\begin{subequations}\label{eq:optmod}
+			\begin{align}
+			\underset{x,p_i,e_i}{\mathrm{min}} & \text{      }e_i\\
+			\mathrm{st} & \text{      }|v_i-v^*| \le 0.2\\
+			& \text{      }|x-x^*| \le 0.3\\
+			& \text{      }|Sv| \le 1\text{x}10^{-8}\\
+			& \text{      }v_i = f(x,p_i) + e\\
+			& \text{      }v_{j\not= i} = f(x,p_{j\not=i})\\
+			& \text{      }x_{min}\le x \le x_{max}\\
+			& \text{      }p_{i,min} \le p_i \le p_{i,max}\\
+			& \text{      }e_{i,min} \le e_i \le e_{i,max}					
+			\end{align}
+		\end{subequations}		
+	\end{center}
+
+	In Equation (\ref{eq:optmod}) above, the index $i$ represents the flux whose parameters are being optimized and index $j$ represents all other fluxes whose parameters are held constant.
+
+	The steady state constraint has been relaxed (from a equality constraint to an inequality constraint) along with permitting the concentrations and the estimated fluxes to deviate significantly from their experimentally observed values (~20 - 30\%).
+
+	The bounds for the flux noise were set at (0,20) in the above problem.
 \end{document}

### 2017-07-05 KineticModel/estimation/docs/formulation/noisy_estimate.tex
changed formulation w/o ss constraints

@@
 	The bounds for the flux noise were set at (0,20) in the above problem.
+
+	\paragraph{Suggested modifications to address infeasibility (July 4)}
+	After discussion with Prof. M, the following changes were made to the problem formulation:
+	\begin{itemize}
+		\item Experimental fluxes may/may not be at steady state.
+Accordingly, using the steady state constraint in the above formulation can render the optimization problem infeasible.
+Suggestion is to remove this constraint(s).
+		\item Include the norm minimization constraint as an objective.
+		\item Add the noise in fluxes and concentrations as part of the bound constraint.
+		\begin{center}
+			\begin{subequations}
+				\begin{align}
+				\underset{x,p_i}{\mathrm{min}} & \text{      }\left\Vert v_i-v^*\right\Vert\\
+				& \text{      }v_i = f(x,p_i)\\
+				& \text{      }x_{min}(1-\epsilon)\le x \le x_{max}(1+\epsilon)\\
+				%& \text{      }v_{min}(1-\epsilon)\le v \le v_{max}(1+\epsilon)\\
+				& \text{      }p_{i,min} \le p_i \le p_{i,max}\\
+				\end{align}
+			\end{subequations}
+		\end{center}
+	\item The noise $\epsilon$ is fixed at 5\% or 10\% to signify experimental noise.
+	\end{itemize}
+
 \end{document}

### 2017-07-05 KineticModel/Kotte2014/KotteMATCONTflux.m
add constitutive expression flux for heterologus expression simulations

@@ -56,6 +56,9 @@ flux(2) = vemax.*(1-1./(1+(KeFBP./M(2)).^ne));
 ratio = 1+M(2)/KFbpFBP;
 flux(3) = vFbpmax.*(ratio-1).*(ratio).^3/(ratio.^4+Lfbp*(1+M(1)./KFbpPEP).^(-4));
 
+% no regulation FBP -  added to test heterologous expression w/o regulation
+% flux(3,:) = vFbpmax.*M(2)./(M(2)+KFbpFBP);
+
 % vEX(PEP)
 flux(4) = vEXmax.*M(1)./(M(1)+KEXPEP);
 

### 2017-07-08 KineticModel/estimation/optim_scripts/nlconsopt.m
deprecates nlconstoptim_flux_noCAS through simplification of inputs/outputs

@@ -1,0 +1,111 @@
+function optsol = nlconsopt(prob,x0,solveropts,miscdata)
+if nargin<4
+    miscdata = [];
+end
+% problem structure
+if isfield(prob,'obj')
+    obj = prob.obj;
+else
+    obj = [];
+end
+if isfield(prob,'nlcons')
+    nlcons = prob.nlcons;
+else
+    nlcons = [];
+end
+if isfield(prob,'lb')
+    lb = prob.lb;
+else
+    lb = [];
+end
+if isfield(prob,'ub')
+    ub = prob.ub;
+else
+    ub = [];
+end
+if isfield(prob,'optimopts')
+    optimopts = prob.optimopts;
+else
+    optimopts = [];
+end
+if isfield(prob,'nlrhs')
+    nlrhs = prob.nlrhs;
+else
+    nlrhs = [];
+end
+if isfield(prob,'nle')
+    nle = prob.nle;
+else
+    nle = [];
+end
+
+% options structure
+if isfield(solveropts,'solver')
+    solver = solveropts.solver;
+else
+    solver = 'nlopt';
+end
+if isfield(solveropts,'opt_alg')
+    opt_alg = solveropts.opt_alg;
+else
+    opt_alg = 'none';
+end
+
+if isempty(obj)
+    error('No given objective function');
+end
+
+tstrt = tic;
+optimopts = optiset('solver',solver,...
+                      'maxiter',5000,...
+                      'maxfeval',500000,...
+                      'tolrfun',1e-6,...
+                      'tolafun',1e-6,...
+                      'display','final');   
+switch(solver)
+    case 'nlopt'        
+        switch(opt_alg)            
+            case 'GN_ISRES'
+                solver_opts = nloptset('algorithm','GN_ISRES'); 
+            otherwise
+                solver_opts =...
+                nloptset('algorithm','AUGLAG','subalgorithm','LN_COBYLA'); 
+        end                
+    case 'ipopt'
+         
+    case 'scip'
+        solver_opts = scipset('scipopts',{'limits/time',1e20});
+        optimopts = optiset(optimopts,'maxnodes',10000000);
+    otherwise
+        error('Nonexistent solver in OPTI');
+end
+optimopts = optiset(optimopts,'solverOpts',solver_opts);   
+
+if ~isempty(nlcons)
+    optim_prob =...
[…]

### 2017-07-08 KineticModel/estimation/scripts-other/gen_noise.m
use wrapper to call optimization code

@@ -31,7 +31,7 @@ noisy_sol = sol;
 pt_p = repmat(odep_bkp,length(pt_val.exp_pval),1);
 pt_p(:,pt_val.exp_pid) = pt_val.exp_pval;
 [noisy_sol.xss,noisy_sol.fss] = addnoise(sol.xss,pt_p);
-noisy_sol.xss = sol.xss + random(pd,3,size(sol.xss,2));
+% noisy_sol.xss = sol.xss + random(pd,3,size(sol.xss,2));
 
 % get only data from one steady state
 pss = ones(length(pt_val.exp_pval),1);
@@ -42,7 +42,8 @@ pss(noisy_sol.xss(1,:)>noisy_sol.xss(2,:)) = 0;
 %                    'fss',{noisy_fss});
 
 % problem defn
-optimdata = struct('nvar',5,'nc',3,'vexp',noisy_sol.fss(:,logical(pss)),'p_id',[1 11]);
+optimdata = struct('nvar',5,'nc',3,'vexp',noisy_sol.fss(:,logical(pss)),...
+                    'p_id',[1 11],'flxid',1);
 
 % set objective
 obj = @(x)objnoisy(x,odep_bkp,optimdata);
@@ -60,18 +61,9 @@ ub(optimdata.nc+1:end) = [10;10];
 % initial values for consrained nl(or quadratic?) optimization
 x0 = [xss1;odep_bkp(optimdata.p_id)'];
 
-optimopts = optiset('solver','scip',...
-                      'maxiter',5000,...
-                      'maxfeval',500000,...
-                      'tolrfun',1e-6,...
-                      'tolafun',1e-6,...
-                      'display','final');   
-optimopts = optiset(optimopts,'maxnodes',10000000);   
-solveropts = scipset('scipopts',{'limits/time',1e20});
-optimopts = optiset(optimopts,'solverOpts',solveropts);
-optim_prob = opti('obj',obj,'bounds',lb,ub,'options',optimopts);
-[xval,fval,exitflag,info] = solve(optim_prob,x0); 
-
+prob = struct('obj',obj,'lb',lb,'ub',ub);
+solveropt = struct('solver','scip');
+optsol = nlconsopt(prob,x0,solveropt);
 
 
 

### 2017-07-08 KineticModel/estimation/optim_scripts/nlconsopt.m
do not test for solution optimality

@@ -23,11 +23,11 @@ if isfield(prob,'ub')
 else
     ub = [];
 end
-if isfield(prob,'optimopts')
-    optimopts = prob.optimopts;
-else
-    optimopts = [];
-end
+% if isfield(prob,'optimopts')
+%     optimopts = prob.optimopts;
+% else
+%     optimopts = [];
+% end
 if isfield(prob,'nlrhs')
     nlrhs = prob.nlrhs;
 else
@@ -91,21 +91,9 @@ else
 end
 [xval,fval,exitflag,info] = solve(optim_prob,x0); 
 
-if exitflag == 1 &&...
-   (strcmpi(info.Status,'Success')||...
-    strcmpi(info.Status,'Optimal')||...
-    strcmpi(info.Status,'Converged / Target Reached')||...
-    strcmpi(info.Status,'Globally Optimal'))
-
-    xopt = xval;
-elseif ~isempty(xval) && ~isnan(fval)
-    xopt = xval;    
-else
-    xopt = [];    
-end
-
 optsol.x0 = x0;
-optsol.xval = xopt;
+optsol.xval = xval;
 optsol.fval = fval;
+optsol.exitflag = exitflag;
+optsol.info = info;
 optsol.time = toc(tstrt);
-

### 2017-07-08 KineticModel/estimation/objectives/objnoisy.m
supply index of objective flux in structure

@@ -9,4 +9,4 @@ vest = getmodelflux(x(1:data.nc),p);
 vest(1) = x(5).*x(3).*p(17)./(p(17)+x(4));
 
 % calc objective
-fx = sqrt(sum((vest(1)-data.vexp(1,:)).^2));
+fx = sqrt(sum((vest(data.flxid)-data.vexp(data.flxid,:)).^2));

### 2017-07-15 KineticModel/estimation/scripts-other/gen_noise.m
supply # multisolve pts in solveropts

@@ -65,7 +65,7 @@ setup_opts.nle = allnle;
 x0 = [noisy_xss(:,2);optimdata.odep(optimdata.p_id)';noisy_fss(optimdata.flxid,2)];
 
 % prob = struct('obj',obj,'nlcons',nlcons,'nlrhs',nlrhs,'nle',nle,'lb',lb,'ub',ub);
-solveropt = struct('solver','ipopt','multi',1);
+solveropt = struct('solver','ipopt','multi',1,'multi_pts',[6 12]);
 optsol = nlconsopt(prob,x0,solveropt,optimdata);
 
 % compare fluxes and concentrations

### 2017-07-15 KineticModel/estimation/optim_scripts/nlconsopt.m
supply # multistart points in options structure

@@ -55,7 +55,11 @@ if isfield(solveropts,'multi')
 else
     multi = 0;
 end
-
+if isfield(solveropts,'multi_pts')
+    multi_pts = solveropts.multi_pts;
+else
+    multi_pts = [5 12];
+end
 if isempty(obj)
     error('No given objective function');
 end
@@ -95,7 +99,7 @@ else
     opti('obj',obj,'bounds',lb,ub,'options',optimopts);    
 end
 if multi
-    [xval,fval,exitflag,info] = multisolve(optim_prob,[],[5 12]);   
+    [xval,fval,exitflag,info] = multisolve(optim_prob,[],multi_pts);   
 else
     [xval,fval,exitflag,info] = solve(optim_prob,x0); 
 end     

### 2017-07-15 KineticModel/estimation/scripts-other/est_flux2.m

@@ -66,7 +66,7 @@ x0 = [noisy_xss(:,2);...
       noisy_fss(optimdata.flxid,2)];
 
 % prob = struct('obj',obj,'nlcons',nlcons,'nlrhs',nlrhs,'nle',nle,'lb',lb,'ub',ub);
-solveropt = struct('solver','ipopt','multi',1);
+solveropt = struct('solver','ipopt','multi',1,'multi_pts',[7 12]);
 optsol = nlconsopt(prob,x0,solveropt,optimdata);
 
 % compare fluxes and concentrations

### 2017-07-15 KineticModel/estimation/docs/formulation/noisy_estimate.tex
clearly specify index dimensions
updates to formulation for estimation

@@
 				\begin{align}
-				\underset{x,p_i}{\mathrm{min}} & \text{      }\left\Vert v_i-v^*\right\Vert\\
-				& \text{      }v_i = f(x,p_i)\\
+				\underset{x,p_j}{\mathrm{min}} & \text{      }\left\Vert v_i-v^*\right\Vert &  i\in\mathcal{R}^n,\text{  }j\in\mathcal{R}^l\\
+				\mathrm{st}& \text{      }v_i = f(x,p_j)\\
 				& \text{      }x_{min}(1-\epsilon)\le x \le x_{max}(1+\epsilon)\\
 				%& \text{      }v_{min}(1-\epsilon)\le v \le v_{max}(1+\epsilon)\\
-				& \text{      }p_{i,min} \le p_i \le p_{i,max}\\
+				& \text{      }p_{j,min} \le p_j \le p_{j,max}
 				\end{align}
@@
 
+	\paragraph{More changes to address nonlinearity due to division(July 13)}
+	Prof. M suggested changing the nonlinear formulation of the flux equations into just a bilinear form by removing the denominator and formulating it as nonlinear algebraic equation.
+	\begin{center}
+		\begin{subequations}
+			\begin{align}
+			\underset{x,p_j,v_i}{\mathrm{min}} & \text{      }\left\Vert v_i-v^*\right\Vert &  i\in\mathcal{R}^n,\text{  }j\in\mathcal{R}^l\\
+			\mathrm{st}& \text{      }N(x,p_j) - v_iD(x,p_j) = 0\\
+			& \text{      }x_{min}(1-\epsilon)\le x \le x_{max}(1+\epsilon)\\			
+			& \text{      }p_{j,min} \le p_j \le p_{j,max}\\
+			& \text{      }v_{i,min}(1-\epsilon)\le v_i \le v_{i,max}(1+\epsilon)
+			\end{align}
+		\end{subequations}
+	\end{center}
+
+	\begin{itemize}
+		\item Also, the initial noise added to the steady state solutions was reduced from 50\% to 5\% or 10\%.
+		\item $\epsilon$ is fixed at 50\%.
+		\item Solution is obtained using multisolve with IPOPT instead of aiming to obtain a global solution with SCIP.
+		\item Using approximately 14 million points for the second phase of multistart in OPTI
+	\end{itemize}
+
 \end{document}

### 2017-07-15 KineticModel/estimation/constraints/consnoisyf3.m
files for estimating flux 3 parameters

@@ -1,0 +1,4 @@
+function fx = consnoisyf3(x,p,data)
+
+ratio = 1+x(2)./x(4);
+fx = x(6).*(ratio-1).*(ratio).^3 - x(end).*(ratio.^4+p(3).*(1+x(1)./x(5)).^(-4));

### 2017-07-15 KineticModel/estimation/scripts-other/est_flux3.m

@@ -1,0 +1,77 @@
+% estimate fluxes for flux 3 in kotte model fdp ---> 
+
+% generate experimental data - get initial ss
+tspan = 0:0.1:100;
+[xdyn,fdyn,xss1,fss1,opts] = run_nonoise(tspan);
+
+% backup parameters and initial conditions
+ival_bkp = opts.x0;
+odep_bkp = opts.odep;
+
+% generate nsmp samples by adding random noise to ss values
+nsmp = 10;
+[noisy_xss,noisy_fss] = addnoise(repmat(xss1,1,nsmp),odep_bkp);
+
+% figure
+% boxplot([noisy_fss';fss1']);
+% figure
+% boxplot([noisy_xss';xss1']);
+
+% perturb system from noisy initial conditions
+pt_sol_id = [1 2 3];
+[exp_sol,noisy_sol] = dopert_noisy(opts,noisy_xss,odep_bkp,pt_sol_id);
+close all
+
+% get only data from one steady state
+pss = ones(1,numel(exp_sol.exp_pval));
+% pss(exp_sol.xss(1,:)>exp_sol.xss(2,:)) = 0;    
+
+% options structure for solving problem
+optimdata = struct('nc',3,'nf',1,'flxid',3,'eps',.5,...
+                    'vexp',exp_sol.fss(:,logical(pss)),...                    
+                    'odep',odep_bkp,...
+                    'wt_xss',noisy_xss(:,1),'wt_fss',noisy_fss(:,1));
+
+% define all objectiv efun handles
+allobjh = {'objnoisy',...
+           [],...
+           'objnoisy',...
+           'objnoisy',...
+           'objnoisy',...
+           []};
+% define all constraint fun handles
+allconsh = {'consnoisyf1',...
+            [],...
+            'consnoisyf3',...
+            'consnoisyf2',...
+            'consnoisyf1',...
+            []};
+% define all rhs for nl cons
+allnlrhs = {0,[],0,0,0,[]};
+% define cons type for nl cons
+allnle = {0,[],0,0,0,[]};
+
+% problem defn
+setup_opts = optimdata;
+setup_opts.obj = allobjh;
+setup_opts.nlcons = allconsh;
+setup_opts.nlrhs = allnlrhs;
+setup_opts.nle = allnle;
+
+[prob,optimdata] = setup_optim_prob(setup_opts);
+
[…]

### 2017-07-15 KineticModel/estimation/scripts-other/gen_noise.m
add all constraint funs to cell array

@@ -43,9 +43,9 @@ allobjh = {'objnoisy',...
 % define all constraint fun handles
 allconsh = {'consnoisyf1',...
             [],...
-            'consnoisyf1',...
-            'consnoisyf1',...
-            'consnoisyf1',...
+            'consnoisyf3',...
+            'consnoisyf2',...
+            'consnoisyf4',...
             []};
 % define all rhs for nl cons
 allnlrhs = {0,[],0,0,0,[]};

### 2017-07-15 KineticModel/estimation/optim_scripts/setup_optim_prob.m
change parameter lb to .05

@@ -56,7 +56,7 @@ end
 lb(1:nc) = exp_xss.*(1-eps);
 ub(1:nc) = exp_xss.*(1+eps);
 % set bounds - parameter
-lb(nc+1:nc+length(p_id)) = .1*ones(length(p_id),1); 
+lb(nc+1:nc+length(p_id)) = .05*ones(length(p_id),1); 
 ub(nc+1:optimdata.nc+length(p_id)) = 10*ones(length(p_id),1);
 % set bounds - flux
 lb(nvar-nf+1:nvar) = exp_fss(flxid)*(1-eps);

### 2017-07-27 KineticModel/estimation/scripts-other/est_flux1.m
change number of multistart points

@@ -64,7 +64,7 @@ setup_opts.nle = allnle;
 x0 = [noisy_xss(:,2);optimdata.odep(optimdata.p_id)';noisy_fss(1,2)];
 
 % prob = struct('obj',obj,'nlcons',nlcons,'nlrhs',nlrhs,'nle',nle,'lb',lb,'ub',ub);
-solveropt = struct('solver','ipopt','multi',1);
+solveropt = struct('solver','ipopt','multi',1,'multi_pts',[7 7]);
 optsol = nlconsopt(prob,x0,solveropt,optimdata);
 
 % compare fluxes and concentrations

### 2017-07-27 KineticModel/estimation/docs/formulation/noisy_estimate.tex
add results from earlier formkulation

@@
 \usepackage{lmodern}
+\usepackage{graphicx}
+\usepackage[space]{grffile}
+\usepackage{subfigure}
+\usepackage{longtable}
 
@@
 
+	\paragraph{Results:}
+	These are the results from the parameter estimation algorithm that we have been running. I have tried estimating parameters for the Kotte model by using noisy perturbation data.
+
+	\begin{figure}[!h]%figure1
+		\centering{\includegraphics[width=0.6\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1_phenotypes.pdf}}
+		\caption{The two stable phenotypes for acetate consumption through the gluconeogenic model.
+a) The high production phenotype which has a high flux throughout the network as a result of a high \textit{pep} concentration that reduces the inhibition on acetate uptake and b) The low production phenotype that is characterized by a relatively low \textit{pep} concentration and consequently increased inhibition on the acetate uptake as a result of a smaller \textit{fdp}.}\label{fig:fig1}
+	\end{figure}
+
+	\begin{table}[!tbhp]
+		\caption{Table showing the perturbed values of all fluxes used for parameter estimation.}
+		\begin{center}				
+			\begin{tabular}{ccc}
+				Designation & Perturbed Fluxes & Perturbed Values\\
+				\hline
+				P1 & $v_1$ & 2\\
+				P2 & $v_2$ & 0.2\\
+				P3 & $v_3$ & 0.5
+			\end{tabular}
+		\end{center}	
+	\label{tab:pval}
+	\end{table}
+
+	\paragraph{Estimation for $v_1$}: The steady state data from the above perturbations in Table \ref{tab:pval} were used to estimate the value of the parameters $k_1^{cat}$ and $K_1^{acetate}$ for flux $v_1$.
+
+	The comparison of the estimated and the experimental (noisy model generated) data are shown below.
+Notice that the model estimates do not match well with the noisy data for perturbation P1 (Table \ref{tab:pval}).
+This is the observed case for all other parameter estimations.
+This could be due to the effect of perturbation to the acetate uptake flux ($v_1$) on the steady states of the system.
+(Note that a higher acetate uptake flux lets the system reach the higher steady state).
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/est_flux1_conc_Jul20}}
+		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$. pep, fdp and enzyme E (Top to bottom)}
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/est_flux1_flux_Jul20}}
+		\caption{plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$.}
+	\end{figure}
+	\clearpage
+
+	\paragraph{Estimation for $v_2$}: The steady state data from the above perturbations in Table \ref{tab:pval} were used to estimate the value of the parameters $V_2^{max}$ and $K_2^{pep}$ for flux $v_2$.
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/est_flux2_conc_Jul20}}
+		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$. pep, fdp and enzyme E (Top to bottom)}
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/est_flux2_flux_Jul20}}
+		\caption{plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$.}
+	\end{figure}
+
+	The estimation of parameter for flux $v_2$ also runs into the same issues mentioned above for flux $v_1$.
+
+	\paragraph{Estimation for $v_3$}: The steady state data from the above perturbations in Table \ref{tab:pval} were used to estimate the value of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ for flux $v_2$.
+
+	The optimization cannot find an optimal solution for this problem.
+The optimization problem using IPOPT with the OPTI Toolbox  ends with a cannot find an optimal solution error despite increasing the number of multistart points considerably.
+
+	\begin{itemize}
+		\item Noise to original steady state was 5\%.
+		\item Bounds for both fluxes and concentrations in the aforementioned formulation were fixed at 50\%.
+	\end{itemize}
+
 \end{document}

### 2017-07-29 KineticModel/estimation/docs/formulation/noisy_estimate.tex
add new formulation that includes pertubed vals as vars

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{array}
@@
 
+\paragraph{Changes in formulation}
+This formulation assumes that the concentrations and the fluxes corresponding to each perturbation are variables in the optimization problem formulated to estimate kinetic parameters.
+
+Let $\mathbf{x}\in\mathbb{R}^m$, $\mathbf{p}\in\mathbb{R}^l$ and $\mathbf{v}\in\mathbb{R}$ be the vector of concentrations, parameters and fluxes respectively.
+If the number of perturbations used for parameter estimation is in $\mathbb{R}^p$, then the vector of concentrations, fluxes and parameters that need to be estimated in the optimization problem changes to $\mathbf{x}\in\mathbb{R}^{mp}, \mathbf{v}\in\mathbb{R}^{p}$ and $\mathbf{p}\in\mathbb{R}^{l}$ respectively.
+
+Let $i\in\mathbb{R}^{mp}$, $j\in\mathbb{R}^{p}$ and $k\in\mathbb{R}^{l}$ represent the indices for concentrations, fluxes and parameters respectively.
+
+\begin{center}
+	\begin{subequations}
+		\begin{align}
+		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert\\
+		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+		\end{align}
+	\end{subequations}
+\end{center}
+
 	\paragraph{Results:}

### 2017-07-29 KineticModel/estimation/docs/formulation/noisy_estimate.tex
copy identifiability analysis text from q document from 2013

@@
 
+\paragraph{Identifiability Analysis:}
+Parameter uncertainties accompanying parameter inference and estimation in traditional deterministic models can also be reduced.
+Parameter uncertainties in deterministic models can be reduced by making sure that parameters in a given model are structurally (identifiability) and practically (estimability) identifiable (Nikerel et al., 2006).
+It has been shown that reducing the number of both structurally and practically non-identifiable parameters can result in reduction of model uncertainty propagated through uncertain parameter estimates (Hengl et al., 2007).
+The effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
+Assume a dynamic model for a model state variable vector x is represented by Equation 2.5 where u is the vector of input trajectories to the model and p is the vector of model parameters, and the observables y obtained from experimental data are represented in Equation 2.6 where ε is the intrinsic noise associated with the data.
+If elements of p are estimated based on the noisy data y, then a parameter pi in p is structurally identifiable if and only if there is a unique solution to the system in Equations 2.5-2.6 (Hengl et al., 2007).
+If many solutions exist for pi, then the parameter is said to be locally identifiable.
+Consequently, absence of any solution would result in the parameter being termed as non-identifiable in the context of Equations 2.5 and 2.6.
+
+Structural identifiability can also be interpreted in an alternative way based on maximum likelihood estimates (MLE) with respect to a given parameter (Raue, et al., 2009).
+A parameter is non-identifiable if a unique minimum, to the maximum likelihood estimate with respect to that parameter does not exist.
+Hence, parameters cannot be truly inferred if they are structurally non-identifiable.
+Several techniques used to differentiate structurally identifiable parameters from non-identifiable parameters were reviewed by McLean \& McAuley (2012).
+The advantages and disadvantages of using these techniques for highly nonlinear models that are frequently encountered in biology were discussed therein as well.
+In contrast to structural parameter identifiability, practical identifiability of a parameter helps in determining the ability and effects of the available experimental data on estimating model parameter values uniquely.
+Practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
+It can be gathered that both the model and the data play an important role in the ability to estimate parameters of a model.
+However, structural identifiability of any parameter in a given model depends on the fundamental biological process being modeled.
+Modifications to model structure to alleviate structural identifiability should not upend the purpose of the model, i.e.
+to represent the biological process and the underlying dynamics.
+Hence improvements in this regard are few and far between as more biological knowledge becomes available.
+Conversely, the ability to improve confidence in parameter estimates by decreasing practical non-identifiability of model parameters is a convenient solution.
+This can be done by improving the quality and quantity of data obtained for parameter estimation.
+Experimental design (ED) makes it possible to predetermine the quality and quantity of measurements obtained from experiments for the purpose of parameter estimation.
+The Fisher information matrix (FIM)-based ED strategies (McLean \& McAuley 2012) are one of the many ED methodologies out there.
+Most if not all are based on senstivity-based measures of the model predicted state variable to each parameter in the model.
+However, biological models are highly nonlinear and are subject to high-dimensional parameterization.
+As a result, increase in network size for large models (genome-scale models for example) increases the computational complexity of calculating such sensitivity matrices required for ED.
+Techniques based on graphical analysis of the sensitivity curves are also rendered moot due to the high-dimensional parameter space (McLean \& McAuley 2012).
+Profile likelihood (PL) based ED for parameter estimation aims to overcome this shortcoming in order to reduce practical as well as structural parameter identifiability.
+A complete description of PL based ED is provided in Appendix C.
+Although PL has been applied to small (4-12 parameters) and medium scale (15-40 parameters) networks (Raue, et al., 2009; Steiert et al., 2012), an application to a large scale network has not been accomplished.
+In addition, currently there is an absence of an ED-based strategy for estimating parameters for large scale transcriptional regulatory and metabolic network models.
+
+PL of any parameter θi,   in a vector θ of parameters is a measure of the log likelihood parameter estimate (LLE) as defined by Equation C.1.
+This methodology proposed by Raue et al.
+(2009) determines the PL of a parameter by minimizing the LLE of a parameter with respect to all other parameters in the model.
+The most important deliverable from this technique is the ability to distinguish between practically identifiable and non-identifiable parameters through confidence interval estimates for the said parameter.
+Confidence intervals are estimated from the PLs according to Equation C.2.
+The LLE of any given parameter can be estimated using logarithm of Equation C.3 that gives the likelihood estimate for a parameter vector θ based on a known vector of observables Y that are embedded with Gaussian noise.
+
+In Equation C.2, Δα is the α quartile of the chi-squared distribution with df degrees of freedom.
+Quantities ym and   in Equation C.3 correspond to the measured and predicted value of observable m respectively at time points tj in a time series.
+The variance of each observable m is written as  .
+Examples of (practically) identifiable and non-identifiable parameters are provided in Figure C.1 adapted from (Raue, et al., 2009).
+A practically identifiable parameter is distinguished from a non-identifiable one by the nature of the PL curve.
+In Panel A, the PL curve is asymptotic giving rise to an unbounded confidence interval for the corresponding parameter θi.
+This essentially would result in a huge parameter space that is not good enough for a confident model.
+Experimental data capable of distinguishing observables based on changes in a given parameter are simulated (Panel B of Figure C.1) and used to calculate a fresh PL for the said parameter in Panel A.
+Accordingly, the nature of the PL curve could change from being asymptotic to quadratic as shown in Panel C.
+This gives a very narrow and well-defined confidence interval for the estimated parameter unlike the case demonstrated in Panel A.
+Thus the nature of experiment to be performed and data to be collected is gleaned at the data simulation stage.
+Experimental data obtained through these designed experiments have been shown to change the characteristic of PL curves in small networks of ~40 parameters (Steiert et al., 2012).
+It is postulated that this way parameter uncertainties can be reduced and more confidence can be infused into model predictions.
+In addition to its utility as an experimental design aid, the algorithmic approach followed for PL estimation enables its integration into the computational model reconstruction and validation paradigm being developed as part of the current proposal.
+
 	\paragraph{Results:}

### 2017-07-29 KineticModel/estimation/scripts-other/gen_noise.m
include new wrapper to call multi start vs single ival

@@ -65,17 +65,10 @@ setup_opts.nle = allnle;
 
 % initial values for consrained nl(or quadratic?) optimization
 x0 = getrandomivals(optimdata,.1,1000);
-optsol = multistart_search(prob,x0,optimdata);
-
-x0 = [optimdata.xexp;optimdata.odep(optimdata.p_id)';optimdata.vexp];
-
-solveropt = struct('solver','ipopt','multi',1,'multi_pts',[2 2]);
-optsol = nlconsopt(prob,x0,solveropt,optimdata);
+solveropt = struct('solver','ipopt','multi',0);
+optsol = choose_nlconsopt(prob,x0,optimdata,solveropt);
 
 % compare fluxes and concentrations
 compare_vals(optsol,noisy_sol,optimdata,opts,pss);
-
-
-
 
 

### 2017-07-29 KineticModel/estimation/optim_scripts/multistart_search.m
add total time calculation

@@ -7,9 +7,11 @@ end
 
 npts = size(x0,2);
 cell_optsol = cell(npts,1);
+ts = tic;
 parfor ix0 = 1:npts % change this between serial/parallel 
     cell_optsol{ix0} = nlconsopt(prob,x0(:,ix0),solveropt,optimdata);     
 end
+fprintf('Total time for completion :%4.2g\n',toc(ts));
 
 % convert cell_optsol to structure array
 optsol = struct();

### 2017-07-29 KineticModel/estimation/optim_scripts/setup_optim_prob.m
set bounds separately for kcat and Km

@@ -64,6 +64,12 @@ ub(1:nc*npert) = xss_exp_v.*(1+eps);
 % set bounds - parameter
 lb(nc*npert+1:nc*npert+length(p_id)) = .05*ones(length(p_id),1); 
 ub(nc*npert+1:nc*npert+length(p_id)) = 5*ones(length(p_id),1);
+% set separate bounds for Km
+lb(nc*npert+1:nc*npert+1) = .05*ones(1,1); 
+ub(nc*npert+1:nc*npert+1) = 1*ones(1,1);
+% set separate bounds for kcat
+lb(nc*npert+2:nc*npert+length(p_id)) = .1*ones(1,1); 
+ub(nc*npert+2:nc*npert+length(p_id)) = 3*ones(1,1);
 % set bounds - flux
 lb(nvar-nf*npert+1:nvar) = vss_exp_v*(1-eps);
 ub(nvar-nf*npert+1:nvar) = vss_exp_v*(1+eps);

### 2017-08-02 KineticModel/estimation/docs/formulation/ident.tex
write-up on identification

\documentclass[10pt]{report}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{tabu}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage[space]{grffile}
\usepackage{subfigure}
\usepackage{longtable}
\usepackage[margin=1.0in]{geometry}
\renewcommand{\baselinestretch}{2.0}

\begin{document}
	\section{Introduction:}
	\subsection{Identifiability analysis: Definitions and Formulations}
	The identifiability of parameters in nonlinear models of physical processes can be classified into two categories: structural and practical identifiability. The effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter. The effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter. Practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model. The aforementioned distinction between practical and structural identifiability of parameters is illustrated in Figure. Thus, the identifiability of parameters in nonlinear models is dependent on the model structure, parameterization, and the quality and quantity of experimental data that is available for the purpose of estimation.

	On the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated. On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.

	We can use a profile likelihood-based approach to determine the structural and practical identifiability of parameters in any any dynamic model.

	We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/ed_with_pl}}
		\caption{Cartoon illustrating the utility of identifiability analysis for experimental design. a) The profile likelihood of a practically non-identifiable parameter that needs to be estimated based on both good and bad experimental data. The changes in the profile likelihood of the parameter when estimated with b) good experimental data and c) bad experimental. The identifiability of the parameter does not change due to the poor quality/quantity of experiments.}\label{fig:edwithpl}
	\end{figure}

	\section{Methods:}
	We use a profile likelihood-based approach to establish structural and practical identifiability of parameters in nonlinear kinetic models of metabolism. Briefly, the approach seeks to establish the existence/non-existence of bounds in confidence intervals for the estimates of parameters in nonlinear models. The profile likelihood is calculated based on Equation \ref{eq:pl} for each parameter $\theta_i$ where $\chi^2(\theta_i)$ is given by Equation \ref{eq:chi2}.
	\begin{align}\label{eq:pl}
	\chi_{PL}^2(\theta_i) = \underset{\theta_{j\ne i}}{\mathrm{min}} \left[\chi^2(\theta)\right]
	\end{align}
	\begin{align}\label{eq:chi2}
	\chi^2(\theta) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
	\end{align}
	Equation \ref{eq:chi2} is typically used as the objective to be minimized in a typical parameter estimation problem where $y_{kl}^*$ is the available experimental time course data for each observable state $k$ at each $l$ time point. The difference between this data and the model estimates at these time points, $y_{kl}$ is weighted by the experimental error $\sigma_{kl}^*$. An algorithm to calculate $\chi_{PL}^2(\theta_i)$ based on Equation \ref{eq:pl} is given below.

	\subsubsection{Algorithm:}	
	\begin{enumerate}
		\item Start by solving the numerical optimization problem in Equation \ref{eq:pl} for a fixed value of $\theta_i$ to determine the initial point on the profile likelihood curve.
		\item Take an increasing/decreasing step $\theta_{step}$ in the direction of $\theta_i$.
		\item Re-optimize $\theta_{j\ne i}$ using Equation \ref{eq:pl}.
		\item Repeat the last two steps until a desired threshold $\Delta_{\alpha}$ is exceeded or a maximal amount of steps is reached.
	\end{enumerate}

	\subsubsection{Adaptive step change:}
	The increasing/decreasing steps in $\theta_i$ can be adapted to the shape of the profile likelihood curve that is encountered while running the above algorithm. Accordingly, the step $\theta_{step}$ should satisfy Equation \ref{eq:adaptivestep} below, where $\theta^{k-1}$ refers to parameter estimates obtained from iteration $k-1$ of the above algorithm, and $q\in \left[0,1\right]$.
	\begin{align}\label{eq:adaptivestep}
	\chi^2\left(\theta^{k-1}+\theta_{step}\right) - \chi^2(\theta^{k-1}) \approx q.\Delta_{\alpha}
	\end{align}
\end{document}

### 2017-08-04 KineticModel/estimation/noiseless_model/dopert_nonoise.m
parameter estimation for non-noisy model

@@ -1,0 +1,33 @@
+function [exp_sol,no_noise_sol] =...
+        dopert_nonoise(opts,xss,fss,odep_bkp,pt_sol_id)
+    
+% generate perturbation data 
+% perturb system from deterministic (non-noisy) initial conditions
+id = 1;
+opts.x0 = xss(:,id);
+opts.tspan = 0:.1:300;
+opts.odep = odep_bkp;
+pt_val = struct('exp_pid',{11,12,13},...
+                'exp_pval',{[2],...
+                            [.2],...
+                            [.5]}); 
+sol = getperturbations(pt_val,@perturb_nonoise,opts);
+
+no_noise_sol = sol(pt_sol_id);
+j = length(pt_sol_id);
+% include wt (unperturbed) data
+no_noise_sol(j+1).xss = xss(:,id);
+no_noise_sol(j+1).fss = fss(:,id);
+no_noise_sol(j+1).exp_pid = 0;
+no_noise_sol(j+1).exp_pval = 0;
+no_noise_sol(j+1).odep = odep_bkp;
+
+% combine all perturbation flux data into a single vector
+exp_sol.xss = cat(2,no_noise_sol.xss);
+exp_sol.fss = cat(2,no_noise_sol.fss);
+exp_sol.exp_pid = cat(2,no_noise_sol.exp_pid);
+exp_sol.exp_pval = cat(2,no_noise_sol.exp_pval);
+exp_sol.odep = cat(1,no_noise_sol.odep);
+
+    
+    

### 2017-08-04 KineticModel/estimation/scripts-other/gen_nonoise.m

@@ -1,0 +1,70 @@
+% script to generate non-noisy data to estimated steady state
+% fluxes and concentrations in kotte model
+
+% generate experimental data - get initial ss
+tspan = 0:0.1:100;
+[xdyn,fdyn,xss,fss,opts] = run_nonoise(tspan);
+
+% backup parameters and initial conditions
+ival_bkp = opts.x0;
+odep_bkp = opts.odep;
+
+% perturb system from non-noisy initial conditions
+pt_sol_id = [1 2 3];
+[exp_sol,no_noise_sol] = dopert_nonoise(opts,xss,fss,odep_bkp,pt_sol_id);
+close all
+
+% get only data from one steady state
+pss = ones(1,numel(exp_sol.exp_pval));
+% pss(1) = 0;
+% pss(exp_sol.xss(1,:)>exp_sol.xss(2,:)) = 0;    
+
+% options structure for solving problem
+optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',1,'eps',.2,...
+                    'vexp',exp_sol.fss(:,logical(pss)),...
+                    'xexp',exp_sol.xss(:,logical(pss)),...
+                    'odep',odep_bkp,...
+                    'wt_xss',xss(:,1),'wt_fss',fss(:,1));
+                
+% define all objectiv efun handles
+allobjh = {'objnoisy',...
+           [],...
+           'objnoisy',...
+           'objnoisy',...
+           'objnoisy',...
+           []};     
+       
+% define all constraint fun handles
+allconsh = {'consnoisyf1',...
+            [],...
+            'consnoisyf3',...
+            'consnoisyf2',...
+            'consnoisyf4',...
+            []};
+% define all rhs for nl cons
+allnlrhs = {[],[],0,0,0,[]};
+% define cons type for nl cons
+allnle = {[],[],0,0,0,[]};       
+
+% problem defn
+setup_opts = optimdata;
+setup_opts.obj = allobjh;
+setup_opts.nlcons = allconsh;
+setup_opts.nlrhs = allnlrhs;
+setup_opts.nle = allnle;
+
+[prob,optimdata] = setup_optim_prob(setup_opts);
+
+% initial values for consrained nl(or quadratic?) optimization
+x0 = getrandomivals(optimdata,.1,1000);
+solveropt = struct('solver','ipopt','multi',0);
+optsol = choose_nlconsopt(prob,x0,optimdata,solveropt);
+
+% combine results for comparison plot
[…]

### 2017-08-04 KineticModel/estimation/bounds/boundsf1.m
new fun to set bounds separately for each flux parameter

@@ -1,0 +1,18 @@
+% change lb and ub for specific parametyers for specific fluxes
+function [new_lb,new_ub] = boundsf1(lb,ub,data)
+
+nc = data.nc;
+npert = data.npert;
+p_id = data.p_id;
+
+% parameter for flux 1
+new_lb = lb;
+new_ub = ub;
+
+% set separate bounds for Km for flux 1
+new_lb(nc*npert+1:nc*npert+1) = .03*ones(1,1); 
+new_ub(nc*npert+1:nc*npert+1) = 1*ones(1,1);
+% set separate bounds for kcat for flux 1
+new_lb(nc*npert+2:nc*npert+length(p_id)) = .1*ones(1,1); 
+new_ub(nc*npert+2:nc*npert+length(p_id)) = 2*ones(1,1);
+

### 2017-08-04 KineticModel/estimation/optim_scripts/setup_optim_prob.m

@@ -61,18 +61,28 @@ end
 % set bounds - concentration
 lb(1:nc*npert) = xss_exp_v.*(1-eps);
 ub(1:nc*npert) = xss_exp_v.*(1+eps);
-% set bounds - parameter
+% set general bounds - parameter - specific bounds set below
 lb(nc*npert+1:nc*npert+length(p_id)) = .05*ones(length(p_id),1); 
 ub(nc*npert+1:nc*npert+length(p_id)) = 5*ones(length(p_id),1);
-% set separate bounds for Km
+% % set separate bounds for Km for flux 1
 % lb(nc*npert+1:nc*npert+1) = .05*ones(1,1); 
 % ub(nc*npert+1:nc*npert+1) = 1*ones(1,1);
-% set separate bounds for kcat
+% % set separate bounds for kcat for flux 1
 % lb(nc*npert+2:nc*npert+length(p_id)) = .1*ones(1,1); 
 % ub(nc*npert+2:nc*npert+length(p_id)) = 3*ones(1,1);
 % set bounds - flux
 lb(nvar-nf*npert+1:nvar) = vss_exp_v*(1-eps);
 ub(nvar-nf*npert+1:nvar) = vss_exp_v*(1+eps);
+
+% setup fresh bounds specifically for each flux parameter dependent on
+% flxid
+if isfield(optimdata,'bounds')
+    boundhs = optimdata.bounds;
+end
+if ~isempty(boundhs{flxid})
+    boundfh = str2func(boundhs{flxid});
+    [lb,ub] = boundfh(lb,ub,newdata);
+end
 
 % setup objectives and cons
 if isfield(optimdata,'obj')

### 2017-08-04 KineticModel/estimation/scripts-other/gen_nonoise.m

@@ -20,12 +20,19 @@ pss = ones(1,numel(exp_sol.exp_pval));
 % pss(exp_sol.xss(1,:)>exp_sol.xss(2,:)) = 0;    
 
 % options structure for solving problem
-optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',1,'eps',.2,...
+optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',1,'eps',.18,...
                     'vexp',exp_sol.fss(:,logical(pss)),...
                     'xexp',exp_sol.xss(:,logical(pss)),...
                     'odep',odep_bkp,...
                     'wt_xss',xss(:,1),'wt_fss',fss(:,1));
                 
+% define all bound fun handles
+allboundh = {'boundsf1',...
+             [],...
+             [],...
+             [],...
+             [],...
+             []};
 % define all objectiv efun handles
 allobjh = {'objnoisy',...
            [],...
@@ -48,6 +55,7 @@ allnle = {[],[],0,0,0,[]};
 
 % problem defn
 setup_opts = optimdata;
+setup_opts.bounds = allboundh;
 setup_opts.obj = allobjh;
 setup_opts.nlcons = allconsh;
 setup_opts.nlrhs = allnlrhs;
@@ -56,7 +64,7 @@ setup_opts.nle = allnle;
 [prob,optimdata] = setup_optim_prob(setup_opts);
 
 % initial values for consrained nl(or quadratic?) optimization
-x0 = getrandomivals(optimdata,.1,1000);
+x0 = getrandomivals(optimdata,.2,1000);
 solveropt = struct('solver','ipopt','multi',0);
 optsol = choose_nlconsopt(prob,x0,optimdata,solveropt);
 

### 2017-08-04 KineticModel/estimation/docs/formulation/ident.tex
write-up w/ figures on identifiability analysis

@@
 
-	We can use a profile likelihood-based approach to determine the structural and practical identifiability of parameters in any any dynamic model.
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.8\textwidth,height=.8\textheight,keepaspectratio]{figures/identifiability}}
+		\caption{The profile likelihood estimates of a) a structurally non-identifiable, b) a practically non-identifiable and c) an identifiable parameter.
+The practically identifiability of a parameter is dependent on the quality and quantity of experimental data used to estimate the parameter.}\label{fig:ident-pl}
+	\end{figure}
 
-	We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.
+	We can use a profile likelihood-based approach to determine the structural and practical identifiability of parameters in any dynamic model.
+The identifiability of parameters is established through the confidence intervals of their estimates, $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
+The confidence intervals of a structurally non-identifiable parameter are unbounded, i.e., $\left[-\infty, +\infty\right]$, while the confidence intervals of a practically non-identifiable parameter are unbounded in at least one direction, i.e., $\left[\sigma_{i}^-,\sigma_{i}^+\right]$ where either $\sigma_{i}^- = -\infty$ or $\sigma_{i}^+ = +\infty$.
+If a parameter's estimates have a finite confidence interval then the parameter is said to be identifiable.
+The visualization of structurally and practically non-identifiable parameters using the profile likelihood approach is illustrated in Figure \ref{fig:ident-pl}.
 
@@
 	\end{figure}
+
+	Due to the dependence of practical parameter identifiability on the experimental data, the profile likelihood approach can be used to design experiments in such a way that the observables that are derived from these experiments can improve the practical identifiability of the parameters.
+We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.
 
@@
 	\end{align}
+
+	\subsubsection{Identifiability analysis of parameters in a kinetic model of gluconeogenesis:}
+	The proposed model for acetate consumption through gluconeogenesis and its corresponding kinetic model is used as a case study to illustrate the utility of identifiability analysis for the design of experiments for estimating parameters in kinetic models of metabolism.
+The kinetic model is described below.
+	\begin{equation}\label{eq:ode1}
+	\frac{d}{dt}pep=v_1-v_2-v_4
+	\end{equation}
+	\begin{equation}\label{eq:ode2}
+	\frac{d}{dt}fdp=v_2-v_3
+	\end{equation}
+	\begin{equation}\label{eq:ode3}
+	\frac{d}{dt}E=v_{e,max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right) - d E
+	\end{equation}
+	The kinetic expressions for fluxes $v_1$ through $v_4$ are given below.
+The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics.
+The acetate flux through $v_1$ is also governed by the quantity of available enzyme E.
+	\begin{equation}\label{eq:flux1}
+	v_1 = k_{1}^{cat}E\frac{acetate}{acetate+K_{1}^{acetate}}
+	\end{equation}	
+	\begin{equation}\label{eq:flux2}
+	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
+	\end{equation}
+	\begin{equation}\label{eq:flux3}
+	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
+	\end{equation}
+	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
+The added flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
+	\begin{equation}\label{eq:flux4}
+	v_4 = k_{4}^{cat}.pep
+	\end{equation}
 \end{document}

### 2017-08-04 KineticModel/estimation/optim_scripts/setup_optim_prob.m
remove commented portions

@@ -64,12 +64,6 @@ ub(1:nc*npert) = xss_exp_v.*(1+eps);
 % set general bounds - parameter - specific bounds set below
 lb(nc*npert+1:nc*npert+length(p_id)) = .05*ones(length(p_id),1); 
 ub(nc*npert+1:nc*npert+length(p_id)) = 5*ones(length(p_id),1);
-% % set separate bounds for Km for flux 1
-% lb(nc*npert+1:nc*npert+1) = .05*ones(1,1); 
-% ub(nc*npert+1:nc*npert+1) = 1*ones(1,1);
-% % set separate bounds for kcat for flux 1
-% lb(nc*npert+2:nc*npert+length(p_id)) = .1*ones(1,1); 
-% ub(nc*npert+2:nc*npert+length(p_id)) = 3*ones(1,1);
 % set bounds - flux
 lb(nvar-nf*npert+1:nvar) = vss_exp_v*(1-eps);
 ub(nvar-nf*npert+1:nvar) = vss_exp_v*(1+eps);

### 2017-08-04 KineticModel/estimation/objectives/objnoisy.m
remove commented portions

@@ -1,12 +1,5 @@
 % objective to minimize the L2-norm of fluxes (quadratic?)
 function fx = objnoisy(x,p,data)
 
-% assign parameter values from variable vector
-% p(data.p_id) = x(data.nc+1:end);
-
-% calculate flux
-% vest = getmodelflux(x(1:data.nc),p);
-% vest(1) = x(5).*x(3).*p(17)./(p(17)+x(4));
-
 % calc objective
 fx = sqrt(sum((x(data.nvar-data.nf*data.npert+1:data.nvar)-data.vexp).^2));

### 2017-08-04 KineticModel/estimation/scripts-other/est_flux2_nn.m
no noise flux2 estimation

@@ -1,0 +1,83 @@
+% estimate fluxes for flux 2 in kotte model pep ---> fdp
+
+% generate experimental data - get initial ss
+tspan = 0:0.1:100;
+[xdyn,fdyn,xss,fss,opts] = run_nonoise(tspan);
+
+% backup parameters and initial conditions
+ival_bkp = opts.x0;
+odep_bkp = opts.odep;
+
+% generate nsmp samples by adding random noise to ss values
+nsmp = 10;
+[noisy_xss,noisy_fss] = addnoise(repmat(xss,1,nsmp),odep_bkp);
+
+% perturb system from non-noisy initial conditions
+pt_sol_id = [1 2 3];
+[exp_sol,no_noise_sol] = dopert_nonoise(opts,xss,fss,odep_bkp,pt_sol_id);
+close all
+
+% get only data from one steady state
+pss = ones(1,numel(exp_sol.exp_pval));
+% pss(exp_sol.xss(1,:)>exp_sol.xss(2,:)) = 0;    
+
+% options structure for solving problem
+optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',4,'eps_v',.1,...
+                    'eps_c',.9,'vexp',exp_sol.fss(:,logical(pss)),...
+                    'xexp',exp_sol.xss(:,logical(pss)),...
+                    'odep',odep_bkp,...
+                    'wt_xss',xss(:,1),'wt_fss',fss(:,1));
+                
+% define all bound fun handles
+allboundh = {'boundsf1',...
+             [],...
+             [],...
+             'boundsf2',...
+             [],...
+             []};
+% define all objectiv efun handles
+allobjh = {'objnoisy',...
+           [],...
+           'objnoisy',...
+           'objnoisy',...
+           'objnoisy',...
+           []};
+% define all constraint fun handles
+allconsh = {'consnoisyf1',...
+            [],...
+            'consnoisyf1',...
+            'consnoisyf2',...
+            'consnoisyf1',...
+            []};
+% define all rhs for nl cons
+allnlrhs = {0,[],0,[],0,[]};
+% define cons type for nl cons
+allnle = {0,[],0,[],0,[]};
+
+% problem defn
+setup_opts = optimdata;
+setup_opts.bounds = allboundh;
+setup_opts.obj = allobjh;
+setup_opts.nlcons = allconsh;
+setup_opts.nlrhs = allnlrhs;
+setup_opts.nle = allnle;
+
[…]

### 2017-08-04 KineticModel/estimation/scripts-other/est_flux1_nn.m
no noise counterparts

@@ -1,0 +1,80 @@
+% script to estimate parameters from noisy data for flux 1 in kotte model
+
+% generate experimental data - get initial ss
+tspan = 0:0.1:100;
+[xdyn,fdyn,xss,fss,opts] = run_nonoise(tspan);
+
+% backup parameters and initial conditions
+ival_bkp = opts.x0;
+odep_bkp = opts.odep;
+
+% perturb system from non-noisy initial conditions
+pt_sol_id = [1 2 3];
+[exp_sol,no_noise_sol] = dopert_nonoise(opts,xss,fss,odep_bkp,pt_sol_id);
+close all
+
+% get only data from one steady state
+pss = ones(1,numel(exp_sol.exp_pval));
+% pss(2) = 0;
+% pss(exp_sol.xss(1,:)>exp_sol.xss(2,:)) = 0;    
+
+% options structure for solving problem
+optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',1,'eps_v',.05,...
+                    'eps_c',.3,'vexp',exp_sol.fss(:,logical(pss)),...
+                    'xexp',exp_sol.xss(:,logical(pss)),...
+                    'odep',odep_bkp,...
+                    'wt_xss',xss(:,1),'wt_fss',fss(:,1));
+
+% define all bound fun handles
+allboundh = {'boundsf1',...
+             [],...
+             [],...
+             [],...
+             [],...
+             []};
+% define all objectiv efun handles
+allobjh = {'objnoisy',...
+           [],...
+           'objnoisy',...
+           'objnoisy',...
+           'objnoisy',...
+           []};
+% define all constraint fun handles
+allconsh = {'consnoisyf1',...
+            [],...
+            'consnoisyf1',...
+            'consnoisyf1',...
+            'consnoisyf1',...
+            []};
+% define all rhs for nl cons
+allnlrhs = {[],[],0,0,0,[]};
+% define cons type for nl cons
+allnle = {[],[],0,0,0,[]};
+
+% problem defn
+setup_opts = optimdata;
+setup_opts.bounds = allboundh;
+setup_opts.obj = allobjh;
+setup_opts.nlcons = allconsh;
+setup_opts.nlrhs = allnlrhs;
+setup_opts.nle = allnle;
+
+[prob,optimdata] = setup_optim_prob(setup_opts);
+
+% initial values for consrained nl(or quadratic?) optimization
+x0 = getrandomivals(optimdata,.2,1000);
[…]

### 2017-08-04 KineticModel/estimation/scripts-other/est_flux3_nn.m

@@ -1,0 +1,79 @@
+% estimate fluxes for flux 3 in kotte model fdp ---> 
+
+% generate experimental data - get initial ss
+tspan = 0:0.1:100;
+[xdyn,fdyn,xss,fss,opts] = run_nonoise(tspan);
+
+% backup parameters and initial conditions
+ival_bkp = opts.x0;
+odep_bkp = opts.odep;
+
+% perturb system from non-noisy initial conditions
+pt_sol_id = [1 2 3];
+[exp_sol,no_noise_sol] = dopert_nonoise(opts,xss,fss,odep_bkp,pt_sol_id);
+close all
+
+% get only data from one steady state
+pss = ones(1,numel(exp_sol.exp_pval));
+% pss(exp_sol.xss(1,:)>exp_sol.xss(2,:)) = 0;    
+
+% options structure for solving problem
+optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',3,'eps_v',.05,...
+                    'eps_c',.5,'vexp',exp_sol.fss(:,logical(pss)),...
+                    'xexp',exp_sol.xss(:,logical(pss)),...
+                    'odep',odep_bkp,...
+                    'wt_xss',xss(:,1),'wt_fss',fss(:,1));
+                
+% define all bound fun handles
+allboundh = {'boundsf1',...
+             [],...
+             'boundsf3',...
+             [],...
+             [],...
+             []};
+% define all objectiv efun handles
+allobjh = {'objnoisy',...
+           [],...
+           'objnoisy',...
+           'objnoisy',...
+           'objnoisy',...
+           []};
+% define all constraint fun handles
+allconsh = {'consnoisyf1',...
+            [],...
+            'consnoisyf3',...
+            'consnoisyf2',...
+            'consnoisyf1',...
+            []};
+% define all rhs for nl cons
+allnlrhs = {0,[],[],0,0,[]};
+% define cons type for nl cons
+allnle = {0,[],[],0,0,[]};
+
+% problem defn
+setup_opts = optimdata;
+setup_opts.bounds = allboundh;
+setup_opts.obj = allobjh;
+setup_opts.nlcons = allconsh;
+setup_opts.nlrhs = allnlrhs;
+setup_opts.nle = allnle;
+
+[prob,optimdata] = setup_optim_prob(setup_opts);
+
+% initial values for consrained nl(or quadratic?) optimization
+x0 = getrandomivals(optimdata,.2,1000);
[…]

### 2017-08-04 KineticModel/estimation/scripts-other/est_flux2.m
no major changes

@@ -27,8 +27,8 @@ pss = ones(1,numel(exp_sol.exp_pval));
 % pss(exp_sol.xss(1,:)>exp_sol.xss(2,:)) = 0;    
 
 % options structure for solving problem
-optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',4,'eps_v',.05,...
-                    'eps_c',.3,'vexp',exp_sol.fss(:,logical(pss)),...
+optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',4,'eps_v',.1,...
+                    'eps_c',.9,'vexp',exp_sol.fss(:,logical(pss)),...
                     'xexp',exp_sol.xss(:,logical(pss)),...
                     'odep',odep_bkp,...
                     'wt_xss',xss(:,1),'wt_fss',fss(:,1));
@@ -70,7 +70,7 @@ setup_opts.nle = allnle;
 [prob,optimdata] = setup_optim_prob(setup_opts);
 
 % initial values for consrained nl(or quadratic?) optimization
-x0 = getrandomivals(optimdata,.1,5000);
+x0 = getrandomivals(optimdata,.2,1000);
 solveropt = struct('solver','ipopt','multi',0);
 optsol = choose_nlconsopt(prob,x0,optimdata,solveropt);
 

### 2017-08-04 KineticModel/estimation/scripts-other/est_flux3.m

@@ -70,7 +70,7 @@ setup_opts.nle = allnle;
 [prob,optimdata] = setup_optim_prob(setup_opts);
 
 % initial values for consrained nl(or quadratic?) optimization
-x0 = getrandomivals(optimdata,.1,5000);
+x0 = getrandomivals(optimdata,.1,1000);
 solveropt = struct('solver','ipopt','multi',0);
 optsol = choose_nlconsopt(prob,x0,optimdata,solveropt);
 

### 2017-08-05 KineticModel/estimation/docs/formulation/ident.tex
no major change
add all estimation formulations

@@
 	\end{equation}
+	Simulated data
+
+	\subsection{Parameter estimation in kinetic models of metabolism:}
+	The formulation for parameter estimation assumes that the concentrations and the fluxes corresponding to each perturbation are variables in the optimization problem formulated to estimate kinetic parameters.
+
+	Let $\mathbf{x}\in\mathbb{R}^m$, $\mathbf{p}\in\mathbb{R}^l$ and $\mathbf{v}\in\mathbb{R}$ be the vector of concentrations, parameters and fluxes respectively.
+If the number of perturbations used for parameter estimation is in $\mathbb{R}^p$, then the vector of concentrations, fluxes and parameters that need to be estimated in the optimization problem changes to $\mathbf{x}\in\mathbb{R}^{mp}, \mathbf{v}\in\mathbb{R}^{p}$ and $\mathbf{p}\in\mathbb{R}^{l}$ respectively.
+
+	Let $i\in\mathbb{R}^{mp}$, $j\in\mathbb{R}^{p}$ and $k\in\mathbb{R}^{l}$ represent the indices for concentrations, fluxes and parameters respectively in Equation (\ref{eq:l2norm-flux}).
+
+	\begin{center}
+		\begin{subequations}\label{eq:l2norm-flux}
+			\begin{align}
+			\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert\\
+			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+			& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+			& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+			\end{align}
+		\end{subequations}
+	\end{center}
+
+	Due to the inability of the above formulation to get feasible results, we modified the problem to include the L2-norm of the concentrations as well in the objective function in Equation (\ref{eq:free-var}a).
+In Equation (\ref{eq:free-var}) the flux and the concentrations are free variables, i.e.,  $\mathbf{x}\in\left[\mathbf{0},+\mathbf{\infty}\right]$.
+	\begin{center}
+		\begin{subequations}\label{eq:free-var}
+			\begin{align}
+			\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
+			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+			& \text{      }0\le x_i \le +\infty & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+			& \text{      }0\le v_j \le +\infty & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+			\end{align}
+		\end{subequations}
+	\end{center}
+After testing the problem in Equation (\ref{eq:free-var}) above for the acetate uptake flux $v_1$, we found that all solutions using IPOPT exceeded the maximum number of iterations (set at 5000) and never produced an optimal solution.
+
+	We further also test a formulation in Equation (\ref{eq:l2norm-conc}) where the flux and concentrations are constrained as earlier with a noise term.	
+	\begin{center}
+		\begin{subequations}\label{eq:l2norm-conc}
+			\begin{align}
+			\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
+			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+			& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+			& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+			\end{align}
+		\end{subequations}
+	\end{center}
+
+	A third formulation where the noise terms ($\epsilon$) used as bounds for the flux and the concentrations are included in the optimization problem as a variable to be penalized is given in Equation (\ref{eq:noise-var}) below.
+
+	\begin{center}
+		\begin{subequations}\label{eq:noise-var}
+			\begin{align}
+			\underset{\mathbf{x},\mathbf{p},\mathbf{v},\epsilon}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert + w_3\epsilon^2\\
+			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+			& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+			& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+			\end{align}
+		\end{subequations}
+	\end{center}
+
 \end{document}

### 2017-08-07 KineticModel/estimation/scripts-other/est_flux3.m
change objs and cons and add figure codes

@@ -27,32 +27,35 @@ pss = ones(1,numel(exp_sol.exp_pval));
 % pss(exp_sol.xss(1,:)>exp_sol.xss(2,:)) = 0;    
 
 % options structure for solving problem
-optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',3,'eps_v',.05,...
-                    'eps_c',.3,'vexp',exp_sol.fss(:,logical(pss)),...
+optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',3,'eps_v',1,...
+                    'eps_c',1,'vexp',exp_sol.fss(:,logical(pss)),...
                     'xexp',exp_sol.xss(:,logical(pss)),...
+                    'flux_wt',1000,'conc_wt',100,...
+                    'eps_c_wt',1000,'eps_v_wt',10000,...
                     'odep',odep_bkp,...
-                    'wt_xss',xss(:,1),'wt_fss',fss(:,1));
+                    'wt_xss',noisy_xss(:,1),'wt_fss',noisy_fss(:,1),...
+                    'type',2);
                 
 % define all bound fun handles
-allboundh = {'boundsf1',...
+allboundh = {[],...
              [],...
              'boundsf3',...
              [],...
              [],...
              []};
 % define all objectiv efun handles
-allobjh = {'objnoisy',...
+allobjh = {[],...
            [],...
-           'objnoisy',...
-           'objnoisy',...
-           'objnoisy',...
+           'obj_typec',...
+           [],...
+           [],...
            []};
 % define all constraint fun handles
-allconsh = {'consnoisyf1',...
+allconsh = {[],...
             [],...
-            'consnoisyf3',...
-            'consnoisyf2',...
-            'consnoisyf1',...
+            'cons_typeb_f3',...
+            [],...
+            [],...
             []};
 % define all rhs for nl cons
 allnlrhs = {0,[],[],0,0,[]};
@@ -78,10 +81,25 @@ optsol = choose_nlconsopt(prob,x0,optimdata,solveropt);
 est_data = combine_results(optsol,opts,noisy_sol,optimdata,pss,pss);
 
 % compare fluxes and concentrations
-compare_vals(est_data,noisy_sol,optimdata,opts,pss);
+hfcv = compare_vals(est_data,noisy_sol,optimdata,opts,pss);
 
[…]

### 2017-08-07 KineticModel/estimation/scripts-other/est_flux2.m
change obj weights

@@ -30,8 +30,8 @@ pss = ones(1,numel(exp_sol.exp_pval));
 optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',4,'eps_v',1,...
                     'eps_c',1,'vexp',exp_sol.fss(:,logical(pss)),...
                     'xexp',exp_sol.xss(:,logical(pss)),...
-                    'flux_wt',100,'conc_wt',100,...
-                    'eps_c_wt',10,'eps_v_wt',10,...
+                    'flux_wt',1000,'conc_wt',1000,...
+                    'eps_c_wt',100,'eps_v_wt',100,...
                     'odep',odep_bkp,...
                     'wt_xss',noisy_xss(:,1),'wt_fss',noisy_fss(:,1),...
                     'type',2);
@@ -89,19 +89,19 @@ hfcv = compare_vals(est_data,noisy_sol,optimdata,opts,pss);
 hfp = compare_pars(est_data);
 
 % save figure files
-dir = 'C:\Users\shyam\Documents\Courses\CHE1125Project\Results\estimation\est_flux2\noisy\typec\';
-set(0,'CurrentFigure',hfcv(1));
-fname = 'est_flux2_conc_aug6';
-print([dir fname],'-depsc','-painters','-loose','-tiff','-r200');
-close(hfcv(1));
-set(0,'CurrentFigure',hfcv(2));
-fname = 'est_flux2_flux_aug6';
-print([dir fname],'-depsc','-painters','-loose','-tiff','-r200');
-close(hfcv(2));
-set(0,'CurrentFigure',hfp);
-fname = 'est_flux2_par_aug6';
-print([dir fname],'-depsc','-painters','-loose','-tiff','-r200');
-close(hfp);
+% dir = 'C:\Users\shyam\Documents\Courses\CHE1125Project\Results\estimation\est_flux2\noisy\typec\';
+% set(0,'CurrentFigure',hfcv(1));
+% fname = 'est_flux2_conc_aug6';
+% print([dir fname],'-depsc','-painters','-loose','-tiff','-r200');
+% close(hfcv(1));
+% set(0,'CurrentFigure',hfcv(2));
+% fname = 'est_flux2_flux_aug6';
+% print([dir fname],'-depsc','-painters','-loose','-tiff','-r200');
+% close(hfcv(2));
+% set(0,'CurrentFigure',hfp);
+% fname = 'est_flux2_par_aug6';
[…]

### 2017-08-07 KineticModel/estimation/docs/formulation/ident.tex
add figures and optimization formulations

@@
 \usepackage{longtable}
+\usepackage{multirow}
 \usepackage[margin=1.0in]{geometry}
@@
 
-	Let $i\in\mathbb{R}^{mp}$, $j\in\mathbb{R}^{p}$ and $k\in\mathbb{R}^{l}$ represent the indices for concentrations, fluxes and parameters respectively in Equation (\ref{eq:l2norm-flux}).
+	Let $i\in\mathbb{R}^{mp}$, $j\in\mathbb{R}^{p}$ and $k\in\mathbb{R}^{l}$ represent the indices for concentrations, fluxes and parameters respectively in Equations (\ref{eq:l2norm-flux})-(\ref{eq:noise-var}).
 
-	\begin{center}
+	\subsubsection{Formulation a):}		
 	\begin{subequations}\label{eq:l2norm-flux}
@@
 	\end{subequations}
+
+	\subsubsection{Formulation b):}	
+	In this formulation, in Equation (\ref{eq:l2norm-conc}), the flux and concentrations are constrained as earlier with a noise term.	
+	\begin{subequations}\label{eq:l2norm-conc}
+		\begin{align}
+		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
+		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+		\end{align}
+	\end{subequations}
+
+	\subsubsection{Formulation c):}
+	A third formulation where the noise terms ($\epsilon$) used as bounds for the flux and the concentrations are included in the optimization objective as a variable to be penalized is given in Equation (\ref{eq:noise-var}) below.
+Notice that the noise term has been separated into two components, $\epsilon_x$ and $\epsilon_v$, corresponding to noise in the concentrations and the fluxes, respectively.
+$w_3$ and $w_4$ are the weights associated with the noise terms in the objective function.
+	\begin{center}
+		\begin{subequations}\label{eq:noise-var}
+			\begin{align}
+			\underset{\mathbf{x},\mathbf{p},\mathbf{v},\epsilon_x,\epsilon_v}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert + w_3\epsilon_x + w_4\epsilon_v\\
+			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+			& \text{      }x^*_i(1-\epsilon_x)\le x_i \le x^*_i(1+\epsilon_x) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+			& \text{      }v^*_j(1-\epsilon_v)\le v_j \le v^*_j(1+\epsilon_v) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+			\end{align}
+		\end{subequations}
 	\end{center}
 
-	Due to the inability of the above formulation to get feasible results, we modified the problem to include the L2-norm of the concentrations as well in the objective function in Equation (\ref{eq:free-var}a).
+	\subsubsection{Formulation d):}
+	Due to the inability of Equation (\ref{eq:l2norm-flux}) to get feasible results, we modified the problem to include the L2-norm of the concentrations as well in the objective function in Equation (\ref{eq:free-var}a).
 In Equation (\ref{eq:free-var}) the flux and the concentrations are free variables, i.e.,  $\mathbf{x}\in\left[\mathbf{0},+\mathbf{\infty}\right]$.
-	\begin{center}
 	\begin{subequations}\label{eq:free-var}
@@
 	\end{subequations}
+
+After testing the problem in Equation (\ref{eq:free-var}) above for the acetate uptake flux $v_1$, we found that all solutions using IPOPT exceeded the maximum number of iterations (set at 5000) and never produced an optimal solution.
+Hence, we use a modified form of the bounds in Equation (\ref{eq:free-var}) above in Equation (\ref{eq:l2norm-conc}).
+
+	\begin{table}[!tbhp]
+		\caption{Table showing the perturbed values of all fluxes used for parameter estimation.}
+		\begin{center}				
+			\begin{tabular}{ccc}
+				Designation & Perturbed Fluxes & Perturbed Values\\
+				\hline
+				P1 & $v_1$ & 2\\
+				P2 & $v_2$ & 0.2\\
+				P3 & $v_3$ & 0.5
+			\end{tabular}
 		\end{center}	
-After testing the problem in Equation (\ref{eq:free-var}) above for the acetate uptake flux $v_1$, we found that all solutions using IPOPT exceeded the maximum number of iterations (set at 5000) and never produced an optimal solution.
+		\label{tab:pval}
+	\end{table}
 
-	We further also test a formulation in Equation (\ref{eq:l2norm-conc}) where the flux and concentrations are constrained as earlier with a noise term.	
+	\section{Results:}
+
+	\begin{table}[!tbhp]
+		\caption{Table showing the ability of different formulations mentioned above to estimate enzyme kinetic parameters for different fluxes based on the perturbations given in the above table.}
 		\begin{center}				
-		\begin{subequations}\label{eq:l2norm-conc}
-			\begin{align}
-			\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
-			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-			& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-			& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-			\end{align}
-		\end{subequations}
+			\begin{tabular}{cccccc}
+				\hline
+				Flux & Data Type & \multicolumn{2}{c}{Formulation (b)} & \multicolumn{2}{c}{Formulation (c)}\\
+				\hline
+				&&$\epsilon_x$(\%)&$\epsilon_v$(\%)&$\epsilon_x$(\%)&$\epsilon_v$(\%)\\
+				\hline
+				\multirow{2}{*}{flux 1} & no noise & 30 & 5 & 33.33 & 0\\
+												  &	noisy & x & x & 0 & 33.33\\
+			  	\hline
+				 \multirow{2}{*}{flux 2} & no noise & 90 & 30 & 0 & 31\\
+				 									& noisy & 85 & 40 & 0 & 30\\
+				\hline				 									
+				\multirow{2}{*}{flux 3} & no noise & 15 & 10 & 17 & 0\\
+													& noisy & x & x & 17 & 0\\
+				\hline			
+			\end{tabular}
 		\end{center}	
+		\label{tab:res1}
+	\end{table}
+	\subsection{Estimation of flux 1:}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_conc_aug6}}
+		\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_conc_aug6}}
+		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_flux_aug6}}
+		\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_flux_aug6}}
+		\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_par_aug6}}
+		\caption{Plot comparing noiseless model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_par_aug6}}
+		\caption{Plot comparing noisy model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
+	\end{figure}
+	\clearpage
 
-	A third formulation where the noise terms ($\epsilon$) used as bounds for the flux and the concentrations are included in the optimization problem as a variable to be penalized is given in Equation (\ref{eq:noise-var}) below.
+	\subsection{Estimation of flux 2:}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_conc_aug6}}
+		\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_conc_aug6}}
+		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_flux_aug6}}
+		\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_flux_aug6}}
+		\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_par_aug6}}
+		\caption{Plot comparing noiseless model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_par_aug6}}
+		\caption{Plot comparing noisy model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
+	\end{figure}
+	\clearpage
 
-	\begin{center}
-		\begin{subequations}\label{eq:noise-var}
-			\begin{align}
-			\underset{\mathbf{x},\mathbf{p},\mathbf{v},\epsilon}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert + w_3\epsilon^2\\
-			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-			& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-			& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-			\end{align}
-		\end{subequations}
-	\end{center}
+	\subsection{Estimation of flux 3:}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_conc_aug6}}
+		\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_conc_aug6}}
+		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_flux_aug6}}
+		\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_flux_aug6}}
+		\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_par_aug6}}
+		\caption{Plot comparing noiseless model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
+	\end{figure}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_par_aug6}}
+		\caption{Plot comparing noisy model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
+	\end{figure}
+	\clearpage
 

### 2017-08-16 KineticModel/estimation/docs/formulation/ident.tex
updated document with corrections from prof. cluett

@@
 
+%\usepackage[font=small,labelfont=bf,labelsep=period]{caption}
+\usepackage[style=authoryear,sorting=nyt,url=false,isbn=false,doi=false,firstinits=true,backend=biber]{biblatex}
+
+\renewcommand{\baselinestretch}{2.0}
+\renewcommand*\contentsname{Table of Contents}
+%\captionsetup{font={stretch=2.0}}
+
+\DeclareNameAlias{default}{last-first}
+
+\DefineBibliographyStrings{english}{%
+	andothers = {\addcomma\addspace\textsc{et\addabbrvspace al}\adddot},
+	and = {\textsc{and}}
+}
+\renewcommand*{\labelnamepunct}{\space\space}
+
+\renewbibmacro{in:}
+{%
+	\ifentrytype{article}{%
+	}{%
+		\printtext{\bibstring{in}\intitlepunct}%
+	}%
+}
+\renewbibmacro*{volume+number}{%
+	\printfield{volume}%
+	\setunit*{\addcomma\space}%
+	\printfield{number}%
+	\setunit{\addcomma\space}}
+
+\DeclareFieldFormat{pages}{#1}
+
+\renewbibmacro*{publisher+location+date}{%
+	\printlist{publisher}%
+	\setunit*{\addcomma\space}%
+	\printlist{location}%
+	\setunit*{\addcomma\space}%
+	\usebibmacro{date}%
+	\newunit}
+
+\renewcommand{\newunitpunct}{\addcomma\space}
+\DeclareFieldFormat[article,inbook,incollection,inproceedings,patent,thesis,unpublished]{title}{#1} 
+\DeclareFieldFormat{year}{#1}
+
+\addbibresource{refs/ident_refs.bib}
+
 \begin{document}
@@
 	\subsection{Identifiability analysis: Definitions and Formulations}
-	The identifiability of parameters in nonlinear models of physical processes can be classified into two categories: structural and practical identifiability.
+	Any nonlinear dynamical system can be represented by a set of states $\mathbf{x}$, observables $\mathbf{y}$ that are dependent on the states, parameters $\mathbf{\mu}$, and inputs $u$ as in Equation (\ref{eq:dynmodel}).
+	\begin{subequations}\label{eq:dynmodel}
+		\begin{align}
+		\dot{\mathbf{x}} = g(\mathbf{x},\mathbf{\mu},u)\\
+		\mathbf{y} = h(\mathbf{x},\mathbf{\mu},u)
+		\end{align}
+	\end{subequations}
+	Identifiability concerns with the ability to determine a unique solution to the problem of estimating parameters $\mathbf{\mu}$ from given data on the system observables $\mathbf{y}$ for inputs $u$ \parencite{McLean2012}.	The identifiability of parameters in nonlinear models of physical processes can be classified into two categories: structural and practical identifiability.
+
+	Any system (Equation \ref{eq:dynmodel}) is said to be structurally or a priori identifiable if, for an input-output mapping defined by $\mathbf{y} = \Phi(\mu,u)$ for at least one input function $u$, any two values of parameters $\mathbf{\mu}_1$ and $\mathbf{\mu}_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
+	\begin{align}\label{eq:stident}
+	\Phi(\mathbf{\mu}_1,u) = \Phi(\mathbf{\mu}_2,u) \iff \mathbf{\mu}_1 = \mathbf{\mu}_2
+	\end{align}
+	Accordingly, any system that has an infinite number of solutions to the parameter estimation problem for all input functions is said to structurally non-identifiable.
 The effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
+
+	When parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
 The effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
-Practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
-The aforementioned distinction between practical and structural identifiability of parameters is illustrated in Figure.
-Thus, the identifiability of parameters in nonlinear models is dependent on the model structure, parameterization, and the quality and quantity of experimental data that is available for the purpose of estimation.
+Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
+	%Thus, the identifiability of parameters in nonlinear models is dependent on the model structure, parameterization, and the quality and quantity of experimental data that is available for the purpose of estimation.
+	Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated.
+On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
 
-	On the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated.
-On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
+	A profile likelihood-based approach can be used to determine both the structural and practical identifiability of parameters in nonlinear dynamical models \parencite{Raue2009a}.
+The identifiability of parameters is established through the confidence intervals of their estimates, $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
+We provide the methods to estimate these confidence intervals below.
+The visualization of structurally and practically non-identifiable parameters using the profile likelihood approach is illustrated in Figure \ref{fig:identtypes}.
+The confidence intervals of a structurally non-identifiable parameter are unbounded, i.e., $\left[-\infty, +\infty\right]$ (Figure \ref{fig:identtypes}a), while the confidence intervals of a practically non-identifiable parameter are unbounded in at least one direction, i.e., $\left[\sigma_{i}^-,\sigma_{i}^+\right]$ where either $\sigma_{i}^- = -\infty$ or $\sigma_{i}^+ = +\infty$ (Figure \ref{fig:identtypes}b).
+If a parameter's estimates have a finite confidence interval then the parameter is said to be identifiable (Figure \ref{fig:identtypes}c).
+Note that the horizontal dotted lines in Figure \ref{fig:identtypes} represent the confidence interval thresholds ($\Delta_{\alpha}$) that are used to establish identifiability.
 
@@
 		\caption{The profile likelihood estimates of a) a structurally non-identifiable, b) a practically non-identifiable and c) an identifiable parameter.
+The horizontal axis represents the changes in the value of the parameter ($\theta_i$) whose identifiability is being determined and the profile likelihood ($\chi^2_{PL}(\theta_i)$) is shown in the vertical axis.
+The confidence interval thresholds ($\Delta_{\alpha}$) used to determine the identifiability of the parameter are denoted by the horizontal dotted line.
 The practically identifiability of a parameter is dependent on the quality and quantity of experimental data used to estimate the parameter.}\label{fig:ident-pl}
-	\end{figure}
+	\end{figure}\label{fig:identtypes}
 
-	We can use a profile likelihood-based approach to determine the structural and practical identifiability of parameters in any dynamic model.
-The identifiability of parameters is established through the confidence intervals of their estimates, $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
-The confidence intervals of a structurally non-identifiable parameter are unbounded, i.e., $\left[-\infty, +\infty\right]$, while the confidence intervals of a practically non-identifiable parameter are unbounded in at least one direction, i.e., $\left[\sigma_{i}^-,\sigma_{i}^+\right]$ where either $\sigma_{i}^- = -\infty$ or $\sigma_{i}^+ = +\infty$.
-If a parameter's estimates have a finite confidence interval then the parameter is said to be identifiable.
-The visualization of structurally and practically non-identifiable parameters using the profile likelihood approach is illustrated in Figure \ref{fig:ident-pl}.
+	Due to the dependence of practical parameter identifiability on the experimental data, the profile likelihood approach can be used to design experiments in such a way that the observables that are derived from these experiments can improve the practical identifiability of the parameters.
+We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.
 
@@
 	\end{figure}
-
-	Due to the dependence of practical parameter identifiability on the experimental data, the profile likelihood approach can be used to design experiments in such a way that the observables that are derived from these experiments can improve the practical identifiability of the parameters.
-We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.
 
@@
 
+	\printbibliography
 \end{document}

### 2017-08-16 KineticModel/estimation/ident/pleCI_finite_sample.m
CI estimation based on PLE - untested

@@ -1,0 +1,39 @@
+% get finite sample confidence intervals (CI) for PLE 
+% does not need hessians of the objective function at the optimal solution
+function [sigma,collect_eps] = pleCI_finite_sample(delta_alpha,chiPLE,thetai)
+
+% theta s.t. obj(theta)-obj(optimal_theta)<delta_alpha
+% get point of intersection of PL curve with threshold line (delta_alpha)
+eps = 1e-1;
+pl_diff = delta_alpha-chiPLE;
+
+while any(pl_diff<eps)
+    eps = eps/5;
+    if ~any(pl_diff<eps)
+        collect_eps = eps*5;
+        collect_pos_first = find(pl_diff<eps*5,1,'first');
+        collect_pos_last = find(pl_diff<eps*5,1,'last');
+    end
+end
+
+sigma = zeros(1,2);
+if collect_pos_first==collect_pos_last
+    % only one confidence interval is finite
+    sigma(1) = thetai(collect_pos_first);
+    sigma(2) = sigma(1); %
+else
+    % finite confidence intervals on both ends
+    % CI lb
+    sigma(1) = thetai(collect_pos_first);
+    sigma(2) = thetai(collect_pos_last);
+end
+        
+
+
+
+
+
+
+
+
+

### 2017-08-20 KineticModel/estimation/compare_vals.m
no major changes

@@ -48,7 +48,7 @@ xss_error = cell(data.nc,1);
 fss_plot = cell(nf,1);
 fss_error = cell(nf,1);
 
-barc = struct();
+% collect all data
 for j = 1:data.nc
     xss_plot{j} = [exp_xss(j,:)' est_xss(j,:)'];
     xss_error{j} = [zeros(1,length(exp_xss(j,:)))' est_xerr(j,:)'];
@@ -58,8 +58,11 @@ for k = 1:nf
     fss_plot{k} = [exp_fss(k,:)' est_fss(k,:)'];
     fss_error{k} = [zeros(1,length(exp_fss(k,:)))' est_ferr(k,:)'];
 end
+
+% plot concentrations
 hfc = figure;
 ahc = zeros(data.nc,1);
+barc = struct();
 for j = 1:data.nc
     ahc(j) = subplot(data.nc,1,j);
     set(ahc(j),'NextPlot','add');

### 2017-08-20 KineticModel/estimation/compare_vals_scatter.m
initial commit - not ready for use

@@ -1,0 +1,52 @@
+% scatter plot of experimental and estimated data
+function compare_vals_scatter(est_sol,exp_sol,data,opts,test_data_id)
+if isempty(est_sol)
+    fprintf('No optimal solution found\n');
+    allfh = [];
+    return
+end
+
+test_data_id = logical(test_data_id);
+all_labels = {'P1','P2','P3','WT'};
+
+% calculate new model fluxes
+% before perturbation - wt fluxes
+% assume wt (unperturbed) data is always included at the end
+wt_exp_xss = exp_sol(end).xss;
+wt_exp_fss = kotte_flux_noCAS(wt_exp_xss,data.odep);
+
+% after perturbation - perturb parameters to ascertain fluxes
+pt_datasize = length(find(test_data_id(1:end-1)));
+nf = size(wt_exp_fss,1);
+
+exp_xss = cat(2,exp_sol.xss);
+exp_fss = cat(2,exp_sol.fss);
+
+% collect all data to plot [p#1 p#2....  wt]  
+% estimated
+est_xss = est_sol.xss;
+est_xerr = est_sol.xerr;
+est_fss = est_sol.fss;
+est_ferr = est_sol.ferr;
+
+if ~test_data_id(end) % wt not included in estimation
+    rel_labels = all_labels(test_data_id);
+else
+    % experimental
+    exp_xss = exp_xss(:,test_data_id);
+    exp_fss = exp_fss(:,test_data_id);
+    % axis labels
+    rel_labels = [all_labels(test_data_id) all_labels(end)];
+end
+nplot = pt_datasize+1;
+for j = 1:data.nc
+    xss_plot{j} = [exp_xss(j,:)' est_xss(j,:)'];
+    xss_error{j} = [zeros(1,length(exp_xss(j,:)))' est_xerr(j,:)'];
+end
+
+for k = 1:nf
+    fss_plot{k} = [exp_fss(k,:)' est_fss(k,:)'];
+    fss_error{k} = [zeros(1,length(exp_fss(k,:)))' est_ferr(k,:)'];
+end
+
+hfc = figure

### 2017-08-20 KineticModel/estimation/scripts-other/est_flux1_nn.m
call to scatter plot

@@ -78,6 +78,8 @@ est_data = combine_results(optsol,opts,no_noise_sol,optimdata,pss,pss);
 % compare fluxes and concentrations
 hfcv = compare_vals(est_data,no_noise_sol,optimdata,opts,pss);
 
+compare_vals_scatter(est_data,no_noise_sol,optimdata,opts,pss);
+
 % compare parameters in parameter space
 hfp = compare_pars(est_data);
 

### 2017-08-20 KineticModel/estimation/docs/formulation/ident.tex
changes based on comments from prof. cluett

@@
-\documentclass[10pt]{report}
+\documentclass[10pt]{article}
 
@@
 
-	A profile likelihood-based approach can be used to determine both the structural and practical identifiability of parameters in nonlinear dynamical models \parencite{Raue2009a}.
+	\section{Methods:}
+	 We use a profile likelihood-based approach \parencite{Raue2009a} to establish structural and practical identifiability of parameters in nonlinear kinetic models of metabolism.
+Briefly, the approach seeks to establish the existence/non-existence of bounds in confidence intervals for the estimates of parameters in nonlinear models.
+The profile likelihood is calculated based on Equation (\ref{eq:pl}) for each parameter $\theta_i$ where $\chi^2(\theta_i)$ is given by Equation (\ref{eq:chi2}).
+	 \begin{align}\label{eq:pl}
+	 \chi_{PL}^2(\theta_i) = \underset{\theta_{j\ne i}}{\mathrm{min}} \left[\chi^2(\theta)\right]
+	 \end{align}
+	 \begin{align}\label{eq:chi2}
+	 \chi^2(\theta) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
+	 \end{align}
+	 In the minimization objective shown in Equation (\ref{eq:chi2}) for parameter estimation, $y_{kl}^*$ is the available experimental time course data for each observable state $k$ at each $l$ time point.
+The difference between the data and the model estimates at these time points, $y_{kl}$ is weighted by the variance in the experimental data $\sigma_{kl}^*$.
+An algorithm to calculate the profile likelihood, $\chi_{PL}^2(\theta_i)$, based on Equation \ref{eq:pl} is given below.
+
 	 The identifiability of parameters is established through the confidence intervals of their estimates, $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
-We provide the methods to estimate these confidence intervals below.
+The likelihood-based confidence interval for any parameter whose profile likelihood is estimated can be written on the basis of a threshold $\Delta_{\alpha}$ in the likelihood as in Equation (\ref{eq:pleci}).
+	 \begin{align}\label{eq:pleci}
+	 \{\theta|\chi^2(\theta)-\chi^2(\hat{\theta})<\Delta_{\alpha}\}
+	 \end{align}
+	 The threshold $\Delta_{\alpha}$ in the likelihood is the 1-$\alpha$ quantile of the $\chi^2$ distribution, represented as $\chi^2(\alpha,df)$.
+The confidence intervals obtained hold for $df$ degrees of freedom.
+For a choice of $df$=1 the confidence intervals will hold for each parameter individually, and confidence intervals that hold jointly for all parameters can be obtained by choosing the number of parameters as $df$.
+
 	The visualization of structurally and practically non-identifiable parameters using the profile likelihood approach is illustrated in Figure \ref{fig:identtypes}.
+The points of intersection between the profile likelihood curves (solid line) with the one parameter likelihood threshold ($\Delta_{\alpha} = \chi^2(\alpha,1)$, dashed line) provide the confidence intervals of the parameter $\theta_i$.
 The confidence intervals of a structurally non-identifiable parameter are unbounded, i.e., $\left[-\infty, +\infty\right]$ (Figure \ref{fig:identtypes}a), while the confidence intervals of a practically non-identifiable parameter are unbounded in at least one direction, i.e., $\left[\sigma_{i}^-,\sigma_{i}^+\right]$ where either $\sigma_{i}^- = -\infty$ or $\sigma_{i}^+ = +\infty$ (Figure \ref{fig:identtypes}b).
@@
 The horizontal axis represents the changes in the value of the parameter ($\theta_i$) whose identifiability is being determined and the profile likelihood ($\chi^2_{PL}(\theta_i)$) is shown in the vertical axis.
-The confidence interval thresholds ($\Delta_{\alpha}$) used to determine the identifiability of the parameter are denoted by the horizontal dotted line.
-The practically identifiability of a parameter is dependent on the quality and quantity of experimental data used to estimate the parameter.}\label{fig:ident-pl}
-	\end{figure}\label{fig:identtypes}
+The confidence interval threshold ($\Delta_{\alpha}$) used to determine the identifiability of the parameter is denoted by the horizontal dotted line.
+Identifiable parameters are distinguished from non-identifiable parameters by the presence of both upper and lower bounds on their confidence interval estimates $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
+}\label{fig:identtypes}
+	\end{figure}
+
+	\subsection{Algorithm to calculate profile likelihood:}		
+	\begin{enumerate}
+		\item Start by solving the numerical optimization problem in Equation \ref{eq:pl} for a fixed value of $\theta_i$ to determine the initial point on the profile likelihood curve.
+		\item Take an increasing/decreasing step $\theta_{step}$ in the direction of $\theta_i$.\\
+		The increasing/decreasing steps in $\theta_i$ can be adapted to the shape of the profile likelihood curve that is encountered while running the algorithm.
+Accordingly, the step $\theta_{step}$ should satisfy Equation (\ref{eq:adaptivestep}) below, where $\theta^{k-1}$ refers to parameter estimates obtained from iteration $k-1$ of the above algorithm and $q\in \left[0,1\right]$.
+		\begin{align}\label{eq:adaptivestep}
+		\chi^2\left(\theta^{k-1}+\theta_{step}\right) - \chi^2(\theta^{k-1}) \approx q.\Delta_{\alpha}
+		\end{align}
+		\item Re-optimize $\theta_{j\ne i}$ using Equation \ref{eq:pl}.
+		\item Repeat the last two steps until a desired threshold $\Delta_{\alpha}$ is exceeded or a maximal amount of steps is reached.
+	\end{enumerate}
 
 	Due to the dependence of practical parameter identifiability on the experimental data, the profile likelihood approach can be used to design experiments in such a way that the observables that are derived from these experiments can improve the practical identifiability of the parameters.
-We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.
+We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.	Assuming a parameter $\theta_i$ is practically non-identifiable (Figure \ref{fig:edwithpl}a), performing a profile-likelihood based identifiability analysis using simulated data can help determine the nature of experiments needed to make the parameter identifiable (Figure \ref{fig:edwithpl}b).
+In contrast, performing non-informative experiments without prior knowledge on their ability to change the identifiability of the parameter may provide data that cannot be used to estimate parameter $\theta_i$ (Figure \ref{fig:edwithpl}c).
 
@@
 The changes in the profile likelihood of the parameter when estimated with b) good experimental data and c) bad experimental.
-The identifiability of the parameter does not change due to the poor quality/quantity of experiments.}\label{fig:edwithpl}
+The practical identifiability of a parameter is dependent on the quality and quantity of experimental data used to estimate the parameter.
+The identifiability of the parameter, as determined using its profile likelihood, does not change due to the poor quality/quantity of experiments.}\label{fig:edwithpl}
 	\end{figure}
 
-	\section{Methods:}
-	We use a profile likelihood-based approach to establish structural and practical identifiability of parameters in nonlinear kinetic models of metabolism.
-Briefly, the approach seeks to establish the existence/non-existence of bounds in confidence intervals for the estimates of parameters in nonlinear models.
-The profile likelihood is calculated based on Equation \ref{eq:pl} for each parameter $\theta_i$ where $\chi^2(\theta_i)$ is given by Equation \ref{eq:chi2}.
-	\begin{align}\label{eq:pl}
-	\chi_{PL}^2(\theta_i) = \underset{\theta_{j\ne i}}{\mathrm{min}} \left[\chi^2(\theta)\right]
-	\end{align}
-	\begin{align}\label{eq:chi2}
-	\chi^2(\theta) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
-	\end{align}
-	Equation \ref{eq:chi2} is typically used as the objective to be minimized in a typical parameter estimation problem where $y_{kl}^*$ is the available experimental time course data for each observable state $k$ at each $l$ time point.
-The difference between this data and the model estimates at these time points, $y_{kl}$ is weighted by the experimental error $\sigma_{kl}^*$.
-An algorithm to calculate $\chi_{PL}^2(\theta_i)$ based on Equation \ref{eq:pl} is given below.
-
-	\subsubsection{Algorithm:}	
-	\begin{enumerate}
-		\item Start by solving the numerical optimization problem in Equation \ref{eq:pl} for a fixed value of $\theta_i$ to determine the initial point on the profile likelihood curve.
-		\item Take an increasing/decreasing step $\theta_{step}$ in the direction of $\theta_i$.
-		\item Re-optimize $\theta_{j\ne i}$ using Equation \ref{eq:pl}.
-		\item Repeat the last two steps until a desired threshold $\Delta_{\alpha}$ is exceeded or a maximal amount of steps is reached.
-	\end{enumerate}
-
-	\subsubsection{Adaptive step change:}
-	The increasing/decreasing steps in $\theta_i$ can be adapted to the shape of the profile likelihood curve that is encountered while running the above algorithm.
-Accordingly, the step $\theta_{step}$ should satisfy Equation \ref{eq:adaptivestep} below, where $\theta^{k-1}$ refers to parameter estimates obtained from iteration $k-1$ of the above algorithm, and $q\in \left[0,1\right]$.
-	\begin{align}\label{eq:adaptivestep}
-	\chi^2\left(\theta^{k-1}+\theta_{step}\right) - \chi^2(\theta^{k-1}) \approx q.\Delta_{\alpha}
-	\end{align}
-
-	\subsubsection{Identifiability analysis of parameters in a kinetic model of gluconeogenesis:}
+	\subsection{Identifiability analysis of parameters in a kinetic model of gluconeogenesis:}
 	The proposed model for acetate consumption through gluconeogenesis and its corresponding kinetic model is used as a case study to illustrate the utility of identifiability analysis for the design of experiments for estimating parameters in kinetic models of metabolism.
@@
 	\end{equation}
-	Simulated data
 
@@
 
-	\subsubsection{Formulation a):}		
-	\begin{subequations}\label{eq:l2norm-flux}
-		\begin{align}
-		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert\\
-		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-		\end{align}
-	\end{subequations}
-
-	\subsubsection{Formulation b):}	
-	In this formulation, in Equation (\ref{eq:l2norm-conc}), the flux and concentrations are constrained as earlier with a noise term.	
-	\begin{subequations}\label{eq:l2norm-conc}
-		\begin{align}
-		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
-		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-		\end{align}
-	\end{subequations}
-
-	\subsubsection{Formulation c):}
-	A third formulation where the noise terms ($\epsilon$) used as bounds for the flux and the concentrations are included in the optimization objective as a variable to be penalized is given in Equation (\ref{eq:noise-var}) below.
-Notice that the noise term has been separated into two components, $\epsilon_x$ and $\epsilon_v$, corresponding to noise in the concentrations and the fluxes, respectively.
-$w_3$ and $w_4$ are the weights associated with the noise terms in the objective function.
+	\paragraph{Formulation c):}
+	In order to estimate kinetic parameters for fluxes in models of metabolism, we propose a formulation accounting for the uncertainty in the concentration and flux estimates.
+The uncertainty terms ($\epsilon$) are used as bounds for the flux and the concentrations and are included in the optimization objective as a variable to be penalized is given in Equation (\ref{eq:noise-var}) below.
+Note that the concentration and the fluxes have different uncertainties, $\epsilon_x$ and $\epsilon_v$ respectively, associated with them.
+$w_1$, $w_2$, $w_3$ and $w_4$ are the weights associated with the four different terms in the objective.
 	\begin{center}
@@
 	\end{center}
-
-	\subsubsection{Formulation d):}
-	Due to the inability of Equation (\ref{eq:l2norm-flux}) to get feasible results, we modified the problem to include the L2-norm of the concentrations as well in the objective function in Equation (\ref{eq:free-var}a).
-In Equation (\ref{eq:free-var}) the flux and the concentrations are free variables, i.e.,  $\mathbf{x}\in\left[\mathbf{0},+\mathbf{\infty}\right]$.
-	\begin{subequations}\label{eq:free-var}
-		\begin{align}
-		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
-		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-		& \text{      }0\le x_i \le +\infty & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-		& \text{      }0\le v_j \le +\infty & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-		\end{align}
-	\end{subequations}
-
-After testing the problem in Equation (\ref{eq:free-var}) above for the acetate uptake flux $v_1$, we found that all solutions using IPOPT exceeded the maximum number of iterations (set at 5000) and never produced an optimal solution.
-Hence, we use a modified form of the bounds in Equation (\ref{eq:free-var}) above in Equation (\ref{eq:l2norm-conc}).
 
@@
 
+	\section{Appendix}
+	\subsection{Formulation a):}		
+	\begin{subequations}\label{eq:l2norm-flux}
+		\begin{align}
+		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert\\
+		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+		\end{align}
+	\end{subequations}
+
+	\subsection{Formulation b):}	
+	In this formulation, in Equation (\ref{eq:l2norm-conc}), the flux and concentrations are constrained as earlier with an uncertainty term.	
+	\begin{subequations}\label{eq:l2norm-conc}
+		\begin{align}
+		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
+		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+		\end{align}
+	\end{subequations}
+
+	\subsection{Formulation d):}
+	Due to the inability of Equation (\ref{eq:l2norm-flux}) to get feasible results, we modified the problem to include the L2-norm of the concentrations as well in the objective function in Equation (\ref{eq:free-var}a).
+In Equation (\ref{eq:free-var}) the flux and the concentrations are free variables, i.e.,  $\mathbf{x}\in\left[\mathbf{0},+\mathbf{\infty}\right]$.
+	\begin{subequations}\label{eq:free-var}
+		\begin{align}
+		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
+		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
+		& \text{      }0\le x_i \le +\infty & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
+		& \text{      }0\le v_j \le +\infty & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
+		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
+		\end{align}
+	\end{subequations}
+
+	After testing the problem in Equation (\ref{eq:free-var}) above for the acetate uptake flux $v_1$, we found that all solutions using IPOPT exceeded the maximum number of iterations (set at 5000) and never produced an optimal solution.
+Hence, we use a modified form of the bounds in Equation (\ref{eq:free-var}) above in Equation (\ref{eq:l2norm-conc}).
+
 	\printbibliography

### 2017-08-20 KineticModel/estimation/ident/syside_test2.m
calc pos change in theta K1ac

@@ -13,7 +13,7 @@ pt_sol_id = [1 2 3];
 close all
 
 optim_opts = struct('pname','K1ac','nc',3,'nf',6,...
-                    'plim',[0.001 5],...
+                    'plim',[0.001 6],...
                     'casmodelfun',@kotteCASident,...
                     'integratorfun','RK4integrator_cas',...
                     'odep',odep_bkp,...
@@ -39,8 +39,8 @@ scale(2) = 1e6;
 p0 = opts.odep(2:13)'./scale;
 
 % call PLE evaluation function
-[PLEvals_neg] =...
-getPLE(thetai_fixed_value,theta_step,p0,opts.odep,delta_alpha,optim_opts,maxiter,3);
+[PLEvals_pos] =...
+getPLE(thetai_fixed_value,theta_step,p0,opts.odep,delta_alpha,optim_opts,maxiter,1);
 
 % figure
 % hold on

### 2017-11-04 ident/mathematica/docs/ident_method_sym.tex
moved relevant mathematica files to ident folder

--- 2017-08-07 KineticModel/estimation/docs/formulation/ident.tex
@@
 
+%\usepackage[font=small,labelfont=bf,labelsep=period]{caption}
+\usepackage[style=authoryear,sorting=nyt,url=false,isbn=false,doi=false,firstinits=true,backend=biber]{biblatex}
+
+\renewcommand{\baselinestretch}{2.0}
+\renewcommand*\contentsname{Table of Contents}
+%\captionsetup{font={stretch=2.0}}
+
+\DeclareNameAlias{default}{last-first}
+
+\DefineBibliographyStrings{english}{%
+	andothers = {\addcomma\addspace\textsc{et\addabbrvspace al}\adddot},
+	and = {\textsc{and}}
+}
+\renewcommand*{\labelnamepunct}{\space\space}
+
+\renewbibmacro{in:}
+{%
+	\ifentrytype{article}{%
+	}{%
+		\printtext{\bibstring{in}\intitlepunct}%
+	}%
+}
+\renewbibmacro*{volume+number}{%
+	\printfield{volume}%
+	\setunit*{\addcomma\space}%
+	\printfield{number}%
+	\setunit{\addcomma\space}}
+
+\DeclareFieldFormat{pages}{#1}
+
+\renewbibmacro*{publisher+location+date}{%
+	\printlist{publisher}%
+	\setunit*{\addcomma\space}%
+	\printlist{location}%
+	\setunit*{\addcomma\space}%
+	\usebibmacro{date}%
+	\newunit}
+
+\renewcommand{\newunitpunct}{\addcomma\space}
+\DeclareFieldFormat[article,inbook,incollection,inproceedings,patent,thesis,unpublished]{title}{#1} 
+\DeclareFieldFormat{year}{#1}
+
+\addbibresource{refs/ident_refs.bib}
+
 \begin{document}
 	\section{Introduction:}
-	\subsection{Identifiability analysis: Definitions and Formulations}
-	The identifiability of parameters in nonlinear models of physical processes can be classified into two categories: structural and practical identifiability.
-The effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
-The effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
-Practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
-The aforementioned distinction between practical and structural identifiability of parameters is illustrated in Figure.
-Thus, the identifiability of parameters in nonlinear models is dependent on the model structure, parameterization, and the quality and quantity of experimental data that is available for the purpose of estimation.
+	\subsection{A method to establish posterior identifiability of metabolic network models:}
+	This document details a method to establish the practical (posterior) identifiability of metabolic network models using the algebraic relationship between fluxes.
+	Every flux, $v$, in a kinetic model of a metabolic network can be expressed as a nonlinear algebraic equation (Equation \ref{eq:flux_gen}).
+The fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ in Equation (\ref{eq:flux_gen}).
+	\begin{align}\label{eq:flux_gen}
+	v = f(\mathbf{x},\mathbf{\theta})
+	\end{align}
+	Given the nonlinear nature of this model, the function $f$ in Equation (\ref{eq:flux_gen}) can expressed, without loss of generality as,
+	\begin{align}\label{eq:nr_dr}
+	v = \frac{N(\mathbf{x},\theta)}{D(\mathbf{x},\theta)}
+	\end{align}
+	where $N(\mathbf{x},\theta)$ is the numerator of $f$, and $D(\mathbf{x},\theta)$ is the denominator of $f$.
 
-	On the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated.
-On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
+	If $\theta \in \mathbb{R}^p$, given a set of experimental measurements for the metabolite concentrations $\mathbf{x}$ and the reaction fluxes $\mathbf{v}$, theoretically, it is possible to choose $p$ sets of data from these measurements to solve for the $p$ parameters in $\theta$.
+However, if any these datasets do not satisfy the condition that $D(\mathbf{x},\theta) \neq 0$, then the number of experiments required to estimate the $p$ parameters in $\theta$ can be established to be greater than $p$.
+An example is shown below.
 
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.8\textwidth,height=.8\textheight,keepaspectratio]{figures/identifiability}}
-		\caption{The profile likelihood estimates of a) a structurally non-identifiable, b) a practically non-identifiable and c) an identifiable parameter.
-The practically identifiability of a parameter is dependent on the quality and quantity of experimental data used to estimate the parameter.}\label{fig:ident-pl}
-	\end{figure}
-
-	We can use a profile likelihood-based approach to determine the structural and practical identifiability of parameters in any dynamic model.
-The identifiability of parameters is established through the confidence intervals of their estimates, $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
-The confidence intervals of a structurally non-identifiable parameter are unbounded, i.e., $\left[-\infty, +\infty\right]$, while the confidence intervals of a practically non-identifiable parameter are unbounded in at least one direction, i.e., $\left[\sigma_{i}^-,\sigma_{i}^+\right]$ where either $\sigma_{i}^- = -\infty$ or $\sigma_{i}^+ = +\infty$.
-If a parameter's estimates have a finite confidence interval then the parameter is said to be identifiable.
-The visualization of structurally and practically non-identifiable parameters using the profile likelihood approach is illustrated in Figure \ref{fig:ident-pl}.
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/ed_with_pl}}
-		\caption{Cartoon illustrating the utility of identifiability analysis for experimental design.
-a) The profile likelihood of a practically non-identifiable parameter that needs to be estimated based on both good and bad experimental data.
-The changes in the profile likelihood of the parameter when estimated with b) good experimental data and c) bad experimental.
-The identifiability of the parameter does not change due to the poor quality/quantity of experiments.}\label{fig:edwithpl}
-	\end{figure}
-
-	Due to the dependence of practical parameter identifiability on the experimental data, the profile likelihood approach can be used to design experiments in such a way that the observables that are derived from these experiments can improve the practical identifiability of the parameters.
-We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.
-
-	\section{Methods:}
-	We use a profile likelihood-based approach to establish structural and practical identifiability of parameters in nonlinear kinetic models of metabolism.
-Briefly, the approach seeks to establish the existence/non-existence of bounds in confidence intervals for the estimates of parameters in nonlinear models.
-The profile likelihood is calculated based on Equation \ref{eq:pl} for each parameter $\theta_i$ where $\chi^2(\theta_i)$ is given by Equation \ref{eq:chi2}.
-	\begin{align}\label{eq:pl}
-	\chi_{PL}^2(\theta_i) = \underset{\theta_{j\ne i}}{\mathrm{min}} \left[\chi^2(\theta)\right]
-	\end{align}
-	\begin{align}\label{eq:chi2}
-	\chi^2(\theta) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
-	\end{align}
-	Equation \ref{eq:chi2} is typically used as the objective to be minimized in a typical parameter estimation problem where $y_{kl}^*$ is the available experimental time course data for each observable state $k$ at each $l$ time point.
-The difference between this data and the model estimates at these time points, $y_{kl}$ is weighted by the experimental error $\sigma_{kl}^*$.
-An algorithm to calculate $\chi_{PL}^2(\theta_i)$ based on Equation \ref{eq:pl} is given below.
-
-	\subsubsection{Algorithm:}	
-	\begin{enumerate}
-		\item Start by solving the numerical optimization problem in Equation \ref{eq:pl} for a fixed value of $\theta_i$ to determine the initial point on the profile likelihood curve.
-		\item Take an increasing/decreasing step $\theta_{step}$ in the direction of $\theta_i$.
-		\item Re-optimize $\theta_{j\ne i}$ using Equation \ref{eq:pl}.
-		\item Repeat the last two steps until a desired threshold $\Delta_{\alpha}$ is exceeded or a maximal amount of steps is reached.
-	\end{enumerate}
-
-	\subsubsection{Adaptive step change:}
-	The increasing/decreasing steps in $\theta_i$ can be adapted to the shape of the profile likelihood curve that is encountered while running the above algorithm.
-Accordingly, the step $\theta_{step}$ should satisfy Equation \ref{eq:adaptivestep} below, where $\theta^{k-1}$ refers to parameter estimates obtained from iteration $k-1$ of the above algorithm, and $q\in \left[0,1\right]$.
-	\begin{align}\label{eq:adaptivestep}
-	\chi^2\left(\theta^{k-1}+\theta_{step}\right) - \chi^2(\theta^{k-1}) \approx q.\Delta_{\alpha}
-	\end{align}
-
-	\subsubsection{Identifiability analysis of parameters in a kinetic model of gluconeogenesis:}
-	The proposed model for acetate consumption through gluconeogenesis and its corresponding kinetic model is used as a case study to illustrate the utility of identifiability analysis for the design of experiments for estimating parameters in kinetic models of metabolism.
-The kinetic model is described below.
-	\begin{equation}\label{eq:ode1}
-	\frac{d}{dt}pep=v_1-v_2-v_4
-	\end{equation}
-	\begin{equation}\label{eq:ode2}
-	\frac{d}{dt}fdp=v_2-v_3
-	\end{equation}
-	\begin{equation}\label{eq:ode3}
-	\frac{d}{dt}E=v_{e,max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right) - d E
-	\end{equation}
-	The kinetic expressions for fluxes $v_1$ through $v_4$ are given below.
-The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics.
-The acetate flux through $v_1$ is also governed by the quantity of available enzyme E.
-	\begin{equation}\label{eq:flux1}
-	v_1 = k_{1}^{cat}E\frac{acetate}{acetate+K_{1}^{acetate}}
-	\end{equation}	
-	\begin{equation}\label{eq:flux2}
-	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
-	\end{equation}
-	\begin{equation}\label{eq:flux3}
-	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
-	\end{equation}
-	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
-The added flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
-	\begin{equation}\label{eq:flux4}
-	v_4 = k_{4}^{cat}.pep
-	\end{equation}
-	Simulated data
-
-	\subsection{Parameter estimation in kinetic models of metabolism:}
-	The formulation for parameter estimation assumes that the concentrations and the fluxes corresponding to each perturbation are variables in the optimization problem formulated to estimate kinetic parameters.
-
-	Let $\mathbf{x}\in\mathbb{R}^m$, $\mathbf{p}\in\mathbb{R}^l$ and $\mathbf{v}\in\mathbb{R}$ be the vector of concentrations, parameters and fluxes respectively.
-If the number of perturbations used for parameter estimation is in $\mathbb{R}^p$, then the vector of concentrations, fluxes and parameters that need to be estimated in the optimization problem changes to $\mathbf{x}\in\mathbb{R}^{mp}, \mathbf{v}\in\mathbb{R}^{p}$ and $\mathbf{p}\in\mathbb{R}^{l}$ respectively.
-
-	Let $i\in\mathbb{R}^{mp}$, $j\in\mathbb{R}^{p}$ and $k\in\mathbb{R}^{l}$ represent the indices for concentrations, fluxes and parameters respectively in Equations (\ref{eq:l2norm-flux})-(\ref{eq:noise-var}).
-
-	\subsubsection{Formulation a):}		
-	\begin{subequations}\label{eq:l2norm-flux}
-		\begin{align}
-		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert\\
-		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-		\end{align}
-	\end{subequations}
-
-	\subsubsection{Formulation b):}	
-	In this formulation, in Equation (\ref{eq:l2norm-conc}), the flux and concentrations are constrained as earlier with a noise term.	
-	\begin{subequations}\label{eq:l2norm-conc}
-		\begin{align}
-		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
-		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-		\end{align}
-	\end{subequations}
-
-	\subsubsection{Formulation c):}
-	A third formulation where the noise terms ($\epsilon$) used as bounds for the flux and the concentrations are included in the optimization objective as a variable to be penalized is given in Equation (\ref{eq:noise-var}) below.
-Notice that the noise term has been separated into two components, $\epsilon_x$ and $\epsilon_v$, corresponding to noise in the concentrations and the fluxes, respectively.
-$w_3$ and $w_4$ are the weights associated with the noise terms in the objective function.
-	\begin{center}
-		\begin{subequations}\label{eq:noise-var}
-			\begin{align}
-			\underset{\mathbf{x},\mathbf{p},\mathbf{v},\epsilon_x,\epsilon_v}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert + w_3\epsilon_x + w_4\epsilon_v\\
-			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-			& \text{      }x^*_i(1-\epsilon_x)\le x_i \le x^*_i(1+\epsilon_x) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-			& \text{      }v^*_j(1-\epsilon_v)\le v_j \le v^*_j(1+\epsilon_v) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-			\end{align}
-		\end{subequations}
-	\end{center}
-
-	\subsubsection{Formulation d):}
-	Due to the inability of Equation (\ref{eq:l2norm-flux}) to get feasible results, we modified the problem to include the L2-norm of the concentrations as well in the objective function in Equation (\ref{eq:free-var}a).
-In Equation (\ref{eq:free-var}) the flux and the concentrations are free variables, i.e.,  $\mathbf{x}\in\left[\mathbf{0},+\mathbf{\infty}\right]$.
-	\begin{subequations}\label{eq:free-var}
-		\begin{align}
-		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
-		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-		& \text{      }0\le x_i \le +\infty & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-		& \text{      }0\le v_j \le +\infty & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-		\end{align}
-	\end{subequations}
-
-After testing the problem in Equation (\ref{eq:free-var}) above for the acetate uptake flux $v_1$, we found that all solutions using IPOPT exceeded the maximum number of iterations (set at 5000) and never produced an optimal solution.
-Hence, we use a modified form of the bounds in Equation (\ref{eq:free-var}) above in Equation (\ref{eq:l2norm-conc}).
-
-	\begin{table}[!tbhp]
-		\caption{Table showing the perturbed values of all fluxes used for parameter estimation.}
-		\begin{center}				
-			\begin{tabular}{ccc}
-				Designation & Perturbed Fluxes & Perturbed Values\\
-				\hline
-				P1 & $v_1$ & 2\\
-				P2 & $v_2$ & 0.2\\
-				P3 & $v_3$ & 0.5
-			\end{tabular}
-		\end{center}	
-		\label{tab:pval}
-	\end{table}
-
-	\section{Results:}
-
-	\begin{table}[!tbhp]
-		\caption{Table showing the ability of different formulations mentioned above to estimate enzyme kinetic parameters for different fluxes based on the perturbations given in the above table.}
-		\begin{center}				
-			\begin{tabular}{cccccc}
-				\hline
-				Flux & Data Type & \multicolumn{2}{c}{Formulation (b)} & \multicolumn{2}{c}{Formulation (c)}\\
-				\hline
-				&&$\epsilon_x$(\%)&$\epsilon_v$(\%)&$\epsilon_x$(\%)&$\epsilon_v$(\%)\\
-				\hline
-				\multirow{2}{*}{flux 1} & no noise & 30 & 5 & 33.33 & 0\\
-												  &	noisy & x & x & 0 & 33.33\\
-			  	\hline
-				 \multirow{2}{*}{flux 2} & no noise & 90 & 30 & 0 & 31\\
-				 									& noisy & 85 & 40 & 0 & 30\\
-				\hline				 									
-				\multirow{2}{*}{flux 3} & no noise & 15 & 10 & 17 & 0\\
-													& noisy & x & x & 17 & 0\\
-				\hline			
-			\end{tabular}
-		\end{center}	
-		\label{tab:res1}
-	\end{table}
-	\subsection{Estimation of flux 1:}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_conc_aug6}}
-		\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_conc_aug6}}
-		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_flux_aug6}}
-		\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_flux_aug6}}
-		\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_par_aug6}}
-		\caption{Plot comparing noiseless model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_par_aug6}}
-		\caption{Plot comparing noisy model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\clearpage
-
-	\subsection{Estimation of flux 2:}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_conc_aug6}}
-		\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_conc_aug6}}
-		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_flux_aug6}}
-		\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_flux_aug6}}
-		\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_par_aug6}}
-		\caption{Plot comparing noiseless model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_par_aug6}}
-		\caption{Plot comparing noisy model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\clearpage
-
-	\subsection{Estimation of flux 3:}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_conc_aug6}}
-		\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_conc_aug6}}
-		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_flux_aug6}}
-		\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_flux_aug6}}
-		\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_par_aug6}}
-		\caption{Plot comparing noiseless model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_par_aug6}}
-		\caption{Plot comparing noisy model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\clearpage
-
+	This analysis can be performed for each flux in a metabolic network independent of all the other fluxes. This enables this method to be scalable to even genome-scale models.
 \end{document}

### 2017-11-04 KineticModel/estimation/docs/formulation/ident.tex

@@
 
+	\subsection{A method to establish posterior identifiability of metabolic network models:}
+	This section details a method to establish the practical (posterior) identifiability of metabolic network models using the algebraic relationship between fluxes.
+	Every flux, $v$, in a kinetic model of a metabolic network can be expressed as a nonlinear algebraic equation (Equation \ref{eq:flux_gen}).
+The fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ in Equation (\ref{eq:flux_gen}).
+	\begin{align}\label{eq:flux_gen}
+		v = f(\mathbf{x},\mathbf{\theta})
+	\end{align}
+	Given the nonlinear nature of this model, the function $f$ in Equation (\ref{eq:flux_gen}) can expressed, without loss of generality as,
+	\begin{align}\label{eq:nr_dr}
+		v = \frac{N(\mathbf{x},\theta)}{D(\mathbf{x},\theta)}
+	\end{align}
+	where $N(\mathbf{x},\theta)$ is the numerator of $f$, and $D(\mathbf{x},\theta)$ is the denominator of $f$.
+
+	If $\theta \in \mathbb{R}^p$, given a set of experimental measurements for the metabolite concentrations $\mathbf{x}$ and the reaction fluxes $\mathbf{v}$, theoretically, it is possible to choose $p$ sets of data from these measurements to solve for the $p$ parameters in $\theta$.
+However, if any these datasets do not satisfy the condition that $D(\mathbf{x},\theta) \neq 0$, then the number of experiments required to estimate the $p$ parameters in $\theta$ can be established to be greater than $p$.
+An example is shown below.
+
+	This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
+This enables this method to be scalable to even genome-scale models.
+The following section demonstrates this methodology for one of the fluxes in the gluconeogenic model of Kotte et al., \parencite{Kotte2014}.
+
 	\subsection{Identifiability analysis of parameters in a kinetic model of gluconeogenesis:}
@@
 	\end{equation}
+
+	We use flux $v_2$ to demonstrate the identifiability analysis method described in the previous section.
+Flux $v_2$ has two parameters, $V_2^{max}$ and $K_2^{pep}$ that need to be estimated from experimental data.
+Here, we assume that at least two different sets of experimental data for the concentrations and fluxes are available.
+Accordingly, we label these dataset as ${pep}^1$, ${v_2}^1$ and ${pep}^2$, ${v_2}^2$ respectively.
+Subsequently, these experimental datasets can be included in the model to form two simultaneous nonlinear algebraic equations in the parameters $V_2^{max}$ and $K_2^{pep}$ (Equation \ref{eq:v2_par}).
+
+	\begin{subequations}\label{eq:v2_par}
+		\begin{align}		
+		V_2^{max} = \frac{v_2^1v_2^2(pep^1-pep^2)}{v_2^2pep^1-v_2^1pep^2}\\
+		K_2^{pep} = \frac{pep^1(v_2^1pep^2-v_2^2pep^2)}{v_2^2pep^1-v_2^1pep^2}
+		\end{align}
+	\end{subequations}
 
@@
 	\subsection{Estimation of flux 1:}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_conc_aug6}}
-		\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_conc_aug6}}
-		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_flux_aug6}}
-		\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_flux_aug6}}
-		\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_par_aug6}}
-		\caption{Plot comparing noiseless model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_par_aug6}}
-		\caption{Plot comparing noisy model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_conc_aug6}}
+	%	\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_conc_aug6}}
+	%	\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_flux_aug6}}
+	%	\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_flux_aug6}}
+	%	\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_par_aug6}}
+	%	\caption{Plot comparing noiseless model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_par_aug6}}
+	%	\caption{Plot comparing noisy model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
+	%\end{figure}
 	\clearpage
@@
 	\subsection{Estimation of flux 2:}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_conc_aug6}}
-		\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_conc_aug6}}
-		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_flux_aug6}}
-		\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_flux_aug6}}
-		\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_par_aug6}}
-		\caption{Plot comparing noiseless model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_par_aug6}}
-		\caption{Plot comparing noisy model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_conc_aug6}}
+	%	\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_conc_aug6}}
+	%	\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_flux_aug6}}
+	%	\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_flux_aug6}}
+	%	\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_par_aug6}}
+	%	\caption{Plot comparing noiseless model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_par_aug6}}
+	%	\caption{Plot comparing noisy model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
+	%\end{figure}
 	\clearpage
@@
 	\subsection{Estimation of flux 3:}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_conc_aug6}}
-		\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_conc_aug6}}
-		\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_flux_aug6}}
-		\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_flux_aug6}}
-		\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_par_aug6}}
-		\caption{Plot comparing noiseless model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_par_aug6}}
-		\caption{Plot comparing noisy model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
-	\end{figure}
-	\clearpage
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_conc_aug6}}
+	%	\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_conc_aug6}}
+	%	\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_flux_aug6}}
+	%	\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_flux_aug6}}
+	%	\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_par_aug6}}
+	%	\caption{Plot comparing noiseless model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
+	%\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_par_aug6}}
+	%	\caption{Plot comparing noisy model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
+	%\end{figure}
+	%\clearpage
 

### 2017-11-09 ident/python2/ss-ident/kotte_model.py
get parameter and perturbation ids for each positively identifiable parameter

@@ -825,15 +825,19 @@ def establish_kotte_flux_identifiability(experimental_data_list):
             parameter_id = 1
     # create signed boolean array for identifiability
     signed_ident_values = np.sign(ident_values)
+    ident_fun_val = []
+    for id in range(0, number_data):
+        ident_fun_val.append(signed_ident_values[id*12:(id+1)*12, -1])
+    p_list = [[p_id for p_id, val in enumerate(data_set) if val > 0] for data_set in ident_fun_val]
+
+    parameter_perturbation_list = []
+    for all_parameter_iter in range(0, 12):
+        parameter_perturbation_list.append([id for id, parameter_ids in enumerate(p_list)
+                                            if all_parameter_iter in parameter_ids])
 
     # identify perturbations that result in positive value for identifiability
     p = [0] + parameters_per_flux
     p = np.cumsum(p).tolist()
-    #for iset in range(0, number_data):
-    #    rel_data = ident_values[0*(iset+1):p[-1]*(iset+1), :]
-    #    [id for id in range(0, len(rel_data)) if rel_data[:, -1]>0]
-
-
 
     fp_list = ['flux{}p{}'.format(f_index + 1, p_index + 1)
                for f_index, p_limit in enumerate(parameters_per_flux)
@@ -847,9 +851,11 @@ def establish_kotte_flux_identifiability(experimental_data_list):
     flux_based_array = np.zeros((number_data*p[-1], 3))
     flux_based_signed_array = np.zeros((number_data*p[-1], 3))
     pos = 0
+    #set = 0
     for fluxid in range(0, number_fluxes):
         all_id_list = []
         all_id_signed_list = []
+        #p_list = ['p{}'.format(p_index+1) for p_index in range(0, parameters_per_flux[fluxid])]
         for id1, id2 in zip(starting_indices[fluxid], ending_indices[fluxid]):
             nrows = len(range(id1, id2))
             flux_based_array[pos:pos+nrows, :] = ident_values[id1:id2, :]

### 2017-11-09 ident/python2/ss-ident/kotte_model.py
remove data rearrangement ordered based on fluxes

@@ -842,40 +842,6 @@ def establish_kotte_flux_identifiability(experimental_data_list):
     fp_list = ['flux{}p{}'.format(f_index + 1, p_index + 1)
                for f_index, p_limit in enumerate(parameters_per_flux)
                for p_index in range(0, p_limit)]
-    p_id_list = [[(p[i - 1] + 12 * id, p[i] + 12 * id)
-                       for id in range(0, number_data)]
-                      for i in range(1, len(p))]
-    starting_indices = [[x[0] for x in p_id] for p_id in p_id_list]
-    ending_indices = [[x[1] for x in p_id] for p_id in p_id_list]
-    flux_based_list = []
-    flux_based_array = np.zeros((number_data*p[-1], 3))
-    flux_based_signed_array = np.zeros((number_data*p[-1], 3))
-    pos = 0
-    #set = 0
-    for fluxid in range(0, number_fluxes):
-        all_id_list = []
-        all_id_signed_list = []
-        #p_list = ['p{}'.format(p_index+1) for p_index in range(0, parameters_per_flux[fluxid])]
-        for id1, id2 in zip(starting_indices[fluxid], ending_indices[fluxid]):
-            nrows = len(range(id1, id2))
-            flux_based_array[pos:pos+nrows, :] = ident_values[id1:id2, :]
-            flux_based_signed_array[pos:pos + nrows, :] = signed_ident_values[id1:id2, :]
-            all_id_list.append(ident_values[id1:id2, :])
-            all_id_signed_list.append(signed_ident_values[id1:id2, :])
-            pos += nrows
-        flux_based_list.append(all_id_list)
-
-    #pos = 0
-    #for id1 in parameters_per_flux:
-    #    rel_data = flux_based_signed_array[pos:pos+id1*number_data, :]
-    #    [id for id in range(0, len(rel_data)) if rel_data[:, -1]>0]
-
-    #perturbation_list = {}
-    #for exp_id in range(0, number_data):
-    #    [exp_id if signed_ident_values[i*12:(i+1)*12]]
-    #for flux_id in range(0, number_fluxes):
-    #    [id for id in range(0, number_data) if flux_based_signed_array[id, -1]>0]
-
 
     perturbation_list_flux_1 = []
     for parameter_id in range(0, 4):

### 2017-11-09 ident/python2/ss-ident/kotte_model.py
create dictionary of identifiable perturbations for each parameter

@@ -832,16 +832,19 @@ def establish_kotte_flux_identifiability(experimental_data_list):
 
     parameter_perturbation_list = []
     for all_parameter_iter in range(0, 12):
-        parameter_perturbation_list.append([id for id, parameter_ids in enumerate(p_list)
-                                            if all_parameter_iter in parameter_ids])
+        parameter_perturbation_list.append(tuple([id for id, parameter_ids in enumerate(p_list)
+                                            if all_parameter_iter in parameter_ids]))
+    fp_list = ['flux{}p{}'.format(f_index + 1, p_index + 1)
+               for f_index, p_limit in enumerate(parameters_per_flux)
+               for p_index in range(0, p_limit)]
+    # build dictionary of results
+    fp_dict = dict(zip(fp_list, parameter_perturbation_list))
 
     # identify perturbations that result in positive value for identifiability
     p = [0] + parameters_per_flux
     p = np.cumsum(p).tolist()
 
-    fp_list = ['flux{}p{}'.format(f_index + 1, p_index + 1)
-               for f_index, p_limit in enumerate(parameters_per_flux)
-               for p_index in range(0, p_limit)]
+
 
     perturbation_list_flux_1 = []
     for parameter_id in range(0, 4):

### 2017-11-09 ident/python2/ss-ident/kotte_model.py
remove old code performing identifiability for each parameter with separate variables

@@ -830,6 +830,7 @@ def establish_kotte_flux_identifiability(experimental_data_list):
         ident_fun_val.append(signed_ident_values[id*12:(id+1)*12, -1])
     p_list = [[p_id for p_id, val in enumerate(data_set) if val > 0] for data_set in ident_fun_val]
 
+    # identify perturbations that result in positive value for identifiability
     parameter_perturbation_list = []
     for all_parameter_iter in range(0, 12):
         parameter_perturbation_list.append(tuple([id for id, parameter_ids in enumerate(p_list)
@@ -838,35 +839,7 @@ def establish_kotte_flux_identifiability(experimental_data_list):
                for f_index, p_limit in enumerate(parameters_per_flux)
                for p_index in range(0, p_limit)]
     # build dictionary of results
-    fp_dict = dict(zip(fp_list, parameter_perturbation_list))
-
-    # identify perturbations that result in positive value for identifiability
-    p = [0] + parameters_per_flux
-    p = np.cumsum(p).tolist()
-
-
-
-    perturbation_list_flux_1 = []
-    for parameter_id in range(0, 4):
-        perturbation_list_flux_1.append(
-            tuple([id for id in range(flux1_sign_values.shape[0]) if (flux1_sign_values[id, parameter_id]>0)]))
-    perturbation_list_flux_2 = []
-    for parameter_id in range(0, 2):
-        perturbation_list_flux_2.append(
-            tuple([id for id in range(flux2_sign_values.shape[0]) if (flux2_sign_values[id, parameter_id]>0)]))
-    perturbation_list_flux_3 = []
-    for parameter_id in range(0, 6):
-        perturbation_list_flux_3.append(
-            tuple([id for id in range(flux3_sign_values.shape[0]) if (flux3_sign_values[id, parameter_id]>0)]))
-
-    # get common perturbation ids between differnt perturbation tuples above
-    perturbation_list = [tuple(perturbation_list_flux_1),
-                         tuple(perturbation_list_flux_2),
-                         tuple(perturbation_list_flux_3)]
-    #list_length = 3
-    #for ilist in perturbation_list:
-    #    number_of_pramaters = len(ilist)
-    #    print number_of_pramaters
+    perturbation_dict = dict(zip(fp_list, parameter_perturbation_list))
 
     # write results to file
[…]

### 2017-11-09 ident/python2/ss-ident/kotte_model.py
move file writing to the end

@@ -808,6 +808,28 @@ def establish_kotte_flux_identifiability(experimental_data_list):
     number_fluxes = len(ident_function_list)
     number_data = len(experimental_data_list)
     ident_values, parameters_per_flux = get_ident_value(ident_function_list, experimental_data_list)
+
+    # create signed boolean array for identifiability
+    signed_ident_values = np.sign(ident_values)
+
+    ident_fun_val = []
+    for id in range(0, number_data):
+        ident_fun_val.append(signed_ident_values[id * 12:(id + 1) * 12, -1])
+    p_list = [[p_id for p_id, val in enumerate(data_set) if val > 0] for data_set in ident_fun_val]
+
+    # identify perturbations that result in positive value for identifiability
+    parameter_perturbation_list = []
+    for all_parameter_iter in range(0, 12):
+        parameter_perturbation_list.append(tuple([id for id, parameter_ids in enumerate(p_list)
+                                                  if all_parameter_iter in parameter_ids]))
+    fp_list = ['flux{}p{}'.format(f_index + 1, p_index + 1)
+               for f_index, p_limit in enumerate(parameters_per_flux)
+               for p_index in range(0, p_limit)]
+
+    # build dictionary of results
+    perturbation_dict = dict(zip(fp_list, parameter_perturbation_list))
+
+    # write/append all data to file
     write_2_file_data = []
     flux_id, parameter_id = 1, 1
     for values in ident_values:
@@ -823,23 +845,8 @@ def establish_kotte_flux_identifiability(experimental_data_list):
             else:
                 flux_id = 1
             parameter_id = 1
-    # create signed boolean array for identifiability
-    signed_ident_values = np.sign(ident_values)
-    ident_fun_val = []
-    for id in range(0, number_data):
-        ident_fun_val.append(signed_ident_values[id*12:(id+1)*12, -1])
-    p_list = [[p_id for p_id, val in enumerate(data_set) if val > 0] for data_set in ident_fun_val]
 
-    # identify perturbations that result in positive value for identifiability
-    parameter_perturbation_list = []
-    for all_parameter_iter in range(0, 12):
-        parameter_perturbation_list.append(tuple([id for id, parameter_ids in enumerate(p_list)
-                                            if all_parameter_iter in parameter_ids]))
-    fp_list = ['flux{}p{}'.format(f_index + 1, p_index + 1)
-               for f_index, p_limit in enumerate(parameters_per_flux)
[…]

### 2017-11-09 ident/python2/ss-ident/kotte_model.py
write perturbation indices to file

@@ -829,14 +829,14 @@ def establish_kotte_flux_identifiability(experimental_data_list):
     # build dictionary of results
     perturbation_dict = dict(zip(fp_list, parameter_perturbation_list))
 
-    # write/append all data to file
+    # create data for write/append all data to file
     write_2_file_data = []
+    write_2_file_data.append(['Identifiable Perturbations'])
     flux_id, parameter_id = 1, 1
-    for values in ident_values:
+    for values in perturbation_dict:
         write_2_file_data.append(['Flux {}, Parameter {}'.format(flux_id, parameter_id)])
-        # print(['Flux {}, Parameter {}'.format(flux_id, parameter_id)])
-        write_2_file_data.append(values)
-        write_2_file_data.append(np.sign(values))
+        write_2_file_data.append(['Identifiable Perturbations'])
+        write_2_file_data.append(perturbation_dict[values])
         if parameter_id < parameters_per_flux[flux_id-1]:
             parameter_id += 1
         else:
@@ -846,7 +846,21 @@ def establish_kotte_flux_identifiability(experimental_data_list):
                 flux_id = 1
             parameter_id = 1
 
-
+    for values in ident_values:
+        write_2_file_data.append(['Flux {}, Parameter {}'.format(flux_id, parameter_id)])
+        # print(['Flux {}, Parameter {}'.format(flux_id, parameter_id)])
+        write_2_file_data.append(values)
+        write_2_file_data.append(np.sign(values))
+        #write_2_file_data.append(['Identifiable Perturbations'])
+        #write_2_file_data.append
+        if parameter_id < parameters_per_flux[flux_id-1]:
+            parameter_id += 1
+        else:
+            if flux_id < number_fluxes:
+                flux_id += 1
+            else:
+                flux_id = 1
+            parameter_id = 1
 
     # write results to file
     path = "~" + "shyam" + r"\Documents\Courses\CHE1125Project\Results\ident\python2\kotte_ident_results.txt"

### 2017-11-09 ident/python2/ss-ident/kotte_model.py
remove code to add ident_values data to file

@@ -835,24 +835,7 @@ def establish_kotte_flux_identifiability(experimental_data_list):
     flux_id, parameter_id = 1, 1
     for values in perturbation_dict:
         write_2_file_data.append(['Flux {}, Parameter {}'.format(flux_id, parameter_id)])
-        write_2_file_data.append(['Identifiable Perturbations'])
         write_2_file_data.append(perturbation_dict[values])
-        if parameter_id < parameters_per_flux[flux_id-1]:
-            parameter_id += 1
-        else:
-            if flux_id < number_fluxes:
-                flux_id += 1
-            else:
-                flux_id = 1
-            parameter_id = 1
-
-    for values in ident_values:
-        write_2_file_data.append(['Flux {}, Parameter {}'.format(flux_id, parameter_id)])
-        # print(['Flux {}, Parameter {}'.format(flux_id, parameter_id)])
-        write_2_file_data.append(values)
-        write_2_file_data.append(np.sign(values))
-        #write_2_file_data.append(['Identifiable Perturbations'])
-        #write_2_file_data.append
         if parameter_id < parameters_per_flux[flux_id-1]:
             parameter_id += 1
         else:

### 2017-11-09 ident/python2/ss-ident/kotte_sym_ident.py
remove all result printing code from script

@@ -39,24 +39,9 @@ experimental_datasets = \
     arrange_experimental_data(noisy_exp_xss, noisy_exp_fss, perturbed_parameter_values, exp_flux_index)
 
 # identifiability for all kotte fluxes
-perturbation_list, identifiability_value_signs = establish_kotte_flux_identifiability(experimental_datasets[0:10])
-print('Perturbations allowing identifiability:\n')
-print 'Flux 1 Parameter 1\n{}\nFlux 1 Parameter 2\n{}\nFlux 1 Parameter 3\n{}\nFlux 1 Parameter 4\n{}\n'.\
-    format(perturbation_list[0][0], perturbation_list[0][1], perturbation_list[0][2], perturbation_list[0][3])
-print 'Flux 2 Parameter 1\n{}\nFlux 2 Parameter 2\n{}\n'.format(perturbation_list[1][0], perturbation_list[1][1])
-print 'Flux 3 Parameter 1\n{}\nFlux 3 Parameter 2\n{}\nFlux 3 Parameter 3\n{}\n' \
-      'Flux 3 Parameter 4\n{}\nFlux 3 Parameter 5\n{}\nFlux 3 Parameter 6\n{}\n'.\
-    format(perturbation_list[2][0], perturbation_list[2][1], perturbation_list[2][2],
-           perturbation_list[2][3], perturbation_list[2][4], perturbation_list[2][5])
-#print('Flux 1 Identifiability:\n')
-#for index in range(1, 100):
-#    print flux1[0][index, :], flux1[1][index, :], flux1[2][index, :]
-#print('Flux 2 Identifiability:\n')
-#for index in range(1, 100):
-#    print flux2[0][index, :], flux2[1][index, :], flux2[2][index, :]
-#print('Flux 3 Identifiability:\n')
-#for index in range(1, 100):
-#    print flux3[0][index, :], flux3[1][index, :], flux3[2][index, :]
+perturbation_list, _ = establish_kotte_flux_identifiability(experimental_datasets[0:10])
+print('Perturbation analysis for identifiability complete.\n')
+
 
 # clear workspace (removes all module names and objects)
 # import sys

### 2017-11-09 ident/python2/ss-ident/kotte_model.py
correct for random key order in dictionary

@@ -833,9 +833,11 @@ def establish_kotte_flux_identifiability(experimental_data_list):
     write_2_file_data = []
     write_2_file_data.append(['Identifiable Perturbations'])
     flux_id, parameter_id = 1, 1
-    for values in perturbation_dict:
+    perturbation_keys = perturbation_dict.keys()
+    for id in range(0, len(perturbation_keys)):
         write_2_file_data.append(['Flux {}, Parameter {}'.format(flux_id, parameter_id)])
-        write_2_file_data.append(perturbation_dict[values])
+        key_id = 'flux{}p{}'.format(flux_id, parameter_id)
+        write_2_file_data.append(perturbation_dict[key_id])
         if parameter_id < parameters_per_flux[flux_id-1]:
             parameter_id += 1
         else:

### 2017-11-09 ident/docs/ident_manuscript.tex
ident manuscript files (figures, text and refs)

--- 2017-11-04 KineticModel/estimation/docs/formulation/ident.tex
@@
 
-	\subsection{Algorithm to calculate profile likelihood:}		
-	\begin{enumerate}
-		\item Start by solving the numerical optimization problem in Equation \ref{eq:pl} for a fixed value of $\theta_i$ to determine the initial point on the profile likelihood curve.
-		\item Take an increasing/decreasing step $\theta_{step}$ in the direction of $\theta_i$.\\
-		The increasing/decreasing steps in $\theta_i$ can be adapted to the shape of the profile likelihood curve that is encountered while running the algorithm.
-Accordingly, the step $\theta_{step}$ should satisfy Equation (\ref{eq:adaptivestep}) below, where $\theta^{k-1}$ refers to parameter estimates obtained from iteration $k-1$ of the above algorithm and $q\in \left[0,1\right]$.
-		\begin{align}\label{eq:adaptivestep}
-		\chi^2\left(\theta^{k-1}+\theta_{step}\right) - \chi^2(\theta^{k-1}) \approx q.\Delta_{\alpha}
-		\end{align}
-		\item Re-optimize $\theta_{j\ne i}$ using Equation \ref{eq:pl}.
-		\item Repeat the last two steps until a desired threshold $\Delta_{\alpha}$ is exceeded or a maximal amount of steps is reached.
-	\end{enumerate}
-
 	Due to the dependence of practical parameter identifiability on the experimental data, the profile likelihood approach can be used to design experiments in such a way that the observables that are derived from these experiments can improve the practical identifiability of the parameters.
@@
 In contrast, performing non-informative experiments without prior knowledge on their ability to change the identifiability of the parameter may provide data that cannot be used to estimate parameter $\theta_i$ (Figure \ref{fig:edwithpl}c).
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/ed_with_pl}}
-		\caption{Cartoon illustrating the utility of identifiability analysis for experimental design.
-a) The profile likelihood of a practically non-identifiable parameter that needs to be estimated based on both good and bad experimental data.
-The changes in the profile likelihood of the parameter when estimated with b) good experimental data and c) bad experimental.
-The practical identifiability of a parameter is dependent on the quality and quantity of experimental data used to estimate the parameter.
-The identifiability of the parameter, as determined using its profile likelihood, does not change due to the poor quality/quantity of experiments.}\label{fig:edwithpl}
-	\end{figure}
 
@@
 
-	\subsection{Parameter estimation in kinetic models of metabolism:}
-	The formulation for parameter estimation assumes that the concentrations and the fluxes corresponding to each perturbation are variables in the optimization problem formulated to estimate kinetic parameters.
-
-	Let $\mathbf{x}\in\mathbb{R}^m$, $\mathbf{p}\in\mathbb{R}^l$ and $\mathbf{v}\in\mathbb{R}$ be the vector of concentrations, parameters and fluxes respectively.
-If the number of perturbations used for parameter estimation is in $\mathbb{R}^p$, then the vector of concentrations, fluxes and parameters that need to be estimated in the optimization problem changes to $\mathbf{x}\in\mathbb{R}^{mp}, \mathbf{v}\in\mathbb{R}^{p}$ and $\mathbf{p}\in\mathbb{R}^{l}$ respectively.
-
-	Let $i\in\mathbb{R}^{mp}$, $j\in\mathbb{R}^{p}$ and $k\in\mathbb{R}^{l}$ represent the indices for concentrations, fluxes and parameters respectively in Equations (\ref{eq:l2norm-flux})-(\ref{eq:noise-var}).
-
-	\paragraph{Formulation c):}
-	In order to estimate kinetic parameters for fluxes in models of metabolism, we propose a formulation accounting for the uncertainty in the concentration and flux estimates.
-The uncertainty terms ($\epsilon$) are used as bounds for the flux and the concentrations and are included in the optimization objective as a variable to be penalized is given in Equation (\ref{eq:noise-var}) below.
-Note that the concentration and the fluxes have different uncertainties, $\epsilon_x$ and $\epsilon_v$ respectively, associated with them.
-$w_1$, $w_2$, $w_3$ and $w_4$ are the weights associated with the four different terms in the objective.
-	\begin{center}
-		\begin{subequations}\label{eq:noise-var}
-			\begin{align}
-			\underset{\mathbf{x},\mathbf{p},\mathbf{v},\epsilon_x,\epsilon_v}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert + w_3\epsilon_x + w_4\epsilon_v\\
-			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-			& \text{      }x^*_i(1-\epsilon_x)\le x_i \le x^*_i(1+\epsilon_x) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-			& \text{      }v^*_j(1-\epsilon_v)\le v_j \le v^*_j(1+\epsilon_v) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-			\end{align}
-		\end{subequations}
-	\end{center}
-
 	\begin{table}[!tbhp]
@@
 
-	\begin{table}[!tbhp]
-		\caption{Table showing the ability of different formulations mentioned above to estimate enzyme kinetic parameters for different fluxes based on the perturbations given in the above table.}
-		\begin{center}				
-			\begin{tabular}{cccccc}
-				\hline
-				Flux & Data Type & \multicolumn{2}{c}{Formulation (b)} & \multicolumn{2}{c}{Formulation (c)}\\
-				\hline
-				&&$\epsilon_x$(\%)&$\epsilon_v$(\%)&$\epsilon_x$(\%)&$\epsilon_v$(\%)\\
-				\hline
-				\multirow{2}{*}{flux 1} & no noise & 30 & 5 & 33.33 & 0\\
-												  &	noisy & x & x & 0 & 33.33\\
-			  	\hline
-				 \multirow{2}{*}{flux 2} & no noise & 90 & 30 & 0 & 31\\
-				 									& noisy & 85 & 40 & 0 & 30\\
-				\hline				 									
-				\multirow{2}{*}{flux 3} & no noise & 15 & 10 & 17 & 0\\
-													& noisy & x & x & 17 & 0\\
-				\hline			
-			\end{tabular}
-		\end{center}	
-		\label{tab:res1}
-	\end{table}
-	\subsection{Estimation of flux 1:}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_conc_aug6}}
-	%	\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_conc_aug6}}
-	%	\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_flux_aug6}}
-	%	\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_flux_aug6}}
-	%	\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_1$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/no_noise/typec/est_flux1_par_aug6}}
-	%	\caption{Plot comparing noiseless model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux1/noisy/typec/est_flux1_par_aug6}}
-	%	\caption{Plot comparing noisy model generated parameters for $v_1$ through a multi-start search algorithm using formulation (c). }
-	%\end{figure}
-	\clearpage
-
-	\subsection{Estimation of flux 2:}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_conc_aug6}}
-	%	\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_conc_aug6}}
-	%	\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_flux_aug6}}
-	%	\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_flux_aug6}}
-	%	\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_2$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/no_noise/typec/est_flux2_par_aug6}}
-	%	\caption{Plot comparing noiseless model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux2/noisy/typec/est_flux2_par_aug6}}
-	%	\caption{Plot comparing noisy model generated parameters for $v_2$ through a multi-start search algorithm using formulation (c). }
-	%\end{figure}
-	\clearpage
-
-	\subsection{Estimation of flux 3:}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_conc_aug6}}
-	%	\caption{Plot comparing noiseless model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_conc_aug6}}
-	%	\caption{Plot comparing noisy model generated steady state metabolite concentrations with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_flux_aug6}}
-	%	\caption{Plot comparing noiseless model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_flux_aug6}}
-	%	\caption{Plot comparing noisy model generated steady state fluxes with model estimates obtained using an optimized model for $v_3$ using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/no_noise/typec/est_flux3_par_aug6}}
-	%	\caption{Plot comparing noiseless model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
-	%\end{figure}
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=.5\textwidth,height=.5\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/estimation/est_flux3/noisy/typec/est_flux3_par_aug6}}
-	%	\caption{Plot comparing noisy model generated parameters for $v_3$ through a multi-start search algorithm using formulation (c). }
-	%\end{figure}
-	%\clearpage
-
-	\section{Appendix}
-	\subsection{Formulation a):}		
-	\begin{subequations}\label{eq:l2norm-flux}
-		\begin{align}
-		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert\\
-		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-		\end{align}
-	\end{subequations}
-
-	\subsection{Formulation b):}	
-	In this formulation, in Equation (\ref{eq:l2norm-conc}), the flux and concentrations are constrained as earlier with an uncertainty term.	
-	\begin{subequations}\label{eq:l2norm-conc}
-		\begin{align}
-		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
-		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-		& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-		& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-		\end{align}
-	\end{subequations}
-
-	\subsection{Formulation d):}
-	Due to the inability of Equation (\ref{eq:l2norm-flux}) to get feasible results, we modified the problem to include the L2-norm of the concentrations as well in the objective function in Equation (\ref{eq:free-var}a).
-In Equation (\ref{eq:free-var}) the flux and the concentrations are free variables, i.e.,  $\mathbf{x}\in\left[\mathbf{0},+\mathbf{\infty}\right]$.
-	\begin{subequations}\label{eq:free-var}
-		\begin{align}
-		\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
-		\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
-		& \text{      }0\le x_i \le +\infty & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
-		& \text{      }0\le v_j \le +\infty & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
-		& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
-		\end{align}
-	\end{subequations}
-
-	After testing the problem in Equation (\ref{eq:free-var}) above for the acetate uptake flux $v_1$, we found that all solutions using IPOPT exceeded the maximum number of iterations (set at 5000) and never produced an optimal solution.
-Hence, we use a modified form of the bounds in Equation (\ref{eq:free-var}) above in Equation (\ref{eq:l2norm-conc}).
-
 	\printbibliography

### 2017-11-13 ident/docs/ident_manuscript.tex
add intro outline

@@
 	\section{Introduction:}
+	\paragraph{Outline:}
+	\begin{itemize}
+		\item How dynamic models of metabolism are defined in biology
+		\item How parameter values is important for model-based prediction and design of metabolism
+		\item importance of parameter identifiability, structural and practical identifiability for dynamic models of metabolism
+		\item work done in the are of structural identifiability (numerical and symbolic/computer algebra methods) - local vs global identifiability
+		\item work done in the area of practical identifiability (numerical and symbolic/computer algebra methods)
+		\item work done in this paper for practical identifiability
+		\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable)
+	\end{itemize}
 	\subsection{Identifiability analysis: Definitions and Formulations}

### 2017-11-14 ident/docs/ident_manuscript.tex
add stements about project in intro

@@
 		\item work done in this paper for practical identifiability
-		\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable)
+		Here, we present a computer algebra-based method to establish practical identifiability of kinetic models of metabolic networks.
+Our method establishes posterior identifiability for each individual flux separately and can potentially be scaled-up to models of large metabolic networks.
+To illustrate this point, we have demonstrated the application of our methodology for the kinetic model of the red blood cell hepatocyte metabolic network.
+		\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable) (move to discussion - may be)
 	\end{itemize}

### 2017-11-15 ident/docs/ident_manuscript.tex
added a complete intro w/o final paragraph

@@
 	\section{Introduction:}	
+	Kinetic models of metabolic networks are typically represented as ordinary differential equations (ODE) wherein the metabolite concentrations, $\mathbf{x}$, are state variables (Equation \ref{eq:kinode}a).
+The changes in $\mathbf{x}$ are expressed as the product of the stoichiometric matrix, $S$, and the vector of the metabolic network reaction fluxes, $\mathbf{v}$.
+	\begin{subequations}\label{eq:kinode}
+		\begin{align}
+		\dot{\mathbf{x}} = S\mathbf{v}\\
+		\mathbf{v} = \mathbf{f}(\mathbf{x}, p, u)
+		\end{align}
+	\end{subequations}	
+	The fluxes on the right hand side of Equation (\ref{eq:kinode}a) can be expressed as nonlinear functions of the states ($\mathbf{x}$), kinetic parameters ($p$) and other input variables ($u$) as in Equation (\ref{eq:kinode}b).
+In silico simulation of Equation (\ref{eq:kinode}) to predict responses of the in vivo system to changes in either the enzyme kinetic parameters ($p$) or other inputs ($u$) requires knowledge of the parameter values under in vivo conditions.
+Accordingly, parameter estimation based on experimentally observed states is an integral component of kinetic modeling of metabolism.
+
+	However, not all states are typically observable through experiments.
+Hence, for parameter estimation, Equation (\ref{eq:kinode}) is usually augmented with a system that defines the relationship between the experimentally observable output variables ($y$) and the states ($x$) as in Equation (\ref{eq:output}).
+The parameters used to establish this relationship in Equation (\ref{eq:output}b), $\mu$, may or may not include system parameters $p$ defined in Equation (\ref{eq:output}a).
+	\begin{subequations}\label{eq:output}
+		\begin{align}
+		\dot{\mathbf{x}} = g(\mathbf{x}, p, u)\\
+		\mathbf{y} = h(\mathbf{x},\mu,u)
+		\end{align}
+	\end{subequations}
+
+	The parameter estimation problem is commonly formulated as a nonlinear programming problem with the objective of minimizing of least square error between the measured ($y_{kl}^*$) and modeled ($y_{kl}$) outputs over the time course ($l = {1, ..., d}$) for which experimental data is collected (Equation \ref{eq:chi2}) (Raue et al., and other parameter estimation/identifiability papers).
+	\begin{align}\label{eq:chi2}
+	\chi^2(\theta) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
+	\end{align}
+	The difference between the data and the model estimate for each state and at each time point is weighted by the variance in the experimental data $\sigma_{kl}^*$ for that corresponding variable and time point.
+For large and computationally intractable optimization problems, parameter estimation can also be posed as a Bayesian problem (Neilsen, Scientific Reports).
+	The ability to determine a unique solution to the parameters ($p$ and $\mu$) in Equation (\ref{eq:output}) through the aforementioned parameter estimation problem is however governed by the identifiability of the parameters in the model \parencite{McLean2012}.
+
+	The identifiability of parameters in nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability. 	
+	Any system (Equation \ref{eq:output}) is said to be structurally identifiable if, for an input-output mapping defined by $\mathbf{y} = \Phi(\mu,u)$ for at least one input function $u$, any two values of parameters $\mathbf{\mu}_1$ and $\mathbf{\mu}_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
+	\begin{align}\label{eq:stident}
+	\Phi(\mathbf{\mu}_1,u) = \Phi(\mathbf{\mu}_2,u) \iff \mathbf{\mu}_1 = \mathbf{\mu}_2
+	\end{align}
+	Accordingly, any system that has an infinite number of solutions to the parameter estimation problem for all input functions is said to structurally non-identifiable.
+Thus, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unobservable system states and the observable system outputs.
+Accordingly, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
+
+	Experimental data from biological systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
+The effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
+Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
+
+	Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated.
+On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
+
+	Algorithms have been extensively developed for establishing structural identifiability of dynamic models, that of biological systems in particular (IEEE Trans paper from 2007), that involve methods based on differential algebra (Glad and Ljung, 1994).
+However, these methods not only scale poorly with increases in size of the modeled system, but also require dynamic time course data of the observable variables of the system.
+While computational burden due to poor scalability can be partly addressed with the current increase in computational power, the ability to obtain dynamic data for establishing identifiability of parameters in kinetic models of metabolism still remains a challenge.
+
+	In this paper, we propose a methodology to establish practical identifiability of parameters in kinetic models of metabolism that addresses the twin issues of scalability as well as data availability.
+
 	\paragraph{Outline:}
@@
 		\item How dynamic models of metabolism are defined in biology
-		\item How parameter values is important for model-based prediction and design of metabolism
+		\item How parameter values are important for model-based prediction and design of metabolism
 		\item importance of parameter identifiability, structural and practical identifiability for dynamic models of metabolism
@@
 		\item work done in the area of practical identifiability (numerical and symbolic/computer algebra methods)
-		\item work done in this paper for practical identifiability
+		\item work done in this paper for practical identifiability\\
 		Here, we present a computer algebra-based method to establish practical identifiability of kinetic models of metabolic networks.
@@
 	\end{itemize}
+
 	\subsection{Identifiability analysis: Definitions and Formulations}
@@
 
-	Any system (Equation \ref{eq:dynmodel}) is said to be structurally or a priori identifiable if, for an input-output mapping defined by $\mathbf{y} = \Phi(\mu,u)$ for at least one input function $u$, any two values of parameters $\mathbf{\mu}_1$ and $\mathbf{\mu}_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
-	\begin{align}\label{eq:stident}
-	\Phi(\mathbf{\mu}_1,u) = \Phi(\mathbf{\mu}_2,u) \iff \mathbf{\mu}_1 = \mathbf{\mu}_2
-	\end{align}
-	Accordingly, any system that has an infinite number of solutions to the parameter estimation problem for all input functions is said to structurally non-identifiable.
-The effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
-
-	When parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
-The effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
-Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
 	%Thus, the identifiability of parameters in nonlinear models is dependent on the model structure, parameterization, and the quality and quantity of experimental data that is available for the purpose of estimation.
-	Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated.
-On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
 
@@
 	 \end{align}
-	 \begin{align}\label{eq:chi2}
-	 \chi^2(\theta) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
-	 \end{align}
+
 	 In the minimization objective shown in Equation (\ref{eq:chi2}) for parameter estimation, $y_{kl}^*$ is the available experimental time course data for each observable state $k$ at each $l$ time point.
@@
 	\section{Results:}
+	\paragraph{Outline:}
+	\begin{itemize}
+		\item parameter estimation is a well developed field typically using minimization of least square error to estimate model parameters from available experimental data
+		\item if parameters are structurally identifiable, it does not guarantee practical identifiability from noisy experimental data
+		\item identifiability dependent on whether given datasets (outputs) for estimation can sufficiently distinguish between different parameter values
+	\end{itemize}
+
+	\paragraph{Sections:}
+	\begin{itemize}
+		\item datasets required for parameter estimation in kinetic models of metabolism (methods?)		
+		\item identifiability in kotte model - scalability, number of experiments required, requirements for time course data(? in the intro)		
+		\item identifiability in large rbc model
+	\end{itemize}
 

### 2017-11-23 ident/docs/ident_manuscript.tex

@@
 	In this paper, we propose a methodology to establish practical identifiability of parameters in kinetic models of metabolism that addresses the twin issues of scalability as well as data availability.
+We illustrate the utility of our methods by demonstrating its ability to establish identifiability for a small network of gluconeogenesis in \textit{E.
+coli} and a relatively larger metabolic network of the human red blood cell hepatocyte.
+
+	Here, we present a computer algebra-based method to establish practical identifiability of kinetic models of metabolic networks.
+Our method establishes posterior identifiability for each individual flux separately and can potentially be scaled-up to models of large metabolic networks.
+To illustrate this point, we have demonstrated the application of our methodology for the kinetic model of the red blood cell hepatocyte metabolic network.
 
@@
 	\begin{itemize}
-		\item How dynamic models of metabolism are defined in biology
-		\item How parameter values are important for model-based prediction and design of metabolism
-		\item importance of parameter identifiability, structural and practical identifiability for dynamic models of metabolism
-		\item work done in the are of structural identifiability (numerical and symbolic/computer algebra methods) - local vs global identifiability
-		\item work done in the area of practical identifiability (numerical and symbolic/computer algebra methods)
+		\item Importance of parameter values for estimating in vivo response using kinetic models of metabolism and consequently for metabolic network design using kinetic models of metabolism
+		\item The need for parameter identifiability to determine unique and true parameter values from observed data
+		\item Types and purpose of identifiability for parameters
+		\item Methods for structural identifiability and existing methods for practical identifiability
+		\item Lack of methods for practical identifiability and consequently experimental design
 		\item work done in this paper for practical identifiability\\			
-		Here, we present a computer algebra-based method to establish practical identifiability of kinetic models of metabolic networks.
-Our method establishes posterior identifiability for each individual flux separately and can potentially be scaled-up to models of large metabolic networks.
-To illustrate this point, we have demonstrated the application of our methodology for the kinetic model of the red blood cell hepatocyte metabolic network.
 		\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable) (move to discussion - may be)

### 2017-11-30 ident/docs/ident_manuscript.tex
change intro and methods

@@
 \section{Introduction:}	
-	Kinetic models of metabolic networks are typically represented as ordinary differential equations (ODE) wherein the metabolite concentrations, $\mathbf{x}$, are state variables (Equation \ref{eq:kinode}a).
-The changes in $\mathbf{x}$ are expressed as the product of the stoichiometric matrix, $S$, and the vector of the metabolic network reaction fluxes, $\mathbf{v}$.
-	\begin{subequations}\label{eq:kinode}
-		\begin{align}
-		\dot{\mathbf{x}} = S\mathbf{v}\\
-		\mathbf{v} = \mathbf{f}(\mathbf{x}, p, u)
-		\end{align}
-	\end{subequations}	
-	The fluxes on the right hand side of Equation (\ref{eq:kinode}a) can be expressed as nonlinear functions of the states ($\mathbf{x}$), kinetic parameters ($p$) and other input variables ($u$) as in Equation (\ref{eq:kinode}b).
-In silico simulation of Equation (\ref{eq:kinode}) to predict responses of the in vivo system to changes in either the enzyme kinetic parameters ($p$) or other inputs ($u$) requires knowledge of the parameter values under in vivo conditions.
-Accordingly, parameter estimation based on experimentally observed states is an integral component of kinetic modeling of metabolism.
-
-	However, not all states are typically observable through experiments.
-Hence, for parameter estimation, Equation (\ref{eq:kinode}) is usually augmented with a system that defines the relationship between the experimentally observable output variables ($y$) and the states ($x$) as in Equation (\ref{eq:output}).
-The parameters used to establish this relationship in Equation (\ref{eq:output}b), $\mu$, may or may not include system parameters $p$ defined in Equation (\ref{eq:output}a).
+The dynamic response of any system can be modeled as a set of ordinary differential equations (ode) (Equation \ref{eq:output}a) in the state variables ($\mathbf{x}$).
+The changes in the state variables are expressed as a function of the state variables, the system parameters ($p$) and the inputs to the system ($u$).
 \begin{subequations}\label{eq:output}
@@
 \end{subequations}
+When not all states are observable, Equation (\ref{eq:output}a) is usually augmented with a system that defines the relationship between the observable output variables ($\mathbf{y}$) and the state variables ($\mathbf{x}$) as in Equation (\ref{eq:output}b).
+The parameters used to establish this relationship (Equation \ref{eq:output}b), $\mu$, may or may not include system parameters $p$ defined in Equation (\ref{eq:output}a).
 
-	The parameter estimation problem is commonly formulated as a nonlinear programming problem with the objective of minimizing of least square error between the measured ($y_{kl}^*$) and modeled ($y_{kl}$) outputs over the time course ($l = {1, ..., d}$) for which experimental data is collected (Equation \ref{eq:chi2}) (Raue et al., and other parameter estimation/identifiability papers).
-	\begin{align}\label{eq:chi2}
+Typically, a nonlinear programming formulation with the objective of minimizing the least square error between the measured ($y_{kl}^*$) and modeled ($y_{kl}$) outputs over the time course ($l = {1, ..., d}$) for which experimental data is collected (Equation \ref{eq:minnlp}) (Raue et al., and other parameter estimation/identifiability papers), is used to estimate parameters $p$ and $\mu$ in Equation \ref{eq:output}.
+\begin{align}\label{eq:minnlp}
 \chi^2(\theta) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
 \end{align}
-	The difference between the data and the model estimate for each state and at each time point is weighted by the variance in the experimental data $\sigma_{kl}^*$ for that corresponding variable and time point.
-For large and computationally intractable optimization problems, parameter estimation can also be posed as a Bayesian problem (Neilsen, Scientific Reports).
-	The ability to determine a unique solution to the parameters ($p$ and $\mu$) in Equation (\ref{eq:output}) through the aforementioned parameter estimation problem is however governed by the identifiability of the parameters in the model \parencite{McLean2012}.
+The difference between the data and the model estimate for each output, and at each time point, is weighted by the variance in the experimental data $\sigma_{kl}^*$ for that corresponding variable and time point.
+The minimization problem is solved subject to constraints on the state and output variables expressed in Equation \ref{eq:output}, as well as bounds on the estimated parameter values.
+The ability to determine a unique solution to the parameter estimation problem is however governed by the identifiability of the parameters in the model \parencite{McLean2012}.
 
@@
 
-	Experimental data from biological systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
+Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
 The effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
@@
 
-	Algorithms have been extensively developed for establishing structural identifiability of dynamic models, that of biological systems in particular (IEEE Trans paper from 2007), that involve methods based on differential algebra (Glad and Ljung, 1994).
+The dynamics of metabolic networks can also be represented by Equation (\ref{eq:output}) wherein the metabolite concentrations, $\mathbf{x}$, are state variables.
+The changes in the metabolite concentrations ($\mathbf{x}$) are expressed as a function of reaction rates dependent on enzyme kinetics, which are in turn expressed as nonlinear functions of the states ($\mathbf{x}$), kinetic parameters ($p$) and other input variables ($u$).
+
+Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems (IEEE Trans paper from 2007).
+Most of these algorithms use methods based on differential algebra (Glad and Ljung, 1994).
+More recent methods take a profile-likelihood-based approach (2012 Paper) to establish both structural and practical identifiability.
 However, these methods not only scale poorly with increases in size of the modeled system, but also require dynamic time course data of the observable variables of the system.
-While computational burden due to poor scalability can be partly addressed with the current increase in computational power, the ability to obtain dynamic data for establishing identifiability of parameters in kinetic models of metabolism still remains a challenge.
+While computational burden due to poor scalability can be partly addressed with the current increases in computational power, the ability to obtain dynamic data for establishing identifiability of parameters in kinetic models of metabolism still remains a challenge.
 
-	In this paper, we propose a methodology to establish practical identifiability of parameters in kinetic models of metabolism that addresses the twin issues of scalability as well as data availability.
-We illustrate the utility of our methods by demonstrating its ability to establish identifiability for a small network of gluconeogenesis in \textit{E.
-coli} and a relatively larger metabolic network of the human red blood cell hepatocyte.
-
-	Here, we present a computer algebra-based method to establish practical identifiability of kinetic models of metabolic networks.
-Our method establishes posterior identifiability for each individual flux separately and can potentially be scaled-up to models of large metabolic networks.
-To illustrate this point, we have demonstrated the application of our methodology for the kinetic model of the red blood cell hepatocyte metabolic network.
+In this paper, we propose a methodology to establish practical identifiability for parameters in kinetic models of metabolism.
+We present a computer algebra-based method that can facilitate experimental design for estimating parameters separately for each individual reaction within a metabolic network based on available steady state experimental data.
+This enables us to address the twin issues of scalability and data availability.
+We illustrate the utility of this method by applying it for a small network of gluconeogenesis in \textit{E.
+coli} and demonstrating our ability to propose experiments that will facilitate parameter estimation for a kinetic model of this network.
+We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
 
@@
 	\item Methods for structural identifiability and existing methods for practical identifiability
-		\item Lack of methods for practical identifiability and consequently experimental design
-		\item work done in this paper for practical identifiability\\			
-		\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable) (move to discussion - may be)
+	\item Lack of methods for practical identifiability and consequently experimental design (not covered?)
+	\item work done in this paper for practical identifiability			
+	\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable) (move to discussion - may be in discussion)
 \end{itemize}
@@
 	\section{Methods:}
-	 We use a profile likelihood-based approach \parencite{Raue2009a} to establish structural and practical identifiability of parameters in nonlinear kinetic models of metabolism.
-Briefly, the approach seeks to establish the existence/non-existence of bounds in confidence intervals for the estimates of parameters in nonlinear models.
-The profile likelihood is calculated based on Equation (\ref{eq:pl}) for each parameter $\theta_i$ where $\chi^2(\theta_i)$ is given by Equation (\ref{eq:chi2}).
-	 \begin{align}\label{eq:pl}
-	 \chi_{PL}^2(\theta_i) = \underset{\theta_{j\ne i}}{\mathrm{min}} \left[\chi^2(\theta)\right]
-	 \end{align}
-
-	 In the minimization objective shown in Equation (\ref{eq:chi2}) for parameter estimation, $y_{kl}^*$ is the available experimental time course data for each observable state $k$ at each $l$ time point.
-The difference between the data and the model estimates at these time points, $y_{kl}$ is weighted by the variance in the experimental data $\sigma_{kl}^*$.
-An algorithm to calculate the profile likelihood, $\chi_{PL}^2(\theta_i)$, based on Equation \ref{eq:pl} is given below.
-
-	 The identifiability of parameters is established through the confidence intervals of their estimates, $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
-The likelihood-based confidence interval for any parameter whose profile likelihood is estimated can be written on the basis of a threshold $\Delta_{\alpha}$ in the likelihood as in Equation (\ref{eq:pleci}).
-	 \begin{align}\label{eq:pleci}
-	 \{\theta|\chi^2(\theta)-\chi^2(\hat{\theta})<\Delta_{\alpha}\}
-	 \end{align}
-	 The threshold $\Delta_{\alpha}$ in the likelihood is the 1-$\alpha$ quantile of the $\chi^2$ distribution, represented as $\chi^2(\alpha,df)$.
-The confidence intervals obtained hold for $df$ degrees of freedom.
-For a choice of $df$=1 the confidence intervals will hold for each parameter individually, and confidence intervals that hold jointly for all parameters can be obtained by choosing the number of parameters as $df$.
-
-	The visualization of structurally and practically non-identifiable parameters using the profile likelihood approach is illustrated in Figure \ref{fig:identtypes}.
-The points of intersection between the profile likelihood curves (solid line) with the one parameter likelihood threshold ($\Delta_{\alpha} = \chi^2(\alpha,1)$, dashed line) provide the confidence intervals of the parameter $\theta_i$.
-The confidence intervals of a structurally non-identifiable parameter are unbounded, i.e., $\left[-\infty, +\infty\right]$ (Figure \ref{fig:identtypes}a), while the confidence intervals of a practically non-identifiable parameter are unbounded in at least one direction, i.e., $\left[\sigma_{i}^-,\sigma_{i}^+\right]$ where either $\sigma_{i}^- = -\infty$ or $\sigma_{i}^+ = +\infty$ (Figure \ref{fig:identtypes}b).
-If a parameter's estimates have a finite confidence interval then the parameter is said to be identifiable (Figure \ref{fig:identtypes}c).
-Note that the horizontal dotted lines in Figure \ref{fig:identtypes} represent the confidence interval thresholds ($\Delta_{\alpha}$) that are used to establish identifiability.
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.8\textwidth,height=.8\textheight,keepaspectratio]{figures/identifiability}}
-		\caption{The profile likelihood estimates of a) a structurally non-identifiable, b) a practically non-identifiable and c) an identifiable parameter.
-The horizontal axis represents the changes in the value of the parameter ($\theta_i$) whose identifiability is being determined and the profile likelihood ($\chi^2_{PL}(\theta_i)$) is shown in the vertical axis.
-The confidence interval threshold ($\Delta_{\alpha}$) used to determine the identifiability of the parameter is denoted by the horizontal dotted line.
-Identifiable parameters are distinguished from non-identifiable parameters by the presence of both upper and lower bounds on their confidence interval estimates $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
-}\label{fig:identtypes}
-	\end{figure}
-
-	Due to the dependence of practical parameter identifiability on the experimental data, the profile likelihood approach can be used to design experiments in such a way that the observables that are derived from these experiments can improve the practical identifiability of the parameters.
-We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.	Assuming a parameter $\theta_i$ is practically non-identifiable (Figure \ref{fig:edwithpl}a), performing a profile-likelihood based identifiability analysis using simulated data can help determine the nature of experiments needed to make the parameter identifiable (Figure \ref{fig:edwithpl}b).
-In contrast, performing non-informative experiments without prior knowledge on their ability to change the identifiability of the parameter may provide data that cannot be used to estimate parameter $\theta_i$ (Figure \ref{fig:edwithpl}c).
-
 	\subsection{A method to establish posterior identifiability of metabolic network models:}
@@
 
+	 We use a profile likelihood-based approach \parencite{Raue2009a} to establish structural and practical identifiability of parameters in nonlinear kinetic models of metabolism.
+Briefly, the approach seeks to establish the existence/non-existence of bounds in confidence intervals for the estimates of parameters in nonlinear models.
+The profile likelihood is calculated based on Equation (\ref{eq:pl}) for each parameter $\theta_i$ where $\chi^2(\theta_i)$ is given by Equation (\ref{eq:chi2}).
+	 \begin{align}\label{eq:pl}
+	 \chi_{PL}^2(\theta_i) = \underset{\theta_{j\ne i}}{\mathrm{min}} \left[\chi^2(\theta)\right]
+	 \end{align}
+
+	 In the minimization objective shown in Equation (\ref{eq:chi2}) for parameter estimation, $y_{kl}^*$ is the available experimental time course data for each observable state $k$ at each $l$ time point.
+The difference between the data and the model estimates at these time points, $y_{kl}$ is weighted by the variance in the experimental data $\sigma_{kl}^*$.
+An algorithm to calculate the profile likelihood, $\chi_{PL}^2(\theta_i)$, based on Equation \ref{eq:pl} is given below.
+
+	 The identifiability of parameters is established through the confidence intervals of their estimates, $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
+The likelihood-based confidence interval for any parameter whose profile likelihood is estimated can be written on the basis of a threshold $\Delta_{\alpha}$ in the likelihood as in Equation (\ref{eq:pleci}).
+	 \begin{align}\label{eq:pleci}
+	 \{\theta|\chi^2(\theta)-\chi^2(\hat{\theta})<\Delta_{\alpha}\}
+	 \end{align}
+	 The threshold $\Delta_{\alpha}$ in the likelihood is the 1-$\alpha$ quantile of the $\chi^2$ distribution, represented as $\chi^2(\alpha,df)$.
+The confidence intervals obtained hold for $df$ degrees of freedom.
+For a choice of $df$=1 the confidence intervals will hold for each parameter individually, and confidence intervals that hold jointly for all parameters can be obtained by choosing the number of parameters as $df$.
+
+	The visualization of structurally and practically non-identifiable parameters using the profile likelihood approach is illustrated in Figure \ref{fig:identtypes}.
+The points of intersection between the profile likelihood curves (solid line) with the one parameter likelihood threshold ($\Delta_{\alpha} = \chi^2(\alpha,1)$, dashed line) provide the confidence intervals of the parameter $\theta_i$.
+The confidence intervals of a structurally non-identifiable parameter are unbounded, i.e., $\left[-\infty, +\infty\right]$ (Figure \ref{fig:identtypes}a), while the confidence intervals of a practically non-identifiable parameter are unbounded in at least one direction, i.e., $\left[\sigma_{i}^-,\sigma_{i}^+\right]$ where either $\sigma_{i}^- = -\infty$ or $\sigma_{i}^+ = +\infty$ (Figure \ref{fig:identtypes}b).
+If a parameter's estimates have a finite confidence interval then the parameter is said to be identifiable (Figure \ref{fig:identtypes}c).
+Note that the horizontal dotted lines in Figure \ref{fig:identtypes} represent the confidence interval thresholds ($\Delta_{\alpha}$) that are used to establish identifiability.
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.8\textwidth,height=.8\textheight,keepaspectratio]{figures/identifiability}}
+		\caption{The profile likelihood estimates of a) a structurally non-identifiable, b) a practically non-identifiable and c) an identifiable parameter.
+The horizontal axis represents the changes in the value of the parameter ($\theta_i$) whose identifiability is being determined and the profile likelihood ($\chi^2_{PL}(\theta_i)$) is shown in the vertical axis.
+The confidence interval threshold ($\Delta_{\alpha}$) used to determine the identifiability of the parameter is denoted by the horizontal dotted line.
+Identifiable parameters are distinguished from non-identifiable parameters by the presence of both upper and lower bounds on their confidence interval estimates $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
+}\label{fig:identtypes}
+	\end{figure}
+
+	Due to the dependence of practical parameter identifiability on the experimental data, the profile likelihood approach can be used to design experiments in such a way that the observables that are derived from these experiments can improve the practical identifiability of the parameters.
+We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.	Assuming a parameter $\theta_i$ is practically non-identifiable (Figure \ref{fig:edwithpl}a), performing a profile-likelihood based identifiability analysis using simulated data can help determine the nature of experiments needed to make the parameter identifiable (Figure \ref{fig:edwithpl}b).
+In contrast, performing non-informative experiments without prior knowledge on their ability to change the identifiability of the parameter may provide data that cannot be used to estimate parameter $\theta_i$ (Figure \ref{fig:edwithpl}c).
+
 	\section{Results:}

### 2017-11-30 ident/docs/ident_manuscript.tex
changes to methods section

@@
 The dynamic response of any system can be modeled as a set of ordinary differential equations (ode) (Equation \ref{eq:output}a) in the state variables ($\mathbf{x}$).
-The changes in the state variables are expressed as a function of the state variables, the system parameters ($p$) and the inputs to the system ($u$).
+The changes in the state variables are expressed as a function of the state variables, the system parameters ($\theta$) and the inputs to the system ($u$).
 \begin{subequations}\label{eq:output}
 	\begin{align}
-	\dot{\mathbf{x}} = g(\mathbf{x}, p, u)\\
+	\dot{\mathbf{x}} = g(\mathbf{x}, \theta, u)\\
 	\mathbf{y} = h(\mathbf{x},\mu,u)
@@
 When not all states are observable, Equation (\ref{eq:output}a) is usually augmented with a system that defines the relationship between the observable output variables ($\mathbf{y}$) and the state variables ($\mathbf{x}$) as in Equation (\ref{eq:output}b).
-The parameters used to establish this relationship (Equation \ref{eq:output}b), $\mu$, may or may not include system parameters $p$ defined in Equation (\ref{eq:output}a).
+The parameters used to establish this relationship (Equation \ref{eq:output}b), $\mu$, may or may not include system parameters $\theta$ defined in Equation (\ref{eq:output}a).
 
-Typically, a nonlinear programming formulation with the objective of minimizing the least square error between the measured ($y_{kl}^*$) and modeled ($y_{kl}$) outputs over the time course ($l = {1, ..., d}$) for which experimental data is collected (Equation \ref{eq:minnlp}) (Raue et al., and other parameter estimation/identifiability papers), is used to estimate parameters $p$ and $\mu$ in Equation \ref{eq:output}.
+Typically, a nonlinear programming formulation with the objective of minimizing the least square error between the measured ($y_{kl}^*$) and modeled ($y_{kl}$) outputs over the time course ($l = {1, ..., d}$) for which experimental data is collected (Equation \ref{eq:minnlp}) (Raue et al., and other parameter estimation/identifiability papers), is used to estimate parameters $\theta$ and $\mu$ in Equation \ref{eq:output}.
 \begin{align}\label{eq:minnlp}
-\chi^2(\theta) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
+\chi^2(\theta,\mu) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
 \end{align}
@@
 The dynamics of metabolic networks can also be represented by Equation (\ref{eq:output}) wherein the metabolite concentrations, $\mathbf{x}$, are state variables.
-The changes in the metabolite concentrations ($\mathbf{x}$) are expressed as a function of reaction rates dependent on enzyme kinetics, which are in turn expressed as nonlinear functions of the states ($\mathbf{x}$), kinetic parameters ($p$) and other input variables ($u$).
+The changes in the metabolite concentrations ($\mathbf{x}$) are expressed as a function of reaction rates dependent on enzyme kinetics, which are in turn expressed as nonlinear functions of the states ($\mathbf{x}$), kinetic parameters ($\theta$) and other input variables ($u$).
 
@@
 
-	\subsection{Identifiability analysis: Definitions and Formulations}
-	Any nonlinear dynamical system can be represented by a set of states $\mathbf{x}$, observables $\mathbf{y}$ that are dependent on the states, parameters $\mathbf{\mu}$, and inputs $u$ as in Equation (\ref{eq:dynmodel}).
-	\begin{subequations}\label{eq:dynmodel}
-		\begin{align}
-		\dot{\mathbf{x}} = g(\mathbf{x},\mathbf{\mu},u)\\
-		\mathbf{y} = h(\mathbf{x},\mathbf{\mu},u)
-		\end{align}
-	\end{subequations}
-	Identifiability concerns with the ability to determine a unique solution to the problem of estimating parameters $\mathbf{\mu}$ from given data on the system observables $\mathbf{y}$ for inputs $u$ \parencite{McLean2012}.	The identifiability of parameters in nonlinear models of physical processes can be classified into two categories: structural and practical identifiability.
-
-	%Thus, the identifiability of parameters in nonlinear models is dependent on the model structure, parameterization, and the quality and quantity of experimental data that is available for the purpose of estimation.
-
 \section{Methods:}
-	\subsection{A method to establish posterior identifiability of metabolic network models:}
-	This section details a method to establish the practical (posterior) identifiability of metabolic network models using the algebraic relationship between fluxes.
-	Every flux, $v$, in a kinetic model of a metabolic network can be expressed as a nonlinear algebraic equation (Equation \ref{eq:flux_gen}).
-The fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ in Equation (\ref{eq:flux_gen}).
-	\begin{align}\label{eq:flux_gen}
-	v = f(\mathbf{x},\mathbf{\theta})
-	\end{align}
-	Given the nonlinear nature of this model, the function $f$ in Equation (\ref{eq:flux_gen}) can expressed, without loss of generality as,
-	\begin{align}\label{eq:nr_dr}
-	v = \frac{N(\mathbf{x},\theta)}{D(\mathbf{x},\theta)}
-	\end{align}
-	where $N(\mathbf{x},\theta)$ is the numerator of $f$, and $D(\mathbf{x},\theta)$ is the denominator of $f$.
-
-	If $\theta \in \mathbb{R}^p$, given a set of experimental measurements for the metabolite concentrations $\mathbf{x}$ and the reaction fluxes $\mathbf{v}$, theoretically, it is possible to choose $p$ sets of data from these measurements to solve for the $p$ parameters in $\theta$.
-However, if any these datasets do not satisfy the condition that $D(\mathbf{x},\theta) \neq 0$, then the number of experiments required to estimate the $p$ parameters in $\theta$ can be established to be greater than $p$.
-An example is shown below.
-
-	This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
-This enables this method to be scalable to even genome-scale models.
-The following section demonstrates this methodology for one of the fluxes in the gluconeogenic model of Kotte et al., \parencite{Kotte2014}.
-
-	\subsection{Identifiability analysis of parameters in a kinetic model of gluconeogenesis:}
+\subsection{Kinetic model of gluconeogenesis in E. coli:}
 The proposed model for acetate consumption through gluconeogenesis and its corresponding kinetic model is used as a case study to illustrate the utility of identifiability analysis for the design of experiments for estimating parameters in kinetic models of metabolism.
@@
 
+\subsection{A method for practical identifiability of kinetic models of metabolism:}
+In this section, we show how practical identifiability of kinetic parameters in a dynamic model of metabolism can be established using the algebraic relationship between fluxes that form the network.
+In a kinetic model, the value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations.
+Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations (Equation \ref{eq:flux_gen}).
+The fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$.
+\begin{align}\label{eq:flux_gen}
+v = f(\mathbf{x},\mathbf{\theta})=\frac{N(\mathbf{x},\theta)}{D(\mathbf{x},\theta)}
+\end{align}
+In Equation \ref{eq:flux_gen}, $N(\mathbf{x},\theta)$ is the numerator of $f$, and $D(\mathbf{x},\theta)$ is the denominator of $f$.
+
+Let $\theta \in \mathbb{R}^p$ for each flux $v_i$ in the network.
+Also, let $E\in \mathbb{R}^n$ (and $j = {1, 2, ..., n}$) be the set of all possible experiments from which measurements of metabolite concentrations $\mathbf{x}_j$ and reaction fluxes $\mathbf{v}_j$ are available.
+For a given flux $v_i$, if all parameters $\theta \in \mathbb{R}^p$ are identifiable, then
+
+if all parameters are identifiable, theoretically, the $P$ unknown parameters in $\theta$ can be estimated by solving $P\subset{1, 2, ..., n}$ nonlinear algebraic equations shown in Equation \ref{eq:nonlineq}.
+\begin{align}\label{eq:nonlineq}
+v_k = \frac{N(\mathbf{x}_k,\theta)}{D(\mathbf{x}_k,\theta)} && \forall k\in{1, 2, ..., k, ..., P}
+\end{align}
+Here, $K = {1, 2, ..., k, ..., P}$ is the index set of $P$ experiments for which metabolite concentration ($\mathbf{x}_K$) and flux ($\mathbf{v}_K$) measurements are available for each metabolite and flux within the metabolic network.
+Parameter $\theta_p\in\theta$ is said to be identifiable if
+
+However, if for any index $l$, $\mathbf{x}_l$ and $\mathbf{v}_l$ do not satisfy the condition that $D(\mathbf{x}_l,\theta) \neq 0$, i.e., at least one flux $v_i$ is not identifiable based on experiment $l$, then the $l$th experiment cannot be used for parameter estimation and should be discarded.
+Consequently, an additional experiment $m \in {1, 2, ..., n}, m\neq l$ is needed to estimate all $p$ parameters in $\theta$.
+This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
+This enables this method to be scalable to even genome-scale models.
+The following section demonstrates this methodology for one of the fluxes in the gluconeogenic model described earlier.
+
+\subsection{Experimental design for parameter estimation in kinetic models of metabolism:}
+The identifiability of each parameter based on each experiment indexed as $j = {1, ..., n}$
+
+\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis:}
+
+\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis:}
+Steady state metabolomics and fluxomics data can be gathered under different physiological conditions by either perturbing the expression levels for different enzymes within a metabolic network, or by changing the substrate concentrations under which the cells grow.
+The aforementioned model of gluconeogenesis has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($V_1^{max}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
+Furthermore, the acetate concentration on which the cell is grown can also be perturbed to measure cellular response to changes in the substrate concentration.
+We use the in silico metabolomics and fluxomics data generated from these perturbation experiments to demonstrate parameter identification with our methodology.
+
+If any data generated from a perturbation experiment $i$ results in nonidentifiability, we eliminate experiment $i$ from the list of experiments that need to be performed for parameter estimation.
+
 We use flux $v_2$ to demonstrate the identifiability analysis method described in the previous section.
@@
 Subsequently, these experimental datasets can be included in the model to form two simultaneous nonlinear algebraic equations in the parameters $V_2^{max}$ and $K_2^{pep}$ (Equation \ref{eq:v2_par}).
-
 \begin{subequations}\label{eq:v2_par}
@@
 \end{table}
+
+\section{Results:}
+\subsection{Getting closed-form expressions for each flux in Kotte model:}
+\subsection{Experiments that can establish identifiability for each flux in Kotte model:}	
+\subsection{Combinations of experiments that will enable identification of all model parameters:}
+\subsection{Expanding methodology to RBC model:}
+
+\paragraph{Outline:}
+\begin{itemize}
+	\item parameter estimation is a well developed field typically using minimization of least square error to estimate model parameters from available experimental data
+	\item if parameters are structurally identifiable, it does not guarantee practical identifiability from noisy experimental data
+	\item identifiability dependent on whether given datasets (outputs) for estimation can sufficiently distinguish between different parameter values
+\end{itemize}
+
+\paragraph{Sections:}
+\begin{itemize}
+	\item datasets required for parameter estimation in kinetic models of metabolism (methods?)		
+	\item identifiability in kotte model - scalability, number of experiments required, requirements for time course data(? in the intro)		
+	\item identifiability in large rbc model
+\end{itemize}
+
+\printbibliography
+\end{document}
+
+\subsection{Identifiability analysis: Definitions and Formulations}
+Any nonlinear dynamical system can be represented by a set of states $\mathbf{x}$, observables $\mathbf{y}$ that are dependent on the states, parameters $\mathbf{\mu}$, and inputs $u$ as in Equation (\ref{eq:dynmodel}).
+\begin{subequations}\label{eq:dynmodel}
+	\begin{align}
+	\dot{\mathbf{x}} = g(\mathbf{x},\mathbf{\mu},u)\\
+	\mathbf{y} = h(\mathbf{x},\mathbf{\mu},u)
+	\end{align}
+\end{subequations}
+Identifiability concerns with the ability to determine a unique solution to the problem of estimating parameters $\mathbf{\mu}$ from given data on the system observables $\mathbf{y}$ for inputs $u$ \parencite{McLean2012}.	The identifiability of parameters in nonlinear models of physical processes can be classified into two categories: structural and practical identifiability.
+
+Given the nonlinear nature of this model, the function $f$ in Equation (\ref{eq:flux_gen}) can expressed, without loss of generality as,
+\begin{align}\label{eq:nr_dr}
+v = \frac{N(\mathbf{x},\theta)}{D(\mathbf{x},\theta)}
+\end{align}
 
@@
 
-	\section{Results:}
-	\paragraph{Outline:}
-	\begin{itemize}
-		\item parameter estimation is a well developed field typically using minimization of least square error to estimate model parameters from available experimental data
-		\item if parameters are structurally identifiable, it does not guarantee practical identifiability from noisy experimental data
-		\item identifiability dependent on whether given datasets (outputs) for estimation can sufficiently distinguish between different parameter values
-	\end{itemize}
-
-	\paragraph{Sections:}
-	\begin{itemize}
-		\item datasets required for parameter estimation in kinetic models of metabolism (methods?)		
-		\item identifiability in kotte model - scalability, number of experiments required, requirements for time course data(? in the intro)		
-		\item identifiability in large rbc model
-	\end{itemize}
-
-	\printbibliography
-\end{document}
+%Thus, the identifiability of parameters in nonlinear models is dependent on the model structure, parameterization, and the quality and quantity of experimental data that is available for the purpose of estimation.

### 2018-01-02 ident/python2/ss-ident/noise_script.py
no major changes

@@ -31,7 +31,7 @@ print('Perturbation analysis for identifiability complete.\n')
 data_list, original_data_ident, combo_data_ident, max_parameter = process_info_sample(ident_details,
                                                                                       experimental_datasets,
                                                                                       perturbation_details)
-# get data for plots
+# get data for plots 
 total_ident_data, fraction_ident_data, all_boolean_p_id = parameter_plot_data_per_sample(original_data_ident, 1)
 # plot
 flux_parameter_plot(total_ident_data, fraction_ident_data)

### 2018-01-02 ident/docs/ident_manuscript.tex

@@
 \end{subequations}
-When not all states are observable, Equation (\ref{eq:output}a) is usually augmented with a system that defines the relationship between the observable output variables ($\mathbf{y}$) and the state variables ($\mathbf{x}$) as in Equation (\ref{eq:output}b).
+When all states are not observable, Equation (\ref{eq:output}a) is usually augmented with a system that defines the relationship between the observable output variables ($\mathbf{y}$) and the state variables ($\mathbf{x}$) as in Equation (\ref{eq:output}b).
 The parameters used to establish this relationship (Equation \ref{eq:output}b), $\mu$, may or may not include system parameters $\theta$ defined in Equation (\ref{eq:output}a).
@@
 \section{Methods:}	
-\subsection{Kinetic model of gluconeogenesis in E. coli:}
+\subsection{A method for practical identifiability of kinetic models of metabolism:}\label{sec:ident}
+In this section, we show how practical identifiability of kinetic parameters in a dynamic model of metabolism can be established using the nonlinear algebraic kinetic rate law formulations that describe the relationship between the concentrations, fluxes and the kinetic parameters of the metabolic network model.
+A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
+In a kinetic model, the value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations.
+Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations (Equation \ref{eq:flux_gen}).
+The fluxes are expressed as a function of the metabolite concentrations $\mathbf{x}$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
+\begin{align}\label{eq:flux_gen}
+v_i = f(\mathbf{x},\mathbf{\theta}, u)
+\end{align}
+Let $\theta \in \mathbb{R}^p$ for each flux $v_i$ in the network.
+For each experiment $j = {1, 2, ..., n}$, we assume that all metabolite concentrations $\mathbf{x}$ and reaction fluxes $\mathbf{v}$ are measurable.
+The pertinent information for each experiment is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
+
+In order to establish the identifiability of kinetic parameters for each flux $v_i$, we describe a computer algebra-based method.
+The primary use of the computer algebra system is to obtain closed form expressions for each parameter in $\theta$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}c).
+This is done by solving a system of nonlinear algebraic equations in $\mathbb{R}^p$, shown in Equation (\ref{eq:nonlineq}).
+\begin{align}\label{eq:nonlineq}
+v_{i, k} = f_k(\mathbf{x}_k,\theta, u_k) && \forall k=\{1, 2, ..., p\}\subset\{1, 2, ..., n\}
+\end{align}
+Each equation in (\ref{eq:nonlineq}), indicated by the index $k$, corresponds to the kinetic rate law expression $f(\mathbf{x},\mathbf{\theta}, u)$ for $v_i$, described earlier in Equation (\ref{eq:flux_gen}), written for concentrations and fluxes obtained from experiment $k$.
+Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$, where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}c).
+\begin{align}\label{eq:theta-eq}
+\theta_k = g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
+\end{align}
+The identifiability of parameter $\theta_k$ for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}c): any parameter $\theta_k$ is said to practically identifiable (practically non-identifiable) if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ ($D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$).
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=.8\textwidth,height=.8\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figures/ident_flow_chart}}
+	\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
+a) The steps for the construction of a kinetic model of a metabolic network are shown.
+The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
+The identifiability of parameters for each flux can be establish independently.
+b) in silico data generation to test the identifiability of a given parameter is an important step that is also required for experimental design using the proposed methodology.
+c) The steps for identifiability analysis for a parameters of a single flux are shown.
+d) If any of the proposed in silico experiments results in non-identifiability of any parameter, then the number and nature of experiments to be performed to establish identifiability of the corresponding parameter can be determined by following the steps shown.}\label{fig:ident-flowchart}
+\end{figure}
+
+\subsection{Experimental design for parameter estimation in kinetic models of metabolism:}
+The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in Section \ref{sec:ident} and demonstrated in Section \ref{sec:example}.
+Subsequently, for any flux $v_i$, if for any $p$ combinations of indices $j$, the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make a parameter $\theta_k$ identifiable.
+Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
+Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
+This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
+In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
+This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
+Hence, our method is theoretically scalable even to genome-scale models.
+We show the application of this design methodology for flux $v_2$ in the gluconeogenic model in Section.
+
+Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability using our methodology.
+
+In the following sections we provide a previously published kinetic model of a small gluconeogenic network, followed by a demonstration of our methodology to establish practical identifiability and experimental design for one of the fluxes in this network.
+
+\subsection{Kinetic model of gluconeogenesis in E. coli:}\label{sec:small-model}
 The proposed model for acetate consumption through gluconeogenesis and its corresponding kinetic model is used as a case study to illustrate the utility of identifiability analysis for the design of experiments for estimating parameters in kinetic models of metabolism.
@@
 
-\subsection{A method for practical identifiability of kinetic models of metabolism:}
-In this section, we show how practical identifiability of kinetic parameters in a dynamic model of metabolism can be established using the algebraic relationship between fluxes that form the network.
-In a kinetic model, the value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations.
-Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations (Equation \ref{eq:flux_gen}).
-The fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$.
-\begin{align}\label{eq:flux_gen}
-v = f(\mathbf{x},\mathbf{\theta})=\frac{N(\mathbf{x},\theta)}{D(\mathbf{x},\theta)}
-\end{align}
-In Equation \ref{eq:flux_gen}, $N(\mathbf{x},\theta)$ is the numerator of $f$, and $D(\mathbf{x},\theta)$ is the denominator of $f$.
-
-Let $\theta \in \mathbb{R}^p$ for each flux $v_i$ in the network.
-Also, let $E\in \mathbb{R}^n$ (and $j = {1, 2, ..., n}$) be the set of all possible experiments from which measurements of metabolite concentrations $\mathbf{x}_j$ and reaction fluxes $\mathbf{v}_j$ are available.
-For a given flux $v_i$, if all parameters $\theta \in \mathbb{R}^p$ are identifiable, then
-
-if all parameters are identifiable, theoretically, the $P$ unknown parameters in $\theta$ can be estimated by solving $P\subset{1, 2, ..., n}$ nonlinear algebraic equations shown in Equation \ref{eq:nonlineq}.
-\begin{align}\label{eq:nonlineq}
-v_k = \frac{N(\mathbf{x}_k,\theta)}{D(\mathbf{x}_k,\theta)} && \forall k\in{1, 2, ..., k, ..., P}
-\end{align}
-Here, $K = {1, 2, ..., k, ..., P}$ is the index set of $P$ experiments for which metabolite concentration ($\mathbf{x}_K$) and flux ($\mathbf{v}_K$) measurements are available for each metabolite and flux within the metabolic network.
-Parameter $\theta_p\in\theta$ is said to be identifiable if
-
-However, if for any index $l$, $\mathbf{x}_l$ and $\mathbf{v}_l$ do not satisfy the condition that $D(\mathbf{x}_l,\theta) \neq 0$, i.e., at least one flux $v_i$ is not identifiable based on experiment $l$, then the $l$th experiment cannot be used for parameter estimation and should be discarded.
-Consequently, an additional experiment $m \in {1, 2, ..., n}, m\neq l$ is needed to estimate all $p$ parameters in $\theta$.
-This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
-This enables this method to be scalable to even genome-scale models.
-The following section demonstrates this methodology for one of the fluxes in the gluconeogenic model described earlier.
-
-\subsection{Experimental design for parameter estimation in kinetic models of metabolism:}
-The identifiability of each parameter based on each experiment indexed as $j = {1, ..., n}$
-
-\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis:}
-
 \subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis:}
@@
 
-If any data generated from a perturbation experiment $i$ results in nonidentifiability, we eliminate experiment $i$ from the list of experiments that need to be performed for parameter estimation.
+\section{Results:}
+First, in Section \ref{sec:example} we demonstrate how our methodology can be applied for one of the fluxes in the gluconeogenic model.
+In Section \ref{sec:closedform} that follows, we discuss the ability to apply our methods to different nonlinear kinetic rate law formulations within the context of the gluconeogenic model.
+Then, in Section we discuss the different experimental design strategies that were found through our identifiability analysis to enable kinetic parameter estimation for the each flux in the model using steady state data.
+Finally, in Section we expand the application of our method to a relatively large (39 reactions) kinetic model of the red blood cell metabolism.
 
-We use flux $v_2$ to demonstrate the identifiability analysis method described in the previous section.
+\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis:}\label{sec:example}
+Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_2$ in the small model of gluconeogenesis described in Section \ref{sec:small-model}.
 Flux $v_2$ has two parameters, $V_2^{max}$ and $K_2^{pep}$ that need to be estimated from experimental data.
 Here, we assume that at least two different sets of experimental data for the concentrations and fluxes are available.
-Accordingly, we label these dataset as ${pep}^1$, ${v_2}^1$ and ${pep}^2$, ${v_2}^2$ respectively.
-Subsequently, these experimental datasets can be included in the model to form two simultaneous nonlinear algebraic equations in the parameters $V_2^{max}$ and $K_2^{pep}$ (Equation \ref{eq:v2_par}).
+We label the concentrations as ${pep}^1$ and ${pep}^2$, and the fluxes as ${v_2}^1$ and ${v_2}^2$ respectively, from each experiment.
+Accordingly, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for flux $v_2$ as follows:
+\begin{subequations}\label{eq:nonlin-flux2}
+	\begin{align}
+	{v_2}^1 = V_{2}^{max}\frac{pep^1}{pep^1+K_{2}^{pep}}\\
+	{v_2}^2 = V_{2}^{max}\frac{pep^2}{pep^2+K_{2}^{pep}}
+	\end{align}
+\end{subequations}
+Solving this simultaneous system of equations, we get two nonlinear algebraic equations in the parameters $V_2^{max}$ and $K_2^{pep}$ based on the form shown earlier in Equation (\ref{eq:theta-eq}).
 \begin{subequations}\label{eq:v2_par}
@@
 \end{subequations}
+In Equation (\ref{eq:v2_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_2^{max}$ (Equation \ref{eq:v2_par}a) and $K_2^{pep}$ (Equation \ref{eq:v2_par}b).
+
+\subsection{Getting closed-form expressions for each flux in Kotte model:}\label{sec:closedform}
+The complexity of the equations in our specific scenario is determined by the complexity of enzyme-metabolite interaction models used describe fluxes in metabolic networks.
+Although computer algebra systems (CAS) are capable of handling complex symbolic calculations, sometimes, the complexity of getting closed form expressions for all kinetic parameters of certain rate law formulations is too much for the CAS to overcome.
+We encountered this scenario in the case of the gluconeogenic model for flux $v_3$ where the kinetics of the allosterically activated reaction are described by the MWC model for allosteric regulation.
+In order to overcome this computational difficulty, we used a convenience kinetic rate law formulation to describe the allosteric interaction in $v_3$.
+We give this formulation in Section.
+
+We believe this problem of computational tractability to occur in other complex kinetic rate law formulations as well. Examples?
+
+\subsection{Experiments that can establish identifiability for each flux in Kotte model:}	
+As indicated previously in Section, we performed different experiments by perturbing the parameters for fluxes $v_1$, $v_2$ and $v_3$ in the model.
+We also perturbed the input substrate concentration to collect data on the concentrations and fluxes within the network.
+We describe the results of using this data to establish the identifiability of various parameters in the kinetic model of the small metabolic network.
+
+\subsection{Combinations of experiments that will enable identification of all model parameters:}
+\subsection{Expanding methodology to RBC model:}
+
+If any data generated from a perturbation experiment $i$ results in nonidentifiability, we eliminate experiment $i$ from the list of experiments that need to be performed for parameter estimation.
 
@@
 \end{table}
-
-\section{Results:}
-\subsection{Getting closed-form expressions for each flux in Kotte model:}
-\subsection{Experiments that can establish identifiability for each flux in Kotte model:}	
-\subsection{Combinations of experiments that will enable identification of all model parameters:}
-\subsection{Expanding methodology to RBC model:}
 

### 2018-01-05 ident/docs/ident_manuscript.tex
changes to intro - follow document outline - simplified and focussed on metabolic network models

@@
 \begin{document}
+\paragraph{Outline:}
+\begin{itemize}
+	\item Importance of parameter values for estimating in vivo response using kinetic models of metabolism and consequently for metabolic network design using kinetic models of metabolism
+	\item The need for parameter identifiability to determine unique and true parameter values from observed data
+	\item Types and purpose of identifiability for parameters
+	\item Methods for structural identifiability and existing methods for practical identifiability
+	\item Lack of methods for practical identifiability and consequently experimental design (not covered?)
+	\item work done in this paper for practical identifiability			
+	\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable) (move to discussion - may be in discussion)
+\end{itemize}	
 \section{Introduction:}
+Kinetic models of metabolism can be used to study the dynamic characteristics of metabolic networks.
+In these models, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
+The matrix $\mathbf{S}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations in the metabolic network.
+\begin{subequations}\label{eq:kinstoich}
+	\begin{align}
+	\dot{x} = \mathbf{S}v\\
+	v = g(x, \theta, u)
+	\end{align}
+\end{subequations}
+The expression for the nonlinear function ($g$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction.
+Accordingly, the reaction fluxes are expressed as a nonlinear function ($g$) of the metabolite concentrations, enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
+
+The ability to predict the steady state and dynamic responses of the metabolic network, under in vivo conditions, to different perturbations is dependent on the numerical values of the enzyme kinetic parameter values ($\theta$) in Equation (\ref{eq:kinstoich}).
+The use of in vitro, or unreliable in vivo parameter estimates, reduces the confidence in the model predicted behaviour.
+The reduction in confidence not only hampers the use of these models to gain insight into the functioning of the metabolic network, but also is an obstacle for using the predicted responses as a basis for designing metabolic networks.
+
+If all intracellular metabolite concentrations can be measured over a time course, and the corresponding data is available, a nonlinear programming problem can be formulated to estimate the enzyme kinetic parameters ($\theta$) in Equation (\ref{eq:kinstoich}).
+The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations can be used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}), with fixed upper ($\theta_u$) and lower ($\theta_l$) bounds for the parameter values (Equation \ref{eq:lsqopt}b).
+\begin{subequations}\label{eq:lsqopt}
+	\begin{align}
+	\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{x_{kl}^*-x_{kl}}{\sigma_{kl}^*}\right)^2\\
+	&\theta_l \le \theta \le \theta_u
+	\end{align}
+\end{subequations}
+
+However, not all metabolite concentrations used in the model in Equation (\ref{eq:kinstoich}) can be measured.
+Additionally, measurable fluxes in the metabolic network also need to be included as part of the parameter estimation problem.
+In such scenarios, the parameter estimation problem is modified to suit a new system of equations shown below (Equation \ref{eq:output}).
+The new system of equations is obtained by augmenting the original system in Equation (\ref{eq:kinstoich}) with Equation (\ref{eq:output}c) that models the relationship between the measurable metabolite concentrations ($y$) and fluxes  and the unmeasurable concentrations ($x$) that are used in the original model in Equation (\ref{eq:kinstoich}) above.
+In Equation (\ref{eq:output}), the measurable concentrations
+\begin{subequations}\label{eq:output}
+	\begin{align}
+	\dot{x} = \mathbf{S}v\\
+	v = g(x, \theta, u)\\
+	\dot{y} = h(x, \mu, u)
+	\end{align}
+\end{subequations}
+The metabolite concentrations ($x$) in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the measurable concentrations $y$ to estimate $\theta$.
+
+\section{old Introduction:}	
 The dynamic response of any system can be modeled as a set of ordinary differential equations (ode) (Equation \ref{eq:output}a) in the state variables ($\mathbf{x}$).
 The changes in the state variables are expressed as a function of the state variables, the system parameters ($\theta$) and the inputs to the system ($u$).
-\begin{subequations}\label{eq:output}
-	\begin{align}
-	\dot{\mathbf{x}} = g(\mathbf{x}, \theta, u)\\
-	\mathbf{y} = h(\mathbf{x},\mu,u)
-	\end{align}
-\end{subequations}
+
 When all states are not observable, Equation (\ref{eq:output}a) is usually augmented with a system that defines the relationship between the observable output variables ($\mathbf{y}$) and the state variables ($\mathbf{x}$) as in Equation (\ref{eq:output}b).
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=.8\textwidth,height=.8\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figures/ident_flow_chart}}
+	\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure3/ident_analysis}}
 	\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.

### 2018-01-05 ident/docs/ident_manuscript.tex
complete intro with outline

--- 2018-01-02 ident/docs/ident_manuscript.tex
@@
 \begin{document}
-\section{Introduction:}	
-The dynamic response of any system can be modeled as a set of ordinary differential equations (ode) (Equation \ref{eq:output}a) in the state variables ($\mathbf{x}$).
-The changes in the state variables are expressed as a function of the state variables, the system parameters ($\theta$) and the inputs to the system ($u$).
-\begin{subequations}\label{eq:output}
-	\begin{align}
-	\dot{\mathbf{x}} = g(\mathbf{x}, \theta, u)\\
-	\mathbf{y} = h(\mathbf{x},\mu,u)
-	\end{align}
-\end{subequations}
-When all states are not observable, Equation (\ref{eq:output}a) is usually augmented with a system that defines the relationship between the observable output variables ($\mathbf{y}$) and the state variables ($\mathbf{x}$) as in Equation (\ref{eq:output}b).
-The parameters used to establish this relationship (Equation \ref{eq:output}b), $\mu$, may or may not include system parameters $\theta$ defined in Equation (\ref{eq:output}a).
-
-Typically, a nonlinear programming formulation with the objective of minimizing the least square error between the measured ($y_{kl}^*$) and modeled ($y_{kl}$) outputs over the time course ($l = {1, ..., d}$) for which experimental data is collected (Equation \ref{eq:minnlp}) (Raue et al., and other parameter estimation/identifiability papers), is used to estimate parameters $\theta$ and $\mu$ in Equation \ref{eq:output}.
-\begin{align}\label{eq:minnlp}
-\chi^2(\theta,\mu) = \sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2
-\end{align}
-The difference between the data and the model estimate for each output, and at each time point, is weighted by the variance in the experimental data $\sigma_{kl}^*$ for that corresponding variable and time point.
-The minimization problem is solved subject to constraints on the state and output variables expressed in Equation \ref{eq:output}, as well as bounds on the estimated parameter values.
-The ability to determine a unique solution to the parameter estimation problem is however governed by the identifiability of the parameters in the model \parencite{McLean2012}.
-
-The identifiability of parameters in nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability. 	
-Any system (Equation \ref{eq:output}) is said to be structurally identifiable if, for an input-output mapping defined by $\mathbf{y} = \Phi(\mu,u)$ for at least one input function $u$, any two values of parameters $\mathbf{\mu}_1$ and $\mathbf{\mu}_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
-\begin{align}\label{eq:stident}
-\Phi(\mathbf{\mu}_1,u) = \Phi(\mathbf{\mu}_2,u) \iff \mathbf{\mu}_1 = \mathbf{\mu}_2
-\end{align}
-Accordingly, any system that has an infinite number of solutions to the parameter estimation problem for all input functions is said to structurally non-identifiable.
-Thus, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unobservable system states and the observable system outputs.
-Accordingly, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
-
-Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
-The effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
-Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
-
-Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated.
-On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
-
-The dynamics of metabolic networks can also be represented by Equation (\ref{eq:output}) wherein the metabolite concentrations, $\mathbf{x}$, are state variables.
-The changes in the metabolite concentrations ($\mathbf{x}$) are expressed as a function of reaction rates dependent on enzyme kinetics, which are in turn expressed as nonlinear functions of the states ($\mathbf{x}$), kinetic parameters ($\theta$) and other input variables ($u$).
-
-Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems (IEEE Trans paper from 2007).
-Most of these algorithms use methods based on differential algebra (Glad and Ljung, 1994).
-More recent methods take a profile-likelihood-based approach (2012 Paper) to establish both structural and practical identifiability.
-However, these methods not only scale poorly with increases in size of the modeled system, but also require dynamic time course data of the observable variables of the system.
-While computational burden due to poor scalability can be partly addressed with the current increases in computational power, the ability to obtain dynamic data for establishing identifiability of parameters in kinetic models of metabolism still remains a challenge.
-
-In this paper, we propose a methodology to establish practical identifiability for parameters in kinetic models of metabolism.
-We present a computer algebra-based method that can facilitate experimental design for estimating parameters separately for each individual reaction within a metabolic network based on available steady state experimental data.
-This enables us to address the twin issues of scalability and data availability.
-We illustrate the utility of this method by applying it for a small network of gluconeogenesis in \textit{E.
-coli} and demonstrating our ability to propose experiments that will facilitate parameter estimation for a kinetic model of this network.
-We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
-
 \paragraph{Outline:}
@@
 \end{itemize}
+\section{Introduction:}
+Kinetic models of metabolism can be used to study the dynamic characteristics of metabolic networks.
+In these models, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
+The matrix $\mathbf{S}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
+\begin{subequations}\label{eq:kinstoich}
+	\begin{align}
+	\dot{x} = \mathbf{S}v\\
+	v = g(x, \theta, u)
+	\end{align}
+\end{subequations}
+The expression for the nonlinear function ($g$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction (older kinetic modeling papers, reviews, Srinivasan et al., 2015).
+Accordingly, $g$ is a nonlinear function of the metabolite concentrations, enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
+
+The ability to predict the steady state and dynamic responses of metabolic networks, under in vivo conditions, to different perturbations is dependent on the numerical values of the enzyme kinetic parameter values ($\theta$) in Equation (\ref{eq:kinstoich}).
+Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where parameter values might play a crucial role (Srinivasan et al., 2017?).
+The use of in vitro, or unreliable in vivo parameter estimates, reduces confidence in the model predicted behaviour.
+Consequently, the reduction in confidence hampers the use of these models to gain insight into the functioning of metabolic networks (ensemble modeling and ORACLE papers).
+The insights gained from the use of kinetic models are subsequently used to design changes to these metabolic networks to achieve various goals.
+These goals could either be to increase metabolite production for biosynthesis of different chemicals (kinetic model for BDO in ecoli, k-ecoli457, Srinivasan et al., 2017) or to find therapeutic targets that can cut-off metabolic pathways to kill cancer cells (new paper on using geneMCS for cancer cell lines).
+Hence, an increase in uncertainty in model predicted responses is also an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve these goals.
+
+If all intracellular metabolite concentrations can be measured over a time course, and the corresponding data is available, a nonlinear programming problem can be formulated to estimate the enzyme kinetic parameters ($\theta$) in Equation (\ref{eq:kinstoich}).
+The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations can be used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}), with fixed upper ($\theta_u$) and lower ($\theta_l$) bounds for the parameter values (Equation \ref{eq:lsqopt}b).
+The difference between the data and the model estimate for each output and at each time point in the objective function is weighted by the variance in the experimental data $\sigma_{kl}^*$ for that corresponding variable and time point.
+\begin{subequations}\label{eq:lsqopt}
+	\begin{align}
+	\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{x_{kl}^*-x_{kl}}{\sigma_{kl}^*}\right)^2\\
+	&\theta_l \le \theta \le \theta_u
+	\end{align}
+\end{subequations}
+
+However, not all metabolite concentrations used in the model (Equation \ref{eq:kinstoich}) can be measured.
+Additionally, measurable fluxes in the metabolic network also need to be included as part of the parameter estimation problem.
+In such scenarios, the parameter estimation problem is modified to suit a new system of equations shown below (Equation \ref{eq:output}).
+The new system of equations is obtained by augmenting the original system (Equation \ref{eq:kinstoich}) with Equation (\ref{eq:output}c) that models the relationship between the measurable metabolite concentrations and fluxes, represented by the variable $y$, and the unmeasured concentrations ($x$) that are used in the original model (Equation \ref{eq:kinstoich}) above.
+Equation (\ref{eq:output}) now has additional parameters $\mu$ that define this relationship and also need to be estimated.
+\begin{subequations}\label{eq:output}
+	\begin{align}
+	\dot{x} = \mathbf{S}v\\
+	v = g(x, y, \theta, u)\\
+	\dot{y} = h(x, y, \mu, u)
+	\end{align}
+\end{subequations}
+In systems identification, the measured concentrations and fluxes, $y$ in Equation (\ref{eq:output}c), are called output or observed variables, and the unmeasured concentrations, $x$ in Equation (\ref{eq:output}a-b), are called the state variables.
+For estimating both $\theta$ and $\mu$, the metabolite concentrations ($x$) in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the output variables $y$.
+
+However, the ability to determine unique solutions to parameters $\theta$ and $\mu$ is governed by the identifiability of these parameters in the model \parencite{McLean2012}.
+The identifiability of parameters in nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability.
+Any system (Equation \ref{eq:output}) is said to be structurally identifiable if, for an input-output mapping defined by $\mathbf{y} = \Phi(\mu,u)$ for at least one input function $u$, any two values of parameters $\mathbf{\mu}_1$ and $\mathbf{\mu}_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
+\begin{align}\label{eq:stident}
+\Phi(\mathbf{\mu}_1,u) = \Phi(\mathbf{\mu}_2,u) \iff \mathbf{\mu}_1 = \mathbf{\mu}_2
+\end{align}
+Accordingly, any system that has an infinite number of solutions to the parameter estimation problem for all input functions is said to structurally non-identifiable.
+Thus, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unobservable state variables and the observable output variables.
+Accordingly, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
+
+Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
+The effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
+Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
+
+Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated.
+On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
+
+Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems (IEEE Trans paper from 2007).
+Most of these algorithms use methods based on differential algebra (Glad and Ljung, 1994).
+More recent methods take a profile-likelihood-based approach (2012 Paper) to establish both structural and practical identifiability.
+However, these methods scale poorly with increase in size of the modeled system.
+The aforementioned methods are also heavily dependent on the availability of dynamic time course data for all the output variables of the system.
+
+Computational burden due to poor scalability can be partly addressed with the current increases in computational power, while current improvements in measuring technologies can enable one to obtain high frequency dynamic data, no scalable methodologies exist to utilize the abundantly available steady state data for parameter estimation and identifiability for large metabolic networks.
+
+In this paper, we propose a methodology to establish practical identifiability for parameters in kinetic models of metabolism.
+We present a computer algebra-based method that can facilitate practical identifiability as well as experimental design for estimating parameters separately for each individual reaction within a metabolic network based on available steady state experimental data.
+This enables us to address the twin issues of scalability and data availability.
+We illustrate the utility of this method by applying it for a small network of gluconeogenesis in \textit{E.
+coli} and demonstrating our ability to propose experiments that will facilitate parameter estimation for a kinetic model of this network.
+We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
 

### 2018-01-09 ident/docs/ident_manuscript.tex
additions to refs and intro section

@@
 \end{subequations}
-The expression for the nonlinear function ($g$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction (older kinetic modeling papers, reviews, Srinivasan et al., 2015).
+The expression for the nonlinear function ($g$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Heijnen2005, Link2014, Machado2011, Srinivasan2015}.
 Accordingly, $g$ is a nonlinear function of the metabolite concentrations, enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
@@
 The ability to predict the steady state and dynamic responses of metabolic networks, under in vivo conditions, to different perturbations is dependent on the numerical values of the enzyme kinetic parameter values ($\theta$) in Equation (\ref{eq:kinstoich}).
-Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where parameter values might play a crucial role (Srinivasan et al., 2017?).
+Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where parameter values might play a crucial role \parencite{Srinivasan2015, Vital-Lopez2006}(Srinivasan et al., 2017?).
 The use of in vitro, or unreliable in vivo parameter estimates, reduces confidence in the model predicted behaviour.
-Consequently, the reduction in confidence hampers the use of these models to gain insight into the functioning of metabolic networks (ensemble modeling and ORACLE papers).
+Consequently, the reduction in confidence hampers the use of these models to gain insight into the functioning of metabolic networks \parencite{Tran2008, Chakrabarti2013a}.
 The insights gained from the use of kinetic models are subsequently used to design changes to these metabolic networks to achieve various goals.
-These goals could either be to increase metabolite production for biosynthesis of different chemicals (kinetic model for BDO in ecoli, k-ecoli457, Srinivasan et al., 2017) or to find therapeutic targets that can cut-off metabolic pathways to kill cancer cells (new paper on using geneMCS for cancer cell lines).
+These goals could either be to increase metabolite production for biosynthesis of different chemicals \parencite{Almquist2014, Khodayari2016, Costa2016, Andreozzi2016}(Srinivasan et al., 2017) or to find therapeutic targets to cure ailments \parencite{Apaolaza2017}.
 Hence, an increase in uncertainty in model predicted responses is also an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve these goals.
 
-If all intracellular metabolite concentrations can be measured over a time course, and the corresponding data is available, a nonlinear programming problem can be formulated to estimate the enzyme kinetic parameters ($\theta$) in Equation (\ref{eq:kinstoich}).
+If all intracellular metabolite concentrations can be measured over a time course, a nonlinear programming problem can be formulated to estimate the enzyme kinetic parameters ($\theta$) in Equation (\ref{eq:kinstoich}), based on the measured data.
 The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations can be used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}), with fixed upper ($\theta_u$) and lower ($\theta_l$) bounds for the parameter values (Equation \ref{eq:lsqopt}b).
@@
 In such scenarios, the parameter estimation problem is modified to suit a new system of equations shown below (Equation \ref{eq:output}).
-The new system of equations is obtained by augmenting the original system (Equation \ref{eq:kinstoich}) with Equation (\ref{eq:output}c) that models the relationship between the measurable metabolite concentrations and fluxes, represented by the variable $y$, and the unmeasured concentrations ($x$) that are used in the original model (Equation \ref{eq:kinstoich}) above.
+The new system of equations is obtained by augmenting the original system (Equation \ref{eq:kinstoich}) with Equation (\ref{eq:output}c) that models the relationship between the measurable metabolite concentrations and fluxes ($y$) and the unmeasured concentrations ($x$) that are used in the original model (Equation \ref{eq:kinstoich}) above.
 Equation (\ref{eq:output}) now has additional parameters $\mu$ that define this relationship and also need to be estimated.
@@
 \end{subequations}
-In systems identification, the measured concentrations and fluxes, $y$ in Equation (\ref{eq:output}c), are called output or observed variables, and the unmeasured concentrations, $x$ in Equation (\ref{eq:output}a-b), are called the state variables.
-For estimating both $\theta$ and $\mu$, the metabolite concentrations ($x$) in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the output variables $y$.
+In systems identification, the measured concentrations and fluxes ($y$) are called output or observed variables, and the unmeasured concentrations ($x$) are called the state variables.
+For estimating both $\theta$ and $\mu$, the metabolite concentrations $x$ in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the output variables $y$.
 
@@
 The identifiability of parameters in nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability.
-Any system (Equation \ref{eq:output}) is said to be structurally identifiable if, for an input-output mapping defined by $\mathbf{y} = \Phi(\mu,u)$ for at least one input function $u$, any two values of parameters $\mathbf{\mu}_1$ and $\mathbf{\mu}_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
+Any system (Equation \ref{eq:output}) is said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\mu,u)$ for at least one input function $u$, any two values of parameters $\mathbf{\mu}_1$ and $\mathbf{\mu}_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
 \begin{align}\label{eq:stident}
@@
 Accordingly, any system that has an infinite number of solutions to the parameter estimation problem for all input functions is said to structurally non-identifiable.
-Thus, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unobservable state variables and the observable output variables.
-Accordingly, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
+So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unobservable state variables and the observable output variables.
+Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
 
 Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
-The effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
+So, the effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
 Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
@@
 The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
-The identifiability of parameters for each flux can be establish independently.
-b) in silico data generation to test the identifiability of a given parameter is an important step that is also required for experimental design using the proposed methodology.
-c) The steps for identifiability analysis for a parameters of a single flux are shown.
-d) If any of the proposed in silico experiments results in non-identifiability of any parameter, then the number and nature of experiments to be performed to establish identifiability of the corresponding parameter can be determined by following the steps shown.}\label{fig:ident-flowchart}
+The identifiability of parameters for each flux can be established independently.
+b) The steps for identifiability analysis for parameters of a single flux are shown.}\label{fig:ident-flowchart}
 \end{figure}

### 2018-01-09 ident/docs/ident_manuscript.tex
complete intro and methods w/o experimental design section

@@
 	\dot{x} = \mathbf{S}v\\
-	v = g(x, \theta, u)
+	v = f(x, \theta, u)
 	\end{align}
 \end{subequations}
-The expression for the nonlinear function ($g$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Heijnen2005, Link2014, Machado2011, Srinivasan2015}.
-Accordingly, $g$ is a nonlinear function of the metabolite concentrations, enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
+The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Heijnen2005, Link2014, Machado2011, Srinivasan2015}.
+Accordingly, $f$ is a nonlinear function of the metabolite concentrations, enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
 
@@
 	\dot{x} = \mathbf{S}v\\
-	v = g(x, y, \theta, u)\\
-	\dot{y} = h(x, y, \mu, u)
+	v = f(x, y, \theta, u)\\
+	\dot{y} = h(x, y, \theta, u)
 	\end{align}
@@
 In systems identification, the measured concentrations and fluxes ($y$) are called output or observed variables, and the unmeasured concentrations ($x$) are called the state variables.
-For estimating both $\theta$ and $\mu$, the metabolite concentrations $x$ in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the output variables $y$.
+For estimating $\theta$, the metabolite concentrations $x$ in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the output variables $y$.
 
-However, the ability to determine unique solutions to parameters $\theta$ and $\mu$ is governed by the identifiability of these parameters in the model \parencite{McLean2012}.
+However, the ability to determine unique solutions to parameters $\theta$ is governed by the identifiability of these parameters in the model \parencite{McLean2012}.
 The identifiability of parameters in nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability.
-Any system (Equation \ref{eq:output}) is said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\mu,u)$ for at least one input function $u$, any two values of parameters $\mathbf{\mu}_1$ and $\mathbf{\mu}_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
+Any system (Equation \ref{eq:output}) is said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
 \begin{align}\label{eq:stident}
-\Phi(\mathbf{\mu}_1,u) = \Phi(\mathbf{\mu}_2,u) \iff \mathbf{\mu}_1 = \mathbf{\mu}_2
+\Phi(\theta_1,u) = \Phi(\theta_2,u) \iff \theta_1 = \theta_2
 \end{align}
-Accordingly, any system that has an infinite number of solutions to the parameter estimation problem for all input functions is said to structurally non-identifiable.
+Accordingly, the system can have a unique solution, a finite number of non-unique solutions or an infinite number of solutions for all input functions, and is said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
 So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unobservable state variables and the observable output variables.
@@
 
-Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated.
+Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated from measurable data.
 On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
 
-Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems (IEEE Trans paper from 2007).
-Most of these algorithms use methods based on differential algebra (Glad and Ljung, 1994).
-More recent methods take a profile-likelihood-based approach (2012 Paper) to establish both structural and practical identifiability.
-However, these methods scale poorly with increase in size of the modeled system.
-The aforementioned methods are also heavily dependent on the availability of dynamic time course data for all the output variables of the system.
+Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems \parencite{Audoly2001}(IEEE Trans paper from 2007).
+Most of these algorithms use methods based on differential algebra \parencite{Ljung1994}.
+More recent methods take a profile-likelihood-based approach to establish both structural and practical identifiability \parencite{Raue2009a}.
+All of these methods are heavily dependent on the availability of dynamic time course data for all the output variables of the system and scale poorly with increase in size of the modeled system \parencite{Bellu2007}.
 
-Computational burden due to poor scalability can be partly addressed with the current increases in computational power, while current improvements in measuring technologies can enable one to obtain high frequency dynamic data, no scalable methodologies exist to utilize the abundantly available steady state data for parameter estimation and identifiability for large metabolic networks.
+Computational burden due to poor scalability can be partly addressed with the current increases in computational power.
+Although, improvements in measurement technologies can enable one to obtain high frequency dynamic data, no scalable methodologies exist to utilize the abundantly available steady state data for parameter estimation and identifiability for large metabolic networks.
 
@@
 \section{Methods:}	
-\subsection{A method for practical identifiability of kinetic models of metabolism:}\label{sec:ident}
-In this section, we show how practical identifiability of kinetic parameters in a dynamic model of metabolism can be established using the nonlinear algebraic kinetic rate law formulations that describe the relationship between the concentrations, fluxes and the kinetic parameters of the metabolic network model.
-A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
-In a kinetic model, the value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations.
-Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations (Equation \ref{eq:flux_gen}).
-The fluxes are expressed as a function of the metabolite concentrations $\mathbf{x}$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
-\begin{align}\label{eq:flux_gen}
-v_i = f(\mathbf{x},\mathbf{\theta}, u)
-\end{align}
-Let $\theta \in \mathbb{R}^p$ for each flux $v_i$ in the network.
-For each experiment $j = {1, 2, ..., n}$, we assume that all metabolite concentrations $\mathbf{x}$ and reaction fluxes $\mathbf{v}$ are measurable.
-The pertinent information for each experiment is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
-
-In order to establish the identifiability of kinetic parameters for each flux $v_i$, we describe a computer algebra-based method.
-The primary use of the computer algebra system is to obtain closed form expressions for each parameter in $\theta$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}c).
-This is done by solving a system of nonlinear algebraic equations in $\mathbb{R}^p$, shown in Equation (\ref{eq:nonlineq}).
-\begin{align}\label{eq:nonlineq}
-v_{i, k} = f_k(\mathbf{x}_k,\theta, u_k) && \forall k=\{1, 2, ..., p\}\subset\{1, 2, ..., n\}
-\end{align}
-Each equation in (\ref{eq:nonlineq}), indicated by the index $k$, corresponds to the kinetic rate law expression $f(\mathbf{x},\mathbf{\theta}, u)$ for $v_i$, described earlier in Equation (\ref{eq:flux_gen}), written for concentrations and fluxes obtained from experiment $k$.
-Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$, where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}c).
-\begin{align}\label{eq:theta-eq}
-\theta_k = g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
-\end{align}
-The identifiability of parameter $\theta_k$ for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}c): any parameter $\theta_k$ is said to practically identifiable (practically non-identifiable) if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ ($D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$).
-
 \begin{figure}[!tbhp]
@@
 \end{figure}	
+\subsection{A method for practical identifiability of kinetic models of metabolism:}\label{sec:ident}
+In this section, we show how practical identifiability of kinetic parameters in a dynamic model of metabolism can be established.
+A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
+In a kinetic model, the value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
+Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations.
+The fluxes are expressed as a function of the metabolite concentrations $\mathbf{x}$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
 
-\subsection{Experimental design for parameter estimation in kinetic models of metabolism:}
-The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in Section \ref{sec:ident} and demonstrated in Section \ref{sec:example}.
-Subsequently, for any flux $v_i$, if for any $p$ combinations of indices $j$, the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make a parameter $\theta_k$ identifiable.
-Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
-Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
-This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
-In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
-This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
-Hence, our method is theoretically scalable even to genome-scale models.
-We show the application of this design methodology for flux $v_2$ in the gluconeogenic model in Section.
+Let $\theta \in \mathbb{R}^p$ in Equation (\ref{eq:output}b) for each flux $v_i$ in the network.
+For each experiment $j = {1, 2, ..., n}$, we assume that all metabolite concentrations $x$ and reaction fluxes $v$ are measurable.
+The pertinent information for each experiment is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
 
-Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability using our methodology.
+In order to establish the identifiability of kinetic parameters for each flux $v_i$, we describe a computer algebra-based method.
+The primary use of the computer algebra system is to obtain closed form expressions for each parameter in $\theta$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
+This is done by solving a system of nonlinear algebraic equations in $\mathbb{R}^p$, shown in Equation (\ref{eq:nonlineq}).
+\begin{align}\label{eq:nonlineq}
+v_{i, k} = f_k(\mathbf{x}_k,\theta, u_k) && \forall k=\{1, 2, ..., p\}\subset\{1, 2, ..., n\}
+\end{align}
+Each equation in (\ref{eq:nonlineq}), indicated by the index $k$, corresponds to the kinetic rate law expression $g(x, \theta, u)$ for $v_i$, described earlier in Equation (\ref{eq:output}b), written for concentrations and fluxes obtained from experiment $k$.
+Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$, where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
+\begin{align}\label{eq:theta-eq}
+\theta_k = g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
+\end{align}
+The identifiability of parameter $\theta_k$ for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable (practically non-identifiable) if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ ($D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$).
+Furthermore, the physical properties of the kinetic parameter values can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
 
-In the following sections we provide a previously published kinetic model of a small gluconeogenic network, followed by a demonstration of our methodology to establish practical identifiability and experimental design for one of the fluxes in this network.
+In the following sections we provide a previously published kinetic model of a small gluconeogenic network, followed by a demonstration of our methodology to establish practical identifiability for one of the fluxes in this network.
 
@@
 The aforementioned model of gluconeogenesis has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($V_1^{max}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
-Furthermore, the acetate concentration on which the cell is grown can also be perturbed to measure cellular response to changes in the substrate concentration.
-We use the in silico metabolomics and fluxomics data generated from these perturbation experiments to demonstrate parameter identification with our methodology.
-
-\section{Results:}
-First, in Section \ref{sec:example} we demonstrate how our methodology can be applied for one of the fluxes in the gluconeogenic model.
-In Section \ref{sec:closedform} that follows, we discuss the ability to apply our methods to different nonlinear kinetic rate law formulations within the context of the gluconeogenic model.
-Then, in Section we discuss the different experimental design strategies that were found through our identifiability analysis to enable kinetic parameter estimation for the each flux in the model using steady state data.
-Finally, in Section we expand the application of our method to a relatively large (39 reactions) kinetic model of the red blood cell metabolism.
+Furthermore, the acetate concentration, that determines the acetate uptake flux $v_1$, can also be perturbed to measure cellular response to changes in the substrate concentration.
+We use the in silico data generated by 18 different experiments wherein these four model parameters ($acetate$, $V_1^{max}$, $V_2^{max}$ and $V_3^{max}$) were perturbed to demonstrate practical parameter identification with our methodology.
+The experiments and the perturbed values of each of the four the parameters are given in the Appendix.
 
@@
 Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_2$ in the small model of gluconeogenesis described in Section \ref{sec:small-model}.
-Flux $v_2$ has two parameters, $V_2^{max}$ and $K_2^{pep}$ that need to be estimated from experimental data.
-Here, we assume that at least two different sets of experimental data for the concentrations and fluxes are available.
-We label the concentrations as ${pep}^1$ and ${pep}^2$, and the fluxes as ${v_2}^1$ and ${v_2}^2$ respectively, from each experiment.
+We choose flux $v_2$, expressed using Michaelis-Menten kinetics for this demonstration.
+$v_2$ has two parameters, $V_2^{max}$ and $K_2^{pep}$ that need to be identifiable so that they can be estimated from experimental data.
+Here, we assume that data (for the concentrations and fluxes) from at least two different sets of experiments is available.
+We label the available concentrations as ${pep}^1$ and ${pep}^2$, and the fluxes as ${v_2}^1$ and ${v_2}^2$ respectively, from each experiment.
 Accordingly, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for flux $v_2$ as follows:
@@
 \end{subequations}
-Solving this simultaneous system of equations, we get two nonlinear algebraic equations in the parameters $V_2^{max}$ and $K_2^{pep}$ based on the form shown earlier in Equation (\ref{eq:theta-eq}).
+Solving this simultaneous system of equations using Mathematica (Wolfram Research, USA), a computer algebra system, we get two nonlinear algebraic equations in the parameters $V_2^{max}$ and $K_2^{pep}$ based on the form shown earlier in Equation (\ref{eq:theta-eq}).
 \begin{subequations}\label{eq:v2_par}
@@
 \end{subequations}
-In Equation (\ref{eq:v2_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_2^{max}$ (Equation \ref{eq:v2_par}a) and $K_2^{pep}$ (Equation \ref{eq:v2_par}b).
+In Equation (\ref{eq:v2_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_2^{max}$ (Equation \ref{eq:v2_par}a) and $K_2^{pep}$ (Equation \ref{eq:v2_par}b) for different available experimental data set combinations.
+Since the enzyme binding constant ($K_2^{pep}$) and maximum reaction rate ($V_2^{max}$) cannot be negative, we can further reduce the criteria for identifiability for both these parameters by saying that the evaluated expressions should be non-negative.
+
+\section{Results:}
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure3/ident_analysis}}
+	\caption{to be filled in}\label{fig:figure1}
+\end{figure}
+
+\subsection{Experimental design for parameter estimation in kinetic models of metabolism:}
+The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in Section \ref{sec:ident} and demonstrated in Section \ref{sec:example}.
+Subsequently, for any flux $v_i$, if for any $p$ combinations of indices $j$, the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make a parameter $\theta_k$ identifiable.
+Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
+Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
+This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
+In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
+This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
+Hence, our method is theoretically scalable even to genome-scale models.
+We show the application of this design methodology for flux $v_2$ in the gluconeogenic model in Section.
+
+Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability using our methodology.
+
+\section{Results:}
+First, in Section \ref{sec:example} we demonstrate how our methodology can be applied for one of the fluxes in the gluconeogenic model.
+In Section \ref{sec:closedform} that follows, we discuss the ability to apply our methods to different nonlinear kinetic rate law formulations within the context of the gluconeogenic model.
+Then, in Section we discuss the different experimental design strategies that were found through our identifiability analysis to enable kinetic parameter estimation for the each flux in the model using steady state data.
+Finally, in Section we expand the application of our method to a relatively large (39 reactions) kinetic model of the red blood cell metabolism.
 

### 2018-01-09 ident/docs/ident_manuscript.tex
figure caption update and methods update

@@
 \subsection{Identifiability of parameters in a kinetic model of gluconeogenesis:}\label{sec:example}
-Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_2$ in the small model of gluconeogenesis described in Section \ref{sec:small-model}.
+Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for fluxes $v_1$, $v_2$ and $v_3$ in the small model of gluconeogenesis described in Section \ref{sec:small-model}.
+
+For flux $v_1$
 We choose flux $v_2$, expressed using Michaelis-Menten kinetics for this demonstration.
@@
 
+We found that the complexity of the MWC kinetic rate law formulation precluded its use to determine closed form expressions for the parameters in $v_3$ in Mathematica.
+Hence, we modeled the allosterically activated flux $v_3$ using the convenience kinetic rate law formulation (reference for convenience kinetics).
+The corresponding expression obtained for $v_3$ is given below (Equation).
+
+Using this expression for our identifiability analysis, we found that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different close form expressions owing to the presence of a square root term in their solutions.
+These distinct expressions are denoted by (1) and (2) following the parameter names throughout the rest of the document.
+
 \section{Results:}
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure3/ident_analysis}}
-	\caption{to be filled in}\label{fig:figure1}
+	\centering{\includegraphics[width=1\textwidth,height=.8\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure1/noise_no_noise_kotte_ident_jan01}}
+	\caption{The number of data sets, made from a combination of 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
+In the presence of additive noise in the flux and concentration data used for identifiability, the number of data sets practically identifying parameters in fluxes d)$v_1$, e)$v_2$ and f)$v_3$ are shown.
+25 samples of additive noise $\epsilon ~ N(0, 0.05)$ were used to generate noisy data for figures in the right hand side panel.}\label{fig:figure1}
 \end{figure}

### 2018-01-10 ident/docs/ident_manuscript.tex
improve methods + strat with results

--- 2018-01-09 ident/docs/ident_manuscript.tex
@@
 \usepackage{amssymb}
-\usepackage{array}
 \usepackage{tabu}
@@
 \begin{equation}\label{eq:flux1}
-v_1 = k_{1}^{cat}E\frac{acetate}{acetate+K_{1}^{acetate}}
+v_1 = k_{1}^{cat}E\frac{ac}{ac+K_{1}^{ac}}
 \end{equation}	
@@
 
+\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis:}\label{sec:example}
+Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for fluxes $v_1$, $v_2$ and $v_3$ in the small model of gluconeogenesis described in Section \ref{sec:small-model}.
+For the purposes of our identifiability methodology, we assume that relevant steady state metabolite concentrations and fluxes can be measured.
+
+In flux $v_1$, the concentration of the enzyme E is used as a variable.
+If the enzyme concentration can be measured, then the expression for $v_1$ given in Equation (\ref{eq:flux1}) can be used for identifiability analysis of parameters $k_1^{cat}$ and $K_1^{ac}$.
+However, if enzyme concentrations are not available, the expression in Equation (\ref{eq:flux1}) can be modified as given in Equation (\ref{eq:flux1a}) below.
+This equation does not make use of enzyme concentrations as variables, and uses $V_1^{max}$ and $K_1^{ac}$ as parameters.
+\begin{align}\label{eq:flux1a}
+v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}}
+\end{align}
+
+We choose this expression for flux $v_1$ (Equation \ref{eq:flux1a}), expressed using Michaelis-Menten kinetics, to demonstrate our method for practical identifiability.
+Both $V_1^{max}$ and $K_1^{ac}$ need to be identifiable so that they can be estimated from experimental data.
+Here, we assume that data (for the concentrations and fluxes) from at least two different sets of experiments is available i.e., in Equation (\ref{eq:nonlineq}) $k = {1, 2}$ .
+We label the available concentrations and fluxes as ${ac}^{(k)}$ and ${v_1}^{(k)}$, respectively.
+Accordingly, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as follows:
+\begin{align}\label{eq:nonlin-flux2}
+{v_1}^{(k)} = V_{1}^{max}\frac{ac^{(k)}}{ac^{(k)}+K_{1}^{ac}} && \forall k=\{1, 2\}
+\end{align}
+
+Solving this simultaneous system of $k$ equations using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic equations in the parameters $V_1^{max}$ and $K_1^{ac}$ based on the form shown earlier in Equation (\ref{eq:theta-eq}).
+\begin{subequations}\label{eq:v1_par}
+	\begin{align}		
+	V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
+	K_1^{ac} = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
+	\end{align}
+\end{subequations}
+In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}$ (Equation \ref{eq:v1_par}b) for different available experimental data set combinations.
+Since the enzyme binding constant ($K_1^{ac}$) and maximum reaction rate ($V_1^{max}$) cannot be negative, we can further reduce the criteria for identifiability for both these parameters by saying that the evaluated expressions should be non-negative.
+
 \subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis:}
@@
 Furthermore, the acetate concentration, that determines the acetate uptake flux $v_1$, can also be perturbed to measure cellular response to changes in the substrate concentration.
-We use the in silico data generated by 18 different experiments wherein these four model parameters ($acetate$, $V_1^{max}$, $V_2^{max}$ and $V_3^{max}$) were perturbed to demonstrate practical parameter identification with our methodology.
+We use the in silico data generated by 18 different experiments wherein these four model parameters ($ac$, $V_1^{max}$, $V_2^{max}$ and $V_3^{max}$) are perturbed to demonstrate practical parameter identification with our methodology.
 The experiments and the perturbed values of each of the four the parameters are given in the Appendix.
 
-\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis:}\label{sec:example}
-Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_2$ in the small model of gluconeogenesis described in Section \ref{sec:small-model}.
-We choose flux $v_2$, expressed using Michaelis-Menten kinetics for this demonstration.
-$v_2$ has two parameters, $V_2^{max}$ and $K_2^{pep}$ that need to be identifiable so that they can be estimated from experimental data.
-Here, we assume that data (for the concentrations and fluxes) from at least two different sets of experiments is available.
-We label the available concentrations as ${pep}^1$ and ${pep}^2$, and the fluxes as ${v_2}^1$ and ${v_2}^2$ respectively, from each experiment.
-Accordingly, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for flux $v_2$ as follows:
-\begin{subequations}\label{eq:nonlin-flux2}
-	\begin{align}
-	{v_2}^1 = V_{2}^{max}\frac{pep^1}{pep^1+K_{2}^{pep}}\\
-	{v_2}^2 = V_{2}^{max}\frac{pep^2}{pep^2+K_{2}^{pep}}
-	\end{align}
-\end{subequations}
-Solving this simultaneous system of equations using Mathematica (Wolfram Research, USA), a computer algebra system, we get two nonlinear algebraic equations in the parameters $V_2^{max}$ and $K_2^{pep}$ based on the form shown earlier in Equation (\ref{eq:theta-eq}).
-\begin{subequations}\label{eq:v2_par}
-	\begin{align}		
-	V_2^{max} = \frac{v_2^1v_2^2(pep^1-pep^2)}{v_2^2pep^1-v_2^1pep^2}\\
-	K_2^{pep} = \frac{pep^1(v_2^1pep^2-v_2^2pep^2)}{v_2^2pep^1-v_2^1pep^2}
-	\end{align}
-\end{subequations}
-In Equation (\ref{eq:v2_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_2^{max}$ (Equation \ref{eq:v2_par}a) and $K_2^{pep}$ (Equation \ref{eq:v2_par}b) for different available experimental data set combinations.
-Since the enzyme binding constant ($K_2^{pep}$) and maximum reaction rate ($V_2^{max}$) cannot be negative, we can further reduce the criteria for identifiability for both these parameters by saying that the evaluated expressions should be non-negative.
+The minimum number of experiments from which data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
+For instance, as demonstrated above, data from two distinct experiments is required for identifying the two parameters of $v_1$.
+In this case $p=2$.
+This also applies for identifying parameters in $v_2$.
+
+Looking at all the three fluxes whose parameters have to be identified, $v_3$ has the maximum number of parameters at three.
+Hence, we choose to select multiple distinct data sets, each with three different experiments to establish the practical identifiability of all the parameters in the model.
+Accordingly, we can select 4896 (18 x 17 x 16) distinct combinations of three experiments from the 18 different experiments mentioned earlier and elaborated in the Appendix.
 
 \section{Results:}
+First in Section \ref{sec:closedform}, we discuss the ability to apply our methods to different nonlinear kinetic rate law formulations within the context of the gluconeogenic model.
+%Then, in Section we discuss the different experimental design strategies that were found through our identifiability analysis to enable kinetic parameter estimation for the each flux in the model using steady state data.
+Finally, in Section we expand the application of our method to a relatively large (39 reactions) kinetic model of the red blood cell metabolism.
+\subsection{Getting closed-form expressions for each flux in Kotte model:}\label{sec:closedform}
+As demonstrated for identifying parameters of $v_1$ earlier in Section \ref{sec:example}, the computational effort required to establish identifiability is governed by the nonlinear complexity of the enzyme kinetic rate law used to model a specific flux.
+Although computer algebra systems are capable of handling complex symbolic calculations, sometimes, the complexity of getting closed form expressions for all kinetic parameters of certain rate law formulations is too much for the computer algebra system to overcome.
+We encounter this scenario when expression for the the allosterically activated reaction $v_3$ is described by the MWC model for allosteric regulation.
+We found that the complexity of the MWC kinetic rate law precluded its use to determine closed form expressions for the parameters in $v_3$ using either Mathematica or SymPy in Python.
+In order to overcome this computational obstacle, we modeled the reaction rate for $v_3$ using the convenience kinetic rate law formulation (reference for convenience kinetics).
+The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
+\begin{align}\label{eq:flux3_convkin}
+content...
+\end{align}
+
+Using this expression for identifiability analysis, we found that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different close form expressions owing to the presence of a square root term in their solutions.
+These distinct expressions are denoted by (1) and (2) following the respective parameter names throughout the rest of the document.
+
+We believe this problem of computational tractability to occur in other complex kinetic rate law formulations as well. Example?
+
+\subsection{Variability in the degree of identifiability of each parameter:}
+We found that the ability of a combination of experiments to identify a particular parameter widely varies depending the the position of the flux in the metabolic network with respect to the inputs and the outputs of the network.
+In Figure \ref{fig:figure1} we show the number of combinations of experiments that are capable of identifying each parameter in each flux.
+In the left panels (Figure \ref{fig:figure1}a-c), the data is show when noise-free data is used for identification.
+The identifiability information for parameters using data with additive noise is shown on the right panel (Figure \ref{fig:figure1}d-f).
+
+Both in the presence as well as absence of noise, the identifiability of parameters in $v_1$ is guaranteed by only about 20\% of the available data combinations (Figure \ref{fig:figure1}a, d).
+As seen in Section \ref{sec:example}, the identifiability expressions for both parameters in $v_1$ are dependent on changes in the both the uptake flux $v_1$ as well as the acetate concentration $ac$ in both the experiments used for identification.
+If either of these values cannot be distinguished between the two experiments whose data is used for identification, then the corresponding combination of experiments can be guaranteed to result in practical non-identifiability of the parameters in $v_1$.
+This is further illustrated in Figure below where the different experiments in each combination are shown.
+
+When the experimental data has no measurement noise associated with it, then $V_2^{max}$ is the only parameter that can be identified by the most number of experimental combinations (89.5\%).
+In contrast to the high identifiability of this parameter, $K_2^{pep}$, the other parameter for $v_2$ is identified by less than 2\% of the available data combinations.
+
+This is followed by $V_3^{max}$ that is capable of being identified by 76.3\% of the available data combinations.
+
+In the small model that we use to elucidate our methodology, we find that only 
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure3/ident_analysis}}
-	\caption{to be filled in}\label{fig:figure1}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure1/noise_no_noise_kotte_ident_jan01}}
+	\caption{The number of data sets, made from a combination of 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
+In the presence of additive noise in the flux and concentration data used for identifiability, the number of data sets practically identifying parameters in fluxes d) $v_1$, e) $v_2$ and f) $v_3$ are shown.
+25 samples of additive noise $\epsilon ~ N(0, 0.05)$ were used to generate noisy data for figures in the right hand side panel.}\label{fig:figure1}
 \end{figure}
+
+\subsection{Utility of data for identifiability purposes:}
+
+\subsection{Experiments that can establish identifiability for each flux in Kotte model:}	
+As indicated previously in Section, we performed different experiments by perturbing the parameters for fluxes $v_1$, $v_2$ and $v_3$ in the model.
+We also perturbed the input substrate concentration to collect data on the concentrations and fluxes within the network.
+We describe the results of using this data to establish the identifiability of various parameters in the kinetic model of the small metabolic network.
 
@@
 
-\section{Results:}
-First, in Section \ref{sec:example} we demonstrate how our methodology can be applied for one of the fluxes in the gluconeogenic model.
-In Section \ref{sec:closedform} that follows, we discuss the ability to apply our methods to different nonlinear kinetic rate law formulations within the context of the gluconeogenic model.
-Then, in Section we discuss the different experimental design strategies that were found through our identifiability analysis to enable kinetic parameter estimation for the each flux in the model using steady state data.
-Finally, in Section we expand the application of our method to a relatively large (39 reactions) kinetic model of the red blood cell metabolism.
+\subsection{Combinations of experiments that will enable identification of all model parameters:}
 
-\subsection{Getting closed-form expressions for each flux in Kotte model:}\label{sec:closedform}
-The complexity of the equations in our specific scenario is determined by the complexity of enzyme-metabolite interaction models used describe fluxes in metabolic networks.
-Although computer algebra systems (CAS) are capable of handling complex symbolic calculations, sometimes, the complexity of getting closed form expressions for all kinetic parameters of certain rate law formulations is too much for the CAS to overcome.
-We encountered this scenario in the case of the gluconeogenic model for flux $v_3$ where the kinetics of the allosterically activated reaction are described by the MWC model for allosteric regulation.
-In order to overcome this computational difficulty, we used a convenience kinetic rate law formulation to describe the allosteric interaction in $v_3$.
-We give this formulation in Section.
-
-We believe this problem of computational tractability to occur in other complex kinetic rate law formulations as well. Examples?
-
-\subsection{Experiments that can establish identifiability for each flux in Kotte model:}	
-As indicated previously in Section, we performed different experiments by perturbing the parameters for fluxes $v_1$, $v_2$ and $v_3$ in the model.
-We also perturbed the input substrate concentration to collect data on the concentrations and fluxes within the network.
-We describe the results of using this data to establish the identifiability of various parameters in the kinetic model of the small metabolic network.
-
-\subsection{Combinations of experiments that will enable identification of all model parameters:}
 \subsection{Expanding methodology to RBC model:}
-
 If any data generated from a perturbation experiment $i$ results in nonidentifiability, we eliminate experiment $i$ from the list of experiments that need to be performed for parameter estimation.
 
+\section*{Appendix}
+\subsection*{In silico perturbation experiments for gluconeogenic model:}
 \begin{table}[!tbhp]
@@
 \end{table}
-
-\paragraph{Outline:}
-\begin{itemize}
-	\item parameter estimation is a well developed field typically using minimization of least square error to estimate model parameters from available experimental data
-	\item if parameters are structurally identifiable, it does not guarantee practical identifiability from noisy experimental data
-	\item identifiability dependent on whether given datasets (outputs) for estimation can sufficiently distinguish between different parameter values
-\end{itemize}
-
-\paragraph{Sections:}
-\begin{itemize}
-	\item datasets required for parameter estimation in kinetic models of metabolism (methods?)		
-	\item identifiability in kotte model - scalability, number of experiments required, requirements for time course data(? in the intro)		
-	\item identifiability in large rbc model
-\end{itemize}
 

### 2018-01-10 ident/docs/ident_manuscript.tex
add new figure for experiments in k1cat identification

@@
 \end{figure}
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/k1cat_no_noise_experiments_jan03}}
+	\caption{The contribution of different experiments types used in combinations that can practically identify parameter $k_1^{cat}$ in $v_1$.
+Panels a) and b) represent the position of the experiment in $k = \{1, 2\}$}\label{fig:figure2}
+\end{figure}
 

### 2018-01-10 ident/docs/ident_manuscript.tex
expanded explanation for figure 1 in results

@@
 \subsection{Variability in the degree of identifiability of each parameter:}
+In Figure \ref{fig:figure1} we show the number of combinations of experiments that are capable of identifying each parameter in each flux.
+In the left panels (Figure \ref{fig:figure1}a-c), the data is shown when noise-free data is used for identification.
+The identifiability information for parameters using data with additive noise is shown on the right panel (Figure \ref{fig:figure1}d-f).
+As seen in these figures, irrespective of the presence or absence of noise in the data, the number of combinations that can identify each parameter varies widely, even for parameters within the same flux.
+
+Since both fluxes $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law, we compare the number of data combinations that are capable of identifying the different parameters used to model these fluxes.
+We observe that the number of data combinations identifying the maximum reaction rates ($V_i^{max}$) for both $v_1$ (Figure \ref{fig:figure1}a and d) and $v_2$ (Figure \ref{fig:figure1}b and e) is always greater for than the number of data combinations that identify the enzyme binding ($K_i$) constants for both of these fluxes, irrespective of the presence of noise in the experimental data.
+
+Although, only about 20\% of the 4896 different data combinations can identify parameters in $v_1$, the difference between the number of data combinations being able to identify the different parameters in $v_1$ is less than 2\% (Figure \ref{fig:figure1}a and d).
+However, the same could not be said for the number of data combinations identifying parameters in $v_2$.
+In contrast to the high number of data combinations identifying $V_2^{max}$ (89.5\%), less than 2\% of the combinations are able to identify $K_2^{pep}$ (Figure \ref{fig:figure1}b and e).
+
+The trend noted earlier for $v_1$ and $v_2$ with regards to the number of data combinations identifying the reaction rate parameters versus the number of data sets identifying the corresponding enzyme binding constants in these fluxes is also seen for $v_3$, i.e., the percentage of data combinations identifying $V_3^{max} (1)$ and $V_3^{max}(2)$ is greater than the percentage of combinations that can identify $K_3^{fdp}(1)$ and $K_3^{fdp}(2)$.
+
 We found that the ability of a combination of experiments to identify a particular parameter widely varies depending the the position of the flux in the metabolic network with respect to the inputs and the outputs of the network.
-In Figure \ref{fig:figure1} we show the number of combinations of experiments that are capable of identifying each parameter in each flux.
-In the left panels (Figure \ref{fig:figure1}a-c), the data is show when noise-free data is used for identification.
-The identifiability information for parameters using data with additive noise is shown on the right panel (Figure \ref{fig:figure1}d-f).
+
+As seen from Figure \ref{fig:figure1},
 

### 2018-01-10 ident/docs/ident_manuscript.tex
updates to intro

@@
 If all intracellular metabolite concentrations can be measured over a time course, a nonlinear programming problem can be formulated to estimate the enzyme kinetic parameters ($\theta$) in Equation (\ref{eq:kinstoich}), based on the measured data.
-The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations can be used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}), with fixed upper ($\theta_u$) and lower ($\theta_l$) bounds for the parameter values (Equation \ref{eq:lsqopt}b).
+The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}), with fixed upper ($\theta_u$) and lower ($\theta_l$) bounds for the parameter values (Equation \ref{eq:lsqopt}b).
 The difference between the data and the model estimate for each output and at each time point in the objective function is weighted by the variance in the experimental data $\sigma_{kl}^*$ for that corresponding variable and time point.
@@
 The new system of equations is obtained by augmenting the original system (Equation \ref{eq:kinstoich}) with Equation (\ref{eq:output}c) that models the relationship between the measurable metabolite concentrations and fluxes ($y$) and the unmeasured concentrations ($x$) that are used in the original model (Equation \ref{eq:kinstoich}) above.
-Equation (\ref{eq:output}) now has additional parameters $\mu$ that define this relationship and also need to be estimated.
+The parameter vector ($\theta$) is augmented with additional parameters that define this relationship.
+These additional parameters also need to be estimated.
 \begin{subequations}\label{eq:output}
@@
 
+Algorithms for structural identification of parameters have 
 Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems \parencite{Audoly2001}(IEEE Trans paper from 2007).
 Most of these algorithms use methods based on differential algebra \parencite{Ljung1994}.
-More recent methods take a profile-likelihood-based approach to establish both structural and practical identifiability \parencite{Raue2009a}.
-All of these methods are heavily dependent on the availability of dynamic time course data for all the output variables of the system and scale poorly with increase in size of the modeled system \parencite{Bellu2007}.
+The underlying motivation for these methods is the need to change the non-convex parameter estimation problem (Equation \ref{eq:lsqopt}) to a convex problem due to inherent difficulties in obtaining global optimal solutions for non-convex problems \parencite{Bellu2007}.
+However, due to the use of computer algebra systems to solve the identifiability problem, these methods are not scalable to large systems like a kinetic model of a large metabolic network .
 
-Computational burden due to poor scalability can be partly addressed with the current increases in computational power.
-Although, improvements in measurement technologies can enable one to obtain high frequency dynamic data, no scalable methodologies exist to utilize the abundantly available steady state data for parameter estimation and identifiability for large metabolic networks.
+Recently proposed methods address the issue of scalability by using a profile-likelihood-based approach to establish structural identifiability \parencite{Raue2009a}.
+Profile likelihood-based methods offer an advantage over the differential algebra methods due to their ability to not only structurally identify parameters, but also help with experimental design through practical identifiability of parameters.
+
+In contrast to differential algebra-based methods, these methods still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters.
+Hence, these methods can still be considered to be un-scalable for structurally identifying parameters in large metabolic networks.
+Nonetheless,
+
+Apart from issues with their scalability, profile likelihood-based methods also depend on the availability of actual measured or simulated dynamic data to establish structural identifiability.
+
+Although, improvements in measurement technologies can enable one to obtain high frequency dynamic data to address the need for time course data required by these methods,
+
+The profile likelihood-based methods and the differential algebra-based methods are heavily dependent on the availability of
+
+ for all the output variables of the system and scale poorly with increase in size of the modeled system.
+
+However, computational burden due to poor scalability can be partly addressed with the current increases in computational power.
+
+, no scalable methodologies exist to utilize the abundantly available steady state data for parameter estimation and identifiability for large metabolic networks.
 

### 2018-01-10 ident/docs/ident_manuscript.tex
complete intro with experimental design focus

--- 2018-01-10 ident/docs/ident_manuscript.tex
@@
 
-Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems \parencite{Audoly2001}(IEEE Trans paper from 2007).
-Most of these algorithms use methods based on differential algebra \parencite{Ljung1994}.
-More recent methods take a profile-likelihood-based approach to establish both structural and practical identifiability \parencite{Raue2009a}.
-All of these methods are heavily dependent on the availability of dynamic time course data for all the output variables of the system and scale poorly with increase in size of the modeled system \parencite{Bellu2007}.
+Methods and tools for structural identification of parameters based on differential algebra \parencite{Ljung1994, Audoly2001, Bellu2007} and profile likelihood \parencite{Raue2009a} are available.
+However, only the profile likelihood-based methods enable experimental design by facilitating practical identification of parameters.
+Nonetheless, these methods still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters, and hence still suffer from all the inherent difficulties associated with obtaining global optimal solutions for non-convex optimization problems.
+This also makes them un-scalable for experimental design and practical identifiability of parameters in kinetic models of large metabolic networks.
 
-Computational burden due to poor scalability can be partly addressed with the current increases in computational power.
-Although, improvements in measurement technologies can enable one to obtain high frequency dynamic data, no scalable methodologies exist to utilize the abundantly available steady state data for parameter estimation and identifiability for large metabolic networks.
-
-In this paper, we propose a methodology to establish practical identifiability for parameters in kinetic models of metabolism.
+In this paper, we propose a scalable methodology to establish practical identifiability for parameters in kinetic models of metabolism using abundantly available steady state concentration and flux data.
 We present a computer algebra-based method that can facilitate practical identifiability as well as experimental design for estimating parameters separately for each individual reaction within a metabolic network based on available steady state experimental data.
-This enables us to address the twin issues of scalability and data availability.
 We illustrate the utility of this method by applying it for a small network of gluconeogenesis in \textit{E.

### 2018-01-10 ident/docs/ident_manuscript.tex
update to results

@@
 If all intracellular metabolite concentrations can be measured over a time course, a nonlinear programming problem can be formulated to estimate the enzyme kinetic parameters ($\theta$) in Equation (\ref{eq:kinstoich}), based on the measured data.
-The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}), with fixed upper ($\theta_u$) and lower ($\theta_l$) bounds for the parameter values (Equation \ref{eq:lsqopt}b).
-The difference between the data and the model estimate for each output and at each time point in the objective function is weighted by the variance in the experimental data $\sigma_{kl}^*$ for that corresponding variable and time point.
+The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration at each time point, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}).
+The parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).
 \begin{subequations}\label{eq:lsqopt}
@@
 However, only the profile likelihood-based methods enable experimental design by facilitating practical identification of parameters.
-Nonetheless, these methods still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters, and hence still suffer from all the inherent difficulties associated with obtaining global optimal solutions for non-convex optimization problems.
-This also makes them un-scalable for experimental design and practical identifiability of parameters in kinetic models of large metabolic networks.
+Nonetheless, this method still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters, and hence still suffers from all the inherent difficulties associated with obtaining global optimal solutions for non-convex optimization problems.
+This also makes it un-scalable for experimental design and practical identifiability of parameters in kinetic models of large metabolic networks.
 
 In this paper, we propose a scalable methodology to establish practical identifiability for parameters in kinetic models of metabolism using abundantly available steady state concentration and flux data.
-We present a computer algebra-based method that can facilitate practical identifiability as well as experimental design for estimating parameters separately for each individual reaction within a metabolic network based on available steady state experimental data.
-We illustrate the utility of this method by applying it for a small network of gluconeogenesis in \textit{E.
+We present a computer algebra-based method that can facilitate experimental design through practical identifiability of parameters separately for each individual reaction within a metabolic network based on available steady state experimental data.
+We illustrate the utility of this method by applying it to a small network of gluconeogenesis in \textit{E.
 coli} and demonstrating our ability to propose experiments that will facilitate parameter estimation for a kinetic model of this network.
@@
 Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations.
-The fluxes are expressed as a function of the metabolite concentrations $\mathbf{x}$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
+The fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
 
@@
 \end{align}
-Each equation in (\ref{eq:nonlineq}), indicated by the index $k$, corresponds to the kinetic rate law expression $g(x, \theta, u)$ for $v_i$, described earlier in Equation (\ref{eq:output}b), written for concentrations and fluxes obtained from experiment $k$.
+Each equation in (\ref{eq:nonlineq}), indicated by the index $k$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for $v_i$, described earlier in Equation (\ref{eq:output}b), written for concentrations and fluxes obtained from experiment $k$.
 Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$, where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
@@
 \subsection{Identifiability of parameters in a kinetic model of gluconeogenesis:}\label{sec:example}
-Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for fluxes $v_1$, $v_2$ and $v_3$ in the small model of gluconeogenesis described in Section \ref{sec:small-model}.
-For the purposes of our identifiability methodology, we assume that relevant steady state metabolite concentrations and fluxes can be measured.
+Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_1$ in the small model of gluconeogenesis described in Section \ref{sec:small-model}.
+For the purposes of this demonstration, we assume that all relevant steady state metabolite concentrations and fluxes can be measured.
 
@@
 \begin{align}\label{eq:nonlin-flux2}
-{v_1}^{(k)} = V_{1}^{max}\frac{ac^{(k)}}{ac^{(k)}+K_{1}^{ac}} && \forall k=\{1, 2\}
+{v_1}^{(k)} = V_{1}^{max}\frac{ac^{(k)}}{ac^{(k)}+K_{1}^{ac}} &&  k=\{1, 2\}
 \end{align}
@@
 In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}$ (Equation \ref{eq:v1_par}b) for different available experimental data set combinations.
-Since the enzyme binding constant ($K_1^{ac}$) and maximum reaction rate ($V_1^{max}$) cannot be negative, we can further reduce the criteria for identifiability for both these parameters by saying that the evaluated expressions should be non-negative.
+Since the enzyme binding constant ($K_1^{ac}$) and maximum reaction rate ($V_1^{max}$) cannot be negative, we can further reduce the criteria for identifiability for both these parameters by saying that the evaluated expressions should be non-negative (Figure \ref{fig:ident-flowchart}b).
 
-\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis:}
+\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis:}\label{sec:experiments}
 Steady state metabolomics and fluxomics data can be gathered under different physiological conditions by either perturbing the expression levels for different enzymes within a metabolic network, or by changing the substrate concentrations under which the cells grow.
@@
 
-Looking at all the three fluxes whose parameters have to be identified, $v_3$ has the maximum number of parameters at three.
+Looking at the three fluxes whose parameters have to be identified, $v_3$ has the maximum number of parameters at three.
 Hence, we choose to select multiple distinct data sets, each with three different experiments to establish the practical identifiability of all the parameters in the model.
 Accordingly, we can select 4896 (18 x 17 x 16) distinct combinations of three experiments from the 18 different experiments mentioned earlier and elaborated in the Appendix.
+
+\subsection{Experimental design through practical parameter identifiability for kinetic models of metabolism:}
+Following the methodology described in Section \ref{sec:ident}, and demonstrated in Section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found from Section \ref{sec:experiments} can be tested for their ability to practically identify parameters for any of the three fluxes in the small metabolic network.
+This step would help distinguish identifiable experiment combinations from combinations that do not practically identify any parameter in the model (Figure).
+Subsequently, it is possible to obtain a collection of experiments that make up all identifiable data combinations that can be performed to obtain the most minimal and informative set of experiments to identify as many parameters as possible.
+Consequently, the set of experiments can be used to estimate the identifiable parameters in the model.
 
@@
 \begin{align}\label{eq:flux3_convkin}
-content...
+	v_3 = V_3^{max}\left(\frac{1}{1 + \frac{K_3^{pep}}{pep}}\right)\left(\frac{\frac{fdp}{K_3^{fdp}}}{1 + \frac{fdp}{K_3^{fdp}}}\right)
 \end{align}
@@
 \subsection{Variability in the degree of identifiability of each parameter:}
-In Figure \ref{fig:figure1} we show the number of combinations of experiments that are capable of identifying each parameter in each flux.
+We use the percentage of data combinations that can identify a parameter as a naive measure of the degree of identifiability of that parameter.
+As an example, if 90\% of the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
+On the other hand if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%, and $\theta_i$ has a higher degree of identifiability than $\theta_j$.
+We can use this criteria to distinguish parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure1/noise_no_noise_kotte_ident_jan01}}
+	\caption{The number of data sets, made from a combination of 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
+In the presence of additive noise in the flux and concentration data used for identifiability, the number of data sets practically identifying parameters in fluxes d) $v_1$, e) $v_2$ and f) $v_3$ are shown.
+25 samples of additive noise $\epsilon ~ N(0, 0.05)$ were used to generate noisy data for figures in the right hand side panel.}\label{fig:figure1}
+\end{figure}
+
+In Figure \ref{fig:figure1} we show the number and percentage of experiment combinations that are capable of identifying each parameter in each flux.
 In the left panels (Figure \ref{fig:figure1}a-c), the data is shown when noise-free data is used for identification.
 The identifiability information for parameters using data with additive noise is shown on the right panel (Figure \ref{fig:figure1}d-f).
-As seen in these figures, irrespective of the presence or absence of noise in the data, the number of combinations that can identify each parameter varies widely, even for parameters within the same flux.
+The data shown in the right hand side panel was generated used 25 different samples of random additive noise selected from a normal distribution with 0 mean 0.05 standard deviation.
+The value used for standard deviation reflects a 5\% noise level.
+As seen in these figures, irrespective of the presence or absence of noise in the data, the number of combinations that can identify each parameter, and concordantly, the degree of identifiability of each parameter varies widely.
+Next, we look at trends in the degree of identifiability of each parameter in each flux.
 
 Since both fluxes $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law, we compare the number of data combinations that are capable of identifying the different parameters used to model these fluxes.
-We observe that the number of data combinations identifying the maximum reaction rates ($V_i^{max}$) for both $v_1$ (Figure \ref{fig:figure1}a and d) and $v_2$ (Figure \ref{fig:figure1}b and e) is always greater for than the number of data combinations that identify the enzyme binding ($K_i$) constants for both of these fluxes, irrespective of the presence of noise in the experimental data.
+We observe that the number of data combinations identifying the maximum reaction rates ($V_i^{max}$) for both $v_1$ (Figure \ref{fig:figure1}a and d) and $v_2$ (Figure \ref{fig:figure1}b and e) is always greater for than the number of data combinations that identify the enzyme binding ($K_i$) constants in the presence as well as absence of noise in the data.
+This trend is also observed for parameters in $v_3$: the percentage of data combinations identifying $V_3^{max} (1)$ and $V_3^{max}(2)$ are greater than the percentage of combinations that can identify $K_3^{fdp}(1)$ and $K_3^{fdp}(2)$ (Figure \ref{fig:figure1}c and f).
+As such all maximum reaction rate parameters in the network have a higher degree of identifiability that their corresponding enzyme binding constants.
 
-Although, only about 20\% of the 4896 different data combinations can identify parameters in $v_1$, the difference between the number of data combinations being able to identify the different parameters in $v_1$ is less than 2\% (Figure \ref{fig:figure1}a and d).
-However, the same could not be said for the number of data combinations identifying parameters in $v_2$.
-In contrast to the high number of data combinations identifying $V_2^{max}$ (89.5\%), less than 2\% of the combinations are able to identify $K_2^{pep}$ (Figure \ref{fig:figure1}b and e).
+The difference between the degrees of identifiability of each parameter within a single flux varies widely depending on the reaction flux.
+For instance, the difference in the number of combinations that can identify the parameters used to model $v_1$ is less than 2\% (Figure \ref{fig:figure1}a and d).
+In contrast, a high number of combinations (89.5\%) can identify $V_2^{max}$ in $v_2$, while less than 2\% of the combinations are able to identify $K_2^{pep}$ (Figure \ref{fig:figure1}b and e).
+This variability in the degrees of identifiability for various parameters modeling the same flux is also seen for $v_3$ (Figure \ref{fig:figure1}c and f): $V_3^{max} (1)$ and $V_3^{max} (2)$ have an average degree of identifiability of 75\% in comparison to the 15\% identifiability of both $K_3^{fdp} (1)$ and $K_3^{fdp} (2)$.
 
-The trend noted earlier for $v_1$ and $v_2$ with regards to the number of data combinations identifying the reaction rate parameters versus the number of data sets identifying the corresponding enzyme binding constants in these fluxes is also seen for $v_3$, i.e., the percentage of data combinations identifying $V_3^{max} (1)$ and $V_3^{max}(2)$ is greater than the percentage of combinations that can identify $K_3^{fdp}(1)$ and $K_3^{fdp}(2)$.
+Also notable is the difference in the degree of identifiability between the maximum reaction rates of the uptake flux $v_1$ and the intracellular fluxes $v_2$ and $v_3$.
+Only about 20\% of the 4896 different data combinations can identify either the maximum reaction rate ($V_1^{max}$) or the enzyme turnover rate ($k_1^{cat}$) in $v_1$ (Figure \ref{fig:figure1}a and d), in contrast to the high number of data combinations capable of identifying $V_2^{max}$ (89.5\%) (Figure \ref{fig:figure1}b and e), $V_3^{max} (1)$ or $V_3^{max} (2)$ (>75\%) (Figure \ref{fig:figure1}c and f).
 
-We found that the ability of a combination of experiments to identify a particular parameter widely varies depending the the position of the flux in the metabolic network with respect to the inputs and the outputs of the network.
-
-As seen from Figure \ref{fig:figure1},
+This informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending the position of the flux in the metabolic network with respect to the inputs and the outputs of the network.
+The dependency can be further elucidated using the closed form solutions obtained for parameters of $v_1$ in Section \ref{sec:example}.
+As seen in Section \ref{sec:example}, the identifiability expressions for both parameters in $v_1$ are dependent on changes in the both the uptake flux $v_1$ as well as the acetate concentration $ac$ in both the experiments used for identification.
+If either of these values cannot be distinguished between the two experiments whose data is used for identification, then the corresponding combination of experiments can be guaranteed to result in practical non-identifiability of the parameters in $v_1$.
+As the uptake flux for the metabolic network that is modeled to be dependent on the substrate concentration, the only way meaningful changes in $v_1$ can be observed are through experiments involving changes either in the substrate concentration or changes in the enzyme concentration for $v_1$.
+In contrast, any experiment that can distinguish between different values of $pep$ and $v_2$ can be used to identify $V_2^{max}$.
 
 Both in the presence as well as absence of noise, the identifiability of parameters in $v_1$ is guaranteed by only about 20\% of the available data combinations (Figure \ref{fig:figure1}a, d).
-As seen in Section \ref{sec:example}, the identifiability expressions for both parameters in $v_1$ are dependent on changes in the both the uptake flux $v_1$ as well as the acetate concentration $ac$ in both the experiments used for identification.
-If either of these values cannot be distinguished between the two experiments whose data is used for identification, then the corresponding combination of experiments can be guaranteed to result in practical non-identifiability of the parameters in $v_1$.
+
  This is further illustrated in Figure below where the different experiments in each combination are shown.
@@
 In the small model that we use to elucidate our methodology, we find that only
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure1/noise_no_noise_kotte_ident_jan01}}
-	\caption{The number of data sets, made from a combination of 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
-In the presence of additive noise in the flux and concentration data used for identifiability, the number of data sets practically identifying parameters in fluxes d) $v_1$, e) $v_2$ and f) $v_3$ are shown.
-25 samples of additive noise $\epsilon ~ N(0, 0.05)$ were used to generate noisy data for figures in the right hand side panel.}\label{fig:figure1}
-\end{figure}
+
 \begin{figure}[!tbhp]

### 2018-01-11 ident/docs/ident_manuscript.tex
more results for section 1

@@
 As an example, if 90\% of the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
-On the other hand if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%, and $\theta_i$ has a higher degree of identifiability than $\theta_j$.
+On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%, and $\theta_i$ has a higher degree of identifiability than $\theta_j$.
 We can use this criteria to distinguish parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
@@
 
-In Figure \ref{fig:figure1} we show the number and percentage of experiment combinations that are capable of identifying each parameter in each flux.
+In Figure \ref{fig:figure1} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
 In the left panels (Figure \ref{fig:figure1}a-c), the data is shown when noise-free data is used for identification.
 The identifiability information for parameters using data with additive noise is shown on the right panel (Figure \ref{fig:figure1}d-f).
-The data shown in the right hand side panel was generated used 25 different samples of random additive noise selected from a normal distribution with 0 mean 0.05 standard deviation.
+The data shown in the right hand side panel was generated using 25 different samples of random additive noise selected from a normal distribution with 0 mean 0.05 standard deviation.
 The value used for standard deviation reflects a 5\% noise level.
@@
 Since both fluxes $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law, we compare the number of data combinations that are capable of identifying the different parameters used to model these fluxes.
-We observe that the number of data combinations identifying the maximum reaction rates ($V_i^{max}$) for both $v_1$ (Figure \ref{fig:figure1}a and d) and $v_2$ (Figure \ref{fig:figure1}b and e) is always greater for than the number of data combinations that identify the enzyme binding ($K_i$) constants in the presence as well as absence of noise in the data.
-This trend is also observed for parameters in $v_3$: the percentage of data combinations identifying $V_3^{max} (1)$ and $V_3^{max}(2)$ are greater than the percentage of combinations that can identify $K_3^{fdp}(1)$ and $K_3^{fdp}(2)$ (Figure \ref{fig:figure1}c and f).
-As such all maximum reaction rate parameters in the network have a higher degree of identifiability that their corresponding enzyme binding constants.
+The number of data combinations identifying the maximum reaction rates ($V_i^{max}$) for both $v_1$ (Figure \ref{fig:figure1}a and d) and $v_2$ (Figure \ref{fig:figure1}b and e) is always greater for than the number of data combinations that identify the enzyme binding ($K_i$) constants in the presence as well as absence of noise in the data.
+Even for parameters in $v_3$ the percentage of data combinations identifying $V_3^{max} (1)$ and $V_3^{max}(2)$ are greater than the percentage of combinations that can identify $K_3^{fdp}(1)$ and $K_3^{fdp}(2)$ (Figure \ref{fig:figure1}c and f).
+Hence, all maximum reaction rate parameters in the network have a higher degree of identifiability that their corresponding enzyme binding constants.
 
 The difference between the degrees of identifiability of each parameter within a single flux varies widely depending on the reaction flux.
-For instance, the difference in the number of combinations that can identify the parameters used to model $v_1$ is less than 2\% (Figure \ref{fig:figure1}a and d).
+For instance, the difference between the number of combinations that can identify the parameters used to model $v_1$ is less than 2\% (Figure \ref{fig:figure1}a and d).
 In contrast, a high number of combinations (89.5\%) can identify $V_2^{max}$ in $v_2$, while less than 2\% of the combinations are able to identify $K_2^{pep}$ (Figure \ref{fig:figure1}b and e).
-This variability in the degrees of identifiability for various parameters modeling the same flux is also seen for $v_3$ (Figure \ref{fig:figure1}c and f): $V_3^{max} (1)$ and $V_3^{max} (2)$ have an average degree of identifiability of 75\% in comparison to the 15\% identifiability of both $K_3^{fdp} (1)$ and $K_3^{fdp} (2)$.
+This variability in the degree of identifiability for various parameters modeling the same flux is also seen for $v_3$ (Figure \ref{fig:figure1}c and f): $V_3^{max} (1)$ and $V_3^{max} (2)$ have an average degree of 75\% in comparison to the 15\% identifiability for both $K_3^{fdp} (1)$ and $K_3^{fdp} (2)$.
 
@@
 
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/k1cat_no_noise_experiments_jan03_ab}}
+	\caption{The contribution of different experiments types used in combinations that can practically identify parameter $k_1^{cat}$ in $v_1$.
+Panels a) and b) represent the position of the experiment in $k = \{1, 2\}$}\label{fig:figure2}
+\end{figure}
+
 This informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending the position of the flux in the metabolic network with respect to the inputs and the outputs of the network.
 The dependency can be further elucidated using the closed form solutions obtained for parameters of $v_1$ in Section \ref{sec:example}.
-As seen in Section \ref{sec:example}, the identifiability expressions for both parameters in $v_1$ are dependent on changes in the both the uptake flux $v_1$ as well as the acetate concentration $ac$ in both the experiments used for identification.
+The identifiability expressions for both parameters in $v_1$ (Equation \ref{eq:v1_par}) are dependent on changes in the both the uptake flux $v_1$ as well as the acetate concentration $ac$ in both the experiments used for identification.
 If either of these values cannot be distinguished between the two experiments whose data is used for identification, then the corresponding combination of experiments can be guaranteed to result in practical non-identifiability of the parameters in $v_1$.
@@
 In the small model that we use to elucidate our methodology, we find that only
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/k1cat_no_noise_experiments_jan03}}
-	\caption{The contribution of different experiments types used in combinations that can practically identify parameter $k_1^{cat}$ in $v_1$.
-Panels a) and b) represent the position of the experiment in $k = \{1, 2\}$}\label{fig:figure2}
-\end{figure}
 

### 2018-01-12 ident/python2/ss-ident/nonlin_solve.py
sole kotte model as nonlinear algebraic equations

@@ -1,0 +1,53 @@
+import numpy as np
+import scipy.optimize as opt
+from kotte_model import kotte_ck_flux
+import matplotlib.pyplot as plt
+
+
+def kotte_nlae(y, p):
+    K1ac, K3fdp, L3fdp, K3pep, K2pep, vemax, Kefdp, ne, d, V4max, k1cat, V3max, V2max, ac = p
+
+    flux_1 = k1cat * y[2] * ac / (ac + K1ac)
+    flux_2 = vemax * (1 - 1 / (1 + (Kefdp / y[1]) ** ne))
+    # convenience kinetics for flux 3
+    fdp_sat = y[1] / K3fdp
+    pep_sat = y[0] / K3pep
+    nr_3 = V3max * fdp_sat
+    dr_3 = 1 + fdp_sat
+    regulation_activate = 1 / (1 + 1 / pep_sat)
+    # regulation_inhibition = 1/(1 + pep_sat) # for future reference
+    flux_3 = regulation_activate * nr_3 / dr_3
+    flux_4 = V2max * y[0] / (y[0] + K2pep)
+    flux_5 = V4max * y[0]
+    flux_6 = d * y[2]
+    flux = np.hstack((flux_1, flux_2, flux_3, flux_4, flux_5, flux_6))
+    # flux = kotte_ck_flux(y, p)
+
+    yd_pep = flux[0] - flux[3] - flux[4]
+    yd_fdp = flux[3] - flux[2]
+    yd_e = flux[1] - flux[5]
+
+    return np.hstack((yd_pep, yd_fdp, yd_e))
+
+
+def solve_nlae(x0, model_parameters):
+    x_sol, info = opt.fsolve(kotte_nlae, x0, model_parameters, xtol=1e-6)
+    return x_sol
+
+
+def parameter_exp(flux, x, ac):
+    v1max = flux[0, 0] * flux[0, 1] * (ac[0] - ac[1])/(-(ac[1] * flux[0, 0] - ac[0] * flux[0, 1]))
+    # k1ac_no_enzyme_numerator_value = ac1 * (ac2 * v11 - ac2 * v12)
+    # k1ac_no_enzyme_denominator_value = -ac2 * v11 + ac1 * v12
+    # k1ac
+    return v1max
+
+
+if __name__ == '__main__':
+    # initial value of concentrations
+    y0 = np.array([5, 1, 1])
+    # initial parameter values
[…]

### 2018-01-12 ident/docs/ident_manuscript.tex
chnages to results - added new section(s)

@@
 We encounter this scenario when expression for the the allosterically activated reaction $v_3$ is described by the MWC model for allosteric regulation.
-We found that the complexity of the MWC kinetic rate law precluded its use to determine closed form expressions for the parameters in $v_3$ using either Mathematica or SymPy in Python.
-In order to overcome this computational obstacle, we modeled the reaction rate for $v_3$ using the convenience kinetic rate law formulation (reference for convenience kinetics).
+We find that the complexity of the MWC kinetic rate law precludes its use to determine closed form expressions for the parameters in $v_3$ using either Mathematica or SymPy in Python.
+In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation (reference for convenience kinetics).
 The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
@@
 
-Using this expression for identifiability analysis, we found that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different close form expressions owing to the presence of a square root term in their solutions.
+Using this expression for identifiability analysis, we find that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different close form expressions owing to the presence of a square root term in their solutions.
 These distinct expressions are denoted by (1) and (2) following the respective parameter names throughout the rest of the document.
 
-We believe this problem of computational tractability to occur in other complex kinetic rate law formulations as well. Example?
+We believe this problem of computational tractability will occur in other complex kinetic rate law formulations as well. Example?
 
-\subsection{Variability in the degree of identifiability of each parameter:}
-We use the percentage of data combinations that can identify a parameter as a naive measure of the degree of identifiability of that parameter.
-As an example, if 90\% of the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
+\subsection{Maximum reaction rates are more identifiable than enzyme binding constants:}
+We use the percentage of data combinations that can identify a parameter as a simple measure of the degree of identifiability of that parameter.
+As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
 On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%, and $\theta_i$ has a higher degree of identifiability than $\theta_j$.
 We can use this criteria to distinguish parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
+
 \begin{figure}[!tbhp]
@@
 In the presence of additive noise in the flux and concentration data used for identifiability, the number of data sets practically identifying parameters in fluxes d) $v_1$, e) $v_2$ and f) $v_3$ are shown.
-25 samples of additive noise $\epsilon ~ N(0, 0.05)$ were used to generate noisy data for figures in the right hand side panel.}\label{fig:figure1}
+25 samples of additive noise $\epsilon ~ N(0, 0.05)$ were used to generate noisy data for figures in the right hand side panel.
+The average and standard deviation values shown for identifiability of parameters with data in the presence of noise were calculated on the basis of the number of data combinations that can identify the respective parameter for each of the 25 noisy data samples.}\label{fig:figure1}
 \end{figure}
@@
 The value used for standard deviation reflects a 5\% noise level.
+The average and standard deviation values shown for identifiability of parameters with data in the presence of noise were calculated on the basis of the number of data combinations that can identify the respective parameter for each of the 25 noisy data samples.
 As seen in these figures, irrespective of the presence or absence of noise in the data, the number of combinations that can identify each parameter, and concordantly, the degree of identifiability of each parameter varies widely.
-Next, we look at trends in the degree of identifiability of each parameter in each flux.
+Next, we look at some of the trends in the degree of identifiability of each parameter in each flux.
 
 Since both fluxes $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law, we compare the number of data combinations that are capable of identifying the different parameters used to model these fluxes.
-The number of data combinations identifying the maximum reaction rates ($V_i^{max}$) for both $v_1$ (Figure \ref{fig:figure1}a and d) and $v_2$ (Figure \ref{fig:figure1}b and e) is always greater for than the number of data combinations that identify the enzyme binding ($K_i$) constants in the presence as well as absence of noise in the data.
+The number of data combinations identifying the maximum reaction rates ($V_i^{max}$) for both $v_1$ (Figure \ref{fig:figure1}a and d) and $v_2$ (Figure \ref{fig:figure1}b and e) is always greater than the number of data combinations that identify the enzyme binding ($K_i$) constants.
+This trend is observed for data combinations with and without noise.
 Even for parameters in $v_3$ the percentage of data combinations identifying $V_3^{max} (1)$ and $V_3^{max}(2)$ are greater than the percentage of combinations that can identify $K_3^{fdp}(1)$ and $K_3^{fdp}(2)$ (Figure \ref{fig:figure1}c and f).
@@
 The difference between the degrees of identifiability of each parameter within a single flux varies widely depending on the reaction flux.
-For instance, the difference between the number of combinations that can identify the parameters used to model $v_1$ is less than 2\% (Figure \ref{fig:figure1}a and d).
+For instance, the difference between the percentage of combinations that can identify the parameters used to model $v_1$ is less than 2\% (Figure \ref{fig:figure1}a and d).
 In contrast, a high number of combinations (89.5\%) can identify $V_2^{max}$ in $v_2$, while less than 2\% of the combinations are able to identify $K_2^{pep}$ (Figure \ref{fig:figure1}b and e).
 This variability in the degree of identifiability for various parameters modeling the same flux is also seen for $v_3$ (Figure \ref{fig:figure1}c and f): $V_3^{max} (1)$ and $V_3^{max} (2)$ have an average degree of 75\% in comparison to the 15\% identifiability for both $K_3^{fdp} (1)$ and $K_3^{fdp} (2)$.
+Again, this observation is common for both data with and without added noise.
 
 Also notable is the difference in the degree of identifiability between the maximum reaction rates of the uptake flux $v_1$ and the intracellular fluxes $v_2$ and $v_3$.
-Only about 20\% of the 4896 different data combinations can identify either the maximum reaction rate ($V_1^{max}$) or the enzyme turnover rate ($k_1^{cat}$) in $v_1$ (Figure \ref{fig:figure1}a and d), in contrast to the high number of data combinations capable of identifying $V_2^{max}$ (89.5\%) (Figure \ref{fig:figure1}b and e), $V_3^{max} (1)$ or $V_3^{max} (2)$ (>75\%) (Figure \ref{fig:figure1}c and f).
+We discuss this difference and its implications for experimental design in the following section.
 
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/k1cat_no_noise_experiments_jan03_ab}}
-	\caption{The contribution of different experiments types used in combinations that can practically identify parameter $k_1^{cat}$ in $v_1$.
-Panels a) and b) represent the position of the experiment in $k = \{1, 2\}$}\label{fig:figure2}
-\end{figure}
+\subsection{Experiments required for identifying parameters depends on the position of the flux in the metabolic network:}
+Only about 20\% of the 4896 different data combinations can identify either the maximum reaction rate ($V_1^{max}$) or the enzyme turnover rate ($k_1^{cat}$) in $v_1$ (Figure \ref{fig:figure1}a and d), in contrast to the high number of data combinations capable of identifying $V_2^{max}$ (89.5\%) (Figure \ref{fig:figure1}b and e), $V_3^{max} (1)$ or $V_3^{max} (2)$ ($>$75\%) (Figure \ref{fig:figure1}c and f).
 
@@
 The dependency can be further elucidated using the closed form solutions obtained for parameters of $v_1$ in Section \ref{sec:example}.
-The identifiability expressions for both parameters in $v_1$ (Equation \ref{eq:v1_par}) are dependent on changes in the both the uptake flux $v_1$ as well as the acetate concentration $ac$ in both the experiments used for identification.
+The identifiability expressions for both parameters in $v_1$ ($V_1^{max}$ and $K_1^{ac}$ in Equation \ref{eq:v1_par}) are dependent on changes in the both the uptake flux $v_1$ as well as the acetate concentration $ac$ in both the experiments used for identification.
 If either of these values cannot be distinguished between the two experiments whose data is used for identification, then the corresponding combination of experiments can be guaranteed to result in practical non-identifiability of the parameters in $v_1$.
@@
 
-Both in the presence as well as absence of noise, the identifiability of parameters in $v_1$ is guaranteed by only about 20\% of the available data combinations (Figure \ref{fig:figure1}a, d).
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/v1max_k1cat_all_experiments_jan01}}
+	\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter $V_1^{max}$ a) experiment 1 ($k=1$), and b) experiment 2 ($k=2$) in combination.
+Different experiments that contribute towards identifiability of $k_1^{cat}$ c) experiment 1 ($k=1$) and d) experiment 2 ($k=2$) in combination.}\label{fig:figure2}
+\end{figure}
 
- This is further illustrated in Figure below where the different experiments in each combination are shown.
+ This is further illustrated in Figure \ref{fig:figure2} where the different experiments in each combination identifying both $V_1^{max}$ and $k_1^{cat}$ are shown.
 

### 2018-01-13 ident/docs/ident_manuscript.tex
additions to sec 3.1

@@
 
-We believe this problem of computational tractability will occur in other complex kinetic rate law formulations as well. Example?
+We believe that this is not a unique problem isolated to the system that we are investigating.
+Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms (Chassagnole, other yeast and e.
+coli kinetic modeling papers).
+However, the identifiability of the parameters used in these models has never been truly examined.
+Any test for identifiability is forgone in favor of direct parameter estimation using a nonlinear least squares optimization approach with powerful computers to overcome difficulties associated with getting global optimal solutions ().
+This is however paving way for probabilistic approaches based on Bayesian techniques (Lars Nielsen papers).
+Furthermore none of these methods address the issue of designing experiments to suit parameter estimation.
+Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data.
+
+We believe this problem of computational tractability will occur in other complex kinetic rate law formulations as well. Example(s)?
 

### 2018-01-16 ident/python2/ss-ident/no_noise_script.py
pass number of parameters of each flux through processing fun call

@@ -40,19 +40,22 @@ experimental_datasets_3_expts = \
 ident_details_3 = flux_ident_3_data_combination(experimental_datasets_3_expts, choose=choose_3)
 
 # identifiability for all kotte fluxes
-ident_details = establish_kotte_flux_identifiability(experimental_datasets, choose=choose)
+# ident_details = establish_kotte_flux_identifiability(experimental_datasets, choose=choose)
 print('Perturbation analysis for identifiability complete.\n')
 
 # data processing
-data_list, original_data_ident, combo_data_ident, max_parameter = process_info_sample(ident_details,
-                                                                                      experimental_datasets,
+data_list, original_data_ident, combo_data_ident, max_parameter = process_info_sample(ident_details_2,
+                                                                                      experimental_datasets_2_expts,
                                                                                       perturbation_details)
 
+number_of_parameters_per_flux_2 = [4, 2]
+number_of_parameters_per_flux_3 = [6]
 # plot results
 # file_destination = 'C:\Users\shyam\Documents\Courses\CHE1125Project\Results\ident\python2\\figure_1'
 # plot parameters for each flux and the number of data sets that enable their identification
 # get data for plots
-total_ident_data, fraction_ident_data, all_boolean_p_id = parameter_plot_data_per_sample(original_data_ident, 1)
+total_ident_data, fraction_ident_data, all_boolean_p_id = \
+    parameter_plot_data_per_sample(original_data_ident, number_of_parameters_per_flux_2, 1)
 # plot
 # flux_parameter_plot(total_ident_data, fraction_ident_data)
 

### 2018-01-16 ident/docs/ident_manuscript.tex
more additions to section 3.3

@@
 If either of these values cannot be distinguished between the two experiments whose data is used for identification, then the corresponding combination of experiments can be guaranteed to result in practical non-identifiability of the parameters in $v_1$.
+Furthermore, the distinct values should satisfy the requirements for practical identifiability mentioned earlier for each parameter (Figure \ref{fig:ident-flowchart}b).
+
 As the uptake flux for the metabolic network that is modeled to be dependent on the substrate concentration, the only way meaningful changes in $v_1$ can be observed are through experiments involving changes either in the substrate concentration or changes in the enzyme concentration for $v_1$.
-In contrast, any experiment that can distinguish between different values of $pep$ and $v_2$ can be used to identify $V_2^{max}$.
+As the change in $v_1$ is directly proportional to changes in acetate concentrations, one can except the conditions for practical identifiability of parameters of $v_1$ (Equation \ref{eq:v1_par}) to be satisfied with relative certainty when experimental data gathered under different acetate concentrations is used for identification.
+This is further illustrated in Figure \ref{fig:figure2} where the different experiments in each combination identifying both $V_1^{max}$ and $k_1^{cat}$ are shown.
+
+However, in certain cases unlike $v_1$, satisfaction of this conditions may not be as easy as one might except due to the nonlinearity in the relationship between different parts of the network.
+
+We see that experiments involving changes to the substrate concentration figure prominently in the data combinations (as experiment 1) used to identify $V_1^{max}$ or $k_1^{cat}$, and the corresponding enzyme binding constant $K_1^{ac}$ (see Supplementary Figure S1).
+Experiments involving perturbations to other parameters appear in higher percentages as experiment 2 in the combination identifying these parameters.
+This can be attributed to the level of change in both $v_1$ as well a
 
@@
 
- This is further illustrated in Figure \ref{fig:figure2} where the different experiments in each combination identifying both $V_1^{max}$ and $k_1^{cat}$ are shown.
+In contrast, any experiment that can distinguish between different values of $pep$ and $v_2$ can be used to identify $V_2^{max}$.
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/v2max_k2pep_no_noise_experiments_jan03_ab}}
+	\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter $V_2^{max}$ a) experiment 1 ($k=1$), and b) experiment 2 ($k=2$) in combination.
+Different experiments that contribute towards identifiability of $K_2^{pep}$ c) experiment 1 ($k=1$) and d) experiment 2 ($k=2$) in combination.}\label{fig:figure3}
+\end{figure}
 

### 2018-01-21 ident/python2/ss-ident/noise_script.py
replicate code from no_noise_script for noisy data - change noise parameter to 1

@@ -1,13 +1,12 @@
 import numpy as np
 from generate_expdata import generate_expdata
 from simulate_data import arrange_experimental_data
-from kotte_model import establish_kotte_flux_identifiability
+from kotte_model import flux_ident_2_data_combination
+from kotte_model import flux_ident_3_data_combination
 from process_ident_data import process_info_sample
-from process_ident_data import parameter_plot_data_per_sample
-from process_ident_data import experiments_per_sample_for_ident
-from process_ident_data import experiment_position_based_info_per_sample
-from plot_ident_results import flux_parameter_plot
-from plot_ident_results import parameter_experiment_type_plot
+from plot_ident_results import data_utility_plot
+from plot_ident_results import parameter_identifibaility_plot
+from plot_ident_results import parameter_experiment_info_plot
 
 
 # generate noisy experimental data for testing identifiability
@@ -22,30 +21,54 @@ exp_xss, exp_fss, exp_ssid, perturbation_details = \
 
 # arrange experimental data to form multiple data sets
 exp_flux_index = np.array([0, 3, 2, 4])
-# choose numbr of experimental datasets for which identifiability is to be calculated
-choose = range(0, 200)
+
+# get combination of 2 experiments and perform identifiability on all fluxes that require 2 data sets
+print('Practical Identifiability Analysis of fluxes with 2 parameters \n')
+choose_2 = range(0, 306) # choose numbr of experimental datasets to use of analysis
 # get combinations of experimental datasets
-experimental_datasets = arrange_experimental_data(exp_xss, exp_fss, perturbation_details, 3, exp_flux_index, choose)
-
-# identifiability for all kotte fluxes
-ident_details = establish_kotte_flux_identifiability(experimental_datasets, choose=choose)
-print('Perturbation analysis for identifiability complete.\n')
-
+experimental_datasets_2_expts = \
+    arrange_experimental_data(exp_xss, exp_fss, perturbation_details, 2, exp_flux_index, choose_2)
+ident_details_2 = flux_ident_2_data_combination(experimental_datasets_2_expts, choose=choose_2, flux_ids=[1, 2])
+print('Identifiability analysis for fluxes with 2 parameters complete.\n')
 # data processing
-data_list, original_data_ident, combo_data_ident, max_parameter = process_info_sample(ident_details,
-                                                                                      experimental_datasets,
-                                                                                      perturbation_details)
-# get data for plots 
[…]

### 2018-01-21 ident/python2/ss-ident/noise_script.py
inc number of samples to 50

@@ -17,7 +17,7 @@ ode_parameter_values = np.array([.1, .1, 4e6, .1, .3, 1.1, .45, 2, .25, .2, 1, 1
 
 # get experimental system steady state data without noise
 exp_xss, exp_fss, exp_ssid, perturbation_details = \
-    generate_expdata(y0, cvode_options, ode_parameter_values, noise=1, number_of_samples=2)
+    generate_expdata(y0, cvode_options, ode_parameter_values, noise=1, number_of_samples=50)
 
 # arrange experimental data to form multiple data sets
 exp_flux_index = np.array([0, 3, 2, 4])

### 2018-01-21 ident/docs/ident_manuscript.tex
no major change
small changes to results in sec 1 and 2

@@
 \section{Methods:}	
+\subsection{A method for practical identifiability of kinetic models of metabolism:}\label{sec:ident}
+In this section, we show how practical identifiability of kinetic parameters in a dynamic model of metabolism can be established.
+A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
+In a kinetic model, the value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
+Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations.
+The fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
+
 \begin{figure}[!tbhp]
@@
 \end{figure}
-\subsection{A method for practical identifiability of kinetic models of metabolism:}\label{sec:ident}
-In this section, we show how practical identifiability of kinetic parameters in a dynamic model of metabolism can be established.
-A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
-In a kinetic model, the value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
-Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations.
-The fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
 
@@
 First in Section \ref{sec:closedform}, we discuss the ability to apply our methods to different nonlinear kinetic rate law formulations within the context of the gluconeogenic model.
-%Then, in Section we discuss the different experimental design strategies that were found through our identifiability analysis to enable kinetic parameter estimation for the each flux in the model using steady state data.
+Then, in Section \ref{sec:trends} we discuss some of the trends we observed in the identifiability of different types of parameters using in silico experimental data from different parameter perturbations.%Then, in Section we discuss the different experimental design strategies that were found through our identifiability analysis to enable kinetic parameter estimation for the each flux in the model using steady state data.
 Finally, in Section we expand the application of our method to a relatively large (39 reactions) kinetic model of the red blood cell metabolism.
 \subsection{Getting closed-form expressions for each flux in Kotte model:}\label{sec:closedform}
-As demonstrated for identifying parameters of $v_1$ earlier in Section \ref{sec:example}, the computational effort required to establish identifiability is governed by the nonlinear complexity of the enzyme kinetic rate law used to model a specific flux.
-Although computer algebra systems are capable of handling complex symbolic calculations, sometimes, the complexity of getting closed form expressions for all kinetic parameters of certain rate law formulations is too much for the computer algebra system to overcome.
-We encounter this scenario when expression for the the allosterically activated reaction $v_3$ is described by the MWC model for allosteric regulation.
-We find that the complexity of the MWC kinetic rate law precludes its use to determine closed form expressions for the parameters in $v_3$ using either Mathematica or SymPy in Python.
+Establishing parameter identifiability for fluxes in kinetic models of metabolism using the methodology we describe in the Methods section is governed by the nonlinear complexity of the enzyme kinetic rate law used to model a specific flux.
+We demonstrate one example of how the methodology works in Section \ref{sec:example} for parameters of flux $v_1$.
+The expression in Equation (\ref{eq:v1_par}) is obtained by using a computer algebra system.
+To recall, in this specific case, a computer algebra system is used to solve for the parameters of a flux described using the Michaelis-Menten kinetic rate law when experimental data from two different experiments is available.
+
+However, we find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica and SymPy in Python).
 In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation (reference for convenience kinetics).
 The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
+
 \begin{align}\label{eq:flux3_convkin}
@@
 Using this expression for identifiability analysis, we find that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different close form expressions owing to the presence of a square root term in their solutions.
-These distinct expressions are denoted by (1) and (2) following the respective parameter names throughout the rest of the document.
+These distinct expressions are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, $K_3^{pep}(1)$, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$.
 
-We believe that this is not a unique problem isolated to the system that we are investigating.
+This may not be a unique problem isolated to the system that we are investigating.
 Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms (Chassagnole, other yeast and e.
@@
 However, the identifiability of the parameters used in these models has never been truly examined.
-Any test for identifiability is forgone in favor of direct parameter estimation using a nonlinear least squares optimization approach with powerful computers to overcome difficulties associated with getting global optimal solutions ().
+If these models are tractable, we believe our methodology can help in elucidating the identifiability of parameters in these models.
+If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation.
+
+%Any test for identifiability is forgone in favor of direct parameter estimation using a nonlinear least squares optimization approach with powerful computers to overcome difficulties associated with getting global optimal solutions ().
 This is however paving way for probabilistic approaches based on Bayesian techniques (Lars Nielsen papers).
-Furthermore none of these methods address the issue of designing experiments to suit parameter estimation.
+Furthermore, none of these methods address the issue of designing experiments to suit parameter estimation.
 Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data.
 
-We believe this problem of computational tractability will occur in other complex kinetic rate law formulations as well. Example(s)?
-
-\subsection{Maximum reaction rates are more identifiable than enzyme binding constants:}
+\subsection{Maximum reaction rates are more identifiable than enzyme binding constants:\label{sec:trends}
 We use the percentage of data combinations that can identify a parameter as a simple measure of the degree of identifiability of that parameter.

### 2018-01-23 ident/docs/ident_manuscript.tex
complete results and not so good discussion with new erfs

@@
 \begin{document}
-\paragraph{Outline:}
-\begin{itemize}
-	\item Importance of parameter values for estimating in vivo response using kinetic models of metabolism and consequently for metabolic network design using kinetic models of metabolism
-	\item The need for parameter identifiability to determine unique and true parameter values from observed data
-	\item Types and purpose of identifiability for parameters
-	\item Methods for structural identifiability and existing methods for practical identifiability
-	\item Lack of methods for practical identifiability and consequently experimental design (not covered?)
-	\item work done in this paper for practical identifiability			
-	\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable) (move to discussion - may be in discussion)
-\end{itemize}
 \section{Introduction:}
@@
 coli} and demonstrating our ability to propose experiments that will facilitate parameter estimation for a kinetic model of this network.
-We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
+%We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
 
@@
 This also applies for identifying parameters in $v_2$.
+Hence, multiple data sets generated using a combination of any two experiments are used to test the identifiability of parameters for $v_1$ and $v_2$.
+The total number of such possible combinations is 306 (18 x 17) from the 18 different experiments mentioned earlier.
+A similar calculation for identifying the three parameters of $v_3$, where $p=3$, leads to 4896 distinct data combinations (18 x 17 x 16) of three experiments.
+% 18 different experiments mentioned earlier and elaborated in the Appendix.
 
-Looking at the three fluxes whose parameters have to be identified, $v_3$ has the maximum number of parameters at three.
-Hence, we choose to select multiple distinct data sets, each with three different experiments to establish the practical identifiability of all the parameters in the model.
-Accordingly, we can select 4896 (18 x 17 x 16) distinct combinations of three experiments from the 18 different experiments mentioned earlier and elaborated in the Appendix.
+\subsection{Degee of Identifiability, a quantitative measure of practical identifiability of parameters:}\label{sec:degree_of_identifiability}
+We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
+We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test practical identifiability) that can identify that parameter.
+As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
+On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
+Furthermore, we can create a hierarchy of practically identifiable parameters using their degree of identifiabilities.
+In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiabilities of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
+Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments, and design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network.
 
-\subsection{Experimental design through practical parameter identifiability for kinetic models of metabolism:}
+\subsection{Experimental design through practical parameter identifiability for kinetic models of metabolism:}\label{sec:experimental_design}
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure3/experimental_design}}
+	\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identifiability of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
+\end{figure}
 Following the methodology described in Section \ref{sec:ident}, and demonstrated in Section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found from Section \ref{sec:experiments} can be tested for their ability to practically identify parameters for any of the three fluxes in the small metabolic network.
 This step would help distinguish identifiable experiment combinations from combinations that do not practically identify any parameter in the model (Figure).
-Subsequently, it is possible to obtain a collection of experiments that make up all identifiable data combinations that can be performed to obtain the most minimal and informative set of experiments to identify as many parameters as possible.
+Subsequently, it is possible to obtain a collection of experiments that make up all identifiable data combinations that can be performed to obtain the most minimal and informative set of experiments to identify as many parameters as possible (Figure \ref{fig:ident-design}).
 Consequently, the set of experiments can be used to estimate the identifiable parameters in the model.
+We explain this formally below.
+
+The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in Section \ref{sec:ident} and demonstrated in Section \ref{sec:example}.
+Subsequently, for any flux $v_i$, if for any $p$ combinations of indices $j$, the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make a parameter $\theta_k$ identifiable.
+Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
+Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
+This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
+In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
+This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
+Hence, our method is theoretically scalable even to genome-scale models.
+
+Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability using our methodology.
 
 \section{Results:}
-First in Section \ref{sec:closedform}, we discuss the ability to apply our methods to different nonlinear kinetic rate law formulations within the context of the gluconeogenic model.
-Then, in Section \ref{sec:trends} we discuss some of the trends we observed in the identifiability of different types of parameters using in silico experimental data from different parameter perturbations.%Then, in Section we discuss the different experimental design strategies that were found through our identifiability analysis to enable kinetic parameter estimation for the each flux in the model using steady state data.
-Finally, in Section we expand the application of our method to a relatively large (39 reactions) kinetic model of the red blood cell metabolism.
-\subsection{Getting closed-form expressions for each flux in Kotte model:}\label{sec:closedform}
+First in Section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in Section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
+We use the gluconeogenic model described in Section \ref{sec:example} as an example.
+Then, in Section \ref{sec:trends} we show how the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetics rate law used to describe the flux.
+In Section \ref{sec:experiment_type} that follows, we discuss the the ability to determine experiments for parameter identifiability for a given flux based on the informativeness of a given experiment.
+In this section, we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identifiability of the parameters for a given flux.
+In Section \ref{sec:noise} we discuss the identifiability of parameters when data with additive noise is used with our methodology to test practical identifiability.
+Finally, in Section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models metabolic networks.
+
+\subsection{Nonlinearity of enzyme kinetic rate law expression affects identifiability analysis:}\label{sec:closedform}
 Establishing parameter identifiability for fluxes in kinetic models of metabolism using the methodology we describe in the Methods section is governed by the nonlinear complexity of the enzyme kinetic rate law used to model a specific flux.
@@
 However, we find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica and SymPy in Python).
-In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation (reference for convenience kinetics).
+In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
 The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
@@
 
-This may not be a unique problem isolated to the system that we are investigating.
-Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms (Chassagnole, other yeast and e.
-coli kinetic modeling papers).
+The impact of nonlinearity of a kinetic rate law expression affecting the practical identifiability of parameters in the expression may not be an unique problem isolated to the system that we are investigating.
+Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms \parencite{Chassagnole2002a, Peskov2012, VanHeerden2014}.
 However, the identifiability of the parameters used in these models has never been truly examined.
 If these models are tractable, we believe our methodology can help in elucidating the identifiability of parameters in these models.
-If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation.
+If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation and model validation.
 
-%Any test for identifiability is forgone in favor of direct parameter estimation using a nonlinear least squares optimization approach with powerful computers to overcome difficulties associated with getting global optimal solutions ().
-This is however paving way for probabilistic approaches based on Bayesian techniques (Lars Nielsen papers).
-Furthermore, none of these methods address the issue of designing experiments to suit parameter estimation.
-Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data.
+Next, we look some of the results we obtained for the practical identifiability of different parameters for all three fluxes $v_1$, $v_2$ and $v_3$ in the small gluconeogenic network model described in Section \ref{sec:small-model}, with the flux for $v_3$ described by Equation (\ref{eq:flux3_convkin}).
 
-\subsection{Maximum reaction rates are more identifiable than enzyme binding constants:\label{sec:trends}
-We use the percentage of data combinations that can identify a parameter as a simple measure of the degree of identifiability of that parameter.
-As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
-On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%, and $\theta_i$ has a higher degree of identifiability than $\theta_j$.
-We can use this criteria to distinguish parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
+\subsection{Maximum reaction rates are more identifiable than enzyme binding constants:}\label{sec:trends}
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure1/v1_v2_v3_identifiability_no_noise}}
+	\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data is shown.
+The percentage of total experimental data combinations that can identify each parameter is specified in/on the right hand side of the bar showing the total number of combinations identifying a given parameter.
+The total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower that the total number of combinations used to test identifiability of parameters for $v_3$ since both $v_1$ and $v_2$ have only two parameters while $v_3$ is modeled with three parameters.
+Hence, the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, while parameter identification in $v_3$ requires data from at least three different experiments.}\label{fig:figure1}
+\end{figure}
+In Figure \ref{fig:figure1} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
+Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
+The three panels in Figure \ref{fig:figure1} represent the degree of identifiability for parameters modeling the three different fluxes of the small network individually.
+Looking at these panels for $v_1$ (Figure \ref{fig:figure1}a), $v_2$ (Figure \ref{fig:figure1}b) and $v_3$ (Figure \ref{fig:figure1}c), the degree of identifiability of the maximum reaction rates ($V_i^{max}$) in each of the three fluxes is higher than the degree of identifiability of the enzyme binding ($K_i$) constants and the allosteric activation constant ($K_3^{pep}$).
+We observe this trend irrespective of the fact that only  $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law (Michaelis-Menten), while the convenience kinetic rate law is used to model $v_3$.
+These observations lead us to conclude that the maximum reaction rate parameters are always more identifiable (as indicated by the higher degree of identifiability) than their enzyme binding constant counterparts irrespective of the enzyme kinetic rate law used to model the corresponding flux.
+
+The most conspicuous of all observations from Figure \ref{fig:figure1} is the difference in the identifiability of parameters based on the flux in which these parameters are present.
+The variation in the degree of identifiability is not only dependent on the type of parameter being identified, but also depends on the  flux whose parameters are being tested for identifiability.
+This can be tied to selecting experimental data that have the ability to satisfy conditions for practical identifiability for a given parameter.
+We discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of the experiment from which data is collected for identifiability in the following section.
+
+\subsection{Experiments required for identifying parameters depends on the position of the flux in the metabolic network:}\label{sec:experiment_type}
+In system identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system.
+Only information obtained from such changes in the input and the corresponding signals can be used to completely identify the system over its entire dynamic range.
+Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
+In linear systems, persistence of excitation of any input signal is guaranteed.
+However, for nonlinear systems, like the metabolic network we deal with in this paper, persistence of excitation of input signals is not guaranteed and should be assessed on a case by case basis.
+The lack of persistence of excitation requires the design of experiments to satisfy the data needs for complete system identification.
+
+In this section, we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model.
+The informativeness of experiments for identifying parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter.
+For example, we demonstrated in Section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
+Thus, in this instance, the informativeness of the experiments (changes in the input acetate concentration) for identifying parameters of $v_1$ is determined by their ability to change the measured value of flux $v_1$.
+Similarly, the identification of parameters for $v_2$ requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}, given both these variables are outputs of the model describing $v_2$.
+
+In Figure \ref{fig:figure1}a we show that only about 20\% of all the experimental data sets satisfy these requirements for $v_1$ while more than 90\% of the available experimental satisfies these requires for $v_2$ (Figure \ref{fig:figure1}b).
+Thus, experimental design becomes crucial to determine the minimum number, as well as the nature of experiments that can help identify parameters for both $v_1$ and $v_2$.
+Recall that we use experimental data from five different experiments to test the practical identifiability of parameters in the model (Section \ref{sec:experiments}).
+In order to determine the type of experiments that help satisfy identifiability conditions for both parameters of $v_1$, we look at the distribution of the five different experiments in all data combinations that can identify these 2 parameters (Figure \ref{fig:figure2}).
 
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure1/noise_no_noise_kotte_ident_jan01}}
-	\caption{The number of data sets, made from a combination of 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
-In the presence of additive noise in the flux and concentration data used for identifiability, the number of data sets practically identifying parameters in fluxes d) $v_1$, e) $v_2$ and f) $v_3$ are shown.
-25 samples of additive noise $\epsilon ~ N(0, 0.05)$ were used to generate noisy data for figures in the right hand side panel.
-The average and standard deviation values shown for identifiability of parameters with data in the presence of noise were calculated on the basis of the number of data combinations that can identify the respective parameter for each of the 25 noisy data samples.}\label{fig:figure1}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/v1max_k1ac_experiments}}
+	\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter $V_1^{max}$ a) experiment 1 ($k=1$), and b) experiment 2 ($k=2$) in combination, and $K_1^{ac}$ c) experiment 1 ($k=1$), and d) experiment 2 ($k=2$) in combination.
+The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:figure2}
 \end{figure}
 
-In Figure \ref{fig:figure1} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
-In the left panels (Figure \ref{fig:figure1}a-c), the data is shown when noise-free data is used for identification.
-The identifiability information for parameters using data with additive noise is shown on the right panel (Figure \ref{fig:figure1}d-f).
-The data shown in the right hand side panel was generated using 25 different samples of random additive noise selected from a normal distribution with 0 mean 0.05 standard deviation.
-The value used for standard deviation reflects a 5\% noise level.
-The average and standard deviation values shown for identifiability of parameters with data in the presence of noise were calculated on the basis of the number of data combinations that can identify the respective parameter for each of the 25 noisy data samples.
-As seen in these figures, irrespective of the presence or absence of noise in the data, the number of combinations that can identify each parameter, and concordantly, the degree of identifiability of each parameter varies widely.
-Next, we look at some of the trends in the degree of identifiability of each parameter in each flux.
+The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$ contribute to a significant part (> 50\%) of the identifiable experimental data combinations in contrast to the other types of experiments.
+This matches with requirements for identifiability on the basis of the informativeness of experiments that was laid out earlier i.e., experiments must be able distinguish between changes in $v_1$ on the basis of changes in acetate.
+This informs why only about 20\% of the 306 different experiment combinations can identify these parameters (Figure \ref{fig:figure1}a).
+In can be gathered that only about 20\% of all the data combinations used to test identifiability are informative enough to distinguish between both acetate as well as $v_1$, and most of these experiments require changes in the acetate concentration to be able to identify $V_1^{max}$ and $K_1^{ac}$ through changes in the input and the output quantities acetate and $v_1$, respectively.
 
-Since both fluxes $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law, we compare the number of data combinations that are capable of identifying the different parameters used to model these fluxes.
-The number of data combinations identifying the maximum reaction rates ($V_i^{max}$) for both $v_1$ (Figure \ref{fig:figure1}a and d) and $v_2$ (Figure \ref{fig:figure1}b and e) is always greater than the number of data combinations that identify the enzyme binding ($K_i$) constants.
-This trend is observed for data combinations with and without noise.
-Even for parameters in $v_3$ the percentage of data combinations identifying $V_3^{max} (1)$ and $V_3^{max}(2)$ are greater than the percentage of combinations that can identify $K_3^{fdp}(1)$ and $K_3^{fdp}(2)$ (Figure \ref{fig:figure1}c and f).
-Hence, all maximum reaction rate parameters in the network have a higher degree of identifiability that their corresponding enzyme binding constants.
+In contrast to $v_1$, we see that the different perturbation experiments have a higher contribution towards experimental data combinations that can identify parameters for $v_2$ (Figure).
+Also, it is important to note that since most experiment types can bring about changes in the dynamics of $v_2$ (changes to \textit{pep} and value of $v_2$), the identifiability of parameters for $v_2$ is relatively higher than the identifiability of $v_1$ parameters (Figure \ref{fig:figure1}b).
+We can further extend these arguments to also justify the observed contribution of experiments towards identifying parameters for $v_3$ (figure not shown).
 
-The difference between the degrees of identifiability of each parameter within a single flux varies widely depending on the reaction flux.
-For instance, the difference between the percentage of combinations that can identify the parameters used to model $v_1$ is less than 2\% (Figure \ref{fig:figure1}a and d).
-In contrast, a high number of combinations (89.5\%) can identify $V_2^{max}$ in $v_2$, while less than 2\% of the combinations are able to identify $K_2^{pep}$ (Figure \ref{fig:figure1}b and e).
-This variability in the degree of identifiability for various parameters modeling the same flux is also seen for $v_3$ (Figure \ref{fig:figure1}c and f): $V_3^{max} (1)$ and $V_3^{max} (2)$ have an average degree of 75\% in comparison to the 15\% identifiability for both $K_3^{fdp} (1)$ and $K_3^{fdp} (2)$.
-Again, this observation is common for both data with and without added noise.
+Alternatively, both the degree of identifiability and the capability of different experiments to identify parameters of different fluxes based on their informativeness can be explained by the position of the flux in the metabolic network.
+We believe that the position of any given in the metabolic network determines the specific experiment that is persistently excitable enough to identify the parameters of that flux.
+This dependency of experiment informativeness on the position of the flux can be further elucidated using $v_1$ and $v_2$ as examples.
+We know from Equation (\ref{eq:v1_par}) that identifiability of $v_1$ requires changes in both acetate and $v_1$.
+We also know, based on our knowledge of the Michaelis-Menten kinetic rate law that changes in the substrate concentration of a reaction can bring about a nonlinear change in the value of the corresponding reaction rate (flux).
+In this specific metabolic network, since the substrate is an input variable to the model and $v_1$ is the corresponding uptake flux, the substrate can be easily perturbed to create persistently excitable experiments to identify parameters in $v_1$.
+We can generalize this observation for the identification of all uptake fluxes in all metabolic networks, i.e., at a minimum, a change in the input substrate concentration may be necessary for an informative experiment to identify the uptake flux parameters.
 
-Also notable is the difference in the degree of identifiability between the maximum reaction rates of the uptake flux $v_1$ and the intracellular fluxes $v_2$ and $v_3$.
-We discuss this difference and its implications for experimental design in the following section.
+On the other hand, the Michaelis-Menten model for $v_2$ also requires changes in \textit{pep} and $v_2$ for persistently excitable experiments to identify $v_2$.
+However, since both of these are system outputs, satisfaction of this condition cannot be guaranteed without an analysis of the dynamics of the metabolic network and how changes in the input (acetate) bring about changes in the two requisite output quantities.
+Previously, a dynamical analysis of the current network has shown that the concentration of \textit{pep}, and hence the flux $v_2$ are functions of not only the input acetate concentration, but also the levels of expression of the different enzymes within the network (Srinivasan et al., 2017).
+Hence, it is theoretically possible for any of the five different experiment types to be persistently excitable to identify $v_2$.
+This is confirmed by the high degree of identifiability for both parameters of $v_2$, where in more than 90\% of all data combinations can identify the parameters.
+Thus, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending the position of the flux in the metabolic network with respect to the inputs and the outputs of the network.
 
-\subsection{Experiments required for identifying parameters depends on the position of the flux in the metabolic network:}
-Only about 20\% of the 4896 different data combinations can identify either the maximum reaction rate ($V_1^{max}$) or the enzyme turnover rate ($k_1^{cat}$) in $v_1$ (Figure \ref{fig:figure1}a and d), in contrast to the high number of data combinations capable of identifying $V_2^{max}$ (89.5\%) (Figure \ref{fig:figure1}b and e), $V_3^{max} (1)$ or $V_3^{max} (2)$ ($>$75\%) (Figure \ref{fig:figure1}c and f).
-
-This informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending the position of the flux in the metabolic network with respect to the inputs and the outputs of the network.
-The dependency can be further elucidated using the closed form solutions obtained for parameters of $v_1$ in Section \ref{sec:example}.
-The identifiability expressions for both parameters in $v_1$ ($V_1^{max}$ and $K_1^{ac}$ in Equation \ref{eq:v1_par}) are dependent on changes in the both the uptake flux $v_1$ as well as the acetate concentration $ac$ in both the experiments used for identification.
-If either of these values cannot be distinguished between the two experiments whose data is used for identification, then the corresponding combination of experiments can be guaranteed to result in practical non-identifiability of the parameters in $v_1$.
-Furthermore, the distinct values should satisfy the requirements for practical identifiability mentioned earlier for each parameter (Figure \ref{fig:ident-flowchart}b).
-
-As the uptake flux for the metabolic network that is modeled to be dependent on the substrate concentration, the only way meaningful changes in $v_1$ can be observed are through experiments involving changes either in the substrate concentration or changes in the enzyme concentration for $v_1$.
-As the change in $v_1$ is directly proportional to changes in acetate concentrations, one can except the conditions for practical identifiability of parameters of $v_1$ (Equation \ref{eq:v1_par}) to be satisfied with relative certainty when experimental data gathered under different acetate concentrations is used for identification.
-This is further illustrated in Figure \ref{fig:figure2} where the different experiments in each combination identifying both $V_1^{max}$ and $k_1^{cat}$ are shown.
-
-However, in certain cases unlike $v_1$, satisfaction of this conditions may not be as easy as one might except due to the nonlinearity in the relationship between different parts of the network.
-
-We see that experiments involving changes to the substrate concentration figure prominently in the data combinations (as experiment 1) used to identify $V_1^{max}$ or $k_1^{cat}$, and the corresponding enzyme binding constant $K_1^{ac}$ (see Supplementary Figure S1).
-Experiments involving perturbations to other parameters appear in higher percentages as experiment 2 in the combination identifying these parameters.
-This can be attributed to the level of change in both $v_1$ as well a
+Nonetheless, all these explanations could only be verified due to the fact that the conditions for identifiability for both $v_1$ and $v_2$ are simple and readable due to the use of the Michaelis-Menten rate law to model these fluxes.
+Instead, in addition to presence of an allosteric interaction, when the convenience kinetic rate law is used to model this activation in $v_3$, we find that the complex expressions obtained for the three parameters preclude such an in-depth insight into the nature of persistently excitable experiments to identify  $v_3$.
+Thus, the only avenue through which experiments can be designed for the identification of parameters of $v_3$ would be through our proposed methodology wherein manual curation of experimental combinations required for identifiability are not necessary.
 
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/v1max_k1cat_all_experiments_jan01}}
-	\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter $V_1^{max}$ a) experiment 1 ($k=1$), and b) experiment 2 ($k=2$) in combination.
-Different experiments that contribute towards identifiability of $k_1^{cat}$ c) experiment 1 ($k=1$) and d) experiment 2 ($k=2$) in combination.}\label{fig:figure2}
-\end{figure}
-
-In contrast, any experiment that can distinguish between different values of $pep$ and $v_2$ can be used to identify $V_2^{max}$.
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/v2max_k2pep_no_noise_experiments_jan03_ab}}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/v2max_k2pep_experiments}}
 	\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter $V_2^{max}$ a) experiment 1 ($k=1$), and b) experiment 2 ($k=2$) in combination.
@@
 
-When the experimental data has no measurement noise associated with it, then $V_2^{max}$ is the only parameter that can be identified by the most number of experimental combinations (89.5\%).
-In contrast to the high identifiability of this parameter, $K_2^{pep}$, the other parameter for $v_2$ is identified by less than 2\% of the available data combinations.
+\subsection{Effect of noise on degree of identifiability is dependent on nonlinearity of enzyme kinetic rate law models:}\label{sec:noise}
+The degree of identifiability for all the parameters shown in Figure \ref{fig:figure1} is obtained using noise-free in silico experimental data.
+However, actual experimental data from biological systems is usually noisy.
+Data gathered from experiments is usually associated with both biological noise attributable to the stochasticity of cellular function, as well as measurement noise associated with the techniques used to obtain concentration and flux measurements.
+In order to test the practical identifiability of parameters under circumstances where the presence of noise in the experimental data can affect the identifiability of the parameters, we apply our methodology on data (concentration and flux) with 5\% additive noise.
+50 different samples of noise with 0 mean and 0.05 standard deviation are drawn from a normal distribution and added to the steady state experimental data generated on the basis of experiments described in Section \ref{sec:experiments} to generate 50 different samples of noisy data.
+Subsequently, all 50 samples are tested for their ability to practically identify parameters for all three fluxes of the small network.
+The corresponding graphs generated for practical identifiability analysis using noisy experimental data are shown in the Appendix.
 
-This is followed by $V_3^{max}$ that is capable of being identified by 76.3\% of the available data combinations.
+We hypothesize that the nonlinearity of the enzyme kinetic rate law used to model a flux, and consequently the complexity of the closed form expression obtained for the various parameters used in the model has an impact on the ability of noisy data to practically identify the parameters.
+The difference in the identifiability attributable to the aforementioned factors can be seen in the difference in the degree of identifiability for the parameters of $v_1$, $v_2$ and $v_3$ in the small metabolic network that we use to demonstrate our methodology.
 
-In the small model that we use to elucidate our methodology, we find that only
+Despite the presence of noise, the degree of identifiability of $v_1$ and $v_2$ (figure not shown) does not change from the case where no noise is present in the experimental data (Figure \ref{fig:figure1}).
+However, the degree of identifiability for parameters of $v_3$ is affected by the presence of noise in the data (Figure).
+This effect is represented by the presence of a non-zero standard deviation and corresponding error bars on the graphs in the degrees of identifiability of parameters modeling $v_3$ (Figure).
+The standard deviation and the error bars represent the presence of data combinations that can identify a given parameter under certain values of noise, but not do so under different noise values.
+In the presence of 5\% additive noise in the data, the conditions under which parameters for $v_1$ and $v_2$ can be identified are strictly satisfied by the same data combinations that can identify these parameters in the absence of noise, i.e., the changes in the acetate concentration (for $v_1$) and \textit{pep} (for $v_2$) are significant enough to enable identifiability.
+The contrinution of different experiment types towards identifiability also remains the same (figure not shown separately).
 
-\subsection{Utility of data for identifiability purposes:}
+However, as stated in the previous section, the nonlinearity of the original kinetic rate law expression for $v_3$, and the complexity of the closed-form expression obtained for the parameters in the rate law model preclude us from testing the presence, and consequently the satisfaction of these conditions for parameters in $v_3$.
+Furthermore, the conditions for identifiability of $v_3$ parameters are not as simple as either requiring just a simple difference in \textit{fdp} or \textit{pep} or that of both concentrations.
+Thus, while under some noise values these identifiability conditions may be satisfied, it may not be possible to satisfy these conditions under other noise values.
+Accordingly, we end up with a distribution of experimental data combinations that can identify $v_3$ parameters depending on noise levels in the data.
+It is significant to note that the variability in the degree of identifiability of parameters is small (< 2\%) when noise is present.
 
-\subsection{Experiments that can establish identifiability for each flux in Kotte model:}	
-As indicated previously in Section, we performed different experiments by perturbing the parameters for fluxes $v_1$, $v_2$ and $v_3$ in the model.
-We also perturbed the input substrate concentration to collect data on the concentrations and fluxes within the network.
-We describe the results of using this data to establish the identifiability of various parameters in the kinetic model of the small metabolic network.
+These observations, in conjunction with the estimation of the degree of identifiability of the different parameters (Figure \ref{fig:figure1}), helps in determining not only the nature of experiments required for parameter identifiability (Figure \ref{fig:figure2}), but also the order which the type of experiments necessary for identifying the entire metabolic network needs to be evaluated.
 
-\subsection{Experimental design for parameter estimation in kinetic models of metabolism:}
-The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in Section \ref{sec:ident} and demonstrated in Section \ref{sec:example}.
-Subsequently, for any flux $v_i$, if for any $p$ combinations of indices $j$, the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make a parameter $\theta_k$ identifiable.
-Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
-Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
-This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
-In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
-This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
-Hence, our method is theoretically scalable even to genome-scale models.
-We show the application of this design methodology for flux $v_2$ in the gluconeogenic model in Section.
+For instance, in the case of the example network that we use in this paper, given the high degree of identifiability for $v_2$ and large spread in the frequency of different experiment types that can contribute to identifiability for $v_2$, the focus should be on first choosing experiments for identifying $v_1$ followed by selecting experiments for identifying $v_3$.
+It is possible that the three or at most four experiments chosen to identify parameters of both $v_1$ and $v_3$ would suffice to identify $v_2$ as well without the need to perform additional experiments.
 
-Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability using our methodology.
+\subsection{Utility of experimental data for practical parameter identification:}\label{sec:data_utility}
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure4/v1_v2_v3_data_utility}}
+	\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
+The total possible number of parameters that can be identified by data from combinations of a) and b) two, and c) three experiments is shown in the horizontal axis.
+The vertical axis represents the total number of combinations that can identify the corresponding number of parameters.
+The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
+A total of 306 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 4896 combinations are used to practical identifiability of parameters in c)$v_3$.
+Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
+\end{figure}
 
-\subsection{Combinations of experiments that will enable identification of all model parameters:}
+Apart from determining experiments that are informative enough to identify parameters for each flux in a kinetic model, experimental design can also be used to design and develop a minimal set of experiments that informative enough to estimate all parameters in a metabolic network model.
+However, there is not only a scarcity of experimental design methods in literature devoted to the purpose of designing experiments for parameter identifiability in kinetic models of metabolism, the motivation to develop such methods is not well understood.
+Hence, we found the need to stress the necessity for experimental design methodologies tailored specifically for metabolic network models.
+We do this by showing how useful the multiple combinations of experimental data sets are towards identifying the maximum number of parameters for each given flux (Figure \ref{fig:figure4}).
 
-\subsection{Expanding methodology to RBC model:}
-If any data generated from a perturbation experiment $i$ results in nonidentifiability, we eliminate experiment $i$ from the list of experiments that need to be performed for parameter estimation.
+In Figure \ref{fig:figure4}, we show the percentage of all data combinations that can estimate a given number of parameters within a given flux.
+For flux $v_1$ (Figure \ref{fig:figure4}a) we see that about 78\% of all data combinations cannot identify a single parameter, while only about 18\% of the data combinations can identify all the four parameters (two parameters each under different model setups) used in the model.
+This directly correlates with the degree of identifiability of the various parameters of $v_1$, shown earlier in Figure \ref{fig:figure1}.
+
+In contrast, given the high degree of identifiability of both parameters for $v_2$, most data combinations (>90\%) are not wasted, and can identify both parameters (Figure \ref{fig:figure4}b).
+Furthermore, the data utility plots for $v_3$ (Figure \ref{fig:figure4}c) show that about 36\% of the 5000 data combinations can identify only two parameters while 6\% of the experiment combinations cannot identify any data set.
+Thus, Figure \ref{fig:figure4} provides evidence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
+Choosing any of the experiment combinations for the 78\% percent that does not identify any parameter in $v_1$, or the 6\% that does not identify any parameters in $v_3$ can lead to waste of resources used to perform these non-informative experiments.
+
+\section{Discussions:}
+Methods have been extensively developed for the establishing structurally identifiability of parameters in nonlinear biological model based on differential algebra techniques \parencite{Ljung1994}.
+However, very few methods exist for testing practical identifiability of parameters in nonlinear models and subsequently design experiments to estimate these parameters.
+One of the examples of a method for practical identifiability is the profile likelihood approach \parencite{Raue2009a}.
+However, methods that can overcome some of the difficulties associated with implementing this method for models of large metabolic network are sorely missing.
+Additionally, any test for identifiability is forgone in favor of direct parameter estimation using a nonlinear least squares optimization approach with powerful computers to overcome difficulties associated with getting global optimal solutions ().
+Deterministic parameter estimation techniques, based on optimization methods are paving way for probabilistic approaches based on Bayesian techniques (Lars Nielsen papers).
+The common thread that connects all these parameter estimation techniques is their failure to address the fundamental need for practically parameter identifiability before actual parameter estimation.
+
+Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data.
+Hence, these methods fail to address the issue of designing experiments to suit parameter estimation, and the need for ways to determine informative experiments remains.
+
+In this paper we have developed a methodology to test the practical identifiability of parameters in kinetic models of metabolism.
+Our methodology enables practical parameter identification for each individual flux within a metabolic network under the assumption that all steady state substrate, product and allosteric effector concentrations and fluxes for a given reaction can be measured.
+Using a small metabolic network for gluconeogenesis, we have demonstrated that the identifiability of parameters for a given flux is dependent on the position of the flux within the metabolic network.
+We have also shown the ability to use our analysis to design the minimal number of experiments that are most informative for identifying all fluxes within a metabolic network.
+
+Our work adds to this existing body of work wherein we develop a method for practical identifiability tailored for use with nonlinear enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks.
+With our work we hope to change the status quo in the application of systems identification techniques for kinetic models of metabolic networks.
+Our methodology fills the niche gap of experimental design for parameter estimation by providing a way to design informative experiments to obtain data required for parameter estimation by spending the least amount of resources.
+
+Based on the results we obtained in establishing the identifiability of parameters modeling all the fluxes of a small metabolic network, in the future we hope to extend this methodology to a larger metabolic network with more reactions.
 
@@
 	\begin{center}				
-		\begin{tabular}{ccc}
-			Designation & Perturbed Fluxes & Perturbed Values\\
+		\begin{tabular}{ccccccc}
 			\hline
-			P1 & $v_1$ & 2\\
-			P2 & $v_2$ & 0.2\\
-			P3 & $v_3$ & 0.5
+			Experiment Type & Perturbed Parameter & \multicolumn{5}{c}{Perturbed Values}\\
+			\hline
+			acetate & $ac$ & \multicolumn{5}{l}{0.1}\\
+			\hline
+			acetate perturbation & $ac$ & 0.5 & 1.0 & & &\\
+			\hline
+			$v_1$ perturbation & $k_1^{cat}$ & 1.1 & 1.5 & 2.0 & 0.9 & 0.5\\
+			\hline
+			$v_3$ perturbation & $V_3^{max}$ & 1.1 & 1.5 & 2.0 & 0.9 & 0.5\\
+			\hline
+			$v_2$ perturbation & $V_2^{max}$ & 1.1 & 1.5 & 2.0 & 0.9 & 0.5\\
 		\end{tabular}
@@
 %Thus, the identifiability of parameters in nonlinear models is dependent on the model structure, parameterization, and the quality and quantity of experimental data that is available for the purpose of estimation.
+
+%In the left panels (Figure \ref{fig:figure1}a-c), the data is shown when noise-free data is used for identification.
+The identifiability information for parameters using data with additive noise is shown on the right panel (Figure \ref{fig:figure1}d-f).
+The data shown in the right hand side panel was generated using 25 different samples of random additive noise selected from a normal distribution with 0 mean 0.05 standard deviation.
+The value used for standard deviation reflects a 5\% noise level.
+The average and standard deviation values shown for identifiability of parameters with data in the presence of noise were calculated on the basis of the number of data combinations that can identify the respective parameter for each of the 25 noisy data samples.
+As seen in these figures, irrespective of the presence or absence of noise in the data, the number of combinations that can identify each parameter, and concordantly, the degree of identifiability of each parameter varies widely.
+Next, we look at some of the trends in the degree of identifiability of each parameter in each flux.
+
+\paragraph{Outline:}
+\begin{itemize}
+	\item Importance of parameter values for estimating in vivo response using kinetic models of metabolism and consequently for metabolic network design using kinetic models of metabolism
+	\item The need for parameter identifiability to determine unique and true parameter values from observed data
+	\item Types and purpose of identifiability for parameters
+	\item Methods for structural identifiability and existing methods for practical identifiability
+	\item Lack of methods for practical identifiability and consequently experimental design (not covered?)
+	\item work done in this paper for practical identifiability			
+	\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable) (move to discussion - may be in discussion)
+\end{itemize}

### 2018-02-06 ident/docs/ident_manuscript.tex
copied text from committee meeting document

@@
 \section{Results:}
-First in Section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in Section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
-We use the gluconeogenic model described in Section \ref{sec:example} as an example.
-Then, in Section \ref{sec:trends} we show how the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetics rate law used to describe the flux.
-In Section \ref{sec:experiment_type} that follows, we discuss the the ability to determine experiments for parameter identifiability for a given flux based on the informativeness of a given experiment.
-In this section, we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identifiability of the parameters for a given flux.
-In Section \ref{sec:noise} we discuss the identifiability of parameters when data with additive noise is used with our methodology to test practical identifiability.
-Finally, in Section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models metabolic networks.
+First in section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
+We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
+Then, in section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models of metabolic networks.
+We do this by looking at the number of combinations that can identify the maximum number of parameters.
+In section \ref{sec:trends} we show that the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetic rate law used to describe the flux.
+In section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment.
+In this section we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux.
+Finally, we discuss some results arising out of the use of our methodology to determine the identifiability of parameters when data with additive noise is used in section \ref{sec:noise}.
 
-\subsection{Nonlinearity of enzyme kinetic rate law expression affects identifiability analysis:}\label{sec:closedform}
-Establishing parameter identifiability for fluxes in kinetic models of metabolism using the methodology we describe in the Methods section is governed by the nonlinear complexity of the enzyme kinetic rate law used to model a specific flux.
-We demonstrate one example of how the methodology works in Section \ref{sec:example} for parameters of flux $v_1$.
+\subsection{Nonlinearity of enzyme kinetic rate law expression affects identifiability analysis}\label{sec:closedform}
+Practical identifiability of parameters in kinetic models of metabolism using the methodology described in section \ref{sec:methods} is governed by the nonlinear complexity of the enzyme kinetic rate law used to model a specific flux.
+We demonstrate one example of how the methodology works in section \ref{sec:example} for parameters of flux $v_1$.
 The expression in Equation (\ref{eq:v1_par}) is obtained by using a computer algebra system.
-To recall, in this specific case, a computer algebra system is used to solve for the parameters of a flux described using the Michaelis-Menten kinetic rate law when experimental data from two different experiments is available.
+To recall, in this specific case, a computer algebra system is used to solve for the parameters of a flux described using the Michaelis-Menten kinetic rate law when data from two different experiments is available.
 
@@
 
-Using this expression for identifiability analysis, we find that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different close form expressions owing to the presence of a square root term in their solutions.
+Using this expression for analysis, we find that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different closed-form expressions owing to the presence of a square root term in their solutions.
 These distinct expressions are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, $K_3^{pep}(1)$, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$.
@@
 However, the identifiability of the parameters used in these models has never been truly examined.
-If these models are tractable, we believe our methodology can help in elucidating the identifiability of parameters in these models.
+Unlike the case of the MWC model, if determining closed-form expressions for the parameters of these models are tractable, we believe our methodology can help in elucidating the identifiability of these parameters.
 If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation and model validation.
 
-Next, we look some of the results we obtained for the practical identifiability of different parameters for all three fluxes $v_1$, $v_2$ and $v_3$ in the small gluconeogenic network model described in Section \ref{sec:small-model}, with the flux for $v_3$ described by Equation (\ref{eq:flux3_convkin}).
+Next, we look at some of the results we obtained for the practical identification of different parameters for all three fluxes $v_1$, $v_2$ and $v_3$ in the small gluconeogenic network model described in section \ref{sec:small-model}, with the flux for $v_3$ described by Equation (\ref{eq:flux3_convkin}).
+We start by looking at one of motivations for the need for designing experiments to collect data for parameter estimation.
 
-\subsection{Maximum reaction rates are more identifiable than enzyme binding constants:}\label{sec:trends}
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure1/v1_v2_v3_identifiability_no_noise}}
-	\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data is shown.
-The percentage of total experimental data combinations that can identify each parameter is specified in/on the right hand side of the bar showing the total number of combinations identifying a given parameter.
-The total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower that the total number of combinations used to test identifiability of parameters for $v_3$ since both $v_1$ and $v_2$ have only two parameters while $v_3$ is modeled with three parameters.
-Hence, the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, while parameter identification in $v_3$ requires data from at least three different experiments.}\label{fig:figure1}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure4/v1_V1max_v2_v3_value1_data_utility}}
+	\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
+The total possible number of parameters that can be identified by data from combinations of a) two, b) two, and c) three experiments is shown on the horizontal axis.
+The vertical axis represents the total number of combinations that can identify the corresponding number of parameters.
+The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
+A total of 306 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 4896 combinations are used to practical identifiability of c) $v_3$.
+Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
+\end{figure}	
+\subsection{Utility of experimental data for practical parameter identification}\label{sec:data_utility}
+Although the need for better parameter estimation methods is well researched, within the kinetic modeling community, the need for experimental design methods to satisfy the data needs for parameter estimation are not considered seriously.
+Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
+
+We do this by showing how useful the multiple combinations of experimental data are towards identifying parameters for each given flux (Figure \ref{fig:figure4}).
+We express the usefulness of any given data combination on the basis of the number of parameters that the combination can identify.
+The higher the number of parameters a data combination can identify, the greater is the usefulness of that combination of experiments.
+
+The number of data combinations that can identify a specific number of parameters within a given flux is shown in Figure \ref{fig:figure4}.
+The percentage of all data combinations that can estimate a given number of parameters within a given flux are also shown in Figure \ref{fig:figure4}.
+For flux $v_1$ (Figure \ref{fig:figure4}a) about 78\% of the 306 data combinations cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$).
+Only a single parameter can be identified by about 2\% of the combinations, while only about 19\% of the data combinations can identify both the parameters used in the model.
+In contrast, most data combinations (>90\%) are not wasted, and can identify both parameters in $v_2$ (Figure \ref{fig:figure4}b).
+Less than 5\% of the experimental data combinations are wasted from not being able to identify any parameter.
+In the case of $v_3$ (Figure \ref{fig:figure4}c), about 45\% of the 4896 data combinations can identify only one parameter, while 13\% of the combinations cannot identify any parameter.
+Only 10\% of combinations can identify $v_3$ completely.
+
+Thus, Figure \ref{fig:figure4} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
+For instance, choosing any of the experiment combinations from the 78\% percent that does not identify any parameter in $v_1$, or the 13\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative.
+This can lead to potential waste of resources used to perform these experiments.
+
+In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:figure4} also informs on the ease of identifying parameters for a single flux.
+For the small network (Figure \ref{fig:network}), we find that more than 90\% of all the data combinations tested for their ability to identify $v_2$ are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:figure4}b).
+In contrast, less than 20\% of the combinations can identify both parameters in $v_1$.
+Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ as well as $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$.
+A careful consideration of experiments for identifying $v_2$ is however not a major concern since there is a very high probability that experiments chosen to identify $v_1$ will also identify $v_2$.
+
+Although the calculation of the utility of experiment combinations in terms of their ability to identify the maximum number of parameters gives an idea on the informativeness of experiments used to gather the data, further analysis from the perspective of identifiability of the parameters of different fluxes is required to be able to design experiments based on this information.
+Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.
+
+\subsection{Maximum reaction rates are more identifiable than enzyme binding constants}\label{sec:trends}
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure1/v1_V1max_v2_v3_identifiability}}
+	\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
+The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
+Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:figure1}
 \end{figure}
 In Figure \ref{fig:figure1} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
-Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
+Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
 The three panels in Figure \ref{fig:figure1} represent the degree of identifiability for parameters modeling the three different fluxes of the small network individually.
-Looking at these panels for $v_1$ (Figure \ref{fig:figure1}a), $v_2$ (Figure \ref{fig:figure1}b) and $v_3$ (Figure \ref{fig:figure1}c), the degree of identifiability of the maximum reaction rates ($V_i^{max}$) in each of the three fluxes is higher than the degree of identifiability of the enzyme binding ($K_i$) constants and the allosteric activation constant ($K_3^{pep}$).
+
+In regards to the identifiability of $v_1$ (Figure \ref{fig:figure1}a), $v_2$ (Figure \ref{fig:figure1}b) and $v_3$ (Figure \ref{fig:figure1}c), the degree of identifiability of the maximum reaction rates ($V_i^{max}$) in each of the three fluxes is higher than the degree of identifiability of the corresponding enzyme binding ($K_i$) constants and the allosteric activation constant ($K_3^{pep}$).
 We observe this trend irrespective of the fact that only  $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law (Michaelis-Menten), while the convenience kinetic rate law is used to model $v_3$.
-These observations lead us to conclude that the maximum reaction rate parameters are always more identifiable (as indicated by the higher degree of identifiability) than their enzyme binding constant counterparts irrespective of the enzyme kinetic rate law used to model the corresponding flux.
+These observations lead us to conclude that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
 
-The most conspicuous of all observations from Figure \ref{fig:figure1} is the difference in the identifiability of parameters based on the flux in which these parameters are present.
-The variation in the degree of identifiability is not only dependent on the type of parameter being identified, but also depends on the  flux whose parameters are being tested for identifiability.
-This can be tied to selecting experimental data that have the ability to satisfy conditions for practical identifiability for a given parameter.
-We discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of the experiment from which data is collected for identifiability in the following section.
-
-\subsection{Experiments required for identifying parameters depends on the position of the flux in the metabolic network:}\label{sec:experiment_type}
-In system identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system.
-Only information obtained from such changes in the input and the corresponding signals can be used to completely identify the system over its entire dynamic range.
+From Figure \ref{fig:figure1} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified.
+In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
+Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range.
 Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
-In linear systems, persistence of excitation of any input signal is guaranteed.
-However, for nonlinear systems, like the metabolic network we deal with in this paper, persistence of excitation of input signals is not guaranteed and should be assessed on a case by case basis.
+In linear systems, persistence of excitation of any input signal can be theoretically guaranteed.
+However, for nonlinear systems, like the metabolic network we deal with in this paper, persistence of excitation of input signals cannot be theoretically guaranteed, and should be assessed on a case by case basis.
 The lack of persistence of excitation requires the design of experiments to satisfy the data needs for complete system identification.
 
+In the case of identifying fluxes within a metabolic network individually, each flux is not independent from the other fluxes that form the metabolic network.
+Each flux is connected to one or more fluxes through the metabolites that not only act as substrates and products, but also as allosteric activators and inhibitors (e.g., effect of \textit{pep} on $v_3$ in Figure \ref{fig:network}).
+Hence, careful consideration of experiments is a necessity so that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network.
+In the next section we discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of experiments from which data is collected for identification.
+
+\subsection{Experiment requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}	
 In this section, we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model.
-The informativeness of experiments for identifying parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter.
-For example, we demonstrated in Section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
-Thus, in this instance, the informativeness of the experiments (changes in the input acetate concentration) for identifying parameters of $v_1$ is determined by their ability to change the measured value of flux $v_1$.
-Similarly, the identification of parameters for $v_2$ requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}, given both these variables are outputs of the model describing $v_2$.
+Earlier, we mentioned that the informativeness of experiments for identifying parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter.
+For example, we demonstrated in section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them (Equation \ref{eq:v1_par}).
+Thus, in this instance, the informativeness of the experiments (changes in the input acetate concentration and $v_1$) for identifying parameters of $v_1$ is determined by the ability of the input change to effect a change in the measured value of $v_1$.
+Similarly, the identification of parameters for $v_2$ requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}.
+%, given both these variables are outputs of the model describing $v_2$.
 
-In Figure \ref{fig:figure1}a we show that only about 20\% of all the experimental data sets satisfy these requirements for $v_1$ while more than 90\% of the available experimental satisfies these requires for $v_2$ (Figure \ref{fig:figure1}b).
-Thus, experimental design becomes crucial to determine the minimum number, as well as the nature of experiments that can help identify parameters for both $v_1$ and $v_2$.
-Recall that we use experimental data from five different experiments to test the practical identifiability of parameters in the model (Section \ref{sec:experiments}).
-In order to determine the type of experiments that help satisfy identifiability conditions for both parameters of $v_1$, we look at the distribution of the five different experiments in all data combinations that can identify these 2 parameters (Figure \ref{fig:figure2}).
+As such, as per Figure \ref{fig:figure1}a only 20\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$.
+Whereas, more than 90\% of the available data combinations satisfy the requirements for identifying $v_2$ (Figure \ref{fig:figure1}b).
+Thus, identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for both $v_1$ and $v_2$.
+Recall that we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Section \ref{sec:experiments}).
+In order to determine the type of experiments that help satisfy identifiability conditions for both parameters of $v_1$, we look at the distribution of the five different types in all data combinations that can identify these 2 parameters (Figure \ref{fig:figure2}a and b).
 
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/v1max_k1ac_experiments}}
-	\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter $V_1^{max}$ a) experiment 1 ($k=1$), and b) experiment 2 ($k=2$) in combination, and $K_1^{ac}$ c) experiment 1 ($k=1$), and d) experiment 2 ($k=2$) in combination.
+	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_only_v2_spider}}
+	\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
 The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:figure2}
@@
 
-The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$ contribute to a significant part (> 50\%) of the identifiable experimental data combinations in contrast to the other types of experiments.
-This matches with requirements for identifiability on the basis of the informativeness of experiments that was laid out earlier i.e., experiments must be able distinguish between changes in $v_1$ on the basis of changes in acetate.
-This informs why only about 20\% of the 306 different experiment combinations can identify these parameters (Figure \ref{fig:figure1}a).
-In can be gathered that only about 20\% of all the data combinations used to test identifiability are informative enough to distinguish between both acetate as well as $v_1$, and most of these experiments require changes in the acetate concentration to be able to identify $V_1^{max}$ and $K_1^{ac}$ through changes in the input and the output quantities acetate and $v_1$, respectively.
+The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments.
+This matches with requirements for identifiability on the basis of the informativeness of experiments that was laid out earlier i.e., experiments must be able distinguish between different values of $v_1$ and acetate.
+Accordingly, only about 20\% of the 306 different experiment combinations used to test identifiability are informative enough to discriminate both acetate as well as $v_1$, and can identify $V_1^{max}$ and $K_1^{ac}(ne)$ (Figure \ref{fig:figure1}a).
 
-In contrast to $v_1$, we see that the different perturbation experiments have a higher contribution towards experimental data combinations that can identify parameters for $v_2$ (Figure).
-Also, it is important to note that since most experiment types can bring about changes in the dynamics of $v_2$ (changes to \textit{pep} and value of $v_2$), the identifiability of parameters for $v_2$ is relatively higher than the identifiability of $v_1$ parameters (Figure \ref{fig:figure1}b).
-We can further extend these arguments to also justify the observed contribution of experiments towards identifying parameters for $v_3$ (figure not shown).
+In contrast to $v_1$, we see that the different enzyme perturbation experiments have a higher contribution towards data combinations that can identify parameters for $v_2$ (Figure \ref{fig:figure2}c and d).
+Hence, in conjunction with the high degree of identifiability seen in Figure \ref{fig:figure1}b for $v_2$, we can deduce that almost all experiment types are persistently excitable for $v_2$ i.e., most experiments can bring about noticeable changes to both \textit{pep} and $v_2$.
+Consequently, in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative enough to identify $v_2$.
+This observations is in line with the one made in section \ref{sec:data_utility} based on the total number of combinations that can identify both parameters in $v_2$.
 
-Alternatively, both the degree of identifiability and the capability of different experiments to identify parameters of different fluxes based on their informativeness can be explained by the position of the flux in the metabolic network.
-We believe that the position of any given in the metabolic network determines the specific experiment that is persistently excitable enough to identify the parameters of that flux.
+Both the degree of identifiability of parameters and the informativeness of the corresponding experiments used to identify them can be explained by the position of the flux in the metabolic network.
+The position of any given flux in the metabolic network determines the specific experiment that is persistently excitable enough to identify the parameters of that flux.
 This dependency of experiment informativeness on the position of the flux can be further elucidated using $v_1$ and $v_2$ as examples.
 We know from Equation (\ref{eq:v1_par}) that identifiability of $v_1$ requires changes in both acetate and $v_1$.
-We also know, based on our knowledge of the Michaelis-Menten kinetic rate law that changes in the substrate concentration of a reaction can bring about a nonlinear change in the value of the corresponding reaction rate (flux).
-In this specific metabolic network, since the substrate is an input variable to the model and $v_1$ is the corresponding uptake flux, the substrate can be easily perturbed to create persistently excitable experiments to identify parameters in $v_1$.
+We also know, based on our knowledge of the Michaelis-Menten kinetic rate law that changes in the substrate concentration of a reaction can bring about a nonlinear change in the value of the corresponding reaction rate.
+In this specific metabolic network, since the substrate is an input variable to the model, and $v_1$ is the corresponding uptake flux, the substrate can be easily perturbed to create persistently excitable experiments to identify parameters in $v_1$.
 We can generalize this observation for the identification of all uptake fluxes in all metabolic networks, i.e., at a minimum, a change in the input substrate concentration may be necessary for an informative experiment to identify the uptake flux parameters.
@@
 However, since both of these are system outputs, satisfaction of this condition cannot be guaranteed without an analysis of the dynamics of the metabolic network and how changes in the input (acetate) bring about changes in the two requisite output quantities.
-Previously, a dynamical analysis of the current network has shown that the concentration of \textit{pep}, and hence the flux $v_2$ are functions of not only the input acetate concentration, but also the levels of expression of the different enzymes within the network (Srinivasan et al., 2017).
+Previous dynamical analysis of the network (Figure \ref{fig:network} in the Appendix) has already established the existence of a functional relationship between \textit{pep} and $v_2$, and the input acetate concentration and the levels of expression of the different enzymes within the network \parencite{Srinivasan2017}.
 Hence, it is theoretically possible for any of the five different experiment types to be persistently excitable to identify $v_2$.
 This is confirmed by the high degree of identifiability for both parameters of $v_2$, where in more than 90\% of all data combinations can identify the parameters.
-Thus, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending the position of the flux in the metabolic network with respect to the inputs and the outputs of the network.
+Thus, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending on the position of the flux in the metabolic network with respect to the inputs and the outputs of the network, as well as the various regulatory interactions present within the network.
+We can extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_3$ as well (Figure \ref{fig:figure2_v3}).
 
-Nonetheless, all these explanations could only be verified due to the fact that the conditions for identifiability for both $v_1$ and $v_2$ are simple and readable due to the use of the Michaelis-Menten rate law to model these fluxes.
-Instead, in addition to presence of an allosteric interaction, when the convenience kinetic rate law is used to model this activation in $v_3$, we find that the complex expressions obtained for the three parameters preclude such an in-depth insight into the nature of persistently excitable experiments to identify  $v_3$.
-Thus, the only avenue through which experiments can be designed for the identification of parameters of $v_3$ would be through our proposed methodology wherein manual curation of experimental combinations required for identifiability are not necessary.
+\subsection{Effect of noise on degree of identifiability is dependent on nonlinearity of enzyme kinetic rate law models}\label{sec:noise}
+In reality, experimental data from biological systems is usually noisy.
+Noise in data gathered from experiments is associated with biological noise attributable to the stochasticity of cellular function, as well as measurement noise associated with the techniques used to obtain concentration and flux measurements.
+However, the previous sections describe practical identifiability for parameters obtained using noise-free in silico experimental data.
+So, in order to illustrate the ability of our method to test the practical identifiability of parameters under realistic circumstances, we apply our methodology on in silico data (concentration and flux) with 5\% additive noise.
 
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure2/v2max_k2pep_experiments}}
-	\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter $V_2^{max}$ a) experiment 1 ($k=1$), and b) experiment 2 ($k=2$) in combination.
-Different experiments that contribute towards identifiability of $K_2^{pep}$ c) experiment 1 ($k=1$) and d) experiment 2 ($k=2$) in combination.}\label{fig:figure3}
-\end{figure}
-
-\subsection{Effect of noise on degree of identifiability is dependent on nonlinearity of enzyme kinetic rate law models:}\label{sec:noise}
-The degree of identifiability for all the parameters shown in Figure \ref{fig:figure1} is obtained using noise-free in silico experimental data.
-However, actual experimental data from biological systems is usually noisy.
-Data gathered from experiments is usually associated with both biological noise attributable to the stochasticity of cellular function, as well as measurement noise associated with the techniques used to obtain concentration and flux measurements.
-In order to test the practical identifiability of parameters under circumstances where the presence of noise in the experimental data can affect the identifiability of the parameters, we apply our methodology on data (concentration and flux) with 5\% additive noise.
-50 different samples of noise with 0 mean and 0.05 standard deviation are drawn from a normal distribution and added to the steady state experimental data generated on the basis of experiments described in Section \ref{sec:experiments} to generate 50 different samples of noisy data.
-Subsequently, all 50 samples are tested for their ability to practically identify parameters for all three fluxes of the small network.
-The corresponding graphs generated for practical identifiability analysis using noisy experimental data are shown in the Appendix.
+50 different samples of noise with 0 mean and 5\% standard deviation are drawn from a normal distribution and added to the steady state experimental data generated on the basis of experiments described in Section \ref{sec:experiments} to generate 50 different samples of noisy data.
+Subsequently, all 50 samples are tested for their ability to practically identify all three fluxes of the small network (Figure \ref{fig:network}).
+The degree of identifiability of the different parameters using noisy experimental data are shown in the Appendix (Figure \ref{fig:figure1_noisy}).
 
@@
 
-Despite the presence of noise, the degree of identifiability of $v_1$ and $v_2$ (figure not shown) does not change from the case where no noise is present in the experimental data (Figure \ref{fig:figure1}).
-However, the degree of identifiability for parameters of $v_3$ is affected by the presence of noise in the data (Figure).
-This effect is represented by the presence of a non-zero standard deviation and corresponding error bars on the graphs in the degrees of identifiability of parameters modeling $v_3$ (Figure).
+Despite the presence of noise, the degree of identifiability of $v_1$ and $v_2$ (Figure \ref{fig:figure1_noisy}a and b in the Appendix) does not change from the case where no noise is present in the experimental data (Figure \ref{fig:figure1}).
+However, the degree of identifiability for parameters of $v_3$ is affected by the presence of noise in the data (Figure \ref{fig:figure1_noisy}c in the Appendix).
+This effect is represented by the presence of a non-zero standard deviation and corresponding error bars on the graphs in the degrees of identifiability of parameters modeling $v_3$ (Figure \ref{fig:figure1_noisy}c in the Appendix).
 The standard deviation and the error bars represent the presence of data combinations that can identify a given parameter under certain values of noise, but not do so under different noise values.
-In the presence of 5\% additive noise in the data, the conditions under which parameters for $v_1$ and $v_2$ can be identified are strictly satisfied by the same data combinations that can identify these parameters in the absence of noise, i.e., the changes in the acetate concentration (for $v_1$) and \textit{pep} (for $v_2$) are significant enough to enable identifiability.
-The contrinution of different experiment types towards identifiability also remains the same (figure not shown separately).
+In the presence of 5\% additive noise in the data, the conditions under which parameters for $v_1$ and $v_2$ can be identified are strictly satisfied by the same data combinations that can identify these parameters in the absence of noise, i.e., the changes in the acetate concentration (for $v_1$) and \textit{pep} (for $v_2$) are significant enough to enable identification.
+The contribution of different experiment types towards identifiability also remains the same (figure not shown separately).
 
@@
 
-These observations, in conjunction with the estimation of the degree of identifiability of the different parameters (Figure \ref{fig:figure1}), helps in determining not only the nature of experiments required for parameter identifiability (Figure \ref{fig:figure2}), but also the order which the type of experiments necessary for identifying the entire metabolic network needs to be evaluated.
+These observations, in conjunction with the estimation of the degree of identifiability of the different parameters (Figure \ref{fig:figure1}) help in determining not only the nature of experiments required for parameter identifiability (Figure \ref{fig:figure2}), but also the order in which the type of experiments necessary for identifying the entire metabolic network needs to be evaluated.
 
@@
 
-\subsection{Utility of experimental data for practical parameter identification:}\label{sec:data_utility}
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure4/v1_v2_v3_data_utility}}
-	\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
-The total possible number of parameters that can be identified by data from combinations of a) and b) two, and c) three experiments is shown in the horizontal axis.
-The vertical axis represents the total number of combinations that can identify the corresponding number of parameters.
-The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 306 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 4896 combinations are used to practical identifiability of parameters in c)$v_3$.
-Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
-\end{figure}
+\section{Discussions}\label{sec:discussion}	
+Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data.
+The presence of noise is typically said to be a significant factor that results in non-identifiability.
+However, the primary reason for non-identifiability of parameters from experimental data can be attributed to the lack of information about the dynamics of the system whose parameters are being estimated within the chosen experimental data.
+The informativeness of experiments can be tied back to their ability to discriminate the dynamics of the system under two or more different input conditions.
+Thus, the presence of noise only serves to exacerbate the inability of experiments to discriminate the dynamics of the systems.
 
-Apart from determining experiments that are informative enough to identify parameters for each flux in a kinetic model, experimental design can also be used to design and develop a minimal set of experiments that informative enough to estimate all parameters in a metabolic network model.
-However, there is not only a scarcity of experimental design methods in literature devoted to the purpose of designing experiments for parameter identifiability in kinetic models of metabolism, the motivation to develop such methods is not well understood.
-Hence, we found the need to stress the necessity for experimental design methodologies tailored specifically for metabolic network models.
-We do this by showing how useful the multiple combinations of experimental data sets are towards identifying the maximum number of parameters for each given flux (Figure \ref{fig:figure4}).
+Although methods have been developed for practical parameter identification and experimental design, existing methods for experimental design based on practical identification of parameters are based on solving nonlinear least squares problems using optimization approaches that cannot guarantee global optimal solutions \parencite{Raue2009a}.
+In this document, we have presented a scalable method to practically identify parameters in kinetic models of metabolism, and use it to design experiments that are minimal and informative for estimating the parameters that does not require solutions to non-convex optimization problems.
+While some authors have eschewed deterministic parameter estimation techniques in favour of Bayesian methods based on probabilistic estimation of parameters \parencite{Saa2016, Saa2016a}, The need to establish practical identifiability of parameters, even for these methods, still stands.
+Additionally, the ability to design experiments to get good priors for Bayesian approaches to parameter estimation can also be fulfilled through practical identifiability analysis.
 
-In Figure \ref{fig:figure4}, we show the percentage of all data combinations that can estimate a given number of parameters within a given flux.
-For flux $v_1$ (Figure \ref{fig:figure4}a) we see that about 78\% of all data combinations cannot identify a single parameter, while only about 18\% of the data combinations can identify all the four parameters (two parameters each under different model setups) used in the model.
-This directly correlates with the degree of identifiability of the various parameters of $v_1$, shown earlier in Figure \ref{fig:figure1}.
-
-In contrast, given the high degree of identifiability of both parameters for $v_2$, most data combinations (>90\%) are not wasted, and can identify both parameters (Figure \ref{fig:figure4}b).
-Furthermore, the data utility plots for $v_3$ (Figure \ref{fig:figure4}c) show that about 36\% of the 5000 data combinations can identify only two parameters while 6\% of the experiment combinations cannot identify any data set.
-Thus, Figure \ref{fig:figure4} provides evidence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
-Choosing any of the experiment combinations for the 78\% percent that does not identify any parameter in $v_1$, or the 6\% that does not identify any parameters in $v_3$ can lead to waste of resources used to perform these non-informative experiments.
-
-\section{Discussions:}
-Methods have been extensively developed for the establishing structurally identifiability of parameters in nonlinear biological model based on differential algebra techniques \parencite{Ljung1994}.
-However, very few methods exist for testing practical identifiability of parameters in nonlinear models and subsequently design experiments to estimate these parameters.
-One of the examples of a method for practical identifiability is the profile likelihood approach \parencite{Raue2009a}.
-However, methods that can overcome some of the difficulties associated with implementing this method for models of large metabolic network are sorely missing.
-Additionally, any test for identifiability is forgone in favor of direct parameter estimation using a nonlinear least squares optimization approach with powerful computers to overcome difficulties associated with getting global optimal solutions ().
-Deterministic parameter estimation techniques, based on optimization methods are paving way for probabilistic approaches based on Bayesian techniques (Lars Nielsen papers).
-The common thread that connects all these parameter estimation techniques is their failure to address the fundamental need for practically parameter identifiability before actual parameter estimation.
-
-Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data.
-Hence, these methods fail to address the issue of designing experiments to suit parameter estimation, and the need for ways to determine informative experiments remains.
-
-In this paper we have developed a methodology to test the practical identifiability of parameters in kinetic models of metabolism.
-Our methodology enables practical parameter identification for each individual flux within a metabolic network under the assumption that all steady state substrate, product and allosteric effector concentrations and fluxes for a given reaction can be measured.
+Due to their fundamental aspect of requiring solutions to non-convex optimization problems, existing methods for practical identifications may not be scalable for identification and experimental design for parameters of large metabolic networks.
+By establishing identifiability for each flux within a metabolic network individually, we hope to overcome the scalability obstacle.
+Furthermore, we believe our method offers an algorithmic alternative to determine persistently excitable experiments that can enable identification of all fluxes within a metabolic network.
 Using a small metabolic network for gluconeogenesis, we have demonstrated that the identifiability of parameters for a given flux is dependent on the position of the flux within the metabolic network.
@@
 Our methodology fills the niche gap of experimental design for parameter estimation by providing a way to design informative experiments to obtain data required for parameter estimation by spending the least amount of resources.
+In the future, we believe our work can be extended and formulated as a mixed integer linear programming problem that can be solved to determine the type and total minimum number of experiments necessary to estimate all parameters in kinetic models of genome-scale metabolic networks.
 
-Based on the results we obtained in establishing the identifiability of parameters modeling all the fluxes of a small metabolic network, in the future we hope to extend this methodology to a larger metabolic network with more reactions.
+%Nonetheless, all these explanations could only be verified due to the fact that the conditions for identifiability for both $v_1$ and $v_2$ are simple and readable due to the use of the Michaelis-Menten rate law to model these fluxes.
+Instead, in addition to presence of an allosteric interaction, when the convenience kinetic rate law is used to model this activation in $v_3$, we find that the complex expressions obtained for the three parameters preclude such an in-depth insight into the nature of persistently excitable experiments to identify  $v_3$.
+Thus, the only avenue through which experiments can be designed for the identification of parameters of $v_3$ would be through our proposed methodology wherein manual curation of experimental combinations required for identifiability are not necessary.
+
+%Based on the results we obtained in establishing the identifiability of parameters modeling all the fluxes of a small metabolic network, in the future we hope to extend this methodology to a larger metabolic network with more reactions.
+%Apart from determining experiments that are informative enough to identify parameters for each flux in a kinetic model, experimental design can also be used to design and develop a minimal set of experiments that informative enough to estimate all parameters in a metabolic network model.
+However, there is not only a scarcity of experimental design methods in literature devoted to the purpose of designing experiments for parameter identifiability in kinetic models of metabolism, the motivation to develop such methods is not well understood.
 

### 2018-03-22 ident/docs/draft_1.tex

--- 2018-02-06 ident/docs/ident_manuscript.tex
@@
 
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+\usepackage{lmodern}
 \usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage{tabu}
-\usepackage{lmodern}
+\usepackage{pifont}
+\usepackage{bm}
 \usepackage{graphicx}
 \usepackage[space]{grffile}
+\usepackage{multicol}
+\usepackage{array}
+\usepackage{tabu}
+\usepackage{ragged2e}
+\usepackage{setspace}
+\usepackage[font=small,labelfont=bf,labelsep=period]{caption}
+%\usepackage{subcaption}
+%\usepackage[CaptionAfterwards]{fltpage}
 \usepackage{subfigure}
-\usepackage{longtable}
-\usepackage{multirow}
+\usepackage{lineno}
+\linenumbers
+\usepackage{tikz}
+\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;} 
 \usepackage[margin=1.0in]{geometry}
-\renewcommand{\baselinestretch}{2.0}
 
-%\usepackage[font=small,labelfont=bf,labelsep=period]{caption}
-\usepackage[style=authoryear,sorting=nyt,url=false,isbn=false,doi=false,firstinits=true,backend=biber]{biblatex}
-
-\renewcommand{\baselinestretch}{2.0}
-\renewcommand*\contentsname{Table of Contents}
-%\captionsetup{font={stretch=2.0}}
+\usepackage[backend=biber,style=authoryear,sorting=nyt,url=false,isbn=false,doi=false,firstinits=true]{biblatex}
 
@@
 	\setunit*{\addcomma\space}%
-	\printlist{location}%
 	\setunit*{\addcomma\space}%
@@
 
+\renewcommand{\baselinestretch}{2.0}
 \addbibresource{refs/ident_refs.bib}
+\captionsetup{font={stretch=2.0}}
 
 \begin{document}
+	\begin{center}
+		\begin{Large}
+			Practical Identification and Experimental Design for Parameter Estimation in Kinetic Models of Metabolism
+		\end{Large}\\
+		Shyam Srinivasan\textsuperscript{a}, William R. Cluett\textsuperscript{a} and Radhakrishnan Mahadevan\textsuperscript{*,a,b}\\
+	\end{center}
+	a - Department of Chemical Engineering and Applied Chemistry, University of Toronto, Toronto, ON, Canada.\\
+	b - Institute for Biomaterials and Biomedical Engineering, University of Toronto, Toronto, ON, Canada.\\
+	{*} Corresponding author
+	\section*{Abstract}
 	\section{Introduction:}
-Kinetic models of metabolism can be used to study the dynamic characteristics of metabolic networks.
+	Kinetic models of metabolism can be used to study the dynamic characteristics of metabolic networks \parencite{Bordbar2015, Andreozzi2016}.
 In these models, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
@@
 	\end{subequations}
-The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Heijnen2005, Link2014, Machado2011, Srinivasan2015}.
+	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Link2014, Srinivasan2015}.
 Accordingly, $f$ is a nonlinear function of the metabolite concentrations, enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
 
-The ability to predict the steady state and dynamic responses of metabolic networks, under in vivo conditions, to different perturbations is dependent on the numerical values of the enzyme kinetic parameter values ($\theta$) in Equation (\ref{eq:kinstoich}).
-Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where parameter values might play a crucial role \parencite{Srinivasan2015, Vital-Lopez2006}(Srinivasan et al., 2017?).
-The use of in vitro, or unreliable in vivo parameter estimates, reduces confidence in the model predicted behaviour.
-Consequently, the reduction in confidence hampers the use of these models to gain insight into the functioning of metabolic networks \parencite{Tran2008, Chakrabarti2013a}.
-The insights gained from the use of kinetic models are subsequently used to design changes to these metabolic networks to achieve various goals.
-These goals could either be to increase metabolite production for biosynthesis of different chemicals \parencite{Almquist2014, Khodayari2016, Costa2016, Andreozzi2016}(Srinivasan et al., 2017) or to find therapeutic targets to cure ailments \parencite{Apaolaza2017}.
-Hence, an increase in uncertainty in model predicted responses is also an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve these goals.
+	Unlike CBMs, where the ability to predict the steady state responses of metabolic networks is dependent on the stoichiometry of the network, the prediction of dynamic responses of metabolic networks to different perturbations using kinetic models of metabolism is also dependent on the numerical values of the enzyme kinetic parameters ($\theta$ in Equation \ref{eq:kinstoich}) \parencite{Andreozzi2016a}.
+Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where parameter values might play a crucial role \parencite{Srinivasan2015, Vital-Lopez2006}.
+The use of in vitro, or unreliable in vivo parameter estimates reduces confidence in the model predicted behaviour \parencite{Andreozzi2016a}.
+Consequently, the reduction in confidence hampers the use of these models to gain insight into the functioning of metabolic networks \parencite{Chakrabarti2013a, Bordbar2015}.
+The corresponding increase in uncertainty in model predicted responses becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design.
 
-If all intracellular metabolite concentrations can be measured over a time course, a nonlinear programming problem can be formulated to estimate the enzyme kinetic parameters ($\theta$) in Equation (\ref{eq:kinstoich}), based on the measured data.
+	Parameter estimation methods based on optimization principles are typically used to determine true parameter values based on available experimental data.
+Under the assumption that all time dependent intracellular metabolite concentrations can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:kinstoich}) to estimate the values of enzyme kinetic parameters, $\theta$, based on the measured data.
 	The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration at each time point, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}).
@@
 
-However, the ability to determine unique solutions to parameters $\theta$ is governed by the identifiability of these parameters in the model \parencite{McLean2012}.
+	The ability to determine unique solutions to parameters $\theta$ is governed by the identifiability of these parameters in the model \parencite{McLean2012, Raue2009a}.
 The identifiability of parameters in nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability.
@@
 
-In this paper, we propose a scalable methodology to establish practical identifiability for parameters in kinetic models of metabolism using abundantly available steady state concentration and flux data.
-We present a computer algebra-based method that can facilitate experimental design through practical identifiability of parameters separately for each individual reaction within a metabolic network based on available steady state experimental data.
-We illustrate the utility of this method by applying it to a small network of gluconeogenesis in \textit{E.
-coli} and demonstrating our ability to propose experiments that will facilitate parameter estimation for a kinetic model of this network.
+	In this paper, we propose a scalable methodology to establish practical identifiability for parameters in kinetic models of metabolism using steady state concentration and flux data.
+We present a computer algebra-based method that can facilitate experimental design through practical identification of parameters separately for each individual reaction within a metabolic network.
+For the purposes of this method we assume that all intracellular metabolite concentrations and fluxes can be measured.
+We identify and design experiments to estimate parameters in a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} to illustrate the utility of our method.
 %We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
 
-\section{Methods:}	
-\subsection{A method for practical identifiability of kinetic models of metabolism:}\label{sec:ident}
-In this section, we show how practical identifiability of kinetic parameters in a dynamic model of metabolism can be established.
+	\section{Methods}\label{sec:methods}
+	\subsection{A method to determine practical identifiability of kinetic models of metabolism}\label{sec:ident}
+	We provide the mathematical framework for practical identification of parameters in kinetic models of metabolism in this section.
 A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
-In a kinetic model, the value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
-Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations.
-The fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
+In kinetic models of metabolic networks (Equation \ref{eq:kinstoich}), the fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
+The value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
+Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations (Figure \ref{fig:ident-flowchart}a).
 
 	\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure3/ident_analysis}}
+		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
 		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
-a) The steps for the construction of a kinetic model of a metabolic network are shown.
+a) The steps for the construction of a kinetic model of a metabolic network.
 The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
 The identifiability of parameters for each flux can be established independently.
-b) The steps for identifiability analysis for parameters of a single flux are shown.}\label{fig:ident-flowchart}
+b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
 	\end{figure}
@@
 	Let $\theta \in \mathbb{R}^p$ in Equation (\ref{eq:output}b) for each flux $v_i$ in the network.
-For each experiment $j = {1, 2, ..., n}$, we assume that all metabolite concentrations $x$ and reaction fluxes $v$ are measurable.
+As stated earlier, for each experiment $j = {1, 2, ..., n}$, we assume that all metabolite concentrations ($x$) and reaction fluxes ($v$) are measurable.
 The pertinent information for each experiment is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
 
-In order to establish the identifiability of kinetic parameters for each flux $v_i$, we describe a computer algebra-based method.
-The primary use of the computer algebra system is to obtain closed form expressions for each parameter in $\theta$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
-This is done by solving a system of nonlinear algebraic equations in $\mathbb{R}^p$, shown in Equation (\ref{eq:nonlineq}).
+	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, we describe a computer algebra-based method.
+The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\theta$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
+This is done by solving a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
 	\begin{align}\label{eq:nonlineq}
@@
 	Each equation in (\ref{eq:nonlineq}), indicated by the index $k$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for $v_i$, described earlier in Equation (\ref{eq:output}b), written for concentrations and fluxes obtained from experiment $k$.
-Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$, where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
+Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
 	\begin{align}\label{eq:theta-eq}
@@
 	\end{align}
-The identifiability of parameter $\theta_k$ for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable (practically non-identifiable) if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ ($D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$).
-Furthermore, the physical properties of the kinetic parameter values can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
+	The identifiability of parameter $\theta_k$ for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
+Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
 
-In the following sections we provide a previously published kinetic model of a small gluconeogenic network, followed by a demonstration of our methodology to establish practical identifiability for one of the fluxes in this network.
+	In the following sections we provide a previously published kinetic model of a small gluconeogenic network (section \ref{sec:small-model}), followed by a demonstration of our methodology to establish practical identifiability for one of the fluxes in this network (section \ref{sec:example}).
 
-\subsection{Kinetic model of gluconeogenesis in E. coli:}\label{sec:small-model}
-The proposed model for acetate consumption through gluconeogenesis and its corresponding kinetic model is used as a case study to illustrate the utility of identifiability analysis for the design of experiments for estimating parameters in kinetic models of metabolism.
+	\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
+	A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate the utility of identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism.
 The kinetic model is described below.
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
+		\caption{The small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
+	\end{figure}	
 	\begin{equation}\label{eq:ode1}
@@
 	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
-The added flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
+The flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
 	\begin{equation}\label{eq:flux4}
@@
 
-\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis:}\label{sec:example}
-Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_1$ in the small model of gluconeogenesis described in Section \ref{sec:small-model}.
-For the purposes of this demonstration, we assume that all relevant steady state metabolite concentrations and fluxes can be measured.
+	\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis}\label{sec:example}
+	Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_1$ in the small model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model}.
 
 	In flux $v_1$, the concentration of the enzyme E is used as a variable.
-If the enzyme concentration can be measured, then the expression for $v_1$ given in Equation (\ref{eq:flux1}) can be used for identifiability analysis of parameters $k_1^{cat}$ and $K_1^{ac}$.
-However, if enzyme concentrations are not available, the expression in Equation (\ref{eq:flux1}) can be modified as given in Equation (\ref{eq:flux1a}) below.
-This equation does not make use of enzyme concentrations as variables, and uses $V_1^{max}$ and $K_1^{ac}$ as parameters.
+Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, the expression previously given for $v_1$ (Equation \ref{eq:flux1}) cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
+So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
+Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
+The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
 	\begin{align}\label{eq:flux1a}
-v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}}
+	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
+	\end{align}		
+	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability.
+
+	As mentioned in section \ref{sec:intro}, practical identifiability is a necessary condition to estimate true values of both $V_1^{max}$ and $K_1^{ac}$ from experimental data.
+Here, we assume that the concentrations and fluxes from at least two different experiments are available i.e., in Equation (\ref{eq:nonlineq}) $k = \{1, 2\}$ .
+We label the available concentrations and fluxes as ${ac}^{(k)}$ and ${v_1}^{(k)}$, respectively.
+Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
+	\begin{align}\label{eq:nonlin-flux2}
+	{v_1}^{(k)} = V_{1}^{max}\frac{ac^{(k)}}{ac^{(k)}+K_{1}^{ac}(ne)} &&  k=\{1, 2\}
 	\end{align}
 
-We choose this expression for flux $v_1$ (Equation \ref{eq:flux1a}), expressed using Michaelis-Menten kinetics, to demonstrate our method for practical identifiability.
-Both $V_1^{max}$ and $K_1^{ac}$ need to be identifiable so that they can be estimated from experimental data.
-Here, we assume that data (for the concentrations and fluxes) from at least two different sets of experiments is available i.e., in Equation (\ref{eq:nonlineq}) $k = {1, 2}$ .
-We label the available concentrations and fluxes as ${ac}^{(k)}$ and ${v_1}^{(k)}$, respectively.
-Accordingly, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as follows:
-\begin{align}\label{eq:nonlin-flux2}
-{v_1}^{(k)} = V_{1}^{max}\frac{ac^{(k)}}{ac^{(k)}+K_{1}^{ac}} &&  k=\{1, 2\}
-\end{align}
-
-Solving this simultaneous system of $k$ equations using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic equations in the parameters $V_1^{max}$ and $K_1^{ac}$ based on the form shown earlier in Equation (\ref{eq:theta-eq}).
+	Solving this simultaneous system of $k$ equations using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
+These expressions have the form previously shown in Equation (\ref{eq:theta-eq}).
 	\begin{subequations}\label{eq:v1_par}
@@
 		V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
-	K_1^{ac} = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
+		K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
 		\end{align}
 	\end{subequations}
-In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}$ (Equation \ref{eq:v1_par}b) for different available experimental data set combinations.
-Since the enzyme binding constant ($K_1^{ac}$) and maximum reaction rate ($V_1^{max}$) cannot be negative, we can further reduce the criteria for identifiability for both these parameters by saying that the evaluated expressions should be non-negative (Figure \ref{fig:ident-flowchart}b).
+	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations.
+Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
 
-\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis:}\label{sec:experiments}
-Steady state metabolomics and fluxomics data can be gathered under different physiological conditions by either perturbing the expression levels for different enzymes within a metabolic network, or by changing the substrate concentrations under which the cells grow.
-The aforementioned model of gluconeogenesis has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($V_1^{max}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
-Furthermore, the acetate concentration, that determines the acetate uptake flux $v_1$, can also be perturbed to measure cellular response to changes in the substrate concentration.
-We use the in silico data generated by 18 different experiments wherein these four model parameters ($ac$, $V_1^{max}$, $V_2^{max}$ and $V_3^{max}$) are perturbed to demonstrate practical parameter identification with our methodology.
-The experiments and the perturbed values of each of the four the parameters are given in the Appendix.
+	\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
+	Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions.
+One way to alter the physiological conditions is to change the substrate concentration under which the cells grow.
+In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$.
+Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.
+
+	Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network.
+The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
+Accordingly, in addition to measuring network response to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
+
+	In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
+The 18 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
+	\begin{table}[!thbp]
+		\caption{Table showing the perturbed values of all parameters used to generate experimental data in silico for testing practical identifiability of all fluxes in the small metabolic network.}
+		\begin{center}				
+			\begin{tabular}{llccccc}
+				\hline
+				Experiment Type & Perturbed Parameter & \multicolumn{5}{c}{Perturbed Values}\\
+				\hline
+				wild type & - & \multicolumn{5}{l}{$acetate$ = 0.1}\\
+				\hline
+				acetate perturbation & $acetate$ & 0.5 & 1.0 & & &\\
+				\hline
+				$v_1$ perturbation & $k_1^{cat}$ & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\
+				\hline
+				$v_2$ perturbation & $V_2^{max}$ & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\				
+				\hline
+				$v_3$ perturbation & $V_3^{max}$ & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\
+			\end{tabular}
+		\end{center}	
+		\label{tab:pval}
+	\end{table}
 
 	The minimum number of experiments from which data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
-For instance, as demonstrated above, data from two distinct experiments is required for identifying the two parameters of $v_1$.
-In this case $p=2$.
-This also applies for identifying parameters in $v_2$.
-Hence, multiple data sets generated using a combination of any two experiments are used to test the identifiability of parameters for $v_1$ and $v_2$.
-The total number of such possible combinations is 306 (18 x 17) from the 18 different experiments mentioned earlier.
-A similar calculation for identifying the three parameters of $v_3$, where $p=3$, leads to 4896 distinct data combinations (18 x 17 x 16) of three experiments.
-% 18 different experiments mentioned earlier and elaborated in the Appendix.
+For instance, as demonstrated above, data from two distinct experiments is required for identifying $v_1$, as $v_1$ has two parameters ($\mathbb{R}^2$).
+$v_2$ is also modeled by a parameter space with dimensions of $\mathbb{R}^2$.
+Hence, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_1$ and $v_2$.
+The total number of such possible combinations is 306 (18 x 17) from the 18 different experiments shown in Table \ref{tab:pval}.
+Similarly, for identifying $v_3$, that is described by three parameters ($\mathbb{R}^3$), we need data from a combination of three different experiments.
+Based on the 18 experiments in Table \ref{tab:pval}, we have 4896 distinct combinations (18 x 17 x 16) of three experiments.
 
-\subsection{Degee of Identifiability, a quantitative measure of practical identifiability of parameters:}\label{sec:degree_of_identifiability}
+	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
 	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
-We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test practical identifiability) that can identify that parameter.
+We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
+
 	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
 On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
-Furthermore, we can create a hierarchy of practically identifiable parameters using their degree of identifiabilities.
-In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiabilities of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
-Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments, and design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network.
+Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
+In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
+Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
+Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
 
-\subsection{Experimental design through practical parameter identifiability for kinetic models of metabolism:}\label{sec:experimental_design}
+	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}
 	\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{C:/Users/shyam/Documents/Courses/CHE1125Project/Results/ident/figure3/experimental_design}}
-	\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identifiability of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
+		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
+		\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
 	\end{figure}
-Following the methodology described in Section \ref{sec:ident}, and demonstrated in Section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found from Section \ref{sec:experiments} can be tested for their ability to practically identify parameters for any of the three fluxes in the small metabolic network.
-This step would help distinguish identifiable experiment combinations from combinations that do not practically identify any parameter in the model (Figure).
-Subsequently, it is possible to obtain a collection of experiments that make up all identifiable data combinations that can be performed to obtain the most minimal and informative set of experiments to identify as many parameters as possible (Figure \ref{fig:ident-design}).
+	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
+This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguishing experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
+In doing so, it is possible to obtain a minimal and informative collection of experiments that can be performed to identify as many model parameters as possible (Figure \ref{fig:ident-design}).
 Consequently, the set of experiments can be used to estimate the identifiable parameters in the model.
-We explain this formally below.
+This is formally explained below.
 
-The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in Section \ref{sec:ident} and demonstrated in Section \ref{sec:example}.
-Subsequently, for any flux $v_i$, if for any $p$ combinations of indices $j$, the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make a parameter $\theta_k$ identifiable.
+	The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in section \ref{sec:ident} and demonstrated in Section \ref{sec:example} (Figure \ref{fig:ident-flowchart}b).
+Subsequently, for any flux $v_i$ and for any $p$ combinations of indices $j$, if the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
 Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
@@
 In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
-This analysis can be performed for each flux in a metabolic network independent of all the other fluxes.
-Hence, our method is theoretically scalable even to genome-scale models.
-
-Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability using our methodology.
+Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
+This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.
 
@@
 
-\section*{Appendix}
-\subsection*{In silico perturbation experiments for gluconeogenic model:}
-\begin{table}[!tbhp]
-	\caption{Table showing the perturbed values of all fluxes used for parameter estimation.}
-	\begin{center}				
-		\begin{tabular}{ccccccc}
-			\hline
-			Experiment Type & Perturbed Parameter & \multicolumn{5}{c}{Perturbed Values}\\
-			\hline
-			acetate & $ac$ & \multicolumn{5}{l}{0.1}\\
-			\hline
-			acetate perturbation & $ac$ & 0.5 & 1.0 & & &\\
-			\hline
-			$v_1$ perturbation & $k_1^{cat}$ & 1.1 & 1.5 & 2.0 & 0.9 & 0.5\\
-			\hline
-			$v_3$ perturbation & $V_3^{max}$ & 1.1 & 1.5 & 2.0 & 0.9 & 0.5\\
-			\hline
-			$v_2$ perturbation & $V_2^{max}$ & 1.1 & 1.5 & 2.0 & 0.9 & 0.5\\
-		\end{tabular}
-	\end{center}	
-	\label{tab:pval}
-\end{table}
-
 	\printbibliography
 \end{document}
-
-\subsection{Identifiability analysis: Definitions and Formulations}
-Any nonlinear dynamical system can be represented by a set of states $\mathbf{x}$, observables $\mathbf{y}$ that are dependent on the states, parameters $\mathbf{\mu}$, and inputs $u$ as in Equation (\ref{eq:dynmodel}).
-\begin{subequations}\label{eq:dynmodel}
-	\begin{align}
-	\dot{\mathbf{x}} = g(\mathbf{x},\mathbf{\mu},u)\\
-	\mathbf{y} = h(\mathbf{x},\mathbf{\mu},u)
-	\end{align}
-\end{subequations}
-Identifiability concerns with the ability to determine a unique solution to the problem of estimating parameters $\mathbf{\mu}$ from given data on the system observables $\mathbf{y}$ for inputs $u$ \parencite{McLean2012}.	The identifiability of parameters in nonlinear models of physical processes can be classified into two categories: structural and practical identifiability.
-
-Given the nonlinear nature of this model, the function $f$ in Equation (\ref{eq:flux_gen}) can expressed, without loss of generality as,
-\begin{align}\label{eq:nr_dr}
-v = \frac{N(\mathbf{x},\theta)}{D(\mathbf{x},\theta)}
-\end{align}
-
-We use a profile likelihood-based approach \parencite{Raue2009a} to establish structural and practical identifiability of parameters in nonlinear kinetic models of metabolism.
-Briefly, the approach seeks to establish the existence/non-existence of bounds in confidence intervals for the estimates of parameters in nonlinear models.
-The profile likelihood is calculated based on Equation (\ref{eq:pl}) for each parameter $\theta_i$ where $\chi^2(\theta_i)$ is given by Equation (\ref{eq:chi2}).
-\begin{align}\label{eq:pl}
-\chi_{PL}^2(\theta_i) = \underset{\theta_{j\ne i}}{\mathrm{min}} \left[\chi^2(\theta)\right]
-\end{align}
-
-In the minimization objective shown in Equation (\ref{eq:chi2}) for parameter estimation, $y_{kl}^*$ is the available experimental time course data for each observable state $k$ at each $l$ time point.
-The difference between the data and the model estimates at these time points, $y_{kl}$ is weighted by the variance in the experimental data $\sigma_{kl}^*$.
-An algorithm to calculate the profile likelihood, $\chi_{PL}^2(\theta_i)$, based on Equation \ref{eq:pl} is given below.
-
-The identifiability of parameters is established through the confidence intervals of their estimates, $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
-The likelihood-based confidence interval for any parameter whose profile likelihood is estimated can be written on the basis of a threshold $\Delta_{\alpha}$ in the likelihood as in Equation (\ref{eq:pleci}).
-\begin{align}\label{eq:pleci}
-\{\theta|\chi^2(\theta)-\chi^2(\hat{\theta})<\Delta_{\alpha}\}
-\end{align}
-The threshold $\Delta_{\alpha}$ in the likelihood is the 1-$\alpha$ quantile of the $\chi^2$ distribution, represented as $\chi^2(\alpha,df)$.
-The confidence intervals obtained hold for $df$ degrees of freedom.
-For a choice of $df$=1 the confidence intervals will hold for each parameter individually, and confidence intervals that hold jointly for all parameters can be obtained by choosing the number of parameters as $df$.
-
-The visualization of structurally and practically non-identifiable parameters using the profile likelihood approach is illustrated in Figure \ref{fig:identtypes}.
-The points of intersection between the profile likelihood curves (solid line) with the one parameter likelihood threshold ($\Delta_{\alpha} = \chi^2(\alpha,1)$, dashed line) provide the confidence intervals of the parameter $\theta_i$.
-The confidence intervals of a structurally non-identifiable parameter are unbounded, i.e., $\left[-\infty, +\infty\right]$ (Figure \ref{fig:identtypes}a), while the confidence intervals of a practically non-identifiable parameter are unbounded in at least one direction, i.e., $\left[\sigma_{i}^-,\sigma_{i}^+\right]$ where either $\sigma_{i}^- = -\infty$ or $\sigma_{i}^+ = +\infty$ (Figure \ref{fig:identtypes}b).
-If a parameter's estimates have a finite confidence interval then the parameter is said to be identifiable (Figure \ref{fig:identtypes}c).
-Note that the horizontal dotted lines in Figure \ref{fig:identtypes} represent the confidence interval thresholds ($\Delta_{\alpha}$) that are used to establish identifiability.
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=.8\textwidth,height=.8\textheight,keepaspectratio]{figures/identifiability}}
-	\caption{The profile likelihood estimates of a) a structurally non-identifiable, b) a practically non-identifiable and c) an identifiable parameter.
-The horizontal axis represents the changes in the value of the parameter ($\theta_i$) whose identifiability is being determined and the profile likelihood ($\chi^2_{PL}(\theta_i)$) is shown in the vertical axis.
-The confidence interval threshold ($\Delta_{\alpha}$) used to determine the identifiability of the parameter is denoted by the horizontal dotted line.
-Identifiable parameters are distinguished from non-identifiable parameters by the presence of both upper and lower bounds on their confidence interval estimates $\left[\sigma_{i}^-,\sigma_{i}^+\right]$.
-}\label{fig:identtypes}
-\end{figure}
-
-Due to the dependence of practical parameter identifiability on the experimental data, the profile likelihood approach can be used to design experiments in such a way that the observables that are derived from these experiments can improve the practical identifiability of the parameters.
-We show how experimental design can have a meaningful impact on parameter identification and estimation in Figure \ref{fig:edwithpl}.	Assuming a parameter $\theta_i$ is practically non-identifiable (Figure \ref{fig:edwithpl}a), performing a profile-likelihood based identifiability analysis using simulated data can help determine the nature of experiments needed to make the parameter identifiable (Figure \ref{fig:edwithpl}b).
-In contrast, performing non-informative experiments without prior knowledge on their ability to change the identifiability of the parameter may provide data that cannot be used to estimate parameter $\theta_i$ (Figure \ref{fig:edwithpl}c).
-
-%Thus, the identifiability of parameters in nonlinear models is dependent on the model structure, parameterization, and the quality and quantity of experimental data that is available for the purpose of estimation.
-
-%In the left panels (Figure \ref{fig:figure1}a-c), the data is shown when noise-free data is used for identification.
-The identifiability information for parameters using data with additive noise is shown on the right panel (Figure \ref{fig:figure1}d-f).
-The data shown in the right hand side panel was generated using 25 different samples of random additive noise selected from a normal distribution with 0 mean 0.05 standard deviation.
-The value used for standard deviation reflects a 5\% noise level.
-The average and standard deviation values shown for identifiability of parameters with data in the presence of noise were calculated on the basis of the number of data combinations that can identify the respective parameter for each of the 25 noisy data samples.
-As seen in these figures, irrespective of the presence or absence of noise in the data, the number of combinations that can identify each parameter, and concordantly, the degree of identifiability of each parameter varies widely.
-Next, we look at some of the trends in the degree of identifiability of each parameter in each flux.
-
-\paragraph{Outline:}
-\begin{itemize}
-	\item Importance of parameter values for estimating in vivo response using kinetic models of metabolism and consequently for metabolic network design using kinetic models of metabolism
-	\item The need for parameter identifiability to determine unique and true parameter values from observed data
-	\item Types and purpose of identifiability for parameters
-	\item Methods for structural identifiability and existing methods for practical identifiability
-	\item Lack of methods for practical identifiability and consequently experimental design (not covered?)
-	\item work done in this paper for practical identifiability			
-	\item scalability of computer algebra-based methods for structural identifiability (using CRNT to reduce networks to make structural identifiability scalable) (move to discussion - may be in discussion)
-\end{itemize}

### 2018-03-27 ident/docs/draft_1.tex
change table 1

@@
 		\begin{center}				
-			\begin{tabular}{llccccc}
+			\begin{tabular}{lp{3.5cm}ccccc}
 				\hline
-				Experiment Type & Perturbed Parameter & \multicolumn{5}{c}{Perturbed Values}\\
+				Experiment Type & Perturbed Parameter (wild type value a.u.)& \multicolumn{5}{c}{Perturbed Values}\\
 				\hline
-				wild type & - & \multicolumn{5}{l}{$acetate$ = 0.1}\\
+				wild type &  &\multicolumn{5}{l}{}\\
+				acetate perturbation & $acetate$ (0.1) & 0.05 & 0.09 & 0.11 & 0.5 & 1.0\\
+				$v_1$ perturbation & $k_1^{cat}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\				
+				$v_2$ perturbation & $V_2^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\								
+				$v_3$ perturbation & $V_3^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\
 				\hline
-				acetate perturbation & $acetate$ & 0.5 & 1.0 & & &\\
-				\hline
-				$v_1$ perturbation & $k_1^{cat}$ & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\
-				\hline
-				$v_2$ perturbation & $V_2^{max}$ & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\				
-				\hline
-				$v_3$ perturbation & $V_3^{max}$ & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\
 			\end{tabular}

### 2018-03-27 ident/docs/draft_1.tex
more changes to table 1

--- 2018-03-22 ident/docs/draft_1.tex
@@
 		\begin{center}				
-			\begin{tabular}{llccccc}
+			\begin{tabular}{lp{3.5cm}cccccp{2cm}}
 				\hline
-				Experiment Type & Perturbed Parameter & \multicolumn{5}{c}{Perturbed Values}\\
+				Experiment Type & Perturbed Parameter (wild type value a.u.)& \multicolumn{5}{c}{Perturbed Values} & Number of Experiments\\
 				\hline
-				wild type & - & \multicolumn{5}{l}{$acetate$ = 0.1}\\
+				wild type &  &\multicolumn{5}{l}{} & 1\\
+				acetate perturbation & $acetate$ (0.1) & 0.05 & 0.09 & 0.2 & 0.5 & 1.0 & 5\\
+				$v_1$ perturbation & $k_1^{cat}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\				
+				$v_2$ perturbation & $V_2^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\								
+				$v_3$ perturbation & $V_3^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\
 				\hline
-				acetate perturbation & $acetate$ & 0.5 & 1.0 & & &\\
+				\multicolumn{7}{l}{Total number of experimental data sets used for practical parameter identification:} & 21\\
 				\hline
-				$v_1$ perturbation & $k_1^{cat}$ & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\
-				\hline
-				$v_2$ perturbation & $V_2^{max}$ & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\				
-				\hline
-				$v_3$ perturbation & $V_3^{max}$ & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 \\
 			\end{tabular}

### 2018-04-02 ident/docs/draft_1.tex
significant changes to methods + move all figures to end of file

@@
 	Kinetic models of metabolism can be used to study the dynamic characteristics of metabolic networks \parencite{Bordbar2015, Andreozzi2016}.
-In these models, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
-The matrix $\mathbf{S}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
-	\begin{subequations}\label{eq:kinstoich}
-		\begin{align}
-		\dot{x} = \mathbf{S}v\\
-		v = f(x, \theta, u)
-		\end{align}
-	\end{subequations}
-	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Link2014, Srinivasan2015}.
-Accordingly, $f$ is a nonlinear function of the metabolite concentrations, enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
 
@@
 	\section{Methods}\label{sec:methods}
+	\subsection{Kinetic models of metabolic networks}\label{sec:kinetic_model}
+	In kinetic models of metabolism, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
+The matrix $\mathbf{S}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
+	\begin{subequations}\label{eq:kinstoich}
+		\begin{align}
+		\dot{x} = \mathbf{S}v\\
+		v = f(x, \theta, u)
+		\end{align}
+	\end{subequations}
+	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Link2014, Srinivasan2015}.
+Accordingly, $f$ is a nonlinear function of the metabolite concentrations ($x$), enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
+
 	\subsection{A method to determine practical identifiability of kinetic models of metabolism}\label{sec:ident}
@@
 A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
-In kinetic models of metabolic networks (Equation \ref{eq:kinstoich}), the fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
-The value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
+As indicated in Figure \ref{fig:ident-flowchart}a, the first step involves the construction of the kinetic model (Equation \ref{eq:kinstoich}) of the metabolic network with $n_r$ reaction fluxes.
+
+	%In kinetic models of metabolic networks (Equation \ref{eq:kinstoich}), the fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
+The value of every flux $v_i$, $i={1, 2, ..., n_r}$, is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
 Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations (Figure \ref{fig:ident-flowchart}a).
 
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
-		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
-a) The steps for the construction of a kinetic model of a metabolic network.
-The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
-The identifiability of parameters for each flux can be established independently.
-b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
-	\end{figure}
+	For each flux $v_i$, $i={1, 2, ..., n_r}$, in the kinetic model, let $\mathbf{\theta} \in \mathbb{R}^p$ in Equation (\ref{eq:kinstoich}b).
+If data from $n_E$ experiments is available for the chosen metabolic network, as stated earlier, for each experiment $j = {1, 2, ..., n_E}$, we assume that all metabolite concentrations ($x$) and reaction fluxes ($v$) are measurable.
+We discuss the implications of relaxing this assumption later.
+The pertinent information for each experiment $j$ is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
 
-	Let $\theta \in \mathbb{R}^p$ in Equation (\ref{eq:output}b) for each flux $v_i$ in the network.
-As stated earlier, for each experiment $j = {1, 2, ..., n}$, we assume that all metabolite concentrations ($x$) and reaction fluxes ($v$) are measurable.
-The pertinent information for each experiment is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
+	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method.
+The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta}$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
+This is done by first selecting a combination of $p$ data sets from the $n_{E}$ available experimental data sets.
+The combination consisting of data from $p$ different experiments is then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
+	\begin{align}\label{eq:nonlineq}
+	v_{i, j} = f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) && \forall j=\{1, 2, ..., p\}\subset\{1, 2, ..., n_E\}
+	\end{align}
+	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$, $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$.
+Note that $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$.
 
-	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, we describe a computer algebra-based method.
-The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\theta$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
-This is done by solving a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
-	\begin{align}\label{eq:nonlineq}
-	v_{i, k} = f_k(\mathbf{x}_k,\theta, u_k) && \forall k=\{1, 2, ..., p\}\subset\{1, 2, ..., n\}
+	Each equation in (\ref{eq:nonlineq}), indicated by the index $j$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for each $v_i$, $i={1, 2, ..., n_r}$, described in Equation (\ref{eq:kinstoich}b), written for concentrations ($\mathbf{x}_j$, $\mathbf{u}_j$) and fluxes ($v_{i,j}$) obtained from experiment $j$.
+Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
+Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote a vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from all chosen experiments $j = {1, 2, ..., p}$, in the selected combination.
+	\begin{align}\label{eq:theta-eq}
+	\theta_j = g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
 	\end{align}
-	Each equation in (\ref{eq:nonlineq}), indicated by the index $k$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for $v_i$, described earlier in Equation (\ref{eq:output}b), written for concentrations and fluxes obtained from experiment $k$.
-Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
-	\begin{align}\label{eq:theta-eq}
-	\theta_k = g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
-	\end{align}
-	The identifiability of parameter $\theta_k$ for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
-Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
+	The identifiability of parameter $\theta_j$ for flux $v_i$ can be established by determining the value of $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_j$ is said to practically identifiable if $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ and practically non-identifiable if $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
+Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
+The solution obtained for $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_j$, and an infinite number of are possible for a non-identifiable $\theta_j$.
+However, if there are multiple but finite solutions for $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, then the corresponding parameter $\theta_j$ is locally identifiable.
 
-	In the following sections we provide a previously published kinetic model of a small gluconeogenic network (section \ref{sec:small-model}), followed by a demonstration of our methodology to establish practical identifiability for one of the fluxes in this network (section \ref{sec:example}).
+	For completeness, in the following sections we provide a previously published kinetic model of a small gluconeogenic network (section \ref{sec:small-model}).
+In the following section we demonstrate our methodology to establish practical identifiability for one of the fluxes in this network (section \ref{sec:example}) (move to results).
 
@@
 The kinetic model is described below.
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
-		\caption{The small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
-	\end{figure}	
+
 	\begin{equation}\label{eq:ode1}
@@
 	\end{equation}
+
+	\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
+	Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions.
+One way to alter the physiological conditions is to change the substrate concentration under which the cells grow.
+In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$.
+Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.
+
+	Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network.
+The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
+Accordingly, in addition to measuring network response to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
+
+	In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
+The 18 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
+	\begin{table}[!thbp]
+		\caption{Table showing the perturbed values of all parameters used to generate experimental data in silico for testing practical identifiability of all fluxes in the small metabolic network.}
+		\begin{center}				
+			\begin{tabular}{lp{3.5cm}cccccp{2cm}}
+				\hline
+				Experiment Type & Perturbed Parameter (wild type value a.u.)& \multicolumn{5}{c}{Perturbed Values a.u.} & Number of Experiments\\
+				\hline
+				wild type &  &\multicolumn{5}{l}{} & 1\\
+				acetate perturbation & $acetate$ (0.1) & 0.05 & 0.09 & 0.2 & 0.5 & 1.0 & 5\\
+				$v_1$ perturbation & $k_1^{cat}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\				
+				$v_2$ perturbation & $V_2^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\								
+				$v_3$ perturbation & $V_3^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\
+				\hline
+				\multicolumn{7}{l}{Total number of experimental data sets used for practical parameter identification:} & 21\\
+				\hline
+			\end{tabular}
+		\end{center}	
+		\label{tab:pval}
+	\end{table}
+
+	The minimum number of experiments from which data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
+For instance, as demonstrated above, data from two distinct experiments is required for identifying $v_1$, as $v_1$ has two parameters ($\mathbb{R}^2$).
+$v_2$ is also modeled by a parameter space with dimensions of $\mathbb{R}^2$.
+Hence, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_1$ and $v_2$.
+The total number of such possible combinations is 420 (21 x 20) from the 18 different experiments shown in Table \ref{tab:pval}.
+Similarly, for identifying $v_3$, that is described by three parameters ($\mathbb{R}^3$), we need data from a combination of three different experiments.
+Based on the 21 experiments in Table \ref{tab:pval}, we have 7980 distinct combinations (21 x 20 x 19) of three experiments.
+
+	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
+	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
+We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
+
+	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
+On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
+Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
+In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
+Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
+Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
+
+	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}
+
+	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
+This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguishing experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
+In doing so, it is possible to obtain a minimal and informative collection of experiments that can be performed to identify as many model parameters as possible (Figure \ref{fig:ident-design}).
+Consequently, the set of experiments can be used to estimate the identifiable parameters in the model.
+This is formally explained below.
+
+	The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in section \ref{sec:ident} and demonstrated in Section \ref{sec:example} (Figure \ref{fig:ident-flowchart}b).
+Subsequently, for any flux $v_i$ and for any $p$ combinations of indices $j$, if the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
+Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
+Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
+This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
+In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
+Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
+This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.
+
+	\section{Results:}
+	First in section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
+We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
+	Then, in section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models of metabolic networks.
+We do this by looking at the number of combinations that can identify the maximum number of parameters.
+	In section \ref{sec:trends} we show that the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetic rate law used to describe the flux.
+	In section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment.
+In this section we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux.
+Finally, we discuss some results arising out of the use of our methodology to determine the identifiability of parameters when data with additive noise is used in section \ref{sec:noise}.
 
@@
 
-	\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
-	Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions.
-One way to alter the physiological conditions is to change the substrate concentration under which the cells grow.
-In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$.
-Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.
-
-	Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network.
-The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
-Accordingly, in addition to measuring network response to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
-
-	In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
-The 18 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
-	\begin{table}[!thbp]
-		\caption{Table showing the perturbed values of all parameters used to generate experimental data in silico for testing practical identifiability of all fluxes in the small metabolic network.}
-		\begin{center}				
-			\begin{tabular}{lp{3.5cm}cccccp{2cm}}
-				\hline
-				Experiment Type & Perturbed Parameter (wild type value a.u.)& \multicolumn{5}{c}{Perturbed Values} & Number of Experiments\\
-				\hline
-				wild type &  &\multicolumn{5}{l}{} & 1\\
-				acetate perturbation & $acetate$ (0.1) & 0.05 & 0.09 & 0.2 & 0.5 & 1.0 & 5\\
-				$v_1$ perturbation & $k_1^{cat}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\				
-				$v_2$ perturbation & $V_2^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\								
-				$v_3$ perturbation & $V_3^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\
-				\hline
-				\multicolumn{7}{l}{Total number of experimental data sets used for practical parameter identification:} & 21\\
-				\hline
-			\end{tabular}
-		\end{center}	
-		\label{tab:pval}
-	\end{table}
-
-	The minimum number of experiments from which data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
-For instance, as demonstrated above, data from two distinct experiments is required for identifying $v_1$, as $v_1$ has two parameters ($\mathbb{R}^2$).
-$v_2$ is also modeled by a parameter space with dimensions of $\mathbb{R}^2$.
-Hence, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_1$ and $v_2$.
-The total number of such possible combinations is 306 (18 x 17) from the 18 different experiments shown in Table \ref{tab:pval}.
-Similarly, for identifying $v_3$, that is described by three parameters ($\mathbb{R}^3$), we need data from a combination of three different experiments.
-Based on the 18 experiments in Table \ref{tab:pval}, we have 4896 distinct combinations (18 x 17 x 16) of three experiments.
-
-	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
-	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
-We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
-
-	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
-On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
-Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
-In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
-Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
-Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
-
-	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
-		\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
-	\end{figure}
-	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
-This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguishing experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
-In doing so, it is possible to obtain a minimal and informative collection of experiments that can be performed to identify as many model parameters as possible (Figure \ref{fig:ident-design}).
-Consequently, the set of experiments can be used to estimate the identifiable parameters in the model.
-This is formally explained below.
-
-	The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in section \ref{sec:ident} and demonstrated in Section \ref{sec:example} (Figure \ref{fig:ident-flowchart}b).
-Subsequently, for any flux $v_i$ and for any $p$ combinations of indices $j$, if the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
-Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
-Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
-This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
-In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
-Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
-This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.
-
-	\section{Results:}
-	First in section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
-We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
-	Then, in section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models of metabolic networks.
-We do this by looking at the number of combinations that can identify the maximum number of parameters.
-	In section \ref{sec:trends} we show that the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetic rate law used to describe the flux.
-	In section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment.
-In this section we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux.
-Finally, we discuss some results arising out of the use of our methodology to determine the identifiability of parameters when data with additive noise is used in section \ref{sec:noise}.
-
 	\subsection{Nonlinearity of enzyme kinetic rate law expression affects identifiability analysis}\label{sec:closedform}
@@
 
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure4/v1_V1max_v2_v3_value1_data_utility}}
-		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
-The total possible number of parameters that can be identified by data from combinations of a) two, b) two, and c) three experiments is shown on the horizontal axis.
-The vertical axis represents the total number of combinations that can identify the corresponding number of parameters.
-The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 306 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 4896 combinations are used to practical identifiability of c) $v_3$.
-Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
-	\end{figure}	
 	\subsection{Utility of experimental data for practical parameter identification}\label{sec:data_utility}
@@
 	\subsection{Maximum reaction rates are more identifiable than enzyme binding constants}\label{sec:trends}	
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure1/v1_V1max_v2_v3_identifiability}}
-		\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
-The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
-Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:figure1}
-	\end{figure}
 	In Figure \ref{fig:figure1} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
@@
 	In order to determine the type of experiments that help satisfy identifiability conditions for both parameters of $v_1$, we look at the distribution of the five different types in all data combinations that can identify these 2 parameters (Figure \ref{fig:figure2}a and b).
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_only_v2_spider}}
-		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:figure2}
-	\end{figure}
 
@@
 	\printbibliography
+
+	\clearpage
+	\section{Figures and Figure Captions:}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
+		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
+a) The steps for the construction of a kinetic model of a metabolic network.
+The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
+The identifiability of parameters for each flux can be established independently.
+b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
+		\caption{The small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
+		\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure4/v1_V1max_v2_v3_value1_data_utility}}
+		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
+The total possible number of parameters that can be identified by data from combinations of a) two, b) two, and c) three experiments is shown on the horizontal axis.
+The vertical axis represents the total number of combinations that can identify the corresponding number of parameters.
+The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
+A total of 306 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 4896 combinations are used to practical identifiability of c) $v_3$.
+Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure1/v1_V1max_v2_v3_identifiability}}
+		\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
+The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
+Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:figure1}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_only_v2_spider}}
+		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
+The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:figure2}
+	\end{figure}
+
 \end{document}

### 2018-04-02 ident/docs/draft_1.tex
changes to methods + move table to end

@@
 	\end{align}
-	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$, $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$.
-Note that $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$.
+	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$, $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$.
 
@@
 Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
-Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote a vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from all chosen experiments $j = {1, 2, ..., p}$, in the selected combination.
+Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from all chosen experiments $j = {1, 2, ..., p}$, in the selected combination.
 	\begin{align}\label{eq:theta-eq}
@@
 	\end{align}
-	The identifiability of parameter $\theta_j$ for flux $v_i$ can be established by determining the value of $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_j$ is said to practically identifiable if $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ and practically non-identifiable if $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
+	The identifiability of parameter $\theta_j$, $j = {1, 2, ..., p}$, for flux $v_i$ can be established by determining the value of $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_j$ is said to practically identifiable if $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, and practically non-identifiable if $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
 Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
-The solution obtained for $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_j$, and an infinite number of are possible for a non-identifiable $\theta_j$.
+The solution obtained for $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_j$, and an infinite number of solutions are possible for a non-identifiable $\theta_j$.
 However, if there are multiple but finite solutions for $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, then the corresponding parameter $\theta_j$ is locally identifiable.
 
-	For completeness, in the following sections we provide a previously published kinetic model of a small gluconeogenic network (section \ref{sec:small-model}).
-In the following section we demonstrate our methodology to establish practical identifiability for one of the fluxes in this network (section \ref{sec:example}) (move to results).
+	%For completeness, in the following sections we provide a previously published kinetic model of a small gluconeogenic network (section \ref{sec:small-model}).
+
+	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
+	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
+We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
+
+	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
+On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
+Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
+In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
+
+	Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
+Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
 
@@
 	In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
-The 18 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
-	\begin{table}[!thbp]
-		\caption{Table showing the perturbed values of all parameters used to generate experimental data in silico for testing practical identifiability of all fluxes in the small metabolic network.}
-		\begin{center}				
-			\begin{tabular}{lp{3.5cm}cccccp{2cm}}
-				\hline
-				Experiment Type & Perturbed Parameter (wild type value a.u.)& \multicolumn{5}{c}{Perturbed Values a.u.} & Number of Experiments\\
-				\hline
-				wild type &  &\multicolumn{5}{l}{} & 1\\
-				acetate perturbation & $acetate$ (0.1) & 0.05 & 0.09 & 0.2 & 0.5 & 1.0 & 5\\
-				$v_1$ perturbation & $k_1^{cat}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\				
-				$v_2$ perturbation & $V_2^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\								
-				$v_3$ perturbation & $V_3^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\
-				\hline
-				\multicolumn{7}{l}{Total number of experimental data sets used for practical parameter identification:} & 21\\
-				\hline
-			\end{tabular}
-		\end{center}	
-		\label{tab:pval}
-	\end{table}
+The 21 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
 
-	The minimum number of experiments from which data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
-For instance, as demonstrated above, data from two distinct experiments is required for identifying $v_1$, as $v_1$ has two parameters ($\mathbb{R}^2$).
-$v_2$ is also modeled by a parameter space with dimensions of $\mathbb{R}^2$.
-Hence, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_1$ and $v_2$.
-The total number of such possible combinations is 420 (21 x 20) from the 18 different experiments shown in Table \ref{tab:pval}.
-Similarly, for identifying $v_3$, that is described by three parameters ($\mathbb{R}^3$), we need data from a combination of three different experiments.
+	As described previously in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, the minimum number of experiments from which steady state data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
+For instance, the flux $v_2$ in the model described in section \ref{sec:small-model} has two parameters: $V_2^{max}$ and $K_2^{pep}$.
+Thus $\theta = \{V_2^{max}, K_2^{pep}\} \in \mathbb{R}^2$.
+So, steady state data from two distinct experiments is required for identifying $v_2$.
+Accordingly, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_2$.
+The total number of such possible combinations is 420 (21 x 20) from the 21 different experiments shown in Table \ref{tab:pval}.
+Assuming we also have data for the enzyme concentration $E$ in the expression for flux $v_1$ (Equation \ref{eq:flux1}), identifying the two parameters $k_1^{cat}$ and $K_1^{ac}$ ($\theta \in \mathbb{R}^2$) also requires steady state experimental data from two different experiments.
+
+	Similarly, for identifying $v_3$ that is described by three parameters ($\theta \in \mathbb{R}^3$), we need data from a combination of three different experiments.
 Based on the 21 experiments in Table \ref{tab:pval}, we have 7980 distinct combinations (21 x 20 x 19) of three experiments.
-
-	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
-	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
-We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
-
-	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
-On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
-Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
-In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
-Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
-Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
+However, it is important to note that not all of the either 420 combinations (for $v_1$ and $v_2$), or the 7980 combinations (for $v_3$), may be suitable for parameter estimation.
+Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers.
+Moreover, among these chosen smaller number of experimental data combinations, our methodology enables us to calculate the number of experimental data combinations, and consequently, the number of experiments that might allow for parameter identification, and subsequent parameter estimation.
+The following section describes a procedure that can be used to design experiments within this context.
 
 	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}	
-
 	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
@@
 We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
+
+	%In the following section we demonstrate our methodology to establish practical identifiability for one of the fluxes in this network (section \ref{sec:example}) (move to results).
+
 	Then, in section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models of metabolic networks.
@@
 	\clearpage
-	\section{Figures and Figure Captions:}
+	\section{Tables, Figures and Figure Captions:}
+	\begin{table}[!thbp]
+		\caption{Table showing the perturbed values of all parameters used to generate experimental data in silico for testing practical identifiability of all fluxes in the small metabolic network.}
+		\begin{center}				
+			\begin{tabular}{lp{3.5cm}cccccp{2cm}}
+				\hline
+				Experiment Type & Perturbed Parameter (wild type value a.u.)& \multicolumn{5}{c}{Perturbed Values a.u.} & Number of Experiments\\
+				\hline
+				wild type &  &\multicolumn{5}{l}{} & 1\\
+				acetate perturbation & $acetate$ (0.1) & 0.05 & 0.09 & 0.2 & 0.5 & 1.0 & 5\\
+				$v_1$ perturbation & $k_1^{cat}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\				
+				$v_2$ perturbation & $V_2^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\								
+				$v_3$ perturbation & $V_3^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\
+				\hline
+				\multicolumn{7}{l}{Total number of experimental data sets used for practical parameter identification:} & 21\\
+				\hline
+			\end{tabular}
+		\end{center}	
+		\label{tab:pval}
+	\end{table}
+
 	\begin{figure}[!tbhp]

### 2018-04-02 ident/docs/draft_1.tex
major changes to results

@@
 	\end{align}
-	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$, $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$.
+	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$, $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$, whose elements are denoted by $\theta_k$.
 
 	Each equation in (\ref{eq:nonlineq}), indicated by the index $j$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for each $v_i$, $i={1, 2, ..., n_r}$, described in Equation (\ref{eq:kinstoich}b), written for concentrations ($\mathbf{x}_j$, $\mathbf{u}_j$) and fluxes ($v_{i,j}$) obtained from experiment $j$.
-Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
+Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^p$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
 Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from all chosen experiments $j = {1, 2, ..., p}$, in the selected combination.
 	\begin{align}\label{eq:theta-eq}
-	\theta_j = g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
+	\theta_k = g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
 	\end{align}
-	The identifiability of parameter $\theta_j$, $j = {1, 2, ..., p}$, for flux $v_i$ can be established by determining the value of $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_j$ is said to practically identifiable if $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, and practically non-identifiable if $D_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
-Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
-The solution obtained for $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_j$, and an infinite number of solutions are possible for a non-identifiable $\theta_j$.
-However, if there are multiple but finite solutions for $g_j(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, then the corresponding parameter $\theta_j$ is locally identifiable.
+	The identifiability of parameter $\theta_k$, $k = {1, 2, ..., p}$, for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
+Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
+The solution obtained for $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_k$, and an infinite number of solutions are possible for a non-identifiable $\theta_k$.
+However, if there are multiple but finite solutions for $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, then the corresponding parameter $\theta_k$ is locally identifiable.
 
@@
 	\begin{equation}\label{eq:ode3}
-	\frac{d}{dt}E=v_{e,max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right) - d E
+	\frac{d}{dt}E=v_5 - d E
 	\end{equation}
-	The kinetic expressions for fluxes $v_1$ through $v_4$ are given below.
+	The kinetic expressions for fluxes $v_1$ through $v_5$ are given below.
 The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics.
@@
 	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
+
+	The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (section \ref{sec:ident}).
+We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica or SymPy in Python).
+In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
+The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
+	\begin{align}\label{eq:flux3_convkin}
+	v_3 = V_3^{max}\left(\frac{1}{1 + \frac{K_3^{pep}}{pep}}\right)\left(\frac{\frac{fdp}{K_3^{fdp}}}{1 + \frac{fdp}{K_3^{fdp}}}\right)
+	\end{align}
+
 	The flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
@@
 	\end{equation}		
+	The production of enzyme E is represented by flux $v_5$.
+This flux is inhibited by \textit{fdp} and modeled using Hill kinetics, where $K_e^{fdp}$ represents the Hill binding constant for the inhibiting metabolite \textit{fdp}, $n_e$ is the Hill exponent, and $V_e^{max}$ is the maximum reaction rate for $v_5$.
+	\begin{align}\label{eq:flux5}
+	v_5 = V_e^{max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right)
+	\end{align}
 
@@
 
-	As described previously in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, the minimum number of experiments from which steady state data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
+	As described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, the minimum number of experiments from which steady state data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
 For instance, the flux $v_2$ in the model described in section \ref{sec:small-model} has two parameters: $V_2^{max}$ and $K_2^{pep}$.
@@
 Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers.
-Moreover, among these chosen smaller number of experimental data combinations, our methodology enables us to calculate the number of experimental data combinations, and consequently, the number of experiments that might allow for parameter identification, and subsequent parameter estimation.
+Moreover, among these chosen smaller number of experimental data combinations, our methodology enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation.
 The following section describes a procedure that can be used to design experiments within this context.
@@
 	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
-This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguishing experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
+This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguish experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
 In doing so, it is possible to obtain a minimal and informative collection of experiments that can be performed to identify as many model parameters as possible (Figure \ref{fig:ident-design}).
-Consequently, the set of experiments can be used to estimate the identifiable parameters in the model.
+Consequently, the set of experiments can be used to estimate all the identifiable parameters in the model.
 This is formally explained below.
 
-	The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in section \ref{sec:ident} and demonstrated in Section \ref{sec:example} (Figure \ref{fig:ident-flowchart}b).
-Subsequently, for any flux $v_i$ and for any $p$ combinations of indices $j$, if the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
-Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
-Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
+	The identifiability of each parameter based on each experiment indexed as $j = \{1, 2, ..., n_E\}$ is established based on the methodology described in section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in section \ref{sec:example}.
+Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
+Consequently, the corresponding experiment cannot be used for estimating parameter $\theta_k$, and needs to be discarded from the set of all necessary experiments.
+Furthermore, another experiment from $j = \{1, ..., n_E\}$ needs to be selected to replace the discarded experiment such that parameter $\theta_k$ is identifiable.
 This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
 In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
-Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
+Note that if none of the $n_E$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n_E+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
 This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.
@@
 	\section{Results:}
+	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model}.
+
 	First in section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
@@
 	\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis}\label{sec:example}	
-	Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_1$ in the small model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model}.
-
-	In flux $v_1$, the concentration of the enzyme E is used as a variable.
-Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, the expression previously given for $v_1$ (Equation \ref{eq:flux1}) cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
+	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}).
+Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression previously given for $v_1$ (Equation \ref{eq:flux1}) cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
 So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
@@
 
-	As mentioned in section \ref{sec:intro}, practical identifiability is a necessary condition to estimate true values of both $V_1^{max}$ and $K_1^{ac}$ from experimental data.
-Here, we assume that the concentrations and fluxes from at least two different experiments are available i.e., in Equation (\ref{eq:nonlineq}) $k = \{1, 2\}$ .
-We label the available concentrations and fluxes as ${ac}^{(k)}$ and ${v_1}^{(k)}$, respectively.
+	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ still holds for $v_1$, as mentioned in section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
+So, from the $n_E = 21$ different experiments described in section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$.
+We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively.
 Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
-	\begin{align}\label{eq:nonlin-flux2}
-	{v_1}^{(k)} = V_{1}^{max}\frac{ac^{(k)}}{ac^{(k)}+K_{1}^{ac}(ne)} &&  k=\{1, 2\}
-	\end{align}
+	\begin{align*}%\label{eq:nonlin-flux2}
+	{v_1}^{(j)} = V_{1}^{max}\frac{ac^{(j)}}{ac^{(j)}+K_{1}^{ac}(ne)} &&  j=\{1, 2\}
+	\end{align*}
 
-	Solving this simultaneous system of $k$ equations using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
-These expressions have the form previously shown in Equation (\ref{eq:theta-eq}).
+	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
+These expressions have the form shown in Equation (\ref{eq:theta-eq}).
 	\begin{subequations}\label{eq:v1_par}
 		\begin{align}		
-		V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
-		K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
+		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
+		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
 		\end{align}
@@
 
+	%As mentioned in section \ref{sec:intro}, practical identifiability is a necessary condition to estimate true values of both $V_1^{max}$ and $K_1^{ac}$ from experimental data. 	
+	%Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_1$ in the small model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model}.
+
 	\subsection{Nonlinearity of enzyme kinetic rate law expression affects identifiability analysis}\label{sec:closedform}
-	Practical identifiability of parameters in kinetic models of metabolism using the methodology described in section \ref{sec:methods} is governed by the nonlinear complexity of the enzyme kinetic rate law used to model a specific flux.
-We demonstrate one example of how the methodology works in section \ref{sec:example} for parameters of flux $v_1$.
-The expression in Equation (\ref{eq:v1_par}) is obtained by using a computer algebra system.
-To recall, in this specific case, a computer algebra system is used to solve for the parameters of a flux described using the Michaelis-Menten kinetic rate law when data from two different experiments is available.
+	In the previous section, we have demonstrated our methodology to practically identify parameters for a flux that is modeled using the Michaelis-Menten (MM) kinetic rate law when steady state concentration and flux data from two different experiments are available.
+However, metabolic fluxes can be described by several different kinetic rate law expressions based on their representation of the physiological interactions that facilitate enzyme catalysis of each flux in a metabolic network.
+For instance, as described in section \ref{sec:small-model}, flux $v_3$, which is allosterically activated by \textit{pep}, is modeled using the MWC kinetics (Equation \ref{eq:flux3}).
+So, we comment on the capability of our methodology to practically identify parameters in cases where metabolic fluxes are modeled using rate laws that are different from the MM model, and consequently differ in terms of their nonlinear complexity.
 
-	However, we find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica and SymPy in Python).
-In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
-The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
+	In using the convenience kinetic rate law model for $v_3$ (Equation \ref{eq:flux3_convkin}) for identifying the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$, we found that each of the three parameters have two different closed-form expressions owing to the presence of a square root term in their solutions.
+Thus, as per the stated definition in section \ref{sec:ident}, the parameters for $v_3$, described using the convenience kinetics rate laws, are only locally identifiable due to the non-unique but finite number of possible solutions for the three parameters.
+The two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, and $K_3^{pep}(1)$ refer to the first set of solutions, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$ refer to the other set of solutions.
 
-	\begin{align}\label{eq:flux3_convkin}
-	v_3 = V_3^{max}\left(\frac{1}{1 + \frac{K_3^{pep}}{pep}}\right)\left(\frac{\frac{fdp}{K_3^{fdp}}}{1 + \frac{fdp}{K_3^{fdp}}}\right)
-	\end{align}
+	However, we found that the local identifiability of the parameters could be resolved to obtain globally identifiable parameters if either the Michaelis binding constant for \textit{fdp}, $K_3^{fdp}$, or the allosteric binding constant for \textit{pep}, $K_3^{pep}$, is known.
+Thus, fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter.
+Although one could argue that a similar approach could be used to identify parameters of $v_3$ written using the MWC model, we found that it was not possible to obtain identifiability expressions for any of the parameters in the MWC model, except for $V_3^{max}$ when all other parameters are assumed to be known.
 
-	Using this expression for analysis, we find that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different closed-form expressions owing to the presence of a square root term in their solutions.
-These distinct expressions are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, $K_3^{pep}(1)$, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$.
+	Similarly, not all parameters in the Hill equation, used to model the inhibition of \textit{fdp} on the production of the enzyme E through flux $v_5$ (Equation \ref{eq:flux5}), can be identified using our methodology.
+We found that the Hill exponent ($n_e$) needs to be fixed to be able to identify the other two parameters, $V_e^{max}$ and $K_e^{fdp}$, in the model.
+Thus, the dimension of the parameter space for $v_5$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have developed in this paper.
+However, on applying the primary test for identifiability ($D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, see section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b), we found only $V_e^{max}$ to globally identifiable.
+We found $K_e^{fdp}$ to be locally identifiable to the possibility of having two different roots.
+As in the case of parameters for $v_3$ when using the convenience kinetics formulation, the presence of a square root term was the reason for the existence of two roots for $K_e^{fdp}$.
+However, on applying the secondary criteria for identifiability ($g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, see section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b), we were able to eliminate the existence of the second root as it is always negative, and therefore not physically relevant.
+Thus, we were able to determine that both parameters modeling $v_5$ ($V_e^{max}$ and $K_e^{fdp}$) are globally practically identifiable under certain conditions.
 
@@
 However, the identifiability of the parameters used in these models has never been truly examined.
+\textcolor{red}{mentione Heijnen papers on resolving identifiability using approximate models here}.
+
 	Unlike the case of the MWC model, if determining closed-form expressions for the parameters of these models are tractable, we believe our methodology can help in elucidating the identifiability of these parameters.
@@
 	Next, we look at some of the results we obtained for the practical identification of different parameters for all three fluxes $v_1$, $v_2$ and $v_3$ in the small gluconeogenic network model described in section \ref{sec:small-model}, with the flux for $v_3$ described by Equation (\ref{eq:flux3_convkin}).
-We start by looking at one of motivations for the need for designing experiments to collect data for parameter estimation.
 
 	\subsection{Utility of experimental data for practical parameter identification}\label{sec:data_utility}
-	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community, the need for experimental design methods to satisfy the data needs for parameter estimation are not considered seriously.
+	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not been well studied (\textcolor{red}{include Rudiyanto Gunawan papers here}).
 Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
@@
 
-	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:figure4} also informs on the ease of identifying parameters for a single flux.
+	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:figure4} also informs the ease of identifying parameters for a single flux.
 For the small network (Figure \ref{fig:network}), we find that more than 90\% of all the data combinations tested for their ability to identify $v_2$ are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:figure4}b).
@@
 Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range.
-Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
+%Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
 In linear systems, persistence of excitation of any input signal can be theoretically guaranteed.
@@
 
-	\subsection{Experiment requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}	
+	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}	
 	In this section, we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model.

### 2018-04-03 ident/docs/draft_1.tex
merge conflict correction from upstream

@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
+		\centering{\includegraphics[width=.9\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
 		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure3/experimental_design}}
 		\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure4/v1_V1max_v2_v3_value1_data_utility}}
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure4/v1_v1max_v2_v3_1_data_utility}}
 		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
 Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
-The total possible number of parameters that can be identified by data from combinations of a) two, b) two, and c) three experiments is shown on the horizontal axis.
-The vertical axis represents the total number of combinations that can identify the corresponding number of parameters.
+The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, and c) three experiments.
 The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 306 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 4896 combinations are used to practical identifiability of c) $v_3$.
+A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 3360 combinations are used to practically identify c) $v_3$.
 Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure1/v1_V1max_v2_v3_identifiability}}
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v1max_v2_v3_1_2_ident}}
 		\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_only_v2_spider}}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_v1max_v2_exp_info}}
 		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
@@
 
+	\clearpage
+
+	\section{Potential Supplementary Information Figures}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_data_utility}}
+		\caption{Same as other data utility figure for v1 with k1cat as parameter (with enzyme data) and second roots of v3.
+Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$.}\label{fig:figure4s}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_k1cat_exp_info}}
+		\caption{missing caption}%\label{fig:figure2}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_1_2_exp_info}}
+		\caption{missing caption}%\label{fig:figure2}
+	\end{figure}
+
+	\clearpage
+
+	\section{Old Figures and Figure Captions}
+
 \end{document}

### 2018-04-03 ident/python2/ss-ident/v5_script.py
change flux equation to valid root only

@@ -57,7 +57,7 @@ print('Practical Identifiability Analysis of v5 with 2 parameters: vemax and Kef
 ident_fun_choice = [2]
 # perform identifiability when v1 is written with k1cat*E in the numerator
 ident_details_v5_root1 = flux_ident_2_data_combination(experimental_datasets_2_expts, choose=combination_choice,
-                                                       flux_ids=[5], flux_choice=[1], ident_fun_choice=ident_fun_choice)
+                                                       flux_ids=[5], flux_choice=[2], ident_fun_choice=ident_fun_choice)
 print('Identifiability analysis of v5 with 2 parameters (vemax and Kefdp) complete.\n')
 
 # data processing - do not combine fluxes

### 2018-04-04 ident/docs/draft_1.tex
added additional figures for v5 and v3 variants where k3pep and k3fdp are constant

@@
 	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v5_1_2_ident}}
+		\caption{identifiability for flux v5 with 2 roots for Kefdp}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v5_1_2_data_utility}}
+		\caption{data utility for flux v5 with 2 roots for Kefdp}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure2/v5_1_2_exp_info}}
+		\caption{experiment information for flux v5 with 2 roots for Kefdp}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
 		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_data_utility}}
@@
 
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_1_2_ident}}
+		\caption{missing caption}%\label{fig:figure2}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v3_1_2_data_utility}}
+		\caption{missing caption}%\label{fig:figure2}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_exp_info}}
+		\caption{missing caption}%\label{fig:figure2}
+	\end{figure}
+
 	\clearpage

### 2018-04-05 ident/docs/draft_1.tex
significant changes to intro based on prof. M's comment

@@
 	\section{Introduction:}
-	Kinetic models of metabolism can be used to study the dynamic characteristics of metabolic networks \parencite{Bordbar2015, Andreozzi2016}.
+	The use of metabolic engineering spans a wide variety of applications, ranging from the design of microorganisms for the biosynthesis of commodity and specialty chemicals \parencite{Andreozzi2016} to engineering mammalian cells as therapeutic targets for cures to various ailments that affect humans \parencite{Bordbar2015,DiFilippo2016,Apaolaza2017}.
+These applications require us to understand the numerous complex interaction, their roles in cell function, and sometimes even the mechanisms behind these interactions.
+Computational models offer an avenue to study and understand these interactions through simplified representations of these interactions.
+They are also used to predict changes in cell function based on changes in the type and nature of the modeled interactions, or finding therapeutic targets to cure various ailments \textcolor{red}{(cite from old draft)}.
 
-	Unlike CBMs, where the ability to predict the steady state responses of metabolic networks is dependent on the stoichiometry of the network, the prediction of dynamic responses of metabolic networks to different perturbations using kinetic models of metabolism is also dependent on the numerical values of the enzyme kinetic parameters ($\theta$ in Equation \ref{eq:kinstoich}) \parencite{Andreozzi2016a}.
+	Constraint-based models (CBMs) of metabolism are one such category of models that are used to improve our understanding of the flow of flux through metabolism by representing it as a stoichiometric network of reactions that operate under a pseudo-steady state assumption.
+The ability of CBMs to shine light on the nonintuitive interactions that govern cellular metabolism is leveraged to predict the effect of different metabolic engineering design decisions on the ability of a cell to grow, or produce a desired metabolite (Bordbar, Monk, et al.
+2014; Maia, M.
+Rocha, and I.
+Rocha 2016).
+These advantages, combined with minimal number of parameters that are required to make CBMs functional, have made them useful to get an initial understanding of metabolism and its relationship to cellular function for various aforementioned applications \textcolor{red}{(CBM use references?
+cMCS, MCS may be?)}.
+
+	However, since metabolism is assumed to operate under a pseudo steady state, the metabolite concentrations within the metabolic network are assumed to be constant, and CBMs of metabolism do not model changes in metabolite concentrations.
+Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various non-catalytic regulatory interactions that are also responsible of metabolic function.
+These shortcomings prevent CBMs from being used to fully understand the steady state as well as the dynamic characteristics of metabolic networks.
+
+	Recently, the focus has shifted towards understanding the importance of the dynamics of metabolism in determining cell function \textcolor{red}{(recent kinetic model review about their utility)}.
+Consequently, it has been shown that using kinetic models that can represent the dynamics of the network can not only help us learn about previously unknown characteristics of metabolism \textcolor{red}{(red blood cell-both old and new, Maranas bifurcation paper)}, but can also improve predictions about the impact of engineering design perturbations on metabolism \textcolor{red}{(ecoli k457, k-optForce, Maranas kinetic model paper)}.
+\parencite{Bordbar2015, Andreozzi2016}.
+These kinetic models of metabolism, in contrast to their constraint-based counterparts, account for the thermodynamic and regulatory interactions that underly metabolic networks in addition to its stoichiometry \parencite{Andreozzi2016a}.
+
+	Kinetic models differ from CBMs is their use of mechanistic and empirical enzyme kinetic rate laws to model enzyme catalyzed fluxes within a metabolic network.
+The other aspect is the fact that kinetic models are heavily parameterized.
+These parameters represent various aspects of the enzyme kinetic rate laws \textcolor{red}{(kinetic model reviews, our review and a more recent one)}.
+Hence, the use of kinetic models requires information on all the enzyme kinetic rate laws that will be used to model all the fluxes within a metabolic network, as well as numerical values for all the parameters used in these rate laws.
+These aspects not only differentiate kinetic models from CBMs of metabolism, but are also two of the major hurdles to their widespread use and adoption.
+
+	Firstly, the development of these enzyme kinetic rate laws is based on observations of enzyme activity that are made in vitro.
+Accordingly, several authors have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions, as opposed to in vitro conditions \textcolor{red}{(Heijnen and Klamt)}.
+This problem is further compounded by the fact that typical parameter values used in kinetic models of metabolism are estimated in vitro, and may not be applicable under in vivo conditions.
+Any available in vivo parameter estimates are associated with large uncertainties.
+Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where parameter values might play a crucial role \parencite{Srinivasan2015, Vital-Lopez2006,Srinivasan2017}.
+Hence, the ability to study the steady state and dynamic responses of metabolic networks, under in vivo conditions, to different perturbations is compromised.
+The use of in vitro, or unreliable in vivo parameter estimates reduces confidence in the model predicted behaviour \parencite{Andreozzi2016a}.
+Consequently, the reduction in confidence hampers the use of these models to gain insight into metabolic network function \parencite{Chakrabarti2013a, Bordbar2015}.
+The corresponding increase in uncertainty in model predicted responses becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design.
+
+	Efforts to improve kinetic models have been dedicated to the development of newer enzyme kinetic models that have a higher correlation with what is observed in vivo.
+Authors have taken different approaches to address this problem.
+Some have focused on the development and improvement of parameter estimation methods based on optimization principles for obtaining true in vivo parameter values based on available in vivo experimental data without changing the underlying kinetic rate laws used in the model \textcolor{red}{Isabel Rocha papers on optimization methods, Japanese group that did parameter estimation using a HPC cluster}.
+Others have sought to completely abandon in vitro kinetic rate law descriptions in favour of approximate kinetic rate law models whose parameters are more informative and can be estimated in a relatively easier fashion \textcolor{red}{(Heijnen and other approximate kinetic modeling formats papers, Kotte 2010)}.
+
+	Despite these efforts, one of persistent problems is the lingering uncertainty in the parameter estimates and the subsequent predictions that are obtained from these methods.
+Some authors have proposed to alleviate this parameter uncertainty by using a Monte Carlo approach to obtaining kinetic models of metabolism \textcolor{red}{(ensemble modeling and ORACLE)}.
+Here, experimental steady state data is used as a reference point and kinetic models are built around this steady state value.
+Although a some have sought to quantify this uncertainty in parameter estimation through methods based on sensitivity analysis \textcolor{red}{(Nathan van Riel and Heijnen papers on uncertainty analysis)}, others have sought to approach parameter estimation from a Bayesian perspective \textcolor{red}{(Lars Nielsen papers)}.
+
+	These new kinetic models are either formulated on the basis of approximate enzyme kinetic rate laws to alleviate
+
+	In the absence of any in vivo parameter estimates, the norm is to either use available in vitro enzyme kinetic parameters, or use different parameter estimation techniques to determine parameter values based on either in vitro or in vivo data.
+So, the ability of these enzyme rate laws, and kinetic models built with these rate laws to study, understand and predict in vivo metabolic network function is limited \textcolor{red}{Heijnen}.
+
+	 but have also focused on methods to obtain parameter values that are applicable under in vivo conditions, as opposed to in vitro enzyme kinetic parameters.
+Research in the former area has resulted in the proposal of numerous approximate kinetic models of metabolism
+
+	the prediction of dynamic responses of metabolic networks to different perturbations using kinetic models of metabolism is dependent on numerous enzyme kinetic parameters.
+that are used to represent the changes in both concentrations and fluxes
+
+	the numerical values of the enzyme kinetic parameters ($\theta$ in Equation \ref{eq:kinstoich}) \parencite{Andreozzi2016a}.
 Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where parameter values might play a crucial role \parencite{Srinivasan2015, Vital-Lopez2006}.
-The use of in vitro, or unreliable in vivo parameter estimates reduces confidence in the model predicted behaviour \parencite{Andreozzi2016a}.
-Consequently, the reduction in confidence hampers the use of these models to gain insight into the functioning of metabolic networks \parencite{Chakrabarti2013a, Bordbar2015}.
-The corresponding increase in uncertainty in model predicted responses becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design.
 
@@
 A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 3360 combinations are used to practically identify c) $v_3$.
-Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
+Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
 	\end{figure}
@@
 The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
-Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:figure1}
+Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:ident}
 	\end{figure}
@@
 		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:figure2}
+The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:exp_info}
 	\end{figure}

### 2018-04-05 ident/docs/draft_1.tex
add new outline to intro

--- 2018-04-03 ident/docs/draft_1.tex
@@
 	\section{Introduction:}
+	\textbf{Outline:}
+	\begin{itemize}
+		\item Need for models in studying biology.
+State various applications e.g., for bio-based production, for medicine (gut microbiome interactions, stem cell engineering, studying various diseases like cancer)
+		\item Constraint-based models as a way to peek into the black box of biology without too many parameters. They however do not account for a lot of factors
+		\item Kinetics models as an improvement over constraint-based models
+		\item Advances in kinetic modeling of metabolism (Reviews?)
+		\item Parameter estimation in kinetic models as part of these advances
+	\end{itemize}
 	Kinetic models of metabolism can be used to study the dynamic characteristics of metabolic networks \parencite{Bordbar2015, Andreozzi2016}.
+In these models, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
+The matrix $\mathbf{S}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
+
+	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Link2014, Srinivasan2015}.
+Accordingly, $f$ is a nonlinear function of the metabolite concentrations, enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
 
@@
 The parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).
-	\begin{subequations}\label{eq:lsqopt}
-		\begin{align}
-		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{x_{kl}^*-x_{kl}}{\sigma_{kl}^*}\right)^2\\
-		&\theta_l \le \theta \le \theta_u
-		\end{align}
-	\end{subequations}
 
@@
 These additional parameters also need to be estimated.
-	\begin{subequations}\label{eq:output}
-		\begin{align}
-		\dot{x} = \mathbf{S}v\\
-		v = f(x, y, \theta, u)\\
-		\dot{y} = h(x, y, \theta, u)
-		\end{align}
-	\end{subequations}
+
 	In systems identification, the measured concentrations and fluxes ($y$) are called output or observed variables, and the unmeasured concentrations ($x$) are called the state variables.
@@
 	\section{Methods}\label{sec:methods}
-	\subsection{Kinetic models of metabolic networks}\label{sec:kinetic_model}
-	In kinetic models of metabolism, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
-The matrix $\mathbf{S}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
+
 	\begin{subequations}\label{eq:kinstoich}
@@
 	\end{subequations}
-	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Link2014, Srinivasan2015}.
-Accordingly, $f$ is a nonlinear function of the metabolite concentrations ($x$), enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
+
+	\begin{subequations}\label{eq:lsqopt}
+		\begin{align}
+		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{x_{kl}^*-x_{kl}}{\sigma_{kl}^*}\right)^2\\
+		&\theta_l \le \theta \le \theta_u
+		\end{align}
+	\end{subequations}
+
+	\begin{subequations}\label{eq:output}
+		\begin{align}
+		\dot{x} = \mathbf{S}v\\
+		v = f(x, y, \theta, u)\\
+		\dot{y} = h(x, y, \theta, u)
+		\end{align}
+	\end{subequations}
 
@@
 A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
-As indicated in Figure \ref{fig:ident-flowchart}a, the first step involves the construction of the kinetic model (Equation \ref{eq:kinstoich}) of the metabolic network with $n_r$ reaction fluxes.
-
-	%In kinetic models of metabolic networks (Equation \ref{eq:kinstoich}), the fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
-The value of every flux $v_i$, $i={1, 2, ..., n_r}$, is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
+In kinetic models of metabolic networks (Equation \ref{eq:kinstoich}), the fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
+The value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
 Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations (Figure \ref{fig:ident-flowchart}a).
 
-	For each flux $v_i$, $i={1, 2, ..., n_r}$, in the kinetic model, let $\mathbf{\theta} \in \mathbb{R}^p$ in Equation (\ref{eq:kinstoich}b).
-If data from $n_E$ experiments is available for the chosen metabolic network, as stated earlier, for each experiment $j = {1, 2, ..., n_E}$, we assume that all metabolite concentrations ($x$) and reaction fluxes ($v$) are measurable.
-We discuss the implications of relaxing this assumption later.
-The pertinent information for each experiment $j$ is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
+		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
+a) The steps for the construction of a kinetic model of a metabolic network.
+The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
+The identifiability of parameters for each flux can be established independently.
+b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
+	\end{figure}
 
-	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method.
-The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta}$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
-This is done by first selecting a combination of $p$ data sets from the $n_{E}$ available experimental data sets.
-The combination consisting of data from $p$ different experiments is then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
+	Let $\theta \in \mathbb{R}^p$ in Equation (\ref{eq:output}b) for each flux $v_i$ in the network.
+As stated earlier, for each experiment $j = {1, 2, ..., n}$, we assume that all metabolite concentrations ($x$) and reaction fluxes ($v$) are measurable.
+The pertinent information for each experiment is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
+
+	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, we describe a computer algebra-based method.
+The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\theta$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
+This is done by solving a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
 	\begin{align}\label{eq:nonlineq}
-	v_{i, j} = f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) && \forall j=\{1, 2, ..., p\}\subset\{1, 2, ..., n_E\}
+	v_{i, k} = f_k(\mathbf{x}_k,\theta, u_k) && \forall k=\{1, 2, ..., p\}\subset\{1, 2, ..., n\}
 	\end{align}
-	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$, $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$, whose elements are denoted by $\theta_k$.
-
-	Each equation in (\ref{eq:nonlineq}), indicated by the index $j$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for each $v_i$, $i={1, 2, ..., n_r}$, described in Equation (\ref{eq:kinstoich}b), written for concentrations ($\mathbf{x}_j$, $\mathbf{u}_j$) and fluxes ($v_{i,j}$) obtained from experiment $j$.
-Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^p$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
-Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from all chosen experiments $j = {1, 2, ..., p}$, in the selected combination.
+	Each equation in (\ref{eq:nonlineq}), indicated by the index $k$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for $v_i$, described earlier in Equation (\ref{eq:output}b), written for concentrations and fluxes obtained from experiment $k$.
+Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
 	\begin{align}\label{eq:theta-eq}
@@
 	\end{align}
-	The identifiability of parameter $\theta_k$, $k = {1, 2, ..., p}$, for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
+	The identifiability of parameter $\theta_k$ for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
 Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
-The solution obtained for $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_k$, and an infinite number of solutions are possible for a non-identifiable $\theta_k$.
-However, if there are multiple but finite solutions for $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, then the corresponding parameter $\theta_k$ is locally identifiable.
 
-	%For completeness, in the following sections we provide a previously published kinetic model of a small gluconeogenic network (section \ref{sec:small-model}).
-
-	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
-	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
-We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
-
-	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
-On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
-Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
-In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
-
-	Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
-Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
+	In the following sections we provide a previously published kinetic model of a small gluconeogenic network (section \ref{sec:small-model}), followed by a demonstration of our methodology to establish practical identifiability for one of the fluxes in this network (section \ref{sec:example}).
 
@@
 The kinetic model is described below.
-
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
+		\caption{The small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
+	\end{figure}	
 	\begin{equation}\label{eq:ode1}
@@
 	\begin{equation}\label{eq:ode3}
-	\frac{d}{dt}E=v_5 - d E
+	\frac{d}{dt}E=v_{e,max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right) - d E
 	\end{equation}
-	The kinetic expressions for fluxes $v_1$ through $v_5$ are given below.
+	The kinetic expressions for fluxes $v_1$ through $v_4$ are given below.
 The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics.
@@
 	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
-
-	The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (section \ref{sec:ident}).
-We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica or SymPy in Python).
-In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
-The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
-	\begin{align}\label{eq:flux3_convkin}
-	v_3 = V_3^{max}\left(\frac{1}{1 + \frac{K_3^{pep}}{pep}}\right)\left(\frac{\frac{fdp}{K_3^{fdp}}}{1 + \frac{fdp}{K_3^{fdp}}}\right)
-	\end{align}
-
 The flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
@@
 	\end{equation}
-	The production of enzyme E is represented by flux $v_5$.
-This flux is inhibited by \textit{fdp} and modeled using Hill kinetics, where $K_e^{fdp}$ represents the Hill binding constant for the inhibiting metabolite \textit{fdp}, $n_e$ is the Hill exponent, and $V_e^{max}$ is the maximum reaction rate for $v_5$.
-	\begin{align}\label{eq:flux5}
-	v_5 = V_e^{max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right)
+
+	\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis}\label{sec:example}
+	Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_1$ in the small model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model}.
+
+	In flux $v_1$, the concentration of the enzyme E is used as a variable.
+Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, the expression previously given for $v_1$ (Equation \ref{eq:flux1}) cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
+So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
+Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
+The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
+	\begin{align}\label{eq:flux1a}
+	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
 	\end{align}		
+	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability.
+
+	As mentioned in section \ref{sec:intro}, practical identifiability is a necessary condition to estimate true values of both $V_1^{max}$ and $K_1^{ac}$ from experimental data.
+Here, we assume that the concentrations and fluxes from at least two different experiments are available i.e., in Equation (\ref{eq:nonlineq}) $k = \{1, 2\}$ .
+We label the available concentrations and fluxes as ${ac}^{(k)}$ and ${v_1}^{(k)}$, respectively.
+Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
+	\begin{align}\label{eq:nonlin-flux2}
+	{v_1}^{(k)} = V_{1}^{max}\frac{ac^{(k)}}{ac^{(k)}+K_{1}^{ac}(ne)} &&  k=\{1, 2\}
+	\end{align}
+
+	Solving this simultaneous system of $k$ equations using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
+These expressions have the form previously shown in Equation (\ref{eq:theta-eq}).
+	\begin{subequations}\label{eq:v1_par}
+		\begin{align}		
+		V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
+		K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
+		\end{align}
+	\end{subequations}
+	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations.
+Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
 
@@
 	In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
-The 21 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
+The 18 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
+	\begin{table}[!thbp]
+		\caption{Table showing the perturbed values of all parameters used to generate experimental data in silico for testing practical identifiability of all fluxes in the small metabolic network.}
+		\begin{center}				
+			\begin{tabular}{lp{3.5cm}cccccp{2cm}}
+				\hline
+				Experiment Type & Perturbed Parameter (wild type value a.u.)& \multicolumn{5}{c}{Perturbed Values a.u.} & Number of Experiments\\
+				\hline
+				wild type &  &\multicolumn{5}{l}{} & 1\\
+				acetate perturbation & $acetate$ (0.1) & 0.05 & 0.09 & 0.2 & 0.5 & 1.0 & 5\\
+				$v_1$ perturbation & $k_1^{cat}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\				
+				$v_2$ perturbation & $V_2^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\								
+				$v_3$ perturbation & $V_3^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\
+				\hline
+				\multicolumn{7}{l}{Total number of experimental data sets used for practical parameter identification:} & 21\\
+				\hline
+			\end{tabular}
+		\end{center}	
+		\label{tab:pval}
+	\end{table}
 
-	As described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, the minimum number of experiments from which steady state data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
-For instance, the flux $v_2$ in the model described in section \ref{sec:small-model} has two parameters: $V_2^{max}$ and $K_2^{pep}$.
-Thus $\theta = \{V_2^{max}, K_2^{pep}\} \in \mathbb{R}^2$.
-So, steady state data from two distinct experiments is required for identifying $v_2$.
-Accordingly, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_2$.
-The total number of such possible combinations is 420 (21 x 20) from the 21 different experiments shown in Table \ref{tab:pval}.
-Assuming we also have data for the enzyme concentration $E$ in the expression for flux $v_1$ (Equation \ref{eq:flux1}), identifying the two parameters $k_1^{cat}$ and $K_1^{ac}$ ($\theta \in \mathbb{R}^2$) also requires steady state experimental data from two different experiments.
+	The minimum number of experiments from which data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
+For instance, as demonstrated above, data from two distinct experiments is required for identifying $v_1$, as $v_1$ has two parameters ($\mathbb{R}^2$).
+$v_2$ is also modeled by a parameter space with dimensions of $\mathbb{R}^2$.
+Hence, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_1$ and $v_2$.
+The total number of such possible combinations is 306 (18 x 17) from the 18 different experiments shown in Table \ref{tab:pval}.
+Similarly, for identifying $v_3$, that is described by three parameters ($\mathbb{R}^3$), we need data from a combination of three different experiments.
+Based on the 18 experiments in Table \ref{tab:pval}, we have 4896 distinct combinations (18 x 17 x 16) of three experiments.
 
-	Similarly, for identifying $v_3$ that is described by three parameters ($\theta \in \mathbb{R}^3$), we need data from a combination of three different experiments.
-Based on the 21 experiments in Table \ref{tab:pval}, we have 7980 distinct combinations (21 x 20 x 19) of three experiments.
-However, it is important to note that not all of the either 420 combinations (for $v_1$ and $v_2$), or the 7980 combinations (for $v_3$), may be suitable for parameter estimation.
-Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers.
-Moreover, among these chosen smaller number of experimental data combinations, our methodology enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation.
-The following section describes a procedure that can be used to design experiments within this context.
+	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
+	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
+We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
+
+	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
+On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
+Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
+In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
+Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
+Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
 
 	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
+		\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
+	\end{figure}
 	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
-This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguish experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
+This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguishing experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
 In doing so, it is possible to obtain a minimal and informative collection of experiments that can be performed to identify as many model parameters as possible (Figure \ref{fig:ident-design}).
-Consequently, the set of experiments can be used to estimate all the identifiable parameters in the model.
+Consequently, the set of experiments can be used to estimate the identifiable parameters in the model.
 This is formally explained below.
 
-	The identifiability of each parameter based on each experiment indexed as $j = \{1, 2, ..., n_E\}$ is established based on the methodology described in section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in section \ref{sec:example}.
-Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
-Consequently, the corresponding experiment cannot be used for estimating parameter $\theta_k$, and needs to be discarded from the set of all necessary experiments.
-Furthermore, another experiment from $j = \{1, ..., n_E\}$ needs to be selected to replace the discarded experiment such that parameter $\theta_k$ is identifiable.
+	The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in section \ref{sec:ident} and demonstrated in Section \ref{sec:example} (Figure \ref{fig:ident-flowchart}b).
+Subsequently, for any flux $v_i$ and for any $p$ combinations of indices $j$, if the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
+Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
+Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
 This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
 In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
-Note that if none of the $n_E$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n_E+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
+Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
 This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.
@@
 	\section{Results:}
-	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model}.
-
 	First in section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
 We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
-
-	%In the following section we demonstrate our methodology to establish practical identifiability for one of the fluxes in this network (section \ref{sec:example}) (move to results).
-
 	Then, in section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models of metabolic networks.
@@
 
-	\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis}\label{sec:example}	
-	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}).
-Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression previously given for $v_1$ (Equation \ref{eq:flux1}) cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
-So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
-Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
-The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
-	\begin{align}\label{eq:flux1a}
-	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
+	\subsection{Nonlinearity of enzyme kinetic rate law expression affects identifiability analysis}\label{sec:closedform}
+	Practical identifiability of parameters in kinetic models of metabolism using the methodology described in section \ref{sec:methods} is governed by the nonlinear complexity of the enzyme kinetic rate law used to model a specific flux.
+We demonstrate one example of how the methodology works in section \ref{sec:example} for parameters of flux $v_1$.
+The expression in Equation (\ref{eq:v1_par}) is obtained by using a computer algebra system.
+To recall, in this specific case, a computer algebra system is used to solve for the parameters of a flux described using the Michaelis-Menten kinetic rate law when data from two different experiments is available.
+
+	However, we find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica and SymPy in Python).
+In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
+The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
+
+	\begin{align}\label{eq:flux3_convkin}
+	v_3 = V_3^{max}\left(\frac{1}{1 + \frac{K_3^{pep}}{pep}}\right)\left(\frac{\frac{fdp}{K_3^{fdp}}}{1 + \frac{fdp}{K_3^{fdp}}}\right)
 	\end{align}
-	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability.
 
-	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ still holds for $v_1$, as mentioned in section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
-So, from the $n_E = 21$ different experiments described in section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$.
-We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively.
-Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
-	\begin{align*}%\label{eq:nonlin-flux2}
-	{v_1}^{(j)} = V_{1}^{max}\frac{ac^{(j)}}{ac^{(j)}+K_{1}^{ac}(ne)} &&  j=\{1, 2\}
-	\end{align*}
-
-	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
-These expressions have the form shown in Equation (\ref{eq:theta-eq}).
-	\begin{subequations}\label{eq:v1_par}
-		\begin{align}		
-		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
-		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
-		\end{align}
-	\end{subequations}
-	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations.
-Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
-
-	%As mentioned in section \ref{sec:intro}, practical identifiability is a necessary condition to estimate true values of both $V_1^{max}$ and $K_1^{ac}$ from experimental data. 	
-	%Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_1$ in the small model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model}.
-
-	\subsection{Nonlinearity of enzyme kinetic rate law expression affects identifiability analysis}\label{sec:closedform}
-	In the previous section, we have demonstrated our methodology to practically identify parameters for a flux that is modeled using the Michaelis-Menten (MM) kinetic rate law when steady state concentration and flux data from two different experiments are available.
-However, metabolic fluxes can be described by several different kinetic rate law expressions based on their representation of the physiological interactions that facilitate enzyme catalysis of each flux in a metabolic network.
-For instance, as described in section \ref{sec:small-model}, flux $v_3$, which is allosterically activated by \textit{pep}, is modeled using the MWC kinetics (Equation \ref{eq:flux3}).
-So, we comment on the capability of our methodology to practically identify parameters in cases where metabolic fluxes are modeled using rate laws that are different from the MM model, and consequently differ in terms of their nonlinear complexity.
-
-	In using the convenience kinetic rate law model for $v_3$ (Equation \ref{eq:flux3_convkin}) for identifying the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$, we found that each of the three parameters have two different closed-form expressions owing to the presence of a square root term in their solutions.
-Thus, as per the stated definition in section \ref{sec:ident}, the parameters for $v_3$, described using the convenience kinetics rate laws, are only locally identifiable due to the non-unique but finite number of possible solutions for the three parameters.
-The two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, and $K_3^{pep}(1)$ refer to the first set of solutions, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$ refer to the other set of solutions.
-
-	However, we found that the local identifiability of the parameters could be resolved to obtain globally identifiable parameters if either the Michaelis binding constant for \textit{fdp}, $K_3^{fdp}$, or the allosteric binding constant for \textit{pep}, $K_3^{pep}$, is known.
-Thus, fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter.
-Although one could argue that a similar approach could be used to identify parameters of $v_3$ written using the MWC model, we found that it was not possible to obtain identifiability expressions for any of the parameters in the MWC model, except for $V_3^{max}$ when all other parameters are assumed to be known.
-
-	Similarly, not all parameters in the Hill equation, used to model the inhibition of \textit{fdp} on the production of the enzyme E through flux $v_5$ (Equation \ref{eq:flux5}), can be identified using our methodology.
-We found that the Hill exponent ($n_e$) needs to be fixed to be able to identify the other two parameters, $V_e^{max}$ and $K_e^{fdp}$, in the model.
-Thus, the dimension of the parameter space for $v_5$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have developed in this paper.
-However, on applying the primary test for identifiability ($D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, see section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b), we found only $V_e^{max}$ to globally identifiable.
-We found $K_e^{fdp}$ to be locally identifiable to the possibility of having two different roots.
-As in the case of parameters for $v_3$ when using the convenience kinetics formulation, the presence of a square root term was the reason for the existence of two roots for $K_e^{fdp}$.
-However, on applying the secondary criteria for identifiability ($g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, see section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b), we were able to eliminate the existence of the second root as it is always negative, and therefore not physically relevant.
-Thus, we were able to determine that both parameters modeling $v_5$ ($V_e^{max}$ and $K_e^{fdp}$) are globally practically identifiable under certain conditions.
+	Using this expression for analysis, we find that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different closed-form expressions owing to the presence of a square root term in their solutions.
+These distinct expressions are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, $K_3^{pep}(1)$, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$.
 
@@
 However, the identifiability of the parameters used in these models has never been truly examined.
-\textcolor{red}{mentione Heijnen papers on resolving identifiability using approximate models here}.
-
 Unlike the case of the MWC model, if determining closed-form expressions for the parameters of these models are tractable, we believe our methodology can help in elucidating the identifiability of these parameters.
@@
 	Next, we look at some of the results we obtained for the practical identification of different parameters for all three fluxes $v_1$, $v_2$ and $v_3$ in the small gluconeogenic network model described in section \ref{sec:small-model}, with the flux for $v_3$ described by Equation (\ref{eq:flux3_convkin}).
+We start by looking at one of motivations for the need for designing experiments to collect data for parameter estimation.
 
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure4/v1_V1max_v2_v3_value1_data_utility}}
+		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
+The total possible number of parameters that can be identified by data from combinations of a) two, b) two, and c) three experiments is shown on the horizontal axis.
+The vertical axis represents the total number of combinations that can identify the corresponding number of parameters.
+The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
+A total of 306 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 4896 combinations are used to practical identifiability of c) $v_3$.
+Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
+	\end{figure}	
 	\subsection{Utility of experimental data for practical parameter identification}\label{sec:data_utility}
-	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not been well studied (\textcolor{red}{include Rudiyanto Gunawan papers here}).
+	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community, the need for experimental design methods to satisfy the data needs for parameter estimation are not considered seriously.
 Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
@@
 
-	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:figure4} also informs the ease of identifying parameters for a single flux.
+	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:figure4} also informs on the ease of identifying parameters for a single flux.
 For the small network (Figure \ref{fig:network}), we find that more than 90\% of all the data combinations tested for their ability to identify $v_2$ are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:figure4}b).
@@
 	\subsection{Maximum reaction rates are more identifiable than enzyme binding constants}\label{sec:trends}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure1/v1_V1max_v2_v3_identifiability}}
+		\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
+The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
+Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:figure1}
+	\end{figure}
 	In Figure \ref{fig:figure1} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
@@
 Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range.
-%Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
+Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
 In linear systems, persistence of excitation of any input signal can be theoretically guaranteed.
@@
 
-	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}	
+	\subsection{Experiment requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}	
 	In this section, we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model.
@@
 	In order to determine the type of experiments that help satisfy identifiability conditions for both parameters of $v_1$, we look at the distribution of the five different types in all data combinations that can identify these 2 parameters (Figure \ref{fig:figure2}a and b).
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_only_v2_spider}}
+		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
+The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:figure2}
+	\end{figure}
 
@@
 	\printbibliography
-
-	\clearpage
-	\section{Tables, Figures and Figure Captions:}
-	\begin{table}[!thbp]
-		\caption{Table showing the perturbed values of all parameters used to generate experimental data in silico for testing practical identifiability of all fluxes in the small metabolic network.}
-		\begin{center}				
-			\begin{tabular}{lp{3.5cm}cccccp{2cm}}
-				\hline
-				Experiment Type & Perturbed Parameter (wild type value a.u.)& \multicolumn{5}{c}{Perturbed Values a.u.} & Number of Experiments\\
-				\hline
-				wild type &  &\multicolumn{5}{l}{} & 1\\
-				acetate perturbation & $acetate$ (0.1) & 0.05 & 0.09 & 0.2 & 0.5 & 1.0 & 5\\
-				$v_1$ perturbation & $k_1^{cat}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\				
-				$v_2$ perturbation & $V_2^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\								
-				$v_3$ perturbation & $V_3^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\
-				\hline
-				\multicolumn{7}{l}{Total number of experimental data sets used for practical parameter identification:} & 21\\
-				\hline
-			\end{tabular}
-		\end{center}	
-		\label{tab:pval}
-	\end{table}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.9\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
-		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
-a) The steps for the construction of a kinetic model of a metabolic network.
-The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
-The identifiability of parameters for each flux can be established independently.
-b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
-		\caption{The small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure3/experimental_design}}
-		\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure4/v1_v1max_v2_v3_1_data_utility}}
-		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
-The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, and c) three experiments.
-The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 3360 combinations are used to practically identify c) $v_3$.
-Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v1max_v2_v3_1_2_ident}}
-		\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
-The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
-Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:figure1}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_v1max_v2_exp_info}}
-		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:figure2}
-	\end{figure}
-
-	\clearpage
-
-	\section{Potential Supplementary Information Figures}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_data_utility}}
-		\caption{Same as other data utility figure for v1 with k1cat as parameter (with enzyme data) and second roots of v3.
-Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$.}\label{fig:figure4s}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_k1cat_exp_info}}
-		\caption{missing caption}%\label{fig:figure2}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_1_2_exp_info}}
-		\caption{missing caption}%\label{fig:figure2}
-	\end{figure}
-
-	\clearpage
-
-	\section{Old Figures and Figure Captions}
-
 \end{document}

### 2018-04-06 ident/docs/draft_1.tex
a lot of changes to intro and methods. Intro nearing completion

--- 2018-04-04 ident/docs/draft_1.tex
@@
 	\section{Introduction:}
-	Kinetic models of metabolism can be used to study the dynamic characteristics of metabolic networks \parencite{Bordbar2015, Andreozzi2016}.
+	The use of metabolic engineering spans a wide variety of applications.
+Some notable examples include the design of microorganisms for the biosynthesis of commodity and specialty chemicals \parencite{Andreozzi2016}, engineering mammalian cells as therapeutic targets for cures to some ailments affecting humans \parencite{DiFilippo2016,Apaolaza2017}, and changing the constituents of the human gut microbial community to cure related diseases \parencite{Zerfab2018}.
+These applications require us to understand the numerous complex interaction, their roles in cell function, and sometimes even the mechanisms behind these interactions \parencite{Pushalkar2018}.
+Computational models offer a systematic way to integrate available experimental data, and to study and understand these interactions through mathematical representations of the biological systems in which these interactions occur \parencite{Bordbar2014a,Saa2017}.
+They are also used to predict changes in cell function based on changes in the type and nature of the modeled interactions \parencite{Andreozzi2016}, or aid in the identification of therapeutic targets for drug discovery and development \parencite{Bordbar2015,Chandrasekaran2017}
 
-	Unlike CBMs, where the ability to predict the steady state responses of metabolic networks is dependent on the stoichiometry of the network, the prediction of dynamic responses of metabolic networks to different perturbations using kinetic models of metabolism is also dependent on the numerical values of the enzyme kinetic parameters ($\theta$ in Equation \ref{eq:kinstoich}) \parencite{Andreozzi2016a}.
-Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where parameter values might play a crucial role \parencite{Srinivasan2015, Vital-Lopez2006}.
-The use of in vitro, or unreliable in vivo parameter estimates reduces confidence in the model predicted behaviour \parencite{Andreozzi2016a}.
-Consequently, the reduction in confidence hampers the use of these models to gain insight into the functioning of metabolic networks \parencite{Chakrabarti2013a, Bordbar2015}.
-The corresponding increase in uncertainty in model predicted responses becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design.
+	Constraint-based models (CBMs) of metabolism are one such category of models that are used to improve our understanding of metabolism by representing it as a stoichiometric network of reactions that operate under a pseudo-steady state assumption \parencite{Bordbar2014a}.
+The ability of CBMs to shine light on the nonintuitive interactions that govern cellular metabolism is leveraged to engineer and asses the impact of designs that alter the ability of a cell to grow, or produce a desired metabolite \parencite{Maia2016}.
+However, in CBMs, since metabolism is assumed to operate under a pseudo steady state, the metabolite concentrations within the metabolic network are assumed to be constant, and changes in metabolite concentrations are not modeled.
+Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various non-catalytic regulatory interactions that are also responsible of metabolic function.
+These shortcomings prevent CBMs from being used to fully understand the steady state as well as the dynamic characteristics of metabolic networks.
 
-	Parameter estimation methods based on optimization principles are typically used to determine true parameter values based on available experimental data.
-Under the assumption that all time dependent intracellular metabolite concentrations can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:kinstoich}) to estimate the values of enzyme kinetic parameters, $\theta$, based on the measured data.
-	The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration at each time point, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}).
-The parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).
-	\begin{subequations}\label{eq:lsqopt}
-		\begin{align}
-		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{x_{kl}^*-x_{kl}}{\sigma_{kl}^*}\right)^2\\
-		&\theta_l \le \theta \le \theta_u
-		\end{align}
-	\end{subequations}
+	Recently, the focus has shifted towards understanding the implications of metabolite concentrations and associated regulatory interactions on the steady state and the dynamic characteristics of metabolism \parencite{Saa2017}.
+Kinetic models accounting for the thermodynamic and regulatory interactions that underly metabolic networks in addition to its stoichiometry are used for this purpose \parencite{Link2014}.
+These models can not only help us learn about previously unknown characteristics of metabolism \parencite{Kotte2014}, but can also improve predictions about the impact of engineering design perturbations on metabolism, and propose alternative designs to achieve various goals \parencite{Andreozzi2016,Khodayari2016}.
 
-	However, not all metabolite concentrations used in the model (Equation \ref{eq:kinstoich}) can be measured.
-Additionally, measurable fluxes in the metabolic network also need to be included as part of the parameter estimation problem.
-In such scenarios, the parameter estimation problem is modified to suit a new system of equations shown below (Equation \ref{eq:output}).
-The new system of equations is obtained by augmenting the original system (Equation \ref{eq:kinstoich}) with Equation (\ref{eq:output}c) that models the relationship between the measurable metabolite concentrations and fluxes ($y$) and the unmeasured concentrations ($x$) that are used in the original model (Equation \ref{eq:kinstoich}) above.
-The parameter vector ($\theta$) is augmented with additional parameters that define this relationship.
-These additional parameters also need to be estimated.
-	\begin{subequations}\label{eq:output}
-		\begin{align}
-		\dot{x} = \mathbf{S}v\\
-		v = f(x, y, \theta, u)\\
-		\dot{y} = h(x, y, \theta, u)
-		\end{align}
-	\end{subequations}
-	In systems identification, the measured concentrations and fluxes ($y$) are called output or observed variables, and the unmeasured concentrations ($x$) are called the state variables.
-For estimating $\theta$, the metabolite concentrations $x$ in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the output variables $y$.
+	Kinetic models differ from CBMs in their use of heavily parameterized mechanistic enzyme kinetic rate laws to model enzyme catalyzed fluxes within a metabolic network.
+These parameters represent various aspects of the enzyme kinetic rate laws \parencite{Srinivasan2015,Saa2017}.
+Hence, the use of kinetic models requires information on all the enzyme kinetic rate laws that will be used to model all the fluxes within a metabolic network, as well as numerical values for all the parameters used in these rate laws.
+Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values might play a crucial role \parencite{Vital-Lopez2006,Srinivasan2017}.
 
-	The ability to determine unique solutions to parameters $\theta$ is governed by the identifiability of these parameters in the model \parencite{McLean2012, Raue2009a}.
-The identifiability of parameters in nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability.
-	Any system (Equation \ref{eq:output}) is said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
-	\begin{align}\label{eq:stident}
-	\Phi(\theta_1,u) = \Phi(\theta_2,u) \iff \theta_1 = \theta_2
-	\end{align}
-	Accordingly, the system can have a unique solution, a finite number of non-unique solutions or an infinite number of solutions for all input functions, and is said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
-So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unobservable state variables and the observable output variables.
-Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
+	The development of these enzyme kinetic rate laws is based on in vitro observations of enzyme activity.
+Accordingly, authors have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions, as opposed to in vitro conditions \parencite{Heijnen2013}.
+This problem is further compounded by the fact that typical parameter values used in kinetic models of metabolism are estimated in vitro, and may not be applicable under in vivo conditions \parencite{Heijnen2005,Smallbone2007}.
+Moreover, estimation of kinetic parameters based on experimental in vivo data becomes computationally expensive for larger models, and any estimated in vivo parameter estimates are associated with large uncertainties \parencite{Link2014}.
+Using in vitro, or unreliable in vivo parameter estimates to study the steady state and dynamic responses of metabolic networks, under in vivo conditions, to different perturbations reduces confidence in the model predicted behaviour \parencite{Andreozzi2016a}.
+Consequently, this hampers the use of these models to gain insight into metabolic network function \parencite{Chakrabarti2013a}, and the increase in prediction uncertainty becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design \parencite{Saa2017}.
 
-	Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
-So, the effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
-Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
+	So, the use of heavily parameterized enzyme kinetic rate laws not only differentiates kinetic models from CBMs of metabolism, but is also a major hurdle to the widespread use and adoption of kinetic models \parencite{Berthoumieux2013,Vasilakou2016}.
+In order to address this problem, authors have proposed two different strategies.
+In one scenario, authors have looked at the development and use of global optimization algorithms for parameter estimation \parencite{Moles2003}.
+In another case, in vitro kinetic rate law descriptions are abandoned in favour of approximate and new empirical enzyme kinetic models whose parameters can not only be estimated in a relatively easier fashion, but also lead to kinetic models that generate information that has a higher correlation with experimental data that is observed in vivo \parencite{Heijnen2005,Smallbone2007}.
 
-	Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated from measurable data.
-On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
+	However, one of persistent problems is the lingering uncertainty in the parameter estimates and the subsequent predictions that are obtained from these methods.
+Authors have proposed to alleviate this parameter uncertainty either by using a Monte Carlo approach to kinetic modeling of metabolism \parencite{Srinivasan2015}, or by quantifying this uncertainty in parameter estimates using different methods \parencite{Vanlier2013}.
+Approaches to parameter estimation from a Bayesian perspective have also been proposed to estimate parameters and quantify the uncertainty in these estimates \parencite{Saa2016}.
 
-	Methods and tools for structural identification of parameters based on differential algebra \parencite{Ljung1994, Audoly2001, Bellu2007} and profile likelihood \parencite{Raue2009a} are available.
-However, only the profile likelihood-based methods enable experimental design by facilitating practical identification of parameters.
-Nonetheless, this method still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters, and hence still suffers from all the inherent difficulties associated with obtaining global optimal solutions for non-convex optimization problems.
-This also makes it un-scalable for experimental design and practical identifiability of parameters in kinetic models of large metabolic networks.
+	In spite of all these developments, model parameter identifiability, a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.
+Briefly, it concerns with the ability to distinguish between any two different parameter values based on differences in the inputs to, and consequently the responses exhibited by the cell.
+Structural identifiability refers to the dependence of the ability to estimate unique parameter values on model structure, due to inherent dependencies between two or more parameters in the model \parencite{Berthoumieux2013}.
+Conversely, when the dependence is on the informativeness of the data used for estimation, and not on the structure of the model, it is referred to as practical identifiability \parencite{Ljung1994}.
 
-	In this paper, we propose a scalable methodology to establish practical identifiability for parameters in kinetic models of metabolism using steady state concentration and flux data.
-We present a computer algebra-based method that can facilitate experimental design through practical identification of parameters separately for each individual reaction within a metabolic network.
+	Significant work has been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Ljung1994,Nikerel2009,Berthoumieux2013,Raue2014}\textcolor{red}{(paper from Rudiyanto Gunawan on model discrimination and sensitivity analysis)}.
+However, the focus on methods for practical identification of parameters in kinetic models of metabolism and its use for experimental design is fairly recent and limited \parencite{Gadkar2005a,Vanlier2014a,Raue2014}.
+For instance some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, and some perform identifiability within a dynamic context \parencite{Gadkar2005a,Nikerel2009,Vanlier2014a}.
+Some other methods also suffer from their inability to be applied to kinetic models of large metabolic networks due to the associated computational burden \parencite{Gadkar2005a,Raue2014}\textcolor{red}{(Banga method using FIM for D-optimal design, ??)}.
+
+	In this paper, we propose a scalable methodology that uses available steady state fluxomics, metabolomics and proteomics data to practically identify parameters for each individual reaction in kinetic models of metabolism.
+We demonstrate how the computer algebra-based method that we have developed can also facilitate the design of experiments that are minimal and informative to generate data required to estimate unique parameter values for all reaction fluxes in a metabolic network.
+In doing so, we not only propose specific types of and nature of perturbations that will provide the most useful data for parameter estimation, but also test the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metbaolic networks.
+
 	 For the purposes of this method we assume that all intracellular metabolite concentrations and fluxes can be measured.
-We identify and design experiments to estimate parameters in a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} to illustrate the utility of our method.
+We illustrate our methodology to identify parameters and design experiments for estimating parameters in a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017}.
 %We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
 
+ %In \textcolor{red}{Section x} we provide detailed definitions for structural and practical identifiability of parameters followed by a step by step description of our proposed methodology for practical parameter identification in kinetic models of metabolism.
+In \textcolor{red}{Section y} we discuss the implications
+
+% We explain these identifiability conditions in detail in the Methods section. 
+%Bayesian approach to identification and experimental design \parencite{Vanlier2012}.
+
 	\section{Methods}\label{sec:methods}
-	\subsection{Kinetic models of metabolic networks}\label{sec:kinetic_model}
+	\subsection{Definition: Structural and Practical parameter identifiability of kinetic models of metabolism}\label{sec:kinetic_model}
 	In kinetic models of metabolism, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
@@
 	\end{subequations}
-	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Link2014, Srinivasan2015}.
+	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Srinivasan2015}.
 Accordingly, $f$ is a nonlinear function of the metabolite concentrations ($x$), enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
 
+	Parameter estimation methods based on optimization principles are typically used to determine true parameter values based on available experimental data.
+Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:kinstoich}) to estimate the values of enzyme kinetic parameters, $\theta$, based on the measured data.
+	\begin{subequations}\label{eq:lsqopt}
+		\begin{align}
+		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2\\
+		&\theta_l \le \theta \le \theta_u
+		\end{align}
+	\end{subequations}	
+	The minimization of least square error between the measured ($y^*$) and modeled ($y$) concentrations and fluxes, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration and flux, at each time point, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}).
+The parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).
+
+	However, not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment.
+This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}.
+In such scenarios, the need to design experiments to effect a change in and discriminate between changes in different concentrations/fluxes becomes necessary.
+
+	As briefly mentioned in the Introduction, in systems identification, the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
+The identifiability of parameters in linear and nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability.
+
+	Mathematically, the parameters $\theta$ in any nonlinear model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
+	\begin{align}\label{eq:stident}
+	\Phi(\theta_1,u) = \Phi(\theta_2,u) \iff \theta_1 = \theta_2
+	\end{align}
+	Accordingly, parameters in $\theta$ can have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, and are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
+So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unmeasured and measured concentrations/fluxes, as well as correlations between different model parameters.
+Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
+
+	Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values with quantifiable uncertainties to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
+So, the effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
+Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
+
+	Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated from measurable data.
+On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
+
 	\subsection{A method to determine practical identifiability of kinetic models of metabolism}\label{sec:ident}
-	We provide the mathematical framework for practical identification of parameters in kinetic models of metabolism in this section.
+	We provide the mathematical framework for \textcolor{red}{practical identification} of parameters in kinetic models of metabolism in this section.
 A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
@@
 Accordingly, in addition to measuring network response to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
-
 In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
@@
 
-	\section{Old Figures and Figure Captions}
+	\section{Old Figures, Figure Captions and other miscellaneous text}	
+	Methods and tools for structural identification of parameters based on differential algebra \parencite{Ljung1994, Audoly2001, Bellu2007} and profile likelihood \parencite{Raue2014} are available.
+However, only the profile likelihood-based methods enable experimental design by facilitating practical identification of parameters.
+Nonetheless, this method still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters, and hence still suffers from all the inherent difficulties associated with obtaining global optimal solutions for non-convex optimization problems.
+This also makes it un-scalable for experimental design and practical identifiability of parameters in kinetic models of large metabolic networks.
 

### 2018-04-06 ident/docs/draft_2.tex
change draft file name from 1 to 2

--- 2018-04-06 ident/docs/draft_1.tex
@@
 Some notable examples include the design of microorganisms for the biosynthesis of commodity and specialty chemicals \parencite{Andreozzi2016}, engineering mammalian cells as therapeutic targets for cures to some ailments affecting humans \parencite{DiFilippo2016,Apaolaza2017}, and changing the constituents of the human gut microbial community to cure related diseases \parencite{Zerfab2018}.
-These applications require us to understand the numerous complex interaction, their roles in cell function, and sometimes even the mechanisms behind these interactions \parencite{Pushalkar2018}.
+These applications require us to understand the numerous complex interaction, their roles in cell function, and sometimes even the mechanisms behind these interactions.
 Computational models offer a systematic way to integrate available experimental data, and to study and understand these interactions through mathematical representations of the biological systems in which these interactions occur \parencite{Bordbar2014a,Saa2017}.
@@
 
-	Constraint-based models (CBMs) of metabolism are one such category of models that are used to improve our understanding of metabolism by representing it as a stoichiometric network of reactions that operate under a pseudo-steady state assumption \parencite{Bordbar2014a}.
+	Constraint-based models (CBMs) of metabolism are used to improve our understanding of metabolism by representing it as a stoichiometric network of reactions that operate under a pseudo-steady state assumption \parencite{Bordbar2014a}.
 The ability of CBMs to shine light on the nonintuitive interactions that govern cellular metabolism is leveraged to engineer and asses the impact of designs that alter the ability of a cell to grow, or produce a desired metabolite \parencite{Maia2016}.
-However, in CBMs, since metabolism is assumed to operate under a pseudo steady state, the metabolite concentrations within the metabolic network are assumed to be constant, and changes in metabolite concentrations are not modeled.
+However, in CBMs, metabolism is assumed to operate under a pseudo steady state.
+Consequently, the metabolite concentrations within the metabolic network are assumed to be constant, and changes in metabolite concentrations are not modeled.
 Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various non-catalytic regulatory interactions that are also responsible of metabolic function.
@@
 
-	Recently, the focus has shifted towards understanding the implications of metabolite concentrations and associated regulatory interactions on the steady state and the dynamic characteristics of metabolism \parencite{Saa2017}.
-Kinetic models accounting for the thermodynamic and regulatory interactions that underly metabolic networks in addition to its stoichiometry are used for this purpose \parencite{Link2014}.
-These models can not only help us learn about previously unknown characteristics of metabolism \parencite{Kotte2014}, but can also improve predictions about the impact of engineering design perturbations on metabolism, and propose alternative designs to achieve various goals \parencite{Andreozzi2016,Khodayari2016}.
+	In contrast, the implications of regulatory interactions and changes in metabolite concentrations on different characteristics of metabolism can be studied using kinetic models of metabolism \parencite{Saa2017}.
+These models account for the thermodynamic and regulatory interactions that underly metabolic networks in addition to its stoichiometry \parencite{Link2014}.
+Kinetic models can not only help us better understand lesser known and understood characteristics of metabolism like bistability \parencite{Kotte2014}, and their role in human health, but can also improve predictions about the impact of engineering design perturbations on metabolism, and propose alternative designs to achieve metabolite production goals \parencite{Khodayari2016}.
 
@@
 Hence, the use of kinetic models requires information on all the enzyme kinetic rate laws that will be used to model all the fluxes within a metabolic network, as well as numerical values for all the parameters used in these rate laws.
-Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values might play a crucial role \parencite{Vital-Lopez2006,Srinivasan2017}.
+Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values might play a crucial role \parencite{Srinivasan2017}.
 

### 2018-04-06 ident/docs/draft_2.tex
more changes to intro

@@
 	In contrast, the implications of regulatory interactions and changes in metabolite concentrations on different characteristics of metabolism can be studied using kinetic models of metabolism \parencite{Saa2017}.
-These models account for the thermodynamic and regulatory interactions that underly metabolic networks in addition to its stoichiometry \parencite{Link2014}.
+These models account for changes in metabolite concentrations subject to thermodynamic and regulatory constraints that underly metabolic networks in addition to its stoichiometry \parencite{Link2014}.
 Kinetic models can not only help us better understand lesser known and understood characteristics of metabolism like bistability \parencite{Kotte2014}, and their role in human health, but can also improve predictions about the impact of engineering design perturbations on metabolism, and propose alternative designs to achieve metabolite production goals \parencite{Khodayari2016}.
@@
 	The development of these enzyme kinetic rate laws is based on in vitro observations of enzyme activity.
-Accordingly, authors have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions, as opposed to in vitro conditions \parencite{Heijnen2013}.
+Accordingly, authors have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions, as opposed to in vitro conditions \parencite{Heijnen2005,Heijnen2013}.
 This problem is further compounded by the fact that typical parameter values used in kinetic models of metabolism are estimated in vitro, and may not be applicable under in vivo conditions \parencite{Heijnen2005,Smallbone2007}.
-Moreover, estimation of kinetic parameters based on experimental in vivo data becomes computationally expensive for larger models, and any estimated in vivo parameter estimates are associated with large uncertainties \parencite{Link2014}.
-Using in vitro, or unreliable in vivo parameter estimates to study the steady state and dynamic responses of metabolic networks, under in vivo conditions, to different perturbations reduces confidence in the model predicted behaviour \parencite{Andreozzi2016a}.
-Consequently, this hampers the use of these models to gain insight into metabolic network function \parencite{Chakrabarti2013a}, and the increase in prediction uncertainty becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design \parencite{Saa2017}.
+Also, parameters estimated based on in vivo experimental data are associated with large uncertainties \parencite{Link2014}.
+Using in vitro, or unreliable in vivo parameter estimates to study the steady state and dynamic responses of metabolic networks, under in vivo conditions, to different perturbations reduces confidence in the model predicted behaviour.
+Consequently, this hampers the use of these models to gain insight into metabolic network function \parencite{Andreozzi2016a,Vasilakou2016}, and the increase in prediction uncertainty becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design \parencite{Saa2017}.
 
-	So, the use of heavily parameterized enzyme kinetic rate laws not only differentiates kinetic models from CBMs of metabolism, but is also a major hurdle to the widespread use and adoption of kinetic models \parencite{Berthoumieux2013,Vasilakou2016}.
-In order to address this problem, authors have proposed two different strategies.
-In one scenario, authors have looked at the development and use of global optimization algorithms for parameter estimation \parencite{Moles2003}.
-In another case, in vitro kinetic rate law descriptions are abandoned in favour of approximate and new empirical enzyme kinetic models whose parameters can not only be estimated in a relatively easier fashion, but also lead to kinetic models that generate information that has a higher correlation with experimental data that is observed in vivo \parencite{Heijnen2005,Smallbone2007}.
+	Although some authors have sought to quantify this uncertainty using different techniques \parencite{Vanlier2013,Andreozzi2016a}, others have proposed to alleviate as well as constrain the uncertainty in parameter estimates and consequent model predictions by using a Monte Carlo approach to kinetic modeling of metabolism that allows for integration of experimentally observed in vivo data \parencite{Srinivasan2015}.
+Bayesian approaches to improve parameter estimation and quantify estimation uncertainty have also been proposed \parencite{Saa2016}.
 
-	However, one of persistent problems is the lingering uncertainty in the parameter estimates and the subsequent predictions that are obtained from these methods.
-Authors have proposed to alleviate this parameter uncertainty either by using a Monte Carlo approach to kinetic modeling of metabolism \parencite{Srinivasan2015}, or by quantifying this uncertainty in parameter estimates using different methods \parencite{Vanlier2013}.
-Approaches to parameter estimation from a Bayesian perspective have also been proposed to estimate parameters and quantify the uncertainty in these estimates \parencite{Saa2016}.
+	In spite of the development of these methods to quantify parameter estimation uncertainty, model parameter identifiability, a necessary, and sometimes sufficient condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.
+Briefly, it concerns with the ability to estimate unique values for all model parameters from observed experimental data.
+In a model, any parameter is said to be structurally or a priori identifiable if its values can be estimated independent of all other model parameters from available experimental data.
+However, if parameters cannot be estimated independent of each other due to redundant model parameterization, or due to the nonlinear relationship between the model parameters, then the parameters are said to be structurally non-identifiable.
+Conversely, if the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is unquantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994}.
 
-	In spite of all these developments, model parameter identifiability, a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.
-Briefly, it concerns with the ability to distinguish between any two different parameter values based on differences in the inputs to, and consequently the responses exhibited by the cell.
-Structural identifiability refers to the dependence of the ability to estimate unique parameter values on model structure, due to inherent dependencies between two or more parameters in the model \parencite{Berthoumieux2013}.
-Conversely, when the dependence is on the informativeness of the data used for estimation, and not on the structure of the model, it is referred to as practical identifiability \parencite{Ljung1994}.
+	Authors have proposed to overcome concerns with parameter identifiability by proposing approximate kinetic models of metabolism that utilize empirical enzyme kinetic rate laws whose parameters have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}.
+Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Ljung1994,Nikerel2009,Berthoumieux2013,Raue2014}\textcolor{red}{(paper from Rudiyanto Gunawan on model discrimination and sensitivity analysis)}.
 
-	Significant work has been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Ljung1994,Nikerel2009,Berthoumieux2013,Raue2014}\textcolor{red}{(paper from Rudiyanto Gunawan on model discrimination and sensitivity analysis)}.
-However, the focus on methods for practical identification of parameters in kinetic models of metabolism and its use for experimental design is fairly recent and limited \parencite{Gadkar2005a,Vanlier2014a,Raue2014}.
-For instance some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, and some perform identifiability within a dynamic context \parencite{Gadkar2005a,Nikerel2009,Vanlier2014a}.
-Some other methods also suffer from their inability to be applied to kinetic models of large metabolic networks due to the associated computational burden \parencite{Gadkar2005a,Raue2014}\textcolor{red}{(Banga method using FIM for D-optimal design, ??)}.
+	Methods to improve practical identifiability through a priori experimental design have also been developed, with focus on kinetic models of metabolism \parencite{Gadkar2005a,Vanlier2014a,Raue2014}.
+Some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, while some of them suffer from computational limitations when applied to kinetic models of large metabolic networks \parencite{Gadkar2005a,Raue2014}\textcolor{red}{(Banga method using FIM for D-optimal design, ??)}.
 
-	In this paper, we propose a scalable methodology that uses available steady state fluxomics, metabolomics and proteomics data to practically identify parameters for each individual reaction in kinetic models of metabolism.
+	In this paper, we propose a scalable methodology that uses available steady state fluxomics, metabolomics and proteomics data to test the practical identifiability of parameters for each individual reaction in kinetic models of metabolism.
 We demonstrate how the computer algebra-based method that we have developed can also facilitate the design of experiments that are minimal and informative to generate data required to estimate unique parameter values for all reaction fluxes in a metabolic network.
-In doing so, we not only propose specific types of and nature of perturbations that will provide the most useful data for parameter estimation, but also test the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metbaolic networks.
+In doing so, we not only propose the number and types of perturbations that will provide the most useful data for parameter estimation, but also test the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks.
 
 	For the purposes of this method we assume that all intracellular metabolite concentrations and fluxes can be measured.
-We illustrate our methodology to identify parameters and design experiments for estimating parameters in a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017}.
+We illustrate our methodology to identify parameters and design experiments to identify parameters in a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017}.
+
+	%some perform identifiability within a dynamic context \parencite{Gadkar2005a,Nikerel2009,Vanlier2014a} and
+
 	  %We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
@@
 	\section{Methods}\label{sec:methods}
-	\subsection{Definition: Structural and Practical parameter identifiability of kinetic models of metabolism}\label{sec:kinetic_model}
-	In kinetic models of metabolism, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
+	\subsection{Parameter estimation for kinetic models of metabolism}\label{sec:kinetic_model}
+	In kinetic models of metabolism, ordinary differential equations (ODE) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
 The matrix $\mathbf{S}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
@@
 This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}.
-In such scenarios, the need to design experiments to effect a change in and discriminate between changes in different concentrations/fluxes becomes necessary.
+In such scenarios, the need to design experiments to effect a change in, and discriminate between changes in different concentrations/fluxes becomes necessary.
 
-	As briefly mentioned in the Introduction, in systems identification, the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
+	\subsection{Structural and practical identifiability of parameters in kinetic models}	
+	As briefly mentioned in the Introduction, the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
 The identifiability of parameters in linear and nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability.

### 2018-04-06 ident/docs/draft_2.tex
changes to methods

@@
 	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Srinivasan2015}.
-Accordingly, $f$ is a nonlinear function of the metabolite concentrations ($x$), enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
+Accordingly, $f$ is a nonlinear function of the vector of metabolite concentrations ($x$), the vector of enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
 
 	Parameter estimation methods based on optimization principles are typically used to determine true parameter values based on available experimental data.
-Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:kinstoich}) to estimate the values of enzyme kinetic parameters, $\theta$, based on the measured data.
+Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:lsqopt}) to estimate the values of enzyme kinetic parameters, $\theta$, based on the measured data.
 	\begin{subequations}\label{eq:lsqopt}
@@
 	\end{subequations}
+	Here $y = \left[x, v\right]^T$ is the vector of both concentrations ($x$) and fluxes ($v$).
 The minimization of least square error between the measured ($y^*$) and modeled ($y$) concentrations and fluxes, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration and flux, at each time point, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}).
@@
 
-	However, not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment.
-This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}.
-In such scenarios, the need to design experiments to effect a change in, and discriminate between changes in different concentrations/fluxes becomes necessary.
+	\subsection{Structural and practical identifiability of parameters in kinetic models}	
+	In the Introduction, we briefly metioned that the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
+Below, we provide a formal definition of structural and practical identifiability of parameters.
 
-	\subsection{Structural and practical identifiability of parameters in kinetic models}	
-	As briefly mentioned in the Introduction, the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
-The identifiability of parameters in linear and nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability.
-
-	Mathematically, the parameters $\theta$ in any nonlinear model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
+	The parameters in $\theta$ in any nonlinear model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \left[x, v\right]^T = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
 	\begin{align}\label{eq:stident}
@@
 	\end{align}
-	Accordingly, parameters in $\theta$ can have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, and are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
-So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unmeasured and measured concentrations/fluxes, as well as correlations between different model parameters.
+	Accordingly, if parameters in $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
+So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unmeasured and measured concentrations/fluxes, as well as correlations between different model parameters \textcolor{red}{Rudiyanto Gunawan paper on model discrimination}.
 Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
 
-	Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values with quantifiable uncertainties to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
-So, the effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
-Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
+	Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
+If a single unique parameter satisfying Equation (\ref{eq:stident}) can be found, then $\theta$ is said to be globally practically identifiable.
+Whereas, if parameter estimates with quantifiable uncertainties can be found, then the $\theta$ is said to be locally identifiable.
+The absence of unique parameter estimates for $\theta$ leads to practical non-identifiability.
+The practical identifiability of a parameter is hence contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
 
-	Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated from measurable data.
+	So, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated from measurable data.
 On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
@@
 	\subsection{A method to determine practical identifiability of kinetic models of metabolism}\label{sec:ident}
-	We provide the mathematical framework for \textcolor{red}{practical identification} of parameters in kinetic models of metabolism in this section.
+	We provide the mathematical framework for \textcolor{red}{identification} of parameters in kinetic models of metabolism in this section.
 A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
@@
 	\end{align}
-	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$, $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$, whose elements are denoted by $\theta_k$.
+	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$.
+$\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$, whose elements are denoted by $\theta_k$.
 
@@
 	\end{align}
+
 	The identifiability of parameter $\theta_k$, $k = {1, 2, ..., p}$, for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
 Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
-The solution obtained for $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_k$, and an infinite number of solutions are possible for a non-identifiable $\theta_k$.
-However, if there are multiple but finite solutions for $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, then the corresponding parameter $\theta_k$ is locally identifiable.
+The solution $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_k$, and an infinite number of solutions are possible for a non-identifiable $\theta_k$.
+However, if there are multiple but finite solutions $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, then the corresponding parameter $\theta_k$ is locally identifiable.
 
@@
 	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}
+
+	However, not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment.
+This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}.
+In such scenarios, the need to design experiments to effect a change in, and discriminate between changes in different concentrations/fluxes becomes necessary.
+
 	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.

### 2018-04-08 ident/docs/draft_2.tex
changes to methods, results and figure 1

@@
 Also, parameters estimated based on in vivo experimental data are associated with large uncertainties \parencite{Link2014}.
-Using in vitro, or unreliable in vivo parameter estimates to study the steady state and dynamic responses of metabolic networks, under in vivo conditions, to different perturbations reduces confidence in the model predicted behaviour.
+Using in vitro, or unreliable in vivo parameter estimates to study the steady state and dynamic responses of metabolic networks to different perturbations, under in vivo conditions, reduces confidence in the model predicted behaviour.
 Consequently, this hampers the use of these models to gain insight into metabolic network function \parencite{Andreozzi2016a,Vasilakou2016}, and the increase in prediction uncertainty becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design \parencite{Saa2017}.
@@
 Briefly, it concerns with the ability to estimate unique values for all model parameters from observed experimental data.
-In a model, any parameter is said to be structurally or a priori identifiable if its values can be estimated independent of all other model parameters from available experimental data.
-However, if parameters cannot be estimated independent of each other due to redundant model parameterization, or due to the nonlinear relationship between the model parameters, then the parameters are said to be structurally non-identifiable.
+In a model, any parameter is said to be structurally or a priori identifiable if its values can be uniquely estimated independent of all other model parameters from available experimental data.
+However, if parameters cannot be uniquely estimated independent of each other due to redundant model parameterization, or due to the nonlinear relationship between the model parameters, then the parameters are said to be structurally non-identifiable.
 Conversely, if the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is unquantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994}.
@@
 	Accordingly, if parameters in $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
-So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unmeasured and measured concentrations/fluxes, as well as correlations between different model parameters \textcolor{red}{Rudiyanto Gunawan paper on model discrimination}.
+So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unmeasured and measured concentrations/fluxes, as well as correlations between different model parameters \textcolor{red}{(Rudiyanto Gunawan paper on model discrimination)}.
 Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
@@
 
-	So, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated from measurable data.
+	So, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data.
 On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
@@
 The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta}$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
-This is done by first selecting a combination of $p$ data sets from the $n_{E}$ available experimental data sets.
-The combination consisting of data from $p$ different experiments is then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
+This is done by first selecting a combination of $p\le n_E$ experimental data.
+The fluxes and concentrations from $p$ different experiments are then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
 	\begin{align}\label{eq:nonlineq}
@@
 Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^p$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
-Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from all chosen experiments $j = {1, 2, ..., p}$, in the selected combination.
+Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from $p$ experiments denoted by the index $j = {1, 2, ..., p}$.
 	\begin{align}\label{eq:theta-eq}
@@
 	\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
-	A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate the utility of identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism.
+	A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism.
 The kinetic model is described below.
@@
 	\end{equation}		
+	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}).
+Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression in Equation (\ref{eq:flux1}) for $v_1$ cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
+So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
+Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
+The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
+	\begin{align}\label{eq:flux1a}
+	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
+	\end{align}		
+	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability. 	
 	\begin{equation}\label{eq:flux2}
@@
 One way to alter the physiological conditions is to change the substrate concentration under which the cells grow.
-In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$.
+In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$ (Equation \ref{eq:flux1}).
 Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.
@@
 The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
-Accordingly, in addition to measuring network response to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
+Accordingly, in addition to measuring network responses to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
 In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
@@
 Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers.
-Moreover, among these chosen smaller number of experimental data combinations, our methodology enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation.
+Moreover, among the chosen smaller number of experimental data combinations, the methodology that we propose enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation.
 The following section describes a procedure that can be used to design experiments within this context.
@@
 
-	However, not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment.
+	Not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment.
 This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}.
@@
 
-	The identifiability of each parameter based on each experiment indexed as $j = \{1, 2, ..., n_E\}$ is established based on the methodology described in section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in section \ref{sec:example}.
-Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
+	The identifiability of each parameter based on each experiment with index $j = {1, 2, ..., n_E}$ is established based on the methodology described in section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in section \ref{sec:example}.
+Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$ (Figure \ref{fig:ident-flowchart}b), then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
 Consequently, the corresponding experiment cannot be used for estimating parameter $\theta_k$, and needs to be discarded from the set of all necessary experiments.
-Furthermore, another experiment from $j = \{1, ..., n_E\}$ needs to be selected to replace the discarded experiment such that parameter $\theta_k$ is identifiable.
+Furthermore, another experiment from $j = {1, ..., n_E}$ needs to be selected to replace the discarded experiment such that parameter $\theta_k$ is identifiable.
 This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
@@
 	\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis}\label{sec:example}	
-	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}).
-Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression previously given for $v_1$ (Equation \ref{eq:flux1}) cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
-So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
-Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
-The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
-	\begin{align}\label{eq:flux1a}
-	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
-	\end{align}		
-	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability.
+	In this section, we illustrate the methodology that we developed to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network}) described in section \ref{sec:small-model}.
+We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration.
 
-	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ still holds for $v_1$, as mentioned in section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
+	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
 So, from the $n_E = 21$ different experiments described in section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$.
@@
 Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
+We apply the method described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}, and illustrated above for $v_1$ to fluxes $v_2$ and $v_3$ in Figure \ref{fig:network}.
+We summarize the results on the identifiability of parameters in these fluxes in Figure \ref{fig:ident} and describe them below.
+
+	In Figure \ref{fig:ident} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
+Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
+The three panels in Figure \ref{fig:ident} represent the degree of identifiability for parameters modeling the three different fluxes of the small network individually.
+
+	\subsection{Maximum reaction rates are more identifiable than enzyme binding constants}\label{sec:trends}
+
+	In regards to the identifiability of $v_1$ (Figure \ref{fig:figure1}a), $v_2$ (Figure \ref{fig:figure1}b) and $v_3$ (Figure \ref{fig:figure1}c), the degree of identifiability of the maximum reaction rates ($V_i^{max}$) in each of the three fluxes is higher than the degree of identifiability of the corresponding enzyme binding ($K_i$) constants and the allosteric activation constant ($K_3^{pep}$).
+We observe this trend irrespective of the fact that only  $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law (Michaelis-Menten), while the convenience kinetic rate law is used to model $v_3$.
+These observations lead us to conclude that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
+
+	From Figure \ref{fig:figure1} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified.
+In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
+Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range.
+%Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
+In linear systems, persistence of excitation of any input signal can be theoretically guaranteed.
+However, for nonlinear systems, like the metabolic network we deal with in this paper, persistence of excitation of input signals cannot be theoretically guaranteed, and should be assessed on a case by case basis.
+The lack of persistence of excitation requires the design of experiments to satisfy the data needs for complete system identification.
+
+	In the case of identifying fluxes within a metabolic network individually, each flux is not independent from the other fluxes that form the metabolic network.
+Each flux is connected to one or more fluxes through the metabolites that not only act as substrates and products, but also as allosteric activators and inhibitors (e.g., effect of \textit{pep} on $v_3$ in Figure \ref{fig:network}).
+Hence, careful consideration of experiments is a necessity so that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network.
+In the next section we discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of experiments from which data is collected for identification.
 
@@
 Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.
-
-	\subsection{Maximum reaction rates are more identifiable than enzyme binding constants}\label{sec:trends}	
-	In Figure \ref{fig:figure1} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
-Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
-The three panels in Figure \ref{fig:figure1} represent the degree of identifiability for parameters modeling the three different fluxes of the small network individually.
-
-	In regards to the identifiability of $v_1$ (Figure \ref{fig:figure1}a), $v_2$ (Figure \ref{fig:figure1}b) and $v_3$ (Figure \ref{fig:figure1}c), the degree of identifiability of the maximum reaction rates ($V_i^{max}$) in each of the three fluxes is higher than the degree of identifiability of the corresponding enzyme binding ($K_i$) constants and the allosteric activation constant ($K_3^{pep}$).
-We observe this trend irrespective of the fact that only  $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law (Michaelis-Menten), while the convenience kinetic rate law is used to model $v_3$.
-These observations lead us to conclude that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
-
-	From Figure \ref{fig:figure1} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified.
-In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
-Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range.
-%Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
-In linear systems, persistence of excitation of any input signal can be theoretically guaranteed.
-However, for nonlinear systems, like the metabolic network we deal with in this paper, persistence of excitation of input signals cannot be theoretically guaranteed, and should be assessed on a case by case basis.
-The lack of persistence of excitation requires the design of experiments to satisfy the data needs for complete system identification.
-
-	In the case of identifying fluxes within a metabolic network individually, each flux is not independent from the other fluxes that form the metabolic network.
-Each flux is connected to one or more fluxes through the metabolites that not only act as substrates and products, but also as allosteric activators and inhibitors (e.g., effect of \textit{pep} on $v_3$ in Figure \ref{fig:network}).
-Hence, careful consideration of experiments is a necessity so that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network.
-In the next section we discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of experiments from which data is collected for identification.
 
@@
 	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_v2_v3_v5_ident}}
+		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$ \textcolor{red}{(need to change figure)} when there is no noise in the input experimental data.
+The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
+The total number of experiments used to identify parameters in $v_1$, $v_2$ and $v_5$ is two.
+The total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is 240, 421 different combinations are used to test identifiability for $v_5$, and 4896 combinations are used to test identifiability of parameters for $v_3$.
+Note that $v_3$ is modeled with three parameters and requires data from at least three different experiments, and also has two possible roots for each of its parameters.
+Results for only one of the two roots indicated by (1) is shown here.}\label{fig:ident}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
 		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure4/v1_v1max_v2_v3_1_data_utility}}
@@
 Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v1max_v2_v3_1_2_ident}}
-		\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
-The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
-Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:ident}
 	\end{figure}

### 2018-04-09 ident/docs/draft_2.tex
significant changes to results - combine and remove old section headings

@@
 	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model}.
+Identifiability results for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ are also discussed in this section.
 
@@
 Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
-We apply the method described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}, and illustrated above for $v_1$ to fluxes $v_2$ and $v_3$ in Figure \ref{fig:network}.
+We apply the method described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}, and illustrated above for $v_1$, to fluxes $v_2$ and $v_3$ of the network shown in Figure \ref{fig:network}.
 We summarize the results on the identifiability of parameters in these fluxes in Figure \ref{fig:ident} and describe them below.
 
-	In Figure \ref{fig:ident} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
+	In Figure \ref{fig:ident}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b), $v_3$ (Figure \ref{fig:ident}c) and $v_5$ (Figure \ref{fig:ident}d) belonging to the network shown in Figure \ref{fig:network}.
 Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
-The three panels in Figure \ref{fig:ident} represent the degree of identifiability for parameters modeling the three different fluxes of the small network individually.
+We see from Figure \ref{fig:ident} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$) for each of the four fluxes is higher than the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law model.
+We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience Kinetic rate law, and $v_5$ is modeled as a Hill equation with inhibition.
+We accordingly summarize that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
 
-	\subsection{Maximum reaction rates are more identifiable than enzyme binding constants}\label{sec:trends}
+	In addition to the observation on the degrees of identifiability of different parameter types, we also see that the enzyme kinetic rate law used to model a flux determines the ability to identify the parameters within the given model.
+For instance, as we state in the Methods section, we could not apply our methodology to determine closed form solution for all the parameters when $v_3$ is modeled using MWC kinetics.
+In contrast, when $v_3$ is modeled using the Convenience Kinetic rate law, we find that the parameters are locally identifiable based on the definition we provide in section \ref{sec:ident}.
+We say locally identifiable due to the fact that all three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two possible roots.
+In Figure \ref{fig:ident}c, we only show one of the two roots for each parameter in $v_3$.
+The identifiability of parameters based on the second root obtained from the computer algebra system are shown in \textcolor{red}{(Figure x)}.
 
-	In regards to the identifiability of $v_1$ (Figure \ref{fig:figure1}a), $v_2$ (Figure \ref{fig:figure1}b) and $v_3$ (Figure \ref{fig:figure1}c), the degree of identifiability of the maximum reaction rates ($V_i^{max}$) in each of the three fluxes is higher than the degree of identifiability of the corresponding enzyme binding ($K_i$) constants and the allosteric activation constant ($K_3^{pep}$).
-We observe this trend irrespective of the fact that only  $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law (Michaelis-Menten), while the convenience kinetic rate law is used to model $v_3$.
-These observations lead us to conclude that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
+	Similarly, when we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system.
+However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$.
+Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have describe in section \ref{sec:ident}.
+Even in this scenario, just as we found for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable.
+$V_e^{max}$ on the other hand has only one possible root, thereby it is a globally identifiable parameter.
+The identifiability of the second root for $K_e^{fdp}$, indicated as $K_e^{fdp}(2)$, is shown in Figure \ref{fig:ident}d.
+The identifiability of the first root of $K_e^{fdp}$, $K_e^{fdp}(1)$, is shown in \textcolor{red}{(Figure y)}.
 
-	From Figure \ref{fig:figure1} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified.
+	While the identifiability of the second roots $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$ for $v_3$ \textcolor{red}{(Figure x)} are similar to that of the first root (Figure \ref{fig:ident}c), this is not the case for $v_5$.
+We would like to point out that this is one of the scenarios where the local identifiability of a parameter like $K_e^{fdp}$ can be expanded to global identifiability by adding physical relevance of parameter values as an additional constraint for practical identifiability ($D_k>0$, see Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident}).
+We found that the first root for $K_e^{fdp}$, $K_e^{fdp}(1)$, never possessed a physically relevant value, and accordingly was never identifiable based on the criteria set out in Figure \ref{fig:ident}b and section \ref{sec:ident} \textcolor{red}{(Figure y)}.
+In contrast, the second root, $K_e^{fdp}(2)$ has an identifiability of 50.0\% (Figure \ref{fig:ident}d).
+Thus, we can get globally identifiable parameters through the addition of constraints related to their physical relevance to locally identifiable parameters.
+
+	However, for $v_3$, we could not resolve local identifiability only through the addition of physical relevance constraints on the parameters.
+This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and \textcolor{red}{Figure x}).
+Instead, $K_3^{fdp}$ becomes globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value (\textcolor{red}{Figure m}), and $K_3^{pep}$ becomes globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (\textcolor{red}{Figure n}).
+Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable.
+This informs us of the structural non-identifiability between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, as well the conditions for global identifiability of parameters in the model for $v_3$.
+
+	 Although one could argue that a similar approach could be used to identify parameters of $v_3$ written using the MWC model, we found that it was not possible to obtain identifiability expressions for any of the parameters in the MWC model, except for $V_3^{max}$ when all other parameters are assumed to be known.
+
+	 	%In using the convenience kinetic rate law model for $v_3$ (Equation \ref{eq:flux3_convkin}) for identifying the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$, we found that each of the three parameters have two different closed-form expressions owing to the presence of a square root term in their solutions.
+Thus, as per the stated definition in section \ref{sec:ident}, the parameters for $v_3$, described using the convenience kinetics rate laws, are only locally identifiable due to the non-unique but finite number of possible solutions for the three parameters.
+The two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, and $K_3^{pep}(1)$ refer to the first set of solutions, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$ refer to the other set of solutions.
+
+	\textcolor{red}{(move to discussion)} The impact of the formulation and nonlinearity of a kinetic rate law expression affecting the practical identifiability of parameters in the expression may not be an unique problem isolated to the system that we are investigating.
+Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms \parencite{Chassagnole2002a, Peskov2012, VanHeerden2014}.
+However, the identifiability of the parameters used in these models has never been truly examined.
+\textcolor{red}{mentione Heijnen papers on resolving identifiability using approximate models here}.
+
+	\textcolor{red}{(Move to discussion)} Unlike the case of the MWC model, if determining closed-form expressions for the parameters of these models are tractable, we believe our methodology can help in elucidating the identifiability of these parameters.
+If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation and model validation.
+
+	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}		
+	From Figure \ref{fig:ident} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified.
 In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
@@
 Each flux is connected to one or more fluxes through the metabolites that not only act as substrates and products, but also as allosteric activators and inhibitors (e.g., effect of \textit{pep} on $v_3$ in Figure \ref{fig:network}).
-Hence, careful consideration of experiments is a necessity so that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network.
+Hence, careful consideration of experiments is a necessity so that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}.
 In the next section we discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of experiments from which data is collected for identification.
-
-	%As mentioned in section \ref{sec:intro}, practical identifiability is a necessary condition to estimate true values of both $V_1^{max}$ and $K_1^{ac}$ from experimental data. 	
-	%Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_1$ in the small model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model}.
-
-	\subsection{Nonlinearity of enzyme kinetic rate law expression affects identifiability analysis}\label{sec:closedform}
-	In the previous section, we have demonstrated our methodology to practically identify parameters for a flux that is modeled using the Michaelis-Menten (MM) kinetic rate law when steady state concentration and flux data from two different experiments are available.
-However, metabolic fluxes can be described by several different kinetic rate law expressions based on their representation of the physiological interactions that facilitate enzyme catalysis of each flux in a metabolic network.
-For instance, as described in section \ref{sec:small-model}, flux $v_3$, which is allosterically activated by \textit{pep}, is modeled using the MWC kinetics (Equation \ref{eq:flux3}).
-So, we comment on the capability of our methodology to practically identify parameters in cases where metabolic fluxes are modeled using rate laws that are different from the MM model, and consequently differ in terms of their nonlinear complexity.
-
-	In using the convenience kinetic rate law model for $v_3$ (Equation \ref{eq:flux3_convkin}) for identifying the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$, we found that each of the three parameters have two different closed-form expressions owing to the presence of a square root term in their solutions.
-Thus, as per the stated definition in section \ref{sec:ident}, the parameters for $v_3$, described using the convenience kinetics rate laws, are only locally identifiable due to the non-unique but finite number of possible solutions for the three parameters.
-The two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, and $K_3^{pep}(1)$ refer to the first set of solutions, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$ refer to the other set of solutions.
-
-	However, we found that the local identifiability of the parameters could be resolved to obtain globally identifiable parameters if either the Michaelis binding constant for \textit{fdp}, $K_3^{fdp}$, or the allosteric binding constant for \textit{pep}, $K_3^{pep}$, is known.
-Thus, fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter.
-Although one could argue that a similar approach could be used to identify parameters of $v_3$ written using the MWC model, we found that it was not possible to obtain identifiability expressions for any of the parameters in the MWC model, except for $V_3^{max}$ when all other parameters are assumed to be known.
-
-	Similarly, not all parameters in the Hill equation, used to model the inhibition of \textit{fdp} on the production of the enzyme E through flux $v_5$ (Equation \ref{eq:flux5}), can be identified using our methodology.
-We found that the Hill exponent ($n_e$) needs to be fixed to be able to identify the other two parameters, $V_e^{max}$ and $K_e^{fdp}$, in the model.
-Thus, the dimension of the parameter space for $v_5$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have developed in this paper.
-However, on applying the primary test for identifiability ($D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, see section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b), we found only $V_e^{max}$ to globally identifiable.
-We found $K_e^{fdp}$ to be locally identifiable to the possibility of having two different roots.
-As in the case of parameters for $v_3$ when using the convenience kinetics formulation, the presence of a square root term was the reason for the existence of two roots for $K_e^{fdp}$.
-However, on applying the secondary criteria for identifiability ($g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, see section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b), we were able to eliminate the existence of the second root as it is always negative, and therefore not physically relevant.
-Thus, we were able to determine that both parameters modeling $v_5$ ($V_e^{max}$ and $K_e^{fdp}$) are globally practically identifiable under certain conditions.
-
-	The impact of nonlinearity of a kinetic rate law expression affecting the practical identifiability of parameters in the expression may not be an unique problem isolated to the system that we are investigating.
-Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms \parencite{Chassagnole2002a, Peskov2012, VanHeerden2014}.
-However, the identifiability of the parameters used in these models has never been truly examined.
-\textcolor{red}{mentione Heijnen papers on resolving identifiability using approximate models here}.
-
-	Unlike the case of the MWC model, if determining closed-form expressions for the parameters of these models are tractable, we believe our methodology can help in elucidating the identifiability of these parameters.
-If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation and model validation.
-
-	Next, we look at some of the results we obtained for the practical identification of different parameters for all three fluxes $v_1$, $v_2$ and $v_3$ in the small gluconeogenic network model described in section \ref{sec:small-model}, with the flux for $v_3$ described by Equation (\ref{eq:flux3_convkin}).
 
@@
 
-	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}	
+	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}
 	In this section, we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model.

### 2018-04-09 ident/docs/draft_2.tex
changes to section 2 of results and figure order

@@
 In Figure \ref{fig:ident}c, we only show one of the two roots for each parameter in $v_3$.
-The identifiability of parameters based on the second root obtained from the computer algebra system are shown in \textcolor{red}{(Figure x)}.
+The identifiability of parameters based on the second root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}b.
 
@@
 The identifiability of the second root for $K_e^{fdp}$, indicated as $K_e^{fdp}(2)$, is shown in Figure \ref{fig:ident}d.
-The identifiability of the first root of $K_e^{fdp}$, $K_e^{fdp}(1)$, is shown in \textcolor{red}{(Figure y)}.
+The identifiability of the first root of $K_e^{fdp}$, $K_e^{fdp}(1)$, is shown in Figure \ref{fig:ident_root2}c.
 
-	While the identifiability of the second roots $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$ for $v_3$ \textcolor{red}{(Figure x)} are similar to that of the first root (Figure \ref{fig:ident}c), this is not the case for $v_5$.
+	While the identifiability of the second roots $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$ for $v_3$ (Figure \ref{fig:ident_root2}b) are similar to that of the first root (Figure \ref{fig:ident}c), this is not the case for $v_5$.
 We would like to point out that this is one of the scenarios where the local identifiability of a parameter like $K_e^{fdp}$ can be expanded to global identifiability by adding physical relevance of parameter values as an additional constraint for practical identifiability ($D_k>0$, see Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident}).
-We found that the first root for $K_e^{fdp}$, $K_e^{fdp}(1)$, never possessed a physically relevant value, and accordingly was never identifiable based on the criteria set out in Figure \ref{fig:ident}b and section \ref{sec:ident} \textcolor{red}{(Figure y)}.
+We found that the first root for $K_e^{fdp}$, $K_e^{fdp}(1)$, never possessed a physically relevant value, and accordingly was never identifiable based on the criteria set out in Figure \ref{fig:ident}b and section \ref{sec:ident} (Figure \ref{fig:ident_root2}c).
 In contrast, the second root, $K_e^{fdp}(2)$ has an identifiability of 50.0\% (Figure \ref{fig:ident}d).
@@
 	However, for $v_3$, we could not resolve local identifiability only through the addition of physical relevance constraints on the parameters.
-This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and \textcolor{red}{Figure x}).
-Instead, $K_3^{fdp}$ becomes globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value (\textcolor{red}{Figure m}), and $K_3^{pep}$ becomes globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (\textcolor{red}{Figure n}).
+This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and Figure \ref{fig:ident_root2}b).
+Instead, $K_3^{fdp}$ becomes globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:ident_v3_var}a), and $K_3^{pep}$ becomes globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:ident_v3_var}b).
 Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable.
@@
 
-	\textcolor{red}{(move to discussion)} The impact of the formulation and nonlinearity of a kinetic rate law expression affecting the practical identifiability of parameters in the expression may not be an unique problem isolated to the system that we are investigating.
-Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms \parencite{Chassagnole2002a, Peskov2012, VanHeerden2014}.
-However, the identifiability of the parameters used in these models has never been truly examined.
-\textcolor{red}{mentione Heijnen papers on resolving identifiability using approximate models here}.
-
-	\textcolor{red}{(Move to discussion)} Unlike the case of the MWC model, if determining closed-form expressions for the parameters of these models are tractable, we believe our methodology can help in elucidating the identifiability of these parameters.
-If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation and model validation.
-
 	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}		
-	From Figure \ref{fig:ident} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified.
+	From Figure \ref{fig:ident} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.
+For example, the maximum reactions rates of $v_2$ and $v_5$ have a degree of identifiability of 100\% (Figure \ref{fig:ident}b and d), while their counterparts $V_1^{max}$ and $V_3^{max}$ in $v_1$ and $v_3$ respectively (Figure \ref{fig:ident}a and c), are only identifiable to a lesser extent.
+	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified.
 In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
@@
 	In the case of identifying fluxes within a metabolic network individually, each flux is not independent from the other fluxes that form the metabolic network.
-Each flux is connected to one or more fluxes through the metabolites that not only act as substrates and products, but also as allosteric activators and inhibitors (e.g., effect of \textit{pep} on $v_3$ in Figure \ref{fig:network}).
+Each flux is connected to one or more fluxes through the metabolites that not only act as substrates and products, but also as allosteric or transcriptional activators and inhibitors (e.g., effect of \textit{pep} on $v_3$, or the effect of \textit{fdp} on $v_5$ and consequently on $v_1$ in Figure \ref{fig:network}).
 Hence, careful consideration of experiments is a necessity so that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}.
-In the next section we discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of experiments from which data is collected for identification.
+So, in this section we discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of experiments from which data is collected for identification.
 
-	\subsection{Utility of experimental data for practical parameter identification}\label{sec:data_utility}
-	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not been well studied (\textcolor{red}{include Rudiyanto Gunawan papers here}).
-Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
-
-	We do this by showing how useful the multiple combinations of experimental data are towards identifying parameters for each given flux (Figure \ref{fig:figure4}).
-We express the usefulness of any given data combination on the basis of the number of parameters that the combination can identify.
-The higher the number of parameters a data combination can identify, the greater is the usefulness of that combination of experiments.
-
-	The number of data combinations that can identify a specific number of parameters within a given flux is shown in Figure \ref{fig:figure4}.
-The percentage of all data combinations that can estimate a given number of parameters within a given flux are also shown in Figure \ref{fig:figure4}.
-For flux $v_1$ (Figure \ref{fig:figure4}a) about 78\% of the 306 data combinations cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$).
-Only a single parameter can be identified by about 2\% of the combinations, while only about 19\% of the data combinations can identify both the parameters used in the model.
-	In contrast, most data combinations (>90\%) are not wasted, and can identify both parameters in $v_2$ (Figure \ref{fig:figure4}b).
-Less than 5\% of the experimental data combinations are wasted from not being able to identify any parameter.
-In the case of $v_3$ (Figure \ref{fig:figure4}c), about 45\% of the 4896 data combinations can identify only one parameter, while 13\% of the combinations cannot identify any parameter.
-Only 10\% of combinations can identify $v_3$ completely.
-
-	Thus, Figure \ref{fig:figure4} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
-For instance, choosing any of the experiment combinations from the 78\% percent that does not identify any parameter in $v_1$, or the 13\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative.
-This can lead to potential waste of resources used to perform these experiments.
-
-	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:figure4} also informs the ease of identifying parameters for a single flux.
-For the small network (Figure \ref{fig:network}), we find that more than 90\% of all the data combinations tested for their ability to identify $v_2$ are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:figure4}b).
-In contrast, less than 20\% of the combinations can identify both parameters in $v_1$.
-Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ as well as $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$.
-A careful consideration of experiments for identifying $v_2$ is however not a major concern since there is a very high probability that experiments chosen to identify $v_1$ will also identify $v_2$.
-
-	Although the calculation of the utility of experiment combinations in terms of their ability to identify the maximum number of parameters gives an idea on the informativeness of experiments used to gather the data, further analysis from the perspective of identifiability of the parameters of different fluxes is required to be able to design experiments based on this information.
-Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.
-
-	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}
-	In this section, we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model.
+	we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model.
 Earlier, we mentioned that the informativeness of experiments for identifying parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter.
@@
 
+	\subsection{Utility of experimental data for practical parameter identification}\label{sec:data_utility}
+	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not been well studied (\textcolor{red}{include Rudiyanto Gunawan papers here}).
+Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
+
+	We do this by showing how useful the multiple combinations of experimental data are towards identifying parameters for each given flux (Figure \ref{fig:figure4}).
+We express the usefulness of any given data combination on the basis of the number of parameters that the combination can identify.
+The higher the number of parameters a data combination can identify, the greater is the usefulness of that combination of experiments.
+
+	The number of data combinations that can identify a specific number of parameters within a given flux is shown in Figure \ref{fig:figure4}.
+The percentage of all data combinations that can estimate a given number of parameters within a given flux are also shown in Figure \ref{fig:figure4}.
+For flux $v_1$ (Figure \ref{fig:figure4}a) about 78\% of the 306 data combinations cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$).
+Only a single parameter can be identified by about 2\% of the combinations, while only about 19\% of the data combinations can identify both the parameters used in the model.
+	In contrast, most data combinations (>90\%) are not wasted, and can identify both parameters in $v_2$ (Figure \ref{fig:figure4}b).
+Less than 5\% of the experimental data combinations are wasted from not being able to identify any parameter.
+In the case of $v_3$ (Figure \ref{fig:figure4}c), about 45\% of the 4896 data combinations can identify only one parameter, while 13\% of the combinations cannot identify any parameter.
+Only 10\% of combinations can identify $v_3$ completely.
+
+	Thus, Figure \ref{fig:figure4} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
+For instance, choosing any of the experiment combinations from the 78\% percent that does not identify any parameter in $v_1$, or the 13\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative.
+This can lead to potential waste of resources used to perform these experiments.
+
+	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:figure4} also informs the ease of identifying parameters for a single flux.
+For the small network (Figure \ref{fig:network}), we find that more than 90\% of all the data combinations tested for their ability to identify $v_2$ are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:figure4}b).
+In contrast, less than 20\% of the combinations can identify both parameters in $v_1$.
+Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ as well as $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$.
+A careful consideration of experiments for identifying $v_2$ is however not a major concern since there is a very high probability that experiments chosen to identify $v_1$ will also identify $v_2$.
+
+	Although the calculation of the utility of experiment combinations in terms of their ability to identify the maximum number of parameters gives an idea on the informativeness of experiments used to gather the data, further analysis from the perspective of identifiability of the parameters of different fluxes is required to be able to design experiments based on this information.
+Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.
+
+	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}
+	In this section,
+
 	\subsection{Effect of noise on degree of identifiability is dependent on nonlinearity of enzyme kinetic rate law models}\label{sec:noise}
@@
 	\section{Discussions}\label{sec:discussion}	
+	\textcolor{red}{(move to discussion)} The impact of the formulation and nonlinearity of a kinetic rate law expression affecting the practical identifiability of parameters in the expression may not be an unique problem isolated to the system that we are investigating.
+Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms \parencite{Chassagnole2002a, Peskov2012, VanHeerden2014}.
+However, the identifiability of the parameters used in these models has never been truly examined.
+\textcolor{red}{mentione Heijnen papers on resolving identifiability using approximate models here}.
+
+	\textcolor{red}{(Move to discussion)} Unlike the case of the MWC model, if determining closed-form expressions for the parameters of these models are tractable, we believe our methodology can help in elucidating the identifiability of these parameters.
+If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation and model validation.
+
 	Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data.
@@
 	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_v1max_v2_exp_info}}
+		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
+The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:exp_info}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
 		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure4/v1_v1max_v2_v3_1_data_utility}}
@@
 
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_v1max_v2_exp_info}}
-		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:exp_info}
-	\end{figure}
-
 	\clearpage
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v5_1_2_ident}}
-		\caption{identifiability for flux v5 with 2 roots for Kefdp}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight, keepaspectratio]{figures/figure1/v1_kcat_v3_2_v5_1_ident}}
+		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_3$ and c) $v_5$ \textcolor{red}{(need to change figure)} when there is no noise in the input experimental data.
+The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
+The total number of experiments used to identify parameters in $v_1$ and $v_5$ is two.
+The total number of data combinations used to identify parameters in fluxes $v_1$ is 240.
+421 different combinations are used to test identifiability for $v_5$, and 4896 combinations are used to test identifiability of parameters for $v_3$.
+Note that $v_3$ is modeled with three parameters and requires data from at least three different experiments, and also has two possible roots for each of its parameters.
+Results for only one of the two roots indicated by (2) is shown here.
+Similarly, results for only one of the two roots of $K_e^{fdp}$, indicated as $K_e^{fdp}(1)$ is shown.}\label{fig:ident_root2}
 	\end{figure}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v5_1_2_data_utility}}
-		\caption{data utility for flux v5 with 2 roots for Kefdp}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_1_2_ident}}
+		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in the convenience kinetic rate law model of flux $v_3$ when a) $K_3^{pep}$ is held constant, and b) $K_3^{fdp}$ is held constant.
+Note that the two identifiable parameters in both panel a and panel b are globally identifiable.
+The degree of identifiability of $V_3^{max}$ and $K_3^{pep}$ are higher when $K_3^{fdp}$ is held constant, than the degree of identifiability of $V_3^{max}$ and $K_3^{fdp}$ when $K_3^{pep}$ is held constant.}\label{fig:ident_v3_var}
 	\end{figure}
@@
 		\caption{experiment information for flux v5 with 2 roots for Kefdp}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_data_utility}}
-		\caption{Same as other data utility figure for v1 with k1cat as parameter (with enzyme data) and second roots of v3.
-Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$.}\label{fig:figure4s}
 	\end{figure}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_1_2_ident}}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_exp_info}}
 		\caption{missing caption}%\label{fig:figure2}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v3_1_2_data_utility}}
-		\caption{missing caption}%\label{fig:figure2}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v5_1_2_data_utility}}
+		\caption{data utility for flux v5 with 2 roots for Kefdp}
 	\end{figure}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_exp_info}}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_data_utility}}
+		\caption{Same as other data utility figure for v1 with k1cat as parameter (with enzyme data) and second roots of v3.
+Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$.}\label{fig:figure4s}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v3_1_2_data_utility}}
 		\caption{missing caption}%\label{fig:figure2}

### 2018-04-09 ident/docs/draft_2.tex
changes to sec 2 of results and changed v5 ident figure

@@
 Identifiability results for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ are also discussed in this section.
+We show that the ability to establish practical identifiability of kinetic parameters in metabolic network models relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
+We also show that the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetic rate law used to describe the flux.
 
-	First in section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
-We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
+	First in section \ref{sec:closedform},  We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
 
@@
 We do this by looking at the number of combinations that can identify the maximum number of parameters.
-	In section \ref{sec:trends} we show that the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetic rate law used to describe the flux.
+	In section \ref{sec:trends}
+
 	In section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment.
@@
 
+	In Figure \ref{fig:ident}a less than 50\% of all data combinations can identify $v_1$.
+Whereas, all available data combinations satisfy the requirements for identifying $v_2$ (Figure \ref{fig:ident}b).
+This difference can be alluded to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter.
+For example we showed in Equation (\ref{eq:v1_par}) that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
+So, in this instance, the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$.
+Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}.
+
 	In addition to the observation on the degrees of identifiability of different parameter types, we also see that the enzyme kinetic rate law used to model a flux determines the ability to identify the parameters within the given model.
@@
 However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$.
-Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have describe in section \ref{sec:ident}.
-Even in this scenario, just as we found for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable.
+Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}.
+Even in this scenario, just as we saw earlier for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable.
 $V_e^{max}$ on the other hand has only one possible root, thereby it is a globally identifiable parameter.
@@
 We would like to point out that this is one of the scenarios where the local identifiability of a parameter like $K_e^{fdp}$ can be expanded to global identifiability by adding physical relevance of parameter values as an additional constraint for practical identifiability ($D_k>0$, see Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident}).
-We found that the first root for $K_e^{fdp}$, $K_e^{fdp}(1)$, never possessed a physically relevant value, and accordingly was never identifiable based on the criteria set out in Figure \ref{fig:ident}b and section \ref{sec:ident} (Figure \ref{fig:ident_root2}c).
+We found that the first root for $K_e^{fdp}$, $K_e^{fdp}(1)$, never possessed a physically relevant value, and accordingly was never identifiable based on the criteria set out in Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident} (Figure \ref{fig:ident_root2}c).
 In contrast, the second root, $K_e^{fdp}(2)$ has an identifiability of 50.0\% (Figure \ref{fig:ident}d).
@@
 
-	we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model.
-Earlier, we mentioned that the informativeness of experiments for identifying parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter.
-For example, we demonstrated in section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them (Equation \ref{eq:v1_par}).
-Thus, in this instance, the informativeness of the experiments (changes in the input acetate concentration and $v_1$) for identifying parameters of $v_1$ is determined by the ability of the input change to effect a change in the measured value of $v_1$.
-Similarly, the identification of parameters for $v_2$ requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}.
-%, given both these variables are outputs of the model describing $v_2$.
+	We demonstrated in section \ref{sec:example}, using $v_1$ and $v_2$ as examples, that the ability of data from a given combination of experiments to identify parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter.
+For instance, identification of $V_1^{max}$ and $K_1^{ac}$ requires that the data from two different experiments used for parameter estimation must have distinct acetate concentrations as well as different uptake fluxes ($v_{1,j}$) between them (Equation \ref{eq:v1_par}).
+Thus, the informativeness of the experiments (changes in the input acetate concentration and $v_1$) for identifying parameters of $v_1$ is determined by the ability of the input change to effect a change in the measured value of $v_1$.	Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.
 
-	As such, as per Figure \ref{fig:figure1}a only 20\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$.
-Whereas, more than 90\% of the available data combinations satisfy the requirements for identifying $v_2$ (Figure \ref{fig:figure1}b).
-Thus, identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for both $v_1$ and $v_2$.
-Recall that we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Section \ref{sec:experiments}).
-	In order to determine the type of experiments that help satisfy identifiability conditions for both parameters of $v_1$, we look at the distribution of the five different types in all data combinations that can identify these 2 parameters (Figure \ref{fig:figure2}a and b).
+	Although we were able to determine the requirements for identifying $v_1$ and $v_2$ based on the closed form expressions of the parameters, this may not be possible when a complex nonlinear rate law is used to model a flux.
+Using the Convenience Kinetic rate law to model $v_3$ is a case in point.
+So, in order to determine the type of experiments that help satisfy identifiability conditions for parameters of any flux modeled by any usable rate law, we look at the distribution of the different types of experiments in all data combinations that can identify the parameters within a flux.
+Recall that we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Section \ref{sec:experiments} and Table \ref{tab:pval}).
+
+	The contribution from different types of experiments towards identification of parameters in different fluxes is given in Figure \ref{fig:exp_info}.
+In Figure \ref{fig:exp_info}a and b, the contribution from different experiment types for identifying $V_1^{max}$ and $K_1^{ac}(ne)$ are given, respectively.
+The contribution of different experiments towards identifiability for $V_2^{max}$ (Figure \ref{fig:exp_info}c) and $K_2^{pep}$ (Figure \ref{fig:exp_info}d) are also shown.
 
@@
 		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_v2_v3_v5_ident}}
-		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$ \textcolor{red}{(need to change figure)} when there is no noise in the input experimental data.
+		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$ when there is no noise in the input experimental data.
 The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.

### 2018-04-09 ident/python2/ss-ident/v5_script.py
no major changes

@@ -58,7 +58,7 @@ print('Practical Identifiability Analysis of v5 with 2 parameters: vemax and Kef
 ident_fun_choice = [2]
 # perform identifiability when v1 is written with k1cat*E in the numerator
 ident_details_v5_root1 = flux_ident_2_data_combination(experimental_datasets_2_expts, choose=combination_choice,
-                                                       flux_ids=[5], flux_choice=[1], ident_fun_choice=ident_fun_choice)
+                                                       flux_ids=[5], flux_choice=[2], ident_fun_choice=ident_fun_choice)
 print('Identifiability analysis of v5 with 2 parameters (vemax and Kefdp) complete.\n')
 
 # data processing - do not combine fluxes

### 2018-04-10 ident/docs/draft_2.tex
changes to section 3.2 of results and figure order

@@
 	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}		
-
 	Not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment.
@@
 
+	Then, in section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment.
+Here we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux.
+
 	First in section \ref{sec:closedform},  We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
@@
 
-	In section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment.
-In this section we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux.
 Finally, we discuss some results arising out of the use of our methodology to determine the identifiability of parameters when data with additive noise is used in section \ref{sec:noise}.
@@
 	The contribution from different types of experiments towards identification of parameters in different fluxes is given in Figure \ref{fig:exp_info}.
-In Figure \ref{fig:exp_info}a and b, the contribution from different experiment types for identifying $V_1^{max}$ and $K_1^{ac}(ne)$ are given, respectively.
+In Figure \ref{fig:exp_info}a and \ref{fig:exp_info}b, the contribution from different experiment types for identifying $V_1^{max}$ and $K_1^{ac}(ne)$ are given, respectively.
 The contribution of different experiments towards identifiability for $V_2^{max}$ (Figure \ref{fig:exp_info}c) and $K_2^{pep}$ (Figure \ref{fig:exp_info}d) are also shown.
 
-	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments.
+	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments (Figure \ref{fig:exp_info}a and b).
 This matches with requirements for identifiability on the basis of the informativeness of experiments that was laid out earlier i.e., experiments must be able distinguish between different values of $v_1$ and acetate.
-Accordingly, only about 20\% of the 306 different experiment combinations used to test identifiability are informative enough to discriminate both acetate as well as $v_1$, and can identify $V_1^{max}$ and $K_1^{ac}(ne)$ (Figure \ref{fig:figure1}a).
+Accordingly, only about 40\% of the 240 different experiment combinations used to test identifiability are informative enough to discriminate both acetate as well as $v_1$, and can identify $V_1^{max}$ and $K_1^{ac}(ne)$ (Figure \ref{fig:ident}a).
 
-	In contrast to $v_1$, we see that the different enzyme perturbation experiments have a higher contribution towards data combinations that can identify parameters for $v_2$ (Figure \ref{fig:figure2}c and d).
-Hence, in conjunction with the high degree of identifiability seen in Figure \ref{fig:figure1}b for $v_2$, we can deduce that almost all experiment types are persistently excitable for $v_2$ i.e., most experiments can bring about noticeable changes to both \textit{pep} and $v_2$.
+	In contrast to $v_1$, we see that the different enzyme perturbation experiments have a higher contribution towards data combinations that can identify parameters for $v_2$ (Figure \ref{fig:exp_info}c and d).
+Hence, in conjunction with the high degree of identifiability seen in Figure \ref{fig:ident}b for $v_2$, we can deduce that almost all experiment types are persistently excitable for $v_2$ i.e., most experiments can bring about noticeable changes to both \textit{pep} and $v_2$.
 Consequently, in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative enough to identify $v_2$.
-This observations is in line with the one made in section \ref{sec:data_utility} based on the total number of combinations that can identify both parameters in $v_2$.
+%This observations is in line with the one made in section \ref{sec:data_utility} based on the total number of combinations that can identify both parameters in $v_2$.
 
@@
 Hence, it is theoretically possible for any of the five different experiment types to be persistently excitable to identify $v_2$.
-This is confirmed by the high degree of identifiability for both parameters of $v_2$, where in more than 90\% of all data combinations can identify the parameters.
+This is confirmed by the high degree of identifiability for both parameters of $v_2$, i.e., all data combinations can identify the parameters.
 Thus, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending on the position of the flux in the metabolic network with respect to the inputs and the outputs of the network, as well as the various regulatory interactions present within the network.
-We can extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_3$ as well (Figure \ref{fig:figure2_v3}).
+
+	We can extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_3$ as well (Figure \ref{fig:v3_exp_info}).
+In this case, the top three panels (Figure \ref{fig:v3_exp_info}a-c) show the distribution of different experiment types for identifying each parameter with respect to their first root identified as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$.
+The bottom panels (Figure \ref{fig:v3_exp_info}d-f) show the corresponding distribution of experiments for the second root of each fo three parameters in $v_3$: $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
+
+	In all of the above scenarios for $v_1$ (Figure \ref{fig:exp_info}a and b), $v_2$ (Figure \ref{fig:exp_info}c and d) and $v_3$ (Figure \ref{fig:v3_exp_info}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar.
+Hence, the green and blue surfaces are superimposed on top of each other.
+This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:v5_exp_info}a and b).
+However, this is not the case with the distribution of experiments required for identifying $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:v5_exp_info}c).
+We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$.
+Also, since $V_e^{max}$ is globally identifiable using any experiment type, the choice and number of experiments required to identify $v_5$ completely hinges upon identifiability of $K_e^{fdp}$ from the chosen experiments.
 
@@
 Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.
-
-	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}
-	In this section,
 
@@
 	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_1_2_exp_info}}
+		\caption{missing caption}\label{fig:v3_exp_info}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
 		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure2/v5_1_2_exp_info}}
-		\caption{experiment information for flux v5 with 2 roots for Kefdp}
+		\caption{experiment information for flux v5 with 2 roots for Kefdp}\label{fig:v5_exp_info}
 	\end{figure}
@@
 		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_k1cat_exp_info}}
-		\caption{missing caption}%\label{fig:figure2}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_1_2_exp_info}}
 		\caption{missing caption}%\label{fig:figure2}

### 2018-04-10 ident/docs/draft_2.tex
changes to results, more figures and SI figures, and figure captions

@@
 Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
-We see from Figure \ref{fig:ident} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$) for each of the four fluxes is higher than the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law model.
+We see from Figure \ref{fig:ident} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$) for each of the four fluxes is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law model.
 We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience Kinetic rate law, and $v_5$ is modeled as a Hill equation with inhibition.
@@
 This difference can be alluded to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter.
-For example we showed in Equation (\ref{eq:v1_par}) that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
+For example, we showed in Equation (\ref{eq:v1_par}) that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
 So, in this instance, the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$.
@@
 Even in this scenario, just as we saw earlier for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable.
-$V_e^{max}$ on the other hand has only one possible root, thereby it is a globally identifiable parameter.
+On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally identifiable parameter.
 The identifiability of the second root for $K_e^{fdp}$, indicated as $K_e^{fdp}(2)$, is shown in Figure \ref{fig:ident}d.
@@
 Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable.
-This informs us of the structural non-identifiability between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, as well the conditions for global identifiability of parameters in the model for $v_3$.
+Thus, the dimension of the parameter space for $v_3$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ for global identifiability.
+This informs us of the correlation between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, which results in structural non-identifiability of parameters in the model.
 
@@
 The two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, and $K_3^{pep}(1)$ refer to the first set of solutions, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$ refer to the other set of solutions.
+
+	\subsection{Rationale for a priori experimental design through practical parameter identification}\label{sec:data_utility}
+	%In section \ref{sec:example}, we described how the practical identifiability of parameters in a kinetic model of a metabolic network depended on the enzyme kinetic rate law used to model the fluxes that constitute the network.
+In this section we look at the rationale for a priori experimental design to collect meaningful data for parameter estimation.
+	%We also showed how the methodology that we propose can help in determining the nature and type of experiments that one needs to perform to identify and estimate parameters in kinetic models in section \ref{sec:experiment_type}.
+	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Banga papers on global optimization methods for parameter estimation, Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not gained enough traction (\textcolor{red}{include Rudiyanto Gunawan papers, D-optimal design using FIM here}).
+Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
+
+	We do this by showing how useful the multiple combinations of experimental data are towards identifying parameters for each flux $v_1$ (Figure \ref{fig:data_utility}a), $v_2$ (Figure \ref{fig:data_utility}b), $v_3$ (Figure \ref{fig:data_utility}c) and $v_5$ (Figure \ref{fig:data_utility}d).
+We express the usefulness of data from any given combination of experiments on the basis of the number of parameters that the data from the combination can identify.
+The higher the number of parameters a combination of experiments can identify, the greater is the usefulness of that combination of experiments.
+The percentage of all data combinations that can estimate a given number of parameters within a given flux is also shown in Figure \ref{fig:data_utility}.
+
+	For flux $v_1$ (Figure \ref{fig:data_utility}a), data from about 52\% of the 240 combinations of experiments cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$).
+Only $V_1^{max}$ can be identified by about 7\% of the combinations (Figure \ref{fig:ident}a), while only about 40\% of the combinations can identify both the parameters used in the model.
+	In contrast, all the available data can identify both parameters in $v_2$ (Figure \ref{fig:data_utility}b).
+This is in agreement with the degree of identifiability of parameters used to model $v_2$ (Figure \ref{fig:ident}b).
+In the case of $v_3$ (Figure \ref{fig:data_utility}c), close to 60\% of the \textcolor{red}{4896 combinations} can identify only one parameter, while less than 10\% of the combinations can identify all three parameters in the model.
+Note that only 5\% of combinations cannot identify any parameter modeling $v_3$.
+
+	Thus, Figure \ref{fig:data_utility} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
+For instance, choosing any of the experiment combinations from the 52\% percent that does not identify any parameter in $v_1$, or the 6\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative.
+This can lead to potential waste of resources used to perform these experiments.
+
+	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:data_utility} also informs the ease of identifying parameters for a single flux.
+For the small network (Figure \ref{fig:network}), we find that all chosen combination of experiments are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:data_utility}b).
+In contrast, only about 40\% of the combinations can identify both parameters in $v_1$.
+Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ and $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$.
+A careful consideration of experiments for identifying $v_2$ is however not a major concern since there is a very high probability that experiments chosen to identify $v_1$ will also identify $v_2$.
+A similar argument can also be for determining experiments that can identify $V_e^{max}$ and $K_e^{fdp}$ in $v_5$: since $V_e^{max}$ has an identifiability of 100\%, the choice of experiments to completely identify parameters in $v_5$ rests with experiments that can identify $K_e^{fdp}$.
+
+	We believe that this information (Figure \ref{fig:data_utility}), in conjunction with the identifiability of each parameter presented in Figure \ref{fig:ident} will give an idea on the hierarchy of identifiability of each parameter, and the need for experimental design to identify each parameter.
+Additionally, having the ability to see the contribution of each type of experiment towards identifiability (Figure \ref{fig:exp_info}) would prove sufficient to decide on the total number and type of experiments that need to be performed to identify all fluxes in a kinetic model of metabolism.
+So, we look at the contribution of different types of experiments towards identifying parameters in each flux in the small network in the following section.
+
+	%Although the calculation of the utility of experiment combinations in terms of their ability to identify the maximum number of parameters gives an idea on the informativeness of experiments used to gather the data, further analysis from the perspective of identifiability of the parameters of different fluxes is required to be able to design experiments based on this information.
+Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.
 
@@
 
-	In contrast to $v_1$, we see that the different enzyme perturbation experiments have a higher contribution towards data combinations that can identify parameters for $v_2$ (Figure \ref{fig:exp_info}c and d).
+	In contrast to $v_1$, we see that both the enzyme perturbation and the acetate perturbation experiments have the same contribution towards combinations that can identify parameters for $v_2$ (Figure \ref{fig:exp_info}c and d).
 Hence, in conjunction with the high degree of identifiability seen in Figure \ref{fig:ident}b for $v_2$, we can deduce that almost all experiment types are persistently excitable for $v_2$ i.e., most experiments can bring about noticeable changes to both \textit{pep} and $v_2$.
@@
 Hence, it is theoretically possible for any of the five different experiment types to be persistently excitable to identify $v_2$.
-This is confirmed by the high degree of identifiability for both parameters of $v_2$, i.e., all data combinations can identify the parameters.
-Thus, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending on the position of the flux in the metabolic network with respect to the inputs and the outputs of the network, as well as the various regulatory interactions present within the network.
+This is confirmed by the high degree of identifiability for both parameters of $v_2$ (Figure \ref{fig:ident}b), i.e., all data combinations can identify the parameters.
+Thus, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending on the position of the flux with respect to the inputs and the outputs of the metabolic network, as well as the various regulatory interactions present within the network.
 
@@
 We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$.
-Also, since $V_e^{max}$ is globally identifiable using any experiment type, the choice and number of experiments required to identify $v_5$ completely hinges upon identifiability of $K_e^{fdp}$ from the chosen experiments.
-
-	\subsection{Utility of experimental data for practical parameter identification}\label{sec:data_utility}
-	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not been well studied (\textcolor{red}{include Rudiyanto Gunawan papers here}).
-Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
-
-	We do this by showing how useful the multiple combinations of experimental data are towards identifying parameters for each given flux (Figure \ref{fig:figure4}).
-We express the usefulness of any given data combination on the basis of the number of parameters that the combination can identify.
-The higher the number of parameters a data combination can identify, the greater is the usefulness of that combination of experiments.
-
-	The number of data combinations that can identify a specific number of parameters within a given flux is shown in Figure \ref{fig:figure4}.
-The percentage of all data combinations that can estimate a given number of parameters within a given flux are also shown in Figure \ref{fig:figure4}.
-For flux $v_1$ (Figure \ref{fig:figure4}a) about 78\% of the 306 data combinations cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$).
-Only a single parameter can be identified by about 2\% of the combinations, while only about 19\% of the data combinations can identify both the parameters used in the model.
-	In contrast, most data combinations (>90\%) are not wasted, and can identify both parameters in $v_2$ (Figure \ref{fig:figure4}b).
-Less than 5\% of the experimental data combinations are wasted from not being able to identify any parameter.
-In the case of $v_3$ (Figure \ref{fig:figure4}c), about 45\% of the 4896 data combinations can identify only one parameter, while 13\% of the combinations cannot identify any parameter.
-Only 10\% of combinations can identify $v_3$ completely.
-
-	Thus, Figure \ref{fig:figure4} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
-For instance, choosing any of the experiment combinations from the 78\% percent that does not identify any parameter in $v_1$, or the 13\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative.
-This can lead to potential waste of resources used to perform these experiments.
-
-	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:figure4} also informs the ease of identifying parameters for a single flux.
-For the small network (Figure \ref{fig:network}), we find that more than 90\% of all the data combinations tested for their ability to identify $v_2$ are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:figure4}b).
-In contrast, less than 20\% of the combinations can identify both parameters in $v_1$.
-Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ as well as $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$.
-A careful consideration of experiments for identifying $v_2$ is however not a major concern since there is a very high probability that experiments chosen to identify $v_1$ will also identify $v_2$.
-
-	Although the calculation of the utility of experiment combinations in terms of their ability to identify the maximum number of parameters gives an idea on the informativeness of experiments used to gather the data, further analysis from the perspective of identifiability of the parameters of different fluxes is required to be able to design experiments based on this information.
-Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.
+Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figure \ref{fig:ident}d), the choice and number of experiments required to identify $v_5$ completely hinges upon identifiability of $K_e^{fdp}$ from the chosen experiments.
 
@@
 	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_data_utility}}
+		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$.
+The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, c) three and d) two experiments.
+The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
+A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, \textcolor{red}{3360 combinations} are used to practically identify c) $v_3$, and 420 combinations are used to analyze the identifiability of d) $v_5$.
+Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
 		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_v1max_v2_exp_info}}
@@
 The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:exp_info}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure4/v1_v1max_v2_v3_1_data_utility}}
-		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
-The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, and c) three experiments.
-The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 3360 combinations are used to practically identify c) $v_3$.
-Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
 	\end{figure}
@@
 	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_v5_1_data_utility}}
+		\caption{Same as other data utility figure for v1 with k1cat as parameter (with enzyme data) and second roots of v3.
+Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$.}\label{fig:other_data_utility}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
 		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_k1cat_exp_info}}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v5_1_2_data_utility}}
-		\caption{data utility for flux v5 with 2 roots for Kefdp}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v3_1_2_data_utility}}
+		\caption{missing caption}%\label{fig:figure2}
 	\end{figure}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_data_utility}}
-		\caption{Same as other data utility figure for v1 with k1cat as parameter (with enzyme data) and second roots of v3.
-Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$.}\label{fig:figure4s}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_v2_v3_1_v5_2_parameter_value}}
+		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$.
+The parameter values for the only the first root of parameters in $v_3$ ($V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$) and the second root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(2)$) are shown.
+The variability in the estimated values of $V_1^{max}$ can be explained by the fact that experimental data for identification is generated using the enzyme concentrations as a variable for $v_1$, while the identification of parameters is done by subsuming the enzyme concentration into the enzyme turn over parameter ($k_1^{cat}$) to obtain $V_1^{max}$, and identifying $V_1^{max}$.
+So, the variability in values of $V_1^{max}$ reflects the variability in the metabolite concentrations (\textit{pep}, \textit{fdp}, \textit{E}) in the experimental data.
+Similarly, data is generated using the Monod-Wyman-Changeaux model for allosteric regulation for $v_3$, and a Convenience Kinetic rate law model is used for model identification purposes.}\label{fig:parameter_value_v1_vmax}
 	\end{figure}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v3_1_2_data_utility}}
-		\caption{missing caption}%\label{fig:figure2}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_experiment_data_dist}}
+		\caption{Variability in the a) metabolite concentrations and b) fluxes between different in silico experiments from which data is used for identifiability analysis.
+The variability in the predicted value of parameter $V_1^{max}$ in flux $v_1$ reflects the variability in the concentration of the metabolites (\textit{pep}, \textit{fdp}) and enzymes (\textit{E}) among the different experiments.}\label{fig:experimental_data_dist}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_k1cat_v3_2_v5_1_parameter_values}}
+		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_3$, and c) $v_5$.
+For $v_1$, we have assumed that enzyme concentration is available and have accordingly identified and estimated $k_1^{cat}$, as opposed to $V_1^{max}$.
+The parameter values for the only the second root of parameters in $v_3$ ($V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$) and the first root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(1)$) are shown.
+Note that $K_e^{fdp}(1)$ is not estimated by any combination of two experiments and $V_e^{max}$ is estimated by all combinations.
+Data is generated using the Monod-Wyman-Changeaux model for allosteric regulation for $v_3$, and a Convenience Kinetic rate law model is used for model identification purposes.}\label{fig:parameter_value_v1_kcat}
+	\end{figure}
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_k1cat_parameter_value}}
+		\caption{Distribution of identified/estimated parameter values for flux $v_1$ from different experimental data combinations.
+The box plot of the estimated values of $V_1^{max}$ and $K_1^{ac}(ne)$ are juxtaposed with the box plots of the calculated values of $k_1^{cat}$ based on the estimated value of $V_1^{max}$ and the enzyme concentrations available from the two different experiments used to estimate $V_1^{max}$ and $K_1^{ac}(ne)$}\label{fig:parameter_value_v1_v1max_kcat}
 	\end{figure}

### 2018-04-11 ident/docs/draft_2.tex
add figure captions for previously barren figures

@@
 However, since both of these are system outputs, satisfaction of this condition cannot be guaranteed without an analysis of the dynamics of the metabolic network and how changes in the input (acetate) bring about changes in the two requisite output quantities.
-Previous dynamical analysis of the network (Figure \ref{fig:network} in the Appendix) has already established the existence of a functional relationship between \textit{pep} and $v_2$, and the input acetate concentration and the levels of expression of the different enzymes within the network \parencite{Srinivasan2017}.
+Previous dynamical analysis of the network (Figure \ref{fig:network}) has already established the existence of a functional relationship between \textit{pep} and $v_2$, and the input acetate concentration and the levels of expression of the different enzymes within the network \parencite{Srinivasan2017}.
 Hence, it is theoretically possible for any of the five different experiment types to be persistently excitable to identify $v_2$.
@@
 		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_v1max_v2_exp_info}}
-		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
+		\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
 The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:exp_info}
@@
 		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_1_2_exp_info}}
-		\caption{missing caption}\label{fig:v3_exp_info}
+		\caption{The contribution of different experiments types used in a combination of three experiments ($j = \{1, 2, 3\}$) that can practically identify parameter a) $V_3^{max}(1)$, b) $K_3^{fdp}(1)$, c) $K_3^{pep}(1)$, d) $V_3^{max}(2)$, e) $K_3^{fdp}(2)$ and f) $K_3^{pep}(2)$ for $v_3$.
+parameters $V_3^{max}$, $K_3^{fdp}$, and $K_3^{pep}$ are locally identifiable and have two possible roots.
+The first set of roots are designated as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$, and the second set of roots are designated as $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
+The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.
+The contribution of the five different experiment types towards identifiability of either of the two roots for each of the three parameters is the same.}\label{fig:v3_exp_info}
 	\end{figure}
@@
 		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure2/v5_1_2_exp_info}}
-		\caption{experiment information for flux v5 with 2 roots for Kefdp}\label{fig:v5_exp_info}
+		\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_e^{max}$ and b) $K_e^{fdp}$ in $v_5$.
+Since $V_e^{max}$ is globally identifiable, it has a unique root.
+Since $K_e^{fdp}$ has two roots, is locally identifiable.
+However, results for the only practically identifiable root, based on the physical relevance constraints used in the analysis, is shown.}\label{fig:v5_exp_info}
 	\end{figure}
@@
 		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_v5_1_data_utility}}
-		\caption{Same as other data utility figure for v1 with k1cat as parameter (with enzyme data) and second roots of v3.
-Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$.}\label{fig:other_data_utility}
+		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$ and c) $v_5$.
+The total number of combinations of experiments is shown in the vertical axis, and the horizontal axis shows the maximum number of parameters that be identified using data from a combination of a) two, b) three and c) two experiments.
+The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
+A total of 240 data combinations are used for identifiability analysis for a) $v_1$, \textcolor{red}{3360 combinations} are used to practically identify b) $v_3$, and 420 combinations are used to analyze the identifiability of c) $v_5$.
+Information processed based on the identifiability of a) $v_1$ in the presence of enzyme concentration data, b) the second root of parameters in $v_3$ and c) the first root of $K_e^{fdp}$ in $v_5$ are shown.}\label{fig:other_data_utility}
 	\end{figure}
@@
 		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_k1cat_exp_info}}
-		\caption{missing caption}%\label{fig:figure2}
+		\caption{Contribution of different experiment types towards identification of a) $k_1^{cat}$ and b) $K_1^{ac}$ in $v_1$ based on using data from two different experiments.}%\label{fig:figure2}
 	\end{figure}
@@
 		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_exp_info}}
-		\caption{missing caption}%\label{fig:figure2}
+		\caption{Contribution of different types of experiments for using data from two different experiments towards identification of a) $V_3^{max}$ and b) $K_3^{fdp}$ in $v_3$ when $K_3^{pep}$ is held constant, and c) $V_3^{max}$ and d) $K_3^{pep}$ when $K_3^{fdp}$ is held constant.}%\label{fig:figure2}
 	\end{figure}
@@
 		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v3_1_2_data_utility}}
-		\caption{missing caption}%\label{fig:figure2}
+		\caption{Utility of data from combinations of two different experiments for identifying two parameters in $v_3$ when a) $K_3^{pep}$ is held constant, and when b) $K_3^{fdp}$ is held constant.
+Notice that a significant number of combinations can identify both $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is known and fixed.}%\label{fig:figure2}
 	\end{figure}

### 2018-04-11 ident/docs/draft_2.tex
moved figure to si file and re-doing the results section

@@
 \usepackage{setspace}
+\usepackage{xr} % package for linking external document references
 \usepackage[font=small,labelfont=bf,labelsep=period]{caption}
@@
 \captionsetup{font={stretch=2.0}}
+\externaldocument{sifigures} % reference to existing external document
 
@@
 
-	\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
-	A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism.
-The kinetic model is described below.
-
-	\begin{equation}\label{eq:ode1}
-	\frac{d}{dt}pep=v_1-v_2-v_4
-	\end{equation}
-	\begin{equation}\label{eq:ode2}
-	\frac{d}{dt}fdp=v_2-v_3
-	\end{equation}
-	\begin{equation}\label{eq:ode3}
-	\frac{d}{dt}E=v_5 - d E
-	\end{equation}
-	The kinetic expressions for fluxes $v_1$ through $v_5$ are given below.
-The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics.
-The acetate flux through $v_1$ is also governed by the quantity of available enzyme E.
-	\begin{equation}\label{eq:flux1}
-	v_1 = k_{1}^{cat}E\frac{ac}{ac+K_{1}^{ac}}
-	\end{equation}		
-	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}).
-Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression in Equation (\ref{eq:flux1}) for $v_1$ cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
-So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
-Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
-The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
-	\begin{align}\label{eq:flux1a}
-	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
-	\end{align}		
-	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability. 	
-	\begin{equation}\label{eq:flux2}
-	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
-	\end{equation}
-	\begin{equation}\label{eq:flux3}
-	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
-	\end{equation}
-	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
-
-	The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (section \ref{sec:ident}).
-We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica or SymPy in Python).
-In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
-The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
-	\begin{align}\label{eq:flux3_convkin}
-	v_3 = V_3^{max}\left(\frac{1}{1 + \frac{K_3^{pep}}{pep}}\right)\left(\frac{\frac{fdp}{K_3^{fdp}}}{1 + \frac{fdp}{K_3^{fdp}}}\right)
-	\end{align}
-
-	The flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
-	\begin{equation}\label{eq:flux4}
-	v_4 = k_{4}^{cat}.pep
-	\end{equation}		
-	The production of enzyme E is represented by flux $v_5$.
-This flux is inhibited by \textit{fdp} and modeled using Hill kinetics, where $K_e^{fdp}$ represents the Hill binding constant for the inhibiting metabolite \textit{fdp}, $n_e$ is the Hill exponent, and $V_e^{max}$ is the maximum reaction rate for $v_5$.
-	\begin{align}\label{eq:flux5}
-	v_5 = V_e^{max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right)
-	\end{align}
-
-	\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
-	Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions.
-One way to alter the physiological conditions is to change the substrate concentration under which the cells grow.
-In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$ (Equation \ref{eq:flux1}).
-Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.
-
-	Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network.
-The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
-Accordingly, in addition to measuring network responses to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
-In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
-The 21 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
-
-	As described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, the minimum number of experiments from which steady state data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
-For instance, the flux $v_2$ in the model described in section \ref{sec:small-model} has two parameters: $V_2^{max}$ and $K_2^{pep}$.
-Thus $\theta = \{V_2^{max}, K_2^{pep}\} \in \mathbb{R}^2$.
-So, steady state data from two distinct experiments is required for identifying $v_2$.
-Accordingly, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_2$.
-The total number of such possible combinations is 420 (21 x 20) from the 21 different experiments shown in Table \ref{tab:pval}.
-Assuming we also have data for the enzyme concentration $E$ in the expression for flux $v_1$ (Equation \ref{eq:flux1}), identifying the two parameters $k_1^{cat}$ and $K_1^{ac}$ ($\theta \in \mathbb{R}^2$) also requires steady state experimental data from two different experiments.
-
-	Similarly, for identifying $v_3$ that is described by three parameters ($\theta \in \mathbb{R}^3$), we need data from a combination of three different experiments.
-Based on the 21 experiments in Table \ref{tab:pval}, we have 7980 distinct combinations (21 x 20 x 19) of three experiments.
-However, it is important to note that not all of the either 420 combinations (for $v_1$ and $v_2$), or the 7980 combinations (for $v_3$), may be suitable for parameter estimation.
-Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers.
-Moreover, among the chosen smaller number of experimental data combinations, the methodology that we propose enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation.
-The following section describes a procedure that can be used to design experiments within this context.
-
 	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}		
@@
 
-	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
+	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} of the Supplementary Information can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
 This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguish experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
@@
 	\section{Results:}
-	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model}.
+	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model} of the Supplementary Information.
 Identifiability results for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ are also discussed in this section.
@@
 Here we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux.
-
-	First in section \ref{sec:closedform},  We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
 
@@
 
-	\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis}\label{sec:example}	
-	In this section, we illustrate the methodology that we developed to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network}) described in section \ref{sec:small-model}.
+	\subsection{Identifying parameters in kinetic models of metabolism: an example}\label{sec:example}	
+	In this section, we illustrate the proposed methodology to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model} of the Supplementary Information).
 We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration.
-
-	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
-So, from the $n_E = 21$ different experiments described in section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$.
+	\begin{align}\tag{S5}
+	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
+	\end{align}		
+	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in Section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
+So, from the $n_E = 21$ different experiments described in Section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$.
 We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively.
@@
 Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
-We apply the method described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}, and illustrated above for $v_1$, to fluxes $v_2$ and $v_3$ of the network shown in Figure \ref{fig:network}.
-We summarize the results on the identifiability of parameters in these fluxes in Figure \ref{fig:ident} and describe them below.
+	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is available, in addition to the measured concentration of the metabolites and fluxes.
+In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$.
+The globally identifiable parameter values for all fluxes modeled as Michaelis-Menten equations in provided in Figure \ref{fig:parameter_value_proof}a and \ref{fig:parameter_value_proof}b.
 
+	We apply the method described in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, and illustrated above for $v_1$, to other fluxes of the network (Figure \ref{fig:network}) not described using Michaelis-Menten equations.
+On applying this method to $v_3$, which is modeled using MWC kinetics, we found that we could not apply our methodology to determine closed form solution for all the parameters in $v_3$.
+However, when $v_3$ is instead modeled using Convenience kinetics, we find that the parameters are locally identifiable based on the definition we provide in Section \ref{sec:ident}.
+All three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ used to model $v_3$ have two possible roots.
+Although we could not determine the identifiable values for these parameters (Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}), $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value.
+
+	(Figure \ref{fig:ident_v3_var}a),  (Figure \ref{fig:ident_v3_var}b).
+
+	However, for $v_3$, we could not resolve local identifiability only through the addition of physical relevance constraints on the parameters.
+This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and Figure \ref{fig:ident_root2}b).
+Instead,
+
+	 Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable.
+Thus, the dimension of the parameter space for $v_3$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ for global identifiability.
+This informs us of the correlation between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, which results in structural non-identifiability of parameters in the model.
+
+	In addition to the observation on the degrees of identifiability of different parameter types, we also see that the enzyme kinetic rate law used to model a flux determines the ability to identify the parameters within the given model.
+For instance, as we state in the Methods section,
+
+	 We say locally identifiable due to the fact that
+
+	 In Figure \ref{fig:ident}c, we only show one of the two roots for each parameter in $v_3$.
+The identifiability of parameters based on the second root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}b.
+
+	 that require only data from two different experiments: $v_2$ and $v_5$. In applying this method to $v_5$
+
+		Similarly, when we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system.
+However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$.
+Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}.
+Even in this scenario, just as we saw earlier for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable.
+On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally identifiable parameter.
+
+	In applying this method to $v_5$ we find that analysis returns two distinct roots for $K_e^{fdp}$, denoted as $K_e^{fdp}(1)$ and $K_e^{fdp}(2)$, thereby making it a locally identifiable parameter.
+However, only one of the two roots ($K_e^{fdp}(2)$) is identifiable based on the physical relevance criteria stated in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b.
+Thus, enforcing this criteria results in global identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$.
+Next, we used data from three different experiments to identify the three parameters modeling flux $v_3$.
+Similar to $v_5$, when the parameters of the Convenience Kinetic rate law modeling $v_3$ are identified, each of the three parameters ($V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$) has two distinct roots that can be written in terms of data from three different experiments.
+Thus, these parameters are locally identifiable based on the definition we provide in section \ref{sec:ident}.
+
+	First, we would like to point out that
+
+	First, in order to show that the methodology predicts exact values of the parameters that are used to generate the in silico experimental data used for identifiability analysis, the predicted values of parameters modeling fluxes $v_1$ ($k_1^{cat}$ and $K_1^{ac}$), $v_2$ ($V_2^{max}$ and $K_2^{pep}$) and $v_5$ ($V_e^{max}$ and $K_e^{fdp}$) are shown in Figure \ref{fig:parameter_value_proof}.
+We found that the use of different enzyme kinetic rate law models (Convenience kinetics and Monod-Wyman-Changeaux kinetics) for $v_3$ had no bearing on the ability to estimate the true parameter values when the parameters are globally identifiable.
+
+	For $v_5$, the identifiability
+
+	 Accordingly, we can estimate unique values for any globally identifiable parameter from any set of experiments that satisfy practical identifiability criteria.
+
+	However, we see from the estimated parameter values for either of the two roots for all three parameters (Figure \ref{fig:parameter_value_v3}a and b) that the total number of distinct values that any of the three parameters can take is less than or equal to the total number of combinations of three experiments from which data is used for identifiability analysis.
+Also of importance is the fact that the experimental data for this analysis is generated using the MWC model for $v_3$.
+So, we hypothesized that the uncertainty seen in the parameter estimates from the identifiability analysis could be attributed to the difference between the kinetic model used to generate the data and the model used for identifiability analysis.
+On using the the Convenience Kinetic model to represent $v_3$ for both data generation as well as identifiability analysis, we see that although the uncertainty in the estimated parameter values is reduced (Figure \ref{fig:parameter_value_v3}c and d), the estimated values are not unique and uncertainty persists.
+Hence, we conclude that under the current scenario these parameters are structurally locally identifiable and practically non-identifiable.
+
+	We summarize the results on the identifiability of parameters in these and other fluxes in the small network in Figure \ref{fig:ident}, and describe them in the following section.
+
+	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}	
 	In Figure \ref{fig:ident}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b), $v_3$ (Figure \ref{fig:ident}c) and $v_5$ (Figure \ref{fig:ident}d) belonging to the network shown in Figure \ref{fig:network}.
@@
 
-	Similarly, when we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system.
-However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$.
-Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}.
-Even in this scenario, just as we saw earlier for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable.
-On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally identifiable parameter.
 The identifiability of the second root for $K_e^{fdp}$, indicated as $K_e^{fdp}(2)$, is shown in Figure \ref{fig:ident}d.
@@
 	\section{Tables, Figures and Figure Captions:}
-	\begin{table}[!thbp]
-		\caption{Table showing the perturbed values of all parameters used to generate experimental data in silico for testing practical identifiability of all fluxes in the small metabolic network.}
-		\begin{center}				
-			\begin{tabular}{lp{3.5cm}cccccp{2cm}}
-				\hline
-				Experiment Type & Perturbed Parameter (wild type value a.u.)& \multicolumn{5}{c}{Perturbed Values a.u.} & Number of Experiments\\
-				\hline
-				wild type &  &\multicolumn{5}{l}{} & 1\\
-				acetate perturbation & $acetate$ (0.1) & 0.05 & 0.09 & 0.2 & 0.5 & 1.0 & 5\\
-				$v_1$ perturbation & $k_1^{cat}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\				
-				$v_2$ perturbation & $V_2^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\								
-				$v_3$ perturbation & $V_3^{max}$ (1) & 0.5 & 0.9 & 1.1 & 1.5 & 2.0 & 5\\
-				\hline
-				\multicolumn{7}{l}{Total number of experimental data sets used for practical parameter identification:} & 21\\
-				\hline
-			\end{tabular}
-		\end{center}	
-		\label{tab:pval}
-	\end{table}
-
 	\begin{figure}[!tbhp]
@@
 b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
-		\caption{The small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
 	\end{figure}
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_experiment_data_dist}}
-		\caption{Variability in the a) metabolite concentrations and b) fluxes between different in silico experiments from which data is used for identifiability analysis.
-The variability in the predicted value of parameter $V_1^{max}$ in flux $v_1$ reflects the variability in the concentration of the metabolites (\textit{pep}, \textit{fdp}) and enzymes (\textit{E}) among the different experiments.}\label{fig:experimental_data_dist}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
 		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_k1cat_v3_2_v5_1_parameter_values}}

### 2018-04-11 ident/docs/sifigures.tex
separate si figure/table/methods file

--- 2018-04-05 ident/docs/draft_1.tex
@@
 \usepackage{setspace}
+\usepackage{xr}
 \usepackage[font=small,labelfont=bf,labelsep=period]{caption}
@@
 \addbibresource{refs/ident_refs.bib}
+
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\figurename}{Supplementary Figure}
+\renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\tablename}{{Supplementary Table}}
+
+\renewcommand\thesection{S\arabic{section}}
+\renewcommand\thesubsection{\thesection.\arabic{subsection}}
+\renewcommand{\theequation}{S\arabic{equation}}
+
+\externaldocument{draft_2} % reference to existing external document
+
 \captionsetup{font={stretch=2.0}}
-
 \begin{document}
@@
 		\begin{Large}
-			Practical Identification and Experimental Design for Parameter Estimation in Kinetic Models of Metabolism
+			Manuscript Title 
 		\end{Large}\\
@@
 	\end{center}	
-	a - Department of Chemical Engineering and Applied Chemistry, University of Toronto, Toronto, ON, Canada.\\
-	b - Institute for Biomaterials and Biomedical Engineering, University of Toronto, Toronto, ON, Canada.\\
-	{*} Corresponding author
-	\section*{Abstract}
-	\section{Introduction:}
-	\textbf{Outline:}
-	\begin{itemize}
-		\item Need for models in studying biology.
-State various applications e.g., for bio-based production, for medicine (gut microbiome interactions, stem cell engineering, studying various diseases like cancer)
-		\item Constraint-based models as a way to peek into the black box of biology without too many parameters. They however do not account for a lot of factors
-		\item Kinetics models as an improvement over constraint-based models
-		\item Advances in kinetic modeling of metabolism (Reviews?)
-		\item Parameter estimation in kinetic models as part of these advances
-	\end{itemize}
-	Kinetic models of metabolism can be used to study the dynamic characteristics of metabolic networks \parencite{Bordbar2015, Andreozzi2016}.
-In these models, ordinary differential equations (ode) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
-The matrix $\mathbf{S}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
-
-	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Link2014, Srinivasan2015}.
-Accordingly, $f$ is a nonlinear function of the metabolite concentrations, enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
-
-	Unlike CBMs, where the ability to predict the steady state responses of metabolic networks is dependent on the stoichiometry of the network, the prediction of dynamic responses of metabolic networks to different perturbations using kinetic models of metabolism is also dependent on the numerical values of the enzyme kinetic parameters ($\theta$ in Equation \ref{eq:kinstoich}) \parencite{Andreozzi2016a}.
-Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where parameter values might play a crucial role \parencite{Srinivasan2015, Vital-Lopez2006}.
-The use of in vitro, or unreliable in vivo parameter estimates reduces confidence in the model predicted behaviour \parencite{Andreozzi2016a}.
-Consequently, the reduction in confidence hampers the use of these models to gain insight into the functioning of metabolic networks \parencite{Chakrabarti2013a, Bordbar2015}.
-The corresponding increase in uncertainty in model predicted responses becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design.
-
-	Parameter estimation methods based on optimization principles are typically used to determine true parameter values based on available experimental data.
-Under the assumption that all time dependent intracellular metabolite concentrations can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:kinstoich}) to estimate the values of enzyme kinetic parameters, $\theta$, based on the measured data.
-	The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration at each time point, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}).
-The parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).
-
-	However, not all metabolite concentrations used in the model (Equation \ref{eq:kinstoich}) can be measured.
-Additionally, measurable fluxes in the metabolic network also need to be included as part of the parameter estimation problem.
-In such scenarios, the parameter estimation problem is modified to suit a new system of equations shown below (Equation \ref{eq:output}).
-The new system of equations is obtained by augmenting the original system (Equation \ref{eq:kinstoich}) with Equation (\ref{eq:output}c) that models the relationship between the measurable metabolite concentrations and fluxes ($y$) and the unmeasured concentrations ($x$) that are used in the original model (Equation \ref{eq:kinstoich}) above.
-The parameter vector ($\theta$) is augmented with additional parameters that define this relationship.
-These additional parameters also need to be estimated.
-
-	In systems identification, the measured concentrations and fluxes ($y$) are called output or observed variables, and the unmeasured concentrations ($x$) are called the state variables.
-For estimating $\theta$, the metabolite concentrations $x$ in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the output variables $y$.
-
-	The ability to determine unique solutions to parameters $\theta$ is governed by the identifiability of these parameters in the model \parencite{McLean2012, Raue2009a}.
-The identifiability of parameters in nonlinear models can be classified into two categories: structural (or a priori) and practical (or posterior) identifiability.
-	Any system (Equation \ref{eq:output}) is said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}) below.
-	\begin{align}\label{eq:stident}
-	\Phi(\theta_1,u) = \Phi(\theta_2,u) \iff \theta_1 = \theta_2
-	\end{align}
-	Accordingly, the system can have a unique solution, a finite number of non-unique solutions or an infinite number of solutions for all input functions, and is said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
-So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unobservable state variables and the observable output variables.
-Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
-
-	Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
-So, the effect of the available experimental data on the ability to estimate unique parameter values is determined by the practical identifiability of the parameter.
-Accordingly, practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
-
-	Thus, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but also are parameterized in such a way that the value of these parameters can be estimated from measurable data.
-On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
-
-	Methods and tools for structural identification of parameters based on differential algebra \parencite{Ljung1994, Audoly2001, Bellu2007} and profile likelihood \parencite{Raue2009a} are available.
-However, only the profile likelihood-based methods enable experimental design by facilitating practical identification of parameters.
-Nonetheless, this method still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters, and hence still suffers from all the inherent difficulties associated with obtaining global optimal solutions for non-convex optimization problems.
-This also makes it un-scalable for experimental design and practical identifiability of parameters in kinetic models of large metabolic networks.
-
-	In this paper, we propose a scalable methodology to establish practical identifiability for parameters in kinetic models of metabolism using steady state concentration and flux data.
-We present a computer algebra-based method that can facilitate experimental design through practical identification of parameters separately for each individual reaction within a metabolic network.
-For the purposes of this method we assume that all intracellular metabolite concentrations and fluxes can be measured.
-We identify and design experiments to estimate parameters in a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} to illustrate the utility of our method.
-%We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
-
-	\section{Methods}\label{sec:methods}
-
-	\begin{subequations}\label{eq:kinstoich}
-		\begin{align}
-		\dot{x} = \mathbf{S}v\\
-		v = f(x, \theta, u)
-		\end{align}
-	\end{subequations}
-
-	\begin{subequations}\label{eq:lsqopt}
-		\begin{align}
-		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{x_{kl}^*-x_{kl}}{\sigma_{kl}^*}\right)^2\\
-		&\theta_l \le \theta \le \theta_u
-		\end{align}
-	\end{subequations}
-
-	\begin{subequations}\label{eq:output}
-		\begin{align}
-		\dot{x} = \mathbf{S}v\\
-		v = f(x, y, \theta, u)\\
-		\dot{y} = h(x, y, \theta, u)
-		\end{align}
-	\end{subequations}
-
-	\subsection{A method to determine practical identifiability of kinetic models of metabolism}\label{sec:ident}
-	We provide the mathematical framework for practical identification of parameters in kinetic models of metabolism in this section.
-A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
-In kinetic models of metabolic networks (Equation \ref{eq:kinstoich}), the fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
-The value of every flux $v_i$ is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
-Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations (Figure \ref{fig:ident-flowchart}a).
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
-		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
-a) The steps for the construction of a kinetic model of a metabolic network.
-The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
-The identifiability of parameters for each flux can be established independently.
-b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
-	\end{figure}
-
-	Let $\theta \in \mathbb{R}^p$ in Equation (\ref{eq:output}b) for each flux $v_i$ in the network.
-As stated earlier, for each experiment $j = {1, 2, ..., n}$, we assume that all metabolite concentrations ($x$) and reaction fluxes ($v$) are measurable.
-The pertinent information for each experiment is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
-
-	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, we describe a computer algebra-based method.
-The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\theta$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
-This is done by solving a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
-	\begin{align}\label{eq:nonlineq}
-	v_{i, k} = f_k(\mathbf{x}_k,\theta, u_k) && \forall k=\{1, 2, ..., p\}\subset\{1, 2, ..., n\}
-	\end{align}
-	Each equation in (\ref{eq:nonlineq}), indicated by the index $k$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for $v_i$, described earlier in Equation (\ref{eq:output}b), written for concentrations and fluxes obtained from experiment $k$.
-Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for parameters in $\theta$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
-	\begin{align}\label{eq:theta-eq}
-	\theta_k = g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = \frac{N_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}{D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})}
-	\end{align}
-	The identifiability of parameter $\theta_k$ for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
-Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
-
-	In the following sections we provide a previously published kinetic model of a small gluconeogenic network (section \ref{sec:small-model}), followed by a demonstration of our methodology to establish practical identifiability for one of the fluxes in this network (section \ref{sec:example}).
-
-	\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
-	A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate the utility of identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism.
-The kinetic model is described below.
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
-		\caption{The small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
-	\end{figure}	
-	\begin{equation}\label{eq:ode1}
-	\frac{d}{dt}pep=v_1-v_2-v_4
-	\end{equation}
-	\begin{equation}\label{eq:ode2}
-	\frac{d}{dt}fdp=v_2-v_3
-	\end{equation}
-	\begin{equation}\label{eq:ode3}
-	\frac{d}{dt}E=v_{e,max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right) - d E
-	\end{equation}
-	The kinetic expressions for fluxes $v_1$ through $v_4$ are given below.
-The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics.
-The acetate flux through $v_1$ is also governed by the quantity of available enzyme E.
-	\begin{equation}\label{eq:flux1}
-	v_1 = k_{1}^{cat}E\frac{ac}{ac+K_{1}^{ac}}
-	\end{equation}	
-	\begin{equation}\label{eq:flux2}
-	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
-	\end{equation}
-	\begin{equation}\label{eq:flux3}
-	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
-	\end{equation}
-	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
-The flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
-	\begin{equation}\label{eq:flux4}
-	v_4 = k_{4}^{cat}.pep
-	\end{equation}
-
-	\subsection{Identifiability of parameters in a kinetic model of gluconeogenesis}\label{sec:example}
-	Here, we demonstrate the use of our computer algebra-based methodology to establish practical identifiability of parameters for flux $v_1$ in the small model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model}.
-
-	In flux $v_1$, the concentration of the enzyme E is used as a variable.
-Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, the expression previously given for $v_1$ (Equation \ref{eq:flux1}) cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
-So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
-Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
-The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
-	\begin{align}\label{eq:flux1a}
-	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
-	\end{align}		
-	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability.
-
-	As mentioned in section \ref{sec:intro}, practical identifiability is a necessary condition to estimate true values of both $V_1^{max}$ and $K_1^{ac}$ from experimental data.
-Here, we assume that the concentrations and fluxes from at least two different experiments are available i.e., in Equation (\ref{eq:nonlineq}) $k = \{1, 2\}$ .
-We label the available concentrations and fluxes as ${ac}^{(k)}$ and ${v_1}^{(k)}$, respectively.
-Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
-	\begin{align}\label{eq:nonlin-flux2}
-	{v_1}^{(k)} = V_{1}^{max}\frac{ac^{(k)}}{ac^{(k)}+K_{1}^{ac}(ne)} &&  k=\{1, 2\}
-	\end{align}
-
-	Solving this simultaneous system of $k$ equations using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
-These expressions have the form previously shown in Equation (\ref{eq:theta-eq}).
-	\begin{subequations}\label{eq:v1_par}
-		\begin{align}		
-		V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
-		K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
-		\end{align}
-	\end{subequations}
-	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations.
-Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
-
-	\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
-	Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions.
-One way to alter the physiological conditions is to change the substrate concentration under which the cells grow.
-In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$.
-Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.
-
-	Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network.
-The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
-Accordingly, in addition to measuring network response to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
-
-	In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
-The 18 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
+\section{Supplementary Tables}
 \begin{table}[!thbp]
@@
 
-	The minimum number of experiments from which data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
-For instance, as demonstrated above, data from two distinct experiments is required for identifying $v_1$, as $v_1$ has two parameters ($\mathbb{R}^2$).
-$v_2$ is also modeled by a parameter space with dimensions of $\mathbb{R}^2$.
-Hence, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_1$ and $v_2$.
-The total number of such possible combinations is 306 (18 x 17) from the 18 different experiments shown in Table \ref{tab:pval}.
-Similarly, for identifying $v_3$, that is described by three parameters ($\mathbb{R}^3$), we need data from a combination of three different experiments.
-Based on the 18 experiments in Table \ref{tab:pval}, we have 4896 distinct combinations (18 x 17 x 16) of three experiments.
+\clearpage
 
-	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
-	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
-We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
+\section{Supplementary Figures}	
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/mwc_ck_experimental_data}}
+	\caption{Variability in the metabolite concentrations and fluxes between different in silico experiments from which data is used for identifiability analysis.
+a) Metabolite (\textit{pep}, \textit{fdp}) and enzyme concentration (\textit{E}), and b) flux data ($v_1$, $v_2$, $v_3$, $v_4$, $v_5$ and $v_6$) are generated using the Monod-Wyman-Changeaux model for $v_3$.
+The Convenience Kientic rate law model is used for $v_3$ to generate in silico c) metabolite and enzyme concentration, and d) flux data through enzyme expression parameter and substrate concentration perturbations.}\label{fig:experimental_data_dist}
+\end{figure}
 
-	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
-On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
-Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
-In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
-Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
-Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_k1cat_v2_v5_2_parameter_values}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, and c) $v_5$.
+For $v_1$, we have assumed that enzyme concentration is available and have accordingly identified and estimated $k_1^{cat}$, as opposed to $V_1^{max}$.
+The parameter values for only the second root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(2)$) is shown, since $K_e^{fdp}(1)$ is not estimated by any combination of two experiments, and $V_e^{max}$ is estimated by all combinations.
+Data is generated using the Monod-Wyman-Changeaux model for allosteric regulation for $v_3$.}\label{fig:parameter_value_proof}
+\end{figure}
 
-	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}
 \begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.6\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
-		\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v3_1_2_ck_1_2_parameter_values}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$ when the a) and b) Monod-Wyman-Changeaux model for allosteric regulation for $v_3$ is used to generate data, c) and d) experimental data is generated using the Convenience Kinetic rate law model with activation for $v_3$.
+The left column corresponds to the first root for all three parameters: $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$, and the second column corresponds to the second root of all three parameters in $v_3$: $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.}\label{fig:parameter_value_v3}
 \end{figure}
-	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
-This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguishing experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
-In doing so, it is possible to obtain a minimal and informative collection of experiments that can be performed to identify as many model parameters as possible (Figure \ref{fig:ident-design}).
-Consequently, the set of experiments can be used to estimate the identifiable parameters in the model.
-This is formally explained below.
 
-	The identifiability of each parameter based on each experiment indexed as $j = \{1, ..., n\}$ is established based on the methodology described previously in section \ref{sec:ident} and demonstrated in Section \ref{sec:example} (Figure \ref{fig:ident-flowchart}b).
-Subsequently, for any flux $v_i$ and for any $p$ combinations of indices $j$, if the experimental concentrations ($\mathbf{x}_j$) and fluxes ($\mathbf{v}_j$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$, i.e., $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ for any $k$, then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
-Consequently, the corresponding experiment cannot be used for parameter estimation and needs to be discarded from the set of all necessary experiments.
-Furthermore, another experiment from $j = \{1, ..., n\}$ needs to be selected such that parameter $\theta_k$ is identifiable.
-This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
-In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
-Note that if none of the $n$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
-This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.
+\clearpage
 
-	\section{Results:}
-	First in section \ref{sec:closedform}, we show that the ability to establish practical identifiability of kinetic parameters in metabolic network models using our methodology, described in section \ref{sec:ident}, relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
-We use the gluconeogenic model (Figure \ref{fig:network}) described in section \ref{sec:example} as an example.
-	Then, in section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models of metabolic networks.
-We do this by looking at the number of combinations that can identify the maximum number of parameters.
-	In section \ref{sec:trends} we show that the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetic rate law used to describe the flux.
-	In section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment.
-In this section we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux.
-Finally, we discuss some results arising out of the use of our methodology to determine the identifiability of parameters when data with additive noise is used in section \ref{sec:noise}.
+\section{Supplementary Methods}
+\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
+A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism.
+The kinetic model is described below.
 
-	\subsection{Nonlinearity of enzyme kinetic rate law expression affects identifiability analysis}\label{sec:closedform}
-	Practical identifiability of parameters in kinetic models of metabolism using the methodology described in section \ref{sec:methods} is governed by the nonlinear complexity of the enzyme kinetic rate law used to model a specific flux.
-We demonstrate one example of how the methodology works in section \ref{sec:example} for parameters of flux $v_1$.
-The expression in Equation (\ref{eq:v1_par}) is obtained by using a computer algebra system.
-To recall, in this specific case, a computer algebra system is used to solve for the parameters of a flux described using the Michaelis-Menten kinetic rate law when data from two different experiments is available.
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
+	\caption{The previously published small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
+\end{figure}
 
-	However, we find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica and SymPy in Python).
+\begin{equation}\label{eq:ode1}
+\frac{d}{dt}pep=v_1-v_2-v_4
+\end{equation}
+\begin{equation}\label{eq:ode2}
+\frac{d}{dt}fdp=v_2-v_3
+\end{equation}
+\begin{equation}\label{eq:ode3}
+\frac{d}{dt}E=v_5 - d E
+\end{equation}
+The kinetic expressions for fluxes $v_1$ through $v_5$ are given below.
+The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics.
+The acetate flux through $v_1$ is also governed by the quantity of available enzyme E.
+\begin{equation}\label{eq:flux1}
+v_1 = k_{1}^{cat}E\frac{ac}{ac+K_{1}^{ac}}
+\end{equation}		
+The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}).
+Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression in Equation (\ref{eq:flux1}) for $v_1$ cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
+So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
+Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
+The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
+\begin{align}\label{eq:flux1a}
+v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
+\end{align}		
+We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability. 	
+\begin{equation}\label{eq:flux2}
+v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
+\end{equation}
+\begin{equation}\label{eq:flux3}
+v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
+\end{equation}
+The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
+
+The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (section \ref{sec:ident}).
+We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica or SymPy in Python).
 In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
 The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
-
 \begin{align}\label{eq:flux3_convkin}
@@
 
-	Using this expression for analysis, we find that each of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different closed-form expressions owing to the presence of a square root term in their solutions.
-These distinct expressions are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, $K_3^{pep}(1)$, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$.
+The flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
+\begin{equation}\label{eq:flux4}
+v_4 = k_{4}^{cat}.pep
+\end{equation}		
+The production of enzyme E is represented by flux $v_5$.
+This flux is inhibited by \textit{fdp} and modeled using Hill kinetics, where $K_e^{fdp}$ represents the Hill binding constant for the inhibiting metabolite \textit{fdp}, $n_e$ is the Hill exponent, and $V_e^{max}$ is the maximum reaction rate for $v_5$.
+\begin{align}\label{eq:flux5}
+v_5 = V_e^{max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right)
+\end{align}
 
-	The impact of nonlinearity of a kinetic rate law expression affecting the practical identifiability of parameters in the expression may not be an unique problem isolated to the system that we are investigating.
-Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms \parencite{Chassagnole2002a, Peskov2012, VanHeerden2014}.
-However, the identifiability of the parameters used in these models has never been truly examined.
-Unlike the case of the MWC model, if determining closed-form expressions for the parameters of these models are tractable, we believe our methodology can help in elucidating the identifiability of these parameters.
-If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation and model validation.
+\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
+Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions.
+One way to alter the physiological conditions is to change the substrate concentration under which the cells grow.
+In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$ (Equation \ref{eq:flux1}).
+Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.
 
-	Next, we look at some of the results we obtained for the practical identification of different parameters for all three fluxes $v_1$, $v_2$ and $v_3$ in the small gluconeogenic network model described in section \ref{sec:small-model}, with the flux for $v_3$ described by Equation (\ref{eq:flux3_convkin}).
-We start by looking at one of motivations for the need for designing experiments to collect data for parameter estimation.
+Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network.
+The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
+Accordingly, in addition to measuring network responses to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
+In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
+The 21 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
+The box plots in Figure \ref{fig:experimental_data_dist} show the distribution in the in silico metabolite and enzyme concentration and flux data obtained from these experiments.
 
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure4/v1_V1max_v2_v3_value1_data_utility}}
-		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$ and c) $v_3$.
-The total possible number of parameters that can be identified by data from combinations of a) two, b) two, and c) three experiments is shown on the horizontal axis.
-The vertical axis represents the total number of combinations that can identify the corresponding number of parameters.
-The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 306 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, and 4896 combinations are used to practical identifiability of c) $v_3$.
-Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:figure4}
-	\end{figure}	
-	\subsection{Utility of experimental data for practical parameter identification}\label{sec:data_utility}
-	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community, the need for experimental design methods to satisfy the data needs for parameter estimation are not considered seriously.
-Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
+As described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, the minimum number of experiments from which steady state data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
+For instance, the flux $v_2$ in the model described in section \ref{sec:small-model} has two parameters: $V_2^{max}$ and $K_2^{pep}$.
+Thus $\theta = \{V_2^{max}, K_2^{pep}\} \in \mathbb{R}^2$.
+So, steady state data from two distinct experiments is required for identifying $v_2$.
+Accordingly, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_2$.
+The total number of such possible combinations is 420 (21 x 20) from the 21 different experiments shown in Table \ref{tab:pval}.
+Assuming we also have data for the enzyme concentration $E$ in the expression for flux $v_1$ (Equation \ref{eq:flux1}), identifying the two parameters $k_1^{cat}$ and $K_1^{ac}$ ($\theta \in \mathbb{R}^2$) also requires steady state experimental data from two different experiments.
 
-	We do this by showing how useful the multiple combinations of experimental data are towards identifying parameters for each given flux (Figure \ref{fig:figure4}).
-We express the usefulness of any given data combination on the basis of the number of parameters that the combination can identify.
-The higher the number of parameters a data combination can identify, the greater is the usefulness of that combination of experiments.
-
-	The number of data combinations that can identify a specific number of parameters within a given flux is shown in Figure \ref{fig:figure4}.
-The percentage of all data combinations that can estimate a given number of parameters within a given flux are also shown in Figure \ref{fig:figure4}.
-For flux $v_1$ (Figure \ref{fig:figure4}a) about 78\% of the 306 data combinations cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$).
-Only a single parameter can be identified by about 2\% of the combinations, while only about 19\% of the data combinations can identify both the parameters used in the model.
-	In contrast, most data combinations (>90\%) are not wasted, and can identify both parameters in $v_2$ (Figure \ref{fig:figure4}b).
-Less than 5\% of the experimental data combinations are wasted from not being able to identify any parameter.
-In the case of $v_3$ (Figure \ref{fig:figure4}c), about 45\% of the 4896 data combinations can identify only one parameter, while 13\% of the combinations cannot identify any parameter.
-Only 10\% of combinations can identify $v_3$ completely.
-
-	Thus, Figure \ref{fig:figure4} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
-For instance, choosing any of the experiment combinations from the 78\% percent that does not identify any parameter in $v_1$, or the 13\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative.
-This can lead to potential waste of resources used to perform these experiments.
-
-	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:figure4} also informs on the ease of identifying parameters for a single flux.
-For the small network (Figure \ref{fig:network}), we find that more than 90\% of all the data combinations tested for their ability to identify $v_2$ are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:figure4}b).
-In contrast, less than 20\% of the combinations can identify both parameters in $v_1$.
-Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ as well as $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$.
-A careful consideration of experiments for identifying $v_2$ is however not a major concern since there is a very high probability that experiments chosen to identify $v_1$ will also identify $v_2$.
-
-	Although the calculation of the utility of experiment combinations in terms of their ability to identify the maximum number of parameters gives an idea on the informativeness of experiments used to gather the data, further analysis from the perspective of identifiability of the parameters of different fluxes is required to be able to design experiments based on this information.
-Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.
-
-	\subsection{Maximum reaction rates are more identifiable than enzyme binding constants}\label{sec:trends}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight]{figures/figure1/v1_V1max_v2_v3_identifiability}}
-		\caption{The number of data combination from 18 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$ and c) $v_3$ when there is no noise in the input experimental data.
-The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
-Since the total number of experiments used to identify parameters in $v_1$ and $v_2$ is two, the total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is lower (at 306) that the total number of combinations used to test identifiability of parameters for $v_3$ (at 4896), which is modeled with three parameters and requires data from at least three different experiments.}\label{fig:figure1}
-	\end{figure}
-	In Figure \ref{fig:figure1} we show the number and percentage of combinations that are capable of identifying each parameter in each flux.
-Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
-The three panels in Figure \ref{fig:figure1} represent the degree of identifiability for parameters modeling the three different fluxes of the small network individually.
-
-	In regards to the identifiability of $v_1$ (Figure \ref{fig:figure1}a), $v_2$ (Figure \ref{fig:figure1}b) and $v_3$ (Figure \ref{fig:figure1}c), the degree of identifiability of the maximum reaction rates ($V_i^{max}$) in each of the three fluxes is higher than the degree of identifiability of the corresponding enzyme binding ($K_i$) constants and the allosteric activation constant ($K_3^{pep}$).
-We observe this trend irrespective of the fact that only  $v_1$ and $v_2$ are represented by the same enzyme kinetic rate law (Michaelis-Menten), while the convenience kinetic rate law is used to model $v_3$.
-These observations lead us to conclude that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
-
-	From Figure \ref{fig:figure1} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified.
-In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
-Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range.
-Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
-In linear systems, persistence of excitation of any input signal can be theoretically guaranteed.
-However, for nonlinear systems, like the metabolic network we deal with in this paper, persistence of excitation of input signals cannot be theoretically guaranteed, and should be assessed on a case by case basis.
-The lack of persistence of excitation requires the design of experiments to satisfy the data needs for complete system identification.
-
-	In the case of identifying fluxes within a metabolic network individually, each flux is not independent from the other fluxes that form the metabolic network.
-Each flux is connected to one or more fluxes through the metabolites that not only act as substrates and products, but also as allosteric activators and inhibitors (e.g., effect of \textit{pep} on $v_3$ in Figure \ref{fig:network}).
-Hence, careful consideration of experiments is a necessity so that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network.
-In the next section we discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of experiments from which data is collected for identification.
-
-	\subsection{Experiment requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}	
-	In this section, we demonstrate a specific case of lack of persistence of excitation and the subsequent experimental design required to overcome this hurdle for parameter identifiability for a small nonlinear metabolic network model.
-Earlier, we mentioned that the informativeness of experiments for identifying parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter.
-For example, we demonstrated in section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them (Equation \ref{eq:v1_par}).
-Thus, in this instance, the informativeness of the experiments (changes in the input acetate concentration and $v_1$) for identifying parameters of $v_1$ is determined by the ability of the input change to effect a change in the measured value of $v_1$.
-Similarly, the identification of parameters for $v_2$ requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}.
-%, given both these variables are outputs of the model describing $v_2$.
-
-	As such, as per Figure \ref{fig:figure1}a only 20\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$.
-Whereas, more than 90\% of the available data combinations satisfy the requirements for identifying $v_2$ (Figure \ref{fig:figure1}b).
-Thus, identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for both $v_1$ and $v_2$.
-Recall that we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Section \ref{sec:experiments}).
-	In order to determine the type of experiments that help satisfy identifiability conditions for both parameters of $v_1$, we look at the distribution of the five different types in all data combinations that can identify these 2 parameters (Figure \ref{fig:figure2}a and b).
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_only_v2_spider}}
-		\caption{The contribution of different experiments types used in a combination of two experiments ($k = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:figure2}
-	\end{figure}
-
-	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments.
-This matches with requirements for identifiability on the basis of the informativeness of experiments that was laid out earlier i.e., experiments must be able distinguish between different values of $v_1$ and acetate.
-Accordingly, only about 20\% of the 306 different experiment combinations used to test identifiability are informative enough to discriminate both acetate as well as $v_1$, and can identify $V_1^{max}$ and $K_1^{ac}(ne)$ (Figure \ref{fig:figure1}a).
-
-	In contrast to $v_1$, we see that the different enzyme perturbation experiments have a higher contribution towards data combinations that can identify parameters for $v_2$ (Figure \ref{fig:figure2}c and d).
-Hence, in conjunction with the high degree of identifiability seen in Figure \ref{fig:figure1}b for $v_2$, we can deduce that almost all experiment types are persistently excitable for $v_2$ i.e., most experiments can bring about noticeable changes to both \textit{pep} and $v_2$.
-Consequently, in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative enough to identify $v_2$.
-This observations is in line with the one made in section \ref{sec:data_utility} based on the total number of combinations that can identify both parameters in $v_2$.
-
-	Both the degree of identifiability of parameters and the informativeness of the corresponding experiments used to identify them can be explained by the position of the flux in the metabolic network.
-The position of any given flux in the metabolic network determines the specific experiment that is persistently excitable enough to identify the parameters of that flux.
-This dependency of experiment informativeness on the position of the flux can be further elucidated using $v_1$ and $v_2$ as examples.
-We know from Equation (\ref{eq:v1_par}) that identifiability of $v_1$ requires changes in both acetate and $v_1$.
-We also know, based on our knowledge of the Michaelis-Menten kinetic rate law that changes in the substrate concentration of a reaction can bring about a nonlinear change in the value of the corresponding reaction rate.
-In this specific metabolic network, since the substrate is an input variable to the model, and $v_1$ is the corresponding uptake flux, the substrate can be easily perturbed to create persistently excitable experiments to identify parameters in $v_1$.
-We can generalize this observation for the identification of all uptake fluxes in all metabolic networks, i.e., at a minimum, a change in the input substrate concentration may be necessary for an informative experiment to identify the uptake flux parameters.
-
-	On the other hand, the Michaelis-Menten model for $v_2$ also requires changes in \textit{pep} and $v_2$ for persistently excitable experiments to identify $v_2$.
-However, since both of these are system outputs, satisfaction of this condition cannot be guaranteed without an analysis of the dynamics of the metabolic network and how changes in the input (acetate) bring about changes in the two requisite output quantities.
-Previous dynamical analysis of the network (Figure \ref{fig:network} in the Appendix) has already established the existence of a functional relationship between \textit{pep} and $v_2$, and the input acetate concentration and the levels of expression of the different enzymes within the network \parencite{Srinivasan2017}.
-Hence, it is theoretically possible for any of the five different experiment types to be persistently excitable to identify $v_2$.
-This is confirmed by the high degree of identifiability for both parameters of $v_2$, where in more than 90\% of all data combinations can identify the parameters.
-Thus, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending on the position of the flux in the metabolic network with respect to the inputs and the outputs of the network, as well as the various regulatory interactions present within the network.
-We can extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_3$ as well (Figure \ref{fig:figure2_v3}).
-
-	\subsection{Effect of noise on degree of identifiability is dependent on nonlinearity of enzyme kinetic rate law models}\label{sec:noise}
-	In reality, experimental data from biological systems is usually noisy.
-Noise in data gathered from experiments is associated with biological noise attributable to the stochasticity of cellular function, as well as measurement noise associated with the techniques used to obtain concentration and flux measurements.
-However, the previous sections describe practical identifiability for parameters obtained using noise-free in silico experimental data.
-So, in order to illustrate the ability of our method to test the practical identifiability of parameters under realistic circumstances, we apply our methodology on in silico data (concentration and flux) with 5\% additive noise.
-
-	50 different samples of noise with 0 mean and 5\% standard deviation are drawn from a normal distribution and added to the steady state experimental data generated on the basis of experiments described in Section \ref{sec:experiments} to generate 50 different samples of noisy data.
-Subsequently, all 50 samples are tested for their ability to practically identify all three fluxes of the small network (Figure \ref{fig:network}).
-The degree of identifiability of the different parameters using noisy experimental data are shown in the Appendix (Figure \ref{fig:figure1_noisy}).
-
-	We hypothesize that the nonlinearity of the enzyme kinetic rate law used to model a flux, and consequently the complexity of the closed form expression obtained for the various parameters used in the model has an impact on the ability of noisy data to practically identify the parameters.
-The difference in the identifiability attributable to the aforementioned factors can be seen in the difference in the degree of identifiability for the parameters of $v_1$, $v_2$ and $v_3$ in the small metabolic network that we use to demonstrate our methodology.
-
-	Despite the presence of noise, the degree of identifiability of $v_1$ and $v_2$ (Figure \ref{fig:figure1_noisy}a and b in the Appendix) does not change from the case where no noise is present in the experimental data (Figure \ref{fig:figure1}).
-However, the degree of identifiability for parameters of $v_3$ is affected by the presence of noise in the data (Figure \ref{fig:figure1_noisy}c in the Appendix).
-This effect is represented by the presence of a non-zero standard deviation and corresponding error bars on the graphs in the degrees of identifiability of parameters modeling $v_3$ (Figure \ref{fig:figure1_noisy}c in the Appendix).
-The standard deviation and the error bars represent the presence of data combinations that can identify a given parameter under certain values of noise, but not do so under different noise values.
-In the presence of 5\% additive noise in the data, the conditions under which parameters for $v_1$ and $v_2$ can be identified are strictly satisfied by the same data combinations that can identify these parameters in the absence of noise, i.e., the changes in the acetate concentration (for $v_1$) and \textit{pep} (for $v_2$) are significant enough to enable identification.
-The contribution of different experiment types towards identifiability also remains the same (figure not shown separately).
-
-	However, as stated in the previous section, the nonlinearity of the original kinetic rate law expression for $v_3$, and the complexity of the closed-form expression obtained for the parameters in the rate law model preclude us from testing the presence, and consequently the satisfaction of these conditions for parameters in $v_3$.
-Furthermore, the conditions for identifiability of $v_3$ parameters are not as simple as either requiring just a simple difference in \textit{fdp} or \textit{pep} or that of both concentrations.
-Thus, while under some noise values these identifiability conditions may be satisfied, it may not be possible to satisfy these conditions under other noise values.
-Accordingly, we end up with a distribution of experimental data combinations that can identify $v_3$ parameters depending on noise levels in the data.
-It is significant to note that the variability in the degree of identifiability of parameters is small (< 2\%) when noise is present.
-
-	These observations, in conjunction with the estimation of the degree of identifiability of the different parameters (Figure \ref{fig:figure1}) help in determining not only the nature of experiments required for parameter identifiability (Figure \ref{fig:figure2}), but also the order in which the type of experiments necessary for identifying the entire metabolic network needs to be evaluated.
-
-	For instance, in the case of the example network that we use in this paper, given the high degree of identifiability for $v_2$ and large spread in the frequency of different experiment types that can contribute to identifiability for $v_2$, the focus should be on first choosing experiments for identifying $v_1$ followed by selecting experiments for identifying $v_3$.
-It is possible that the three or at most four experiments chosen to identify parameters of both $v_1$ and $v_3$ would suffice to identify $v_2$ as well without the need to perform additional experiments.
-
-	\section{Discussions}\label{sec:discussion}	
-	Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data.
-The presence of noise is typically said to be a significant factor that results in non-identifiability.
-However, the primary reason for non-identifiability of parameters from experimental data can be attributed to the lack of information about the dynamics of the system whose parameters are being estimated within the chosen experimental data.
-The informativeness of experiments can be tied back to their ability to discriminate the dynamics of the system under two or more different input conditions.
-Thus, the presence of noise only serves to exacerbate the inability of experiments to discriminate the dynamics of the systems.
-
-	Although methods have been developed for practical parameter identification and experimental design, existing methods for experimental design based on practical identification of parameters are based on solving nonlinear least squares problems using optimization approaches that cannot guarantee global optimal solutions \parencite{Raue2009a}.
-In this document, we have presented a scalable method to practically identify parameters in kinetic models of metabolism, and use it to design experiments that are minimal and informative for estimating the parameters that does not require solutions to non-convex optimization problems.
-While some authors have eschewed deterministic parameter estimation techniques in favour of Bayesian methods based on probabilistic estimation of parameters \parencite{Saa2016, Saa2016a}, The need to establish practical identifiability of parameters, even for these methods, still stands.
-Additionally, the ability to design experiments to get good priors for Bayesian approaches to parameter estimation can also be fulfilled through practical identifiability analysis.
-
-	Due to their fundamental aspect of requiring solutions to non-convex optimization problems, existing methods for practical identifications may not be scalable for identification and experimental design for parameters of large metabolic networks.
-By establishing identifiability for each flux within a metabolic network individually, we hope to overcome the scalability obstacle.
-Furthermore, we believe our method offers an algorithmic alternative to determine persistently excitable experiments that can enable identification of all fluxes within a metabolic network.
-Using a small metabolic network for gluconeogenesis, we have demonstrated that the identifiability of parameters for a given flux is dependent on the position of the flux within the metabolic network.
-We have also shown the ability to use our analysis to design the minimal number of experiments that are most informative for identifying all fluxes within a metabolic network.
-
-	Our work adds to this existing body of work wherein we develop a method for practical identifiability tailored for use with nonlinear enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks.
-With our work we hope to change the status quo in the application of systems identification techniques for kinetic models of metabolic networks.
-Our methodology fills the niche gap of experimental design for parameter estimation by providing a way to design informative experiments to obtain data required for parameter estimation by spending the least amount of resources.
-	In the future, we believe our work can be extended and formulated as a mixed integer linear programming problem that can be solved to determine the type and total minimum number of experiments necessary to estimate all parameters in kinetic models of genome-scale metabolic networks.
-
-	%Nonetheless, all these explanations could only be verified due to the fact that the conditions for identifiability for both $v_1$ and $v_2$ are simple and readable due to the use of the Michaelis-Menten rate law to model these fluxes.
-Instead, in addition to presence of an allosteric interaction, when the convenience kinetic rate law is used to model this activation in $v_3$, we find that the complex expressions obtained for the three parameters preclude such an in-depth insight into the nature of persistently excitable experiments to identify  $v_3$.
-Thus, the only avenue through which experiments can be designed for the identification of parameters of $v_3$ would be through our proposed methodology wherein manual curation of experimental combinations required for identifiability are not necessary.
-
-	%Based on the results we obtained in establishing the identifiability of parameters modeling all the fluxes of a small metabolic network, in the future we hope to extend this methodology to a larger metabolic network with more reactions.
-	%Apart from determining experiments that are informative enough to identify parameters for each flux in a kinetic model, experimental design can also be used to design and develop a minimal set of experiments that informative enough to estimate all parameters in a metabolic network model.
-However, there is not only a scarcity of experimental design methods in literature devoted to the purpose of designing experiments for parameter identifiability in kinetic models of metabolism, the motivation to develop such methods is not well understood.
+Similarly, for identifying $v_3$ that is described by three parameters ($\theta \in \mathbb{R}^3$), we need data from a combination of three different experiments.
+Based on the 21 experiments in Table \ref{tab:pval}, we have 7980 distinct combinations (21 x 20 x 19) of three experiments.
+However, it is important to note that not all of the either 420 combinations (for $v_1$ and $v_2$), or the 7980 combinations (for $v_3$), may be suitable for parameter estimation.
+Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers.
+Moreover, among the chosen smaller number of experimental data combinations, the methodology that we propose enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation.
 
 \printbibliography
+
 \end{document}

### 2018-04-12 ident/docs/draft_2.tex
more figures and info on structurally and practically identifiability

@@
 	We apply the method described in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, and illustrated above for $v_1$, to other fluxes of the network (Figure \ref{fig:network}) not described using Michaelis-Menten equations.
-On applying this method to $v_3$, which is modeled using MWC kinetics, we found that we could not apply our methodology to determine closed form solution for all the parameters in $v_3$.
+On applying this method to $v_3$, which is modeled using MWC kinetics, we find that we could not apply our methodology to determine closed form solution for all the parameters in $v_3$.
 However, when $v_3$ is instead modeled using Convenience kinetics, we find that the parameters are locally identifiable based on the definition we provide in Section \ref{sec:ident}.
 All three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ used to model $v_3$ have two possible roots.
-Although we could not determine the identifiable values for these parameters (Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}), $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value.
+Although we could not determine the identifiable values for these parameters (Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}), $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:parameter_value_v3_var}).
 
+	We also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters.
+For example, in the specific case of $v_3$, we see that globally structurally identifiable parameters $V_3^{max}$ and $K_3^{fdp}$ (when $K_3^{pep}$ is fixed) are practically non-identifiable when data generated from experiments using the MWC kinetics is used to model $v_3$ (Figure \ref{fig:parameter_value_v3_var}a).
+The same is true of globally structurally identifiable parameters $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is fixed (Figure \ref{fig:parameter_value_v3_var}b).
+Instead, when Convenience kinetics is used to generate in silico data, under the aforementioned conditions all the parameters are globally structurally and practically identifiable (Figure \ref{fig:parameter_value_v3_var}c and d).
+
+	Similarly, we applied the demonstrated methodology to flux $v_5$ that describes a transcription/translation reaction using Hill kinetics, based on only the available metabolomics and the fluxomics information.
+When we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system.
+However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$.
+Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}.
+In this instance, the identification required data from two different experiments.
+Again, just as we see for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable.
+On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally structurally identifiable parameter.
+However, only one of the two roots ($K_e^{fdp}(2)$) is practically identifiable based on the physical relevance criteria stated in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b.
+Thus, enforcing this criteria results in global practical identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:parameter_value_proof}c).
+
+	\textcolor{red}{summarize and start next section on identifiability here}
 	(Figure \ref{fig:ident_v3_var}a),  (Figure \ref{fig:ident_v3_var}b).
@@
 
-		Similarly, when we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system.
-However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$.
-Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}.
-Even in this scenario, just as we saw earlier for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable.
-On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally identifiable parameter.
+	In applying this method to $v_5$ we find that the analysis returns two distinct roots for $K_e^{fdp}$, denoted as $K_e^{fdp}(1)$ and $K_e^{fdp}(2)$, thereby making it a locally identifiable parameter.
 
-	In applying this method to $v_5$ we find that analysis returns two distinct roots for $K_e^{fdp}$, denoted as $K_e^{fdp}(1)$ and $K_e^{fdp}(2)$, thereby making it a locally identifiable parameter.
-However, only one of the two roots ($K_e^{fdp}(2)$) is identifiable based on the physical relevance criteria stated in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b.
-Thus, enforcing this criteria results in global identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$.
 	Next, we used data from three different experiments to identify the three parameters modeling flux $v_3$.

### 2018-04-12 ident/docs/sifigures.tex
added new figure

@@
 
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v3_var_1_2_ck_1_2_parameter_value}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$ when the a) and b) Monod-Wyman-Changeaux model for allosteric regulation for $v_3$ is used to generate data, c) and d) experimental data is generated using the Convenience Kinetic rate law model with activation for $v_3$.
+The left column shows the globally identifiable roots of $V_3^{max}$ and $K_3^{fdp}$ when $K_3^{pep}$ is held constant, and the right column shows the globally identifiable root of $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is held constant.}\label{fig:parameter_value_v3_var}
+\end{figure}
+
 \clearpage

### 2018-04-12 ident/docs/draft_2.tex
more figure and results changes

@@
 All three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ used to model $v_3$ have two possible roots.
-Although we could not determine the identifiable values for these parameters (Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}), $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:parameter_value_v3_var}).
+We still could not determine identifiable values for these parameters (Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}).
+In the box plot in Figure \ref{fig:parameter_value_v3}a and b we show all the possible estimated parameter values for both roots of all three parameters modeling $v_3$, when experimental data is generated using the MWC model for $v_3$.
+Due to the numerous estimated parameter values, we can designate these parameters are practically non-identifiable.
+Although using the Convenience kinetics model for $v_3$ to generate the in silico experimental data leads to a reduction in the uncertainty of the estimated parameters (\ref{fig:parameter_value_v3}c and d), we still end up with non-unique parameter estimates.
+This leads is to conclude that under the current conditions, flux $v_3$ is structurally locally identifiable and practically non-identifiable.
 
-	We also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters.
+	However, $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:parameter_value_v3_var}).
+
+	Using Figure \ref{fig:parameter_value_v3_var}, we also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters.
 For example, in the specific case of $v_3$, we see that globally structurally identifiable parameters $V_3^{max}$ and $K_3^{fdp}$ (when $K_3^{pep}$ is fixed) are practically non-identifiable when data generated from experiments using the MWC kinetics is used to model $v_3$ (Figure \ref{fig:parameter_value_v3_var}a).
@@
 However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$.
-Thus, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}.
+Thus, similar to $v_3$ earlier, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}.
 In this instance, the identification required data from two different experiments.
@@
 
-	\textcolor{red}{summarize and start next section on identifiability here}
-	(Figure \ref{fig:ident_v3_var}a),  (Figure \ref{fig:ident_v3_var}b).
+	\textcolor{red}{summarize and start next section on identifiability here}So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism.
+Although in some instances (e.g., $v_5$) local identifiability could be resolved to obtain global identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$).
+Moreover, we also see that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information.
+We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$.
+Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to parameter non-identifiability or local identifiability.
+This can lead to uncertainty in parameter estimates made from in vivo experimental data.
 
-	However, for $v_3$, we could not resolve local identifiability only through the addition of physical relevance constraints on the parameters.
-This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and Figure \ref{fig:ident_root2}b).
-Instead,
-
-	 Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable.
-Thus, the dimension of the parameter space for $v_3$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ for global identifiability.
-This informs us of the correlation between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, which results in structural non-identifiability of parameters in the model.
-
-	In addition to the observation on the degrees of identifiability of different parameter types, we also see that the enzyme kinetic rate law used to model a flux determines the ability to identify the parameters within the given model.
-For instance, as we state in the Methods section,
-
-	 We say locally identifiable due to the fact that
-
-	 In Figure \ref{fig:ident}c, we only show one of the two roots for each parameter in $v_3$.
-The identifiability of parameters based on the second root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}b.
-
-	 that require only data from two different experiments: $v_2$ and $v_5$. In applying this method to $v_5$
-
-	In applying this method to $v_5$ we find that the analysis returns two distinct roots for $K_e^{fdp}$, denoted as $K_e^{fdp}(1)$ and $K_e^{fdp}(2)$, thereby making it a locally identifiable parameter.
-
-	Next, we used data from three different experiments to identify the three parameters modeling flux $v_3$.
-Similar to $v_5$, when the parameters of the Convenience Kinetic rate law modeling $v_3$ are identified, each of the three parameters ($V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$) has two distinct roots that can be written in terms of data from three different experiments.
-Thus, these parameters are locally identifiable based on the definition we provide in section \ref{sec:ident}.
-
-	First, we would like to point out that
-
-	First, in order to show that the methodology predicts exact values of the parameters that are used to generate the in silico experimental data used for identifiability analysis, the predicted values of parameters modeling fluxes $v_1$ ($k_1^{cat}$ and $K_1^{ac}$), $v_2$ ($V_2^{max}$ and $K_2^{pep}$) and $v_5$ ($V_e^{max}$ and $K_e^{fdp}$) are shown in Figure \ref{fig:parameter_value_proof}.
-We found that the use of different enzyme kinetic rate law models (Convenience kinetics and Monod-Wyman-Changeaux kinetics) for $v_3$ had no bearing on the ability to estimate the true parameter values when the parameters are globally identifiable.
-
-	For $v_5$, the identifiability
-
-	 Accordingly, we can estimate unique values for any globally identifiable parameter from any set of experiments that satisfy practical identifiability criteria.
-
-	However, we see from the estimated parameter values for either of the two roots for all three parameters (Figure \ref{fig:parameter_value_v3}a and b) that the total number of distinct values that any of the three parameters can take is less than or equal to the total number of combinations of three experiments from which data is used for identifiability analysis.
-Also of importance is the fact that the experimental data for this analysis is generated using the MWC model for $v_3$.
-So, we hypothesized that the uncertainty seen in the parameter estimates from the identifiability analysis could be attributed to the difference between the kinetic model used to generate the data and the model used for identifiability analysis.
-On using the the Convenience Kinetic model to represent $v_3$ for both data generation as well as identifiability analysis, we see that although the uncertainty in the estimated parameter values is reduced (Figure \ref{fig:parameter_value_v3}c and d), the estimated values are not unique and uncertainty persists.
-Hence, we conclude that under the current scenario these parameters are structurally locally identifiable and practically non-identifiable.
-
-	We summarize the results on the identifiability of parameters in these and other fluxes in the small network in Figure \ref{fig:ident}, and describe them in the following section.
+	In the following section, we describe and summarize the results on the identifiability of parameters in these fluxes ($v_1$, $v_2$, $v_3$ and $v_5$) from the network shown in Figure \ref{fig:network}.
 
 	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}	
-	In Figure \ref{fig:ident}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b), $v_3$ (Figure \ref{fig:ident}c) and $v_5$ (Figure \ref{fig:ident}d) belonging to the network shown in Figure \ref{fig:network}.
-Based on the definition given in section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
+	In Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a and e), $v_2$ (Figure \ref{fig:ident}b and f), $v_3$ (Figure \ref{fig:ident}c and g) and $v_5$ (Figure \ref{fig:ident}d and h) belonging to the network shown in Figure \ref{fig:network}.
+Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
 We see from Figure \ref{fig:ident} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$) for each of the four fluxes is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law model.
 We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience Kinetic rate law, and $v_5$ is modeled as a Hill equation with inhibition.
-We accordingly summarize that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
+We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
 
@@
 The two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, and $K_3^{pep}(1)$ refer to the first set of solutions, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$ refer to the other set of solutions.
+
+	(Figure \ref{fig:ident_v3_var}a),  (Figure \ref{fig:ident_v3_var}b).
+
+	However, for $v_3$, we could not resolve local identifiability only through the addition of physical relevance constraints on the parameters.
+This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and Figure \ref{fig:ident_root2}b).
+Instead,
+
+	 Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable.
+Thus, the dimension of the parameter space for $v_3$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ for global identifiability.
+This informs us of the correlation between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, which results in structural non-identifiability of parameters in the model.
+
+	In addition to the observation on the degrees of identifiability of different parameter types, we also see that the enzyme kinetic rate law used to model a flux determines the ability to identify the parameters within the given model.
+For instance, as we state in the Methods section,
+
+	 We say locally identifiable due to the fact that
+
+	 In Figure \ref{fig:ident}c, we only show one of the two roots for each parameter in $v_3$.
+The identifiability of parameters based on the second root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}b.
+
+	 Accordingly, we can estimate unique values for any globally identifiable parameter from any set of experiments that satisfy practical identifiability criteria.
 
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure1/v1_v2_v3_v5_ident}}
-		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$ when there is no noise in the input experimental data.
-The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
-The total number of experiments used to identify parameters in $v_1$, $v_2$ and $v_5$ is two.
-The total number of data combinations used to identify parameters in fluxes $v_1$ and $v_2$ is 240, 421 different combinations are used to test identifiability for $v_5$, and 4896 combinations are used to test identifiability of parameters for $v_3$.
-Note that $v_3$ is modeled with three parameters and requires data from at least three different experiments, and also has two possible roots for each of its parameters.
-Results for only one of the two roots indicated by (1) is shown here.}\label{fig:ident}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.7\textheight,keepaspectratio]{figures/figure1/v1_V1max_v2_v3_1_v5_mwc_ck_ident}}
+		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) and e) $v_1$, b)  and f) $v_2$, c) and g) $v_3$, and d) and h) $v_5$ when there is no noise in the input experimental data.
+The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, 3360 for $v_3$ and 421 for $v_5$) that can identify each parameter is also specified.
+$v_1$, $v_2$ and $v_5$ require data from two experiments for analysis, and $v_3$ requires data from three experiments.
+Results for only one of the two possible roots are shown for $v_3$ and $K_e^{fdp}$ in $v_5$.
+The results obtained using data from the MWC model for $v_3$ are presented in the panels on the left hand side column and the panels on the right hand column show results obtained using data derived from the Convenience kinetics model for $v_3$.}\label{fig:ident}
 	\end{figure}
@@
 	\section{Potential Supplementary Information Figures}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight, keepaspectratio]{figures/figure1/v1_kcat_v3_2_v5_1_ident}}
-		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_3$ and c) $v_5$ \textcolor{red}{(need to change figure)} when there is no noise in the input experimental data.
-The percentage of total experimental data combinations that can identify each parameter is specified either in, or right next to the bar showing the total number of combinations identifying a given parameter.
-The total number of experiments used to identify parameters in $v_1$ and $v_5$ is two.
-The total number of data combinations used to identify parameters in fluxes $v_1$ is 240.
-421 different combinations are used to test identifiability for $v_5$, and 4896 combinations are used to test identifiability of parameters for $v_3$.
-Note that $v_3$ is modeled with three parameters and requires data from at least three different experiments, and also has two possible roots for each of its parameters.
-Results for only one of the two roots indicated by (2) is shown here.
-Similarly, results for only one of the two roots of $K_e^{fdp}$, indicated as $K_e^{fdp}(1)$ is shown.}\label{fig:ident_root2}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_1_2_ident}}
-		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in the convenience kinetic rate law model of flux $v_3$ when a) $K_3^{pep}$ is held constant, and b) $K_3^{fdp}$ is held constant.
-Note that the two identifiable parameters in both panel a and panel b are globally identifiable.
-The degree of identifiability of $V_3^{max}$ and $K_3^{pep}$ are higher when $K_3^{fdp}$ is held constant, than the degree of identifiability of $V_3^{max}$ and $K_3^{fdp}$ when $K_3^{pep}$ is held constant.}\label{fig:ident_v3_var}
-	\end{figure}
 
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_v2_v3_1_v5_2_parameter_value}}
-		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$.
-The parameter values for the only the first root of parameters in $v_3$ ($V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$) and the second root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(2)$) are shown.
-The variability in the estimated values of $V_1^{max}$ can be explained by the fact that experimental data for identification is generated using the enzyme concentrations as a variable for $v_1$, while the identification of parameters is done by subsuming the enzyme concentration into the enzyme turn over parameter ($k_1^{cat}$) to obtain $V_1^{max}$, and identifying $V_1^{max}$.
-So, the variability in values of $V_1^{max}$ reflects the variability in the metabolite concentrations (\textit{pep}, \textit{fdp}, \textit{E}) in the experimental data.
-Similarly, data is generated using the Monod-Wyman-Changeaux model for allosteric regulation for $v_3$, and a Convenience Kinetic rate law model is used for model identification purposes.}\label{fig:parameter_value_v1_vmax}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_k1cat_v3_2_v5_1_parameter_values}}
-		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_3$, and c) $v_5$.
-For $v_1$, we have assumed that enzyme concentration is available and have accordingly identified and estimated $k_1^{cat}$, as opposed to $V_1^{max}$.
-The parameter values for the only the second root of parameters in $v_3$ ($V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$) and the first root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(1)$) are shown.
-Note that $K_e^{fdp}(1)$ is not estimated by any combination of two experiments and $V_e^{max}$ is estimated by all combinations.
-Data is generated using the Monod-Wyman-Changeaux model for allosteric regulation for $v_3$, and a Convenience Kinetic rate law model is used for model identification purposes.}\label{fig:parameter_value_v1_kcat}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
 		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_k1cat_parameter_value}}

### 2018-04-12 ident/docs/draft_2.tex
add figure references

@@
 Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
-We see from Figure \ref{fig:ident} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$) for each of the four fluxes is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law model.
+We see from Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$), whenever they are identifiable for each of the four fluxes, is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law models.
 We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience Kinetic rate law, and $v_5$ is modeled as a Hill equation with inhibition.
@@
 
-	In Figure \ref{fig:ident}a less than 50\% of all data combinations can identify $v_1$.
-Whereas, all available data combinations satisfy the requirements for identifying $v_2$ (Figure \ref{fig:ident}b).
+	In Figures \ref{fig:ident}a and \ref{fig:ident}e less than 50\% of all data combinations can identify $v_1$.
+Whereas, all available data combinations satisfy the requirements for identifying $v_2$ (Figures \ref{fig:ident}b and f).
+This observation is true when data from either the MWC model (Figures \ref{fig:ident}a and b) or the Convenience kinetics model for $v_3$ (Figures \ref{fig:ident}e and \ref{fig:ident}f) is used.
 This difference can be alluded to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter.

### 2018-04-12 ident/python2/ss-ident/v1_Vmax_ck_script.py
changes to experiments included for plotting experimental data distribution

@@ -58,7 +58,7 @@ print('Practical Identifiability Analysis of v1 with 2 parameters: V1max and K1a
 ident_fun_choice = [0]
 # perform identifiability when v1 is written with V1max in the numerator
 ident_details_v1_V1max = flux_ident_2_data_combination(experimental_datasets_2_expts, choose=combination_choice,
-                                                       flux_ids=[1], flux_choice=[3], ident_fun_choice=ident_fun_choice)
+                                                       flux_ids=[1], flux_choice=[1], ident_fun_choice=ident_fun_choice)
 print('Identifiability analysis of v1 with 2 parameters (V1max and K1ac) complete.\n')
 
 # data processing - do not combine fluxes
@@ -80,7 +80,9 @@ parameter_experiment_info_spider(experiment_info_v1_V1max)
 plot_parameter_values(true_value_v1_V1max)
 
 # plot distribution of experimental data (concentration and fluxes)
-plot_experiment_data_dist(exp_xss, exp_fss, experiment_choice=[0, 1, 2, 3, 4, 5,
+plot_experiment_data_dist(exp_xss, exp_fss, experiment_choice=[0,
+                                                               1, 2, 3, 4, 5,
+                                                               6, 7, 8, 9, 10,
                                                                11, 12, 13, 14, 15,
                                                                16, 17, 18, 19, 20])
 

### 2018-04-12 ident/python2/ss-ident/v1_Vmax_script.py

@@ -58,7 +58,7 @@ print('Practical Identifiability Analysis of v1 with 2 parameters: k1cat and K1a
 ident_fun_choice = [0]
 # perform identifiability when v1 is written with k1cat*E in the numerator
 ident_details_v1_k1cat = flux_ident_2_data_combination(experimental_datasets_2_expts, choose=combination_choice,
-                                                       flux_ids=[1], flux_choice=[3], ident_fun_choice=ident_fun_choice)
+                                                       flux_ids=[1], flux_choice=[1], ident_fun_choice=ident_fun_choice)
 print('Identifiability analysis of v1 with 2 parameters (k1cat and K1ac) complete.\n')
 
 # data processing - do not combine fluxes
@@ -81,7 +81,9 @@ parameter_experiment_info_spider(experiment_info_v1_k1cat)
 plot_parameter_values(true_value_v1_k1cat)
 
 # plot distribution of experimental data (concentration and fluxes)
-plot_experiment_data_dist(exp_xss, exp_fss, experiment_choice=[0, 1, 2, 3, 4, 5,
+plot_experiment_data_dist(exp_xss, exp_fss, experiment_choice=[0,
+                                                               1, 2, 3, 4, 5,
+                                                               6, 7, 8, 9, 10,
                                                                11, 12, 13, 14, 15,
                                                                16, 17, 18, 19, 20])
 

### 2018-04-12 ident/python2/ss-ident/v5_ck_script.py
new script for v5 ident when CK is used for v3

@@ -1,0 +1,86 @@
+import numpy as np
+from generate_expdata import generate_expdata
+from simulate_data import arrange_experimental_data
+from kotte_model import flux_ident_2_data_combination
+from process_ident_data import process_info_sample
+from plot_ident_results import data_utility_plot
+from plot_ident_results import plot_parameter_values
+from plot_ident_results import parameter_identifibaility_plot
+# from plot_ident_results import parameter_experiment_info_plot
+from plot_ident_results import parameter_experiment_info_spider
+
+
+# generate noisy experimental data for testing identifiability
+y0 = np.array([5, 1, 1])
+# default parameter values
+cvode_options = ('Newton', 'Adams', 1e-10, 1e-10, 200)
+ode_parameter_values = {"K1ac": np.array([.1]),
+                        "K3fdp": np.array([.1]),
+                        "L3fdp": np.array([4e6]),
+                        "K3pep": np.array([.1]),
+                        "K2pep": np.array([.3]),
+                        "vemax": np.array([1.1]),
+                        "Kefdp": np.array([.45]),
+                        "ne": np.array([2]),
+                        "d": np.array([.25]),
+                        "V4max": np.array([.2]),
+                        "k1cat": np.array([1]),
+                        "V3max": np.array([1]),
+                        "V2max": np.array([1]),
+                        "ac": np.array([.1])}
+
+# get experimental system steady state data without noise using MWC kinetics for v3 (kinetics = 1)
+exp_xss, exp_fss, exp_ssid, perturbation_details = \
+    generate_expdata(y0, cvode_options, ode_parameter_values, noise=0, kinetics=2, dynamic_plot=0,
+                     perturbation_plot=0)
+
+# arrange experimental data to form multiple data sets
+exp_flux_index = np.array([0, 3, 2, 4, 1, 5])
+
+# get combination of 2 experiments and perform identifiability on all fluxes that require 2 data sets
+# get combinations of experimental datasets
+experimental_datasets_2_expts, \
+    experiment_choice, combination_choice = arrange_experimental_data(exp_xss, exp_fss, perturbation_details,
+                                                                      experiments_per_set=2, flux_id=exp_flux_index,
+                                                                      experiment_choice=[0, 1, 2, 3, 4, 5,
+                                                                                         6, 7, 8, 9, 10,
+                                                                                         11, 12, 13, 14, 15,
[…]

### 2018-04-12 ident/docs/sifigures.tex

@@
 
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight, keepaspectratio]{figures/figure1/v1_k1cat_v3_2_v5_1_ident}}
+	\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) and d) $v_1$, b) and e) $v_3$, and c) and f) $v_5$ when there is no noise in the input experimental data.
+The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, 3360 for $v_3$ and 421 for $v_5$) that can identify each parameter is also specified.
+$v_1$, $v_2$ and $v_5$ require data from two experiments for analysis, and $v_3$ requires data from three experiments.
+Results for only one of the two possible roots are shown for $v_3$ and $K_e^{fdp}$ in $v_5$.
+The results obtained using data from the MWC model for $v_3$ are presented in the panels on the left hand side column and the panels on the right hand column show results obtained using data derived from the Convenience kinetics model for $v_3$.}\label{fig:ident_root2}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_1_2_mwc_ck_ident}}
+	\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in the convenience kinetic rate law model of flux $v_3$ when a) and c) $K_3^{pep}$ is held constant, and b) and d) $K_3^{fdp}$ is held constant.
+Results obtained using data from the MWC model for $v_3$ are presented in the panels on the left column and the panels on the right column show results obtained using data derived from the Convenience kinetics model for $v_3$.
+Note that all the identifiable parameters in the four panels are globally identifiable.
+The degree of identifiability of $V_3^{max}$ and $K_3^{pep}$ are higher when $K_3^{fdp}$ is held constant, than the degree of identifiability of $V_3^{max}$ and $K_3^{fdp}$ when $K_3^{pep}$ is held constant.}\label{fig:ident_v3_var}
+\end{figure}
+
 \clearpage
@@
 
-The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (section \ref{sec:ident}).
+The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (Section \ref{sec:ident}).
 We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica or SymPy in Python).

### 2018-04-13 ident/docs/draft_2.tex
changes to section 2 of results

@@
 We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience Kinetic rate law, and $v_5$ is modeled as a Hill equation with inhibition.
+This observation is also true irrespective of the kinetic rate law used to model $v_3$ (MWC vs Convenience kinetics) to generate in silico experimental data.
 We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
 
-	In Figures \ref{fig:ident}a and \ref{fig:ident}e less than 50\% of all data combinations can identify $v_1$.
-Whereas, all available data combinations satisfy the requirements for identifying $v_2$ (Figures \ref{fig:ident}b and f).
-This observation is true when data from either the MWC model (Figures \ref{fig:ident}a and b) or the Convenience kinetics model for $v_3$ (Figures \ref{fig:ident}e and \ref{fig:ident}f) is used.
-This difference can be alluded to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter.
-For example, we showed in Equation (\ref{eq:v1_par}) that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
+	When data based on the MWC model for $v_3$ is used for identifying parameters, less that 50\% of all data combinations can identify $v_1$ (Figure \ref{fig:ident}a).
+Even though the degree of identifiability of $K_1^{ac}(ne)$ is significantly different when data derived from the Convenience kinetics model for $v_3$ is used for identification (Figures \ref{fig:ident}e), the degree of identifiability is still less that 50\% for both parameters.
+In contrast, all available data combinations (from the MWC model and the Convenience kinetics model) satisfy the requirements for identifying $v_2$ (Figures \ref{fig:ident}b and f).
+
+	The difference in the degrees of identifiability between $v_1$ (Figure \ref{fig:ident}a) and $v_2$ (Figure \ref{fig:ident}b) can be attributed to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter.
+To elaborate, we showed in Equation (\ref{eq:v1_par}) that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
 So, in this instance, the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$.
@@
 
-	In addition to the observation on the degrees of identifiability of different parameter types, we also see that the enzyme kinetic rate law used to model a flux determines the ability to identify the parameters within the given model.
-For instance, as we state in the Methods section, we could not apply our methodology to determine closed form solution for all the parameters when $v_3$ is modeled using MWC kinetics.
+	The difference in the degree of identifiability of $v_1$ that comes about from the use of experimental data from different rate law models used to model $v_3$ (Figures \ref{fig:ident}a and e) can be again attributed to the usefulness of experimental data.
+However, the difference in data usefulness between similar experiments due to differences in the kinetic rate law used for $v_3$ could be due to the difference in the dynamics of the network represented by either of the two models.
+The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets and its subsequent effect of parameter identifiability.
+Nonetheless, we would like to note that these differences in the dynamics contributed by the difference between the rate laws can be resolved with more information i.e., experimental proteomic data that provide protein concentrations for different experiments.
+We earlier saw that relaxing the assumption on the available experiment data, i.e., assuming the availability of protein concentration (\textit{E}), in addition to the metabolite concentrations (\textit{pep} and \textit{fdp}) and fluxes, results in removal of uncertainty and global identifiability of $v_1$ (Supplementary Figure \ref{fig:parameter_value_proof}).
+Additionally, we also see an improvement in the degree of identifiability of the corresponding parameters $k_1^{cat}$ and $K_1^{ac}$ (Supplementary Figure \ref{fig:ident_root2}) when enzyme concentrations are available.
+
+	We would also like to point out that a difference in either the ability to estimate true parameter values, or the the identifiability of parameter modeling $v_2$ does not arise due to difference in the kinetic models used to obtain in silico experimental data.
+This could because of the fact that any and all available information that is required for the identification of parameters in $v_2$ is encompassed in the steady state data used for identifiability analysis.
+
 	 In contrast, when $v_3$ is modeled using the Convenience Kinetic rate law, we find that the parameters are locally identifiable based on the definition we provide in section \ref{sec:ident}.

### 2018-04-13 ident/docs/draft_2.tex
completed section 3.2

@@
 
-	\textcolor{red}{summarize and start next section on identifiability here}So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism.
+	So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism.
 Although in some instances (e.g., $v_5$) local identifiability could be resolved to obtain global identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$).
@@
 
-	In the following section, we describe and summarize the results on the identifiability of parameters in these fluxes ($v_1$, $v_2$, $v_3$ and $v_5$) from the network shown in Figure \ref{fig:network}.
+	In the following section, we describe and summarize the results on the identifiability of parameters in these fluxes ($v_1$, $v_2$, $v_3$ and $v_5$) based on the expressions and the identifiable roots we discussed in this section.
 
@@
 We see from Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$), whenever they are identifiable for each of the four fluxes, is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law models.
-We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience Kinetic rate law, and $v_5$ is modeled as a Hill equation with inhibition.
-This observation is also true irrespective of the kinetic rate law used to model $v_3$ (MWC vs Convenience kinetics) to generate in silico experimental data.
+We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition.
+This observation is also true irrespective of the kinetic rate law used to model $v_3$ (MWC in Figure \ref{fig:ident}a-d vs Convenience kinetics in Figure \ref{fig:ident}e-h) to generate in silico experimental data.
 We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
 
-	When data based on the MWC model for $v_3$ is used for identifying parameters, less that 50\% of all data combinations can identify $v_1$ (Figure \ref{fig:ident}a).
-Even though the degree of identifiability of $K_1^{ac}(ne)$ is significantly different when data derived from the Convenience kinetics model for $v_3$ is used for identification (Figures \ref{fig:ident}e), the degree of identifiability is still less that 50\% for both parameters.
-In contrast, all available data combinations (from the MWC model and the Convenience kinetics model) satisfy the requirements for identifying $v_2$ (Figures \ref{fig:ident}b and f).
+	When data based on the MWC model for $v_3$ is used for identifying parameters, less that 50\% of all data combinations can identify $v_1$ (47.5\% and 40.0\% in Figure \ref{fig:ident}a).
+Even though the degree of identifiability of $K_1^{ac}(ne)$ is significantly different when data derived from the Convenience kinetics model for $v_3$ is used for identification (25\% in Figures \ref{fig:ident}e), the degree of identifiability is still less that 50\% for both parameters.
+In contrast, all available data combinations (from the MWC model and the Convenience kinetics model) satisfy the requirements for identifying $v_2$ (100\% in Figures \ref{fig:ident}b and f).
 
 	The difference in the degrees of identifiability between $v_1$ (Figure \ref{fig:ident}a) and $v_2$ (Figure \ref{fig:ident}b) can be attributed to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter.
-To elaborate, we showed in Equation (\ref{eq:v1_par}) that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
-So, in this instance, the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$.
+To elaborate, we showed in Equation (\ref{eq:v1_par}) and Section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
+So, in this instance, recall that the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$.
 Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}.
+We elaborate on the type and nature of experiments that are useful and informative for parameter identification in $v_1$ and $v_2$ later in Section x.
 
 	The difference in the degree of identifiability of $v_1$ that comes about from the use of experimental data from different rate law models used to model $v_3$ (Figures \ref{fig:ident}a and e) can be again attributed to the usefulness of experimental data.
-However, the difference in data usefulness between similar experiments due to differences in the kinetic rate law used for $v_3$ could be due to the difference in the dynamics of the network represented by either of the two models.
-The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets and its subsequent effect of parameter identifiability.
-Nonetheless, we would like to note that these differences in the dynamics contributed by the difference between the rate laws can be resolved with more information i.e., experimental proteomic data that provide protein concentrations for different experiments.
+However, the difference in data usefulness between similar experiments arising due to differences in the kinetic rate law used for $v_3$ could be due to the difference in the dynamics of the network represented by either of the two models.
+The interaction between different components of the network (Figure \ref{fig:network}) as exemplified by the different models plays an important role in determining the dynamic and the steady state response of the network.
+The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets, and its subsequent effect of parameter identifiability.
+
+	Nonetheless, we would like to note that these variations in the dynamics contributed by the difference between the rate laws can be resolved with more information i.e., experimental proteomic data that provide protein concentrations for different experiments.
 We earlier saw that relaxing the assumption on the available experiment data, i.e., assuming the availability of protein concentration (\textit{E}), in addition to the metabolite concentrations (\textit{pep} and \textit{fdp}) and fluxes, results in removal of uncertainty and global identifiability of $v_1$ (Supplementary Figure \ref{fig:parameter_value_proof}).
-Additionally, we also see an improvement in the degree of identifiability of the corresponding parameters $k_1^{cat}$ and $K_1^{ac}$ (Supplementary Figure \ref{fig:ident_root2}) when enzyme concentrations are available.
+Recall that any predicted parameter value for both $k_1^{cat}$ and $V_1^{max}$ matches with the exact known value irrespective of the experiment (MWC vs Convenience kinetics) from which data is obtained when enzyme concentrations are included for identifiability (Supplementary Figure \ref{fig:parameter_value_proof}a).
+Additionally, we also see an improvement in the degree of identifiability of the corresponding parameters $k_1^{cat}$ and $K_1^{ac}$ (Supplementary Figure \ref{fig:ident_root2}a and b) when enzyme concentrations are available.
 
@@
 
-	 In contrast, when $v_3$ is modeled using the Convenience Kinetic rate law, we find that the parameters are locally identifiable based on the definition we provide in section \ref{sec:ident}.
-We say locally identifiable due to the fact that all three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two possible roots.
-In Figure \ref{fig:ident}c, we only show one of the two roots for each parameter in $v_3$.
-The identifiability of parameters based on the second root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}b.
+	However, unlike $v_2$, we see that the degree of identifiability of both locally identifiable roots of parameters in $v_3$ changes with the change in the data (MWC vs Convenience kinetics) used for the analysis (Figure \ref{fig:ident}c and g).
+Despite the difference in the dynamics represented by the different data sets, and its impact on the usefulness of data sets for identifiability purposes that was explained earlier, we could also allude to the structural non-identifiability between the parameters in $v_3$ for the impact that it would have on the practical identifiability of the parameters using in silico data from different sources.
+In order to ascertain this, we plot the degrees of identifiability of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ under conditions when they are globally identifiable (Section \ref{sec:example}: $V_3^{max}$ and $K_3^{fdp}$ are globally identifiable when $K_3^{pep}$ is constant, and $V_3^{max}$, $K_3^{pep}$are globally identifiable when $K_3^{fdp}$ is known and fixed) in Supplementary Figure \ref{fig:ident_v3_var}.
+Now that structurally non-identifiability between parameters has been alleviated, we see that the uncertainty introduced by the data obtained from the MWC model results in only a fraction of data sets being useful to estimate parameters (Supplementary Figure \ref{fig:ident_v3_var}a and b).
+This ties back with our earlier observation that these parameters are not practically identifiable from data sets obtained using the MWC model (Supplementary Figure \ref{fig:parameter_value_v3_var}a and b).
+However, we see from Supplementary Figures \ref{fig:ident_v3_var}c and d that the globally structurally identifiable parameters are also practically identifiable based on the data obtained from the Convenience kinetics model.
+This can be corroborated by the parameter values shown in Supplementary Figure \ref{fig:parameter_value_v3_var}c and d, where data from the Convenience kinetics model is used for analysis.
 
-	The identifiability of the second root for $K_e^{fdp}$, indicated as $K_e^{fdp}(2)$, is shown in Figure \ref{fig:ident}d.
-The identifiability of the first root of $K_e^{fdp}$, $K_e^{fdp}(1)$, is shown in Figure \ref{fig:ident_root2}c.
-
-	While the identifiability of the second roots $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$ for $v_3$ (Figure \ref{fig:ident_root2}b) are similar to that of the first root (Figure \ref{fig:ident}c), this is not the case for $v_5$.
-We would like to point out that this is one of the scenarios where the local identifiability of a parameter like $K_e^{fdp}$ can be expanded to global identifiability by adding physical relevance of parameter values as an additional constraint for practical identifiability ($D_k>0$, see Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident}).
-We found that the first root for $K_e^{fdp}$, $K_e^{fdp}(1)$, never possessed a physically relevant value, and accordingly was never identifiable based on the criteria set out in Figure \ref{fig:ident-flowchart}b and section \ref{sec:ident} (Figure \ref{fig:ident_root2}c).
-In contrast, the second root, $K_e^{fdp}(2)$ has an identifiability of 50.0\% (Figure \ref{fig:ident}d).
-Thus, we can get globally identifiable parameters through the addition of constraints related to their physical relevance to locally identifiable parameters.
-
-	However, for $v_3$, we could not resolve local identifiability only through the addition of physical relevance constraints on the parameters.
-This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and Figure \ref{fig:ident_root2}b).
-Instead, $K_3^{fdp}$ becomes globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:ident_v3_var}a), and $K_3^{pep}$ becomes globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:ident_v3_var}b).
-Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable.
-Thus, the dimension of the parameter space for $v_3$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ for global identifiability.
-This informs us of the correlation between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, which results in structural non-identifiability of parameters in the model.
-
-	Although one could argue that a similar approach could be used to identify parameters of $v_3$ written using the MWC model, we found that it was not possible to obtain identifiability expressions for any of the parameters in the MWC model, except for $V_3^{max}$ when all other parameters are assumed to be known.
+	We did not observe any differences in the degree of identifiability of parameters in $v_5$ between the analyses that used data from either the MWC model or the Convenience kinetics model for $v_3$ (Figure \ref{fig:ident}d and h).
+The identifiability of parameters based on the first root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}c and f.
 
@@
 The two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, and $K_3^{pep}(1)$ refer to the first set of solutions, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$ refer to the other set of solutions.
-
-	(Figure \ref{fig:ident_v3_var}a),  (Figure \ref{fig:ident_v3_var}b).
-
-	However, for $v_3$, we could not resolve local identifiability only through the addition of physical relevance constraints on the parameters.
-This is seen from the similarity in the degree of identifiability for both the roots (Figure \ref{fig:ident}c and Figure \ref{fig:ident_root2}b).
-Instead,
-
-	 Fixing one of the two parameters as a known quantity results in unique expressions for both $V_3^{max}$ and the other parameter, rendering them both globally identifiable.
-Thus, the dimension of the parameter space for $v_3$ has to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ for global identifiability.
-This informs us of the correlation between $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience Kinetic rate law for allosterically activated flux $v_3$, which results in structural non-identifiability of parameters in the model.
-
-	In addition to the observation on the degrees of identifiability of different parameter types, we also see that the enzyme kinetic rate law used to model a flux determines the ability to identify the parameters within the given model.
-For instance, as we state in the Methods section,
-
-	 We say locally identifiable due to the fact that
-
-	 In Figure \ref{fig:ident}c, we only show one of the two roots for each parameter in $v_3$.
-The identifiability of parameters based on the second root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}b.
-
-	 Accordingly, we can estimate unique values for any globally identifiable parameter from any set of experiments that satisfy practical identifiability criteria.
 
@@
 
-	For flux $v_1$ (Figure \ref{fig:data_utility}a), data from about 52\% of the 240 combinations of experiments cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$).
-Only $V_1^{max}$ can be identified by about 7\% of the combinations (Figure \ref{fig:ident}a), while only about 40\% of the combinations can identify both the parameters used in the model.
-	In contrast, all the available data can identify both parameters in $v_2$ (Figure \ref{fig:data_utility}b).
-This is in agreement with the degree of identifiability of parameters used to model $v_2$ (Figure \ref{fig:ident}b).
-In the case of $v_3$ (Figure \ref{fig:data_utility}c), close to 60\% of the \textcolor{red}{4896 combinations} can identify only one parameter, while less than 10\% of the combinations can identify all three parameters in the model.
-Note that only 5\% of combinations cannot identify any parameter modeling $v_3$.
+	For $v_1$ (Figure \ref{fig:data_utility}a), data from about 50\% of the 240 combinations of experiments cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$).
+In contrast, data from only about 5\% of the experiments is useless towards identifying any parameter in $v_3$, while there are no useless combinations of experiments from which data cannot be used to identify either $v_2$, or $v_5$.
+These data are also in agreement with the degree of identifiability observed for parameters in all these fluxes (Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}).
+Thus, when numerous experiments are possible to obtain steady state data to estimate parameters for a small metabolic network, there is a very high possibility of performing experiments that could provide non-informative data that can result in parameter non-identifiability without a priori experimental design.
 
-	Thus, Figure \ref{fig:data_utility} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
-For instance, choosing any of the experiment combinations from the 52\% percent that does not identify any parameter in $v_1$, or the 6\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative.
-This can lead to potential waste of resources used to perform these experiments.
+	We see that in instances where the identifiability of the fluxes differ based on the data set used (data from MWC model vs data from Convenience kinetics model), the utility of the data sets towards identifying parameters are changes.
+For instance, when data from the MWC model is used to identify parameters in $v_3$ (Figure \ref{fig:data_utility}c), less than 10\% of the data sets can identify all three parameters.
+Whereas, when data from the Convenience kinetics model is used for the same purpose, slightly more than 50\% of the combinations can identify all three parameters.
+Note that this observations should be considered in light of the structural local identifiability and the practical non-identifiability that we presented earlier in Sections \ref{sec:example} and \ref{sec:ident_rate_laws}.
 
-	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:data_utility} also informs the ease of identifying parameters for a single flux.
-For the small network (Figure \ref{fig:network}), we find that all chosen combination of experiments are actually capable of identifying $V_2^{max}$ and $K_2^{pep}$ (Figure \ref{fig:data_utility}b).
-In contrast, only about 40\% of the combinations can identify both parameters in $v_1$.
+	Although one could argue based on earlier observations that uncertainty in the model from which data is obtained, and the structural non-identifiability of some of the parameters in the models of the aforementioned fluxes that there may not be a need for experimental design if in vivo model uncertainty and structural non-identifiability can be resolved, one needs to only look at Supplementary Figure \ref{fig:other_data_utility}a and d.
+In these figures we show the utility of multiple combinations of metabolite and protein concentration, and flux data towards identifying $k_1^{cat}$ and $K_1^{ac}$ in $v_1$ when data from the MWC model and Convenience kinetics model are used, respectively.
+We see that about 45\% of the 240 combinations of data from 21 different experiments cannot identify either one of the two parameters.
+
+	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:data_utility} and Supplementary Figure \ref{fig:other_data_utility} also informs the ease of identifying parameters for a single flux.
+In conjunction with the degrees of identifiability shown in Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we see that identifying $v_2$ is much easier than identifying $v_1$.
 Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ and $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$.
@@
 
-	We believe that this information (Figure \ref{fig:data_utility}), in conjunction with the identifiability of each parameter presented in Figure \ref{fig:ident} will give an idea on the hierarchy of identifiability of each parameter, and the need for experimental design to identify each parameter.
-Additionally, having the ability to see the contribution of each type of experiment towards identifiability (Figure \ref{fig:exp_info}) would prove sufficient to decide on the total number and type of experiments that need to be performed to identify all fluxes in a kinetic model of metabolism.
+	So, if we can establish a hierarchy of identifiable parameters and the corresponding fluxes they model, this section gives credence to the idea that a priori experimental design is necessary, even when all possible in vivo measurements can be obtained, to get the most informative data for parameter identification and estimation at minimum cost.
+Additionally, having the ability to see the contribution of each type of experiment towards identifiability would prove sufficient to decide on the total number and type of experiments that need to be performed to identify all fluxes in a kinetic model of metabolism.
 So, we look at the contribution of different types of experiments towards identifying parameters in each flux in the small network in the following section.
 
-	%Although the calculation of the utility of experiment combinations in terms of their ability to identify the maximum number of parameters gives an idea on the informativeness of experiments used to gather the data, further analysis from the perspective of identifiability of the parameters of different fluxes is required to be able to design experiments based on this information.
-Hence, we look at the degree of identifiability (see section \ref{sec:degree_of_identifiability}) of each parameter of each flux in the following section to be able to classify parameters, and consequently fluxes, from most identifiable to least identifiable.
-
 	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}		
-	From Figure \ref{fig:ident} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.
+	From Figures \ref{fig:ident} and \ref{sec:data_utility} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.
 For example, the maximum reactions rates of $v_2$ and $v_5$ have a degree of identifiability of 100\% (Figure \ref{fig:ident}b and d), while their counterparts $V_1^{max}$ and $V_3^{max}$ in $v_1$ and $v_3$ respectively (Figure \ref{fig:ident}a and c), are only identifiable to a lesser extent.
@@
 Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figure \ref{fig:ident}d), the choice and number of experiments required to identify $v_5$ completely hinges upon identifiability of $K_e^{fdp}$ from the chosen experiments.
+
+	%Thus, Figure \ref{fig:data_utility} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
+For instance, choosing any of the experiment combinations from the 52\% percent that does not identify any parameter in $v_1$, or the 6\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative.
+This can lead to potential waste of resources used to perform these experiments.
 
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_data_utility}}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_mwc_ck_data_utility}}
 		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
@@
 The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, \textcolor{red}{3360 combinations} are used to practically identify c) $v_3$, and 420 combinations are used to analyze the identifiability of d) $v_5$.
+A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, 3360 combinations are used to practically identify c) $v_3$, and 420 combinations are used to analyze the identifiability of d) $v_5$.
 Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
@@
 However, results for the only practically identifiable root, based on the physical relevance constraints used in the analysis, is shown.}\label{fig:v5_exp_info}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_v5_1_data_utility}}
-		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$ and c) $v_5$.
-The total number of combinations of experiments is shown in the vertical axis, and the horizontal axis shows the maximum number of parameters that be identified using data from a combination of a) two, b) three and c) two experiments.
-The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 240 data combinations are used for identifiability analysis for a) $v_1$, \textcolor{red}{3360 combinations} are used to practically identify b) $v_3$, and 420 combinations are used to analyze the identifiability of c) $v_5$.
-Information processed based on the identifiability of a) $v_1$ in the presence of enzyme concentration data, b) the second root of parameters in $v_3$ and c) the first root of $K_e^{fdp}$ in $v_5$ are shown.}\label{fig:other_data_utility}
 	\end{figure}

### 2018-04-13 ident/docs/sifigures.tex
new figures on data utility

@@
 
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight, keepaspectratio]{figures/figure4/v1_k1cat_v3_2_v5_1_mwc_ck_data_utility}}
+	\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_3$ and c) $v_5$.
+The total number of combinations of experiments is shown in the vertical axis, and the horizontal axis shows the maximum number of parameters that be identified using data from a combination of a) two, b) three and c) two experiments.
+The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
+A total of 240 data combinations are used for identifiability analysis for a) $v_1$, \textcolor{red}{3360 combinations} are used to practically identify b) $v_3$, and 420 combinations are used to analyze the identifiability of c) $v_5$.
+Information processed based on the identifiability of a) $v_1$ in the presence of enzyme concentration data, b) the second root of parameters in $v_3$ and c) the first root of $K_e^{fdp}$ in $v_5$ are shown.}\label{fig:other_data_utility}
+\end{figure}
+
 \clearpage

### 2018-04-14 ident/docs/draft_2.tex
complete results except noise section

@@
 	\subsection{Identifying parameters in kinetic models of metabolism: an example}\label{sec:example}	
-	In this section, we illustrate the proposed methodology to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model} of the Supplementary Information).
+	In this section, we illustrate the proposed methodology step by step to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model} of the Supplementary Information).
 We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration.
@@
 In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$.
-The globally identifiable parameter values for all fluxes modeled as Michaelis-Menten equations in provided in Figure \ref{fig:parameter_value_proof}a and \ref{fig:parameter_value_proof}b.
+Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) above is substituted with $V_1^{max} = k_1^{cat}E$ instead.
+The globally identifiable parameter values for all fluxes modeled as Michaelis-Menten equations ($v_1$ and $v_2$) in provided in Supplementary Figure \ref{fig:parameter_value_proof}a and \ref{fig:parameter_value_proof}b.
 
-	We apply the method described in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, and illustrated above for $v_1$, to other fluxes of the network (Figure \ref{fig:network}) not described using Michaelis-Menten equations.
+	We apply the method described in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, and illustrated above for $v_1$, to other fluxes of the network (Figure \ref{fig:network}) not described using Michaelis-Menten equations ($v_3$ and $v_5$).
 On applying this method to $v_3$, which is modeled using MWC kinetics, we find that we could not apply our methodology to determine closed form solution for all the parameters in $v_3$.
@@
 All three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ used to model $v_3$ have two possible roots.
-We still could not determine identifiable values for these parameters (Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}).
-In the box plot in Figure \ref{fig:parameter_value_v3}a and b we show all the possible estimated parameter values for both roots of all three parameters modeling $v_3$, when experimental data is generated using the MWC model for $v_3$.
+We still could not determine identifiable values for these parameters (Supplementary Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}).
+In the box plot in Supplementary Figure \ref{fig:parameter_value_v3}a and b we show all the possible estimated parameter values for both roots of all three parameters modeling $v_3$, when experimental data is generated using the MWC model for $v_3$.
 Due to the numerous estimated parameter values, we can designate these parameters are practically non-identifiable.
-Although using the Convenience kinetics model for $v_3$ to generate the in silico experimental data leads to a reduction in the uncertainty of the estimated parameters (\ref{fig:parameter_value_v3}c and d), we still end up with non-unique parameter estimates.
+Although using the Convenience kinetics model for $v_3$ to generate the in silico experimental data leads to a reduction in the uncertainty of the estimated parameters (Supplementary Figure \ref{fig:parameter_value_v3}c and d), we still end up with non-unique parameter estimates.
 This leads is to conclude that under the current conditions, flux $v_3$ is structurally locally identifiable and practically non-identifiable.
 
-	However, $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Figure \ref{fig:parameter_value_v3_var}).
+	However, $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Supplementary Figure \ref{fig:parameter_value_v3_var}).
 
-	Using Figure \ref{fig:parameter_value_v3_var}, we also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters.
-For example, in the specific case of $v_3$, we see that globally structurally identifiable parameters $V_3^{max}$ and $K_3^{fdp}$ (when $K_3^{pep}$ is fixed) are practically non-identifiable when data generated from experiments using the MWC kinetics is used to model $v_3$ (Figure \ref{fig:parameter_value_v3_var}a).
-The same is true of globally structurally identifiable parameters $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is fixed (Figure \ref{fig:parameter_value_v3_var}b).
-Instead, when Convenience kinetics is used to generate in silico data, under the aforementioned conditions all the parameters are globally structurally and practically identifiable (Figure \ref{fig:parameter_value_v3_var}c and d).
+	Through Supplementary Figure \ref{fig:parameter_value_v3_var}, we also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters.
+For example, in the specific case of $v_3$, we see that globally structurally identifiable parameters $V_3^{max}$ and $K_3^{fdp}$ (when $K_3^{pep}$ is fixed) are practically non-identifiable when data generated from experiments using the MWC kinetics is used to model $v_3$ (Supplementary Figure \ref{fig:parameter_value_v3_var}a).
+The same is true of globally structurally identifiable parameters $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is fixed (Supplementary Figure \ref{fig:parameter_value_v3_var}b).
+Instead, when Convenience kinetics is used to generate in silico data, under the aforementioned conditions all the parameters are globally structurally and practically identifiable (Supplementary Figure \ref{fig:parameter_value_v3_var}c and d).
 
-	Similarly, we applied the demonstrated methodology to flux $v_5$ that describes a transcription/translation reaction using Hill kinetics, based on only the available metabolomics and the fluxomics information.
+	Similarly, we applied the demonstrated methodology to identify flux $v_5$ that describes a transcription/translation reaction using Hill kinetics, based on only the available metabolomics and the fluxomics information.
 When we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system.
 However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$.
-Thus, similar to $v_3$ earlier, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in section \ref{sec:ident}.
+Thus, similar to $v_3$ earlier, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in Section \ref{sec:ident}.
 In this instance, the identification required data from two different experiments.
-Again, just as we see for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally identifiable.
+Again, just as we see for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally structurally identifiable.
 On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally structurally identifiable parameter.
 However, only one of the two roots ($K_e^{fdp}(2)$) is practically identifiable based on the physical relevance criteria stated in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b.
-Thus, enforcing this criteria results in global practical identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:parameter_value_proof}c).
+Thus, enforcing this criteria results in global practical identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$ (Supplementary Figure \ref{fig:parameter_value_proof}c).
 
-	So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism.
-Although in some instances (e.g., $v_5$) local identifiability could be resolved to obtain global identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$).
+	\textcolor{red}{(should this be moved to discussions?)}So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism.
+Although in some instances (e.g., $v_5$) local practical identifiability could be resolved to obtain global practical identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$).
 Moreover, we also see that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information.
 We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$.
-Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to parameter non-identifiability or local identifiability.
+Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to practical parameter non-identifiability or local identifiability.
 This can lead to uncertainty in parameter estimates made from in vivo experimental data.
@@
 We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition.
-This observation is also true irrespective of the kinetic rate law used to model $v_3$ (MWC in Figure \ref{fig:ident}a-d vs Convenience kinetics in Figure \ref{fig:ident}e-h) to generate in silico experimental data.
+The fact that in silico experimental data is generated using different kinetic rate law models for $v_3$ (MWC in Figure \ref{fig:ident}a-d vs Convenience kinetics in Figure \ref{fig:ident}e-h) has no bearing on this observation.
 We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
@@
 Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}.
-We elaborate on the type and nature of experiments that are useful and informative for parameter identification in $v_1$ and $v_2$ later in Section x.
+We elaborate on the type and nature of experiments that are useful and informative for parameter identification in $v_1$ and $v_2$ later in Section \ref{sec:experiment_type}.
 
-	The difference in the degree of identifiability of $v_1$ that comes about from the use of experimental data from different rate law models used to model $v_3$ (Figures \ref{fig:ident}a and e) can be again attributed to the usefulness of experimental data.
+	The difference in the degree of identifiability of $v_1$ that comes about from the use of experimental data from different rate laws used to model $v_3$ (Figures \ref{fig:ident}a and e) can be again attributed to the usefulness of experimental data.
 However, the difference in data usefulness between similar experiments arising due to differences in the kinetic rate law used for $v_3$ could be due to the difference in the dynamics of the network represented by either of the two models.
-The interaction between different components of the network (Figure \ref{fig:network}) as exemplified by the different models plays an important role in determining the dynamic and the steady state response of the network.
-The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets, and its subsequent effect of parameter identifiability.
+The interaction between different components of the network (Supplementary Figure \ref{fig:network}) as exemplified by the different models plays an important role in determining the dynamic and the steady state response of the network.
+The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Supplementary Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets, and its subsequent effect of parameter identifiability.
 
-	Nonetheless, we would like to note that these variations in the dynamics contributed by the difference between the rate laws can be resolved with more information i.e., experimental proteomic data that provide protein concentrations for different experiments.
+	Nonetheless, we would like to note that these variations in the dynamics contributed by the difference between the rate laws can be resolved with more information, i.e., experimental proteomic data that provide protein concentrations for different experiments.
 We earlier saw that relaxing the assumption on the available experiment data, i.e., assuming the availability of protein concentration (\textit{E}), in addition to the metabolite concentrations (\textit{pep} and \textit{fdp}) and fluxes, results in removal of uncertainty and global identifiability of $v_1$ (Supplementary Figure \ref{fig:parameter_value_proof}).
-Recall that any predicted parameter value for both $k_1^{cat}$ and $V_1^{max}$ matches with the exact known value irrespective of the experiment (MWC vs Convenience kinetics) from which data is obtained when enzyme concentrations are included for identifiability (Supplementary Figure \ref{fig:parameter_value_proof}a).
+Recall that any predicted parameter value for both $k_1^{cat}$ and $K_1^{ac}$ matches with the exact known value irrespective of the model (MWC vs Convenience kinetics) from which experimental data is obtained when enzyme concentrations are included for identifiability (Supplementary Figure \ref{fig:parameter_value_proof}a).
 Additionally, we also see an improvement in the degree of identifiability of the corresponding parameters $k_1^{cat}$ and $K_1^{ac}$ (Supplementary Figure \ref{fig:ident_root2}a and b) when enzyme concentrations are available.
@@
 	We would also like to point out that a difference in either the ability to estimate true parameter values, or the the identifiability of parameter modeling $v_2$ does not arise due to difference in the kinetic models used to obtain in silico experimental data.
-This could because of the fact that any and all available information that is required for the identification of parameters in $v_2$ is encompassed in the steady state data used for identifiability analysis.
+This could be because of the fact that any and all available information that is required for the identification of parameters in $v_2$ is encompassed in the steady state data used for identifiability analysis.
 
-	However, unlike $v_2$, we see that the degree of identifiability of both locally identifiable roots of parameters in $v_3$ changes with the change in the data (MWC vs Convenience kinetics) used for the analysis (Figure \ref{fig:ident}c and g).
+	However, unlike $v_2$, we see that the degree of identifiability of both locally identifiable roots of parameters in $v_3$ changes with the change in the data (MWC vs Convenience kinetics) used for the analysis (Figures \ref{fig:ident}c and g and Supplementary Figures \ref{fig:ident_root2}b and e).
 Despite the difference in the dynamics represented by the different data sets, and its impact on the usefulness of data sets for identifiability purposes that was explained earlier, we could also allude to the structural non-identifiability between the parameters in $v_3$ for the impact that it would have on the practical identifiability of the parameters using in silico data from different sources.
-In order to ascertain this, we plot the degrees of identifiability of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ under conditions when they are globally identifiable (Section \ref{sec:example}: $V_3^{max}$ and $K_3^{fdp}$ are globally identifiable when $K_3^{pep}$ is constant, and $V_3^{max}$, $K_3^{pep}$are globally identifiable when $K_3^{fdp}$ is known and fixed) in Supplementary Figure \ref{fig:ident_v3_var}.
-Now that structurally non-identifiability between parameters has been alleviated, we see that the uncertainty introduced by the data obtained from the MWC model results in only a fraction of data sets being useful to estimate parameters (Supplementary Figure \ref{fig:ident_v3_var}a and b).
-This ties back with our earlier observation that these parameters are not practically identifiable from data sets obtained using the MWC model (Supplementary Figure \ref{fig:parameter_value_v3_var}a and b).
+In order to ascertain this, we plot the degrees of identifiability of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ under conditions when they are globally identifiable in Supplementary Figure \ref{fig:ident_v3_var}.
+To recall from Section \ref{sec:example}, $V_3^{max}$ and $K_3^{fdp}$ are globally identifiable when $K_3^{pep}$ is constant, and $V_3^{max}$ and $K_3^{pep}$ are globally identifiable when $K_3^{fdp}$ is known and fixed.
+Now that structurally non-identifiability between parameters has been alleviated, we see that the uncertainty introduced by the data obtained from the MWC model still results in a degree of identifiability of only about 5\% for $K_3^{fdp}$ (Supplementary Figure \ref{fig:ident_v3_var}a).
+However, when $K_3^{fdp}$ is held constant all MWC model-based experimental datasets contribute to local practical identifiability of parameters $V_3^{max}$ and $K_3^{pep}$ (Supplementary Figure \ref{fig:ident_v3_var}c), albeit with a bot of uncertainty (Supplementary Figures \ref{fig:ident_v3_var}b).
+This ties back with our earlier observation that these parameters are not globally practically identifiable from data sets obtained using the MWC model (Supplementary Figure \ref{fig:parameter_value_v3_var}a and b).
 However, we see from Supplementary Figures \ref{fig:ident_v3_var}c and d that the globally structurally identifiable parameters are also practically identifiable based on the data obtained from the Convenience kinetics model.
@@
 	We did not observe any differences in the degree of identifiability of parameters in $v_5$ between the analyses that used data from either the MWC model or the Convenience kinetics model for $v_3$ (Figure \ref{fig:ident}d and h).
-The identifiability of parameters based on the first root obtained from the computer algebra system are shown in Figure \ref{fig:ident_root2}c and f.
+The identifiability of parameters based on the first root obtained from the computer algebra system are shown in Supplementary Figure \ref{fig:ident_root2}c and f.
+We use a similar reasoning to that we suggested for identifiability of $v_2$ above to justify our results.
 
@@
 	%We also showed how the methodology that we propose can help in determining the nature and type of experiments that one needs to perform to identify and estimate parameters in kinetic models in section \ref{sec:experiment_type}.
-	Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Banga papers on global optimization methods for parameter estimation, Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not gained enough traction (\textcolor{red}{include Rudiyanto Gunawan papers, D-optimal design using FIM here}).
+	\textcolor{red}{(should this section moved to SI?)}Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Banga papers on global optimization methods for parameter estimation, Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not gained enough traction (\textcolor{red}{include Rudiyanto Gunawan papers, D-optimal design using FIM here}).
 Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
@@
 These data are also in agreement with the degree of identifiability observed for parameters in all these fluxes (Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}).
-Thus, when numerous experiments are possible to obtain steady state data to estimate parameters for a small metabolic network, there is a very high possibility of performing experiments that could provide non-informative data that can result in parameter non-identifiability without a priori experimental design.
+Thus, when it is possible to obtain steady state data from numerous experiments to estimate parameters for a small metabolic network, there is a very high possibility of performing experiments that could provide non-informative data that can result in parameter non-identifiability without a priori experimental design.
 
@@
 Whereas, when data from the Convenience kinetics model is used for the same purpose, slightly more than 50\% of the combinations can identify all three parameters.
-Note that this observations should be considered in light of the structural local identifiability and the practical non-identifiability that we presented earlier in Sections \ref{sec:example} and \ref{sec:ident_rate_laws}.
+Note that this observations should be considered in light of the structural local identifiability and the practical non-identifiability of these parameters that we mentioned earlier in Sections \ref{sec:example} and \ref{sec:ident_rate_laws}.
 
@@
 In these figures we show the utility of multiple combinations of metabolite and protein concentration, and flux data towards identifying $k_1^{cat}$ and $K_1^{ac}$ in $v_1$ when data from the MWC model and Convenience kinetics model are used, respectively.
-We see that about 45\% of the 240 combinations of data from 21 different experiments cannot identify either one of the two parameters.
+We see that about 45\% of the 240 combinations of data from 21 different experiments cannot identify either one of the two parameters, when these parameters are globally structurally and practically identifiable.
 
@@
 
-	So, if we can establish a hierarchy of identifiable parameters and the corresponding fluxes they model, this section gives credence to the idea that a priori experimental design is necessary, even when all possible in vivo measurements can be obtained, to get the most informative data for parameter identification and estimation at minimum cost.
+	Accordingly, if we can establish a hierarchy of identifiable parameters and the corresponding fluxes they model, this section gives credence to the idea that a priori experimental design is necessary, even when all possible in vivo measurements can be obtained, to get the most informative data for parameter identification and estimation at minimum cost.
 Additionally, having the ability to see the contribution of each type of experiment towards identifiability would prove sufficient to decide on the total number and type of experiments that need to be performed to identify all fluxes in a kinetic model of metabolism.
-So, we look at the contribution of different types of experiments towards identifying parameters in each flux in the small network in the following section.
+So, next we look at the contribution of different types of experiments towards identifying parameters in each flux in the small network.
 
 	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}	
-	From Figures \ref{fig:ident} and \ref{sec:data_utility} we see that the variation in the degree of identifiability is not only dependent on the type of parameter being identified, but is also associated with the flux that these parameters model.
-For example, the maximum reactions rates of $v_2$ and $v_5$ have a degree of identifiability of 100\% (Figure \ref{fig:ident}b and d), while their counterparts $V_1^{max}$ and $V_3^{max}$ in $v_1$ and $v_3$ respectively (Figure \ref{fig:ident}a and c), are only identifiable to a lesser extent.
-	In systems identification terminology, this can be tied to selecting experiments that are persistently excitable for the flux being identified.
+	In the previous sections, we showed how ability to structurally identify parameters in a kinetic model of metabolism is dependent on the kinetic rate law used to model a specific flux (Section \ref{sec:example}).
+We also discussed how the the informativeness of experiments used to practically identify parameters in kinetic models is not only dependent on the in silico rate law, but is also a function of the uncertainty in the experimental data brought about by the differences between the in silico rate law and the in vivo rate law from which data is obtained (Section \ref{sec:ident_rate_laws}).
+Subsequently, we also showed the need for a priori experimental design to generate informative data to identify parameters in kinetic models.
+
+	In systems identification terminology, all the aforementioned observations regarding data requirements for parameters identification can be tied to selecting experiments that are persistently excitable for the flux being identified.
 In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
@@
 	We demonstrated in section \ref{sec:example}, using $v_1$ and $v_2$ as examples, that the ability of data from a given combination of experiments to identify parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter.
-For instance, identification of $V_1^{max}$ and $K_1^{ac}$ requires that the data from two different experiments used for parameter estimation must have distinct acetate concentrations as well as different uptake fluxes ($v_{1,j}$) between them (Equation \ref{eq:v1_par}).
-Thus, the informativeness of the experiments (changes in the input acetate concentration and $v_1$) for identifying parameters of $v_1$ is determined by the ability of the input change to effect a change in the measured value of $v_1$.	Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.
-
 Although we were able to determine the requirements for identifying $v_1$ and $v_2$ based on the closed form expressions of the parameters, this may not be possible when a complex nonlinear rate law is used to model a flux.
@@
 The contribution of different experiments towards identifiability for $V_2^{max}$ (Figure \ref{fig:exp_info}c) and $K_2^{pep}$ (Figure \ref{fig:exp_info}d) are also shown.
+Note that the contributions are only shown for data sets obtained using the MWC model in Figure \ref{fig:exp_info}.
+Experiment type contributions towards identifiability based on data sets obtained from the Convenience kinetics model for $v_3$ are shown separately in Supplementary Figure x.
 
-	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments (Figure \ref{fig:exp_info}a and b).
-This matches with requirements for identifiability on the basis of the informativeness of experiments that was laid out earlier i.e., experiments must be able distinguish between different values of $v_1$ and acetate.
-Accordingly, only about 40\% of the 240 different experiment combinations used to test identifiability are informative enough to discriminate both acetate as well as $v_1$, and can identify $V_1^{max}$ and $K_1^{ac}(ne)$ (Figure \ref{fig:ident}a).
+	In corroboration of what we discussed in Section \ref{sec:example}, i.e., experiments must be able distinguish between different values of $v_1$ and acetate, the contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments (Figure \ref{fig:exp_info}a and b).
+
+	For instance, identification of $V_1^{max}$ and $K_1^{ac}$ requires that the data from two different experiments used for parameter estimation must have distinct acetate concentrations as well as different uptake fluxes ($v_{1,j}$) between them (Equation \ref{eq:v1_par}).
+Thus, the informativeness of the experiments (changes in the input acetate concentration and $v_1$) for identifying parameters of $v_1$ is determined by the ability of the input change to effect a change in the measured value of $v_1$.	Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.
 
@@
 Hence, in conjunction with the high degree of identifiability seen in Figure \ref{fig:ident}b for $v_2$, we can deduce that almost all experiment types are persistently excitable for $v_2$ i.e., most experiments can bring about noticeable changes to both \textit{pep} and $v_2$.
-Consequently, in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative enough to identify $v_2$.
-%This observations is in line with the one made in section \ref{sec:data_utility} based on the total number of combinations that can identify both parameters in $v_2$.
+Consequently, this confirms our earlier hypothesis that in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative enough to identify $v_2$.
 
@@
 
-	%Thus, Figure \ref{fig:data_utility} gives credence to the idea that careful experimental design is necessary to minimize usage of resources devoted to performing experiments for parameter estimation.
-For instance, choosing any of the experiment combinations from the 52\% percent that does not identify any parameter in $v_1$, or the 6\% that does not identify any parameter in $v_3$ can lead to experiments that are not informative.
-This can lead to potential waste of resources used to perform these experiments.
-
 	\subsection{Effect of noise on degree of identifiability is dependent on nonlinearity of enzyme kinetic rate law models}\label{sec:noise}
-	In reality, experimental data from biological systems is usually noisy.
+	\textcolor{red}{(Should this section be moved to the SI as well?)}In reality, experimental data from biological systems is usually noisy.
 Noise in data gathered from experiments is associated with biological noise attributable to the stochasticity of cellular function, as well as measurement noise associated with the techniques used to obtain concentration and flux measurements.

### 2018-04-15 ident/docs/draft_2.tex
changes to figures and moved all si figures to si figure file

@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_v1max_v2_exp_info}}
-		\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$ and d) $K_2^{pep}$.
+		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_v2_v5_2_exp_info}}
+		\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$, d) $K_2^{pep}$, e) $V_e^{max}$ and f) $K_e^{fdp}(2)$.
 The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:exp_info}
-	\end{figure}
-
-	\clearpage
-
-	\section{Potential Supplementary Information Figures}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_1_2_exp_info}}
-		\caption{The contribution of different experiments types used in a combination of three experiments ($j = \{1, 2, 3\}$) that can practically identify parameter a) $V_3^{max}(1)$, b) $K_3^{fdp}(1)$, c) $K_3^{pep}(1)$, d) $V_3^{max}(2)$, e) $K_3^{fdp}(2)$ and f) $K_3^{pep}(2)$ for $v_3$.
-parameters $V_3^{max}$, $K_3^{fdp}$, and $K_3^{pep}$ are locally identifiable and have two possible roots.
-The first set of roots are designated as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$, and the second set of roots are designated as $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.
-The contribution of the five different experiment types towards identifiability of either of the two roots for each of the three parameters is the same.}\label{fig:v3_exp_info}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure2/v5_1_2_exp_info}}
-		\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_e^{max}$ and b) $K_e^{fdp}$ in $v_5$.
-Since $V_e^{max}$ is globally identifiable, it has a unique root.
-Since $K_e^{fdp}$ has two roots, is locally identifiable.
-However, results for the only practically identifiable root, based on the physical relevance constraints used in the analysis, is shown.}\label{fig:v5_exp_info}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_k1cat_exp_info}}
-		\caption{Contribution of different experiment types towards identification of a) $k_1^{cat}$ and b) $K_1^{ac}$ in $v_1$ based on using data from two different experiments.}%\label{fig:figure2}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_exp_info}}
-		\caption{Contribution of different types of experiments for using data from two different experiments towards identification of a) $V_3^{max}$ and b) $K_3^{fdp}$ in $v_3$ when $K_3^{pep}$ is held constant, and c) $V_3^{max}$ and d) $K_3^{pep}$ when $K_3^{fdp}$ is held constant.}%\label{fig:figure2}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v3_1_2_data_utility}}
-		\caption{Utility of data from combinations of two different experiments for identifying two parameters in $v_3$ when a) $K_3^{pep}$ is held constant, and when b) $K_3^{fdp}$ is held constant.
-Notice that a significant number of combinations can identify both $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is known and fixed.}%\label{fig:figure2}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_k1cat_parameter_value}}
-		\caption{Distribution of identified/estimated parameter values for flux $v_1$ from different experimental data combinations.
-The box plot of the estimated values of $V_1^{max}$ and $K_1^{ac}(ne)$ are juxtaposed with the box plots of the calculated values of $k_1^{cat}$ based on the estimated value of $V_1^{max}$ and the enzyme concentrations available from the two different experiments used to estimate $V_1^{max}$ and $K_1^{ac}(ne)$}\label{fig:parameter_value_v1_v1max_kcat}
 	\end{figure}

### 2018-04-15 ident/docs/sifigures.tex
most likely has all needed si figures

@@
 
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v3_1_2_data_utility}}
+	\caption{Utility of data from combinations of two different experiments for identifying two parameters in $v_3$ when a) $K_3^{pep}$ is held constant, and when b) $K_3^{fdp}$ is held constant.
+Notice that a significant number of combinations can identify both $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is known and fixed.}%\label{fig:figure2}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_k1cat_exp_info}}
+	\caption{Contribution of different experiment types towards identification of a) $k_1^{cat}$ and b) $K_1^{ac}$ in $v_1$ based on using data from two different experiments.
+Data from MWC model.}%\label{fig:figure2}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure2/v1_V1max_v2_v5_2_ck_exp_info}}
+	\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_e^{max}$ and b) $K_e^{fdp}$ in $v_5$.
+Since $V_e^{max}$ is globally identifiable, it has a unique root.
+Since $K_e^{fdp}$ has two roots, is locally identifiable.
+However, results for the only practically identifiable root, based on the physical relevance constraints used in the analysis, is shown.
+Data from Convenience kinetics model is used.}%\label{fig:v5_exp_info}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_1_2_exp_info}}
+	\caption{The contribution of different experiments types used in a combination of three experiments ($j = \{1, 2, 3\}$) that can practically identify parameter a) $V_3^{max}(1)$, b) $K_3^{fdp}(1)$, c) $K_3^{pep}(1)$, d) $V_3^{max}(2)$, e) $K_3^{fdp}(2)$ and f) $K_3^{pep}(2)$ for $v_3$.
+parameters $V_3^{max}$, $K_3^{fdp}$, and $K_3^{pep}$ are locally identifiable and have two possible roots.
+The first set of roots are designated as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$, and the second set of roots are designated as $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
+The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.
+The contribution of the five different experiment types towards identifiability of either of the two roots for each of the three parameters is the same.}\label{fig:v3_exp_info}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_exp_info}}
+	\caption{Contribution of different types of experiments for using data from two different experiments towards identification of a) $V_3^{max}$ and b) $K_3^{fdp}$ in $v_3$ when $K_3^{pep}$ is held constant, and c) $V_3^{max}$ and d) $K_3^{pep}$ when $K_3^{fdp}$ is held constant.}%\label{fig:figure2}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_k1cat_parameter_value}}
+	\caption{Distribution of identified/estimated parameter values for flux $v_1$ from different experimental data combinations.
+The box plot of the estimated values of $V_1^{max}$ and $K_1^{ac}(ne)$ are juxtaposed with the box plots of the calculated values of $k_1^{cat}$ based on the estimated value of $V_1^{max}$ and the enzyme concentrations available from the two different experiments used to estimate $V_1^{max}$ and $K_1^{ac}(ne)$}\label{fig:parameter_value_v1_v1max_kcat}
+\end{figure}
+
 \clearpage

### 2018-04-16 ident/docs/draft_2.tex
minor changes to results

@@
 	In corroboration of what we discussed in Section \ref{sec:example}, i.e., experiments must be able distinguish between different values of $v_1$ and acetate, the contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments (Figure \ref{fig:exp_info}a and b).
-
-	For instance, identification of $V_1^{max}$ and $K_1^{ac}$ requires that the data from two different experiments used for parameter estimation must have distinct acetate concentrations as well as different uptake fluxes ($v_{1,j}$) between them (Equation \ref{eq:v1_par}).
-Thus, the informativeness of the experiments (changes in the input acetate concentration and $v_1$) for identifying parameters of $v_1$ is determined by the ability of the input change to effect a change in the measured value of $v_1$.	Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.
+Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.
 

### 2018-04-16 ident/python2/ss-ident/add_noise.py
get noise added samples as list of 2-d arrays

@@ -14,9 +14,10 @@ def add_noise(data, default_shape, number_of_samples=1):
         rows, columns = default_shape
     noise = prng_1.normal(0, .05, (rows, columns, number_of_samples))
     # concentration_noise = np.random.normal(0, .05, [rows, columns])
-    noisy_data = np.zeros((rows, columns, number_of_samples))
-    for i_sample in range(0, number_of_samples):
-        noisy_data[:, :, i_sample] = data * (1 + noise[:, :, i_sample])
+    # noisy_data = np.zeros((rows, columns, number_of_samples))
+    noisy_data = [data * (1 + noise[:, :, i_sample]) for i_sample in range(0, number_of_samples)]
+    # for i_sample in range(0, number_of_samples):
+    #     noisy_data[:, :, i_sample] = data * (1 + noise[:, :, i_sample])
     return noisy_data
 
 

### 2018-04-16 ident/python2/ss-ident/generate_noisy_data.py
work with list of 2-d arrays instead of 3-d arrays

@@ -100,17 +100,20 @@ def generate_noisy_data(y0, all_options, kinetics, number_of_samples=1):
     all_sample_noisy_bistable = []
     for i_sample in range(0, number_of_samples):
         # info on bistability
-        if noisy_concentration_dynamic[-1, 0, i_sample] > noisy_concentration_dynamic[-1, 1, i_sample]:
+        if noisy_concentration_dynamic[i_sample][-1, 0] > noisy_concentration_dynamic[i_sample][-1, 1]:
             bistable = 1
-        elif noisy_concentration_dynamic[-1, 0, i_sample] < noisy_concentration_dynamic[-1, 1, i_sample]:
+        elif noisy_concentration_dynamic[i_sample][-1, 0] < noisy_concentration_dynamic[i_sample][-1, 1]:
             bistable = 2
         else:
             bistable = 0
         all_sample_noisy_bistable.append(bistable)
     noisy_dynamic_info = {"y": noisy_concentration_dynamic, "flux": noisy_flux_dynamic,
                           "time": dynamic_info["time"], "ssid": all_sample_noisy_bistable}
-    noisy_steady_state_info = {"y": noisy_concentration_dynamic[-1, :, :],
-                               "flux": noisy_flux_dynamic[-1, :, :], "ssid": all_sample_noisy_bistable}
+    noisy_steady_state_info = {"y": [noisy_concentration_dynamic[i_sample][-1, :]
+                                     for i_sample in range(0, number_of_samples)],
+                               "flux": [noisy_flux_dynamic[i_sample][-1, :]
+                                        for i_sample in range(0, number_of_samples)],
+                               "ssid": all_sample_noisy_bistable}
         # all_sample_noisy_dynamic_info.append(noisy_dynamic_info)
         # all_sample_noisy_steady_state_info.append(noisy_steady_state_info)
 

### 2018-04-16 ident/python2/ss-ident/plot_ident_results.py
create exception when data with noise is plotted

@@ -350,7 +350,7 @@ def line_plots_2d(axis_obj, x_data, y_data, x_label=(), y_label=()):
     return None
 
 
-def plot_dynamic_sim_concentrations(dynamic_data, multiple=0):
+def plot_dynamic_sim_concentrations(dynamic_data, multiple=0, noise=0):
     """plot dynamic time course profiles of all concentrations present in y vector.
     parameter multiple = 1 when multiple sets of plots are plotted in a subplot"""
     if multiple:
@@ -383,7 +383,11 @@ def plot_dynamic_sim_concentrations(dynamic_data, multiple=0):
         f, axis_obj = plt.subplots(1, 1, figsize=(6, 4), dpi=100, facecolor='w', edgecolor='k')
         x_data = dynamic_data["time"]
         y_data = dynamic_data["y"]
-        line_plots_2d(axis_obj, x_data, y_data, x_label='Time (s)', y_label='Concentrations (a.u.)')
+        if noise:
+            # plt only one of many noisy slutions (Figure out a way to plot all)
+            line_plots_2d(axis_obj, x_data, y_data[0], x_label='Time (s)', y_label='Concentrations (a.u.)')
+        else:
+            line_plots_2d(axis_obj, x_data, y_data, x_label='Time (s)', y_label='Concentrations (a.u.)')
         axis_obj.set_title('Dynamic Concentrations')
     plt.show()
     return None

### 2018-04-16 ident/python2/ss-ident/generate_expdata.py
add noise parameter to dynamic plot call

@@ -10,6 +10,7 @@ def initialize_to_ss(y0, cvode_options, ode_parameter_values, noise=1, kinetics=
     if noise:
         # get initial noisy system steady state
         initial_options = (cvode_options, ode_parameter_values)
+        # use non-noisy initial ss (wt) data
         noisy_initial_ss, noisy_initial_dyn, _, _ = generate_noisy_data(y0, initial_options, kinetics=kinetics)
         return noisy_initial_ss, noisy_initial_dyn
     else:
@@ -46,7 +47,7 @@ def generate_expdata(y0, cvode_options, ode_parameter_values, number_of_samples=
 
     # plot all dynamic courses
     if dynamic_plot:
-        plot_dynamic_sim_concentrations(initial_dyn)
+        plot_dynamic_sim_concentrations(initial_dyn, noise=noise)
 
     # all parameter perturbations
     parameter_perturbation = [{"ac": 0}, {"ac": 1}, {"ac": 4}, {"ac": 9}, {"ac": -.1}, {"ac": -.5},

### 2018-04-16 ident/python2/ss-ident/generate_expdata.py
use only data from first ss sample for perturbation analysis + replace for loops with list comprehension for collecting final perturbation data

@@ -27,8 +27,8 @@ def perturb_parameters(initial_ss, parameter_perturbations, cvode_options, ode_p
     if noise:
         perturbation_options = {'ode_parameters': ode_parameter_values, 'cvode_options': cvode_options}
         noisy_ss, noisy_dynamic, perturbation_details, _, dynamic_info = \
-            run_noisy_parameter_perturbation(parameter_perturbations, initial_ss["y"], perturbation_options,
-                                             number_of_samples)
+            run_noisy_parameter_perturbation(parameter_perturbations, initial_ss["y"][0], perturbation_options,
+                                             kinetics=kinetics, number_of_samples=number_of_samples)
         return noisy_ss, perturbation_details
     else:
         perturbation_options = {'ode_parameters': ode_parameter_values, 'cvode_options': cvode_options}
@@ -65,25 +65,17 @@ def generate_expdata(y0, cvode_options, ode_parameter_values, number_of_samples=
                                                                 number_of_samples, noise=noise, kinetics=kinetics,
                                                                 dynamic_plot=perturbation_plot)
 
-    all_sample_exp_xss = []
-    all_sample_exp_fss = []
-    all_sample_exp_ssid = []
-    for i_sample in range(0, number_of_samples):
-        exp_xss = []
-        exp_fss = []
-        exp_ssid = []
-        for ss_values in perturbed_ss:
-            try:
-                exp_xss.append(ss_values["y"][:, i_sample])
-                exp_fss.append(ss_values["flux"][:, i_sample])
-                exp_ssid.append(ss_values["ssid"][i_sample])
-            except IndexError:
-                exp_xss.append(ss_values["y"][:, ])
-                exp_fss.append(ss_values["flux"][:, ])
-                exp_ssid.append(ss_values["ssid"])
-        all_sample_exp_xss.append(exp_xss)
-        all_sample_exp_fss.append(exp_fss)
-        all_sample_exp_ssid.append(exp_ssid)
+    if noise:
+        all_sample_exp_xss = [[i_perturbation_data["y"][i_sample] for i_perturbation_data in perturbed_ss]
+                              for i_sample in range(0, number_of_samples)]
+        all_sample_exp_fss = [[i_perturbation_data["flux"][i_sample] for i_perturbation_data in perturbed_ss]
+                              for i_sample in range(0, number_of_samples)]
[…]

### 2018-04-16 ident/python2/ss-ident/generate_noisy_data.py
convert noisy code to use parameter dicts instead of tuples and arrays similar to non-noisy code

@@ -177,31 +177,33 @@ def run_no_noise_parameter_perturbation(parameter_perturbation, y0, other_option
     return ss_info, dynamic_info, perturbation_details
 
 
-def run_noisy_parameter_perturbation(parameter_perturbation, y0, other_options, number_of_samples=1, plot_arg=0):
+def run_noisy_parameter_perturbation(parameter_perturbation, y0, other_options, kinetics=1,
+                                     number_of_samples=1, plot_arg=0):
     """run parameter perturbations based on tuple input parameter perturbation
     with first position of tuple being parameter id and second index being
     parameter value"""
 
-    ode_parameters = tuple(other_options['ode_parameters'])
+    ode_parameters = other_options['ode_parameters']
     cvode_options = other_options['cvode_options']
     noisy_ss = []
     noisy_dynamic = []
     ss_info = []
     dynamic_info = []
     experiment_info_boolean = []
-    perturbed_parameter = np.zeros((len(parameter_perturbation), len(ode_parameters)))
-    perturbation_indices = np.zeros((len(parameter_perturbation), 2))
+    perturbed_parameter = []
+    perturbation_indices = []
     perturbation_ssid = np.zeros((len(parameter_perturbation), 2))
 
-    for index, p_value in enumerate(parameter_perturbation):
+    for index, perturbation_value in enumerate(parameter_perturbation):
         print('Perturbation {}\n'.format(index+1))
-        parameter_id, parameter_change = p_value
-        changed_ode_parameter = np.array(ode_parameters[:])
-        changed_ode_parameter[parameter_id-1] = changed_ode_parameter[parameter_id-1]*(1 + parameter_change)
+        parameter_name = perturbation_value.keys()[0]
+        parameter_change = np.array(perturbation_value.values()[0])
+        changed_ode_parameter = deepcopy(ode_parameters)
+        changed_ode_parameter[parameter_name] = changed_ode_parameter[parameter_name] * (1 + parameter_change)
         all_options = (cvode_options, changed_ode_parameter)
         # generate data using MWC Kinetics
         noisy_ss_iter, noisy_dynamic_iter, ss_iter, dynamic_iter = \
-            generate_noisy_data(y0, all_options, 1, number_of_samples)
+            generate_noisy_data(y0, all_options, kinetics=kinetics, number_of_samples=number_of_samples)
 
         noisy_ss.append(noisy_ss_iter)
         noisy_dynamic.append(noisy_dynamic_iter)
[…]

### 2018-04-16 ident/python2/ss-ident/v3_var_ck_script.py
no major change

@@ -56,7 +56,7 @@ experiment_type_indices = [[0],
                            [16, 17, 18, 19, 20]]
 # perform identifiability when v3 parameters are written for root (1)
 ident_details_v3_root1 = flux_ident_2_data_combination(experimental_datasets_2_expts, choose=combination_choice,
-                                                       flux_ids=[6], flux_choice=[2], ident_fun_choice=ident_fun_choice)
+                                                       flux_ids=[6], flux_choice=[1], ident_fun_choice=ident_fun_choice)
 print('Identifiability analysis for v3 with 3 parameters (V3max, K3fdp and K3pep) complete.\n')
 # data processing
 data_list_v3_root1, max_parameter_v3_root1, true_value_v3_root1, experiment_info_v3_root1, \

### 2018-04-16 ident/python2/ss-ident/v1_Vmax_noise_script.py
script for noisy data ident for v1 with Vmax model

@@ -1,0 +1,91 @@
+import numpy as np
+from generate_expdata import generate_expdata
+from simulate_data import arrange_experimental_data
+from kotte_model import flux_ident_2_data_combination
+from process_ident_data import process_info_sample
+from plot_ident_results import parameter_identifibaility_plot
+from plot_ident_results import data_utility_plot
+from plot_ident_results import plot_parameter_values
+from plot_ident_results import parameter_experiment_info_spider
+from plot_ident_results import plot_experiment_data_dist
+
+# generate noisy experimental data for testing identifiability
+y0 = np.array([5, 1, 1])
+# default parameter values
+cvode_options = ('Newton', 'Adams', 1e-10, 1e-10, 200)
+ode_parameter_values = {"K1ac": np.array([.1]),
+                        "K3fdp": np.array([.1]),
+                        "L3fdp": np.array([4e6]),
+                        "K3pep": np.array([.1]),
+                        "K2pep": np.array([.3]),
+                        "vemax": np.array([1.1]),
+                        "Kefdp": np.array([.45]),
+                        "ne": np.array([2]),
+                        "d": np.array([.25]),
+                        "V4max": np.array([.2]),
+                        "k1cat": np.array([1]),
+                        "V3max": np.array([1]),
+                        "V2max": np.array([1]),
+                        "ac": np.array([.1])}
+
+# get experimental system steady state data without noise
+exp_xss, exp_fss, exp_ssid, perturbation_details = \
+    generate_expdata(y0, cvode_options, ode_parameter_values, noise=1, number_of_samples=5, kinetics=1,
+                     dynamic_plot=1, perturbation_plot=0)
+
+# arrange experimental data to form multiple data sets
+exp_flux_index = np.array([0, 3, 2, 4, 1, 5])
+
+# get combination of 2 experiments and perform identifiability on all fluxes that require 2 data sets
+# get combinations of experimental datasets
+experimental_datasets_2_expts, \
+    experiment_choice, combination_choice = arrange_experimental_data(exp_xss, exp_fss, perturbation_details,
+                                                                      experiments_per_set=2, flux_id=exp_flux_index,
+                                                                      experiment_choice=[0, 1, 2, 3, 4, 5,
+                                                                                         11, 12, 13, 14, 15,
+                                                                                         16, 17, 18, 19, 20])
+
[…]

### 2018-04-16 ident/docs/draft_2.tex
a more polished discussion - still needs work

@@
 
-	\textcolor{red}{(should this be moved to discussions?)}So, in summary, we find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism.
-Although in some instances (e.g., $v_5$) local practical identifiability could be resolved to obtain global practical identifiability using constraints on the values of the parameters such that they are physically relevant, this may not be possible in other scenarios (e.g., $v_3$).
-Moreover, we also see that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information.
-We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$.
-Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to practical parameter non-identifiability or local identifiability.
-This can lead to uncertainty in parameter estimates made from in vivo experimental data.
-
 	In the following section, we describe and summarize the results on the identifiability of parameters in these fluxes ($v_1$, $v_2$, $v_3$ and $v_5$) based on the expressions and the identifiable roots we discussed in this section.
@@
 Note that the contributions are only shown for data sets obtained using the MWC model in Figure \ref{fig:exp_info}.
-Experiment type contributions towards identifiability based on data sets obtained from the Convenience kinetics model for $v_3$ are shown separately in Supplementary Figure x.
+Experiment type contributions towards identifiability based on data sets obtained from the Convenience kinetics model for $v_3$ are shown separately in Supplementary Figure .
 
-	In corroboration of what we discussed in Section \ref{sec:example}, i.e., experiments must be able distinguish between different values of $v_1$ and acetate, the contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments (Figure \ref{fig:exp_info}a and b).
+	In agreement with what we discussed in Section \ref{sec:example}, i.e., experiments must be able distinguish between different values of $v_1$ and acetate, the contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments (Figure \ref{fig:exp_info}a and b).
 Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.
@@
 	On the other hand, the Michaelis-Menten model for $v_2$ also requires changes in \textit{pep} and $v_2$ for persistently excitable experiments to identify $v_2$.
-However, since both of these are system outputs, satisfaction of this condition cannot be guaranteed without an analysis of the dynamics of the metabolic network and how changes in the input (acetate) bring about changes in the two requisite output quantities.
+However, since both of these are system outputs, satisfaction of this condition cannot be guaranteed without an analysis of the dynamics of the metabolic network, and how changes in the input (acetate) bring about changes in the two requisite output quantities.
 Previous dynamical analysis of the network (Figure \ref{fig:network}) has already established the existence of a functional relationship between \textit{pep} and $v_2$, and the input acetate concentration and the levels of expression of the different enzymes within the network \parencite{Srinivasan2017}.
@@
 
-	We can extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_3$ as well (Figure \ref{fig:v3_exp_info}).
-In this case, the top three panels (Figure \ref{fig:v3_exp_info}a-c) show the distribution of different experiment types for identifying each parameter with respect to their first root identified as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$.
-The bottom panels (Figure \ref{fig:v3_exp_info}d-f) show the corresponding distribution of experiments for the second root of each fo three parameters in $v_3$: $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
+	We can extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_3$ as well (Supplementary Figure \ref{fig:v3_exp_info}).
+In this case, the top three panels (Supplementary Figure \ref{fig:v3_exp_info}a-c) show the distribution of different experiment types for identifying each parameter with respect to their first root identified as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$.
+The bottom panels (Supplementary Figure \ref{fig:v3_exp_info}d-f) show the corresponding distribution of experiments for the second root of each fo three parameters in $v_3$: $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
 
-	In all of the above scenarios for $v_1$ (Figure \ref{fig:exp_info}a and b), $v_2$ (Figure \ref{fig:exp_info}c and d) and $v_3$ (Figure \ref{fig:v3_exp_info}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar.
+	In all of the above scenarios for $v_1$ (Figure \ref{fig:exp_info}a and b), $v_2$ (Figure \ref{fig:exp_info}c and d) and $v_3$ (Supplementary Figure \ref{fig:v3_exp_info}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar.
 Hence, the green and blue surfaces are superimposed on top of each other.
-This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:v5_exp_info}a and b).
-However, this is not the case with the distribution of experiments required for identifying $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:v5_exp_info}c).
+This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:exp_info}e).
+However, this is not the case with the distribution of experiments required for identifying $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:exp_info}f).
 We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$.
@@
 	\section{Discussions}\label{sec:discussion}	
-	\textcolor{red}{(move to discussion)} The impact of the formulation and nonlinearity of a kinetic rate law expression affecting the practical identifiability of parameters in the expression may not be an unique problem isolated to the system that we are investigating.
-Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms \parencite{Chassagnole2002a, Peskov2012, VanHeerden2014}.
-However, the identifiability of the parameters used in these models has never been truly examined.
-\textcolor{red}{mentione Heijnen papers on resolving identifiability using approximate models here}.
-
-	\textcolor{red}{(Move to discussion)} Unlike the case of the MWC model, if determining closed-form expressions for the parameters of these models are tractable, we believe our methodology can help in elucidating the identifiability of these parameters.
-If not, metabolic network fluxes can be expressed using alternative kinetic rate law models whose parameters can be tested for identifiability, and subsequently experiments can be designed for their estimation and model validation.
-
 	Parameter estimation for kinetic models has always focused on the ability to estimate parameters from existing data without the need for additional experiments, which might not be always possible if parameters are not identifiable from existing experimental data.
 The presence of noise is typically said to be a significant factor that results in non-identifiability.
-However, the primary reason for non-identifiability of parameters from experimental data can be attributed to the lack of information about the dynamics of the system whose parameters are being estimated within the chosen experimental data.
+However, there different reasons for non-identifiability of parameters that we show with our work.
+First, non-identifiability could be structural to the model used to represent the flux, and cannot be alleviated without reduction in the parameter space.
+Otherwise, non-identifiability of parameters can be attributed to the lack of information about the dynamics of the system whose parameters are being estimated within the chosen experimental data.
 The informativeness of experiments can be tied back to their ability to discriminate the dynamics of the system under two or more different input conditions.
@@
 
-	Although methods have been developed for practical parameter identification and experimental design, existing methods for experimental design based on practical identification of parameters are based on solving nonlinear least squares problems using optimization approaches that cannot guarantee global optimal solutions \parencite{Raue2009a}.
+	Previously, methods have been developed for practical parameter identification and experimental design for kinetic models of metabolism.
+These methods for experimental design based on practical identification of parameters rely on solving nonlinear least squares problems using optimization approaches that cannot guarantee global optimal solutions \parencite{Raue2009a}, or calculating the Fischer Information Matrix (FIM) to obtain information on the structural and practical identifiability of parameters in kinetic models.
+Either of these types of methods become computationally cumbersome for models of large genome-scale, or even central carbon scale metabolic networks.
+Some authors have eschewed deterministic parameter estimation techniques in favour of Bayesian methods based on probabilistic estimation of parameters and experimental design \parencite{Saa2016, Saa2016a} that has the possibility of overcoming some of the issues with the deterministic techniques.
+
 	In this document, we have presented a scalable method to practically identify parameters in kinetic models of metabolism, and use it to design experiments that are minimal and informative for estimating the parameters that does not require solutions to non-convex optimization problems.
-While some authors have eschewed deterministic parameter estimation techniques in favour of Bayesian methods based on probabilistic estimation of parameters \parencite{Saa2016, Saa2016a}, The need to establish practical identifiability of parameters, even for these methods, still stands.
-Additionally, the ability to design experiments to get good priors for Bayesian approaches to parameter estimation can also be fulfilled through practical identifiability analysis.
-
-	Due to their fundamental aspect of requiring solutions to non-convex optimization problems, existing methods for practical identifications may not be scalable for identification and experimental design for parameters of large metabolic networks.
 By establishing identifiability for each flux within a metabolic network individually, we hope to overcome the scalability obstacle.
@@
 We have also shown the ability to use our analysis to design the minimal number of experiments that are most informative for identifying all fluxes within a metabolic network.
+
+	We find that the identifiability of parameters in kinetic models of metabolism using steady state information is dependent on the kinetic rate law used to model the fluxes within metabolism.
+The impact of the formulation and nonlinearity of a kinetic rate law expression affecting the practical identifiability of parameters in the expression may not be an unique problem isolated to the system that we are investigating.
+Complicated expressions for describing fluxes have been extensively used to model observed experimental data for different fluxes in a variety of organisms \parencite{Chassagnole2002a, Peskov2012, VanHeerden2014}.
+However, authors have favored working with approximate kinetic models of metabolism whose parameters are easily identifiable and estimable instead of trying to establish the identifiability of the parameters used in these models \textcolor{red}{(mention Heijnen papers on resolving identifiability using approximate models here)}.
+
+	We have shown that in some instances (e.g., $v_5$) local practical identifiability could be resolved to obtain global practical identifiability using constraints on the values of the parameters such that they are physically relevant.
+We have also shown that the structural identifiability of the parameters in any given kinetic rate law model has a bearing on the ability to determine the practical identifiability of parameters using steady state metabolomic, fluxomic and proteomic information.
+We find that these can sometimes be resolved by reducing the dimension of the parameter space that is being identified: $\theta \in \mathbb{R}^3$ to $\theta \in \mathbb{R}^2$ for both $v_3$ and $v_2$.
+Additionally, we would also like to point out that discrepancies between in vivo kinetic rate law from which typical experimental data is obtained, and the in vitro rate law used in kinetic models can itself lead to practical parameter non-identifiability or local identifiability.
+This can lead to uncertainty in parameter estimates made from in vivo experimental data.
 
@@
 	Methods and tools for structural identification of parameters based on differential algebra \parencite{Ljung1994, Audoly2001, Bellu2007} and profile likelihood \parencite{Raue2014} are available.
-However, only the profile likelihood-based methods enable experimental design by facilitating practical identification of parameters.
-Nonetheless, this method still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters, and hence still suffers from all the inherent difficulties associated with obtaining global optimal solutions for non-convex optimization problems.
-This also makes it un-scalable for experimental design and practical identifiability of parameters in kinetic models of large metabolic networks.
 

### 2018-04-16 ident/python2/ss-ident/v1_noise_script.py
noise script for v1 with kcat model

@@ -1,0 +1,92 @@
+import numpy as np
+from generate_expdata import generate_expdata
+from simulate_data import arrange_experimental_data
+from kotte_model import flux_ident_2_data_combination
+from process_ident_data import process_info_sample
+from plot_ident_results import parameter_identifibaility_plot
+from plot_ident_results import data_utility_plot
+from plot_ident_results import plot_parameter_values
+from plot_ident_results import parameter_experiment_info_spider
+from plot_ident_results import plot_experiment_data_dist
+
+
+# generate noisy experimental data for testing identifiability
+y0 = np.array([5, 1, 1])
+# default parameter values
+cvode_options = ('Newton', 'Adams', 1e-10, 1e-10, 200)
+ode_parameter_values = {"K1ac": np.array([.1]),
+                        "K3fdp": np.array([.1]),
+                        "L3fdp": np.array([4e6]),
+                        "K3pep": np.array([.1]),
+                        "K2pep": np.array([.3]),
+                        "vemax": np.array([1.1]),
+                        "Kefdp": np.array([.45]),
+                        "ne": np.array([2]),
+                        "d": np.array([.25]),
+                        "V4max": np.array([.2]),
+                        "k1cat": np.array([1]),
+                        "V3max": np.array([1]),
+                        "V2max": np.array([1]),
+                        "ac": np.array([.1])}
+
+# get experimental system steady state data without noise
+exp_xss, exp_fss, exp_ssid, perturbation_details = \
+    generate_expdata(y0, cvode_options, ode_parameter_values, noise=1, number_of_samples=100, kinetics=1,
+                     dynamic_plot=1, perturbation_plot=0)
+
+# arrange experimental data to form multiple data sets
+exp_flux_index = np.array([0, 3, 2, 4, 1, 5])
+
+# get combination of 2 experiments and perform identifiability on all fluxes that require 2 data sets
+# get combinations of experimental datasets
+experimental_datasets_2_expts, \
+    experiment_choice, combination_choice = arrange_experimental_data(exp_xss, exp_fss, perturbation_details,
+                                                                      experiments_per_set=2, flux_id=exp_flux_index,
+                                                                      experiment_choice=[0, 1, 2, 3, 4, 5,
+                                                                                         11, 12, 13, 14, 15,
+                                                                                         16, 17, 18, 19, 20])
+
[…]

### 2018-04-16 ident/python2/ss-ident/v1_Vmax_noise_script.py
change number of sample to 100

@@ -30,7 +30,7 @@ ode_parameter_values = {"K1ac": np.array([.1]),
 
 # get experimental system steady state data without noise
 exp_xss, exp_fss, exp_ssid, perturbation_details = \
-    generate_expdata(y0, cvode_options, ode_parameter_values, noise=1, number_of_samples=5, kinetics=1,
+    generate_expdata(y0, cvode_options, ode_parameter_values, noise=1, number_of_samples=100, kinetics=1,
                      dynamic_plot=1, perturbation_plot=0)
 
 # arrange experimental data to form multiple data sets

### 2018-04-16 ident/docs/sifigures.tex

@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure4/v3_1_2_data_utility}}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure4/v3_var_1_2_mwc_ck_data_utility}}
 	\caption{Utility of data from combinations of two different experiments for identifying two parameters in $v_3$ when a) $K_3^{pep}$ is held constant, and when b) $K_3^{fdp}$ is held constant.
-Notice that a significant number of combinations can identify both $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is known and fixed.}%\label{fig:figure2}
+Notice that a significant number of combinations can identify both $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is known and fixed.
+c and d data is from convenience kinetics model.}%\label{fig:figure2}
 \end{figure}
@@
 \begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_ck_1_2_exp_info}}
+	\caption{The contribution of different experiments types used in a combination of three experiments ($j = \{1, 2, 3\}$) that can practically identify parameter a) $V_3^{max}(1)$, b) $K_3^{fdp}(1)$, c) $K_3^{pep}(1)$, d) $V_3^{max}(2)$, e) $K_3^{fdp}(2)$ and f) $K_3^{pep}(2)$ for $v_3$.
+parameters $V_3^{max}$, $K_3^{fdp}$, and $K_3^{pep}$ are locally identifiable and have two possible roots.
+The first set of roots are designated as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$, and the second set of roots are designated as $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
+The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.
+The contribution of the five different experiment types towards identifiability of either of the two roots for each of the three parameters is the same.
+Data from Convenience kinetics model.}
+\end{figure}
+
+\begin{figure}[!tbhp]
 	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_exp_info}}
 	\caption{Contribution of different types of experiments for using data from two different experiments towards identification of a) $V_3^{max}$ and b) $K_3^{fdp}$ in $v_3$ when $K_3^{pep}$ is held constant, and c) $V_3^{max}$ and d) $K_3^{pep}$ when $K_3^{fdp}$ is held constant.}%\label{fig:figure2}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_ck_exp_info}}
+	\caption{Contribution of different types of experiments for using data from two different experiments towards identification of a) $V_3^{max}$ and b) $K_3^{fdp}$ in $v_3$ when $K_3^{pep}$ is held constant, and c) $V_3^{max}$ and d) $K_3^{pep}$ when $K_3^{fdp}$ is held constant.
+Data from Convenience Kinetic model is used.}%\label{fig:figure2}
 \end{figure}

### 2018-04-23 ident/docs/draft_3.tex
3rd draft based on prof.m's comments on draft 2

--- 2018-04-16 ident/docs/draft_2.tex
@@
 
+	\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
+	A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism.
+The kinetic model is described below.
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
+		\caption{The previously published small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
+	\end{figure}
+
+	\begin{equation}\label{eq:ode1}
+	\frac{d}{dt}pep=v_1-v_2-v_4
+	\end{equation}
+	\begin{equation}\label{eq:ode2}
+	\frac{d}{dt}fdp=v_2-v_3
+	\end{equation}
+	\begin{equation}\label{eq:ode3}
+	\frac{d}{dt}E=v_5 - d E
+	\end{equation}
+	The kinetic expressions for fluxes $v_1$ through $v_5$ are given below.
+The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics.
+The acetate flux through $v_1$ is also governed by the quantity of available enzyme E.
+	\begin{equation}\label{eq:flux1}
+	v_1 = k_{1}^{cat}E\frac{ac}{ac+K_{1}^{ac}}
+	\end{equation}		
+	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}).
+Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression in Equation (\ref{eq:flux1}) for $v_1$ cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
+So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
+Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
+The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
+	\begin{align}\label{eq:flux1a}
+	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
+	\end{align}		
+	We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability. 	
+	\begin{equation}\label{eq:flux2}
+	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
+	\end{equation}
+	\begin{equation}\label{eq:flux3}
+	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
+	\end{equation}
+	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
+
+	The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (Section \ref{sec:ident}).
+We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica or SymPy in Python).
+In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
+The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
+	\begin{align}\label{eq:flux3_convkin}
+	v_3 = V_3^{max}\left(\frac{1}{1 + \frac{K_3^{pep}}{pep}}\right)\left(\frac{\frac{fdp}{K_3^{fdp}}}{1 + \frac{fdp}{K_3^{fdp}}}\right)
+	\end{align}
+
+	The flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
+	\begin{equation}\label{eq:flux4}
+	v_4 = V_{4}^{max}.pep
+	\end{equation}		
+	The production of enzyme E is represented by flux $v_5$.
+The inhibition of this flux by \textit{fdp} is modeled using Hill kinetics, where $K_e^{fdp}$ represents the Hill binding constant for the inhibiting metabolite \textit{fdp}, $n_e$ is the Hill exponent, and $V_e^{max}$ is the maximum reaction rate for $v_5$.
+	\begin{align}\label{eq:flux5}
+	v_5 = V_e^{max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right)
+	\end{align}
+
 	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}		
@@
 	\section{Results:}
-	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model} of the Supplementary Information.
-Identifiability results for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ are also discussed in this section.
-We show that the ability to establish practical identifiability of kinetic parameters in metabolic network models relies upon the nonlinearity of the kinetic rate law formulations used to describe the fluxes.
-We also show that the degree of identifiability of the maximum reaction rate parameter is always higher than the degree of identifiability of the corresponding enzyme binding parameter irrespective of enzyme kinetic rate law used to describe the flux.
-
-	Then, in section \ref{sec:experiment_type} we discuss the ability to determine the type of experiments for parameter identification for a given flux based on the informativeness of a given type of experiment.
-Here we show how the informativeness of a given type of experiment to identify a specific flux can be deduced from its contribution towards the practical identification of the parameters for a given flux.
-
-	%In the following section we demonstrate our methodology to establish practical identifiability for one of the fluxes in this network (section \ref{sec:example}) (move to results).
-
-	Then, in section \ref{sec:data_utility} we provide a motivation for the need for experimental design, especially to identify kinetic models of metabolic networks.
-We do this by looking at the number of combinations that can identify the maximum number of parameters.
-	In section \ref{sec:trends}
-
-	Finally, we discuss some results arising out of the use of our methodology to determine the identifiability of parameters when data with additive noise is used in section \ref{sec:noise}.
+	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model}.
+In Section \ref{sec:proof} that follows, we show how the demonstrated methodology is capable of identifying and estimating parameters for fluxes $v_1$, $v_2$, and $v_5$ using steady state flux values and metabolite concentrations.
+We also discuss the ability of the proposed methodology to determine the structural and practical non-identifiability of parameters modeling $v_3$ in this section.
+The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:experiment_type} that follows.
 
 	\subsection{Identifying parameters in kinetic models of metabolism: an example}\label{sec:example}	
-	In this section, we illustrate the proposed methodology step by step to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model} of the Supplementary Information).
-We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration.
-	\begin{align}\tag{S5}
-	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
-	\end{align}		
-	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in Section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
-So, from the $n_E = 21$ different experiments described in Section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$.
-We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively.
-Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
-	\begin{align*}%\label{eq:nonlin-flux2}
-	{v_1}^{(j)} = V_{1}^{max}\frac{ac^{(j)}}{ac^{(j)}+K_{1}^{ac}(ne)} &&  j=\{1, 2\}
-	\end{align*}
-
-	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
-These expressions have the form shown in Equation (\ref{eq:theta-eq}).
-	\begin{subequations}\label{eq:v1_par}
-		\begin{align}		
-		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
-		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
-		\end{align}
-	\end{subequations}
-	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations.
-Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
-	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is available, in addition to the measured concentration of the metabolites and fluxes.
-In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$.
-Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) above is substituted with $V_1^{max} = k_1^{cat}E$ instead.
-The globally identifiable parameter values for all fluxes modeled as Michaelis-Menten equations ($v_1$ and $v_2$) in provided in Supplementary Figure \ref{fig:parameter_value_proof}a and \ref{fig:parameter_value_proof}b.
-
-	We apply the method described in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, and illustrated above for $v_1$, to other fluxes of the network (Figure \ref{fig:network}) not described using Michaelis-Menten equations ($v_3$ and $v_5$).
-On applying this method to $v_3$, which is modeled using MWC kinetics, we find that we could not apply our methodology to determine closed form solution for all the parameters in $v_3$.
-However, when $v_3$ is instead modeled using Convenience kinetics, we find that the parameters are locally identifiable based on the definition we provide in Section \ref{sec:ident}.
-All three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ used to model $v_3$ have two possible roots.
-We still could not determine identifiable values for these parameters (Supplementary Figure \ref{fig:parameter_value_v3}) using the constraints on the physical relevance of their values (see Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}).
-In the box plot in Supplementary Figure \ref{fig:parameter_value_v3}a and b we show all the possible estimated parameter values for both roots of all three parameters modeling $v_3$, when experimental data is generated using the MWC model for $v_3$.
-Due to the numerous estimated parameter values, we can designate these parameters are practically non-identifiable.
-Although using the Convenience kinetics model for $v_3$ to generate the in silico experimental data leads to a reduction in the uncertainty of the estimated parameters (Supplementary Figure \ref{fig:parameter_value_v3}c and d), we still end up with non-unique parameter estimates.
-This leads is to conclude that under the current conditions, flux $v_3$ is structurally locally identifiable and practically non-identifiable.
-
-	However, $K_3^{fdp}$ and $V_3^{max}$ become globally identifiable when $K_3^{pep}$ is assumed to be known and fixed at a constant value, and $K_3^{pep}$ and $V_3^{max}$ become globally identifiable when $K_3^{fdp}$ is assumed to be known and fixed at a constant value (Supplementary Figure \ref{fig:parameter_value_v3_var}).
-
-	Through Supplementary Figure \ref{fig:parameter_value_v3_var}, we also want to bring attention to the fact that parameter non-identifiability remains when there is a mismatch between the kinetic rate law used to generate the in silico data, and the kinetic rate law used to identify parameters.
-For example, in the specific case of $v_3$, we see that globally structurally identifiable parameters $V_3^{max}$ and $K_3^{fdp}$ (when $K_3^{pep}$ is fixed) are practically non-identifiable when data generated from experiments using the MWC kinetics is used to model $v_3$ (Supplementary Figure \ref{fig:parameter_value_v3_var}a).
-The same is true of globally structurally identifiable parameters $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is fixed (Supplementary Figure \ref{fig:parameter_value_v3_var}b).
-Instead, when Convenience kinetics is used to generate in silico data, under the aforementioned conditions all the parameters are globally structurally and practically identifiable (Supplementary Figure \ref{fig:parameter_value_v3_var}c and d).
-
-	Similarly, we applied the demonstrated methodology to identify flux $v_5$ that describes a transcription/translation reaction using Hill kinetics, based on only the available metabolomics and the fluxomics information.
-When we tried identifying the three parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$, we could not obtain closed form solutions for these parameters using the computer algebra system.
-However, when we fixed the Hill coefficient $n_e$ as a known constant parameter, we were able to obtain closed form solutions for $V_e^{max}$ and $K_e^{fdp}$.
-Thus, similar to $v_3$ earlier, the dimension of the parameter space for $v_5$ had to be reduced from $\mathbb{R}^3$ to $\mathbb{R}^2$ to enable the application of the procedure that we have described in Section \ref{sec:ident}.
-In this instance, the identification required data from two different experiments.
-Again, just as we see for $v_3$, there are two different roots for $K_e^{fdp}$, making it locally structurally identifiable.
-On the other hand, $V_e^{max}$ has only one possible root, thereby it is a globally structurally identifiable parameter.
-However, only one of the two roots ($K_e^{fdp}(2)$) is practically identifiable based on the physical relevance criteria stated in Section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b.
-Thus, enforcing this criteria results in global practical identifiability of both $V_e{max}$ and $K_e^{fdp}$ in $v_5$ (Supplementary Figure \ref{fig:parameter_value_proof}c).
-
-	In the following section, we describe and summarize the results on the identifiability of parameters in these fluxes ($v_1$, $v_2$, $v_3$ and $v_5$) based on the expressions and the identifiable roots we discussed in this section.
-
+	\subsection{Relationship between structurally and practically parameter identifiability}\label{sec:proof}
 	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}			
-	In Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we show the number and percentage of experimental data combinations that are capable of identifying each parameter in each of the four fluxes $v_1$ (Figure \ref{fig:ident}a and e), $v_2$ (Figure \ref{fig:ident}b and f), $v_3$ (Figure \ref{fig:ident}c and g) and $v_5$ (Figure \ref{fig:ident}d and h) belonging to the network shown in Figure \ref{fig:network}.
-Based on the definition given in Section \ref{sec:degree_of_identifiability}, the percentages refer to the degree of identifiability of each parameter.
-We see from Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2} that the degree of identifiability of the maximum reaction rates ($V_i^{max}$), whenever they are identifiable for each of the four fluxes, is greater than or equal to the degree of identifiability of the corresponding enzyme binding ($K_i$) constants, or the allosteric activation constant ($K_3^{pep}$), in the respective reaction rate law models.
-We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition.
-The fact that in silico experimental data is generated using different kinetic rate law models for $v_3$ (MWC in Figure \ref{fig:ident}a-d vs Convenience kinetics in Figure \ref{fig:ident}e-h) has no bearing on this observation.
-We accordingly surmise that the maximum reaction rate parameters are always more identifiable (as indicated by their higher degree of identifiability) than their enzyme binding constant counterparts, irrespective of the enzyme kinetic rate law used to model the corresponding flux.
-
-	When data based on the MWC model for $v_3$ is used for identifying parameters, less that 50\% of all data combinations can identify $v_1$ (47.5\% and 40.0\% in Figure \ref{fig:ident}a).
-Even though the degree of identifiability of $K_1^{ac}(ne)$ is significantly different when data derived from the Convenience kinetics model for $v_3$ is used for identification (25\% in Figures \ref{fig:ident}e), the degree of identifiability is still less that 50\% for both parameters.
-In contrast, all available data combinations (from the MWC model and the Convenience kinetics model) satisfy the requirements for identifying $v_2$ (100\% in Figures \ref{fig:ident}b and f).
-
-	The difference in the degrees of identifiability between $v_1$ (Figure \ref{fig:ident}a) and $v_2$ (Figure \ref{fig:ident}b) can be attributed to the ability of data from different combinations of experiments to satisfy the conditions determined for practical identifiability of that parameter.
-To elaborate, we showed in Equation (\ref{eq:v1_par}) and Section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $V_1^{max}$ and $K_1^{ac}$ in $v_1$, the experiments must have distinct acetate concentrations as well as a different uptake flux $v_1$ between them.
-So, in this instance, recall that the ability of a combination of experiments to be useful is determined by the ability of the input change (changes in the input acetate concentration) to effect a change in the measured value of $v_1$.
-Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires the experiments to distinguish between values of both $v_2$ as well as \textit{pep}.
-We elaborate on the type and nature of experiments that are useful and informative for parameter identification in $v_1$ and $v_2$ later in Section \ref{sec:experiment_type}.
-
-	The difference in the degree of identifiability of $v_1$ that comes about from the use of experimental data from different rate laws used to model $v_3$ (Figures \ref{fig:ident}a and e) can be again attributed to the usefulness of experimental data.
-However, the difference in data usefulness between similar experiments arising due to differences in the kinetic rate law used for $v_3$ could be due to the difference in the dynamics of the network represented by either of the two models.
-The interaction between different components of the network (Supplementary Figure \ref{fig:network}) as exemplified by the different models plays an important role in determining the dynamic and the steady state response of the network.
-The difference in dynamics in turn contributes to different experimentally observed values of metabolite concentrations and fluxes (Supplementary Figure \ref{fig:experimental_data_dist}), which in turn changes the usefulness of the given combination of datasets, and its subsequent effect of parameter identifiability.
-
-	Nonetheless, we would like to note that these variations in the dynamics contributed by the difference between the rate laws can be resolved with more information, i.e., experimental proteomic data that provide protein concentrations for different experiments.
-We earlier saw that relaxing the assumption on the available experiment data, i.e., assuming the availability of protein concentration (\textit{E}), in addition to the metabolite concentrations (\textit{pep} and \textit{fdp}) and fluxes, results in removal of uncertainty and global identifiability of $v_1$ (Supplementary Figure \ref{fig:parameter_value_proof}).
-Recall that any predicted parameter value for both $k_1^{cat}$ and $K_1^{ac}$ matches with the exact known value irrespective of the model (MWC vs Convenience kinetics) from which experimental data is obtained when enzyme concentrations are included for identifiability (Supplementary Figure \ref{fig:parameter_value_proof}a).
-Additionally, we also see an improvement in the degree of identifiability of the corresponding parameters $k_1^{cat}$ and $K_1^{ac}$ (Supplementary Figure \ref{fig:ident_root2}a and b) when enzyme concentrations are available.
-
-	We would also like to point out that a difference in either the ability to estimate true parameter values, or the the identifiability of parameter modeling $v_2$ does not arise due to difference in the kinetic models used to obtain in silico experimental data.
-This could be because of the fact that any and all available information that is required for the identification of parameters in $v_2$ is encompassed in the steady state data used for identifiability analysis.
-
-	However, unlike $v_2$, we see that the degree of identifiability of both locally identifiable roots of parameters in $v_3$ changes with the change in the data (MWC vs Convenience kinetics) used for the analysis (Figures \ref{fig:ident}c and g and Supplementary Figures \ref{fig:ident_root2}b and e).
-Despite the difference in the dynamics represented by the different data sets, and its impact on the usefulness of data sets for identifiability purposes that was explained earlier, we could also allude to the structural non-identifiability between the parameters in $v_3$ for the impact that it would have on the practical identifiability of the parameters using in silico data from different sources.
-In order to ascertain this, we plot the degrees of identifiability of the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ under conditions when they are globally identifiable in Supplementary Figure \ref{fig:ident_v3_var}.
-To recall from Section \ref{sec:example}, $V_3^{max}$ and $K_3^{fdp}$ are globally identifiable when $K_3^{pep}$ is constant, and $V_3^{max}$ and $K_3^{pep}$ are globally identifiable when $K_3^{fdp}$ is known and fixed.
-Now that structurally non-identifiability between parameters has been alleviated, we see that the uncertainty introduced by the data obtained from the MWC model still results in a degree of identifiability of only about 5\% for $K_3^{fdp}$ (Supplementary Figure \ref{fig:ident_v3_var}a).
-However, when $K_3^{fdp}$ is held constant all MWC model-based experimental datasets contribute to local practical identifiability of parameters $V_3^{max}$ and $K_3^{pep}$ (Supplementary Figure \ref{fig:ident_v3_var}c), albeit with a bot of uncertainty (Supplementary Figures \ref{fig:ident_v3_var}b).
-This ties back with our earlier observation that these parameters are not globally practically identifiable from data sets obtained using the MWC model (Supplementary Figure \ref{fig:parameter_value_v3_var}a and b).
-However, we see from Supplementary Figures \ref{fig:ident_v3_var}c and d that the globally structurally identifiable parameters are also practically identifiable based on the data obtained from the Convenience kinetics model.
-This can be corroborated by the parameter values shown in Supplementary Figure \ref{fig:parameter_value_v3_var}c and d, where data from the Convenience kinetics model is used for analysis.
-
-	We did not observe any differences in the degree of identifiability of parameters in $v_5$ between the analyses that used data from either the MWC model or the Convenience kinetics model for $v_3$ (Figure \ref{fig:ident}d and h).
-The identifiability of parameters based on the first root obtained from the computer algebra system are shown in Supplementary Figure \ref{fig:ident_root2}c and f.
-We use a similar reasoning to that we suggested for identifiability of $v_2$ above to justify our results.
-
-	%In using the convenience kinetic rate law model for $v_3$ (Equation \ref{eq:flux3_convkin}) for identifying the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$, we found that each of the three parameters have two different closed-form expressions owing to the presence of a square root term in their solutions.
-Thus, as per the stated definition in section \ref{sec:ident}, the parameters for $v_3$, described using the convenience kinetics rate laws, are only locally identifiable due to the non-unique but finite number of possible solutions for the three parameters.
-The two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ are denoted by (1) and (2) following the respective parameter names throughout the rest of the document: $V_3^{max}(1)$, $K_3^{fdp}(1)$, and $K_3^{pep}(1)$ refer to the first set of solutions, and $V_3^{max}(2)$, $K_3^{fdp}(2)$, $K_3^{pep}(2)$ refer to the other set of solutions.
-
 	\subsection{Rationale for a priori experimental design through practical parameter identification}\label{sec:data_utility}	
-	%In section \ref{sec:example}, we described how the practical identifiability of parameters in a kinetic model of a metabolic network depended on the enzyme kinetic rate law used to model the fluxes that constitute the network.
-In this section we look at the rationale for a priori experimental design to collect meaningful data for parameter estimation.
-	%We also showed how the methodology that we propose can help in determining the nature and type of experiments that one needs to perform to identify and estimate parameters in kinetic models in section \ref{sec:experiment_type}.
-	\textcolor{red}{(should this section moved to SI?)}Although the need for better parameter estimation methods is well researched, within the kinetic modeling community (\textcolor{red}{include references related to Banga papers on global optimization methods for parameter estimation, Rudiyanto Gunawan, Lennart Ljung and Heijnen papers here}), the need for experimental design methods to satisfy the data needs for parameter estimation has not gained enough traction (\textcolor{red}{include Rudiyanto Gunawan papers, D-optimal design using FIM here}).
-Hence, we find the need to stress the necessity for experimental design methodologies, in this case, tailored specifically for metabolic network models.
-
-	We do this by showing how useful the multiple combinations of experimental data are towards identifying parameters for each flux $v_1$ (Figure \ref{fig:data_utility}a), $v_2$ (Figure \ref{fig:data_utility}b), $v_3$ (Figure \ref{fig:data_utility}c) and $v_5$ (Figure \ref{fig:data_utility}d).
-We express the usefulness of data from any given combination of experiments on the basis of the number of parameters that the data from the combination can identify.
-The higher the number of parameters a combination of experiments can identify, the greater is the usefulness of that combination of experiments.
-The percentage of all data combinations that can estimate a given number of parameters within a given flux is also shown in Figure \ref{fig:data_utility}.
-
-	For $v_1$ (Figure \ref{fig:data_utility}a), data from about 50\% of the 240 combinations of experiments cannot identify any parameter used to model the flux ($V_1^{max}$ and $K_1^{ac}(ne)$).
-In contrast, data from only about 5\% of the experiments is useless towards identifying any parameter in $v_3$, while there are no useless combinations of experiments from which data cannot be used to identify either $v_2$, or $v_5$.
-These data are also in agreement with the degree of identifiability observed for parameters in all these fluxes (Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}).
-Thus, when it is possible to obtain steady state data from numerous experiments to estimate parameters for a small metabolic network, there is a very high possibility of performing experiments that could provide non-informative data that can result in parameter non-identifiability without a priori experimental design.
-
-	We see that in instances where the identifiability of the fluxes differ based on the data set used (data from MWC model vs data from Convenience kinetics model), the utility of the data sets towards identifying parameters are changes.
-For instance, when data from the MWC model is used to identify parameters in $v_3$ (Figure \ref{fig:data_utility}c), less than 10\% of the data sets can identify all three parameters.
-Whereas, when data from the Convenience kinetics model is used for the same purpose, slightly more than 50\% of the combinations can identify all three parameters.
-Note that this observations should be considered in light of the structural local identifiability and the practical non-identifiability of these parameters that we mentioned earlier in Sections \ref{sec:example} and \ref{sec:ident_rate_laws}.
-
-	Although one could argue based on earlier observations that uncertainty in the model from which data is obtained, and the structural non-identifiability of some of the parameters in the models of the aforementioned fluxes that there may not be a need for experimental design if in vivo model uncertainty and structural non-identifiability can be resolved, one needs to only look at Supplementary Figure \ref{fig:other_data_utility}a and d.
-In these figures we show the utility of multiple combinations of metabolite and protein concentration, and flux data towards identifying $k_1^{cat}$ and $K_1^{ac}$ in $v_1$ when data from the MWC model and Convenience kinetics model are used, respectively.
-We see that about 45\% of the 240 combinations of data from 21 different experiments cannot identify either one of the two parameters, when these parameters are globally structurally and practically identifiable.
-
-	In addition to providing information on the fraction of experiments that are informative and those that are not informative, Figure \ref{fig:data_utility} and Supplementary Figure \ref{fig:other_data_utility} also informs the ease of identifying parameters for a single flux.
-In conjunction with the degrees of identifiability shown in Figure \ref{fig:ident} and Supplementary Figure \ref{fig:ident_root2}, we see that identifying $v_2$ is much easier than identifying $v_1$.
-Since minimization of resource utilization requires us to use the least number of experiments to identify both $v_1$ and $v_2$, this indicates the need to carefully choose combinations of experiments that can identify $v_1$.
-A careful consideration of experiments for identifying $v_2$ is however not a major concern since there is a very high probability that experiments chosen to identify $v_1$ will also identify $v_2$.
-A similar argument can also be for determining experiments that can identify $V_e^{max}$ and $K_e^{fdp}$ in $v_5$: since $V_e^{max}$ has an identifiability of 100\%, the choice of experiments to completely identify parameters in $v_5$ rests with experiments that can identify $K_e^{fdp}$.
-
-	Accordingly, if we can establish a hierarchy of identifiable parameters and the corresponding fluxes they model, this section gives credence to the idea that a priori experimental design is necessary, even when all possible in vivo measurements can be obtained, to get the most informative data for parameter identification and estimation at minimum cost.
-Additionally, having the ability to see the contribution of each type of experiment towards identifiability would prove sufficient to decide on the total number and type of experiments that need to be performed to identify all fluxes in a kinetic model of metabolism.
-So, next we look at the contribution of different types of experiments towards identifying parameters in each flux in the small network.
-
 	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}
-	In the previous sections, we showed how ability to structurally identify parameters in a kinetic model of metabolism is dependent on the kinetic rate law used to model a specific flux (Section \ref{sec:example}).
-We also discussed how the the informativeness of experiments used to practically identify parameters in kinetic models is not only dependent on the in silico rate law, but is also a function of the uncertainty in the experimental data brought about by the differences between the in silico rate law and the in vivo rate law from which data is obtained (Section \ref{sec:ident_rate_laws}).
-Subsequently, we also showed the need for a priori experimental design to generate informative data to identify parameters in kinetic models.
-
-	In systems identification terminology, all the aforementioned observations regarding data requirements for parameters identification can be tied to selecting experiments that are persistently excitable for the flux being identified.
-In systems identification, any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
-Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range.
-%Input signals that guarantee to fully excite the dynamics of the system are termed as persistently excitable signals.
-In linear systems, persistence of excitation of any input signal can be theoretically guaranteed.
-However, for nonlinear systems, like the metabolic network we deal with in this paper, persistence of excitation of input signals cannot be theoretically guaranteed, and should be assessed on a case by case basis.
-The lack of persistence of excitation requires the design of experiments to satisfy the data needs for complete system identification.
-
-	In the case of identifying fluxes within a metabolic network individually, each flux is not independent from the other fluxes that form the metabolic network.
-Each flux is connected to one or more fluxes through the metabolites that not only act as substrates and products, but also as allosteric or transcriptional activators and inhibitors (e.g., effect of \textit{pep} on $v_3$, or the effect of \textit{fdp} on $v_5$ and consequently on $v_1$ in Figure \ref{fig:network}).
-Hence, careful consideration of experiments is a necessity so that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}.
-So, in this section we discuss the dependence of the variation in the degree of identifiability for parameters of different fluxes from the point of view of the informativeness of experiments from which data is collected for identification.
-
-	We demonstrated in section \ref{sec:example}, using $v_1$ and $v_2$ as examples, that the ability of data from a given combination of experiments to identify parameters in a given flux depends on its ability to satisfy the conditions determined for practical identifiability of that parameter.
-Although we were able to determine the requirements for identifying $v_1$ and $v_2$ based on the closed form expressions of the parameters, this may not be possible when a complex nonlinear rate law is used to model a flux.
-Using the Convenience Kinetic rate law to model $v_3$ is a case in point.
-So, in order to determine the type of experiments that help satisfy identifiability conditions for parameters of any flux modeled by any usable rate law, we look at the distribution of the different types of experiments in all data combinations that can identify the parameters within a flux.
-Recall that we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Section \ref{sec:experiments} and Table \ref{tab:pval}).
-
-	The contribution from different types of experiments towards identification of parameters in different fluxes is given in Figure \ref{fig:exp_info}.
-In Figure \ref{fig:exp_info}a and \ref{fig:exp_info}b, the contribution from different experiment types for identifying $V_1^{max}$ and $K_1^{ac}(ne)$ are given, respectively.
-The contribution of different experiments towards identifiability for $V_2^{max}$ (Figure \ref{fig:exp_info}c) and $K_2^{pep}$ (Figure \ref{fig:exp_info}d) are also shown.
-Note that the contributions are only shown for data sets obtained using the MWC model in Figure \ref{fig:exp_info}.
-Experiment type contributions towards identifiability based on data sets obtained from the Convenience kinetics model for $v_3$ are shown separately in Supplementary Figure .
-
-	In agreement with what we discussed in Section \ref{sec:example}, i.e., experiments must be able distinguish between different values of $v_1$ and acetate, the contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations in comparison to the other types of experiments (Figure \ref{fig:exp_info}a and b).
-Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.
-
-	In contrast to $v_1$, we see that both the enzyme perturbation and the acetate perturbation experiments have the same contribution towards combinations that can identify parameters for $v_2$ (Figure \ref{fig:exp_info}c and d).
-Hence, in conjunction with the high degree of identifiability seen in Figure \ref{fig:ident}b for $v_2$, we can deduce that almost all experiment types are persistently excitable for $v_2$ i.e., most experiments can bring about noticeable changes to both \textit{pep} and $v_2$.
-Consequently, this confirms our earlier hypothesis that in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative enough to identify $v_2$.
-
-	Both the degree of identifiability of parameters and the informativeness of the corresponding experiments used to identify them can be explained by the position of the flux in the metabolic network.
-The position of any given flux in the metabolic network determines the specific experiment that is persistently excitable enough to identify the parameters of that flux.
-This dependency of experiment informativeness on the position of the flux can be further elucidated using $v_1$ and $v_2$ as examples.
-We know from Equation (\ref{eq:v1_par}) that identifiability of $v_1$ requires changes in both acetate and $v_1$.
-We also know, based on our knowledge of the Michaelis-Menten kinetic rate law that changes in the substrate concentration of a reaction can bring about a nonlinear change in the value of the corresponding reaction rate.
-In this specific metabolic network, since the substrate is an input variable to the model, and $v_1$ is the corresponding uptake flux, the substrate can be easily perturbed to create persistently excitable experiments to identify parameters in $v_1$.
-We can generalize this observation for the identification of all uptake fluxes in all metabolic networks, i.e., at a minimum, a change in the input substrate concentration may be necessary for an informative experiment to identify the uptake flux parameters.
-
-	On the other hand, the Michaelis-Menten model for $v_2$ also requires changes in \textit{pep} and $v_2$ for persistently excitable experiments to identify $v_2$.
-However, since both of these are system outputs, satisfaction of this condition cannot be guaranteed without an analysis of the dynamics of the metabolic network, and how changes in the input (acetate) bring about changes in the two requisite output quantities.
-Previous dynamical analysis of the network (Figure \ref{fig:network}) has already established the existence of a functional relationship between \textit{pep} and $v_2$, and the input acetate concentration and the levels of expression of the different enzymes within the network \parencite{Srinivasan2017}.
-Hence, it is theoretically possible for any of the five different experiment types to be persistently excitable to identify $v_2$.
-This is confirmed by the high degree of identifiability for both parameters of $v_2$ (Figure \ref{fig:ident}b), i.e., all data combinations can identify the parameters.
-Thus, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending on the position of the flux with respect to the inputs and the outputs of the metabolic network, as well as the various regulatory interactions present within the network.
-
-	We can extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_3$ as well (Supplementary Figure \ref{fig:v3_exp_info}).
-In this case, the top three panels (Supplementary Figure \ref{fig:v3_exp_info}a-c) show the distribution of different experiment types for identifying each parameter with respect to their first root identified as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$.
-The bottom panels (Supplementary Figure \ref{fig:v3_exp_info}d-f) show the corresponding distribution of experiments for the second root of each fo three parameters in $v_3$: $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
-
-	In all of the above scenarios for $v_1$ (Figure \ref{fig:exp_info}a and b), $v_2$ (Figure \ref{fig:exp_info}c and d) and $v_3$ (Supplementary Figure \ref{fig:v3_exp_info}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar.
-Hence, the green and blue surfaces are superimposed on top of each other.
-This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:exp_info}e).
-However, this is not the case with the distribution of experiments required for identifying $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:exp_info}f).
-We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$.
-Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figure \ref{fig:ident}d), the choice and number of experiments required to identify $v_5$ completely hinges upon identifiability of $K_e^{fdp}$ from the chosen experiments.
 
 	\subsection{Effect of noise on degree of identifiability is dependent on nonlinearity of enzyme kinetic rate law models}\label{sec:noise}
-	\textcolor{red}{(Should this section be moved to the SI as well?)}In reality, experimental data from biological systems is usually noisy.
-Noise in data gathered from experiments is associated with biological noise attributable to the stochasticity of cellular function, as well as measurement noise associated with the techniques used to obtain concentration and flux measurements.
-However, the previous sections describe practical identifiability for parameters obtained using noise-free in silico experimental data.
-So, in order to illustrate the ability of our method to test the practical identifiability of parameters under realistic circumstances, we apply our methodology on in silico data (concentration and flux) with 5\% additive noise.
-
-	50 different samples of noise with 0 mean and 5\% standard deviation are drawn from a normal distribution and added to the steady state experimental data generated on the basis of experiments described in Section \ref{sec:experiments} to generate 50 different samples of noisy data.
-Subsequently, all 50 samples are tested for their ability to practically identify all three fluxes of the small network (Figure \ref{fig:network}).
-The degree of identifiability of the different parameters using noisy experimental data are shown in the Appendix (Figure \ref{fig:figure1_noisy}).
-
-	We hypothesize that the nonlinearity of the enzyme kinetic rate law used to model a flux, and consequently the complexity of the closed form expression obtained for the various parameters used in the model has an impact on the ability of noisy data to practically identify the parameters.
-The difference in the identifiability attributable to the aforementioned factors can be seen in the difference in the degree of identifiability for the parameters of $v_1$, $v_2$ and $v_3$ in the small metabolic network that we use to demonstrate our methodology.
-
-	Despite the presence of noise, the degree of identifiability of $v_1$ and $v_2$ (Figure \ref{fig:figure1_noisy}a and b in the Appendix) does not change from the case where no noise is present in the experimental data (Figure \ref{fig:figure1}).
-However, the degree of identifiability for parameters of $v_3$ is affected by the presence of noise in the data (Figure \ref{fig:figure1_noisy}c in the Appendix).
-This effect is represented by the presence of a non-zero standard deviation and corresponding error bars on the graphs in the degrees of identifiability of parameters modeling $v_3$ (Figure \ref{fig:figure1_noisy}c in the Appendix).
-The standard deviation and the error bars represent the presence of data combinations that can identify a given parameter under certain values of noise, but not do so under different noise values.
-In the presence of 5\% additive noise in the data, the conditions under which parameters for $v_1$ and $v_2$ can be identified are strictly satisfied by the same data combinations that can identify these parameters in the absence of noise, i.e., the changes in the acetate concentration (for $v_1$) and \textit{pep} (for $v_2$) are significant enough to enable identification.
-The contribution of different experiment types towards identifiability also remains the same (figure not shown separately).
-
-	However, as stated in the previous section, the nonlinearity of the original kinetic rate law expression for $v_3$, and the complexity of the closed-form expression obtained for the parameters in the rate law model preclude us from testing the presence, and consequently the satisfaction of these conditions for parameters in $v_3$.
-Furthermore, the conditions for identifiability of $v_3$ parameters are not as simple as either requiring just a simple difference in \textit{fdp} or \textit{pep} or that of both concentrations.
-Thus, while under some noise values these identifiability conditions may be satisfied, it may not be possible to satisfy these conditions under other noise values.
-Accordingly, we end up with a distribution of experimental data combinations that can identify $v_3$ parameters depending on noise levels in the data.
-It is significant to note that the variability in the degree of identifiability of parameters is small (< 2\%) when noise is present.
-
-	These observations, in conjunction with the estimation of the degree of identifiability of the different parameters (Figure \ref{fig:figure1}) help in determining not only the nature of experiments required for parameter identifiability (Figure \ref{fig:figure2}), but also the order in which the type of experiments necessary for identifying the entire metabolic network needs to be evaluated.
-
-	For instance, in the case of the example network that we use in this paper, given the high degree of identifiability for $v_2$ and large spread in the frequency of different experiment types that can contribute to identifiability for $v_2$, the focus should be on first choosing experiments for identifying $v_1$ followed by selecting experiments for identifying $v_3$.
-It is possible that the three or at most four experiments chosen to identify parameters of both $v_1$ and $v_3$ would suffice to identify $v_2$ as well without the need to perform additional experiments.
 

### 2018-04-24 ident/docs/sifigures.tex
move model from SI methods to main methods

@@
 \section{Supplementary Methods}
-\subsection{Kinetic model of gluconeogenesis in \textit{E. coli}}\label{sec:small-model}
-A previously proposed kinetic model \parencite{Kotte2014, Srinivasan2017} for acetate consumption through gluconeogenesis (Figure \ref{fig:network}) is used as a case study to illustrate identifiability analysis for experimental design for parameter estimation in kinetic models of metabolism.
-The kinetic model is described below.
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=.3\textwidth,height=.6\textheight,keepaspectratio]{figures/figure5/Figure1_NetworkA}}
-	\caption{The previously published small metabolic network for gluconeogenesis used to demonstrate our practical identifiability method for kinetic models of metabolism.}\label{fig:network}
-\end{figure}
-
-\begin{equation}\label{eq:ode1}
-\frac{d}{dt}pep=v_1-v_2-v_4
-\end{equation}
-\begin{equation}\label{eq:ode2}
-\frac{d}{dt}fdp=v_2-v_3
-\end{equation}
-\begin{equation}\label{eq:ode3}
-\frac{d}{dt}E=v_5 - d E
-\end{equation}
-The kinetic expressions for fluxes $v_1$ through $v_5$ are given below.
-The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics.
-The acetate flux through $v_1$ is also governed by the quantity of available enzyme E.
-\begin{equation}\label{eq:flux1}
-v_1 = k_{1}^{cat}E\frac{ac}{ac+K_{1}^{ac}}
-\end{equation}		
-The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}).
-Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression in Equation (\ref{eq:flux1}) for $v_1$ cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
-So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
-Consequently $k_1^{cat}$ is replaced by $V_1^{max}$ as a parameter to describe $v_1$.
-The corresponding enzyme binding constant is denoted as $K_1^{ac} (ne)$ to distinguish it from the enzyme binding constant calculated in the presence of measured enzyme concentration data.
-\begin{align}\label{eq:flux1a}
-v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
-\end{align}		
-We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) to demonstrate our method for practical identifiability. 	
-\begin{equation}\label{eq:flux2}
-v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
-\end{equation}
-\begin{equation}\label{eq:flux3}
-v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
-\end{equation}
-The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$.
-
-The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (Section \ref{sec:ident}).
-We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica or SymPy in Python).
-In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
-The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
-\begin{align}\label{eq:flux3_convkin}
-v_3 = V_3^{max}\left(\frac{1}{1 + \frac{K_3^{pep}}{pep}}\right)\left(\frac{\frac{fdp}{K_3^{fdp}}}{1 + \frac{fdp}{K_3^{fdp}}}\right)
-\end{align}
-
-The flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
-\begin{equation}\label{eq:flux4}
-v_4 = k_{4}^{cat}.pep
-\end{equation}		
-The production of enzyme E is represented by flux $v_5$.
-This flux is inhibited by \textit{fdp} and modeled using Hill kinetics, where $K_e^{fdp}$ represents the Hill binding constant for the inhibiting metabolite \textit{fdp}, $n_e$ is the Hill exponent, and $V_e^{max}$ is the maximum reaction rate for $v_5$.
-\begin{align}\label{eq:flux5}
-v_5 = V_e^{max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right)
-\end{align}
-
 \subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}

### 2018-04-24 ident/docs/draft_3.tex
add figure for 3.1

@@
 
-	%In kinetic models of metabolic networks (Equation \ref{eq:kinstoich}), the fluxes are expressed as a function of the metabolite concentrations $x$ and the kinetic parameters $\theta$ (Figure \ref{fig:ident-flowchart}a).
-The value of every flux $v_i$, $i={1, 2, ..., n_r}$, is expressed using one of the many available enzyme kinetic formulations (Equation \ref{eq:output}b).
-Without loss of generality, all of these kinetic formulations can be expressed as nonlinear algebraic equations (Figure \ref{fig:ident-flowchart}a).
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.9\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
+		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
+a) The steps for the construction of a kinetic model of a metabolic network.
+The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
+The identifiability of parameters for each flux can be established independently.
+b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
+	\end{figure}
 
@@
 	\section{Results:}
-	First, in section \ref{sec:example}, we demonstrate the use of the methodology that we described in section \ref{sec:ident} to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in section \ref{sec:small-model}.
-In Section \ref{sec:proof} that follows, we show how the demonstrated methodology is capable of identifying and estimating parameters for fluxes $v_1$, $v_2$, and $v_5$ using steady state flux values and metabolite concentrations.
-We also discuss the ability of the proposed methodology to determine the structural and practical non-identifiability of parameters modeling $v_3$ in this section.
-The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:experiment_type} that follows.
+	First, in Section \ref{sec:example}, we demonstrate the use of the methodology that we described in section 2.1 to practically identify parameters in flux v1 of the small gluconeogenic network (Figure \ref{fig:network}) model given in Section \ref{sec:small-model}.
+In Section 3.2 that follows, we show how the demonstrated methodology is capable of identifying and estimating parameters for fluxes $v_1$, $v_2$ and $v_5$ using steady state flux values and metabolite concentrations.
+We discuss the ability of the proposed methodology to determine the structural and practical non-identifiability of parameters modeling $v_3$ in Section \ref{sec:proof}.
+The contribution of the uncertainty in the data (due to differences between in vivo and in vitro kinetics) towards identifying parameters in enzyme kinetic models is discussed in Section \ref{sec:uncertainty}.
+The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:design} that follows.
 
 	\subsection{Identifying parameters in kinetic models of metabolism: an example}\label{sec:example}	
+	In this section, we illustrate the proposed methodology step by step to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model} of the Supplementary Information).
+We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration.
+	\begin{align}\tag{S5}
+	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
+	\end{align}		
+	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in Section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
+So, from the $n_E = 21$ different experiments described in Section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$.
+We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively.
+Then, the nonlinear algebraic equations shown in Equation (\ref{eq:nonlineq}) can be formulated for $v_1$ as:
+	\begin{align*}%\label{eq:nonlin-flux2}
+	{v_1}^{(j)} = V_{1}^{max}\frac{ac^{(j)}}{ac^{(j)}+K_{1}^{ac}(ne)} &&  j=\{1, 2\}
+	\end{align*}
+
+	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
+These expressions have the form shown in Equation (\ref{eq:theta-eq}).
+	\begin{subequations}\label{eq:v1_par}
+		\begin{align}		
+		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
+		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
+		\end{align}
+	\end{subequations}
+	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations.
+Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
+
+	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is also available, in addition to the measured concentration of the metabolites and fluxes.
+In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$.
+Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) above is substituted with $V_1^{max} = k_1^{cat}E$ instead.
+
+	\subsection{Section Title}\label{sec:initial_analysis}
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=.8\textwidth,height=.6\textheight,keepaspectratio]{figures/figure6/v1_k1cat_v2_v5_2_v3_ck_1_parameter_values}}
+		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, c) $v_5$, and d)$v_3$.
+For $v_1$, we have assumed that enzyme concentration is available and have accordingly identified and estimated $k_1^{cat}$, as opposed to $V_1^{max}$.
+The parameter values for only the second root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(2)$) is shown, since $K_e^{fdp}(1)$ is not estimated by any combination of two experiments, and $V_e^{max}$ is estimated by all combinations.
+Only one of the two roots for $v_3$ is shown in panel d.
+The estimated data for the second root has a similar distribution to that of the first root and is shown in the Supplementary Information.
+Data is generated using the Convenience Kinetic model for allosteric regulation for $v_3$.}\label{fig:ident_values}
+	\end{figure}
 	\subsection{Relationship between structurally and practically parameter identifiability}\label{sec:proof}
-	\subsection{Degree of identifiability of different enzyme kinetic rate laws}\label{sec:ident_rate_laws}			
-	\subsection{Rationale for a priori experimental design through practical parameter identification}\label{sec:data_utility}	
-	\subsection{Experimental requirements for identifying parameters depend on the position of the flux in the metabolic network}\label{sec:experiment_type}
-
-	\subsection{Effect of noise on degree of identifiability is dependent on nonlinearity of enzyme kinetic rate law models}\label{sec:noise}
+	This section will deal with structural and practical identifiability of parameters in v1, v2, v3 and v5.
+This section will only focus on the identifiability and not issues related to experimental design.
+Focus will be on v3 and v5 (which have non-identifiability issues).
+Also, we could use a better name for this section.
+	\subsection{Parameter non-identifiability due to uncertainty in Experimental Data}\label{sec:uncertainty}
+	This will cover issues related to the model used for v3. Convenience kientic model vs MWC model for data generation. Should this uncertainty also include noise?			
+	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
+	Final section talking about distribution of experiments to obtain informative data.
 
@@
 	\section{Tables, Figures and Figure Captions:}
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.9\textwidth,height=.6\textheight,keepaspectratio]{figures/figure3/ident_analysis}}
-		\caption{A flow diagram showing the methodology developed to establish practical identifiability of parameters in kinetic models of metabolism.
-a) The steps for the construction of a kinetic model of a metabolic network.
-The choice of rate law formulations to describe metabolic fluxes influences the identification methodology.
-The identifiability of parameters for each flux can be established independently.
-b) The steps for practical identifiability analysis for parameters of a single flux.}\label{fig:ident-flowchart}
-	\end{figure}
 

### 2018-04-24 ident/docs/draft_3.tex
complete section 3.1

@@
 
-	\subsection{Structural and practical identifiability of parameters in kinetic models}	
+	\subsection{Structural and practical identifiability of parameters in kinetic models}\label{sec:ident_def}
 	In the Introduction, we briefly metioned that the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
@@
 
-	\section{Results:}
-	First, in Section \ref{sec:example}, we demonstrate the use of the methodology that we described in section 2.1 to practically identify parameters in flux v1 of the small gluconeogenic network (Figure \ref{fig:network}) model given in Section \ref{sec:small-model}.
-In Section 3.2 that follows, we show how the demonstrated methodology is capable of identifying and estimating parameters for fluxes $v_1$, $v_2$ and $v_5$ using steady state flux values and metabolite concentrations.
-We discuss the ability of the proposed methodology to determine the structural and practical non-identifiability of parameters modeling $v_3$ in Section \ref{sec:proof}.
+	\section{Results}
+	First, in Section \ref{sec:example}, we demonstrate the use of the methodology that we described in Section 2.1 to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in Section \ref{sec:small-model}.
+We discuss the ability of the proposed methodology to determine the structural and practical non-identifiability of parameters modeling $v_1$, $v_3$ and $v_5$ in Section \ref{sec:proof}.
+	In Section \ref{sec:initial_analysis} that follows, we show how the demonstrated methodology is capable of identifying and estimating parameters for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ using steady state flux values and metabolite concentrations.
 	The contribution of the uncertainty in the data (due to differences between in vivo and in vitro kinetics) towards identifying parameters in enzyme kinetic models is discussed in Section \ref{sec:uncertainty}.
@@
 	\end{subequations}
-	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations.
+	To test the practical identifiability of the parameters in Equation \ref{eq:v1_par}, we substitute any suitable in silico experimental data and determine the value of the denominator of the right hand side expression.
 Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
+The parameter values that are obtained for $V_1^{max}$ and $K_1^{ac} (ne)$ by substituting in silico steady state experimental data are shown in Supplementary Figure \ref{fig:v1_v1max_ck_values}.
 
-	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is also available, in addition to the measured concentration of the metabolites and fluxes.
+	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is also available, in addition to the measured metabolite concentrations and fluxes.
 In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$.
-Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) above is substituted with $V_1^{max} = k_1^{cat}E$ instead.
+Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) is substituted with $V_1^{max} = k_1^{cat}E$ instead.
+The corresponding identifiability expressions for $k_1^{cat}$ and $K_1^{ac}$ are given in Equation (\ref{eq:v1_k1cat}).
+	\begin{subequations}\label{eq:v1_k1cat}
+		\begin{align}
+		k_1^{cat} = \frac{ v_1^{(1)}v_1^{(2)} \left( ac^{(1)}- ac^{(2)}\right)}{v_1^{(2)}ac^{(1)}E^{(1)} - v_1^{(1)}ac^{(2)}E^{(2)}}\\
+		K_1^{ac} = \frac{ac^{(1)}ac^{(2)}\left(v_1^{(1)}E^{(2)} - v_1^{(2)}E^{(1)}\right)}{v_1^{(2)}ac^{(1)}E^{(1)} - v_1^{(1)}ac^{(2)}E^{(2)}}
+		\end{align}
+	\end{subequations}	
+	We show the parameter value for $k_1^{cat}$ and $K_1^{ac}$ that are obtained through the practical identifiability analysis in Figure \ref{fig:ident_values}a when in silico experimental data is substituted in Equation (\ref{eq:v1_k1cat}).
+Through Equation (\ref{eq:v1_k1cat}) and Figure \ref{fig:ident_values}a we are able to show that the uncertainty in the experimental flux values can be resolved through the incorporation of the available enzyme concentrations to get good estimations for the enzyme kinetic parameters.
+Thus, having more experimental information can help resolve practical identifiability.
 
-	\subsection{Section Title}\label{sec:initial_analysis}
 	\begin{figure}[!tbhp]
@@
 	\end{figure}
-	\subsection{Relationship between structurally and practically parameter identifiability}\label{sec:proof}
+
+	In the following section we present results from the identifiability analysis of fluxes $v_2$, $v_3$ and $v_5$ in the small metabolic network (Figure \ref{fig:network}), using the methodology (Figure \ref{fig:ident-flowchart}) that we have demonstrated above for $v_1$.
+
+	\subsection{Relationship between structural and practical parameter identifiability}\label{sec:proof}
+	As mentioned in Section \ref{sec:ident_def}, by definition, it should be possible to estimate unique parameter values based on the model structure and available experimental data for any structurally and practically identifiable parameter.
+For the proposed methodology (Figure \ref{fig:ident-flowchart}) to work, it should be possible to obtain closed form solutions for each parameter in the enzyme kinetic model for each flux as shown in Equation (\ref{eq:theta-eq}).
+Since the ability to obtain closed-form solutions for each parameter is dependent on the model structure, any parameter that has non-unique closed-form solutions can be called a structurally non-identifiable parameter.
+However, if the number of solutions that the parameter has are finite, then the parameter is only locally structurally identifiable.
+
+	We demonstrated the identifiability of parameters modeling $v_1$ in Section \ref{sec:example}.
+We have shown that the parameters have only one unique closed-form solution, and accordingly are structurally identifiable.
+Since $v_2$ is also expressed using the Michaelis-Menten model, just like $v_1$, we find that the parameters ($V_2^{max}$ and $K_2^{pep}$) are also structurally identifiable.
+The closed-form expressions for these parameters are similar to the ones shown in Equation \ref{eq:v1_par}, with \textit{ac} replaced by \textit{pep}, and $v_1$ replaced by $v_2$.
+
+	As shown in Figure \ref{fig:ident-flowchart} and illustrated for $v_1$ in Section \ref{sec:example}, to determine the practical identifiability of parameters we test for the existence of a non-zero denominator of the closed-form expressions of the parameters.
+Additionally, we also check for the physiological feasibility of the parameter values (Figure \ref{fig:ident-flowchart}b) that are obtained through this analysis.
+If the resulting parameter values obtained from unique closed-form expressions and various combinations of experimental data are also unique, then the parameter is both structurally and practically identifiable.
+Accordingly, both $v_1$ and $v_2$ are not only structurally identifiable due to the presence of unique closed-form expressions for their parameters, they are also practically identifiable because the parameters in the respective models possess unique values based on distinct combinations of experimental data (Figure \ref{fig:ident_values}a and b).
+
+	However, we find that the parameters used to model $v_3$ using the Convenience kinetics rate law, and $v_5$ using the Hill kinetic rate law are not structural identifiable under certain conditions.
+First, for $v_3$, we find that the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different closed-form solutions.
+Thus, based on the presence of non-unique but finite number of possible solutions for these parameters we can classify $v_3$ as a locally structurally identifiable flux.
+
+	We would like to point out that we used the Convenience kinetics rate law model for $v_3$ due to the inability of the computer algebra system to generate closed-form expressions for parameters in the Monod-Wyman-Chageaux model (Section \ref{sec:small-model}).
+However, we find that the Convenience kinetics model is not structurally identifiable in its current form.
+In order to alleviate this issue, we reduced the dimension of the parameter space for $v_3$.
+Originally, $\theta \in \mathbb{R}^3$ for $v_3$.
+By reducing the dimension of $\theta$ to $\mathbb{R}^2$, we were able to obtain a structurally identifiable model for $v_3$.
+To reduce the dimension of the parameter space for $v_3$, we fix either $K_3^{fdp}$ or $K_3^{pep}$ as a known quantity, and identify the other unfixed parameter along with $V_3^{max}$.
+This results in unique closed-form expressions for both $V_3^{max}$ and the other unfixed parameter ($K_3^{pep}$ or $K_3^{fdp}$).
+
+	$v_5$ describes a transcription/translation reaction using Hill kinetics.
+We apply our proposed methodology to identify parameters modeling $v_5$ using only the available experimental data on the metabolite concentrations and the fluxes within the metabolic network.
+We could not obtain closed form solutions for parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$ using the computer algebra system.
+So, as illustrated earlier for $v_3$, we resorted to reducing the dimension of the parameter space by fixing one of the three parameters, the Hill coefficient $n_e$.
+With a fixed and known $n_e$, $K_e^{fdp}$ has two possible closed-form solutions, making it a locally structurally identifiable parameter.
+On the other hand, $V_e^{max}$ has only one unique closed-form solution, and therefore is structurally identifiable.
+
 	This section will deal with structural and practical identifiability of parameters in v1, v2, v3 and v5.
@@
 Also, we could use a better name for this section.
+
+	\subsection{Section Title}\label{sec:initial_analysis}
+
 	\subsection{Parameter non-identifiability due to uncertainty in Experimental Data}\label{sec:uncertainty}

### 2018-04-24 ident/docs/draft_3.tex
complete sections 3.1, 3.2 and 3.3

@@
 	First, in Section \ref{sec:example}, we demonstrate the use of the methodology that we described in Section 2.1 to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in Section \ref{sec:small-model}.
-We discuss the ability of the proposed methodology to determine the structural and practical non-identifiability of parameters modeling $v_1$, $v_3$ and $v_5$ in Section \ref{sec:proof}.
-	In Section \ref{sec:initial_analysis} that follows, we show how the demonstrated methodology is capable of identifying and estimating parameters for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ using steady state flux values and metabolite concentrations.
+	We discuss the ability of the proposed methodology to determine the structural identifiability of parameters modeling $v_1$, $v_3$ and $v_5$ in Section \ref{sec:proof}. 	
+	In Section \ref{sec:initial_analysis} that follows, we show how the demonstrated methodology is capable of practically identifying and estimating parameters for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ using steady state flux values and metabolite concentrations.
+	The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:design} that follows.
 	The contribution of the uncertainty in the data (due to differences between in vivo and in vitro kinetics) towards identifying parameters in enzyme kinetic models is discussed in Section \ref{sec:uncertainty}.
-The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:design} that follows.
 
@@
 The parameter values that are obtained for $V_1^{max}$ and $K_1^{ac} (ne)$ by substituting in silico steady state experimental data are shown in Supplementary Figure \ref{fig:v1_v1max_ck_values}.
+Due to the numerous possible parameter values seen in Supplementary Figure \ref{fig:v1_v1max_ck_values}, we can conclude that both $V_1^{max}$ and $K_1^{ac}$ are practically non-identifiable.
 
@@
 	We show the parameter value for $k_1^{cat}$ and $K_1^{ac}$ that are obtained through the practical identifiability analysis in Figure \ref{fig:ident_values}a when in silico experimental data is substituted in Equation (\ref{eq:v1_k1cat}).
-Through Equation (\ref{eq:v1_k1cat}) and Figure \ref{fig:ident_values}a we are able to show that the uncertainty in the experimental flux values can be resolved through the incorporation of the available enzyme concentrations to get good estimations for the enzyme kinetic parameters.
+Through Equation (\ref{eq:v1_k1cat}) and Figure \ref{fig:ident_values}a we are able to show that the uncertainty in the parameter estimates (Supplementary Figure \ref{fig:v1_v1max_ck_values}) can be resolved through the incorporation of the available enzyme concentrations.
 Thus, having more experimental information can help resolve practical identifiability.
@@
 
-	\subsection{Relationship between structural and practical parameter identifiability}\label{sec:proof}
-	As mentioned in Section \ref{sec:ident_def}, by definition, it should be possible to estimate unique parameter values based on the model structure and available experimental data for any structurally and practically identifiable parameter.
-For the proposed methodology (Figure \ref{fig:ident-flowchart}) to work, it should be possible to obtain closed form solutions for each parameter in the enzyme kinetic model for each flux as shown in Equation (\ref{eq:theta-eq}).
+	\subsection{Establishing Structural identifiability of parameters based on closed-form solutions}\label{sec:proof}
+	As mentioned in Section \ref{sec:ident_def}, by definition, unique parameter values based on the model structure are possible for any structurally identifiable parameter.
+Also, for the proposed methodology (Figure \ref{fig:ident-flowchart}) to work, it should be possible to obtain closed form solutions for each parameter in the enzyme kinetic model for each flux as shown in Equation (\ref{eq:theta-eq}).
 Since the ability to obtain closed-form solutions for each parameter is dependent on the model structure, any parameter that has non-unique closed-form solutions can be called a structurally non-identifiable parameter.
@@
 
-	We demonstrated the identifiability of parameters modeling $v_1$ in Section \ref{sec:example}.
+	We demonstrated the structural identifiability of parameters modeling $v_1$ in Section \ref{sec:example}.
 We have shown that the parameters have only one unique closed-form solution, and accordingly are structurally identifiable.
@@
 The closed-form expressions for these parameters are similar to the ones shown in Equation \ref{eq:v1_par}, with \textit{ac} replaced by \textit{pep}, and $v_1$ replaced by $v_2$.
-
-	As shown in Figure \ref{fig:ident-flowchart} and illustrated for $v_1$ in Section \ref{sec:example}, to determine the practical identifiability of parameters we test for the existence of a non-zero denominator of the closed-form expressions of the parameters.
-Additionally, we also check for the physiological feasibility of the parameter values (Figure \ref{fig:ident-flowchart}b) that are obtained through this analysis.
-If the resulting parameter values obtained from unique closed-form expressions and various combinations of experimental data are also unique, then the parameter is both structurally and practically identifiable.
-Accordingly, both $v_1$ and $v_2$ are not only structurally identifiable due to the presence of unique closed-form expressions for their parameters, they are also practically identifiable because the parameters in the respective models possess unique values based on distinct combinations of experimental data (Figure \ref{fig:ident_values}a and b).
 
@@
 
-	$v_5$ describes a transcription/translation reaction using Hill kinetics.
+	While $v_3$ is an allosterically regulated metabolic flux, $v_5$ describes a transcription/translation reaction using Hill kinetics.
 We apply our proposed methodology to identify parameters modeling $v_5$ using only the available experimental data on the metabolite concentrations and the fluxes within the metabolic network.
 We could not obtain closed form solutions for parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$ using the computer algebra system.
-So, as illustrated earlier for $v_3$, we resorted to reducing the dimension of the parameter space by fixing one of the three parameters, the Hill coefficient $n_e$.
+So, instead of changing the model as we did for $v_3$, we resorted to reducing the dimension of the parameter space by fixing one of the three parameters, the Hill coefficient $n_e$.
+We illustrated this for reducing the parameter space of the Convenience kinetic model for $v_3$ earlier.
 With a fixed and known $n_e$, $K_e^{fdp}$ has two possible closed-form solutions, making it a locally structurally identifiable parameter.
@@
 
-	This section will deal with structural and practical identifiability of parameters in v1, v2, v3 and v5.
-This section will only focus on the identifiability and not issues related to experimental design.
-Focus will be on v3 and v5 (which have non-identifiability issues).
-Also, we could use a better name for this section.
+	With this we have established conditions for structural identifiability of all the major fluxes in the small metabolic network (Figure \ref{fig:network}).
+We have shown how our proposed methodology can be used to establish conditions for structural identifiability using steady state information on the model variables.
+We next discuss the practical identifiability of the parameters in $v_2$, $v_3$ and $v_5$ whose parameters are structurally identifiable only under certain conditions.
 
-	\subsection{Section Title}\label{sec:initial_analysis}
+	\subsection{Relationship between structural and practical parameter identifiability}\label{sec:initial_analysis}
+	Together with the definition for structural identifiability we presented in Section \ref{sec:ident_def}, we also introduced the concept of practical parameter identifiability.
+To recall, we mentioned that  it should be possible to estimate unique parameter values based on all available experimental data for any practically identifiable parameter.
+As shown in Figure \ref{fig:ident-flowchart} and illustrated for $v_1$ in Section \ref{sec:example}, to determine the practical identifiability of parameters we test for the existence of a non-zero denominator of the closed-form expressions of the parameters.
+We also reduce the possible space within which a parameter could be practically identifiable by checking for the physiological feasibility of the parameter values that are obtained through this analysis (Figure \ref{fig:ident-flowchart}b).
+If the resulting parameter values obtained from various combinations of experimental data for each closed-form expression are unique, then the parameter is practically identifiable.
+However, if a non-unique number of parameter values are possible from multiple combinations of experimental steady state data, then the parameter is said to be practically non-identifiable.
+In conjunction with the conditions for structural identifiability demonstrated earlier in Section \ref{sec:proof}, if the parameter has only one unique closed-form expression, and its value is also unique, then the parameter is both structurally and practically identifiable.
+If either of these conditions are not satisfied, the parameters can be either locally structurally or practically identifiable or non-identifiable.
+
+	Accordingly, both $v_1$ and $v_2$ are not only structurally identifiable due to the presence of unique closed-form expressions for their parameters, they are also practically identifiable because the parameters in the respective models possess unique values based on distinct combinations of experimental data (Figure \ref{fig:ident_values}a and b).
+
+	Regarding $v_5$, we showed earlier in Section \ref{sec:proof} that the identifiability of $v_5$ can be analyzed only when the Hill coefficient $n_e$ is held constant.
+So, in subsequent discussions, the dimension of the $v_5$ parameter space is kept at $\mathbb{R}^2$ by fixing the value of $n_e$.
+Under these conditions, we find that the structurally identifiable parameter $V_e^{max}$ is also practically identifiable, i.e., it has only one unique value based on all available in silico experimental data (Figure \ref{fig:ident_values}c).
+However, recall that unlike $V_e^{max}$, $K_e^{fdp}$ is only locally structurally identifiable as $K_e^{fdp}$ has two possible closed-form expressions.
+Nonetheless, despite its local structural identifiability, we find that the $K_e^{fdp}$ is also practically identifiable, like $V_e^{max}$, with only one unique parameter value (Figure \ref{fig:ident_values}c).
+
+	We find that the practical identifiability of $v_5$, despite the local structural identifiability of one of its parameters, is due to the enforcement of the physiological relevance criteria on the parameters i.e., only one of the two closed-form expressions for $K_e^{fdp}$ is physiologically relevant.
+The other solution always acquires a negative value that has to physiological meaning.
+Thus, by reducing the practically identifiable space of parameters, we have shown that our methodology can establish global practical identifiability even when the parameters are only locally structurally identifiable.
+
+	Similar to $K_e^{fdp}$ in $v_5$, we also explained the local structural identifiability of $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ modeling $v_3$ in Section \ref{sec:proof}.
+These parameters have two possible closed-form expressions.
+In Figure \ref{fig:ident_values}d we show that the numerical values for one of the two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$.
+The numerical values for the other root of the three parameters is presented in Supplementary Figure, and they have a similar distribution.
+Based on the prior definition we discussed for practically identifiable parameters, the numerous possible values that the three parameters can acquire (Figure \ref{fig:ident_values}d) lead us to conclude that the parameters in $v_3$ are practically non-identifiable when they are only structurally locally identifiable.
+Note that these parameters are practically non-identifiable even after the reduction in the practically identifiable parameter space through application of the physiological relevance constraint (Figure \ref{fig:ident-flowchart}b).
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=0.8\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v3_var_1_2_ck_parameter_value}}
+		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$.
+The globally identifiable parameter values of a)  $V_3^{max}$ and $K_3^{fdp}$ when $K_3^{pep}$ is held constant, and b) $V_3^{max}$ and $K_3^{pep}$ when $K_3^{fdp}$ is held constant.}\label{fig:v3_var_ck_values}
+	\end{figure}
+
+	So, the observation of practical non-identifiability for $v_3$ in the presence of local structural identifiability is in contrast to what we observed for $v_5$ earlier.
+However, we find that $v_3$ is practically identifiable when its parameters are also structurally identifiable (Figure \ref{fig:v3_var_ck_values}).
+Earlier in Section \ref{sec:proof} we had mentioned that $V_3^{max}$ and $K_3^{fdp}$ are structurally identifiable only when $K_3^{pep}$ is fixed, and $V_3^{max}$ and $K_3^{pep}$ are structurally identifiable when $K_3^{fdp}$ is fixed.
+Under these scenarios we find the structurally identifiable parameters to also be practically identifiable (Figure \ref{fig:v3_var_ck_values}).
+
+	Thus, in conjunction with the practical identifiability of $v_5$ established earlier, we see that it is possible to delineate between structural and practical identifiability of parameters in kinetic models of metabolism only under certain conditions, and not in others.
+
+	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
+	Final section talking about distribution of experiments to obtain informative data.
+
+	If the denominator of the closed-form expression is zero, then the experimental data set concerned is said to be incapable of practically identifying the said parameter, and the data from that combination of experiments is considered non-informative.
+
+	In addition, we also reduce the possible space within which a parameter could be practically identifiable by checking for the physiological feasibility of the parameter values that are obtained through this analysis (Figure \ref{fig:ident-flowchart}b).
+If the resulting parameter values obtained from various combinations of experimental data for each closed-form expression are not physiologically feasible, then the experimental data set concerned is said to be incapable of practically identifying the said parameter, and the data from that combination of experiments is considered non-informative.
 
 	\subsection{Parameter non-identifiability due to uncertainty in Experimental Data}\label{sec:uncertainty}
+
 	This will cover issues related to the model used for v3. Convenience kientic model vs MWC model for data generation. Should this uncertainty also include noise?
-	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
-	Final section talking about distribution of experiments to obtain informative data.
 

### 2018-04-25 ident/docs/draft_3.tex
new figure for 3.5

@@
 
-	Thus, in conjunction with the practical identifiability of $v_5$ established earlier, we see that it is possible to delineate between structural and practical identifiability of parameters in kinetic models of metabolism only under certain conditions, and not in others.
+	In conjunction with the practical identifiability of $v_5$ established earlier, we see that it is possible to delineate between structural and practical identifiability of parameters in kinetic models of metabolism only under certain conditions, and not in others.
 
 	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
+	The analysis of parameter practical identifiability can be used to gather information on the type of experiments that can provide useful data for parameter estimation.
+For instance, during practical identifiability analysis, if the denominator of the closed-form expression is zero, then the experimental data set concerned is said to be incapable of practically identifying the said parameter, and the data from that combination of experiments is considered non-informative.
+We can also determine whether a particular combination of experiments is informative or not based on the physiological feasibility of parameter values obtained from various combinations of experimental data for each closed-form expression (Figure \ref{fig:ident-flowchart}b).
+If parameter values are not physiologically feasible, then the experimental data set concerned is said to be incapable of practically identifying the said parameter, and the data from that combination of experiments is considered non-informative.
+
+	Based on the aforementioned criteria, we can determine the number of different experimental data combinations that can practically identify each parameter in each flux.
+This is shown in Figure \ref{fig:ident}.
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v2_v5_2_ck_ident_exp_info}}
+		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_5$ when there is no noise in the input experimental data.
+The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, and 421 for $v_5$) that can identify each parameter is also specified.
+$v_1$, $v_2$ and $v_5$ require data from two experiments for analysis.
+The contribution of different experiment type towards identifying each parameter is shown in the spider plots.}\label{fig:ident}
+	\end{figure}
+
 	Final section talking about distribution of experiments to obtain informative data.
-
-	If the denominator of the closed-form expression is zero, then the experimental data set concerned is said to be incapable of practically identifying the said parameter, and the data from that combination of experiments is considered non-informative.
-
-	In addition, we also reduce the possible space within which a parameter could be practically identifiable by checking for the physiological feasibility of the parameter values that are obtained through this analysis (Figure \ref{fig:ident-flowchart}b).
-If the resulting parameter values obtained from various combinations of experimental data for each closed-form expression are not physiologically feasible, then the experimental data set concerned is said to be incapable of practically identifying the said parameter, and the data from that combination of experiments is considered non-informative.
 
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.7\textheight,keepaspectratio]{figures/figure1/v1_V1max_v2_v3_1_v5_mwc_ck_ident}}
-		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) and e) $v_1$, b)  and f) $v_2$, c) and g) $v_3$, and d) and h) $v_5$ when there is no noise in the input experimental data.
-The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, 3360 for $v_3$ and 421 for $v_5$) that can identify each parameter is also specified.
-$v_1$, $v_2$ and $v_5$ require data from two experiments for analysis, and $v_3$ requires data from three experiments.
-Results for only one of the two possible roots are shown for $v_3$ and $K_e^{fdp}$ in $v_5$.
-The results obtained using data from the MWC model for $v_3$ are presented in the panels on the left hand side column and the panels on the right hand column show results obtained using data derived from the Convenience kinetics model for $v_3$.}\label{fig:ident}
-	\end{figure}
-
-	\begin{figure}[!tbhp]
 		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_mwc_ck_data_utility}}

### 2018-04-25 ident/docs/sifigures.tex
new si figure 1
add new figures and remove old figures

@@
 \begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_ck_parameter_values}}
+	\caption{}\label{fig:v1_v1max_ck_values}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v1_v1max_ck_ident_exp_info}}
+	\caption{}\label{fig:v1_v1max_ck_ident}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_ck_1_ident}}
+	\caption{}\label{fig:v3_1_ck_ident}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_var_1_2_ck_ident}}
+	\caption{}\label{fig:v3_var_1_2_ck_ident}
+\end{figure}
+
+\begin{figure}[!tbhp]
 	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/mwc_ck_experimental_data}}
@@
 The Convenience Kientic rate law model is used for $v_3$ to generate in silico c) metabolite and enzyme concentration, and d) flux data through enzyme expression parameter and substrate concentration perturbations.}\label{fig:experimental_data_dist}
-\end{figure}
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_k1cat_v2_v5_2_parameter_values}}
-	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, and c) $v_5$.
-For $v_1$, we have assumed that enzyme concentration is available and have accordingly identified and estimated $k_1^{cat}$, as opposed to $V_1^{max}$.
-The parameter values for only the second root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(2)$) is shown, since $K_e^{fdp}(1)$ is not estimated by any combination of two experiments, and $V_e^{max}$ is estimated by all combinations.
-Data is generated using the Monod-Wyman-Changeaux model for allosteric regulation for $v_3$.}\label{fig:parameter_value_proof}
 \end{figure}

### 2018-04-26 ident/docs/draft_3.tex
complete 3.4. Only 3.5 uncertainty section remains

@@
 We illustrated this for reducing the parameter space of the Convenience kinetic model for $v_3$ earlier.
-With a fixed and known $n_e$, $K_e^{fdp}$ has two possible closed-form solutions, making it a locally structurally identifiable parameter.
+With a fixed and known $n_e$, $K_e^{fdp}$ has two possible closed-form solutions, which make it a locally structurally identifiable parameter.
 On the other hand, $V_e^{max}$ has only one unique closed-form solution, and therefore is structurally identifiable.
 
-	With this we have established conditions for structural identifiability of all the major fluxes in the small metabolic network (Figure \ref{fig:network}).
+	We have now established conditions for structural identifiability of all the major fluxes in the small metabolic network (Figure \ref{fig:network}).
 We have shown how our proposed methodology can be used to establish conditions for structural identifiability using steady state information on the model variables.
@@
 Under these conditions, we find that the structurally identifiable parameter $V_e^{max}$ is also practically identifiable, i.e., it has only one unique value based on all available in silico experimental data (Figure \ref{fig:ident_values}c).
-However, recall that unlike $V_e^{max}$, $K_e^{fdp}$ is only locally structurally identifiable as $K_e^{fdp}$ has two possible closed-form expressions.
+However, recall that unlike $V_e^{max}$, $K_e^{fdp}$ is only locally structurally identifiable as it has two possible closed-form expressions.
 Nonetheless, despite its local structural identifiability, we find that the $K_e^{fdp}$ is also practically identifiable, like $V_e^{max}$, with only one unique parameter value (Figure \ref{fig:ident_values}c).
@@
 	We find that the practical identifiability of $v_5$, despite the local structural identifiability of one of its parameters, is due to the enforcement of the physiological relevance criteria on the parameters i.e., only one of the two closed-form expressions for $K_e^{fdp}$ is physiologically relevant.
-The other solution always acquires a negative value that has to physiological meaning.
+The other solution always acquires a negative value that has no physiological meaning.
 Thus, by reducing the practically identifiable space of parameters, we have shown that our methodology can establish global practical identifiability even when the parameters are only locally structurally identifiable.
@@
 These parameters have two possible closed-form expressions.
-In Figure \ref{fig:ident_values}d we show that the numerical values for one of the two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$.
-The numerical values for the other root of the three parameters is presented in Supplementary Figure, and they have a similar distribution.
-Based on the prior definition we discussed for practically identifiable parameters, the numerous possible values that the three parameters can acquire (Figure \ref{fig:ident_values}d) lead us to conclude that the parameters in $v_3$ are practically non-identifiable when they are only structurally locally identifiable.
-Note that these parameters are practically non-identifiable even after the reduction in the practically identifiable parameter space through application of the physiological relevance constraint (Figure \ref{fig:ident-flowchart}b).
+In Figure \ref{fig:ident_values}d we show the numerical values for one of the two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$.
+The numerical values for the second closed-form expressions of the three parameters is presented in Supplementary Figure, and they also have a similar distribution.
+Based on the prior definition for practically identifiable parameters, the numerous possible values that the three parameters can acquire (Figure \ref{fig:ident_values}d) lead us to conclude that the parameters in $v_3$ are practically non-identifiable when they are only structurally locally identifiable.
+Also, these parameters are practically non-identifiable even after the reduction in the practically identifiable parameter space through application of the physiological relevance condition for practical identifiability (Figure \ref{fig:ident-flowchart}b).
 
@@
 	The analysis of parameter practical identifiability can be used to gather information on the type of experiments that can provide useful data for parameter estimation.
-For instance, during practical identifiability analysis, if the denominator of the closed-form expression is zero, then the experimental data set concerned is said to be incapable of practically identifying the said parameter, and the data from that combination of experiments is considered non-informative.
-We can also determine whether a particular combination of experiments is informative or not based on the physiological feasibility of parameter values obtained from various combinations of experimental data for each closed-form expression (Figure \ref{fig:ident-flowchart}b).
-If parameter values are not physiologically feasible, then the experimental data set concerned is said to be incapable of practically identifying the said parameter, and the data from that combination of experiments is considered non-informative.
-
-	Based on the aforementioned criteria, we can determine the number of different experimental data combinations that can practically identify each parameter in each flux.
-This is shown in Figure \ref{fig:ident}.
+For instance, during practical identifiability analysis, if either the denominator of the closed-form expression is zero, or if the parameter values that are obtained are not physiologically feasible (Figure \ref{fig:ident-flowchart}b), then the experimental data set concerned is said to be incapable of practically identifying the said parameter.
+Consequently, the data from that combination of experiments is considered non-informative.
+When this analysis is repeated for multiple combinations of steady state data from the 21 different in silico experiments, we can determine the number of different experimental data combinations that can practically identify each parameter in each flux (Figure \ref{fig:ident}).
 
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v2_v5_2_ck_ident_exp_info}}
-		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, c) $v_5$ when there is no noise in the input experimental data.
+		\centering{\includegraphics[width=0.6\textwidth,height=0.4\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v2_v5_2_ck_ident}}
+		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, and c) $v_5$ when there is no noise in the input experimental data.
 The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, and 421 for $v_5$) that can identify each parameter is also specified.
@@
 
-	Final section talking about distribution of experiments to obtain informative data.
+	As described in Section \ref{sec:degree_of_identifiability}, the information on the number of experimental data sets that can practically identify each parameter can be used to determine the degree of identifiability of the corresponding parameters.
+Subsequently, this information can be used to classify parameters based on their ease of identifiability.
+
+	In Figure \ref{fig:ident} we show the number and percentage of experimental data combinations that are capable of identifying each parameter in flux $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b) and $v_5$ (Figure \ref{fig:ident}c).
+The number and percentage of combinations identifying $V_1^{max}$ and $K_1^{ac}(ne)$ in $v_1$, and the closed-form solutions of the three parameters of $v_3$ are shown in Supplementary Figures \ref{fig:v1_v1max_ck_ident} and \ref{fig:v3_1_ck_ident}, respectively.
+The degree of identifiability of each parameter is also given in these figures as percentages.
+
+	Based on their degrees of identifiability, we see that the maximum reaction rates ($V_i^{max}$) are more or similarly identifiable in comparison to the corresponding enzyme binding ($K_i$) constants or the activation/inhibition constants, in the respective reaction rate law models (Figure \ref{fig:ident}, Supplementary Figures \ref{fig:v1_v1max_ck_ident} and \ref{fig:v3_1_ck_ident}).
+We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition.
+
+	It is important to recall that both $v_1$ (Supplementary Figure \ref{fig:v1_v1max_ck_values}) and $v_3$ (Figure \ref{fig:ident_values}d) are not practically identifiable (Section \ref{sec:initial_analysis}).
+While $v_1$ becomes practically identifiable (Figure \ref{fig:ident_values}a) if enzyme concentrations are utilized to alleviate uncertainties in the enzyme turn over rates ($k_1^{cat}$), $v_3$ becomes practically identifiable only when the dimension of the parameter space is reduced (Figure \ref{fig:v3_var_ck_values}).
+In these cases the degree of identifiability (Supplementary Figure \ref{fig:v1_v1max_ck_ident} for $v_1$ and Supplementary Figure \ref{fig:v3_1_ck_ident} for $v_3$) refers to the number of experimental data sets that can determine physiologically relevant values for the corresponding parameters.
+So, we have shown for $v_1$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $v_3$ (Supplementary Figure \ref{fig:v3_1_ck_ident}) that a greater number of experimental data sets can predict physiologically relevant values for $V_1^{max}$ and $V_3^{max}$ than for $K_1^{ac}(ne)$ and $K_3^{fdp}$ or $K_3^{pep}$, respectively.
+
+	The degree of identifiability of $v_3$, when its parameters are structurally and practically identifiable (Sections \ref{sec:proof} and \ref{sec:initial_analysis}) are shown in Supplementary Figure \ref{fig:v3_var_1_2_ck_ident}.
+Accordingly, we find that in all cases other than $V_1^{max}$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $k_1^{cat}$ in $v_1$ (Figure \ref{fig:ident}a), all data sets used to test practical identifiability can determine unique values for parameters when the corresponding parameter is structurally identifiable.
+
+	We can attribute the difference in the degree of identifiability between $v_1$ (Figure \ref{fig:ident}a and Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and the other fluxes ($v_2$, $v_3$ and $v_5$) to the ability of data from different combinations of experiments to satisfy the conditions for practical identifiability of that parameter, that can be determined a priori.
+In systems identification terminology, data requirements for parameters identification can be tied to selecting experiments that are persistently excitable for the flux being identified.
+Any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
+Only information obtained from such changes in the input can be used to completely identify the system over its entire dynamic range.
+So, the ability of data from a combination of different experiments to practically identify parameters of a given flux is governed by the ability of the experiment to generate distinct measured concentrations and fluxes that will satisfy the identifiability conditions.
+
+	In turn, the degree of identifiability of parameters and the informativeness of the corresponding experiments used to identify them can be explained by the position of the flux in the metabolic network.
+The position of any given flux in the metabolic network determines the specific experiment that is persistently excitable enough to identify the parameters of that flux.
+This dependency can be further elucidated using $v_1$ and $v_2$ as examples.
+
+	We know from Equation (\ref{eq:v1_k1cat}) and Section \ref{sec:example} that for a combination of any two experiments to be capable of identifying $v_1$, the experiments must generate data that have distinct acetate concentrations, \textit{E} and $v_1$.
+We also know, based on our knowledge of the Michaelis-Menten kinetic rate law that changes in the substrate concentration of a reaction can bring about a nonlinear change in the value of the corresponding reaction rate.
+So, in this instance, since the substrate is an input variable to the model, and $v_1$ is the corresponding uptake flux and \textit{E} is a system variable, the substrate can be easily perturbed to create persistently excitable experiments to identify parameters in $v_1$.
+We see the consequence of this requirement in the degree of identifiability of $k_1^{cat}$ and $K_1^{ac}$ (Figure \ref{fig:ident}a).
+We can generalize this observation for the identification of all uptake fluxes in all metabolic networks, i.e., at a minimum, a change in the input substrate concentration may be necessary for an informative experiment to identify the uptake flux parameters.
+
+	Similarly, the identification of parameters for $v_2$ (Figure \ref{fig:ident}b) requires that persistently excitable experiments distinguish between values of both $v_2$ as well as \textit{pep}.
+However, since both of these are system outputs, satisfaction of this condition cannot be guaranteed without an analysis of the dynamics of the metabolic network, and how changes in the input (acetate) bring about changes in the two requisite output quantities.
+Previous dynamical analysis of the network (Figure \ref{fig:network}) has already established the existence of a functional relationship between \textit{pep} and $v_2$, and the input acetate concentration and the levels of expression of the different enzymes within the network \parencite{Srinivasan2017}.
+The 100\% degree of identifiability seen for $v_2$ (Figure \ref{fig:ident}b) confirms the theoretical possibility for any type of perturbation experiment to be persistently excitable to identify $v_2$.
+Overall, this analysis informs us that the degree of identifiability and consequently, the type of experiments needed to identify different parameters varies widely depending on the position of the flux with respect to the inputs and the outputs of the metabolic network, as well as the various regulatory interactions present within the network (e.g., effect of \textit{pep} on $v_3$, or the effect of \textit{fdp} on $v_5$ and consequently on $v_1$ in Figure \ref{fig:network}).
+
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure2/v1_k1cat_v2_v5_2_ck_exp_info}}
+		\caption{Frequency of each type of perturbation experiment in data sets that can identify parameters in fluxes a) $v_1$, b) $v_2$ and c)$v_5$.
+The frequencies are represented as a percentage of the total number of experiments present within all the data sets that can practically identify the parameters.
+The frequency of experiments that occur as the first experiment in a combination of two experiments are shown as a blue surface, and the frequency of experiments that occur as the second experiment in the combination are shown as a green surface.}\label{fig:expinfo}
+	\end{figure}
+
+	From the above example we can summarize that identification of individual fluxes within a metabolic network necessitates a careful consideration of experiments such that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}.
+To facilitate the design of experiments based on their ability to satisfy requirements for practical identifiability of parameters, we determine the occurrence of each type of steady state perturbation experiment within combinations that can practically identify each parameter (Figure \ref{fig:expinfo}, Supplementary Figures \ref{fig:v1_v1max_ck_ident},  \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).
+So, with our proposed methodology it is possible to identify the types of perturbation experiments that would be informative for identifying each parameter in each flux with steady state concentration and flux data.
+
+	As mentioned in Supplementary Section \ref{sec:experiments}, we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Table \ref{tab:pval}).
+In Figures \ref{fig:expinfo}a, \ref{fig:expinfo}b and \ref{fig:expinfo}c, the contribution from different experiment types for identifying parameters in $v_1$, $v_2$ and $v_5$ are respectively shown as spider plots.
+
+	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations for $v_1$ in comparison to the other types of experiments (Figure \ref{fig:expinfo}a and Supplementary Figure \ref{fig:v1_v1max_ck_ident}).
+This is in agreement with the condition for identifiability that we discussed earlier (Equations \ref{eq:v1_k1cat} and \ref{eq:v1_par}).
+Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we also say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.
+
+	With $v_2$, we see that the enzyme perturbations as well as the acetate perturbation experiments have similar contributions towards datasets that can identify $v_2$ (Figure \ref{fig:expinfo}b).
+This is also supports our arguments made earlier with regards to the identifiability conditions for $v_2$ and the reasons for the difference in the type of experiments that are informative between $v_1$ and $v_2$.
+Accordingly, we find that in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative to identify $v_2$.
+
+	We can also extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_5$ (Figure \ref{fig:expinfo}c), or determining physiologically relevant parameter values for structurally locally identifiable parameters of $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).
+
+	In all of the above scenarios for $v_1$ (Figure \ref{fig:expinfo}a), $v_2$ (Figure \ref{fig:expinfo}b) and $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar.
+Hence, the green/blue/yellow surfaces in the spider plots are superimposed upon each other.
+This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:expinfo}c).
+However, this is not the case for $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:expinfo}c).
+We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$.
+Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figures \ref{fig:ident}c and \ref{fig:expinfo}c), the choice and number of experiments required to identify $v_5$ completely hinges upon the identifiability of $K_e^{fdp}$ from the chosen experiments.
 
@@
 
-	\clearpage
-	\section{Tables, Figures and Figure Captions:}
+	%\clearpage
+	%\section{Tables, Figures and Figure Captions:}
 
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight,keepaspectratio]{figures/figure3/experimental_design}}
-		\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
-	\end{figure}
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_v2_v5_2_exp_info}}
+	%	\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$, d) $K_2^{pep}$, e) $V_e^{max}$ and f) $K_e^{fdp}(2)$.
+The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:exp_info}
+	%\end{figure}
 
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_mwc_ck_data_utility}}
-		\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$.
-The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, c) three and d) two experiments.
-The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, 3360 combinations are used to practically identify c) $v_3$, and 420 combinations are used to analyze the identifiability of d) $v_5$.
-Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
-	\end{figure}
+	%\clearpage
 
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_v2_v5_2_exp_info}}
-		\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$, d) $K_2^{pep}$, e) $V_e^{max}$ and f) $K_e^{fdp}(2)$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:exp_info}
-	\end{figure}
-
-	\clearpage
-
-	\section{Old Figures, Figure Captions and other miscellaneous text}	
-	Methods and tools for structural identification of parameters based on differential algebra \parencite{Ljung1994, Audoly2001, Bellu2007} and profile likelihood \parencite{Raue2014} are available.
+	%\section{Old Figures, Figure Captions and other miscellaneous text}	
+	%Methods and tools for structural identification of parameters based on differential algebra \parencite{Ljung1994, Audoly2001, Bellu2007} and profile likelihood \parencite{Raue2014} are available.
 

### 2018-04-26 ident/docs/sifigures.tex
remove unnecessary figures and combine all useful figures

@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_ck_parameter_values}}
+	\centering{\includegraphics[width=0.6\textwidth,height=0.2\textheight, keepaspectratio]{figures/figure6/v1_v1max_ck_parameter_values}}
 	\caption{}\label{fig:v1_v1max_ck_values}
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_ck_1_ident}}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure1/v3_ck_1_2_ident_exp_info}}
 	\caption{}\label{fig:v3_1_ck_ident}
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_var_1_2_ck_ident}}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_var_1_2_ck_ident_exp_info}}
 	\caption{}\label{fig:v3_var_1_2_ck_ident}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=0.6\textwidth,height=0.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
+	\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_mwc_ck_data_utility}}
+	\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$.
+The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, c) three and d) two experiments.
+The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
+A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, 3360 combinations are used to practically identify c) $v_3$, and 420 combinations are used to analyze the identifiability of d) $v_5$.
+Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
 \end{figure}
@@
 
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight, keepaspectratio]{figures/figure1/v1_k1cat_v3_2_v5_1_ident}}
-	\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) and d) $v_1$, b) and e) $v_3$, and c) and f) $v_5$ when there is no noise in the input experimental data.
+%\begin{figure}[!tbhp]
+	%\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight, keepaspectratio]{figures/figure1/v1_k1cat_v3_2_v5_1_ident}}
+	%\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) and d) $v_1$, b) and e) $v_3$, and c) and f) $v_5$ when there is no noise in the input experimental data.
 The percentage of total combinations of experimental data used for analysis (240 for $v_1$ and $v_2$, 3360 for $v_3$ and 421 for $v_5$) that can identify each parameter is also specified.
@@
 The results obtained using data from the MWC model for $v_3$ are presented in the panels on the left hand side column and the panels on the right hand column show results obtained using data derived from the Convenience kinetics model for $v_3$.}\label{fig:ident_root2}
-\end{figure}
+%\end{figure}
 
@@
 Data from MWC model.}%\label{fig:figure2}
-\end{figure}
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure2/v1_V1max_v2_v5_2_ck_exp_info}}
-	\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_e^{max}$ and b) $K_e^{fdp}$ in $v_5$.
-Since $V_e^{max}$ is globally identifiable, it has a unique root.
-Since $K_e^{fdp}$ has two roots, is locally identifiable.
-However, results for the only practically identifiable root, based on the physical relevance constraints used in the analysis, is shown.
-Data from Convenience kinetics model is used.}%\label{fig:v5_exp_info}
-\end{figure}
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_1_2_exp_info}}
-	\caption{The contribution of different experiments types used in a combination of three experiments ($j = \{1, 2, 3\}$) that can practically identify parameter a) $V_3^{max}(1)$, b) $K_3^{fdp}(1)$, c) $K_3^{pep}(1)$, d) $V_3^{max}(2)$, e) $K_3^{fdp}(2)$ and f) $K_3^{pep}(2)$ for $v_3$.
-parameters $V_3^{max}$, $K_3^{fdp}$, and $K_3^{pep}$ are locally identifiable and have two possible roots.
-The first set of roots are designated as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$, and the second set of roots are designated as $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.
-The contribution of the five different experiment types towards identifiability of either of the two roots for each of the three parameters is the same.}\label{fig:v3_exp_info}
-\end{figure}
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure2/v3_ck_1_2_exp_info}}
-	\caption{The contribution of different experiments types used in a combination of three experiments ($j = \{1, 2, 3\}$) that can practically identify parameter a) $V_3^{max}(1)$, b) $K_3^{fdp}(1)$, c) $K_3^{pep}(1)$, d) $V_3^{max}(2)$, e) $K_3^{fdp}(2)$ and f) $K_3^{pep}(2)$ for $v_3$.
-parameters $V_3^{max}$, $K_3^{fdp}$, and $K_3^{pep}$ are locally identifiable and have two possible roots.
-The first set of roots are designated as $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$, and the second set of roots are designated as $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.
-The contribution of the five different experiment types towards identifiability of either of the two roots for each of the three parameters is the same.
-Data from Convenience kinetics model.}
-\end{figure}
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_exp_info}}
-	\caption{Contribution of different types of experiments for using data from two different experiments towards identification of a) $V_3^{max}$ and b) $K_3^{fdp}$ in $v_3$ when $K_3^{pep}$ is held constant, and c) $V_3^{max}$ and d) $K_3^{pep}$ when $K_3^{fdp}$ is held constant.}%\label{fig:figure2}
-\end{figure}
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v3_var_1_2_ck_exp_info}}
-	\caption{Contribution of different types of experiments for using data from two different experiments towards identification of a) $V_3^{max}$ and b) $K_3^{fdp}$ in $v_3$ when $K_3^{pep}$ is held constant, and c) $V_3^{max}$ and d) $K_3^{pep}$ when $K_3^{fdp}$ is held constant.
-Data from Convenience Kinetic model is used.}%\label{fig:figure2}
 \end{figure}

### 2018-04-26 ident/python2/ss-ident/v3_ck_noise_script.py
pass data to be plotted as indices as opposed to end point integer

@@ -74,7 +74,7 @@ parameter_identifibaility_plot(max_parameter_v3_root1, noise=1)
 # identifiable data combinations for each parameter
 parameter_experiment_info_spider(experiment_info_v3_root1, noise=1)
 # plot true parameter values and determined parameter values
-plot_parameter_values(true_value_v3_root1, noise=1, data_sets_to_plot=5000)
+plot_parameter_values(true_value_v3_root1, noise=1, data_sets_to_plot=range(0, 2000))
 # plot utility of data sets (number of data sets identifying n, n-1, n-2, ...., 1, 0 parameters
 data_utility_plot(data_list_v3_root1, noise=1)
 
@@ -98,7 +98,7 @@ parameter_identifibaility_plot(max_parameter_v3_root2, noise=1)
 # identifiable data combinations for each parameter
 parameter_experiment_info_spider(experiment_info_v3_root2, noise=1)
 # plot true parameter values and determined parameter values
-plot_parameter_values(true_value_v3_root2, noise=1, data_sets_to_plot=500)
+plot_parameter_values(true_value_v3_root2, noise=1, data_sets_to_plot=range(0, 2000))
 # plot utility of data sets (number of data sets identifying n, n-1, n-2, ...., 1, 0 parameters
 data_utility_plot(data_list_v3_root2, noise=1)
 

### 2018-04-26 ident/python2/ss-ident/generate_expdata.py
use noise standard deviation as user defined/default input arg for data generation

@@ -5,13 +5,14 @@ from generate_noisy_data import run_noisy_parameter_perturbation
 from plot_ident_results import plot_dynamic_sim_concentrations
 
 
-def initialize_to_ss(y0, cvode_options, ode_parameter_values, noise=1, kinetics=1):
+def initialize_to_ss(y0, cvode_options, ode_parameter_values, noise=1, kinetics=1, noise_std=0.05):
     """initialize system to ss from any y0 with given options and system parameters"""
     if noise:
         # get initial noisy system steady state
         initial_options = (cvode_options, ode_parameter_values)
         # use non-noisy initial ss (wt) data
-        noisy_initial_ss, noisy_initial_dyn, _, _ = generate_noisy_data(y0, initial_options, kinetics=kinetics)
+        noisy_initial_ss, noisy_initial_dyn, _, _ = generate_noisy_data(y0, initial_options, kinetics=kinetics,
+                                                                        noise_std=noise_std)
         return noisy_initial_ss, noisy_initial_dyn
     else:
         # get initial noisy system steady state
@@ -21,14 +22,15 @@ def initialize_to_ss(y0, cvode_options, ode_parameter_values, noise=1, kinetics=
 
 
 def perturb_parameters(initial_ss, parameter_perturbations, cvode_options, ode_parameter_values,
-                       number_of_samples=1, noise=1, kinetics=1, dynamic_plot=0):
+                       number_of_samples=1, noise=1, kinetics=1, dynamic_plot=0, noise_std=0.05):
     """perform parameter perturbations from given initial ss"""
 
     if noise:
         perturbation_options = {'ode_parameters': ode_parameter_values, 'cvode_options': cvode_options}
         noisy_ss, noisy_dynamic, perturbation_details, _, dynamic_info = \
             run_noisy_parameter_perturbation(parameter_perturbations, initial_ss["y"][0], perturbation_options,
-                                             kinetics=kinetics, number_of_samples=number_of_samples)
+                                             kinetics=kinetics, number_of_samples=number_of_samples,
+                                             noise_std=noise_std)
         return noisy_ss, perturbation_details
     else:
         perturbation_options = {'ode_parameters': ode_parameter_values, 'cvode_options': cvode_options}
@@ -39,11 +41,12 @@ def perturb_parameters(initial_ss, parameter_perturbations, cvode_options, ode_p
 
 
[…]

### 2018-04-26 ident/python2/ss-ident/v3_ck_noise_script.py

@@ -31,7 +31,7 @@ ode_parameter_values = {"K1ac": np.array([.1]),
 # get experimental system steady state data without noise
 exp_xss, exp_fss, exp_ssid, perturbation_details = \
     generate_expdata(y0, cvode_options, ode_parameter_values, noise=1, number_of_samples=50, kinetics=2,
-                     dynamic_plot=1, perturbation_plot=0)
+                     dynamic_plot=1, perturbation_plot=0, noise_std=0.05)
 
 # arrange experimental data to form multiple data sets
 exp_flux_index = np.array([0, 3, 2, 4, 1, 5])

### 2018-04-26 ident/docs/draft_3.tex
make intro and methods corrections

@@
 
-	The development of these enzyme kinetic rate laws is based on in vitro observations of enzyme activity.
-Accordingly, authors have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions, as opposed to in vitro conditions \parencite{Heijnen2005,Heijnen2013}.
-This problem is further compounded by the fact that typical parameter values used in kinetic models of metabolism are estimated in vitro, and may not be applicable under in vivo conditions \parencite{Heijnen2005,Smallbone2007}.
-Also, parameters estimated based on in vivo experimental data are associated with large uncertainties \parencite{Link2014}.
-Using in vitro, or unreliable in vivo parameter estimates to study the steady state and dynamic responses of metabolic networks to different perturbations, under in vivo conditions, reduces confidence in the model predicted behaviour.
-Consequently, this hampers the use of these models to gain insight into metabolic network function \parencite{Andreozzi2016a,Vasilakou2016}, and the increase in prediction uncertainty becomes an obstacle for using the predicted responses as a basis for designing the metabolic networks to achieve any of the aforementioned goals of metabolic engineering and design \parencite{Saa2017}.
+	Despite its importance, the parameterization of kinetic models is still a problem for which solutions are a subject of debate within the modeling community.
+Typically, enzyme kinetic rate laws are parameterized based on in vitro observations of enzyme activity, as opposed to observations made under in vivo conditions \parencite{Heijnen2005,Smallbone2007}.
+However, some researchers have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions, as opposed to in vitro conditions \parencite{Heijnen2005,Heijnen2013}.
+On the other hand, some reports have shown that despite the large uncertainties associated with parameters estimated based on in vivo experimental data \parencite{Link2014}, in vitro parameter estimates are a reasonable approximation of values that would be applicable under in vivo conditions \textcolor{red}{(Ron Milo paper comparing in vitro vs invivo enzyme turn over rates in PLoS Computational Biology)}.
 
-	Although some authors have sought to quantify this uncertainty using different techniques \parencite{Vanlier2013,Andreozzi2016a}, others have proposed to alleviate as well as constrain the uncertainty in parameter estimates and consequent model predictions by using a Monte Carlo approach to kinetic modeling of metabolism that allows for integration of experimentally observed in vivo data \parencite{Srinivasan2015}.
+	 Nevertheless, some authors have sought to quantify the uncertainty in in vivo parameter estimates using different techniques \parencite{Vanlier2013,Andreozzi2016a}, while others have proposed to alleviate as well as constrain the uncertainty in parameter estimates and consequent model predictions by using a Monte Carlo approach to kinetic modeling of metabolism.
+These approaches also allow for the integration of experimentally observed in vivo data.
+ORACLE \textcolor{red}{(Wang and Haztzimanikatis, 2004)} and Ensemble modeling \textcolor{red}{(Tan and Liao, 2008)} are two examples of such an approach.
+These and other Monte Carlo kinetic modeling methods have been previously reviewed \parencite{Srinivasan2015}.
 Bayesian approaches to improve parameter estimation and quantify estimation uncertainty have also been proposed \parencite{Saa2016}.
@@
 In doing so, we not only propose the number and types of perturbations that will provide the most useful data for parameter estimation, but also test the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks.
-
-	For the purposes of this method we assume that all intracellular metabolite concentrations and fluxes can be measured.
-We illustrate our methodology to identify parameters and design experiments to identify parameters in a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017}.
+We illustrate our methodology to identify parameters and design experiments to identify parameters in a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} under the assumption that all intracellular metabolite concentrations and fluxes can be measured.
 
@@
 	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
-On the other hand, if only 50\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.5 or 50\%.
+On the other hand, if only 10\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.1 or 10\%.
 Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
-In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 50\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
+In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 10\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
 

### 2018-04-27 ident/docs/draft_3.tex
completed section 3.5 on uncertainty and minor corrections to other sections

@@
 If data from $n_E$ experiments is available for the chosen metabolic network, as stated earlier, for each experiment $j = {1, 2, ..., n_E}$, we assume that all metabolite concentrations ($x$) and reaction fluxes ($v$) are measurable.
-We discuss the implications of relaxing this assumption later.
+We discuss the implications of relaxing this assumption in the results section.
 The pertinent information for each experiment $j$ is available as a vector of concentrations and fluxes, $\mathbf{x}_j$ and $\mathbf{v}_j$, respectively (Figure \ref{fig:ident-flowchart}b).
@@
 	\end{align}
-	Here, $v_{i,j}$ refers to the value of the flux $v_i$ obtained from experiment $j$.
+	Here, $v_{i,j}$ refers to the flux $v_i$ obtained from experiment $j$.
 $\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$, whose elements are denoted by $\theta_k$.
@@
 	The identifiability of parameter $\theta_k$, $k = {1, 2, ..., p}$, for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
-Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a non-negative value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
+Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a positive value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
 The solution $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_k$, and an infinite number of solutions are possible for a non-identifiable $\theta_k$.
@@
 
-	Following the methodology described in section \ref{sec:ident}, and demonstrated in section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations found based on experiments described in section \ref{sec:experiments} of the Supplementary Information can be tested for their ability to practically identify any of the three fluxes in the small metabolic network.
-This step would determine the degree of identifiability (defined in section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguish experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
+	Following the methodology described in Section \ref{sec:ident}, and demonstrated in Section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations of data sets obtained from experiments described in Section \ref{sec:experiments} of the Supplementary Information can be tested for their ability to practically identify any of the fluxes in the small metabolic network.
+This step would determine the degree of identifiability (defined in Section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguish experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
 In doing so, it is possible to obtain a minimal and informative collection of experiments that can be performed to identify as many model parameters as possible (Figure \ref{fig:ident-design}).
@@
 
-	The identifiability of each parameter based on each experiment with index $j = {1, 2, ..., n_E}$ is established based on the methodology described in section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in section \ref{sec:example}.
+	The identifiability of each parameter based on each experiment with index $j = {1, 2, ..., n_E}$ is established based on the methodology described in Section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in Section \ref{sec:example}.
 Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$ (Figure \ref{fig:ident-flowchart}b), then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
@@
 	In Section \ref{sec:initial_analysis} that follows, we show how the demonstrated methodology is capable of practically identifying and estimating parameters for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ using steady state flux values and metabolite concentrations.
-	The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:design} that follows.
-	The contribution of the uncertainty in the data (due to differences between in vivo and in vitro kinetics) towards identifying parameters in enzyme kinetic models is discussed in Section \ref{sec:uncertainty}.
+	The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:design}.
+	The contribution of the uncertainty in the data arising from either the differences between in vivo and in vitro kinetics, or the noise present in experimentally measured quantities towards identifying parameters in enzyme kinetic models is discussed finally in Section \ref{sec:uncertainty}.
 
 	\subsection{Identifying parameters in kinetic models of metabolism: an example}\label{sec:example}	
-	In this section, we illustrate the proposed methodology step by step to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model} of the Supplementary Information).
+	In this section, we illustrate the proposed methodology step by step to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model}).
 We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration.
-	\begin{align}\tag{S5}
-	v_1 = V_1^{max}\frac{ac}{ac+K_{1}^{ac}(ne)}
-	\end{align}		
+
 	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in Section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
@@
 	To test the practical identifiability of the parameters in Equation \ref{eq:v1_par}, we substitute any suitable in silico experimental data and determine the value of the denominator of the right hand side expression.
-Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).
+Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be positive (Figure \ref{fig:ident-flowchart}b).
 The parameter values that are obtained for $V_1^{max}$ and $K_1^{ac} (ne)$ by substituting in silico steady state experimental data are shown in Supplementary Figure \ref{fig:v1_v1max_ck_values}.
@@
 	\subsection{Establishing Structural identifiability of parameters based on closed-form solutions}\label{sec:proof}
-	As mentioned in Section \ref{sec:ident_def}, by definition, unique parameter values based on the model structure are possible for any structurally identifiable parameter.
-Also, for the proposed methodology (Figure \ref{fig:ident-flowchart}) to work, it should be possible to obtain closed form solutions for each parameter in the enzyme kinetic model for each flux as shown in Equation (\ref{eq:theta-eq}).
+	For the proposed methodology (Figure \ref{fig:ident-flowchart}) to work, it should be possible to obtain closed form solutions for each parameter in the enzyme kinetic model for each flux as shown in Equation (\ref{eq:theta-eq}).
 Since the ability to obtain closed-form solutions for each parameter is dependent on the model structure, any parameter that has non-unique closed-form solutions can be called a structurally non-identifiable parameter.
@@
 
-	However, we find that the parameters used to model $v_3$ using the Convenience kinetics rate law, and $v_5$ using the Hill kinetic rate law are not structural identifiable under certain conditions.
+	However, we find that the parameters used to model $v_3$ using the Convenience kinetics rate law, and $v_5$ using the Hill kinetic rate law are not structurally identifiable as they are given in Section \ref{sec:small-model}.
+
 	First, for $v_3$, we find that the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different closed-form solutions.
 Thus, based on the presence of non-unique but finite number of possible solutions for these parameters we can classify $v_3$ as a locally structurally identifiable flux.
-
-	We would like to point out that we used the Convenience kinetics rate law model for $v_3$ due to the inability of the computer algebra system to generate closed-form expressions for parameters in the Monod-Wyman-Chageaux model (Section \ref{sec:small-model}).
-However, we find that the Convenience kinetics model is not structurally identifiable in its current form.
-In order to alleviate this issue, we reduced the dimension of the parameter space for $v_3$.
+In order to alleviate local structural identifiability, we reduced the dimension of the parameter space for $v_3$.
 Originally, $\theta \in \mathbb{R}^3$ for $v_3$.
@@
 
+	%We would like to point out that we used the Convenience kinetics rate law model for $v_3$ due to the inability of the computer algebra system to generate closed-form expressions for parameters in the Monod-Wyman-Chageaux model (Section \ref{sec:small-model}).
+However, we find that the Convenience kinetics model is not structurally identifiable in its current form.
+
 	While $v_3$ is an allosterically regulated metabolic flux, $v_5$ describes a transcription/translation reaction using Hill kinetics.
@@
 We could not obtain closed form solutions for parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$ using the computer algebra system.
-So, instead of changing the model as we did for $v_3$, we resorted to reducing the dimension of the parameter space by fixing one of the three parameters, the Hill coefficient $n_e$.
-We illustrated this for reducing the parameter space of the Convenience kinetic model for $v_3$ earlier.
+So, instead of changing the model as we did for $v_3$ (see Section \ref{sec:small-model}), we resorted to reducing the dimension of the parameter space by fixing one of the three parameters, the Hill coefficient $n_e$.
+We illustrated the consequence of reducing the dimension of the parameter space of the Convenience kinetic model for $v_3$ earlier.
 With a fixed and known $n_e$, $K_e^{fdp}$ has two possible closed-form solutions, which make it a locally structurally identifiable parameter.
@@
 	\subsection{Relationship between structural and practical parameter identifiability}\label{sec:initial_analysis}
-	Together with the definition for structural identifiability we presented in Section \ref{sec:ident_def}, we also introduced the concept of practical parameter identifiability.
+	We mention in Section \ref{sec:ident_def} that, by definition, unique parameter values based on the model structure are possible for any structurally identifiable parameter.
+Together with this definition for structural identifiability, we also introduced the concept of practical parameter identifiability.
 To recall, we mentioned that  it should be possible to estimate unique parameter values based on all available experimental data for any practically identifiable parameter.
+
 	As shown in Figure \ref{fig:ident-flowchart} and illustrated for $v_1$ in Section \ref{sec:example}, to determine the practical identifiability of parameters we test for the existence of a non-zero denominator of the closed-form expressions of the parameters.
@@
 These parameters have two possible closed-form expressions.
-In Figure \ref{fig:ident_values}d we show the numerical values for one of the two possible solutions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$.
+In Figure \ref{fig:ident_values}d we show the numerical values for one of the two possible closed-form expressions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$.
 The numerical values for the second closed-form expressions of the three parameters is presented in Supplementary Figure, and they also have a similar distribution.
-Based on the prior definition for practically identifiable parameters, the numerous possible values that the three parameters can acquire (Figure \ref{fig:ident_values}d) lead us to conclude that the parameters in $v_3$ are practically non-identifiable when they are only structurally locally identifiable.
-Also, these parameters are practically non-identifiable even after the reduction in the practically identifiable parameter space through application of the physiological relevance condition for practical identifiability (Figure \ref{fig:ident-flowchart}b).
+Based on the prior definition for practically identifiable parameters, the numerous possible values that the three parameters can acquire (Figure \ref{fig:ident_values}d) leads us to conclude that the parameters in $v_3$ are practically non-identifiable when they are only structurally locally identifiable.
+Also, unlike $v_5$, which is practically identifiable in the presence of local structural identifiability, parameters for $v_3$ are practically non-identifiable even after the reduction in the practically identifiable parameter space realized using the physiological relevance condition (Figure \ref{fig:ident-flowchart}b).
 
@@
 
-	So, the observation of practical non-identifiability for $v_3$ in the presence of local structural identifiability is in contrast to what we observed for $v_5$ earlier.
+	%So, the observation of practical non-identifiability for $v_3$ in the presence of local structural identifiability is in contrast to what we observed for $v_5$ earlier. 
 	However, we find that $v_3$ is practically identifiable when its parameters are also structurally identifiable (Figure \ref{fig:v3_var_ck_values}).
@@
 	The analysis of parameter practical identifiability can be used to gather information on the type of experiments that can provide useful data for parameter estimation.
-For instance, during practical identifiability analysis, if either the denominator of the closed-form expression is zero, or if the parameter values that are obtained are not physiologically feasible (Figure \ref{fig:ident-flowchart}b), then the experimental data set concerned is said to be incapable of practically identifying the said parameter.
+For instance, during practical identifiability analysis, if either the denominator of the closed-form expression is zero, or if the parameter values that are obtained are not physiologically feasible (Figure \ref{fig:ident-flowchart}b), then the experimental data set concerned is said to be incapable of practically identifying that said parameter.
 Consequently, the data from that combination of experiments is considered non-informative.
@@
 
-	In Figure \ref{fig:ident} we show the number and percentage of experimental data combinations that are capable of identifying each parameter in flux $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b) and $v_5$ (Figure \ref{fig:ident}c).
-The number and percentage of combinations identifying $V_1^{max}$ and $K_1^{ac}(ne)$ in $v_1$, and the closed-form solutions of the three parameters of $v_3$ are shown in Supplementary Figures \ref{fig:v1_v1max_ck_ident} and \ref{fig:v3_1_ck_ident}, respectively.
+	In Figure \ref{fig:ident} we show the number of experimental data combinations that are capable of identifying each parameter, and consequently, the degree of identifiability of each parameter (percentage experimental data combinations that are capable of identifying each parameter) in flux $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b) and $v_5$ (Figure \ref{fig:ident}c).
+The degree of identifiability for $V_1^{max}$ and $K_1^{ac}(ne)$ in $v_1$ is shown in Supplementary Figures \ref{fig:v1_v1max_ck_ident}, and the degree of identifiability for both closed-form expressions of the three parameters of $v_3$ are shown in Supplementary Figure \ref{fig:v3_1_ck_ident}.
 The degree of identifiability of each parameter is also given in these figures as percentages.
-
-	Based on their degrees of identifiability, we see that the maximum reaction rates ($V_i^{max}$) are more or similarly identifiable in comparison to the corresponding enzyme binding ($K_i$) constants or the activation/inhibition constants, in the respective reaction rate law models (Figure \ref{fig:ident}, Supplementary Figures \ref{fig:v1_v1max_ck_ident} and \ref{fig:v3_1_ck_ident}).
-We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition.
 
@@
 In these cases the degree of identifiability (Supplementary Figure \ref{fig:v1_v1max_ck_ident} for $v_1$ and Supplementary Figure \ref{fig:v3_1_ck_ident} for $v_3$) refers to the number of experimental data sets that can determine physiologically relevant values for the corresponding parameters.
-So, we have shown for $v_1$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $v_3$ (Supplementary Figure \ref{fig:v3_1_ck_ident}) that a greater number of experimental data sets can predict physiologically relevant values for $V_1^{max}$ and $V_3^{max}$ than for $K_1^{ac}(ne)$ and $K_3^{fdp}$ or $K_3^{pep}$, respectively.
+
+	Based on their degrees of identifiability, we see that the maximum reaction rates ($V_i^{max}$) are more or similarly identifiable in comparison to the corresponding enzyme binding ($K_i$) constants or the activation/inhibition constants, in the respective reaction rate law models (Figure \ref{fig:ident}, Supplementary Figures \ref{fig:v1_v1max_ck_ident} and \ref{fig:v3_1_ck_ident}).
+We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition.
+As mentioned earlier, for $v_1$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $v_3$ (Supplementary Figure \ref{fig:v3_1_ck_ident}) we have shown that a greater number of experimental data sets can predict physiologically relevant or non-zero positive values for $V_1^{max}$ and $V_3^{max}$ than for $K_1^{ac}(ne)$ and $K_3^{fdp}$ or $K_3^{pep}$, respectively.
 
 	The degree of identifiability of $v_3$, when its parameters are structurally and practically identifiable (Sections \ref{sec:proof} and \ref{sec:initial_analysis}) are shown in Supplementary Figure \ref{fig:v3_var_1_2_ck_ident}.
-Accordingly, we find that in all cases other than $V_1^{max}$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $k_1^{cat}$ in $v_1$ (Figure \ref{fig:ident}a), all data sets used to test practical identifiability can determine unique values for parameters when the corresponding parameter is structurally identifiable.
+Accordingly, we find that with the exception of $V_1^{max}$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $k_1^{cat}$ in $v_1$ (Figure \ref{fig:ident}a), all data sets used to test practical identifiability can determine unique values for parameters when the corresponding parameter is structurally identifiable.
 
 	We can attribute the difference in the degree of identifiability between $v_1$ (Figure \ref{fig:ident}a and Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and the other fluxes ($v_2$, $v_3$ and $v_5$) to the ability of data from different combinations of experiments to satisfy the conditions for practical identifiability of that parameter, that can be determined a priori.
-In systems identification terminology, data requirements for parameters identification can be tied to selecting experiments that are persistently excitable for the flux being identified.
+In systems identification terminology, data requirements for parameter identification can be tied to selecting experiments that are persistently excitable for the flux being identified.
 Any input signal should be rich or informative enough to guarantee full excitement of the dynamics of the system \parencite{Ljung1994}.
@@
 
-	As mentioned in Supplementary Section \ref{sec:experiments}, we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Table \ref{tab:pval}).
+	As mentioned in Supplementary Section \ref{sec:experiments}, we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Supplementary Table \ref{tab:pval}).
 In Figures \ref{fig:expinfo}a, \ref{fig:expinfo}b and \ref{fig:expinfo}c, the contribution from different experiment types for identifying parameters in $v_1$, $v_2$ and $v_5$ are respectively shown as spider plots.
@@
 This is in agreement with the condition for identifiability that we discussed earlier (Equations \ref{eq:v1_k1cat} and \ref{eq:v1_par}).
-Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we also say that identifiability analysis is crucial to determine the minimum number of experiments, as well as the nature of experiments that can help identify parameters for $v_1$.
+Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we also say that identifiability analysis is crucial to determine the minimum number of experiments, along with the nature of experiments that can help identify parameters for $v_1$.
 
 	With $v_2$, we see that the enzyme perturbations as well as the acetate perturbation experiments have similar contributions towards datasets that can identify $v_2$ (Figure \ref{fig:expinfo}b).
-This is also supports our arguments made earlier with regards to the identifiability conditions for $v_2$ and the reasons for the difference in the type of experiments that are informative between $v_1$ and $v_2$.
+This also supports our arguments made earlier with regards to the identifiability conditions for $v_2$, and the reasons for the difference in the type of experiments that are informative between $v_1$ and $v_2$.
 Accordingly, we find that in comparison to selecting experiments to identify $v_1$, there is very little restriction on the types of experiments that are informative to identify $v_2$.
@@
 
+	So, now we have established a hierarchy of parameters based on their identifiability.
+Parameters for $v_2$ and $V_e^{max}$ in $v_5$ that are most identifiable are at the top of the hierarchy, while $v_1$ and $K_e^{fdp}$ in $v_5$ fall at the bottom of the list as the experiments needed for their identification require careful consideration.
+When $v_3$ is structurally identifiable, their parameters also do not require any experimental design considerations.
+
 	\subsection{Parameter non-identifiability due to uncertainty in Experimental Data}\label{sec:uncertainty}
+	In all the aforementioned scenarios, the kinetic rate law from which data is derived is known and same as the model for which parameters are estimated.
+However, in reality, the kinetic rate law based on which metabolic networks function and from which in vivo experimental data is extracted is mostly unknown.
+The rate laws are primarily inferred through the parameter estimation procedure.
+This is one of the motivations for the development of approximate kinetic rate law models \parencite{Heijnen2013,Smallbone2007,Berthoumieux2013}.
+	So, there is a need to see if the methodology that we have developed here is capable of handling the uncertainty that arises due to the mismatch between the model and the data used to identify and estimate the parameters in the model.
 
-	This will cover issues related to the model used for v3. Convenience kientic model vs MWC model for data generation. Should this uncertainty also include noise?
+	The scope within which we have defined the model (Section \ref{sec:small-model}) makes such an analysis possible by changing the enzyme kinetic rate law used to describe $v_3$.
+Note that the original description \parencite{Kotte2014,Srinivasan2017} of the network (Figure \ref{fig:network}) uses the Monod-Wyman-Chageaux (MWC) model to describe the flux through $v_3$.
+Whereas, so far we have used a Convenience kinetic rate law description for both data generation as well as identifiability analysis.
+To determine the ability of our methodology to handle the in vivo-in vitro model uncertainty, we use the MWC model description to generate the in silico experimental data.
+This data will then be used to identify parameters in all the fluxes, including $v_3$ that is described by the Convenience kinetic model.
+
+	First, we find that the spread in the estimated Convenience kinetics parameter values, when $v_3$ is only locally structurally identifiable, is much larger than when there is mismatch between the model generating the data and the model that is being identified (Supplementary Figure).
+A more important observation is that even when the parameters are structurally identifiable in $v_3$ (achieved by assuming either $K_3^{fdp}$ or $K_3^{pep}$ as a known constant), they can at most only be locally practically identifiable.
+This is shown by the spread in the estimate values of the structurally identifiable parameters when steady state data based on the MWC model is used in Supplementary Figure.
+
+	Second, note that the dynamics of the network as represented by an MWC model for $v_3$ are different from the dynamic characteristics expressed when a Convenience kinetics model is used instead to describe $v_3$.
+Thus, this can bring about a change in the steady state concentrations and fluxes observed for the various in silico experiments listed in Supplementary Table \ref{tab:pval}.
+For instance, since the enzyme concentration \textit{E} is dependent on the dynamics of the network, the uptake flux $v_1$ can be different between the two models for the same acetate concentration (Equation \ref{eq:flux1}).
+Consequently, as the enzyme concentration \textit{E} is not part of the closed-form expression for $V_1^{max}$ and $K_1^{ac}(ne)$ in Equation \ref{eq:v1_par}), the difference in the steady state data used for identification can result in a change in the spread (\textcolor{red}{uncertainty}) observed for estimated values of $V_1^{max}$ and $K_1^{ac}(ne)$ (Supplementary Figure).
+Thus, while quantifiable, the uncertainty due to mismatch in the in vivo and in vitro information will carry over to the estimated parameters.
+
+	However, this issue can be resolved if more in vivo information is used for parameter identification.
+We first observe this scenario when Equation (\ref{eq:v1_k1cat}), which includes \textit{E}, is used to identify $k_1^{cat}$ and $K_1^{ac}$: these parameters are practically identifiable even when in silico steady state data from a mismatched model is used for identification (Supplementary Figure).
+We also observe this with the identification of $v_2$ and $v_5$ (Supplementary Figure).
+For these two fluxes all available and necessary steady state information are part of their identifiability expressions, thereby leaving no room for any uncertainties to propagate from the data through the practical identification process.
+
+	Apart from the mismatch between the in vivo and the in vitro enzyme kinetic rate laws, uncertainty in experimental data also arises due to the presence of noise in the measured experimental data.
+This noise could be attributed to the measurement error commonly encountered in process analytics.
+In order to test the robustness of our methodology to practically identify parameters using steady state data with measurement errors, we used in silico experimental data with 5\% additive noise for practical identification, instead of the noise-free data that we have used so far.
+
+	We found that in every case where parameters are structurally identifiable, the noise did not have any effect on the identifiability of the parameters or their estimated values.
+We also found that inclusion of all necessary data (e.g., the presence or lack thereof of enzyme concentration \textit{E} for $v_1$) can alleviate issues related to using experimental data with errors for identification and estimation: the degrees of identifiability of $V_1^{max}$ and $K_1^{ac}(ne)$ had non-zero standard deviations associated with them, but the degrees of identifiability of $k_1^{cat}$ and $K_1^{ac}$ did not.
+Using a similar reasoning to the earlier scenario in the presence of mismatches between in vitro and in vivo model, we can say that there is no room for any uncertainties to propagate from the noisy data when all necessary steady state information for identification is available.
+Thus, both $v_2$ and $v_5$ also did not show any differences in either their degree of identifiability or their estimated parameter values.
+
+	However, for $v_3$, whose parameters are only locally structurally identifiable, we found small non-zero standard deviations in the degrees of identifiabilities (Supplementary Figure) when noisy data is used.
+Although this was seen due to the differences in the number of data combinations that can estimate positive values for each of the three parameters between different noisy experimental data sets, we observe that the standard deviation in the estimated parameter values for each data, between different samples of noisy experimental data, is small (Supplementary Figure).
 

### 2018-06-04 ident/docs/draft_3.tex
first correction set from wrc

@@
 Some notable examples include the design of microorganisms for the biosynthesis of commodity and specialty chemicals \parencite{Andreozzi2016}, engineering mammalian cells as therapeutic targets for cures to some ailments affecting humans \parencite{DiFilippo2016,Apaolaza2017}, and changing the constituents of the human gut microbial community to cure related diseases \parencite{Zerfab2018}.
-These applications require us to understand the numerous complex interaction, their roles in cell function, and sometimes even the mechanisms behind these interactions.
+These applications require us to understand the numerous complex interactions, their roles in cell function, and sometimes even the mechanisms behind these interactions.
 Computational models offer a systematic way to integrate available experimental data, and to study and understand these interactions through mathematical representations of the biological systems in which these interactions occur \parencite{Bordbar2014a,Saa2017}.
@@
 Consequently, the metabolite concentrations within the metabolic network are assumed to be constant, and changes in metabolite concentrations are not modeled.
-Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various non-catalytic regulatory interactions that are also responsible of metabolic function.
+Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various non-catalytic regulatory interactions that are also responsible for metabolic function.
 These shortcomings prevent CBMs from being used to fully understand the steady state as well as the dynamic characteristics of metabolic networks.
 
-	In contrast, the implications of regulatory interactions and changes in metabolite concentrations on different characteristics of metabolism can be studied using kinetic models of metabolism \parencite{Saa2017}.
+	In contrast, the effects of regulatory interactions and changes in metabolite concentrations on different characteristics of metabolism can be studied using kinetic models of metabolism \parencite{Saa2017}.
 These models account for changes in metabolite concentrations subject to thermodynamic and regulatory constraints that underly metabolic networks in addition to its stoichiometry \parencite{Link2014}.
-Kinetic models can not only help us better understand lesser known and understood characteristics of metabolism like bistability \parencite{Kotte2014}, and their role in human health, but can also improve predictions about the impact of engineering design perturbations on metabolism, and propose alternative designs to achieve metabolite production goals \parencite{Khodayari2016}.
+Kinetic models can not only help us better understand lesser known and understood characteristics of metabolism like bistability \parencite{Kotte2014}, and their role in human health, but can also improve predictions about the impact of engineering design perturbations on metabolism, and help propose alternative designs to achieve metabolite production goals \parencite{Khodayari2016}.
 
-	Kinetic models differ from CBMs in their use of heavily parameterized mechanistic enzyme kinetic rate laws to model enzyme catalyzed fluxes within a metabolic network.
-These parameters represent various aspects of the enzyme kinetic rate laws \parencite{Srinivasan2015,Saa2017}.
-Hence, the use of kinetic models requires information on all the enzyme kinetic rate laws that will be used to model all the fluxes within a metabolic network, as well as numerical values for all the parameters used in these rate laws.
-Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values might play a crucial role \parencite{Srinivasan2017}.
+	Kinetic models differ from CBMs in their use of mechanistic enzyme kinetic rate laws to model enzyme catalyzed fluxes within a metabolic network \parencite{Srinivasan2015,Saa2017}.
+The use of kinetic models requires information on the enzyme kinetic rate laws that will be used to model the fluxes within a metabolic network, as well as numerical values for the parameters used in these rate laws.
+Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values play a crucial role \parencite{Srinivasan2017}.
 
-	Despite its importance, the parameterization of kinetic models is still a problem for which solutions are a subject of debate within the modeling community.
+	Despite their importance, the parameterization of kinetic models is still a problem for which solutions are a subject of debate within the modeling community.
 Typically, enzyme kinetic rate laws are parameterized based on in vitro observations of enzyme activity, as opposed to observations made under in vivo conditions \parencite{Heijnen2005,Smallbone2007}.
@@
 
-	 Nevertheless, some authors have sought to quantify the uncertainty in in vivo parameter estimates using different techniques \parencite{Vanlier2013,Andreozzi2016a}, while others have proposed to alleviate as well as constrain the uncertainty in parameter estimates and consequent model predictions by using a Monte Carlo approach to kinetic modeling of metabolism.
+	 Authors have sought to constrain and determine uncertainty in parameter estimates and consequent model predictions by using Monte Carlo approaches for kinetic modeling of metabolism \parencite{Andreozzi2016a}.
 These approaches also allow for the integration of experimentally observed in vivo data.
-ORACLE \textcolor{red}{(Wang and Haztzimanikatis, 2004)} and Ensemble modeling \textcolor{red}{(Tan and Liao, 2008)} are two examples of such an approach.
+ORACLE \textcolor{red}{(Wang and Haztzimanikatis, 2004)} and Ensemble modeling \textcolor{red}{(Tan and Liao, 2008)}, are two such examples.
 These and other Monte Carlo kinetic modeling methods have been previously reviewed \parencite{Srinivasan2015}.
 	 Bayesian approaches to improve parameter estimation and quantify estimation uncertainty have also been proposed \parencite{Saa2016}.
+\cite{Vanlier2013} provide a review of different Bayesian approaches to quantify parameter uncertainty.
 
-	In spite of the development of these methods to quantify parameter estimation uncertainty, model parameter identifiability, a necessary, and sometimes sufficient condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.
-Briefly, it concerns with the ability to estimate unique values for all model parameters from observed experimental data.
-In a model, any parameter is said to be structurally or a priori identifiable if its values can be uniquely estimated independent of all other model parameters from available experimental data.
-However, if parameters cannot be uniquely estimated independent of each other due to redundant model parameterization, or due to the nonlinear relationship between the model parameters, then the parameters are said to be structurally non-identifiable.
+	Related to the development of these methods to quantify parameter estimation uncertainty, the importance of model parameter identifiability, i.e., the condition that needs to be satisfied to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.
+In a model, a parameter is said to be structurally or a priori identifiable if its values can be uniquely estimated from available experimental data, independent of all other model parameters.
+However, if a parameter cannot be uniquely estimated due to redundant model parameterization, or due to the nonlinear relationships between the model parameters, then the parameter is said to be structurally non-identifiable.
 Conversely, if the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is unquantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994}.
 
-	Authors have proposed to overcome concerns with parameter identifiability by proposing approximate kinetic models of metabolism that utilize empirical enzyme kinetic rate laws whose parameters have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}.
+	Authors have proposed ways to assess parameter identifiability by proposing approximate kinetic models of metabolism that utilize empirical enzyme kinetic rate laws with parameters that have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}.
 Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Ljung1994,Nikerel2009,Berthoumieux2013,Raue2014}\textcolor{red}{(paper from Rudiyanto Gunawan on model discrimination and sensitivity analysis)}.
 
-	Methods to improve practical identifiability through a priori experimental design have also been developed, with focus on kinetic models of metabolism \parencite{Gadkar2005a,Vanlier2014a,Raue2014}.
+	Methods to improve practical identifiability through a priori experimental design have also been developed, with a focus on kinetic models of metabolism \parencite{Gadkar2005a,Vanlier2014a,Raue2014}.
 Some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, while some of them suffer from computational limitations when applied to kinetic models of large metabolic networks \parencite{Gadkar2005a,Raue2014}\textcolor{red}{(Banga method using FIM for D-optimal design, ??)}.
@@
 	In this paper, we propose a scalable methodology that uses available steady state fluxomics, metabolomics and proteomics data to test the practical identifiability of parameters for each individual reaction in kinetic models of metabolism.
-We demonstrate how the computer algebra-based method that we have developed can also facilitate the design of experiments that are minimal and informative to generate data required to estimate unique parameter values for all reaction fluxes in a metabolic network.
-In doing so, we not only propose the number and types of perturbations that will provide the most useful data for parameter estimation, but also test the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks.
-We illustrate our methodology to identify parameters and design experiments to identify parameters in a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} under the assumption that all intracellular metabolite concentrations and fluxes can be measured.
-
-	%some perform identifiability within a dynamic context \parencite{Gadkar2005a,Nikerel2009,Vanlier2014a} and
-
-	  %We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
-
- %In \textcolor{red}{Section x} we provide detailed definitions for structural and practical identifiability of parameters followed by a step by step description of our proposed methodology for practical parameter identification in kinetic models of metabolism.
-In \textcolor{red}{Section y} we discuss the implications
-
-% We explain these identifiability conditions in detail in the Methods section. 
-%Bayesian approach to identification and experimental design \parencite{Vanlier2012}.
+We demonstrate how the computer algebra-based method that we have developed to perform this test can also facilitate the design of experiments for generating the data required to estimate unique parameter values for all reaction fluxes in a metabolic network.
+In doing so, we are able to propose the number and types of perturbations that will provide the most useful data for parameter estimation, as well as test the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks.
+We illustrate our methodology using a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} under the assumption that all intracellular metabolite concentrations and fluxes can be measured.
 
 	\section{Methods}\label{sec:methods}
+	In Section \ref{sec:kinetic_model} we present the typical structure of a kinetic model of metabolism and a preliminary description of the least squares method by which the model parameters are estimated from experimental data.
+In Section \ref{sec:ident_def} we provide formal mathematical definitions for structural and practical identifiability.
+The computer-algebra system based method to assess identifiability that we have developed is described in Section \ref{sec:ident} that follows.
+We also provide a description of a numerical equivalent for the computer algebra-based method to estimate parameters numerically in Section \ref{sec:numerical_method}.
+In Section \ref{sec:degree_of_identifiability}, we define a quantitative metric to describe the identifiability of parameters, and in Section \ref{sec:small-model} a complete description of the small metabolic network used to demonstrate the methodology that we have developed is provided.
+Finally, a description of how the method can be used for experimental design for parameter estimation is given in Section \ref{sec:experimental_design}.
+
 	\subsection{Parameter estimation for kinetic models of metabolism}\label{sec:kinetic_model}
-	In kinetic models of metabolism, ordinary differential equations (ODE) are used to express the rate of change of metabolite concentrations ($x$) as a function of the reaction fluxes ($v$) in the metabolic network (Equation \ref{eq:kinstoich}).
-The matrix $\mathbf{S}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
+	In kinetic models of metabolism, ordinary differential equations (ODE) are used to express the rate of change of metabolite concentrations ($x\in\mathbb{R}^{n_x}$) as a function of the reaction fluxes ($v\in\mathbb{R}^{n_r}$) in the metabolic network (Equation \ref{eq:kinstoich}).
+The matrix $\mathbf{S}\in\mathbb{R}^{n_x \text{x} n_r}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
 	\begin{subequations}\label{eq:kinstoich}
@@
 	\end{subequations}
-	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Srinivasan2015}.
-Accordingly, $f$ is a nonlinear function of the vector of metabolite concentrations ($x$), the vector of enzyme kinetic parameters ($\theta$) and other input concentrations ($u$).
+	The expression for the nonlinear function ($f$) used to describe each reaction flux $v_i$ in $v$, $i={1, 2, ..., n_r}$, in a particular kinetic model (Equation \ref{eq:kinstoich}b) is dependent on the enzyme kinetic mechanism that is used to model the reaction \parencite{Srinivasan2015}.
+Accordingly, $f$ is typically a nonlinear function of the vector of metabolite concentrations ($x\in \mathbb{R}^{n_x}$), the vector of enzyme kinetic parameters ($\theta\in\mathbb{R}^{n_p}$) and other input concentrations ($u \in \mathbb{R}^{n_u}$).
 
-	Parameter estimation methods based on optimization principles are typically used to determine true parameter values based on available experimental data.
-Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:lsqopt}) to estimate the values of enzyme kinetic parameters, $\theta$, based on the measured data.
+	Parameter estimation methods based on optimization principles are used to determine true parameter values from experimental data.
+Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:lsqopt}) to estimate the values of enzyme kinetic parameters, $\theta \in \mathbb{R}^{n_p}$, based on the measured data.
 	\begin{subequations}\label{eq:lsqopt}
 		\begin{align}
-		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2\\
+		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{n_x + n_r}\sum_{l=1}^{n_t}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2\\
 		&\theta_l \le \theta \le \theta_u
@@
 	\end{subequations}
-	Here $y = \left[x, v\right]^T$ is the vector of both concentrations ($x$) and fluxes ($v$).
-The minimization of least square error between the measured ($y^*$) and modeled ($y$) concentrations and fluxes, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration and flux, at each time point, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}).
-The parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).
+	Here $y \in \mathbb{R}^{n_x + n_r}$ is the combined vector of concentrations ($x\in \mathbb{R}^{n_x}$) and fluxes ($v \in \mathbb{R}^{n_r}$), at each time point $l={1, 2, ..., n_t}$.
+The minimization of the least squares error between the measured ($y^*$) and modeled ($y$) concentrations and fluxes, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration and flux, at each time point $l={1, 2, ..., n_t}$, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem.
+The least squares parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).
 
 	\subsection{Structural and practical identifiability of parameters in kinetic models}\label{sec:ident_def}
-	In the Introduction, we briefly metioned that the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
+	In the Introduction, we briefly mentioned that the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
 Below, we provide a formal definition of structural and practical identifiability of parameters.
 
-	The parameters in $\theta$ in any nonlinear model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \left[x, v\right]^T = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
+	The parameters in $\theta$ in any nonlinear model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
 	\begin{align}\label{eq:stident}
@@
 
-	\subsection{A method to determine practical identifiability of kinetic models of metabolism}\label{sec:ident}
+	\subsection{A method to determine identifiability of kinetic models of metabolism}\label{sec:ident}
 	We provide the mathematical framework for \textcolor{red}{identification} of parameters in kinetic models of metabolism in this section.
@@
 	For each flux $v_i$, $i={1, 2, ..., n_r}$, in the kinetic model, let $\mathbf{\theta} \in \mathbb{R}^p$ in Equation (\ref{eq:kinstoich}b).
-If data from $n_E$ experiments is available for the chosen metabolic network, as stated earlier, for each experiment $j = {1, 2, ..., n_E}$, we assume that all metabolite concentrations ($x$) and reaction fluxes ($v$) are measurable.
+If data from $n_E$ experiments is available for the chosen metabolic network, as stated earlier, for each experiment $j = {1, 2, ..., n_E}$, we assume that all metabolite concentrations ($x\in\mathbb{R}^{n_x}$) and reaction fluxes ($v\in\mathbb{R}^{n_r}$) are measurable.
 We discuss the implications of relaxing this assumption in the results section.
@@
 	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method.
-The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta}$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
+The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta} \in \mathbb{R}^p$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
 This is done by first selecting a combination of $p\le n_E$ experimental data.
@@
 
-	%For completeness, in the following sections we provide a previously published kinetic model of a small gluconeogenic network (section \ref{sec:small-model}).
+	\subsection{Numerical parameter identification and estimation}\label{sec:numerical_method}
+	The identification methodology described in the previous section can be extended to be used in conjunction with an optimization-based procedure for numerical estimation of parameters from experimental data.
+The $\mathbb{R}^p$ nonlinear algebraic equations formulated for each flux $v_i$ (Equation \ref{eq:nonlineq}) can be written as shown in Equation (\ref{eq:numerical_opt}) below to be solved as an optimization problem.
+Minimization of the sum of the absolute error values is used as the objective, and the $\mathbb{R}^p$ nonlinear algebraic equations are provided as inequality constraints.
+
+	\begin{subequations}\label{eq:numerical_opt}
+		\begin{align}
+		\underset{\theta, \varepsilon}{min} & \sum_{j=1}^{p} \left|\varepsilon_j\right|\\
+		& v_{i, j} - f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) \le \varepsilon_j && j=\{1, 2, ..., p\}\\
+		& \theta_l \le \theta_j \le \theta_u && j=\{1, 2, ..., p\}
+		\end{align}
+	\end{subequations}
+
+	The formulation of the above optimization problem is similar to the least squares problem shown earlier in Equation (\ref{eq:lsqopt}).
+Notice that the parameter values are bounded above and below by $\theta_u$ and $\theta_l$, respectively.
+These bounds can be specified to be arbitrarily large to include all possible parameter values.
 

### 2018-06-04 ident/python2/ss-ident/v3_ck_numerical_ident.py
comment all calculation code - leave only plotting codes uncommented

@@ -31,16 +31,16 @@ optim_options = {"solver": "ipopt",
 #                  "opts": {"qpsol": "qpoases"}}
 initial_value = [[80, 80, 80, 0, 0, 0]]  # [100, 400, 400, 0, 0, 0], [50, 50, 400, 0, 0, 0]
 # randomized_initial_values = generate_random_initial_conditions(initial_value, 10, negative=1)
-problem = {"lbx": 6 * [0],
-           "ubx": [100, 100, 100, 1e-20, 1e-20, 1e-20],
-           "lbg": 3 * [0],
-           "ubg": 3 * [0]}
-from numerical_ident import solve_multiple_initial_conditions
-all_sol_df, _ = solve_multiple_initial_conditions(all_initial_conditions=initial_value,
-                                                  experimental_data=all_exp_data, chosen_fun=1, prob=problem,
-                                                  optim_options=optim_options, number_of_parameters=3, flux_id=3,
-                                                  flux_choice=[3], exp_df=arranged_data_df,
-                                                  file_name=storage_file_name)
+# problem = {"lbx": 6 * [0],
+#            "ubx": [100, 100, 100, 1e-20, 1e-20, 1e-20],
+#            "lbg": 3 * [0],
+#            "ubg": 3 * [0]}
+# from numerical_ident import solve_multiple_initial_conditions
+# all_sol_df, _ = solve_multiple_initial_conditions(all_initial_conditions=initial_value,
+#                                                   experimental_data=all_exp_data, chosen_fun=1, prob=problem,
+#                                                   optim_options=optim_options, number_of_parameters=3, flux_id=3,
+#                                                   flux_choice=[3], exp_df=arranged_data_df,
+#                                                   file_name=storage_file_name)
 
 index_labels = ['sample_name', 'data_set_id']
 numerical_ident_df = retrieve_experimental_data_from_file(data_file_name=storage_file_name,
@@ -50,24 +50,24 @@ all_parameter_info = process_opt_solution(numerical_ident_df, arranged_data_df, 
 # get default parameter values
 default_parameter_values = true_parameter_values()
 
-# get parameter value plot
-# parameter_values_plot(all_parameter_info, default_parameter_values, box=False, violin=True)
-
 # extract all parameter values
-from process_ident_data import get_parameter_value
-validation_info = get_parameter_value(all_parameter_info, numerical_ident_df)
[…]

### 2018-06-04 ident/python2/ss-ident/numerical_ident.py
new option for mwc model with l1 and l2 objectives and change fun name for mwc estimation

@@ -130,7 +130,7 @@ def v3_numerical_ck_all(chosen_data):
     return nlp
 
 
-def v3_numerical_problem(chosen_data):
+def v3_mwc_numerical_problem(chosen_data):
     """fun to calculate objectives and constraints for numerical identification of v3
         using MWC kinetics"""
 
@@ -166,6 +166,51 @@ def v3_numerical_problem(chosen_data):
     objective = error[0] ** 2
     for i_error_id in range(1, number_data):
         objective += error[i_error_id] ** 2
+
+    # get complete set of all optimization variables vertcat([parameters, error])
+    x = casadi.vertcat(*[v3max, k3fdp, k3pep, l3fdp, error])
+
+    nlp = {'x': x, 'f': objective, 'g': constraints}
+
+    return nlp
+
+
+def v3_mwc_numerical_problem_l1_obj(chosen_data):
+    """fun to calculate objectives and constraints for numerical identification of v3
+            using MWC kinetics"""
+
+    number_data = len(chosen_data)
+    # symbolic variables for use with casadi
+    v3max = casadi.SX.sym("v3max")
+    k3fdp = casadi.SX.sym("k3fdp")
+    k3pep = casadi.SX.sym("k3pep")
+    l3fdp = casadi.SX.sym("l3fdp")
+    error = casadi.SX.sym("error", number_data)
+
+    # fdp_sat = 1 + y[1] / K3fdp
+    # pep_sat = 1 + y[0] / K3pep
+    # flux_3 = V3max * (fdp_sat - 1) * (fdp_sat ** 3) / (fdp_sat ** 4 + L3fdp * (pep_sat ** (-4)))
+
+    all_constraints = []
+    for i_data_id, i_data in enumerate(chosen_data):
+        _, pep, fdp, _, _, _, v3, _, _, _ = list(i_data)
+
+        # flux equation for each experiment i_data in chosen_data
+        fdp_sat = 1 + fdp / k3fdp
+        pep_sat = 1 + pep / k3pep
+        nr_3 = v3max * (fdp_sat - 1) * (fdp_sat ** 3)
+        dr_3 = (fdp_sat ** 4 + l3fdp * 1e6 * pep_sat ** (-4))
+        flux_3 = nr_3 / dr_3
+        # formulate constraint for each experimental data set
+        all_constraints.append(flux_3 - v3 - error[i_data_id])
+
+    # create casadi array of constraints
+    constraints = casadi.vertcat(*all_constraints)
+
[…]

### 2018-06-04 ident/python2/ss-ident/v3_mwc_numerical_ident.py

@@ -1,0 +1,85 @@
+from create_experiment_data import retrieve_experimental_data_from_file
+from identifiability_analysis import data_numerical_ident
+from numerical_ident import process_opt_solution
+from names_strings import true_parameter_values
+from validate_estimation import process_validation_info
+from plot_ident_results import parameter_values_plot
+from plot_ident_results import validation_plot
+import os.path
+
+
+# create data for identifiability analysis
+# from create_experiment_data import create_data_for_flux
+# create_data_for_flux(flux_id='v3', noise=0, number_samples=1)
+
+# extract experimental data from file
+new_data_file_name = os.path.join(os.getcwd(), 'exp/exp_v3_3_experiments')
+index_labels = ['sample_name', 'data_set_id', 'experiment_id']
+arranged_data_df = retrieve_experimental_data_from_file(data_file_name=new_data_file_name,
+                                                        multi_index_label=index_labels)
+
+# perform identifiability when v3 parameters are written using convenience kinetics
+# get combination of 3 experiments and perform identifiability on all fluxes that require 3 data sets
+storage_file_name = os.path.join(os.getcwd(), 'ident/ident_numerical_v3_l1_obj')
+# lexographic ordering of df indices
+all_exp_data = data_numerical_ident(arranged_data_df, 'sample_0')
+
+# NLP solver options
+optim_options = {"solver": "ipopt",
+                 "opts": {"ipopt.tol": 1e-18}}
+# optim_options = {"solver": "sqpmethod",\
+#                  "opts": {"qpsol": "qpoases"}}
+initial_value = [[80, 80, 80, 0, 0, 0]]  # [100, 400, 400, 0, 0, 0], [50, 50, 400, 0, 0, 0]
+# randomized_initial_values = generate_random_initial_conditions(initial_value, 10, negative=1)
+problem = {"lbx": 8 * [0],
+           "ubx": [2, 1, 1, 5, 1e-20, 1e-20, 1e-20, 1e-20],
+           "lbg": 4 * [0],
+           "ubg": 4 * [0]}
+from numerical_ident import solve_multiple_initial_conditions
+all_sol_df, _ = solve_multiple_initial_conditions(all_initial_conditions=initial_value,
+                                                  experimental_data=all_exp_data, chosen_fun=1, prob=problem,
+                                                  optim_options=optim_options, number_of_parameters=3, flux_id=3,
+                                                  flux_choice=[3], exp_df=arranged_data_df,
[…]

### 2018-06-04 ident/python2/ss-ident/v3_mwc_numerical_ident.py
change initial conditions

@@ -29,7 +29,7 @@ optim_options = {"solver": "ipopt",
                  "opts": {"ipopt.tol": 1e-18}}
 # optim_options = {"solver": "sqpmethod",\
 #                  "opts": {"qpsol": "qpoases"}}
-initial_value = [[80, 80, 80, 0, 0, 0]]  # [100, 400, 400, 0, 0, 0], [50, 50, 400, 0, 0, 0]
+initial_value = [[80, 80, 80, 80, 0, 0, 0, 0]]  # [100, 400, 400, 0, 0, 0], [50, 50, 400, 0, 0, 0]
 # randomized_initial_values = generate_random_initial_conditions(initial_value, 10, negative=1)
 problem = {"lbx": 8 * [0],
            "ubx": [2, 1, 1, 5, 1e-20, 1e-20, 1e-20, 1e-20],

### 2018-06-07 ident/docs/draft_4.tex
new draft with corrections from WRC

--- 2018-06-04 ident/docs/draft_3.tex
@@
 
-	Constraint-based models (CBMs) of metabolism are used to improve our understanding of metabolism by representing it as a stoichiometric network of reactions that operate under a pseudo-steady state assumption \parencite{Bordbar2014a}.
+	Constraint-based models (CBMs) of metabolism are used to improve our understanding of metabolism by representing it as a stoichiometric network of reactions \parencite{Bordbar2014a}.
 The ability of CBMs to shine light on the nonintuitive interactions that govern cellular metabolism is leveraged to engineer and asses the impact of designs that alter the ability of a cell to grow, or produce a desired metabolite \parencite{Maia2016}.
@@
 	In contrast, the effects of regulatory interactions and changes in metabolite concentrations on different characteristics of metabolism can be studied using kinetic models of metabolism \parencite{Saa2017}.
-These models account for changes in metabolite concentrations subject to thermodynamic and regulatory constraints that underly metabolic networks in addition to its stoichiometry \parencite{Link2014}.
+These models account for changes in metabolite concentrations subject to thermodynamic and regulatory constraints that underly metabolic networks in addition to their stoichiometry \parencite{Link2014}.
 Kinetic models can not only help us better understand lesser known and understood characteristics of metabolism like bistability \parencite{Kotte2014}, and their role in human health, but can also improve predictions about the impact of engineering design perturbations on metabolism, and help propose alternative designs to achieve metabolite production goals \parencite{Khodayari2016}.
 
-	Kinetic models differ from CBMs in their use of mechanistic enzyme kinetic rate laws to model enzyme catalyzed fluxes within a metabolic network \parencite{Srinivasan2015,Saa2017}.
+	Kinetic models differ from CBMs in their use of mechanistic enzyme kinetics to model enzyme catalyzed fluxes within a metabolic network \parencite{Srinivasan2015,Saa2017}.
 The use of kinetic models requires information on the enzyme kinetic rate laws that will be used to model the fluxes within a metabolic network, as well as numerical values for the parameters used in these rate laws.
@@
 Typically, enzyme kinetic rate laws are parameterized based on in vitro observations of enzyme activity, as opposed to observations made under in vivo conditions \parencite{Heijnen2005,Smallbone2007}.
-However, some researchers have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions, as opposed to in vitro conditions \parencite{Heijnen2005,Heijnen2013}.
-On the other hand, some reports have shown that despite the large uncertainties associated with parameters estimated based on in vivo experimental data \parencite{Link2014}, in vitro parameter estimates are a reasonable approximation of values that would be applicable under in vivo conditions \textcolor{red}{(Ron Milo paper comparing in vitro vs invivo enzyme turn over rates in PLoS Computational Biology)}.
+However, some researchers have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions \parencite{Heijnen2005,Heijnen2013}.
+On the other hand, some reports have shown that despite the large uncertainties associated with parameters estimated based on in vivo experimental data \parencite{Link2014}, in vitro parameter estimates are a reasonable approximation of values that would be applicable under in vivo conditions \parencite{Davidi2016}.
 
-	 Authors have sought to constrain and determine uncertainty in parameter estimates and consequent model predictions by using Monte Carlo approaches for kinetic modeling of metabolism \parencite{Andreozzi2016a}.
+	 Authors have sought to constrain and determine uncertainty in parameter estimates and associated model predictions by using Monte Carlo approaches for kinetic modeling of metabolism \parencite{Andreozzi2016a}.
 These approaches also allow for the integration of experimentally observed in vivo data.
-ORACLE \textcolor{red}{(Wang and Haztzimanikatis, 2004)} and Ensemble modeling \textcolor{red}{(Tan and Liao, 2008)}, are two such examples.
+ORACLE \parencite{Wang2004} and Ensemble modeling \parencite{Tran2008}, are two such examples.
 These and other Monte Carlo kinetic modeling methods have been previously reviewed \parencite{Srinivasan2015}.
 	 Bayesian approaches to improve parameter estimation and quantify estimation uncertainty have also been proposed \parencite{Saa2016}.
-\cite{Vanlier2013} provide a review of different Bayesian approaches to quantify parameter uncertainty.
+\textcite{Vanlier2013} provide a review of different Bayesian approaches to quantify parameter uncertainty.
 
-	Related to the development of these methods to quantify parameter estimation uncertainty, the importance of model parameter identifiability, i.e., the condition that needs to be satisfied to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.
-In a model, a parameter is said to be structurally or a priori identifiable if its values can be uniquely estimated from available experimental data, independent of all other model parameters.
-However, if a parameter cannot be uniquely estimated due to redundant model parameterization, or due to the nonlinear relationships between the model parameters, then the parameter is said to be structurally non-identifiable.
-Conversely, if the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is unquantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994}.
+	Related to the development of these methods to quantify parameter estimation uncertainty, the importance of model parameter identifiability, i.e., a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.
+
+	Briefly, unique values can only be estimated for structurally identifiable parameters.
+If a parameter cannot be uniquely estimated due to its nonlinear relationships with other parameters (redundant model parameterization), or due to the inability to measure all possible metabolite concentrations (presence of unobservable system states), then the parameter is said to be structurally non-identifiable.
+Similarly, practical identifiability of a parameter concerns with the informativeness of the experimental data used for parameter estimation, and the ability to quantify any existing uncertainty in the parameter estimates.
+If the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is not quantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994,Rodriguez-Fernandez2006}.
 
 	Authors have proposed ways to assess parameter identifiability by proposing approximate kinetic models of metabolism that utilize empirical enzyme kinetic rate laws with parameters that have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}.
-Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Ljung1994,Nikerel2009,Berthoumieux2013,Raue2014}\textcolor{red}{(paper from Rudiyanto Gunawan on model discrimination and sensitivity analysis)}.
+Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Ljung1994,Nikerel2009,Berthoumieux2013,Raue2014}.
+\textcite{Chis2011b} provide a comparison of the computational performance and implementation aspects of some of these methods.
 
-	Methods to improve practical identifiability through a priori experimental design have also been developed, with a focus on kinetic models of metabolism \parencite{Gadkar2005a,Vanlier2014a,Raue2014}.
-Some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, while some of them suffer from computational limitations when applied to kinetic models of large metabolic networks \parencite{Gadkar2005a,Raue2014}\textcolor{red}{(Banga method using FIM for D-optimal design, ??)}.
+	Methods to improve practical identifiability through a priori experimental design have also been developed, with a focus on kinetic models of metabolism \parencite{Gadkar2005a,Rodriguez-Fernandez2006,Vanlier2014a,Raue2014}.
+Some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, while some others suffer from computational limitations when applied to kinetic models of large metabolic networks \parencite{Gadkar2005a,Raue2014}.
 
-	In this paper, we propose a scalable methodology that uses available steady state fluxomics, metabolomics and proteomics data to test the practical identifiability of parameters for each individual reaction in kinetic models of metabolism.
-We demonstrate how the computer algebra-based method that we have developed to perform this test can also facilitate the design of experiments for generating the data required to estimate unique parameter values for all reaction fluxes in a metabolic network.
-In doing so, we are able to propose the number and types of perturbations that will provide the most useful data for parameter estimation, as well as test the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks.
+	Most of the aforementioned methods assess identifiability (structural or practical) in the presence of dynamic time course concentration data.
+In this paper, we show that the identifiability of parameters for each individual reaction in kinetic models of metabolism can be done using just steady state fluxomics, metabolomics and proteomics data, and propose a methodology to do so.
+We also demonstrate how this method can also facilitate the design of experiments for generating the necessary steady state data to estimate unique parameter values for all reaction fluxes in a metabolic network.
+In doing so, we are able to propose the number and types of perturbations that will provide the most useful data for parameter estimation.
+We also assess the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks using steady state data.
 We illustrate our methodology using a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} under the assumption that all intracellular metabolite concentrations and fluxes can be measured.
@@
 
-	Parameter estimation methods based on optimization principles are used to determine true parameter values from experimental data.
+	Parameter estimation methods based on optimization principles are used to determine parameter estimates from experimental data.
 Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:lsqopt}) to estimate the values of enzyme kinetic parameters, $\theta \in \mathbb{R}^{n_p}$, based on the measured data.
@@
 
+	Typically, not all metabolite concentrations and fluxes can be measured.
+Under these conditions, the formulation of the kinetic model and the parameter estimation problem shown in Equations (\ref{eq:kinstoich}) and (\ref{eq:lsqopt}) is different.
+These formulations are ubiquitously presented in other papers \parencite{Gadkar2005a,Rodriguez-Fernandez2006}.
+Hence, these will not be discussed in this paper.
+
 	\subsection{Structural and practical identifiability of parameters in kinetic models}\label{sec:ident_def}
@@
 
-	The parameters in $\theta$ in any nonlinear model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
+	The parameters in $\theta$ in any model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
 	\begin{align}\label{eq:stident}
@@
 	Accordingly, if parameters in $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
-So, the structural identifiability of parameters in a dynamic model helps establish the presence or absence of a relationship between the unmeasured and measured concentrations/fluxes, as well as correlations between different model parameters \textcolor{red}{(Rudiyanto Gunawan paper on model discrimination)}.
+So first, structural identifiability of parameters in a dynamic model helps establish the presence or absence correlations between different model parameters \parencite{Rodriguez-Fernandez2006}.
+Second, in instances where not all concentrations and fluxes can be measured, structural identification can also elucidate relationships between the unmeasurable and measured concentrations/fluxes.
 Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
 
-	Experimental data from many physical systems is usually noisy, and when parameters are estimated on the basis of noisy data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
-If a single unique parameter satisfying Equation (\ref{eq:stident}) can be found, then $\theta$ is said to be globally practically identifiable.
-Whereas, if parameter estimates with quantifiable uncertainties can be found, then the $\theta$ is said to be locally identifiable.
-The absence of unique parameter estimates for $\theta$ leads to practical non-identifiability.
+	Experimental data from many physical systems is usually noisy, and sometimes may not contain all requisite information pertaining to the system.
+When parameters are estimated on the basis of this data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
+If a unique parameter value satisfying Equation (\ref{eq:stident}) can be estimated, then $\theta$ is said to be globally practically identifiable.
+Whereas, if parameter estimates with quantifiable uncertainties can be found, then $\theta$ is said to be locally identifiable.
+No unique parameter estimates can be found if $\theta$ is practical non-identifiable.
 The practical identifiability of a parameter is hence contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
 
-	So, on the one hand, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data.
+	Therefore, in conclusion, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data.
 On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
 
-	\subsection{A method to determine identifiability of kinetic models of metabolism}\label{sec:ident}
-	We provide the mathematical framework for \textcolor{red}{identification} of parameters in kinetic models of metabolism in this section.
+	\subsection{A method to determine structural and practical identifiability of kinetic models of metabolism}\label{sec:ident}
+	In this section, we provide the mathematical framework for \textcolor{red}{identification} of parameters in kinetic models of metabolism.
 A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
@@
 
-	%Nonetheless, all these explanations could only be verified due to the fact that the conditions for identifiability for both $v_1$ and $v_2$ are simple and readable due to the use of the Michaelis-Menten rate law to model these fluxes.
-Instead, in addition to presence of an allosteric interaction, when the convenience kinetic rate law is used to model this activation in $v_3$, we find that the complex expressions obtained for the three parameters preclude such an in-depth insight into the nature of persistently excitable experiments to identify  $v_3$.
-Thus, the only avenue through which experiments can be designed for the identification of parameters of $v_3$ would be through our proposed methodology wherein manual curation of experimental combinations required for identifiability are not necessary.
-
-	%Based on the results we obtained in establishing the identifiability of parameters modeling all the fluxes of a small metabolic network, in the future we hope to extend this methodology to a larger metabolic network with more reactions.
-	%Apart from determining experiments that are informative enough to identify parameters for each flux in a kinetic model, experimental design can also be used to design and develop a minimal set of experiments that informative enough to estimate all parameters in a metabolic network model.
-However, there is not only a scarcity of experimental design methods in literature devoted to the purpose of designing experiments for parameter identifiability in kinetic models of metabolism, the motivation to develop such methods is not well understood.
-
 	\printbibliography	
-
-	%\clearpage
-	%\section{Tables, Figures and Figure Captions:}
-
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_V1max_v2_v5_2_exp_info}}
-	%	\caption{The contribution of different experiments types used in a combination of two experiments ($j = \{1, 2\}$) that can practically identify parameter a) $V_1^{max}$, b) $K_1^{ac}$, c) $V_2^{max}$, d) $K_2^{pep}$, e) $V_e^{max}$ and f) $K_e^{fdp}(2)$.
-The percentages reflect the fractional contribution of each experiment type towards all identifiable data combinations.}\label{fig:exp_info}
-	%\end{figure}
-
-	%\clearpage
-
-	%\section{Old Figures, Figure Captions and other miscellaneous text}	
-	%Methods and tools for structural identification of parameters based on differential algebra \parencite{Ljung1994, Audoly2001, Bellu2007} and profile likelihood \parencite{Raue2014} are available.
-
 \end{document}

### 2018-06-08 ident/docs/draft_4.tex
significant changes to intro and methods based on comments from WRC

@@
 
-	Kinetic models differ from CBMs in their use of mechanistic enzyme kinetics to model enzyme catalyzed fluxes within a metabolic network \parencite{Srinivasan2015,Saa2017}.
-The use of kinetic models requires information on the enzyme kinetic rate laws that will be used to model the fluxes within a metabolic network, as well as numerical values for the parameters used in these rate laws.
+	Kinetic models differ from CBMs in their use of mechanistic enzyme kinetics to model the fluxes within a metabolic network \parencite{Srinivasan2015,Saa2017}.
+The use of kinetic models requires information on the enzyme kinetic rate laws that will be used to model the fluxes, as well as numerical values for the parameters used in these rate laws.
 Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values play a crucial role \parencite{Srinivasan2017}.
@@
 These approaches also allow for the integration of experimentally observed in vivo data.
-ORACLE \parencite{Wang2004} and Ensemble modeling \parencite{Tran2008}, are two such examples.
-These and other Monte Carlo kinetic modeling methods have been previously reviewed \parencite{Srinivasan2015}.
+ORACLE \parencite{Wang2004} and Ensemble modeling \parencite{Tran2008} are two such examples.
+These, and other Monte Carlo kinetic modeling methods have been previously reviewed \parencite{Srinivasan2015}.
 Bayesian approaches to improve parameter estimation and quantify estimation uncertainty have also been proposed \parencite{Saa2016}.
@@
 
-	Related to the development of these methods to quantify parameter estimation uncertainty, the importance of model parameter identifiability, i.e., a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Ljung1994,Berthoumieux2013}.
+	However, the importance of model parameter identifiability, i.e., a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Rodriguez-Fernandez2006,Berthoumieux2013}.
+Unique parameter estimates from experimental data can only be obtained for structurally and practically identifiable parameters.
 
-	Briefly, unique values can only be estimated for structurally identifiable parameters.
-If a parameter cannot be uniquely estimated due to its nonlinear relationships with other parameters (redundant model parameterization), or due to the inability to measure all possible metabolite concentrations (presence of unobservable system states), then the parameter is said to be structurally non-identifiable.
-Similarly, practical identifiability of a parameter concerns with the informativeness of the experimental data used for parameter estimation, and the ability to quantify any existing uncertainty in the parameter estimates.
-If the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, and the uncertainty in parameter estimates is not quantifiable, the parameter is said to be practically non-identifiable \parencite{Ljung1994,Rodriguez-Fernandez2006}.
+	Briefly, it should be possible to obtain unique estimates for structurally identifiable parameters, just based on the model structure and parameterization, irrespective of the type and nature of experimental data available.
+If parameters cannot be uniquely estimated due to their nonlinear relationships with other parameters (redundant model parameterization), or due to the lack of a relationship between the measured concentrations/fluxes and the unmeasurable concentrations/fluxes (presence of unobservable system states) modeled by the parameters, then these parameters are said to be structurally non-identifiable.
+
+	If the experimental data used for parameter estimation is informative so as to facilitate obtaining unique parameter estimates, or the quantification of uncertainties in the parameter estimates, then the parameters are said to be practically identifiable.
+However, if the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, or the uncertainty in parameter estimates is not quantifiable, the parameters are said to be practically non-identifiable.
 
 	Authors have proposed ways to assess parameter identifiability by proposing approximate kinetic models of metabolism that utilize empirical enzyme kinetic rate laws with parameters that have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}.
-Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Ljung1994,Nikerel2009,Berthoumieux2013,Raue2014}.
+Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Nikerel2009,Berthoumieux2013,Raue2014}.
 \textcite{Chis2011b} provide a comparison of the computational performance and implementation aspects of some of these methods.
@@
 Some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, while some others suffer from computational limitations when applied to kinetic models of large metabolic networks \parencite{Gadkar2005a,Raue2014}.
+	Assessing the identifiability (structural or practical) of parameters using any of the aforementioned or other methods is done under the assumption that only dynamic time course concentration data is available and can be used for parameter estimation.
+The ability to use steady state experimental data to assess parameter identifiability is rarely addressed, and very few methodology have been proposed to fill this gap \textcolor{red}{(paper that proposes both steady state and dynamic data use for a priori experimental design resulting in reduced model sizes)}
 
-	Most of the aforementioned methods assess identifiability (structural or practical) in the presence of dynamic time course concentration data.
-In this paper, we show that the identifiability of parameters for each individual reaction in kinetic models of metabolism can be done using just steady state fluxomics, metabolomics and proteomics data, and propose a methodology to do so.
-We also demonstrate how this method can also facilitate the design of experiments for generating the necessary steady state data to estimate unique parameter values for all reaction fluxes in a metabolic network.
+	In this paper, we show that the identifiability of most parameters in a metabolic network can be done for each reaction individually using steady state fluxomics, metabolomics and proteomics data, and propose a methodology to do so.
+We also demonstrate how this method can facilitate the design of experiments for generating the necessary steady state data to estimate unique parameter values for all reaction fluxes in a metabolic network.
 In doing so, we are able to propose the number and types of perturbations that will provide the most useful data for parameter estimation.
-We also assess the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks using steady state data.
+We also comment on the identifiability of different enzyme kinetic rate laws that are typically used to model fluxes in metabolic networks using steady state data.
 We illustrate our methodology using a small metabolic network model of glucoeneogenesis in \textit{Escherichia coli} \parencite{Kotte2014, Srinivasan2017} under the assumption that all intracellular metabolite concentrations and fluxes can be measured.
@@
 	\section{Methods}\label{sec:methods}
-	In Section \ref{sec:kinetic_model} we present the typical structure of a kinetic model of metabolism and a preliminary description of the least squares method by which the model parameters are estimated from experimental data.
-In Section \ref{sec:ident_def} we provide formal mathematical definitions for structural and practical identifiability.
+	In Section \ref{sec:ident_def} we present typical structures of kinetic models of metabolism and pertinent formal mathematical definitions for structural and practical identifiability.
 The computer-algebra system based method to assess identifiability that we have developed is described in Section \ref{sec:ident} that follows.
@@
 
-	\subsection{Parameter estimation for kinetic models of metabolism}\label{sec:kinetic_model}
-	In kinetic models of metabolism, ordinary differential equations (ODE) are used to express the rate of change of metabolite concentrations ($x\in\mathbb{R}^{n_x}$) as a function of the reaction fluxes ($v\in\mathbb{R}^{n_r}$) in the metabolic network (Equation \ref{eq:kinstoich}).
+	%\subsection{Parameter estimation for kinetic models of metabolism}\label{sec:kinetic_model}	
+	%Parameter estimation methods based on optimization principles are used to determine parameter estimates from experimental data.
+Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:lsqopt}) to estimate the values of enzyme kinetic parameters, $\theta \in \mathbb{R}^{n_p}$, based on the measured data.
+	%\begin{subequations}\label{eq:lsqopt}
+	%	\begin{align}
+	%	\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{n_x + n_r}\sum_{l=1}^{n_t}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2\\
+	%	&\theta_l \le \theta \le \theta_u
+	%	\end{align}
+	%\end{subequations}
+	%Here $y \in \mathbb{R}^{n_x + n_r}$ is the combined vector of concentrations ($x\in \mathbb{R}^{n_x}$) and fluxes ($v \in \mathbb{R}^{n_r}$), at each time point $l={1, 2, ..., n_t}$.
+The minimization of the least squares error between the measured ($y^*$) and modeled ($y$) concentrations and fluxes, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration and flux, at each time point $l={1, 2, ..., n_t}$, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem.
+The least squares parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).
+
+	\subsection{Structural and practical identifiability of parameters in kinetic models}\label{sec:ident_def}
+	Ordinary differential equations (ODE) are used in kinetic models of metabolism to express the rate of change of metabolite concentrations ($x\in\mathbb{R}^{n_x}$) as a function of the reaction fluxes ($v\in\mathbb{R}^{n_r}$) in the metabolic network (Equation \ref{eq:kinstoich}).
 The matrix $\mathbf{S}\in\mathbb{R}^{n_x \text{x} n_r}$ in Equation (\ref{eq:kinstoich}a) defines the stoichiometric relationship between the fluxes and the concentrations of the metabolic network.
@@
 
-	Parameter estimation methods based on optimization principles are used to determine parameter estimates from experimental data.
-Under the assumption that all intracellular metabolite concentrations and fluxes can be measured, a parameter estimation problem can be formulated as a nonlinear programming problem (Equation \ref{eq:lsqopt}) to estimate the values of enzyme kinetic parameters, $\theta \in \mathbb{R}^{n_p}$, based on the measured data.
-	\begin{subequations}\label{eq:lsqopt}
-		\begin{align}
-		\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{n_x + n_r}\sum_{l=1}^{n_t}\left(\frac{y_{kl}^*-y_{kl}}{\sigma_{kl}^*}\right)^2\\
-		&\theta_l \le \theta \le \theta_u
-		\end{align}
-	\end{subequations}
-	Here $y \in \mathbb{R}^{n_x + n_r}$ is the combined vector of concentrations ($x\in \mathbb{R}^{n_x}$) and fluxes ($v \in \mathbb{R}^{n_r}$), at each time point $l={1, 2, ..., n_t}$.
-The minimization of the least squares error between the measured ($y^*$) and modeled ($y$) concentrations and fluxes, weighted by the variance in the experimental data $\sigma_{kl}^*$ for each concentration and flux, at each time point $l={1, 2, ..., n_t}$, is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem.
-The least squares parameter values are determined within fixed upper ($\theta_u$) and lower ($\theta_l$) bounds (Equation \ref{eq:lsqopt}b).
-
-	Typically, not all metabolite concentrations and fluxes can be measured.
-Under these conditions, the formulation of the kinetic model and the parameter estimation problem shown in Equations (\ref{eq:kinstoich}) and (\ref{eq:lsqopt}) is different.
-These formulations are ubiquitously presented in other papers \parencite{Gadkar2005a,Rodriguez-Fernandez2006}.
-Hence, these will not be discussed in this paper.
-
-	\subsection{Structural and practical identifiability of parameters in kinetic models}\label{sec:ident_def}
-	In the Introduction, we briefly mentioned that the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
+	In the Introduction, we briefly mentioned that the ability to estimate unique values for parameters $\theta$ from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}.
 Below, we provide a formal definition of structural and practical identifiability of parameters.
 
-	The parameters in $\theta$ in any model (Equation \ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two values of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
+	If we assume that all concentrations ($x\in \mathbb{R}^{n_x}$) and fluxes ($v \in \mathbb{R}^{n_r}$) can be measured, and accordingly classify them as model outputs $y \in \mathbb{R}^{n_x + n_r}$, the parameters $\theta$ in Equation (\ref{eq:kinstoich}) are said to be structurally identifiable if, for an input-output mapping defined by $y = \Phi(\theta,u)$ for at least one input function $u$, any two sets of parameters $\theta_1$ and $\theta_2$ satisfy the relationship in Equation (\ref{eq:stident}):
 	\begin{align}\label{eq:stident}
@@
 	\end{align}
-	Accordingly, if parameters in $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
+	Accordingly, if the parameters $\theta$ have a unique value, a finite number of non-unique values or an infinite number of values for all input functions, they are said to be structurally globally identifiable, locally identifiable or non-identifiable, respectively.
 So first, structural identifiability of parameters in a dynamic model helps establish the presence or absence correlations between different model parameters \parencite{Rodriguez-Fernandez2006}.
-Second, in instances where not all concentrations and fluxes can be measured, structural identification can also elucidate relationships between the unmeasurable and measured concentrations/fluxes.
-Consequently, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter.
 
-	Experimental data from many physical systems is usually noisy, and sometimes may not contain all requisite information pertaining to the system.
+	Typically, not all metabolite concentrations and fluxes can be measured.
+Under these conditions, the formulation of the kinetic model (Equation \ref{eq:kinstoich}) used for parameter estimation purposes slightly varies, as shown in Equation (\ref{eq:output}) below \parencite{Gadkar2005a,Rodriguez-Fernandez2006}.
+	\begin{subequations}\label{eq:output}
+		\begin{align}
+		\dot{x} = \mathbf{S}v\\
+		v = f(x, \theta, u)\\
+		y = g(x, \theta, u)
+		\end{align}
+	\end{subequations}
+	Here the output $y$ in Equation (\ref{eq:output}c) represents concentrations and fluxes that can actually be measured and used for parameter estimation.
+Moreover, $\theta$ is also augmented with additional parameters that connect the measurable quantities with the unmeasurable quantities.
+In these instances, where not all concentrations and fluxes can be measured, the structural identifiability of parameters $\theta$ can also elucidate the presence or absence of correlations between the unmeasurable (unobservable states) and measured concentrations/fluxes (observable outputs).
+However, the definition of structural identifiability pertaining to the relationship between the observable outputs and unobservable states is not relevant to the materials presented in this paper.
+
+	Experimental data (concentration or fluxes) from many physical systems is usually noisy, and sometimes may not contain all requisite information pertaining to the system.
 When parameters are estimated on the basis of this data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
-If a unique parameter value satisfying Equation (\ref{eq:stident}) can be estimated, then $\theta$ is said to be globally practically identifiable.
+If unique parameter values satisfying Equation (\ref{eq:stident}) can be estimated, then $\theta$ is said to be globally practically identifiable.
 Whereas, if parameter estimates with quantifiable uncertainties can be found, then $\theta$ is said to be locally identifiable.
 No unique parameter estimates can be found if $\theta$ is practical non-identifiable.
-The practical identifiability of a parameter is hence contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
 
-	Therefore, in conclusion, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data.
+	In summary, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter, and the practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
+	Therefore, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data.
 On the other hand, establishing practical identifiability of parameters in any model helps design experiments that are minimal, informative and useful for parameter estimation.
@@
 Consequently, the data from that combination of experiments is considered non-informative.
-When this analysis is repeated for multiple combinations of steady state data from the 21 different in silico experiments, we can determine the number of different experimental data combinations that can practically identify each parameter in each flux (Figure \ref{fig:ident}).
+When this analysis is repeated for multiple combinations of steady state data from the 21 different in silico experiments, we can determine the number of different experimental data combinations that can practically identify each parameter in each flux (Figure \ref{fig:ident}a - c).
 
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=0.6\textwidth,height=0.4\textheight,keepaspectratio]{figures/figure1/v1_k1cat_v2_v5_2_ck_ident}}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.9\textheight,keepaspectratio]{figures/figure1/v1_kcat_v2_v5_2_exp_info}}
 		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, and c) $v_5$ when there is no noise in the input experimental data.
@@
 $v_1$, $v_2$ and $v_5$ require data from two experiments for analysis.
-The contribution of different experiment type towards identifying each parameter is shown in the spider plots.}\label{fig:ident}
+The contribution of different experiment type towards identifying each parameter is shown in the spider plots for d) $v_1$, e) $v_2$ and f) $v_5$.
+The choice of perturbation chosen for the first experiment (blue) determines the choice of the second perturbation (green) required for identifying the parameters.}\label{fig:ident}
 	\end{figure}
@@
 
-	In Figure \ref{fig:ident} we show the number of experimental data combinations that are capable of identifying each parameter, and consequently, the degree of identifiability of each parameter (percentage experimental data combinations that are capable of identifying each parameter) in flux $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b) and $v_5$ (Figure \ref{fig:ident}c).
+	In Figure \ref{fig:ident}a-c we show the number of experimental data combinations that are capable of identifying each parameter, and consequently, the degree of identifiability of each parameter (percentage experimental data combinations that are capable of identifying each parameter) in flux $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b) and $v_5$ (Figure \ref{fig:ident}c).
 The degree of identifiability for $V_1^{max}$ and $K_1^{ac}(ne)$ in $v_1$ is shown in Supplementary Figures \ref{fig:v1_v1max_ck_ident}, and the degree of identifiability for both closed-form expressions of the three parameters of $v_3$ are shown in Supplementary Figure \ref{fig:v3_1_ck_ident}.
@@
 
-	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure2/v1_k1cat_v2_v5_2_ck_exp_info}}
-		\caption{Frequency of each type of perturbation experiment in data sets that can identify parameters in fluxes a) $v_1$, b) $v_2$ and c)$v_5$.
+	%\begin{figure}[!tbhp]
+	%	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure2/v1_k1cat_v2_v5_2_ck_exp_info}}
+	%	\caption{Frequency of each type of perturbation experiment in data sets that can identify parameters in fluxes a) $v_1$, b) $v_2$ and c)$v_5$.
 The frequencies are represented as a percentage of the total number of experiments present within all the data sets that can practically identify the parameters.
 The frequency of experiments that occur as the first experiment in a combination of two experiments are shown as a blue surface, and the frequency of experiments that occur as the second experiment in the combination are shown as a green surface.}\label{fig:expinfo}
-	\end{figure}
+	%\end{figure}
 
 	From the above example we can summarize that identification of individual fluxes within a metabolic network necessitates a careful consideration of experiments such that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}.
-To facilitate the design of experiments based on their ability to satisfy requirements for practical identifiability of parameters, we determine the occurrence of each type of steady state perturbation experiment within combinations that can practically identify each parameter (Figure \ref{fig:expinfo}, Supplementary Figures \ref{fig:v1_v1max_ck_ident},  \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).
+To facilitate the design of experiments based on their ability to satisfy requirements for practical identifiability of parameters, we determine the occurrence of each type of steady state perturbation experiment within combinations that can practically identify each parameter (Figure \ref{fig:ident}d-f, Supplementary Figures \ref{fig:v1_v1max_ck_ident},  \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).
 So, with our proposed methodology it is possible to identify the types of perturbation experiments that would be informative for identifying each parameter in each flux with steady state concentration and flux data.
@@
 	As mentioned in Supplementary Section \ref{sec:experiments}, we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Supplementary Table \ref{tab:pval}).
-In Figures \ref{fig:expinfo}a, \ref{fig:expinfo}b and \ref{fig:expinfo}c, the contribution from different experiment types for identifying parameters in $v_1$, $v_2$ and $v_5$ are respectively shown as spider plots.
+In Figures \ref{fig:ident}d-f, the contribution from different experiment types for identifying parameters in $v_1$, $v_2$ and $v_5$ are respectively shown as spider plots.
 
-	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations for $v_1$ in comparison to the other types of experiments (Figure \ref{fig:expinfo}a and Supplementary Figure \ref{fig:v1_v1max_ck_ident}).
+	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations for $v_1$ in comparison to the other types of experiments (Figure \ref{fig:ident}d and Supplementary Figure \ref{fig:v1_v1max_ck_ident}).
 This is in agreement with the condition for identifiability that we discussed earlier (Equations \ref{eq:v1_k1cat} and \ref{eq:v1_par}).
@@
 
-	With $v_2$, we see that the enzyme perturbations as well as the acetate perturbation experiments have similar contributions towards datasets that can identify $v_2$ (Figure \ref{fig:expinfo}b).
+	With $v_2$, we see that the enzyme perturbations as well as the acetate perturbation experiments have similar contributions towards datasets that can identify $v_2$ (Figure \ref{fig:ident}e).
 This also supports our arguments made earlier with regards to the identifiability conditions for $v_2$, and the reasons for the difference in the type of experiments that are informative between $v_1$ and $v_2$.
@@
 
-	We can also extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_5$ (Figure \ref{fig:expinfo}c), or determining physiologically relevant parameter values for structurally locally identifiable parameters of $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).
+	We can also extend these observations to justify the observed contribution of experiments towards identifying parameters for $v_5$ (Figure \ref{fig:ident}f), or determining physiologically relevant parameter values for structurally locally identifiable parameters of $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).
 
-	In all of the above scenarios for $v_1$ (Figure \ref{fig:expinfo}a), $v_2$ (Figure \ref{fig:expinfo}b) and $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar.
+	In all of the above scenarios for $v_1$ (Figure \ref{fig:ident}d), $v_2$ (Figure \ref{fig:ident}e) and $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar.
 Hence, the green/blue/yellow surfaces in the spider plots are superimposed upon each other.
-This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:expinfo}c).
-However, this is not the case for $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:expinfo}c).
+This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:ident}f).
+However, this is not the case for $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:ident}f).
 We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$.
-Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figures \ref{fig:ident}c and \ref{fig:expinfo}c), the choice and number of experiments required to identify $v_5$ completely hinges upon the identifiability of $K_e^{fdp}$ from the chosen experiments.
+Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figures \ref{fig:ident}c and f), the choice and number of experiments required to identify $v_5$ completely hinges upon the identifiability of $K_e^{fdp}$ from the chosen experiments.
 

### 2018-06-08 ident/python2/ss-ident/plot_ident_results.py
change color of means to black and meadians to red in violin plots

@@ -43,6 +43,9 @@ def plot_on_axis_object_violin(axis_object, distribution_data):
     """create violin plot of distributions"""
     # axis_object = plt.Subplot(fig_object, grid_object)
     violin_object = axis_object.violinplot(distribution_data, showmeans=True, showmedians=True)
+    # set meadian and mean line colors
+    violin_object['cmeans'].set_color('k')
+    violin_object['cmedians'].set_color('r')
 
     # Remove top axes and right axes ticks
     axis_object.get_xaxis().tick_bottom()

### 2018-06-08 ident/python2/ss-ident/v1_Vmax_ck_script.py
change file names to be platform independent

@@ -6,6 +6,7 @@ from plot_ident_results import identifiability_plot
 from plot_ident_results import parameter_values_plot
 from plot_ident_results import validation_plot
 from names_strings import true_parameter_values
+import os.path
 
 
 # create data for identifiability analysis
@@ -13,16 +14,14 @@ from names_strings import true_parameter_values
 # create_data_for_flux(flux_id='v1', noise=0, number_samples=1)
 
 # extract experimental data from file
-new_data_file_name = 'C:\Users\shyam\Documents\Courses\CHE1125Project\IntegratedModels' \
-                     '\ident\python2\ss-ident\exp_v1_2_experiments'
+new_data_file_name = os.path.join(os.getcwd(), 'exp/exp_v1_2_experiments')
 index_labels = ['sample_name', 'data_set_id', 'experiment_id']
 arranged_data_df = retrieve_experimental_data_from_file(data_file_name=new_data_file_name,
                                                         multi_index_label=index_labels)
 
 # perform identifiability when v1 parameters with combination of 2 experiments
 # print('Practical Identifiability Analysis of v1 with 2 parameters: V1max and K1ac\n')
-storage_file_name = 'C:\Users\shyam\Documents\Courses\CHE1125Project\IntegratedModels' \
-                     '\ident\python2\ss-ident\ident_v1_vmax'
+storage_file_name = os.path.join(os.getcwd(), 'ident/ident_v1_vmax')
 # choose identifiability functions to test
 # ident_fun_choice = [0]
 # # test identifiability and store data to file
@@ -38,8 +37,7 @@ all_parameter_info = process_ident(ident_df, arranged_data_df)
 
 # run dynamic simulations to obtain ss data based on estimated parameter values
 # get info from data sets that identify all 3 parameters
-validation_file_name = 'C:\Users\shyam\Documents\Courses\CHE1125Project\IntegratedModels\ident\python2' \
-                       '\ss-ident\ident_validate_v1_vmax'
+validation_file_name = os.path.join(os.getcwd(), 'validate\ident_validate_v1_vmax')
 from process_ident_data import get_parameter_value
 validation_info = get_parameter_value(all_parameter_info, ident_df)
 # # get default parameter values
@@ -61,8 +59,7 @@ validate_index_labels = ['estimate_id', 'sample_name', 'data_set_id', 'experimen
 validate_df = retrieve_experimental_data_from_file(data_file_name=validation_file_name,
                                                    multi_index_label=validate_index_labels)
[…]

### 2018-06-08 ident/docs/sifigures.tex
change external document
switched figures and added figure captions

@@
 
-\externaldocument{draft_2} % reference to existing external document
+\externaldocument{draft_3} % reference to existing external document
 
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=0.6\textwidth,height=0.2\textheight, keepaspectratio]{figures/figure6/v1_v1max_ck_parameter_values}}
-	\caption{}\label{fig:v1_v1max_ck_values}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/figure6/v1_kcat_ck_parameter_dist}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_1$ when there is no noise in the input experimental data.
+The top panel shows a violin plot of the distribution of parameter values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.}\label{fig:v1_kcat_ck_values}
 \end{figure}
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v1_v1max_ck_ident_exp_info}}
-	\caption{}\label{fig:v1_v1max_ck_ident}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/figure6/v1_vmax_ck_parameter_values}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_1$, which is described using as a Michaelis-Menten model, without using the enzyme concentration $E$ for identification when there is no noise in the input experimental data.
+Here $V_1^{max} = k_1^{cat}E$.
+The top panel shows a violin plot of the distribution of parameter values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.}\label{fig:v1_v1max_ck_values}
 \end{figure}
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure1/v3_ck_1_2_ident_exp_info}}
-	\caption{}\label{fig:v3_1_ck_ident}
+	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure1/v1_vmax_exp_info}}
+	\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in flux $v_1$, which is described using as a Michaelis-Menten model, without using the enzyme concentration $E$ for identification when there is no noise in the input experimental data.
+Here $V_1^{max} = k_1^{cat}E$.
+The percentage of total combinations of experimental data used for analysis (240 for $v_1$) that can identify each parameter is also specified.
+$v_1$ requires data from two experiments for analysis.
+The contribution of different experiment type towards identifying each parameter is shown in the spider plots for b) $v_1$.
+The choice of perturbation chosen for the first experiment (blue) determines the choice of the second perturbation (green) required for identifying the parameters.}\label{fig:v1_v1max_ck_ident}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure1/v3_ck_exp_info}}
+	\caption{a) The number of data combination from 21 different in silico experiments that can practically identify each parameter in flux $v_3$ when there is no noise in the input experimental data.
+The percentage of total combinations of experimental data used for analysis (3360) that can identify each parameter is also specified.
+$v_3$ required data from three experiments for analysis.
+The contribution of different experiment type towards identifying each parameter is shown in the spider plots for b) $V_3^{max}$, c) $K_3^{fdp}$ and d) $K_3^{pep}$.
+The choice of perturbation chosen for the first experiment (blue) determines the choice of the second (green) and the third perturbation (yellow) required for identifying the parameters.}\label{fig:v3_1_ck_ident}
 \end{figure}

### 2018-06-08 ident/docs/draft_4.tex
changed figure for parameter distribution

@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=.8\textwidth,height=.6\textheight,keepaspectratio]{figures/figure6/v1_k1cat_v2_v5_2_v3_ck_1_parameter_values}}
+		\centering{\includegraphics[width=1.0\textwidth,height=.8\textheight,keepaspectratio]{figures/figure6/v1_kcat_v2_v5_2_v3_1_parameter_dist}}
 		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, c) $v_5$, and d)$v_3$.
@@
 Only one of the two roots for $v_3$ is shown in panel d.
-The estimated data for the second root has a similar distribution to that of the first root and is shown in the Supplementary Information.
+The estimated data for the second root has a similar distribution to that of the first root.
 Data is generated using the Convenience Kinetic model for allosteric regulation for  $v_3$.}\label{fig:ident_values}

### 2018-06-09 ident/docs/sifigures.tex
new figures for experimental data dist for ck model

@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/figure6/v1_vmax_ck_parameter_values}}
+	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure6/v1_vmax_ck_parameter_values}}
 	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_1$, which is described using as a Michaelis-Menten model, without using the enzyme concentration $E$ for identification when there is no noise in the input experimental data.
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/mwc_ck_experimental_data}}
-	\caption{Variability in the metabolite concentrations and fluxes between different in silico experiments from which data is used for identifiability analysis.
-a) Metabolite (\textit{pep}, \textit{fdp}) and enzyme concentration (\textit{E}), and b) flux data ($v_1$, $v_2$, $v_3$, $v_4$, $v_5$ and $v_6$) are generated using the Monod-Wyman-Changeaux model for $v_3$.
-The Convenience Kientic rate law model is used for $v_3$ to generate in silico c) metabolite and enzyme concentration, and d) flux data through enzyme expression parameter and substrate concentration perturbations.}\label{fig:experimental_data_dist}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/figure6/exp_concentration_dist}}
+	\caption{Variability in the metabolite concentrations between different in silico experiments from which data is used for identifiability analysis.
+The top panel shows a violin plot of the distribution of concentration values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.
+Data are are generated using the Convenience Kinetics model for $v_3$.
+}\label{fig:experimental_data_dist}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/figure6/exp_flux_dist}}
+	\caption{Variability in the fluxes between different in silico experiments from which data is used for identifiability analysis.
+The top panel shows a violin plot of the distribution of flux values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.
+Data are are generated using the Convenience Kinetics model for $v_3$.}\label{fig:exp_flux_dist}
 \end{figure}

### 2018-06-11 ident/python2/ss-ident/v3_ck_noise_script.py
v3 ident from noisy data using 500 samples with new script

@@ -1,20 +1,20 @@
-import numpy as np
 from create_experiment_data import retrieve_experimental_data_from_file
 from process_ident_data import process_ident
-from process_ident_data import get_parameter_value
+from validate_estimation import process_validation_info
+from plot_ident_results import exp_info_plot
+from plot_ident_results import identifiability_plot
+from plot_ident_results import parameter_values_plot
+from plot_ident_results import validation_plot
 from names_strings import true_parameter_values
-from validate_estimation import validate_model
-from plot_ident_results import data_utility_plot
-from plot_ident_results import plot_parameter_values
+import os.path
 
 
 # create data for identifiability analysis
-# from create_experiment_data import create_data_for_flux
-# create_data_for_flux(flux_id='v3', noise=1, number_samples=5)
+from create_experiment_data import create_data_for_flux
+create_data_for_flux(flux_id='v3', noise=1, number_samples=500)
 
 # extract data from file
-new_data_file_name = 'C:\Users\shyam\Documents\Courses\CHE1125Project\IntegratedModels' \
-                     '\ident\python2\ss-ident\exp_v3_3_experiments_noise_5_samples'
+new_data_file_name = os.path.join(os.getcwd(), 'exp/exp_v3_3_experiments_noise_500_samples')
 index_labels = ['sample_name', 'data_set_id', 'experiment_id']
 arranged_data_df = retrieve_experimental_data_from_file(data_file_name=new_data_file_name,
                                                         multi_index_label=index_labels)
@@ -22,14 +22,13 @@ arranged_data_df = retrieve_experimental_data_from_file(data_file_name=new_data_
 # perform identifiability when v3 parameters are written for root (1)
 # get combination of 3 experiments and perform identifiability on all fluxes that require 3 data sets
 # print('Practical Identifiability Analysis of v3 with 3 parameters: V3max, K3fdp and K3pep \n')
-storage_file_name = 'C:\Users\shyam\Documents\Courses\CHE1125Project\IntegratedModels' \
-                     '\ident\python2\ss-ident\ident_v3_noise_root_1'
+storage_file_name = os.path.join(os.getcwd(), 'ident/ident_v3_noise_root_1')
 # choose identifiability functions to test
 ident_fun_choice = [0]
 # test identifiability and store data to file
-# from kotte_model import flux_ident_3_data_combination
[…]

### 2018-06-11 ident/docs/draft_4.tex
remove tikz and subfigure from preamble
changes to figures and addition of figure captions for sifigures

@@
 %\usepackage[CaptionAfterwards]{fltpage}
-\usepackage{subfigure}
 \usepackage{lineno}
@@
 \usepackage{tikz}
-\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;} 
 \usepackage[margin=1.0in]{geometry}
@@
 If unique parameter values satisfying Equation (\ref{eq:stident}) can be estimated, then $\theta$ is said to be globally practically identifiable.
-Whereas, if parameter estimates with quantifiable uncertainties can be found, then $\theta$ is said to be locally identifiable.
+Whereas, if parameter estimates with quantifiable uncertainties that satisfy Equation (\ref{eq:stident}) can be found, then $\theta$ is said to be locally identifiable.
 No unique parameter estimates can be found if $\theta$ is practical non-identifiable.
@@
 	\subsection{A method to determine structural and practical identifiability of kinetic models of metabolism}\label{sec:ident}
-	In this section, we provide the mathematical framework for \textcolor{red}{identification} of parameters in kinetic models of metabolism.
-A summary of the methodology in the form of a flow diagram is shown in Figure \ref{fig:ident-flowchart}.
-As indicated in Figure \ref{fig:ident-flowchart}a, the first step involves the construction of the kinetic model (Equation \ref{eq:kinstoich}) of the metabolic network with $n_r$ reaction fluxes.
+	In this section, we provide the mathematical framework for determining the identifiability of parameters in kinetic models of metabolism.
+The methodology in shown as a flow diagram in Figure \ref{fig:ident-flowchart}.
 
@@
 
-	For each flux $v_i$, $i={1, 2, ..., n_r}$, in the kinetic model, let $\mathbf{\theta} \in \mathbb{R}^p$ in Equation (\ref{eq:kinstoich}b).
+	As indicated in Figure \ref{fig:ident-flowchart}a, the first step involves the construction of the kinetic model (Equation \ref{eq:kinstoich}) of the metabolic network with $n_r$ reaction fluxes.
+For each flux $v_i$, $i={1, 2, ..., n_r}$, in the kinetic model, let $\mathbf{\theta} \in \mathbb{R}^{n_p}$ in Equation (\ref{eq:kinstoich}b).
 If data from $n_E$ experiments is available for the chosen metabolic network, as stated earlier, for each experiment $j = {1, 2, ..., n_E}$, we assume that all metabolite concentrations ($x\in\mathbb{R}^{n_x}$) and reaction fluxes ($v\in\mathbb{R}^{n_r}$) are measurable.
@@
 
-	In order to establish the practical identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method.
+	In order to establish the identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method.
 The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta} \in \mathbb{R}^p$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
-This is done by first selecting a combination of $p\le n_E$ experimental data.
-The fluxes and concentrations from $p$ different experiments are then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^p$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
+This is done by first selecting a combination of $n_p\le n_E$ experimental data.
+The fluxes and concentrations from $n_p$ different experiments are then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^{n_p}$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
 	\begin{align}\label{eq:nonlineq}
@@
 	Here, $v_{i,j}$ refers to the flux $v_i$ obtained from experiment $j$.
-$\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^p$, whose elements are denoted by $\theta_k$.
+$\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^{n_p}$, whose elements are denoted by $\theta_k$.
 
 	Each equation in (\ref{eq:nonlineq}), indicated by the index $j$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for each $v_i$, $i={1, 2, ..., n_r}$, described in Equation (\ref{eq:kinstoich}b), written for concentrations ($\mathbf{x}_j$, $\mathbf{u}_j$) and fluxes ($v_{i,j}$) obtained from experiment $j$.
-Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^p$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^p$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
+Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^{n_p}$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^{n_p}$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
 Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from $p$ experiments denoted by the index $j = {1, 2, ..., p}$.
@@
 		\begin{align}
-		\underset{\theta, \varepsilon}{min} & \sum_{j=1}^{p} \left|\varepsilon_j\right|\\
-		& v_{i, j} - f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) \le \varepsilon_j && j=\{1, 2, ..., p\}\\
-		& \theta_l \le \theta_j \le \theta_u && j=\{1, 2, ..., p\}
+		\underset{\theta, \varepsilon}{min} & \sum_{j=1}^{n_p} \left|\varepsilon_j\right|\\
+		& v_{i, j} - f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) \le \varepsilon_j && j=\{1, 2, ..., n_p\}\\
+		& \theta_l \le \theta_j \le \theta_u && j=\{1, 2, ..., n_p\}
 		\end{align}
@@
 
-	The formulation of the above optimization problem is similar to the least squares problem shown earlier in Equation (\ref{eq:lsqopt}).
 	Notice that the parameter values are bounded above and below by $\theta_u$ and $\theta_l$, respectively.
@@
 	The identifiability of each parameter based on each experiment with index $j = {1, 2, ..., n_E}$ is established based on the methodology described in Section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in Section \ref{sec:example}.
-Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^p$ (Figure \ref{fig:ident-flowchart}b), then at least one of the $p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
+Subsequently, for any flux $v_i$, and for any combination of $n_p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., n_p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^{n_p}$ (Figure \ref{fig:ident-flowchart}b), then at least one of the $n_p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
 Consequently, the corresponding experiment cannot be used for estimating parameter $\theta_k$, and needs to be discarded from the set of all necessary experiments.
@@
 This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
-In doing so, we can arrive at a set of $p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
+In doing so, we can arrive at a set of $n_p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
 Note that if none of the $n_E$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n_E+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=.8\textheight,keepaspectratio]{figures/figure6/v1_kcat_v2_v5_2_v3_1_parameter_dist}}
+		\centering{\includegraphics[width=1.0\textwidth,height=.8\textheight,keepaspectratio]{figures/parameter_dist/v1_kcat_v2_v5_2_v3_1_parameter_dist}}
 		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, c) $v_5$, and d)$v_3$.
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=0.8\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v3_var_1_2_ck_parameter_value}}
+		\centering{\includegraphics[width=0.8\textwidth,height=0.8\textheight, keepaspectratio]{figures/parameter_dist/v3_var_parameter_dist}}
 		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$.
@@
 	\begin{figure}[!tbhp]
-		\centering{\includegraphics[width=1.0\textwidth,height=0.9\textheight,keepaspectratio]{figures/figure1/v1_kcat_v2_v5_2_exp_info}}
+		\centering{\includegraphics[width=1.0\textwidth,height=0.9\textheight,keepaspectratio]{figures/ident_exp_info/v1_kcat_v2_v5_2_exp_info}}
 		\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in fluxes a) $v_1$, b) $v_2$, and c) $v_5$ when there is no noise in the input experimental data.

### 2018-06-11 ident/docs/sifigures.tex
changes to figures and addition of figure captions for sifigures

@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/figure6/v1_kcat_ck_parameter_dist}}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/parameter_dist/v1_kcat_ck_parameter_dist}}
 	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_1$ when there is no noise in the input experimental data.
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure6/v1_vmax_ck_parameter_values}}
+	\centering{\includegraphics[width=0.8\textwidth,height=0.8\textheight, keepaspectratio]{figures/parameter_dist/v2_parameter_dist}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_2$ when there is no noise in the input experimental data.
+The top panel shows a violin plot of the distribution of parameter values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.}\label{fig:v2_ck_values}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/parameter_dist/v3_parameter_dist}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$ when there is no noise in the input experimental data, and $v_3$ is expressed using the Convenience Kinetics model.
+The top panel shows a violin plot of the distribution of parameter values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.}\label{fig:v3_ck_values}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/parameter_dist/v5_parameter_dist}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_5$ when there is no noise in the input experimental data.
+The top panel shows a violin plot of the distribution of parameter values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.}\label{fig:v5_ck_values}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/parameter_dist/v1_vmax_ck_parameter_values}}
 	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_1$, which is described using as a Michaelis-Menten model, without using the enzyme concentration $E$ for identification when there is no noise in the input experimental data.
@@
 The means are indicated by a black line and the median values are indicated by a red line.
-The bottom panels show the corresponding distributions as histograms.}\label{fig:v1_v1max_ck_values}
+The bottom panels show the corresponding distributions as histograms.}\label{fig:v5_v1max_ck_values}
 \end{figure}
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/figure1/v1_vmax_exp_info}}
+	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/ident_exp_info/v1_vmax_exp_info}}
 	\caption{The number of data combination from 21 different in silico experiments that can practically identify each parameter in flux $v_1$, which is described using as a Michaelis-Menten model, without using the enzyme concentration $E$ for identification when there is no noise in the input experimental data.
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure1/v3_ck_exp_info}}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/ident_exp_info/v3_ck_exp_info}}
 	\caption{a) The number of data combination from 21 different in silico experiments that can practically identify each parameter in flux $v_3$ when there is no noise in the input experimental data.
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure1/v3_var_1_2_ck_ident_exp_info}}
-	\caption{}\label{fig:v3_var_1_2_ck_ident}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/parameter_dist/v3_var_k3fdp_parameter_dist}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$ (with constant $K_3^{pep}$) when there is no noise in the input experimental data.
+The top panel shows a violin plot of the distribution of parameter values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.}\label{fig:v3_var_k3fdp_values}
 \end{figure}
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=0.6\textwidth,height=0.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
-	\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/parameter_dist/v3_var_k3pep_parameter_dist}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$ (with constant $K_3^{fdp}$) when there is no noise in the input experimental data.
+The top panel shows a violin plot of the distribution of parameter values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.}\label{fig:v3_var_k3pep_values}
 \end{figure}
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_mwc_ck_data_utility}}
-	\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
-Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$.
-The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, c) three and d) two experiments.
-The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
-A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, 3360 combinations are used to practically identify c) $v_3$, and 420 combinations are used to analyze the identifiability of d) $v_5$.
-Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
+	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/ident_exp_info/v3_var_exp_info}}
+	\caption{}\label{fig:v3_var_1_2_ck_ident}
 \end{figure}
@@
 Data are are generated using the Convenience Kinetics model for $v_3$.}\label{fig:exp_flux_dist}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=0.6\textwidth,height=0.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
+	\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_mwc_ck_data_utility}}
+	\caption{Utility of experimental data combinations on the basis of their ability to identify the most number of parameters.
+Information is shown for parameters modeling fluxes a) $v_1$, b) $v_2$, c) $v_3$ and d) $v_5$.
+The total number of combinations of experimental data is shown in the vertical axis and the horizontal axis shows the total possible number of parameters that can be identified by data from combinations of a) two, b) two, c) three and d) two experiments.
+The percentages shown in the plots represent the fraction of the total combinations used to test identifiability of parameters for a given flux.
+A total of 240 data combinations are used for identifiability analysis for a) $v_1$ and b) $v_2$, 3360 combinations are used to practically identify c) $v_3$, and 420 combinations are used to analyze the identifiability of d) $v_5$.
+Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
 \end{figure}
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/figure2/v1_k1cat_exp_info}}
-	\caption{Contribution of different experiment types towards identification of a) $k_1^{cat}$ and b) $K_1^{ac}$ in $v_1$ based on using data from two different experiments.
-Data from MWC model.}%\label{fig:figure2}
-\end{figure}
-
-\begin{figure}[!tbhp]
 	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v1_v1max_k1cat_parameter_value}}

### 2018-06-11 ident/docs/sifigures.tex
add new figures and remove old ones

@@
 \begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.8\textheight, keepaspectratio]{figures/parameter_dist/v3_mwc_ck_parameter_dist}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$ when there is no noise in the input experimental data.
+Here, the data is generated using the MWC model for $v_3$ and the identification is carried out for the Convenience Kinetic model of $v_3$.
+The top panel shows a violin plot of the distribution of parameter values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.
+The mismatch between the data and the model used for identification results in large standard deviations for the estimated parameters}\label{fig:v3_mwc_ck_values}
+\end{figure}
+
+\begin{figure}[!tbhp]
 	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/ident_exp_info/v1_vmax_exp_info}}
@@
 The choice of perturbation chosen for the first experiment (blue) determines the choice of the second (green) and the third perturbation (yellow) required for identifying the parameters.}\label{fig:v3_1_ck_ident}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=0.6\textheight, keepaspectratio]{figures/ident_exp_info/v3_mwc_ck_exp_info}}
+	\caption{a) The number of data combination from 21 different in silico experiments that can practically identify each parameter in flux $v_3$ when there is no noise in the input experimental data.
+Here the data generated using the MWC model for $v_3$ is used to identify parameters in the Convenience Kinetic model of $v_3$.
+The percentage of total combinations of experimental data used for analysis (3360) that can identify each parameter is also specified.
+$v_3$ required data from three experiments for analysis.
+The contribution of different experiment type towards identifying each parameter is shown in the spider plots for b) $V_3^{max}$, c) $K_3^{fdp}$ and d) $K_3^{pep}$.
+The choice of perturbation chosen for the first experiment (blue) determines the choice of the second (green) and the third perturbation (yellow) required for identifying the parameters.}\label{fig:v3_mwc_ck_ident}
 \end{figure}
@@
 Section \ref{sec:experiments} provides more details on how the combinations of experimental data are generated.}\label{fig:data_utility}
-\end{figure}
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=0.3\textheight, keepaspectratio]{figures/figure6/v3_1_2_ck_1_2_parameter_values}}
-	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_3$ when the a) and b) Monod-Wyman-Changeaux model for allosteric regulation for $v_3$ is used to generate data, c) and d) experimental data is generated using the Convenience Kinetic rate law model with activation for $v_3$.
-The left column corresponds to the first root for all three parameters: $V_3^{max}(1)$, $K_3^{fdp}(1)$ and $K_3^{pep}(1)$, and the second column corresponds to the second root of all three parameters in $v_3$: $V_3^{max}(2)$, $K_3^{fdp}(2)$ and $K_3^{pep}(2)$.}\label{fig:parameter_value_v3}
 \end{figure}

### 2018-06-12 ident/docs/draft_4.tex
significant changes to methods and results and change in order of sections and figures in SI

@@
 The computer-algebra system based method to assess identifiability that we have developed is described in Section \ref{sec:ident} that follows.
-We also provide a description of a numerical equivalent for the computer algebra-based method to estimate parameters numerically in Section \ref{sec:numerical_method}.
-In Section \ref{sec:degree_of_identifiability}, we define a quantitative metric to describe the identifiability of parameters, and in Section \ref{sec:small-model} a complete description of the small metabolic network used to demonstrate the methodology that we have developed is provided.
-Finally, a description of how the method can be used for experimental design for parameter estimation is given in Section \ref{sec:experimental_design}.
+In Section \ref{sec:degree_of_identifiability}, we define a quantitative metric to describe the identifiability of parameters, followed by a description of how the method can be used for experimental design for parameter estimation is given in Section \ref{sec:experimental_design}.
+We provide a description of a numerical equivalent for the computer algebra-based method to estimate parameters numerically in Section \ref{sec:numerical_method} (\textcolor{red}{should this be moved to SI?}).
+Finally, in Section \ref{sec:small-model} a complete description of the small metabolic network used to demonstrate the methodology that we have developed is provided.
 
@@
 	Here the output $y$ in Equation (\ref{eq:output}c) represents concentrations and fluxes that can actually be measured and used for parameter estimation.
-Moreover, $\theta$ is also augmented with additional parameters that connect the measurable quantities with the unmeasurable quantities.
+Moreover, $\theta$ is also augmented with additional parameters that relate the measurable quantities to the unmeasurable quantities.
 In these instances, where not all concentrations and fluxes can be measured, the structural identifiability of parameters $\theta$ can also elucidate the presence or absence of correlations between the unmeasurable (unobservable states) and measured concentrations/fluxes (observable outputs).
-However, the definition of structural identifiability pertaining to the relationship between the observable outputs and unobservable states is not relevant to the materials presented in this paper.
+However, this definition of structural identifiability pertaining to the relationship between the observable outputs and unobservable states is not relevant to the materials presented in this paper.
 
 	Experimental data (concentration or fluxes) from many physical systems is usually noisy, and sometimes may not contain all requisite information pertaining to the system.
-When parameters are estimated on the basis of this data, the ability to estimate unique parameter values to satisfy Equation (\ref{eq:stident}) is referred to as practical identifiability.
-If unique parameter values satisfying Equation (\ref{eq:stident}) can be estimated, then $\theta$ is said to be globally practically identifiable.
+If unique parameter values satisfying Equation (\ref{eq:stident}) can be estimated on the basis of the noisy data, then $\theta$ is said to be globally practically identifiable.
 Whereas, if parameter estimates with quantifiable uncertainties that satisfy Equation (\ref{eq:stident}) can be found, then $\theta$ is said to be locally identifiable.
@@
 
-	In summary, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter, and the practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter as opposed to the structure and parameterization of the model.
+	In summary, the effect of model structure and parameterization on the ability to infer true parameter values from experimental data is determined by the structural identifiability of the parameter, and the practical identifiability of a parameter is contingent upon the nature, quality and quantity of data available to estimate the parameter, as opposed to the structure and parameterization of the model.
 	Therefore, establishing the structural identifiability of parameters enables one to propose models that are not only appropriate representations of physical processes, but are also parameterized in such a way that the value of these parameters can be estimated from measurable data.
@@
 	\subsection{A method to determine structural and practical identifiability of kinetic models of metabolism}\label{sec:ident}
-	In this section, we provide the mathematical framework for determining the identifiability of parameters in kinetic models of metabolism.
-The methodology in shown as a flow diagram in Figure \ref{fig:ident-flowchart}.
+	We provide the mathematical framework for determining the identifiability of parameters in kinetic models of metabolism as a flow diagram in Figure \ref{fig:ident-flowchart}.
 
@@
 
-	As indicated in Figure \ref{fig:ident-flowchart}a, the first step involves the construction of the kinetic model (Equation \ref{eq:kinstoich}) of the metabolic network with $n_r$ reaction fluxes.
-For each flux $v_i$, $i={1, 2, ..., n_r}$, in the kinetic model, let $\mathbf{\theta} \in \mathbb{R}^{n_p}$ in Equation (\ref{eq:kinstoich}b).
+	The first step (Figure \ref{fig:ident-flowchart}a) involves the construction of the kinetic model (Equation \ref{eq:kinstoich}) of the metabolic network.
+For each flux $v_i$, $i={1, 2, ..., n_r}$, in the kinetic model, let $\mathbf{\theta} \in \mathbb{R}^{p}$ in Equation (\ref{eq:kinstoich}b).
 If data from $n_E$ experiments is available for the chosen metabolic network, as stated earlier, for each experiment $j = {1, 2, ..., n_E}$, we assume that all metabolite concentrations ($x\in\mathbb{R}^{n_x}$) and reaction fluxes ($v\in\mathbb{R}^{n_r}$) are measurable.
@@
 
-	In order to establish the identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method.
+	In order to establish the structural and practical identifiability of kinetic parameters for each flux $v_i$, $i={1, 2, ..., n_r}$, we describe a computer algebra-based method.
 The primary use of the computer algebra system is to obtain closed-form expressions for each parameter in $\mathbf{\theta} \in \mathbb{R}^p$ for each flux $v_i$ (Figure \ref{fig:ident-flowchart}b).
-This is done by first selecting a combination of $n_p\le n_E$ experimental data.
-The fluxes and concentrations from $n_p$ different experiments are then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^{n_p}$ for each flux $v_i$, shown in Equation (\ref{eq:nonlineq}).
+This is done by first selecting a combination of $p\le n_E$ experiments.
+The fluxes and concentrations from these $p$ different experiments are then used to formulate a system of nonlinear algebraic equations in $\mathbb{R}^{p}$ for each flux $v_i$, as shown in Equation (\ref{eq:nonlineq}).
 	\begin{align}\label{eq:nonlineq}
@@
 	\end{align}
-	Here, $v_{i,j}$ refers to the flux $v_i$ obtained from experiment $j$.
-$\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^{n_p}$, whose elements are denoted by $\theta_k$.
+	Here, $v_{i,j}$ refers to the measured value of flux $v_i$ from experiment $j$.
+$\mathbf{x}_j$ and $\mathbf{u}_j$ are the vector of metabolite and other input concentrations from each experiment $j$, and $\mathbf{\theta}$ is a vector in $\mathbb{R}^{p}$, whose elements are denoted by $\theta_k$.
 
 	Each equation in (\ref{eq:nonlineq}), indicated by the index $j$, corresponds to the kinetic rate law expression $f(x, \theta, u)$ for each $v_i$, $i={1, 2, ..., n_r}$, described in Equation (\ref{eq:kinstoich}b), written for concentrations ($\mathbf{x}_j$, $\mathbf{u}_j$) and fluxes ($v_{i,j}$) obtained from experiment $j$.
-Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^{n_p}$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^{n_p}$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
-Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from $p$ experiments denoted by the index $j = {1, 2, ..., p}$.
+Solving the system in Equation (\ref{eq:nonlineq}) results in $\mathbb{R}^{p}$ nonlinear expressions for each parameter $\theta_k$ in $\theta \in \mathbb{R}^{p}$ (Equation \ref{eq:theta-eq}), where $N(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the numerator of $g$, and $D(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ is the denominator of $g$ (Figure \ref{fig:ident-flowchart}b).
+Note that $\mathbf{v}_i$, $\mathbf{x}$ and $\mathbf{u}$ are used to denote vector of vectors of fluxes for reaction $i$ ($\mathbf{v}_i$), metabolite ($\mathbf{x}$) and input ($\mathbf{u}$) concentrations, respectively, obtained from $p$ experiments, each denoted by the index $j = {1, 2, ..., p}$.
 	\begin{align}\label{eq:theta-eq}
@@
 
-	The identifiability of parameter $\theta_k$, $k = {1, 2, ..., p}$, for flux $v_i$ can be established by determining the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b): any parameter $\theta_k$ is said to practically identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, and practically non-identifiable if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$.
-Furthermore, the physical properties of the kinetic parameters can be used to distinguish between identifiable and non-identifiable parameter values by designating only parameters with a positive value of $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as identifiable (Figure \ref{fig:ident-flowchart}b).
-The solution $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for an identifiable $\theta_k$, and an infinite number of solutions are possible for a non-identifiable $\theta_k$.
-However, if there are multiple but finite solutions $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, then the corresponding parameter $\theta_k$ is locally identifiable.
+	Consistent with the definitions presented in Section \ref{sec:ident_def} above, we can define structural identifiability for a parameter $\theta_k$ as follows: if a unique solution (Equation \ref{eq:theta-eq}) exists for $\theta_k$, then $\theta_k$ is said to be structurally identifiable.
+However, if multiple finite solutions exists, then $\theta_k$ is only locally structurally identifiable.
+In the absence of any solution, $\theta_k$ is designated as a structurally non-identifiable parameter.
+
+	Similarly, any parameter $\theta_k$ is said to practically identifiable if the solution $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}) is unique for any given experimental data set.
+An infinite number of solutions are possible for a practically non-identifiable $\theta_k$ for a given experimental data set.
+However, if there are multiple but finite number of solutions $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ for the same set of experimental data, then the corresponding parameter $\theta_k$ is locally practically identifiable.
+
+	The number of solutions, and consequently, the practically identifiability of a parameter from a given data set can also be constrained on the basis of the physiological relevance of the parameter estimate.
+We designate only parameters with unique physiologically relevant solutions $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ as practically identifiable (Figure \ref{fig:ident-flowchart}b).
+Parameters with multiple physiologically relevant solutions, or no physiologically relevant solution are designated as locally practically identifiable and practically non-identifiable parameters, respectively.
+For instance, since enzyme kinetic parameters represent metabolite-enzyme affinities to either the catalytic or the regulatory sites of an enzyme, we consider only non-zero positive parameter estimates to be physiologically relevant.
+Thus, parameters with multiple solutions ($g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$) can be still be designated as practically identifiable if there is only one unique positive (physiologically relevant) solution among the infinitely many solutions.
+
+	Consequently, the enzyme kinetic model for each flux $v_i$ can be designated as structurally (or practically) identifiable if every parameter $\theta_k$, $\theta \in \mathbb{R}^{p}$, modeling flux $v_i$ is also structurally (or practically) identifiable.
+However, if at least one parameter $\theta_k$ for flux $v_i$ is either locally structurally (or practically) identifiable or non-identifiable, then the model for flux $v_i$ is also said to be locally structurally (or practically) identifiable or non-identifiable, respectively.
+
+	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
+	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
+We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
+
+	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
+On the other hand, if only 10\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.1 or 10\%.
+Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
+In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 10\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
+
+	Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
+Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
+
+	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}		
+	Not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment.
+This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}.
+In such scenarios, the need to design experiments to effect a change in, and discriminate between changes in different concentrations/fluxes becomes necessary.
+
+	Following the methodology described in Section \ref{sec:ident}, and demonstrated in Section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations of data sets obtained from experiments described in Section \ref{sec:experiments} of the Supplementary Information can be tested for their ability to practically identify any of the fluxes in the small metabolic network.
+This step would determine the degree of identifiability (defined in Section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model.
+
+	As mentioned in Section \ref{sec:ident}, the determination of practical identifiability of a parameter for a given set of experimental data requires determining the number and nature of possible solutions for each parameter $\theta_k$ in Equation (\ref{eq:theta-eq}).
+	The ability of a given set of experiments to satisfy conditions for practical identifiability can also be tested a priori without determining the number and nature of solutions $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ in Equation (\ref{eq:theta-eq}).
+This can be done by checking for the existence of solution $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$, i.e., the value of $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ (Figure \ref{fig:ident-flowchart}b).
+Combinations of experiments for which the corresponding $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u}) = 0$ can be designated as non-informative for practically identifying $\theta_k$.
+On the other hand, if $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$, then the relevant experiments are informative, and can be used for determining practical identifiability.
+In summary, for practically identifiable parameters, the corresponding combinations of experiments should satisfy $D_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})\neq0$ and $g_k(\mathbf{v}_i, \mathbf{x}, \mathbf{u})$ should be unique and physiologically relevant.
+
+	We formally explain how the aforementioned criteria can be used to obtain a minimal and informative collection of experiments from which data can be used to identify and estimate as many model parameters as possible (Figure \ref{fig:ident-design}).
+The identifiability of each parameter based on each experiment with index $j = {1, 2, ..., n_E}$ is established based on the methodology summarized in Figure \ref{fig:ident-flowchart}b, and demonstrated in Section \ref{sec:example}.
+Subsequently, for any flux $v_i$, and for any combination of $p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., p}$) do not satisfy the conditions for identifiability for any parameter $\theta_k$, $\theta\in\mathbb{R}^{p}$ (Figure \ref{fig:ident-flowchart}b), then at least one of the $p$ experiments needs to be replaced to make parameter $\theta_k$ identifiable.
+Consequently, the corresponding experiment cannot be used for estimating parameter $\theta_k$, and can be discarded from the set of all necessary experiments.
+Furthermore, another experiment from $j = {1, ..., n_E}$ needs to be selected to replace the discarded experiment such that parameter $\theta_k$ is identifiable.
+This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
+In doing so, we can arrive at a set of experiments that will always result in practically identifiable parameters for flux $v_i$.
+Note that if none of the $n_E$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n_E+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
+This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.
 
@@
 These bounds can be specified to be arbitrarily large to include all possible parameter values.
-
-	\subsection{Degree of identifiability: A quantitative measure of practical identifiability}\label{sec:degree_of_identifiability}
-	We express the practical identifiability of kinetic parameters using a simple quantitative term called the degree of identifiability.
-We describe the degree of identifiability of any single parameter as the percentage of all data combinations (used to test for practical identifiability) that can identify that parameter.
-
-	As an example, if 90\% of all the experimental data combinations used for testing can identify a parameter $\theta_i$, then the degree of identifiability of $\theta_i$ is said to be 0.9 or 90\%.
-On the other hand, if only 10\% of the combinations can identify another parameter $\theta_j$, then $\theta_j$ has a degree of identifiability of 0.1 or 10\%.
-Furthermore, we can create a hierarchy of practically identifiable parameters using their degrees of identifiability.
-In the above instance of the two parameters $\theta_i$ and $\theta_j$ that have degrees of identifiability of 90\% and 10\% respectively, $\theta_i$ is classified to be more identifiable than $\theta_j$ due to its relatively higher degree of identifiability.
-
-	Determining this hierarchy of identifiable parameters can help in distinguishing parameters that can be identified by any type and any combination of experiments from parameters that can be identified by only a select type and combination of experiments.
-Such a classification can subsequently be used to design minimal sets of experiments that can practically identify all kinetic parameters used to model a metabolic network, going from the least identifiable parameter to the most identifiable parameter.
 
@@
 	The model for flux $v_1$ of the small network (Figure \ref{fig:network}), uses the concentration of the enzyme E as a variable (Equation \ref{eq:flux1}).
-Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzymes (again the details on relaxing this assumption are discussed later), the expression in Equation (\ref{eq:flux1}) for $v_1$ cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
+Since we assume that steady state experimental information is only available for metabolite concentrations and fluxes, and not for enzyme concentration (again, the details on relaxing this assumption are discussed later), the expression in Equation (\ref{eq:flux1}) for $v_1$ cannot be used for identifying parameters $k_1^{cat}$ and $K_1^{ac}$.
 So, we modify the Michaelis-Menten kinetic rate law expression to eliminate the enzyme concentration E as a variable in Equation (\ref{eq:flux1a}).
@@
 	The practically identifiability of parameters of a given flux are determined by solving a system of nonlinear algebraic equations using a computer algebra system (Section \ref{sec:ident}).
-We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a computer algebra system (Mathematica or SymPy in Python).
-In order to overcome this computational obstacle, we model the reaction rate for $v_3$ using the convenience kinetic rate law formulation \parencite{Liebermeister2006}.
-The corresponding expression obtained for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
+We find that the nonlinearity of the MWC kinetic rate law used to model the allosteric regulation of $v_3$ makes it computationally intractable for determining the closed form expressions of the three parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ using a general purpose computer algebra system (Mathematica or SymPy in Python).
+We sought to overcome this computational obstacle by modeling flux $v_3$ using the convenience kinetic rate law \parencite{Liebermeister2006}.
+The corresponding expression for $v_3$ is given below (Equation \ref{eq:flux3_convkin}).
 	\begin{align}\label{eq:flux3_convkin}
@@
 
-	\subsection{Experimental design through practical parameter identification}\label{sec:experimental_design}		
-	Not all metabolite concentrations and fluxes in the model (Equation \ref{eq:kinstoich}) change for any random experiment.
-This makes unambiguous estimation of parameters impossible, either due to the inherent correlation between changes in different concentrations or fluxes, or due to the homeostasis of the concentrations and fluxes under the chosen experimental conditions \parencite{Heijnen2013}.
-In such scenarios, the need to design experiments to effect a change in, and discriminate between changes in different concentrations/fluxes becomes necessary.
-
-	Following the methodology described in Section \ref{sec:ident}, and demonstrated in Section \ref{sec:example} for a single flux using data from a combination of two different experiments, all distinct combinations of data sets obtained from experiments described in Section \ref{sec:experiments} of the Supplementary Information can be tested for their ability to practically identify any of the fluxes in the small metabolic network.
-This step would determine the degree of identifiability (defined in Section \ref{sec:degree_of_identifiability}) of each parameter in each flux in the model, and help distinguish experiment combinations that contribute to identifiability from combinations that do not practically identify any parameter in the model (Figure \ref{fig:ident-flowchart}b).
-In doing so, it is possible to obtain a minimal and informative collection of experiments that can be performed to identify as many model parameters as possible (Figure \ref{fig:ident-design}).
-Consequently, the set of experiments can be used to estimate all the identifiable parameters in the model.
-This is formally explained below.
-
-	The identifiability of each parameter based on each experiment with index $j = {1, 2, ..., n_E}$ is established based on the methodology described in Section \ref{sec:ident} (Figure \ref{fig:ident-flowchart}b), and demonstrated in Section \ref{sec:example}.
-Subsequently, for any flux $v_i$, and for any combination of $n_p$ experimental data sets, if the experimental concentrations and fluxes ($\mathbf{x}_j$ and $\mathbf{v}_j$, respectively, where $j = {1, 2,..., n_p}$) do not satisfy the condition for identifiability for any parameter $\theta_k$ in $\theta\in\mathbb{R}^{n_p}$ (Figure \ref{fig:ident-flowchart}b), then at least one of the $n_p$ experiments needs to be changed to make parameter $\theta_k$ identifiable.
-Consequently, the corresponding experiment cannot be used for estimating parameter $\theta_k$, and needs to be discarded from the set of all necessary experiments.
-Furthermore, another experiment from $j = {1, ..., n_E}$ needs to be selected to replace the discarded experiment such that parameter $\theta_k$ is identifiable.
-This process has to be repeated until all parameters in $\theta\in\mathbb{R}^p$ are identifiable for flux $v_i$.
-In doing so, we can arrive at a set of $n_p$ experiments that will always result in practically identifiable parameters for flux $v_i$.
-Note that if none of the $n_E$ pre-selected experiments satisfy the identifiability condition, then we can design an $(n_E+1)^{th}$ experiment that can replace one of the experiments that causes practical non-identifiability.
-This analysis can be performed for each flux in a metabolic network independent of all the other fluxes, making it theoretically scalable even to genome-scale models of metabolism.
-
 	\section{Results}
@@
 	\subsection{Identifying parameters in kinetic models of metabolism: an example}\label{sec:example}	
-	In this section, we illustrate the proposed methodology step by step to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model}).
+	We illustrate the proposed methodology, step by step, to identify parameters of flux $v_1$ in the small metabolic network (Figure \ref{fig:network} and Section \ref{sec:small-model}).
 We choose the expression for flux $v_1$ given in Equation (\ref{eq:flux1a}) for this demonstration.
 
-	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in Section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
-So, from the $n_E = 21$ different experiments described in Section \ref{sec:experiments} and Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$.
+	Since $\theta = \{V_1^{max}, K_1^{ac} (ne)\} \in \mathbb{R}^2$ for $v_1$, as mentioned in Supplementary Section \ref{sec:experiments}, we need steady state concentration and flux measurements from at least two different experiments.
+So, from the $n_E = 21$ different experiments described in Supplementary Section \ref{sec:experiments} and Supplementary Table \ref{tab:pval}, we can choose multiple combinations of $p = 2$ experiments to satisfy the data requirements for identifying $v_1$ i.e., in Equation (\ref{eq:nonlineq}) $j = \{1, 2\}$.
 We label the available concentrations and fluxes as ${ac}^{(j)}$ and ${v_1}^{(j)}$, respectively.
@@
 
-	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
+	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a general purpose computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b).
 These expressions have the form shown in Equation (\ref{eq:theta-eq}).
@@
 	\end{subequations}
-	To test the practical identifiability of the parameters in Equation \ref{eq:v1_par}, we substitute any suitable in silico experimental data and determine the value of the denominator of the right hand side expression.
-Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be positive (Figure \ref{fig:ident-flowchart}b).
-The parameter values that are obtained for $V_1^{max}$ and $K_1^{ac} (ne)$ by substituting in silico steady state experimental data are shown in Supplementary Figure \ref{fig:v1_v1max_ck_values}.
-Due to the numerous possible parameter values seen in Supplementary Figure \ref{fig:v1_v1max_ck_values}, we can conclude that both $V_1^{max}$ and $K_1^{ac}$ are practically non-identifiable.
+	The presence of a single unique expression for both $V_1^{max}$ and $K_1^{ac} (ne)$ shown above indicates that these parameters are structurally identifiable.
 
-	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is also available, in addition to the measured metabolite concentrations and fluxes.
-In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$.
-Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) is substituted with $V_1^{max} = k_1^{cat}E$ instead.
-The corresponding identifiability expressions for $k_1^{cat}$ and $K_1^{ac}$ are given in Equation (\ref{eq:v1_k1cat}).
+	As described in Section \ref{sec:ident}, we use multiple combinations of in silico experimental data with Equation \ref{eq:v1_par} to assess the practical identifiability of $V_1^{max}$ and $K_1^{ac} (ne)$.
+First we determine the value of the denominator of the right hand side expression in Equation (\ref{eq:v1_par}).
+A non-zero value could potentially indicate an identifiable parameter.
+In addition, since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, the values obtained from Equation (\ref{eq:v1_par}) should also be positive for these parameters to be classified as practically identifiable (Figure \ref{fig:ident-flowchart}b).
+All estimates for $V_1^{max}$ and $K_1^{ac} (ne)$ from all (240) possible combinations of experimental data are shown in Supplementary Figure \ref{fig:v1_v1max_ck_values}.
+Due to the numerous possible estimates for both parameters (Supplementary Figure \ref{fig:v1_v1max_ck_values}), we can conclude that they are practically non-identifiable.
+
+	So far, we have assumed that only information on metabolite concentrations and metabolic fluxes is available in the form of omics data sets.
+However, protein concentration measurements can also be obtained along with measured metabolite concentrations and fluxes.
+To ascertain the impact of such a scenario on parameter identifiability, we demonstrate the proposed methodology for identifying flux $v_1$ when it is modeled by Equation (\ref{eq:flux1}), and the practical identifiability is assessed in the presence of concentration measurements for the enzyme \textit{E}.
+Note that the parameter $V_1^{max}$ in Equation (\ref{eq:flux1a}) is substituted with $V_1^{max} = k_1^{cat}E$ in Equation (\ref{eq:flux1}), and the dimension of the parameter space for $v_1$ is still $\mathbb{R}^2$.
+The corresponding identifiability expressions for $k_1^{cat}$ and $K_1^{ac}$ are given in Equation (\ref{eq:v1_k1cat}) below.
 	\begin{subequations}\label{eq:v1_k1cat}
@@
 	\end{subequations}		
-	We show the parameter value for $k_1^{cat}$ and $K_1^{ac}$ that are obtained through the practical identifiability analysis in Figure \ref{fig:ident_values}a when in silico experimental data is substituted in Equation (\ref{eq:v1_k1cat}).
-Through Equation (\ref{eq:v1_k1cat}) and Figure \ref{fig:ident_values}a we are able to show that the uncertainty in the parameter estimates (Supplementary Figure \ref{fig:v1_v1max_ck_values}) can be resolved through the incorporation of the available enzyme concentrations.
-Thus, having more experimental information can help resolve practical identifiability.
+	This establishes the structural identifiability of the parameters.
+
+	Following the procedure described in Section \ref{sec:ident} and demonstrated above for $V_1^{max}$, we are able to show that both $k_1^{cat}$ and $K_1^{ac}$ have unique estimates (Figure \ref{fig:ident_values}a), and are hence practically identifiable.
+Accordingly, we can summarize that the practical non-identifiability of $v_1$ (due to the uncertainty in the estimates of $V_1^{max}$ and $K_1^{ac}$ seen in Supplementary Figure \ref{fig:v1_v1max_ck_values}) can be mitigated through the incorporation of additional experimental data combined with a finer resolution of the model parameters.
+In this instance, the additional data takes the form of enzyme concentrations.
 
@@
 
-	In the following section we present results from the identifiability analysis of fluxes $v_2$, $v_3$ and $v_5$ in the small metabolic network (Figure \ref{fig:network}), using the methodology (Figure \ref{fig:ident-flowchart}) that we have demonstrated above for $v_1$.
+	\subsection{Establishing structural identifiability of parameters based on closed-form solutions}\label{sec:proof}
+	Here, we discuss results from the identifiability analysis of fluxes $v_2$, $v_3$ and $v_5$ in the small metabolic network (Figure \ref{fig:network}), using the methodology (Figure \ref{fig:ident-flowchart}) that we have demonstrated above for flux $v_1$
 
-	\subsection{Establishing Structural identifiability of parameters based on closed-form solutions}\label{sec:proof}
-	For the proposed methodology (Figure \ref{fig:ident-flowchart}) to work, it should be possible to obtain closed form solutions for each parameter in the enzyme kinetic model for each flux as shown in Equation (\ref{eq:theta-eq}).
-Since the ability to obtain closed-form solutions for each parameter is dependent on the model structure, any parameter that has non-unique closed-form solutions can be called a structurally non-identifiable parameter.
-However, if the number of solutions that the parameter has are finite, then the parameter is only locally structurally identifiable.
+	In order to assess the identifiability of each flux in a metabolic network, we demonstrated in Section \ref{sec:example} that it should be possible to obtain closed form solutions (Equation \ref{eq:theta-eq}) for each parameter in the enzyme kinetic model for each flux (Figure \ref{fig:ident-flowchart}).
+Subsequently, we also illustrated the ability to establish structural identifiability based on the number of closed-form solutions obtained for each parameter by using flux $v_1$ as an example (Section \ref{sec:example}).
+Recall that flux $v_1$, in both forms (Equations \ref{eq:flux1} and \ref{eq:flux1a}), is expressed as a Michaelis-Menten model.
+Hence, by extension, we can conclude that $v_2$, which is also modeled as Michaelis-Menten flux, is also structurally identifiable with unique expressions for $V_2^{max}$ and $K_2^{pep}$.
+The closed-form expressions for these parameters are similar to the ones shown in Equation \ref{eq:v1_par}, with \textit{ac} replaced by \textit{pep}, and $v_1$ replaced by $v_2$.
+	However, we find that Convenience kinetics rate law model for $v_3$ (Equation \ref{eq:flux3_convkin}), and Hill kinetic rate law model for $v_5$ (Equation \ref{eq:flux5}) are not structurally identifiable.
 
-	We demonstrated the structural identifiability of parameters modeling $v_1$ in Section \ref{sec:example}.
-We have shown that the parameters have only one unique closed-form solution, and accordingly are structurally identifiable.
-Since $v_2$ is also expressed using the Michaelis-Menten model, just like $v_1$, we find that the parameters ($V_2^{max}$ and $K_2^{pep}$) are also structurally identifiable.
-The closed-form expressions for these parameters are similar to the ones shown in Equation \ref{eq:v1_par}, with \textit{ac} replaced by \textit{pep}, and $v_1$ replaced by $v_2$.
+	In the case of $v_3$, we find that the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have non-unique but a finite number of solutions, i.e., we get two distinct solutions for each parameter.
+So, $v_3$ is only locally structurally identifiable.
+We tried establishing global structural identifiability by reducing the dimension of the parameter space $\theta$ for $v_3$ from $\mathbb{R}^3$ to $\mathbb{R}^2$.
+We do so by fixing either $K_3^{fdp}$ or $K_3^{pep}$ as a known quantity, and identify the other unfixed parameter along with $V_3^{max}$.
+This results in unique expressions for both $V_3^{max}$ and the other unfixed parameter ($K_3^{pep}$ or $K_3^{fdp}$).
+Thus, by reducing the dimension of $\theta$ to $\mathbb{R}^2$, we were able to obtain a structurally identifiable model for $v_3$.
 
-	However, we find that the parameters used to model $v_3$ using the Convenience kinetics rate law, and $v_5$ using the Hill kinetic rate law are not structurally identifiable as they are given in Section \ref{sec:small-model}.
+	Next, we describe results obtained in determining the identifiability of $v_5$, which describes a transcription/translation reaction using Hill kinetics (Equation \ref{eq:flux5}), using only the available experimental data on the metabolite concentrations and the fluxes within the metabolic network.
+	Recall that we use the Convenience kinetic model as opposed to the MWC model for $v_3$ due to the inability of the computer algebra system to solve for the parameters in the MWC model.
+Similarly, for $v_5$, the computer algebra system could not solve for all three model parameters: $V_e^{max}$, $K_e^{fdp}$ and $n_e$.
+So, instead changing the model (as in $v_3$), we resorted to reducing the dimension of the parameter space by fixing one of the three parameters, the Hill coefficient $n_e$.
+With a fixed and known $n_e$, $K_e^{fdp}$ has two possible solutions, and $V_e^{max}$ has only one unique solution.
+Thus, $v_5$ is locally structurally identifiable.
 
-	First, for $v_3$, we find that the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different closed-form solutions.
-Thus, based on the presence of non-unique but finite number of possible solutions for these parameters we can classify $v_3$ as a locally structurally identifiable flux.
-In order to alleviate local structural identifiability, we reduced the dimension of the parameter space for $v_3$.
-Originally, $\theta \in \mathbb{R}^3$ for $v_3$.
-By reducing the dimension of $\theta$ to $\mathbb{R}^2$, we were able to obtain a structurally identifiable model for $v_3$.
-To reduce the dimension of the parameter space for $v_3$, we fix either $K_3^{fdp}$ or $K_3^{pep}$ as a known quantity, and identify the other unfixed parameter along with $V_3^{max}$.
-This results in unique closed-form expressions for both $V_3^{max}$ and the other unfixed parameter ($K_3^{pep}$ or $K_3^{fdp}$).
-
-	%We would like to point out that we used the Convenience kinetics rate law model for $v_3$ due to the inability of the computer algebra system to generate closed-form expressions for parameters in the Monod-Wyman-Chageaux model (Section \ref{sec:small-model}).
-However, we find that the Convenience kinetics model is not structurally identifiable in its current form.
-
-	While $v_3$ is an allosterically regulated metabolic flux, $v_5$ describes a transcription/translation reaction using Hill kinetics.
-We apply our proposed methodology to identify parameters modeling $v_5$ using only the available experimental data on the metabolite concentrations and the fluxes within the metabolic network.
-We could not obtain closed form solutions for parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$ using the computer algebra system.
-So, instead of changing the model as we did for $v_3$ (see Section \ref{sec:small-model}), we resorted to reducing the dimension of the parameter space by fixing one of the three parameters, the Hill coefficient $n_e$.
-We illustrated the consequence of reducing the dimension of the parameter space of the Convenience kinetic model for $v_3$ earlier.
-With a fixed and known $n_e$, $K_e^{fdp}$ has two possible closed-form solutions, which make it a locally structurally identifiable parameter.
-On the other hand, $V_e^{max}$ has only one unique closed-form solution, and therefore is structurally identifiable.
-
-	We have now established conditions for structural identifiability of all the major fluxes in the small metabolic network (Figure \ref{fig:network}).
-We have shown how our proposed methodology can be used to establish conditions for structural identifiability using steady state information on the model variables.
-We next discuss the practical identifiability of the parameters in $v_2$, $v_3$ and $v_5$ whose parameters are structurally identifiable only under certain conditions.
+	Thus, in the process of establishing the structural identifiability of $v_3$ and $v_5$ we have shown that our methodology is capable of extracting the existence of meaningful relationships between different enzyme kinetic parameters using only steady state concentrations and fluxes.
+We have also shown how the existence of these relationships affects the structural identifiability of the enzyme kinetic models for $v_3$ and $v_5$, and under what conditions these relationships can be removed to make some of the most prevalent enzyme kinetic models structurally identifiable.
 
 	\subsection{Relationship between structural and practical parameter identifiability}\label{sec:initial_analysis}
-	We mention in Section \ref{sec:ident_def} that, by definition, unique parameter values based on the model structure are possible for any structurally identifiable parameter.
-Together with this definition for structural identifiability, we also introduced the concept of practical parameter identifiability.
-To recall, we mentioned that  it should be possible to estimate unique parameter values based on all available experimental data for any practically identifiable parameter.
+	Having established conditions for structural identifiability of fluxes $v_1$, $v_2$, $v_3$ and $v_5$ in the previous sections, we look at the practical identifiability of these flux models, and the impact that it has on the choice of experiments chosen to estimate the parameters in these models.
 
-	As shown in Figure \ref{fig:ident-flowchart} and illustrated for $v_1$ in Section \ref{sec:example}, to determine the practical identifiability of parameters we test for the existence of a non-zero denominator of the closed-form expressions of the parameters.
-We also reduce the possible space within which a parameter could be practically identifiable by checking for the physiological feasibility of the parameter values that are obtained through this analysis (Figure \ref{fig:ident-flowchart}b).
-If the resulting parameter values obtained from various combinations of experimental data for each closed-form expression are unique, then the parameter is practically identifiable.
-However, if a non-unique number of parameter values are possible from multiple combinations of experimental steady state data, then the parameter is said to be practically non-identifiable.
-In conjunction with the conditions for structural identifiability demonstrated earlier in Section \ref{sec:proof}, if the parameter has only one unique closed-form expression, and its value is also unique, then the parameter is both structurally and practically identifiable.
-If either of these conditions are not satisfied, the parameters can be either locally structurally or practically identifiable or non-identifiable.
+	We find that both $v_1$ and $v_2$, which are structurally identifiable (Section \ref{sec:proof}), are also practically identifiable.
+The parameters in the respective models possess unique values based on distinct combinations of experimental data (Figure \ref{fig:ident_values}a and b and Supplementary Figures \ref{fig:v1_kcat_ck_values} and \ref{fig:v2_ck_values}).
 
-	Accordingly, both $v_1$ and $v_2$ are not only structurally identifiable due to the presence of unique closed-form expressions for their parameters, they are also practically identifiable because the parameters in the respective models possess unique values based on distinct combinations of experimental data (Figure \ref{fig:ident_values}a and b).
+	It is well understood in the systems identification community that structural identifiability is a necessary condition for practical identifiability ().
+So, in the case of $v_3$, which is only locally structurally identifiable, we would expect it to at least be locally practically identifiable.
+However, we not only find that the parameter estimates for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ in the Convenience kinetics model for $v_3$ are not unique (Figure \ref{fig:ident_values}c), we also observe large uncertainties in these estimates (Figure \ref{fig:ident_values}c and Supplementary Figure \ref{fig:v3_ck_values}).
+These observations lead us to conclude that when using steady state data to establish practical identifiability, the local structural identifiability of $v_3$ does not guarantee its practical identifiability, local or otherwise.
 
-	Regarding $v_5$, we showed earlier in Section \ref{sec:proof} that the identifiability of $v_5$ can be analyzed only when the Hill coefficient $n_e$ is held constant.
-So, in subsequent discussions, the dimension of the $v_5$ parameter space is kept at $\mathbb{R}^2$ by fixing the value of $n_e$.
-Under these conditions, we find that the structurally identifiable parameter $V_e^{max}$ is also practically identifiable, i.e., it has only one unique value based on all available in silico experimental data (Figure \ref{fig:ident_values}c).
-However, recall that unlike $V_e^{max}$, $K_e^{fdp}$ is only locally structurally identifiable as it has two possible closed-form expressions.
-Nonetheless, despite its local structural identifiability, we find that the $K_e^{fdp}$ is also practically identifiable, like $V_e^{max}$, with only one unique parameter value (Figure \ref{fig:ident_values}c).
-
-	We find that the practical identifiability of $v_5$, despite the local structural identifiability of one of its parameters, is due to the enforcement of the physiological relevance criteria on the parameters i.e., only one of the two closed-form expressions for $K_e^{fdp}$ is physiologically relevant.
-The other solution always acquires a negative value that has no physiological meaning.
-Thus, by reducing the practically identifiable space of parameters, we have shown that our methodology can establish global practical identifiability even when the parameters are only locally structurally identifiable.
-
-	Similar to $K_e^{fdp}$ in $v_5$, we also explained the local structural identifiability of $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ modeling $v_3$ in Section \ref{sec:proof}.
-These parameters have two possible closed-form expressions.
-In Figure \ref{fig:ident_values}d we show the numerical values for one of the two possible closed-form expressions for $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$.
-The numerical values for the second closed-form expressions of the three parameters is presented in Supplementary Figure, and they also have a similar distribution.
-Based on the prior definition for practically identifiable parameters, the numerous possible values that the three parameters can acquire (Figure \ref{fig:ident_values}d) leads us to conclude that the parameters in $v_3$ are practically non-identifiable when they are only structurally locally identifiable.
-Also, unlike $v_5$, which is practically identifiable in the presence of local structural identifiability, parameters for $v_3$ are practically non-identifiable even after the reduction in the practically identifiable parameter space realized using the physiological relevance condition (Figure \ref{fig:ident-flowchart}b).
+	However, we find $v_3$ to be practically identifiable when its parameters are also structurally identifiable (Figure \ref{fig:v3_var_ck_values}).
+Earlier in Section \ref{sec:proof} we had mentioned that $V_3^{max}$ and $K_3^{fdp}$ are structurally identifiable only when the dimension of the parameter space is $\mathbb{R}^2$, and not in $\mathbb{R}^3$, i.e., either $K_3^{pep}$ of $K_3^{fdp}$ is fixed to identify $V_3^{max}$ and the other enzyme affinity constant.
+Under these scenarios we find both parameters in $\mathbb{R}^2$ to be structurally and practically identifiable (Figure \ref{fig:v3_var_ck_values}).
 
@@
 
-	%So, the observation of practical non-identifiability for $v_3$ in the presence of local structural identifiability is in contrast to what we observed for $v_5$ earlier. 
-	However, we find that $v_3$ is practically identifiable when its parameters are also structurally identifiable (Figure \ref{fig:v3_var_ck_values}).
-Earlier in Section \ref{sec:proof} we had mentioned that $V_3^{max}$ and $K_3^{fdp}$ are structurally identifiable only when $K_3^{pep}$ is fixed, and $V_3^{max}$ and $K_3^{pep}$ are structurally identifiable when $K_3^{fdp}$ is fixed.
-Under these scenarios we find the structurally identifiable parameters to also be practically identifiable (Figure \ref{fig:v3_var_ck_values}).
+	Regarding $v_5$, we showed earlier in Section \ref{sec:proof} that $v_5$ is structurally identifiable only when the Hill coefficient $n_e$ is held constant.
+Given the need for structural identifiability to study practically identifiability, in subsequent discussions, the dimension of the $v_5$ parameter space is kept at $\mathbb{R}^2$ by fixing the value of $n_e$.
+Under these conditions, we find $V_e^{max}$ to be both structurally and practically identifiable (Figure \ref{fig:ident_values}d and Supplementary Figure \ref{fig:v5_ck_values}).
+However, recall from Section \ref{sec:proof} that unlike $V_e^{max}$, $K_e^{fdp}$ is only locally structurally identifiable as it has two possible closed-form expressions.
+Nonetheless, despite its local structural identifiability, we find that the $K_e^{fdp}$ is also practically identifiable, like $V_e^{max}$, with only one unique parameter value (Figure \ref{fig:ident_values}d and Supplementary Figure \ref{fig:v5_ck_values}).
+Thus, $v_5$ is practically identifiable despite its local structural identifiability.
+This can primarily be attributed to the enforcement of the physiological relevance criteria on the parameters i.e., only one of the two closed-form expressions for $K_e^{fdp}$ is physiologically relevant for all available experimental data sets.
+	 The other solution always acquires a negative value that has no physiological meaning.
+Thus, by reducing the practically identifiable space of parameters, we have shown that our methodology can establish global practical identifiability even when the parameters are only locally structurally identifiable, again using only steady state experimental data.
 
-	In conjunction with the practical identifiability of $v_5$ established earlier, we see that it is possible to delineate between structural and practical identifiability of parameters in kinetic models of metabolism only under certain conditions, and not in others.
+	 To summarize, $v_3$ in practically non-identifiable when it is only locally structurally identifiable.
+In contrast, $v_5$ is practically identifiable even in the presence of local structural identifiability.
+Hence, using $v_3$ and $v_5$ as examples we have shown that in using steady state data to determine both structural and practical identifiability in kinetic models of metabolism, it is possible to establish a relationship between structural and practical identifiability only under certain conditions, and not in others.
 
 	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
-	The analysis of parameter practical identifiability can be used to gather information on the type of experiments that can provide useful data for parameter estimation.
-For instance, during practical identifiability analysis, if either the denominator of the closed-form expression is zero, or if the parameter values that are obtained are not physiologically feasible (Figure \ref{fig:ident-flowchart}b), then the experimental data set concerned is said to be incapable of practically identifying that said parameter.
-Consequently, the data from that combination of experiments is considered non-informative.
-When this analysis is repeated for multiple combinations of steady state data from the 21 different in silico experiments, we can determine the number of different experimental data combinations that can practically identify each parameter in each flux (Figure \ref{fig:ident}a - c).
+	In Section \ref{sec:experimental_design} and Figure \ref{fig:ident-flowchart}b we describe how the analysis of practical identifiability of a parameter can be used to gather information on the type of experiments that can provide useful data for parameter estimation.
+In Section \ref{sec:degree_of_identifiability}, we also describe how this analysis allows us to arrange parameters hierarchically on the basis of their degree of identifiability.
+In this section we show how these two methods combined can be used to design experiments to provide informative data for parameter estimation in the kinetic model of small metabolic network (Figure \ref{fig:network}).
 
@@
 
-	As described in Section \ref{sec:degree_of_identifiability}, the information on the number of experimental data sets that can practically identify each parameter can be used to determine the degree of identifiability of the corresponding parameters.
-Subsequently, this information can be used to classify parameters based on their ease of identifiability.
+	In Figure \ref{fig:ident}a-c we show the degree of identifiability (percentage of experimental data combinations that are capable of identifying each parameter) of each parameter in each of three fluxes $v_1$, $v_2$ and $v_5$, respectively.
+The degree of identifiability of $v_1$ based on Equation (\ref{eq:flux1a}) is provided in Supplementary Figure \ref{fig:v1_v1max_ck_ident}.
+Note that we perform this analysis only for conditions under which the structural and practical identifiability of these fluxes can be established without doubt.
+So, based on the small uncertainties observed for $V_1^{max}$ and $K_1^{ac} (ne)$ (Supplementary Figure \ref{fig:v1_v1max_ck_values}), we classify the corresponding model of $v_1$ (Equation \ref{eq:flux1a}) as locally practically identifiable.
+Although, this analysis can be performed for fluxes that are not practically identifiable, i.e., $v_3$, we doubt the meaningfulness of such an analysis (Supplementary Figure \ref{fig:v3_1_ck_ident}).
+It is important to note that in scenarios where fluxes are only locally practically identifiable the degree of identifiability (Supplementary Figure \ref{fig:v1_v1max_ck_ident} for $v_1$ and Supplementary Figure \ref{fig:v3_1_ck_ident} for $v_3$) refers to the number of experimental data sets that can determine physiologically relevant values for the corresponding parameters.
 
-	In Figure \ref{fig:ident}a-c we show the number of experimental data combinations that are capable of identifying each parameter, and consequently, the degree of identifiability of each parameter (percentage experimental data combinations that are capable of identifying each parameter) in flux $v_1$ (Figure \ref{fig:ident}a), $v_2$ (Figure \ref{fig:ident}b) and $v_5$ (Figure \ref{fig:ident}c).
-The degree of identifiability for $V_1^{max}$ and $K_1^{ac}(ne)$ in $v_1$ is shown in Supplementary Figures \ref{fig:v1_v1max_ck_ident}, and the degree of identifiability for both closed-form expressions of the three parameters of $v_3$ are shown in Supplementary Figure \ref{fig:v3_1_ck_ident}.
-The degree of identifiability of each parameter is also given in these figures as percentages.
+	First, we see that the maximum reaction rates ($V_i^{max}$) are more or similarly identifiable in comparison to the corresponding enzyme binding ($K_i$) constants or the activation/inhibition constants, in the respective reaction rate law models (Figure \ref{fig:ident}, Supplementary Figures \ref{fig:v1_v1max_ck_ident} and \ref{fig:v3_1_ck_ident}).
+This is true despite the fact that the different fluxes are modeled using different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition.
 
-	It is important to recall that both $v_1$ (Supplementary Figure \ref{fig:v1_v1max_ck_values}) and $v_3$ (Figure \ref{fig:ident_values}d) are not practically identifiable (Section \ref{sec:initial_analysis}).
-While $v_1$ becomes practically identifiable (Figure \ref{fig:ident_values}a) if enzyme concentrations are utilized to alleviate uncertainties in the enzyme turn over rates ($k_1^{cat}$), $v_3$ becomes practically identifiable only when the dimension of the parameter space is reduced (Figure \ref{fig:v3_var_ck_values}).
-In these cases the degree of identifiability (Supplementary Figure \ref{fig:v1_v1max_ck_ident} for $v_1$ and Supplementary Figure \ref{fig:v3_1_ck_ident} for $v_3$) refers to the number of experimental data sets that can determine physiologically relevant values for the corresponding parameters.
+	The degree of identifiability of parameters in $v_3$, when it is structurally and practically identifiable (Sections \ref{sec:proof} and \ref{sec:initial_analysis}) is shown in Supplementary Figure \ref{fig:v3_var_1_2_ck_ident}.
+In conjunction with the degree of identifiability of other parameters (Figures \ref{fig:ident} and Supplementary Figure \ref{fig:v1_v1max_ck_ident}), we find that with the exception of $V_1^{max}$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $k_1^{cat}$ in $v_1$ (Figure \ref{fig:ident}a), all data sets used to test practical identifiability can determine unique values for parameters when the corresponding parameter is structurally identifiable.
 
-	Based on their degrees of identifiability, we see that the maximum reaction rates ($V_i^{max}$) are more or similarly identifiable in comparison to the corresponding enzyme binding ($K_i$) constants or the activation/inhibition constants, in the respective reaction rate law models (Figure \ref{fig:ident}, Supplementary Figures \ref{fig:v1_v1max_ck_ident} and \ref{fig:v3_1_ck_ident}).
-We make this observation despite the fact that the four fluxes are modeled using three different enzyme kinetic rate laws: $v_1$ and $v_2$ are modeled using the Michaelis-Menten rate law, $v_3$ is modeled using the Convenience kinetic rate law and $v_5$ is modeled as a Hill equation with inhibition.
-As mentioned earlier, for $v_1$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $v_3$ (Supplementary Figure \ref{fig:v3_1_ck_ident}) we have shown that a greater number of experimental data sets can predict physiologically relevant or non-zero positive values for $V_1^{max}$ and $V_3^{max}$ than for $K_1^{ac}(ne)$ and $K_3^{fdp}$ or $K_3^{pep}$, respectively.
-
-	The degree of identifiability of $v_3$, when its parameters are structurally and practically identifiable (Sections \ref{sec:proof} and \ref{sec:initial_analysis}) are shown in Supplementary Figure \ref{fig:v3_var_1_2_ck_ident}.
-Accordingly, we find that with the exception of $V_1^{max}$ (Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and $k_1^{cat}$ in $v_1$ (Figure \ref{fig:ident}a), all data sets used to test practical identifiability can determine unique values for parameters when the corresponding parameter is structurally identifiable.
-
-	We can attribute the difference in the degree of identifiability between $v_1$ (Figure \ref{fig:ident}a and Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and the other fluxes ($v_2$, $v_3$ and $v_5$) to the ability of data from different combinations of experiments to satisfy the conditions for practical identifiability of that parameter, that can be determined a priori.
+	We can attribute the difference in the degree of identifiability between $v_1$ (Figure \ref{fig:ident}a and Supplementary Figure \ref{fig:v1_v1max_ck_ident}) and the other fluxes ($v_2$, $v_3$ and $v_5$) to the ability of data from different combinations of experiments to satisfy the conditions for practical identifiability of that parameter.
 In systems identification terminology, data requirements for parameter identification can be tied to selecting experiments that are persistently excitable for the flux being identified.
@@
 
-	%\begin{figure}[!tbhp]
-	%	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure2/v1_k1cat_v2_v5_2_ck_exp_info}}
-	%	\caption{Frequency of each type of perturbation experiment in data sets that can identify parameters in fluxes a) $v_1$, b) $v_2$ and c)$v_5$.
-The frequencies are represented as a percentage of the total number of experiments present within all the data sets that can practically identify the parameters.
-The frequency of experiments that occur as the first experiment in a combination of two experiments are shown as a blue surface, and the frequency of experiments that occur as the second experiment in the combination are shown as a green surface.}\label{fig:expinfo}
-	%\end{figure}
+	From the above example we can summarize that identification of individual fluxes within a metabolic network necessitates a careful consideration of experiments such that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}.
 
-	From the above example we can summarize that identification of individual fluxes within a metabolic network necessitates a careful consideration of experiments such that the data acquired can satisfy conditions for practical identifiability for all parameters modeling a flux, and subsequently, all fluxes within a network \parencite{Heijnen2013}.
 	To facilitate the design of experiments based on their ability to satisfy requirements for practical identifiability of parameters, we determine the occurrence of each type of steady state perturbation experiment within combinations that can practically identify each parameter (Figure \ref{fig:ident}d-f, Supplementary Figures \ref{fig:v1_v1max_ck_ident},  \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}).
 So, with our proposed methodology it is possible to identify the types of perturbation experiments that would be informative for identifying each parameter in each flux with steady state concentration and flux data.
+In these figures the contribution from different experiment types for identifying parameters in $v_1$, $v_2$ and $v_5$ are respectively shown as spider plots.
 
-	As mentioned in Supplementary Section \ref{sec:experiments}, we use experimental data from five different types of experiments to test the practical identifiability of parameters in the model (Supplementary Table \ref{tab:pval}).
-In Figures \ref{fig:ident}d-f, the contribution from different experiment types for identifying parameters in $v_1$, $v_2$ and $v_5$ are respectively shown as spider plots.
-
-	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part (> 50\%) of the identifiable experimental data combinations for $v_1$ in comparison to the other types of experiments (Figure \ref{fig:ident}d and Supplementary Figure \ref{fig:v1_v1max_ck_ident}).
-This is in agreement with the condition for identifiability that we discussed earlier (Equations \ref{eq:v1_k1cat} and \ref{eq:v1_par}).
-Since less than 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we also say that identifiability analysis is crucial to determine the minimum number of experiments, along with the nature of experiments that can help identify parameters for $v_1$.
+	The contribution of experiments that involve changes in the acetate concentrations, which consequently bring about changes in the value of $v_1$, contribute to a significant part ($\approx$80\%) of the identifiable experimental data combinations for $v_1$ in comparison to the other types of experiments (Figure \ref{fig:ident}d and Supplementary Figure \ref{fig:v1_v1max_ck_ident}).
+This is in agreement with the condition for identifiability that we discussed earlier and showed mathematically (Equations \ref{eq:v1_k1cat} and \ref{eq:v1_par}).
+Since only about 50\% of all data combinations can satisfy these requirements, and can consequently identify $v_1$ (Figure \ref{fig:ident}a), we also say that identifiability analysis is crucial to determine the minimum number of experiments, along with the nature of experiments that can help identify parameters for $v_1$.
 
@@
 
-	In all of the above scenarios for $v_1$ (Figure \ref{fig:ident}d), $v_2$ (Figure \ref{fig:ident}e) and $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}), the distribution of experiment types between the two ($v_1$ and $v_2$) or three ($v_3$) required experiments is quite similar.
-Hence, the green/blue/yellow surfaces in the spider plots are superimposed upon each other.
-This is also seen for experiments identifying $V_e^{max}$ in $v_5$ (Figure \ref{fig:ident}f).
-However, this is not the case for $K_e^{fdp}$ in $v_5$ (Figure \ref{fig:ident}f).
-We see that when two experiments are required to identify $K_e^{fdp}$, the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa, so that data with enough information is available for the identification of $K_e^{fdp}$.
-Also, since $V_e^{max}$ is globally identifiable using any experiment type (Figures \ref{fig:ident}c and f), the choice and number of experiments required to identify $v_5$ completely hinges upon the identifiability of $K_e^{fdp}$ from the chosen experiments.
+	In all of the above scenarios for $v_1$ (Figure \ref{fig:ident}d), $v_2$ (Figure \ref{fig:ident}e), $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident} and \ref{fig:v3_var_1_2_ck_ident}) and $v_5$ (Figure \ref{fig:ident}f), the distribution of experiment types between second, and in the case of $v_3$, the third experiments is dependent on the choice of the preceding experiment, i.e., the choice of the first experiment has a bearing on the choice of the second experiment, and vice-versa.
 
-	So, now we have established a hierarchy of parameters based on their identifiability.
-Parameters for $v_2$ and $V_e^{max}$ in $v_5$ that are most identifiable are at the top of the hierarchy, while $v_1$ and $K_e^{fdp}$ in $v_5$ fall at the bottom of the list as the experiments needed for their identification require careful consideration.
-When $v_3$ is structurally identifiable, their parameters also do not require any experimental design considerations.
+	With this information, and in conjunction with the degree of identifiability of each and every parameter in the model, we can theoretically determine the experiments needed to identify all model parameters.
+We design experiments in ascending order of degree of identifiability.
+Thus, we need to choose experiments that can identify $K_e^{fdp}$ (Figure \ref{fig:ident}c) followed by experiments for identification of $k_1^{cat}$ and $K_1^{ac}$.
+Since, between fluxes $v_1$ and $v_5$, $v_1$ has a smaller compliment of useful experiments, we choose an experiment involving change in the acetate concentration as the first experiment.
+Since the spider plots (Figure \ref{fig:ident}a and c) show that a combination of data from the wild type experiment and one of the five acetate concentration perturbation experiments to be sufficient to identify both $v_5$ and $v_1$, we can now identify a total of four parameters for both $v_1$ and $v_5$ from only two experiments.
+Given the ability to identify $v_2$ from any of these experiments (Figure \ref{fig:ident}e), we can theoretically identify six different modeling three different fluxes using just two different experiments.
+Extending this idea to determine physiologically feasible estimates for $v_3$ (Supplementary Figures \ref{fig:v3_1_ck_ident}), we hypothesize that would require one more experiment to identify all four fluxes in the small network.
 

### 2018-06-12 ident/docs/sifigures.tex
significant changes to methods and results and change in order of sections and figures in SI

@@
 
-\externaldocument{draft_3} % reference to existing external document
+\externaldocument{draft_4} % reference to existing external document
 
@@
 	\end{center}
-\section{Supplementary Tables}
+
+\section{Supplementary Methods}
+\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
+Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions.
+One way to alter the physiological conditions is to change the substrate concentration under which the cells grow.
+In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$ (Equation \ref{eq:flux1}).
+Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.
+
+Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network.
+The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
+Accordingly, in addition to measuring network responses to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
+In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
+The 21 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
+The box plots in Figure \ref{fig:experimental_data_dist} show the distribution in the in silico metabolite and enzyme concentration and flux data obtained from these experiments.
+
+As described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, the minimum number of experiments from which steady state data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
+For instance, the flux $v_2$ in the model described in section \ref{sec:small-model} has two parameters: $V_2^{max}$ and $K_2^{pep}$.
+Thus $\theta = \{V_2^{max}, K_2^{pep}\} \in \mathbb{R}^2$.
+So, steady state data from two distinct experiments is required for identifying $v_2$.
+Accordingly, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_2$.
+The total number of such possible combinations is 420 (21 x 20) from the 21 different experiments shown in Table \ref{tab:pval}.
+Assuming we also have data for the enzyme concentration $E$ in the expression for flux $v_1$ (Equation \ref{eq:flux1}), identifying the two parameters $k_1^{cat}$ and $K_1^{ac}$ ($\theta \in \mathbb{R}^2$) also requires steady state experimental data from two different experiments.
+
+Similarly, for identifying $v_3$ that is described by three parameters ($\theta \in \mathbb{R}^3$), we need data from a combination of three different experiments.
+Based on the 21 experiments in Table \ref{tab:pval}, we have 7980 distinct combinations (21 x 20 x 19) of three experiments.
+However, it is important to note that not all of the either 420 combinations (for $v_1$ and $v_2$), or the 7980 combinations (for $v_3$), may be suitable for parameter estimation.
+Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers.
+Moreover, among the chosen smaller number of experimental data combinations, the methodology that we propose enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation.
+
 \begin{table}[!thbp]
@@
 \section{Supplementary Figures}	
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=0.6\textwidth,height=0.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
+	\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
+\end{figure}
+
+\begin{figure}[!tbhp]
+	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/parameter_dist/v1_vmax_ck_parameter_values}}
+	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_1$, which is described using as a Michaelis-Menten model, without using the enzyme concentration $E$ for identification when there is no noise in the input experimental data.
+Here $V_1^{max} = k_1^{cat}E$.
+The top panel shows a violin plot of the distribution of parameter values.
+The means are indicated by a black line and the median values are indicated by a red line.
+The bottom panels show the corresponding distributions as histograms.}\label{fig:v1_v1max_ck_values}
+\end{figure}
+
 \begin{figure}[!tbhp]
@@
 The bottom panels show the corresponding distributions as histograms.}\label{fig:v5_ck_values}
-\end{figure}
-
-\begin{figure}[!tbhp]
-	\centering{\includegraphics[width=1.0\textwidth,height=1.0\textheight, keepaspectratio]{figures/parameter_dist/v1_vmax_ck_parameter_values}}
-	\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux $v_1$, which is described using as a Michaelis-Menten model, without using the enzyme concentration $E$ for identification when there is no noise in the input experimental data.
-Here $V_1^{max} = k_1^{cat}E$.
-The top panel shows a violin plot of the distribution of parameter values.
-The means are indicated by a black line and the median values are indicated by a red line.
-The bottom panels show the corresponding distributions as histograms.}\label{fig:v5_v1max_ck_values}
 \end{figure}
@@
 \begin{figure}[!tbhp]
-	\centering{\includegraphics[width=0.6\textwidth,height=0.6\textheight,keepaspectratio]{figures/figure3/experimental_design}}
-	\caption{Flow diagram showing a method for experimental design that uses our methodology for practical identification of parameters to determine the number and type of experiments required to identify all fluxes within a given metabolic network.}\label{fig:ident-design}
-\end{figure}
-
-\begin{figure}[!tbhp]
 	\centering{\includegraphics[width=1.0\textwidth,height=0.5\textheight,keepaspectratio]{figures/figure4/v1_V1max_v2_v3_1_v5_2_mwc_ck_data_utility}}
@@
 
-\section{Supplementary Methods}
-\subsection{Data for establishing parameter identifiability in kinetic model of gluconeogenesis}\label{sec:experiments}
-Steady state information on the metabolome and the fluxome can be gathered under different physiological conditions.
-One way to alter the physiological conditions is to change the substrate concentration under which the cells grow.
-In the small metabolic network (Figure \ref{fig:network}), the acetate concentration plays the role of a substrate, and  determines the acetate uptake flux $v_1$ (Equation \ref{eq:flux1}).
-Thus, steady state metabolite concentrations and fluxes can be calculated under various acetate concentrations to form different experimental data combinations that measure cellular response to changes in the substrate concentration.
-
-Physiological changes can also be brought about by perturbing the expression levels for different enzymes within a metabolic network.
-The model of gluconeogenesis (Figure \ref{fig:network}) described in section \ref{sec:small-model} has three different fluxes ($v_1$, $v_2$ and $v_3$) whose enzyme expression parameters ($k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) can be perturbed to simulate the repression and over expression of the corresponding enzymes.
-Accordingly, in addition to measuring network responses to substrate perturbations, changes in steady state concentrations and fluxes can also be observed for enzyme expression perturbations.
-In total, based on the above discussion, we can perturb four different model parameters ($acetate$, $k_1^{cat}$, $V_2^{max}$ and $V_3^{max}$) to obtain experimental data in silico.
-The 21 different parameter values used to generate experimental data are given in Table \ref{tab:pval}.
-The box plots in Figure \ref{fig:experimental_data_dist} show the distribution in the in silico metabolite and enzyme concentration and flux data obtained from these experiments.
-
-As described in section \ref{sec:ident} and Figure \ref{fig:ident-flowchart}b, the minimum number of experiments from which steady state data is required for identifying all the parameters of a given flux is determined by dimension $\mathbb{R}^p$ of the parameter space of a chosen flux $v_i$.
-For instance, the flux $v_2$ in the model described in section \ref{sec:small-model} has two parameters: $V_2^{max}$ and $K_2^{pep}$.
-Thus $\theta = \{V_2^{max}, K_2^{pep}\} \in \mathbb{R}^2$.
-So, steady state data from two distinct experiments is required for identifying $v_2$.
-Accordingly, multiple combinations of data generated from any two different experiments are used to test the identifiability of $v_2$.
-The total number of such possible combinations is 420 (21 x 20) from the 21 different experiments shown in Table \ref{tab:pval}.
-Assuming we also have data for the enzyme concentration $E$ in the expression for flux $v_1$ (Equation \ref{eq:flux1}), identifying the two parameters $k_1^{cat}$ and $K_1^{ac}$ ($\theta \in \mathbb{R}^2$) also requires steady state experimental data from two different experiments.
-
-Similarly, for identifying $v_3$ that is described by three parameters ($\theta \in \mathbb{R}^3$), we need data from a combination of three different experiments.
-Based on the 21 experiments in Table \ref{tab:pval}, we have 7980 distinct combinations (21 x 20 x 19) of three experiments.
-However, it is important to note that not all of the either 420 combinations (for $v_1$ and $v_2$), or the 7980 combinations (for $v_3$), may be suitable for parameter estimation.
-Hence, the actual number of data combinations that can be used for practically identifying parameters in these fluxes would be smaller than the aforementioned numbers.
-Moreover, among the chosen smaller number of experimental data combinations, the methodology that we propose enables us to determine the number and nature of experimental data combinations, and consequently, the experiments that might allow for parameter identification, and subsequent parameter estimation.
-
 \printbibliography

### 2018-06-20 ident/docs/draft_4.tex
changes to intro (based on WRC comments) and results

@@
 Consequently, the metabolite concentrations within the metabolic network are assumed to be constant, and changes in metabolite concentrations are not modeled.
-Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various non-catalytic regulatory interactions that are also responsible for metabolic function.
+Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various dynamic interactions that are also responsible for metabolic function.
 These shortcomings prevent CBMs from being used to fully understand the steady state as well as the dynamic characteristics of metabolic networks.
@@
 These models account for changes in metabolite concentrations subject to thermodynamic and regulatory constraints that underly metabolic networks in addition to their stoichiometry \parencite{Link2014}.
-Kinetic models can not only help us better understand lesser known and understood characteristics of metabolism like bistability \parencite{Kotte2014}, and their role in human health, but can also improve predictions about the impact of engineering design perturbations on metabolism, and help propose alternative designs to achieve metabolite production goals \parencite{Khodayari2016}.
+As a result kinetic models can help us better understand lesser known and understood characteristics of metabolism like bistability \parencite{Kotte2014}, and their role in human health.
+They can also improve predictions about the impact of engineering design perturbations on metabolism, and help propose alternative designs to achieve metabolite production goals \parencite{Khodayari2016}.
 
 	Kinetic models differ from CBMs in their use of mechanistic enzyme kinetics to model the fluxes within a metabolic network \parencite{Srinivasan2015,Saa2017}.
-The use of kinetic models requires information on the enzyme kinetic rate laws that will be used to model the fluxes, as well as numerical values for the parameters used in these rate laws.
-Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values play a crucial role \parencite{Srinivasan2017}.
+The use of kinetic models requires information on the enzyme kinetic rate laws that are used to model the fluxes, as well as numerical values for the parameters used in these rate laws.
+In addition to the structure of the network represented by kinetic models, the parameter values play a crucial role in analyzing the ability of a metabolic network to exhibit different dynamic characteristics like multiple steady states and oscillations \parencite{Srinivasan2017}.
 
@@
 Typically, enzyme kinetic rate laws are parameterized based on in vitro observations of enzyme activity, as opposed to observations made under in vivo conditions \parencite{Heijnen2005,Smallbone2007}.
-However, some researchers have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions \parencite{Heijnen2005,Heijnen2013}.
-On the other hand, some reports have shown that despite the large uncertainties associated with parameters estimated based on in vivo experimental data \parencite{Link2014}, in vitro parameter estimates are a reasonable approximation of values that would be applicable under in vivo conditions \parencite{Davidi2016}.
+Although some researchers have questioned their relevance for gleaning information on the dynamics of metabolism under in vivo conditions \parencite{Heijnen2005,Heijnen2013}, some others have shown that in vitro parameter estimates are a reasonable approximation of values that would be applicable under in vivo conditions \parencite{Davidi2016}.
+
+	%On the other hand, some reports have shown that despite the large uncertainties associated with parameters estimated based on in vivo experimental data \parencite{Link2014},
 
@@
 
-	However, the importance of model parameter identifiability, i.e., a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Rodriguez-Fernandez2006,Berthoumieux2013}.
+	The importance of model parameter identifiability, i.e., a necessary condition to estimate unique kinetic parameter values from experimental data, is often overlooked \parencite{Rodriguez-Fernandez2006,Berthoumieux2013}.
 Unique parameter estimates from experimental data can only be obtained for structurally and practically identifiable parameters.
 
-	Briefly, it should be possible to obtain unique estimates for structurally identifiable parameters, just based on the model structure and parameterization, irrespective of the type and nature of experimental data available.
+	To be structurally identifiable, a parameter should be capable of being uniquely estimated, irrespective of the type and nature of available experimental data.
+This aspect of a parameter is related to the structure of the model and its parameterization.
+For instance, a parameter is structurally non-identifiable if it is correlated to another model parameter due to redundant model parameterization, and consequently cannot be estimated uniquely from any available experimental data.
+
+	To be practically identifiable, a parameter should be structurally identifiable, and it should be possible to obtain unique parameter estimates from available experimental data.
+Unique parameters estimates for structurally identifiable parameters cannot be determined from experimental data that contribute to practical non-identifiability of the parameter.
+This is attributed to the inability of the data to capture the requisite information needed for estimating the parameter.
+
+	%Briefly, it should be possible to obtain unique estimates for structurally identifiable parameters, just based on the model structure and parameterization, irrespective of the type and nature of experimental data available.
 If parameters cannot be uniquely estimated due to their nonlinear relationships with other parameters (redundant model parameterization), or due to the lack of a relationship between the measured concentrations/fluxes and the unmeasurable concentrations/fluxes (presence of unobservable system states) modeled by the parameters, then these parameters are said to be structurally non-identifiable.
 
-	If the experimental data used for parameter estimation is informative so as to facilitate obtaining unique parameter estimates, or the quantification of uncertainties in the parameter estimates, then the parameters are said to be practically identifiable.
+	%If the experimental data used for parameter estimation is informative so as to facilitate obtaining unique parameter estimates, or the quantification of uncertainties in the parameter estimates, then the parameters are said to be practically identifiable.
 However, if the ability to estimate unique parameter values is compromised due to the inability of the available data to capture the requisite information needed to estimate the parameters in the modeled system, or the uncertainty in parameter estimates is not quantifiable, the parameters are said to be practically non-identifiable.
 
-	Authors have proposed ways to assess parameter identifiability by proposing approximate kinetic models of metabolism that utilize empirical enzyme kinetic rate laws with parameters that have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}.
+	Authors have developed approximate kinetic models of metabolism to assess parameter identifiability.
+These models utilize empirical enzyme kinetic rate laws with parameters that have physical significance, and are identifiable \parencite{Heijnen2005,Smallbone2007}.
 Significant work has also been done towards the development of methods for structural identification of parameters in kinetic models of metabolism \parencite{Nikerel2009,Berthoumieux2013,Raue2014}.
 \textcite{Chis2011b} provide a comparison of the computational performance and implementation aspects of some of these methods.
+	Methods to improve practical identifiability through a priori experimental design have also been developed, with a focus on kinetic models of metabolism \parencite{Gadkar2005a,Rodriguez-Fernandez2006,Vanlier2014a,Raue2014}.
 
-	Methods to improve practical identifiability through a priori experimental design have also been developed, with a focus on kinetic models of metabolism \parencite{Gadkar2005a,Rodriguez-Fernandez2006,Vanlier2014a,Raue2014}.
-Some of these methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, while some others suffer from computational limitations when applied to kinetic models of large metabolic networks \parencite{Gadkar2005a,Raue2014}.
-	Assessing the identifiability (structural or practical) of parameters using any of the aforementioned or other methods is done under the assumption that only dynamic time course concentration data is available and can be used for parameter estimation.
-The ability to use steady state experimental data to assess parameter identifiability is rarely addressed, and very few methodology have been proposed to fill this gap \textcolor{red}{(paper that proposes both steady state and dynamic data use for a priori experimental design resulting in reduced model sizes)}
+	Despite the availability of these methods, the question that has not been answered is the quality and the quantity of experimental data that is required to uniquely estimate parameters in kinetic models of metabolism.
+In this paper, we seek to answer this question in the context of identifiability.
+We present a methodology to determine the conditions under which we can identify parameters, and estimate the number of experimental data sets required for the complete estimation of all parameters in a kinetic model.
+In doing so, we show that the number of experimental data sets required to identify all parameters is bounded by the maximum number of independent parameters in a rate law, if all concentrations and fluxes can be measured.
 
-	In this paper, we show that the identifiability of most parameters in a metabolic network can be done for each reaction individually using steady state fluxomics, metabolomics and proteomics data, and propose a methodology to do so.
-We also demonstrate how this method can facilitate the design of experiments for generating the necessary steady state data to estimate unique parameter values for all reaction fluxes in a metabolic network.
+	We describe the framework for assessing the identifiability of parameters in kinetic models of metabolism in the Methods section.
+We demonstrate the utility of this methodology by illustrating its application to identify parameters in the kinetic model of a small metabolic network.
+In doing so, we also illustrate how some of the metrics that we have developed for assessing practical identifiability of parameters can be used to determine the number and nature of experiments required to generate informative data sets required for parameter estimation.
+	%Since some of the aforementioned methods \parencite{Gadkar2005a,Raue2014} suffer from computational limitations when applied to kinetic models of large metabolic networks, we also improve computational tractability by enabling structural and practical identification of parameters for each reaction in the network individually using steady state data only.
+
+	%Also, the ability to use steady state experimental data to assess parameter identifiability is rarely addressed \textcolor{red}{(paper that proposes both steady state and dynamic data use for a priori experimental design resulting in reduced model sizes)}, and the assumption that only dynamic time course concentration data is available, and can be used for identifiability analysis and parameter estimation is prevalent is all of the above methods.
+
+	%. Also, some of the aforementioned methods are limited by their applicability to approximate kinetic models only \parencite{Nikerel2009,Berthoumieux2013}, and some others
+
+	%We seek to address these issues by first developing a framework to determine the structural and practical identifiability of parameters in kinetic models of metabolism by using only steady state concentration and flux measurements.
+We describe this framework in the Methods section.
+We also demonstrate how identifiability analysis of kinetic models of metabolism with steady state data can be computationally tractable by demonstrating how the proposed methodology can be applied to identify parameters .
+
+	%The application of the methodology is illustrated using one flux from a small metabolic network as an example in Section \ref{sec:example}. In order to enable us to determine the
+
+	%, our proposed methodology can be applied to a large class of kinetic rate laws that are used in the construction of kinetic models of metabolism. Some of these methods also
+
+	%We propose to avoid these computational limitations associated with scale-up
+
+	%We also demonstrate how this method can facilitate the design of experiments for generating the necessary steady state data to estimate unique parameter values for all reaction fluxes in a metabolic network.
+
+	%In this paper, we show that the identifiability of , and propose a methodology to do so.
 In doing so, we are able to propose the number and types of perturbations that will provide the most useful data for parameter estimation.
@@
 In Section \ref{sec:degree_of_identifiability}, we define a quantitative metric to describe the identifiability of parameters, followed by a description of how the method can be used for experimental design for parameter estimation is given in Section \ref{sec:experimental_design}.
-We provide a description of a numerical equivalent for the computer algebra-based method to estimate parameters numerically in Section \ref{sec:numerical_method} (\textcolor{red}{should this be moved to SI?}).
 Finally, in Section \ref{sec:small-model} a complete description of the small metabolic network used to demonstrate the methodology that we have developed is provided.
+%We provide a description of a numerical equivalent for the computer algebra-based method to estimate parameters numerically in Section \ref{sec:numerical_method} (\textcolor{red}{should this be moved to SI?}).
 
@@
 
-	\subsection{Numerical parameter identification and estimation}\label{sec:numerical_method}
-	The identification methodology described in the previous section can be extended to be used in conjunction with an optimization-based procedure for numerical estimation of parameters from experimental data.
+	%\subsection{Numerical parameter identification and estimation}\label{sec:numerical_method}
+	%The identification methodology described in the previous section can be extended to be used in conjunction with an optimization-based procedure for numerical estimation of parameters from experimental data.
 The $\mathbb{R}^p$ nonlinear algebraic equations formulated for each flux $v_i$ (Equation \ref{eq:nonlineq}) can be written as shown in Equation (\ref{eq:numerical_opt}) below to be solved as an optimization problem.
@@
 
-	\begin{subequations}\label{eq:numerical_opt}
-		\begin{align}
-		\underset{\theta, \varepsilon}{min} & \sum_{j=1}^{n_p} \left|\varepsilon_j\right|\\
-		& v_{i, j} - f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) \le \varepsilon_j && j=\{1, 2, ..., n_p\}\\
-		& \theta_l \le \theta_j \le \theta_u && j=\{1, 2, ..., n_p\}
-		\end{align}
-	\end{subequations}
+	%\begin{subequations}\label{eq:numerical_opt}
+	%	\begin{align}
+	%	\underset{\theta, \varepsilon}{min} & \sum_{j=1}^{n_p} \left|\varepsilon_j\right|\\
+	%	& v_{i, j} - f_j(\mathbf{x}_j,\mathbf{\theta}, \mathbf{u}_j) \le \varepsilon_j && j=\{1, 2, ..., n_p\}\\
+	%	& \theta_l \le \theta_j \le \theta_u && j=\{1, 2, ..., n_p\}
+	%	\end{align}
+	%\end{subequations}
 
-	Notice that the parameter values are bounded above and below by $\theta_u$ and $\theta_l$, respectively.
+	%Notice that the parameter values are bounded above and below by $\theta_u$ and $\theta_l$, respectively.
 These bounds can be specified to be arbitrarily large to include all possible parameter values.
@@
 
-	First, we find that the spread in the estimated Convenience kinetics parameter values, when $v_3$ is only locally structurally identifiable, is much larger than when there is mismatch between the model generating the data and the model that is being identified (Supplementary Figure).
+	First, we find that the spread in the estimated Convenience kinetics parameter values, when $v_3$ is only locally structurally identifiable, is much larger than when there is mismatch between the model generating the data and the model that is being identified (Supplementary Figure \ref{fig:v3_mwc_ck_values}).
 A more important observation is that even when the parameters are structurally identifiable in $v_3$ (achieved by assuming either $K_3^{fdp}$ or $K_3^{pep}$ as a known constant), they can at most only be locally practically identifiable.
@@
 
-	Apart from the mismatch between the in vivo and the in vitro enzyme kinetic rate laws, uncertainty in experimental data also arises due to the presence of noise in the measured experimental data.
+	%Apart from the mismatch between the in vivo and the in vitro enzyme kinetic rate laws, uncertainty in experimental data also arises due to the presence of noise in the measured experimental data.
 This noise could be attributed to the measurement error commonly encountered in process analytics.
@@
 
-	We found that in every case where parameters are structurally identifiable, the noise did not have any effect on the identifiability of the parameters or their estimated values.
+	%We found that in every case where parameters are structurally identifiable, the noise did not have any effect on the identifiability of the parameters or their estimated values.
 We also found that inclusion of all necessary data (e.g., the presence or lack thereof of enzyme concentration \textit{E} for $v_1$) can alleviate issues related to using experimental data with errors for identification and estimation: the degrees of identifiability of $V_1^{max}$ and $K_1^{ac}(ne)$ had non-zero standard deviations associated with them, but the degrees of identifiability of $k_1^{cat}$ and $K_1^{ac}$ did not.
@@
 
-	However, for $v_3$, whose parameters are only locally structurally identifiable, we found small non-zero standard deviations in the degrees of identifiabilities (Supplementary Figure) when noisy data is used.
+	%However, for $v_3$, whose parameters are only locally structurally identifiable, we found small non-zero standard deviations in the degrees of identifiabilities (Supplementary Figure) when noisy data is used.
 Although this was seen due to the differences in the number of data combinations that can estimate positive values for each of the three parameters between different noisy experimental data sets, we observe that the standard deviation in the estimated parameter values for each data, between different samples of noisy experimental data, is small (Supplementary Figure).

### 2018-06-20 ident/python2/ss-ident/plot_ident_results.py
add additional argument to determine number of subplots

@@ -583,14 +583,14 @@ def plot_dynamic_sims(dynamic_data, multiple=0, concentrations=1, fluxes=0):
     return None
 
 
-def plot_exp_details(df, color_bar=False, set_palette=True):
+def plot_exp_details(df, max_number_experiments=3, color_bar=False, set_palette=True):
     """plot occurrence of each experiment as a heat map"""
     # set color palette for seaborn heat map
     if set_palette:
         c_pal = sns.color_palette('Blues', 2)
     else:
         c_pal = sns.color_palette('Blues', 16)
-    f, ax = plt.subplots(3, 1, squeeze=True)
+    f, ax = plt.subplots(max_number_experiments, 1, squeeze=True)
 
     idx = pd.IndexSlice
     for i_exp, i_exp_pos in enumerate(df.columns.levels[0]):

### 2018-06-20 ident/python2/ss-ident/all_flux_ident_script.py
split ident results for fluxes with 2 and 3 experiments into separate plots and fun calls

@@ -13,20 +13,31 @@ v1_file_name = os.path.join(os.getcwd(), 'ident/ident_v1_kcat')
 v2_file_name = os.path.join(os.getcwd(), 'ident/ident_v2')
 v3_file_name = os.path.join(os.getcwd(), 'ident/ident_v3_root_1')
 v5_file_name = os.path.join(os.getcwd(), 'ident/ident_v5_root_2')
-file_name_list = [v1_file_name, v2_file_name, v3_file_name, v5_file_name]
+exp_2_file_name_list = [v1_file_name, v2_file_name, v5_file_name]
 
-write_to_file_name = os.path.join(os.getcwd(), 'ident/ident_experiments')
+write_to_file_name = os.path.join(os.getcwd(), 'ident/ident_2_experiments')
 
-df = exp_design_info(list_of_files=file_name_list, original_experiment_file=original_experiment_file,
-                     write_to_file_name=write_to_file_name, max_number_experiments=3)
+df_2 = exp_design_info(list_of_files=exp_2_file_name_list, original_experiment_file=original_experiment_file,
+                       write_to_file_name=write_to_file_name, max_number_experiments=2)
+plot_exp_details(df_2, max_number_experiments=2, color_bar=True, set_palette=False)
+
+# plot only logical only (experiment is involved/not involved)
+logical_df = df_2.applymap(logical_values)
+plot_exp_details(logical_df, max_number_experiments=2)
+
 
 # idx = pd.IndexSlice
 # new_df = df.loc[idx[:, :], idx['experiment_0_id', :]]
 # iris = sns.load_dataset('iris')
 
+exp_3_file_name_list = [v3_file_name]
+write_to_file_name = os.path.join(os.getcwd(), 'ident/ident_3_experiments')
+df_3 = exp_design_info(list_of_files=exp_3_file_name_list, original_experiment_file=original_experiment_file,
+                       write_to_file_name=write_to_file_name, max_number_experiments=3)
+plot_exp_details(df_3, max_number_experiments=3, color_bar=True, set_palette=False)
 
-logical_df = df.applymap(logical_values)
-plot_exp_details(logical_df)
-plot_exp_details(df, color_bar=True, set_palette=False)
+# plot only logical only (experiment is involved/not involved)
+logical_df = df_3.applymap(logical_values)
+plot_exp_details(logical_df, max_number_experiments=3)
 
 print('Run Complete\n')

### 2018-07-17 ident/python2/ss-ident/process_exp_details.py
convert zip to py3 stadard and remove bps

@@ -84,10 +84,9 @@ def exp_design_info(list_of_files, original_experiment_file, write_to_file_name,
                                   parameter_names=all_parameter_names)
     all_frequency = [i_parameter["frequency"] for i_parameter in all_parameter_info]
     col_ind_tuple = zip(all_parameter_info[0]['experiment_pos_id'], all_parameter_info[0]['experiment_id'])
-    import pdb;pdb.set_trace()
-    col_ind = pd.MultiIndex.from_tuples([i_tuple for i_tuple in col_ind_tuple], names=['experiment_pos_id', 'experiment_id'])
+    col_ind = pd.MultiIndex.from_tuples(list(col_ind_tuple), names=['experiment_pos_id', 'experiment_id'])
     flux_tuple = zip(all_parameter_flux, all_parameter_names)
-    row_ind = pd.MultiIndex.from_tuples(flux_tuple, names=['flux_name', 'parameter_name'])
+    row_ind = pd.MultiIndex.from_tuples(list(flux_tuple), names=['flux_name', 'parameter_name'])
     df = pd.DataFrame(all_frequency, index=row_ind, columns=col_ind)
     df.to_csv(write_to_file_name, index_label=['flux_name', 'parameter_name'])
     return df

### 2018-07-17 ident/python2/ss-ident/parallel_validate.py
change case of a in agg

@@ -9,7 +9,7 @@ import numpy as np
 import kotte_model
 import os.path
 import matplotlib
-matplotlib.use('Agg')
+matplotlib.use('agg')
 
 
 class ParallelValidate(object):

### 2018-07-17 ident/python2/ss-ident/process_exp_details.py
add bps

@@ -89,6 +89,7 @@ def exp_design_info(list_of_files, original_experiment_file, write_to_file_name,
     row_ind = pd.MultiIndex.from_tuples(list(flux_tuple), names=['flux_name', 'parameter_name'])
     df = pd.DataFrame(all_frequency, index=row_ind, columns=col_ind)
     df.to_csv(write_to_file_name, index_label=['flux_name', 'parameter_name'])
+    import pdb;pdb.set_trace()
     return df
 
 

### 2018-07-17 ident/python2/ss-ident/plot_ident_results.py
change sns color map to matplotlib colormap

@@ -2,6 +2,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import matplotlib.font_manager as fnt
+from matplotlib.colors import ListedColormap
 import pandas as pd
 import seaborn as sns
 from names_strings import variable_name
@@ -597,7 +598,8 @@ def plot_exp_details(df, color_bar=False, set_palette=True):
     for i_exp, i_exp_pos in enumerate(df.columns.levels[0]):
         i_exp_pos_df = df.loc[:, idx[i_exp_pos, :]]
         i_exp_pos_df.columns = i_exp_pos_df.columns.droplevel(level=0)
-        sns.heatmap(i_exp_pos_df, square=True, linewidth=0.5, cmap=c_pal, ax=ax[i_exp], cbar=color_bar)
+        sns.heatmap(i_exp_pos_df, square=True, linewidth=0.5, cmap=ListedColormap(c_pal.as_hex()), ax=ax[i_exp],
+                    cbar=color_bar)
         for item in ax[i_exp].get_xticklabels():
             item.set_rotation(90)
         # for item in ax[i_exp].get_yticklabels():

### 2018-07-17 ident/python2/ss-ident/plot_ident_results.py
rotate x-tick labels by 90

@@ -602,8 +602,8 @@ def plot_exp_details(df, color_bar=False, set_palette=True):
                     cbar=color_bar)
         for item in ax[i_exp].get_xticklabels():
             item.set_rotation(90)
-        # for item in ax[i_exp].get_yticklabels():
-        #     item.set_rotation(90)
+        for item in ax[i_exp].get_yticklabels():
+            item.set_rotation(90)
 
     # remove axis labels for shared x-axis
     for axis in range(0, len(df.columns.levels[0]) - 1):

### 2018-07-17 ident/python2/ss-ident/plot_ident_results.py
rotate x-tick labels by 270

@@ -603,7 +603,7 @@ def plot_exp_details(df, color_bar=False, set_palette=True):
         for item in ax[i_exp].get_xticklabels():
             item.set_rotation(90)
         for item in ax[i_exp].get_yticklabels():
-            item.set_rotation(90)
+            item.set_rotation(270)
 
     # remove axis labels for shared x-axis
     for axis in range(0, len(df.columns.levels[0]) - 1):

### 2018-07-17 ident/python2/ss-ident/plot_ident_results.py
dont rotate axis labels

@@ -602,8 +602,8 @@ def plot_exp_details(df, color_bar=False, set_palette=True):
                     cbar=color_bar)
         for item in ax[i_exp].get_xticklabels():
             item.set_rotation(90)
-        for item in ax[i_exp].get_yticklabels():
-            item.set_rotation(270)
+        # for item in ax[i_exp].get_yticklabels():
+        #     item.set_rotation(270)
 
     # remove axis labels for shared x-axis
     for axis in range(0, len(df.columns.levels[0]) - 1):

### 2018-07-17 ident/python2/ss-ident/plot_ident_results.py
rotate x-tick labels by 180

@@ -602,8 +602,8 @@ def plot_exp_details(df, color_bar=False, set_palette=True):
                     cbar=color_bar)
         for item in ax[i_exp].get_xticklabels():
             item.set_rotation(90)
-        # for item in ax[i_exp].get_yticklabels():
-        #     item.set_rotation(270)
+        for item in ax[i_exp].get_yticklabels():
+            item.set_rotation(180)
 
     # remove axis labels for shared x-axis
     for axis in range(0, len(df.columns.levels[0]) - 1):

### 2018-07-17 ident/python2/ss-ident/plot_ident_results.py
rotate x-tick labels by 0

@@ -603,7 +603,7 @@ def plot_exp_details(df, color_bar=False, set_palette=True):
         for item in ax[i_exp].get_xticklabels():
             item.set_rotation(90)
         for item in ax[i_exp].get_yticklabels():
-            item.set_rotation(180)
+            item.set_rotation(0)
 
     # remove axis labels for shared x-axis
     for axis in range(0, len(df.columns.levels[0]) - 1):

### 2018-07-17 ident/python2/ss-ident/plot_ident_results.py
fix matplotlib backend by calling matplotlib.use in every file

@@ -1,4 +1,6 @@
 import numpy as np
+import matplotlib
+matplotlib.use('agg')
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import matplotlib.font_manager as fnt

### 2018-07-17 ident/python2/ss-ident/run_ident.py

@@ -6,6 +6,8 @@ from collections import defaultdict
 import pandas as pd
 import numpy as np
 import itertools as it
+import matplotlib
+matplotlib.use('agg')
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import matplotlib.font_manager as fnt

### 2018-07-17 ident/python2/ss-ident/run_validate.py

@@ -1,3 +1,5 @@
+import matplotlib
+matplotlib.use('agg')
 from run_sims import ModelSim
 from names_strings import variable_name
 from plot_ident_results import plot_on_axis_object_box, plot_on_axis_object_violin

### 2018-07-17 ident/python2/ss-ident/ident_script.py
remove file - overridden by parallel_ident

@@ -1,68 +1,0 @@
-from names_strings import true_parameter_values
-from run_ident import ModelIdent
-from run_validation import ValidateSim
-from mpi4py import MPI
-import numpy as np
-import kotte_model
-import os.path
-
-
-name = MPI.Get_processor_name()
-rank = MPI.COMM_WORLD.Get_rank()
-size = MPI.COMM_WORLD.Get_size()
-print('I am  %s rank %d (total %d)' % (name, rank, size))
-
-if rank == 0:
-    # extract experimental data from file
-    default_parameter_values = true_parameter_values()
-
-    v1_ident = ModelIdent(ident_fun=kotte_model.flux_1_kcat_ident,
-                          arranged_data_file_name=os.path.join(os.getcwd(), 'exp/exp_v1_2_experiments'),
-                          ident_data_file_name=os.path.join(os.getcwd(), 'ident/ident_v1_kcat'),
-                          **{'original_exp_file': os.path.join(os.getcwd(), 'exp/experiments'),
-                             'flux_id': 1, 'flux_choice': 2,
-                             'values_figure': os.path.join(os.getcwd(), 'results/v1_kcat_parameter_values.eps'),
-                             'ident_figure': os.path.join(os.getcwd(), 'results/v1_kcat_ident.eps'),
-                             'exp_figure': os.path.join(os.getcwd(), 'results/v1_kcat_exp.eps'),
-                             'figure_format': 'eps'})
-    # test identifiability
-    print('Practical Identifiability Analysis of v1 with 2 parameters: k1cat and K1ac\n')
-    ident_data_df = v1_ident.perform_ident(name, rank, size)
-
-    import pdb;pdb.set_trace()
-    v1_ident.process_ident()
-
-    v1_ident.get_parameter_value()
-
-    # validate estimated parameter values
-    user_ode_opts = {'iter': 'Newton', 'discr': 'Adams', 'atol': 1e-10, 'rtol': 1e-10,
-                     'time_points': 200, 'display_progress': True, 'verbosity': 30}
-    # initial ss to begin all simulations from
-    y0 = np.array([5, 1, 1])
-    # get and set true parameter values, if available separately
-    default_parameters = true_parameter_values()
-
-    validate_obj = ValidateSim(kotte_model.kotte_ck_ode, kotte_model.kotte_ck_flux, **{'kinetics': 2,
-                                                                                       'ode_opts': user_ode_opts,
-                                                                                       't_final': 200,
-                                                                                       'wt_y0': y0,
[…]

### 2018-07-17 ident/docs/draft_4.tex

@@
 
+	\begin{figure}[!tbhp]
+		\centering{\includegraphics[width=1\textwidth,height=.8\textheight,keepaspectratio]{figures/summary_fig/all_flux_exp_ident}}
+		\caption{The frequency and occurrence of each experiment that produce data sets that can identify each parameter in each of the four fluxes a) $v_1$, $v_2$, $v_5$ and b) $v_3$ in the small metabolic network are shown.}\label{fig:ident-expt}
+	\end{figure}
+
 	%Apart from the mismatch between the in vivo and the in vitro enzyme kinetic rate laws, uncertainty in experimental data also arises due to the presence of noise in the measured experimental data.

### 2018-07-17 KineticModel/estimation/scripts-other/est_flux1.m
remove all unnecessary files

@@ -1,111 +1,0 @@
-% script to estimate parameters from noisy data for flux 1 in kotte model
-
-% generate experimental data - get initial ss
-tspan = 0:0.1:100;
-[xdyn,fdyn,xss1,fss1,opts] = run_nonoise(tspan);
-
-% backup parameters and initial conditions
-ival_bkp = opts.x0;
-odep_bkp = opts.odep;
-
-% generate nsmp samples by adding random noise to ss values
-nsmp = 10;
-[noisy_xss,noisy_fss] = addnoise(repmat(xss1,1,nsmp),odep_bkp);
-
-% figure
-% boxplot([noisy_fss';fss1']);
-% figure
-% boxplot([noisy_xss';xss1']);
-
-% perturb system from noisy initial conditions
-pt_sol_id = [1 2 3];
-[exp_sol,noisy_sol] = dopert_noisy(opts,noisy_xss,noisy_fss,odep_bkp,pt_sol_id);
-close all
-
-% get only data from one steady state
-pss = ones(1,numel(exp_sol.exp_pval));
-% pss(2) = 0;
-% pss(exp_sol.xss(1,:)>exp_sol.xss(2,:)) = 0;    
-
-% options structure for solving problem
-optimdata = struct('nc',3,'nflx',6,'nf',1,'flxid',1,'eps_v',1,...
-                    'eps_c',1,'vexp',exp_sol.fss(:,logical(pss)),...
-                    'xexp',exp_sol.xss(:,logical(pss)),...
-                    'flux_wt',100,'conc_wt',100,...
-                    'eps_c_wt',1,'eps_v_wt',1000,...
-                    'odep',odep_bkp,...
-                    'wt_xss',noisy_xss(:,1),'wt_fss',noisy_fss(:,1),...
-                    'type',2);
-expdata = struct('vexp',exp_sol.fss(:,logical(pss)),...
-                'xexp',exp_sol.xss(:,logical(pss)));              
-
-% define all bound fun handles
-allboundh = {'boundsf1',...
-             [],...
-             [],...
-             [],...
-             [],...
-             []};
-% define all objectiv efun handles
-allobjh = {'obj_typec',...
-           [],...
-           [],...
-           [],...
-           [],...
-           []};
-% define all constraint fun handles
-allconsh = {'cons_typec_f1',...
-            [],...
-            [],...
-            [],...
[…]
